### 2015-02-25 latex/paper.Rnw
TeX

\documentclass[12pt,a4paper]{article}
\usepackage{amsmath}
\begin{document}
%\SweaveOpts{concordance=TRUE}

\title{Sequential rank agreement methods for comparison of ranked lists}
\author{}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

Fra Kaspers PhD. Ikke rettet til endnu:

Comparison of ranked lists, though not a recently emerged subject has been receiving increased attention with the use of internet search engines, but lists of this kind are found in many areas. The advances in biotechnology have renewed this interest since many genetic studies are producing vast amounts of data that need to be sorted and evaluated. For example, studies of gene expression often rank genes according to their difference in expression across samples, while search engine results are presented with the best match first. Common for these types of lists is that it is up to the investigator to decide which k is relevant for a given problem when comparing top k lists. Typically, the items in the lists will be more similar at the top of the lists, since strong effects ensure that an item are more likely to be identified rather than further down where items consisting purely or mostly of noise are arranged more randomly.

\section{Methods}
Let there be given $L$ lists, $L\geq2$ of length $P$ that each rank the same set of $P$ items (variables) such that the first element of a list $l$ is the item with rank 1, the second element of $l$ is the item with rank 2 etc. Without any loss of generality we can assume that the predictors are numbered from $1$ to $P$ so that an ordered list, $l$, essentially is a permutation of the numbers from $1$ to $P$. This gives rise to a bijectiuve permutation function, $\pi_l(r) \in \{1, \ldots, P\}$ which returns the predictor that was found at rank $r$ in list $l$.
Conversely, we can define the rank function
\begin{equation}
R_l(p) = \pi_l^{-1}(p)
\end{equation}
which returns the rank given to item $p$ in list $l$ (see Figure~\ref{fig:example} for a schematic example of these two functions).

\begin{figure}[tb]

\begin{center}
\begin{minipage}{4cm}
a)
\begin{tabular}{cccc} \hline\hline
  & \multicolumn{3}{c}{List} \\
Rank  & 1 & 2 & 3 \\ \hline
1 & A & A & B \\
2 & B & C & A \\
3 & C & D & E \\
4 & D & B & C \\
5 & E & E & D \\ \hline
\end{tabular}
\end{minipage}
\begin{minipage}{4cm}b)
\begin{tabular}{cccc} \hline\hline
  & \multicolumn{3}{c}{List} \\
Item  & 1 & 2 & 3 \\ \hline
A & 1 & 1 & 2 \\
B & 2 & 4 & 1 \\
C & 3 & 2 & 4 \\
D & 4 & 3 & 5 \\
E & 5 & 5 & 3 \\ \hline
\end{tabular}
\end{minipage}
\begin{minipage}{4cm}c)
\begin{tabular}{cc} \hline\hline
Depth   &     \\
$d$  &  $S_d$ \\ \hline
1 & $\{$A, B$\}$\\
2 & $\{$A, B, C$\}$ \\
3 & $\{$A, B, C, D, E$\}$ \\
4 & $\{$A, B, C, D, E$\}$ \\
5 & $\{$A, B, C, D, E$\}$ \\ \hline
\end{tabular}
\end{minipage}
\end{center}
\caption{Example set of ranked lists. a) presents the ranked list of predictors/items for each of three lists while b) presents the ranks obtained by each predictor/item in each of the three lists.}
\label{fig:example}
\end{figure}

For a rank $d$ we define the set of items that have been ranked less than or equal to $d$ in any of the $L$ lists as
\begin{equation}
S_d = \{\pi_l(r) ; r \leq d, l = 1, \ldots, L \}.
\end{equation}
$S_d$ contains the set of items found in the $L$ top $d$ lists.
Finally, we define $A(p)$ as the measure of agreement of the rankings given to item $p$ in the $L$ lists,
\begin{equation}
A(p) = f(R_1(p), \ldots, R_L(p)),
\end{equation}
for a given pre-specified function $f$. Throughout this paper we will use the sample standard error as our function $f$,
$$f(R_1(p), \ldots, R_L(p)) = \sqrt{\frac{\sum_{i=1}^L (R_i(p) - \bar{R}(p))^2}{L-1}},
$$
but other choices could be made (see the discussion). The sample standard error has a nice interpretation as the average distance the individual rankings for item $p$ is from the average ranking of item $p$. Thus the investigator can base his or her assessment  ... on a ranking

... define the \emph{sequential rank agreement} as the pooled standard deviation of the items found in the set $S_d$:
\begin{equation}
\textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)A(p)^2}{(L-1)\cdot|S_d|}}.
\end{equation}

\subsection{All lists fully observed}
We shall start by the simplest case where all $L$ lists are fully observed.

\subsection{Analysis of top $k$ lists}
Not uncommon for lists to be

censored

Let $\Lambda_l, l=1, \ldots, L$ be the set of items found in list $l$ so $\Lambda_l$ is the top $k_l$ list of items from list $l$ where $k_l = |\Lambda_l|$. Note that we observe the top $k$ items for each of the $L$ lists if $k_1=\cdots=k_L=k$. For censored lists the rank function becomes
\begin{equation}
R_l(p) \text{ is } \left\{\begin{array}{cl} \pi_l^{-1}(p) & \text{ for } p\in \Lambda_l \\ 
> k_l & \text{ for } p \not\in \Lambda_l\end{array}\right.
\end{equation}
where we only know that the rank for the unobserved items in list $l$ must be larger than the largest rank observed in that list.

In the case of censored lists it only makes sense to look at depths where we have corresponding observations so the largest rank we should consider will be 
\begin{equation}
d \leq \max(k_1, \ldots, k_L).
\end{equation}

We cannot directly compute $A(p)$ for all predictors because we only observe a censored version of $R$ for some of the lists. Instead we assume that the rank assigned to predictor $p$ in list $l$ is uniformly distributed among the ranks that have been unassigned for list $l$. The ranks are clearly not independent since each of the lists essentially contains full set of ranks

\begin{equation}
A(p) = \frac{\sum_{r_1; r_1\in R_1(p)}  \cdots \sum_{r_L; r_L\in R_L(p)} f(r_1, \ldots, r_L)}{\prod_l |R_l(p)|}
\end{equation}

Notationen er ikke helt skarp da R jo returnerer delvist et tal og delvist en mængde, men det giver næppe anledning til misforståelser

\section{Applications}

\subsection{Evaluating results from top-$k$ lists}

\subsection{Comparing results across different method}

\subsection{Stability of results}

Bootstrap across a single method and compare results. Discuss collinearity

\section{Discussion}

Mention/discuss different measures.

\end{document}

### 2015-02-27 latex/paper.tex
TeX
compactification of notation

--- 2015-02-25 latex/paper.Rnw
@@
-\documentclass[12pt,a4paper]{article}
+\documentclass[12pt,a4paper]{article}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\usepackage{subcaption}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage{amsmath}
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 
-Fra Kaspers PhD. Ikke rettet til endnu:
+Ranked lists occur in many applications of statistics. Regression
+methods rank predictor variables according to magnitude of association
+with outcome, prediction models rank subjects according to their risk
+of an event, and genetic studies rank genes according to their
+difference in expression across samples. A common research question is
+where to stop, i.e., to decide the maximal significant rank.  In the
+three examples this would correspond to the number of significantly
+associated predictor variables, the number of patients at high risk of
+an event, and the number of genes that are worth to pursue in further
+experiments, respectively.
 
-Comparison of ranked lists, though not a recently emerged subject has been receiving increased attention with the use of internet search engines, but lists of this kind are found in many areas.
-The advances in biotechnology have renewed this interest since many genetic studies are producing vast amounts of data that need to be sorted and evaluated.
-For example, studies of gene expression often rank genes according to their difference in expression across samples, while search engine results are presented with the best match first.
-Common for these types of lists is that it is up to the investigator to decide which k is relevant for a given problem when comparing top k lists.
-Typically, the items in the lists will be more similar at the top of the lists, since strong effects ensure that an item are more likely to be identified rather than further down where items consisting purely or mostly of noise are arranged more randomly.
+In this article we describe some new breakthrough tools for measuring
+agreement across a set of lists which soon will enter the state-of-the
+art. The methods should be useful whenever there are multiple rankings
+of the same list. The idea is to define agreement based on the ranks
+of the first k elements in each list.
 
 \section{Methods}
-Let there be given $L$ lists, $L\geq2$ of length $P$ that each rank the same set of $P$ items (variables) such that the first element of a list $l$ is the item with rank 1, the second element of $l$ is the item with rank 2 etc.
-Without any loss of generality we can assume that the predictors are numbered from $1$ to $P$ so that an ordered list, $l$, essentially is a permutation of the numbers from $1$ to $P$.
-This gives rise to a bijectiuve permutation function, $\pi_l(r) \in \{1, \ldots, P\}$ which returns the predictor that was found at rank $r$ in list $l$.
-Conversely, we can define the rank function
-\begin{equation}
-R_l(p) = \pi_l^{-1}(p)
-\end{equation}
-which returns the rank given to item $p$ in list $l$ (see Figure~\ref{fig:example} for a schematic example of these two functions).
 
-\begin{figure}[tb]
+Consider a set of $P$ different items $X=\{X_1,\dots,X_P\}$. An
+ordered list is a permutation of the numbers from $1$ to $P$, i.e., a
+mapping $R:X\mapsto \{1,\dots,P\}$, such that $1\le R(X_p)\le P$ is
+the rank of item $X_p$ in the list. Note that $R$ is equivalent to a
+permutation function (a bijection from $\{1,\dots,P\}$ to
+itself). Thus it has an inverse mapping $\pi=R^{-1}$ which assings to
+a rank $r\in\{1,\dots,P\}$ the item $\pi(r)$ found at that rank. The
+methods described below work for a set of $L$ lists $R_1,\dots,R_L$,
+$L\geq2$. Panels (a) and (b) of Table~\ref{tab:example} show a
+schematic example of these mappings.
 
+\begin{table}[tb]
+  \caption{Example set of ranked lists.
+(a) shows the ranked list of items for each of three lists, (b) presents the ranks obtained by each item in each of the three lists and (c) shows the cumulative set of items found between rank 1 and a given depths in the three lists.}
 \begin{center}
-\begin{minipage}{4cm}
-a)
-\begin{tabular}{cccc} \hline\hline
-  & \multicolumn{3}{c}{List} \\
-Rank  & 1 & 2 & 3 \\ \hline
+  \begin{subtable}{4cm}%
+    \caption{}
+      \begin{tabular}{cccc}
+        \hline\hline % & \multicolumn{3}{c}{List} \\
+        Rank & $\pi_1$ & $\pi_2$ & $\pi_3$ \\ \hline
         1 & A & A & B \\
@@
     \end{tabular}
-\end{minipage}
-\begin{minipage}{4cm}b)
+  \end{subtable}
+% \hfill
+\hspace{1em}
+%%% 
+  \begin{subtable}{4cm}%
+    \caption{}
     \begin{tabular}{cccc} \hline\hline
-  & \multicolumn{3}{c}{List} \\
-Item  & 1 & 2 & 3 \\ \hline
+  % & \multicolumn{3}{c}{List} \\
+    Item & $R_1$ & $R_2$ & $R_3$ \\ \hline
     A & 1 & 1 & 2 \\
@@
   \end{tabular}
-\end{minipage}
-\begin{minipage}{4cm}c)
+\end{subtable}
+% \hfill
+\hspace{1em}
+%%%
+\begin{subtable}{4cm}%
+  \caption{}
 \begin{tabular}{cc} \hline\hline
-Depth   &     \\
-$d$  &  $S_d$ \\ \hline
+% Depth   &     \\
+Depth  &  $S_d$ \\ \hline
 1 & $\{$A, B$\}$\\
@@
 \end{tabular}
-\end{minipage}
+\end{subtable}
 \end{center}
-\caption{Example set of ranked lists. a) presents the ranked list of predictors/items for each of three lists while b) presents the ranks obtained by each predictor/item in each of the three lists.}
-\label{fig:example}
-\end{figure}
+\label{tab:example}
+\end{table}
 
-For a rank $d$ we define the set of items that have been ranked less than or equal to $d$ in any of the $L$ lists as
+The agreement of the lists regarding the rank given to an item can be
+measured by
+\begin{equation}
+  A(p) = f(R_1(p), \ldots, R_L(p)),
+\end{equation}
+for a distance function $f$. Throughout this paper we will use the
+sample standard error as our function $f$,
+$$f(R_1(p), \ldots, R_L(p)) = \sqrt{\frac{\sum_{i=1}^L (R_i(p) - \bar{R}(p))^2}{L-1}},
+$$
+but other choices could be made (see the discussion). The sample
+standard error has an interpretation as the average distance of the
+individual rankings of the lists from the average ranking.
+
+We now describe what is exemplified in Panel (c) of Table
+\ref{tab:example} and how it can be used to define \emph{sequential
+  rank agreement}. For an integer $1\le d\le P$ we define the unique
+set of items found in the $L$ top $d$ parts of the lists, i.e., the
+set of items ranked less than or equal to $d$ in any of the lists:
 \begin{equation}
@@
 \end{equation}
-$S_d$ contains the set of items found in the $L$ top $d$ lists.
-Finally, we define $A(p)$ as the measure of agreement of the rankings given to item $p$ in the $L$ lists,
+The \emph{sequential rank agreement} is the pooled standard deviation
+of the items found in the set $S_d$:
 \begin{equation}
-A(p) = f(R_1(p), \ldots, R_L(p)),
-\end{equation}
-for a given pre-specified function $f$. Throughout this paper we will use the sample standard error as our function $f$,
-$$f(R_1(p), \ldots, R_L(p)) = \sqrt{\frac{\sum_{i=1}^L (R_i(p) - \bar{R}(p))^2}{L-1}},
-$$
-but other choices could be made (see the discussion).
-The sample standard error has a nice interpretation as the average distance the individual rankings for item $p$ is from the average ranking of item $p$.
-Thus the investigator can base his or her assessment  ...
-on a ranking
-
-... define the \emph{sequential rank agreement} as the pooled standard deviation of the items found in the set $S_d$:
-\begin{equation}
-\textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)A(p)^2}{(L-1)\cdot|S_d|}}.
+\textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)A(p)^2}{(L-1) |S_d|}}.
 \end{equation}

### 2015-02-27 R/plot.agreement.R

@@ -8,12 +8,10 @@
 ##' @param ylab y-axis lab
 ##' @param ... Processed by \code{\link{prodlim::SmartControl}}.
 ##' @return Graph
-##' @seealso 
 ##' @examples
 ##' a <- agreement(list(c(1,2,3,4,5,7,6,8,9,10,11,12,13), c(5,11,4,7,8,3,12,13,6,10,9,2,1)))
 ##' plot(a)
-#' @method plot agreement
-#' @S3method plot agreement
+#' @export
 ##' @author Thomas A. Gerds <tag@@biostat.ku.dk>
 plot.agreement <- function(x,xlim,ylim,xlab="List depth",ylab="Agreement",add=FALSE,...){
     if (missing(xlim)) xlim=c(0,x$depth)

### 2015-02-27 latex/paper.tex
arbejds session towards state-of-the art

@@
 Consider a set of $P$ different items $X=\{X_1,\dots,X_P\}$. An
-ordered list is a permutation of the numbers from $1$ to $P$, i.e., a
-mapping $R:X\mapsto \{1,\dots,P\}$, such that $1\le R(X_p)\le P$ is
-the rank of item $X_p$ in the list. Note that $R$ is equivalent to a
-permutation function (a bijection from $\{1,\dots,P\}$ to
-itself). Thus it has an inverse mapping $\pi=R^{-1}$ which assings to
-a rank $r\in\{1,\dots,P\}$ the item $\pi(r)$ found at that rank. The
-methods described below work for a set of $L$ lists $R_1,\dots,R_L$,
-$L\geq2$. Panels (a) and (b) of Table~\ref{tab:example} show a
+ordered list is a permutation function, $R: \{1,\dots,P\}\to
+\{1,\dots,P\}$, such that $R(X_p)$ is the rank of item $X_p$ in the
+list. The inverse mapping $\pi=R^{-1}$ assigns to rank
+$r\in\{1,\dots,P\}$ the item $\pi(r)$ found at that rank. The methods
+described below work for a set of $L$ lists $R_1,\dots,R_L$,
+$L\geq2$. We denote $\pi_l=R_l^{-1}$ for the corresponding inverse
+mappings. Panels (a) and (b) of Table~\ref{tab:example} show a
 schematic example of these mappings.
@@
   \caption{Example set of ranked lists.
-(a) shows the ranked list of items for each of three lists, (b) presents the ranks obtained by each item in each of the three lists and (c) shows the cumulative set of items found between rank 1 and a given depths in the three lists.}
+(a) shows the ranked list of items for each of three lists, (b) presents the ranks obtained by each item in each of the three lists and (c) shows the cumulative set of items up to a given depth in the three lists.}
 \begin{center}
@@
 for a distance function $f$. Throughout this paper we will use the
-sample standard error as our function $f$,
-$$f(R_1(p), \ldots, R_L(p)) = \sqrt{\frac{\sum_{i=1}^L (R_i(p) - \bar{R}(p))^2}{L-1}},
+sample standard error as our function $f$ and hence use
+$$A(p) = \sqrt{\frac{\sum_{i=1}^L (R_i(p) - \bar{R}(p))^2}{L-1}},
 $$
@@
 \begin{equation}
-\textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)A(p)^2}{(L-1) |S_d|}}.
+\textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)A(p)^2}{(L-1)|S_d|}}.
 \end{equation}
@@
 
-Let $\Lambda_l, l=1, \ldots, L$ be the set of items found in list $l$ so $\Lambda_l$ is the top $k_l$ list of items from list $l$ where $k_l = |\Lambda_l|$.
-Note that we observe the top $k$ items for each of the $L$ lists if $k_1=\cdots=k_L=k$.
-For censored lists the rank function becomes
+Let $\Lambda_l, l=1, \ldots, L$ be the set of items found in list $l$
+so $\Lambda_l$ is the top $k_l$ list of items from list $l$ where $k_l
+= |\Lambda_l|$. Note that we observe the top $k$ items for each of the
+$L$ lists if $k_1=\cdots=k_L=k$. For censored lists the rank function
+becomes
 \begin{equation}
-R_l(p) \text{ is } \left\{\begin{array}{cl} \pi_l^{-1}(p) & \text{ for } p\in \Lambda_l \\ 
-> k_l & \text{ for } p \not\in \Lambda_l\end{array}\right.
+\tilde R_l(p) = \left\{\begin{array}{cl} \{\pi_l^{-1}(p)\} & \text{ for } p\in \Lambda_l \\ 
+\{k_l+1,\dots,P\} & \text{ for } p \not\in \Lambda_l\end{array}\right.
 \end{equation}
@@
 
-In the case of censored lists it only makes sense to look at depths where we have corresponding observations so the largest rank we should consider will be 
+In the case of censored lists it is sufficient (FIXME: requires
+argument) to look at depths where we have corresponding observations
+so the largest rank we should consider will be
 \begin{equation}
@@
 
-We cannot directly compute $A(p)$ for all predictors because we only observe a censored version of $R$ for some of the lists.
+We cannot directly compute $A(p)$ for all predictors because we only observe a censored version of $\tilde R$ for some of the lists.
 Instead we assume that the rank assigned to predictor $p$ in list $l$ is uniformly distributed among the ranks that have been unassigned for list $l$.
@@
 \begin{equation}
-A(p) = \frac{\sum_{r_1; r_1\in R_1(p)}  \cdots \sum_{r_L; r_L\in R_L(p)} f(r_1, \ldots, r_L)}{\prod_l |R_l(p)|}
+\tilde A(p) = \frac{\sum_{r_1; r_1\in \tilde R_1(p)}  \cdots \sum_{r_L; r_L\in \tilde R_L(p)} A(p)}{\prod_l |\tilde R_l(p)|}
 \end{equation}
 
-Notationen er ikke helt skarp da R jo returnerer delvist et tal og delvist en mængde, men det giver næppe anledning til misforståelser
+FIXME: If instead of running through all elements of $\tilde R_1(p)$
+one would use the average rank in $\tilde R_1(p)$ we would end up with
+a too small variance.
+
+\section{Benchmarks}
 

### 2015-02-27 R/plot.agreement.R

@@ -9,13 +9,27 @@
 ##' @param ... Processed by \code{\link{prodlim::SmartControl}}.
 ##' @return Graph
 ##' @examples
-##' a <- agreement(list(c(1,2,3,4,5,7,6,8,9,10,11,12,13), c(5,11,4,7,8,3,12,13,6,10,9,2,1)))
+##' R1=c(1,2,3,4,5,7,6,8,9,10,11,12,13)
+##' R2=c(5,11,4,7,8,3,12,13,6,10,9,2,1)
+##' a <- agreement(list(R1,R2))
 ##' plot(a)
+##' arand = colMeans(do.call("rbind",lapply(1:20,function(b){
+##'     agreement(list(sample(R1),sample(R1)))$agreement
+##' })))
+##' lines(1:length(R1),arand,col=2,lwd=3)
+##' 
+##' l <- c(1,2,3,4,5,7,6,8,9,10,11,12,13)
+##' l <- 0:100
+##' aa <- sapply(1:20,function(i){
+##'     agreement(list(sample(l),sample(l),sample(l)))$agreement[1]
+##' })
+##' c(mean(aa),sd(aa))
+##' 
 #' @export
 ##' @author Thomas A. Gerds <tag@@biostat.ku.dk>
 plot.agreement <- function(x,xlim,ylim,xlab="List depth",ylab="Agreement",add=FALSE,...){
-    if (missing(xlim)) xlim=c(0,x$depth)
-    if (missing(ylim)) ylim=range(x$agreement)
+    if (missing(xlim)) xlim=c(1,x$depth)
+    if (missing(ylim)) ylim=c(0,max(x$agreement))
     axis1.DefaultArgs <- list()
     axis2.DefaultArgs <- list(side=2)
     plot.DefaultArgs <- list(x=0,y=0,type = "n",ylim = ylim,xlim = xlim,xlab = xlab,ylab = ylab)

### 2015-02-27 latex/paper.Rnw

@@
 \documentclass[12pt,a4paper]{article}
+\usepackage[]{graphicx}
 \usepackage{amsmath}
+\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\usepackage{subcaption}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
 \begin{document}
@@
 
-Fra Kaspers PhD. Ikke rettet til endnu:
+Ranked lists occur in many applications of statistics. Regression
+methods rank predictor variables according to magnitude of association
+with outcome, prediction models rank subjects according to their risk
+of an event, and genetic studies rank genes according to their
+difference in expression across samples. A common research question is
+where to stop, i.e., to decide the maximal significant rank.  In the
+three examples this would correspond to the number of significantly
+associated predictor variables, the number of patients at high risk of
+an event, and the number of genes that are worth to pursue in further
+experiments, respectively.
 
-Comparison of ranked lists, though not a recently emerged subject has been receiving increased attention with the use of internet search engines, but lists of this kind are found in many areas.
-The advances in biotechnology have renewed this interest since many genetic studies are producing vast amounts of data that need to be sorted and evaluated.
-For example, studies of gene expression often rank genes according to their difference in expression across samples, while search engine results are presented with the best match first.
-Common for these types of lists is that it is up to the investigator to decide which k is relevant for a given problem when comparing top k lists.
-Typically, the items in the lists will be more similar at the top of the lists, since strong effects ensure that an item are more likely to be identified rather than further down where items consisting purely or mostly of noise are arranged more randomly.
+In this article we describe some new breakthrough tools for measuring
+agreement across a set of lists which soon will enter the state-of-the
+art. The methods should be useful whenever there are multiple rankings
+of the same list. The idea is to define agreement based on the ranks
+of the first k elements in each list.
 
 \section{Methods}
-Let there be given $L$ lists, $L\geq2$ of length $P$ that each rank the same set of $P$ items (variables) such that the first element of a list $l$ is the item with rank 1, the second element of $l$ is the item with rank 2 etc.
-Without any loss of generality we can assume that the predictors are numbered from $1$ to $P$ so that an ordered list, $l$, essentially is a permutation of the numbers from $1$ to $P$.
-This gives rise to a bijectiuve permutation function, $\pi_l(r) \in \{1, \ldots, P\}$ which returns the predictor that was found at rank $r$ in list $l$.
-Conversely, we can define the rank function
-\begin{equation}
-R_l(p) = \pi_l^{-1}(p)
-\end{equation}
-which returns the rank given to item $p$ in list $l$ (see Figure~\ref{fig:example} for a schematic example of these two functions).
 
-\begin{figure}[tb]
+Consider a set of $P$ different items $X=\{X_1,\dots,X_P\}$. An
+ordered list is a permutation function, $R: \{1,\dots,P\}\to
+\{1,\dots,P\}$, such that $R(X_p)$ is the rank of item $X_p$ in the
+list. The inverse mapping $\pi=R^{-1}$ assigns to rank
+$r\in\{1,\dots,P\}$ the item $\pi(r)$ found at that rank. The methods
+described below work for a set of $L$ lists $R_1,\dots,R_L$,
+$L\geq2$. We denote $\pi_l=R_l^{-1}$ for the corresponding inverse
+mappings. Panels (a) and (b) of Table~\ref{tab:example} show a
+schematic example of these mappings.
 
+\begin{table}[tb]
+  \caption{Example set of ranked lists.
+(a) shows the ranked list of items for each of three lists, (b) presents the ranks obtained by each item in each of the three lists and (c) shows the cumulative set of items up to a given depth in the three lists.}
 \begin{center}
-\begin{minipage}{4cm}
-a)
-\begin{tabular}{cccc} \hline\hline
-  & \multicolumn{3}{c}{List} \\
-Rank  & 1 & 2 & 3 \\ \hline
+  \begin{subtable}{4cm}%
+    \caption{}
+      \begin{tabular}{cccc}
+        \hline\hline % & \multicolumn{3}{c}{List} \\
+        Rank & $\pi_1$ & $\pi_2$ & $\pi_3$ \\ \hline
         1 & A & A & B \\
@@
 \end{tabular}
-\end{minipage}
+\end{subtable}
 \end{center}
-\caption{Example set of ranked lists. a) presents the ranked list of predictors/items for each of three lists while b) presents the ranks obtained by each predictor/item in each of the three lists.}
-\label{fig:example}
-\end{figure}
+\label{tab:example}
+\end{table}
 
-For a rank $d$ we define the set of items that have been ranked less than or equal to $d$ in any of the $L$ lists as
+The agreement of the lists regarding the rank given to an item can be
+measured by
+\begin{equation}
+  A(p) = f(R_1(p), \ldots, R_L(p)),
+\end{equation}
+for a distance function $f$. Throughout this paper we will use the
+sample standard error as our function $f$ and hence use
+$$A(p) = \sqrt{\frac{\sum_{i=1}^L (R_i(p) - \bar{R}(p))^2}{L-1}},
+$$
+but other choices could be made (see the discussion). The sample
+standard error has an interpretation as the average distance of the
+individual rankings of the lists from the average ranking.
+
+We now describe what is exemplified in Panel (c) of Table
+\ref{tab:example} and how it can be used to define \emph{sequential
+  rank agreement}. For an integer $1\le d\le P$ we define the unique
+set of items found in the $L$ top $d$ parts of the lists, i.e., the
+set of items ranked less than or equal to $d$ in any of the lists:
 \begin{equation}
@@
 \end{equation}
-$S_d$ contains the set of items found in the $L$ top $d$ lists.
-Finally, we define $A(p)$ as the measure of agreement of the rankings given to item $p$ in the $L$ lists,
+The \emph{sequential rank agreement} is the pooled standard deviation
+of the items found in the set $S_d$:
 \begin{equation}
-A(p) = f(R_1(p), \ldots, R_L(p)),
-\end{equation}
-for a given pre-specified function $f$. Throughout this paper we will use the sample standard error as our function $f$,
-$$f(R_1(p), \ldots, R_L(p)) = \sqrt{\frac{\sum_{i=1}^L (R_i(p) - \bar{R}(p))^2}{L-1}},
-$$
-but other choices could be made (see the discussion).
-The sample standard error has a nice interpretation as the average distance the individual rankings for item $p$ is from the average ranking of item $p$.
-Thus the investigator can base his or her assessment  ...
-on a ranking
-
-... define the \emph{sequential rank agreement} as the pooled standard deviation of the items found in the set $S_d$:
-\begin{equation}
-\textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)A(p)^2}{(L-1)\cdot|S_d|}}.
+\textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)A(p)^2}{(L-1)|S_d|}}.
 \end{equation}

### 2015-03-20 R/InformationRetrieval.R
Diverse

@@ -1,0 +1,10 @@
+#' 
+#' 
+#' 
+#' 
+#' 
+
+PrecisionAtK <- function()
+{
+  
+}

### 2015-03-20 R/sra.R

@@ -1,0 +1,79 @@
+sra <- function(object, metric=var, depth=NULL, listlength=NULL, B=10) {
+    metric <- match.fun(metric)
+    
+    # Make sure that the input object ends up as a matrix with integer columns all
+    # consisting of elements from 1 and up to listlength
+    if (is.matrix(object))
+        rankmat <- object
+    else
+        rankmat <- as.matrix(do.call("cbind",object))
+    
+    if (missing(depth)) depth <- NROW(rankmat)
+    
+    # Now do sanity check of input
+    # Check if there are dupliates in the columns
+    if (any(apply(rankmat, 2, duplicated, incomparables=c(0, NA)))) {
+      stop("an element must only occur exactly once in each list")
+    }
+    # Check and compute the maximum list length
+    ll <- max(c(NROW(rankmat), max(rankmat, na.rm=TRUE)))
+    if (is.null(listlength)) {
+      listlength <- ll
+    }    
+    if (listlength<ll) {
+      stop("You are requesting a list length that is smaller than the largest element found")
+    }    
+    # Should probably also check integers and such
+        
+    # Compute the full metric for each item in a full sized list
+    # The speed here could be increased (for example by converting everything to itemlist scale first)
+    full.list.metric <- function(m, metric) {
+      itemmatrix <- apply(m, 2, order)
+      apply(itemmatrix, 1, metric)
+    }
+    resample <- function(x, ...) x[sample.int(length(x), ...)]
+
+    
+    # General approach
+    # 1) Fill in missing items in each list
+    # 2) Compute the metric for the full lists
+    # 3) Compute agreement 
+    # 4) Finally average everything
+
+    nseq <- seq(listlength)
+    nlists <- NCOL(rankmat)
+  
+
+    # Compute a list of missing items for each list
+    missing.items <- lapply(as.data.frame(rankmat), function(x) { nseq[-x] })
+    
+    fullrankmat <- matrix(rep(0, nlists*listlength), ncol=nlists)
+    fullrankmat[1:NROW(rankmat),] <- rankmat
+    maxdepth <- min(nrow(fullrankmat), ifelse(is.null(depth), nrow(fullrankmat), depth))
+
+      
+    tmpres <- sapply(1:B, function(i) {
+      # Ad 1)
+      for (j in 1:nlists) {
+          if (length(missing.items[[j]])>0) {
[…]

### 2015-03-20 src/sra.cpp
RCpp

@@ -1,0 +1,156 @@
+#include <Rcpp.h>
+using namespace Rcpp;
+
+
+// wrapper around R's RNG such that we get a uniform distribution over
+// [0,n) as required by the STL algorithm
+inline int randWrapper(const int n) { return floor(unif_rand()*n); }
+
+IntegerVector randomShuffle(IntegerVector a) {
+    // already added by sourceCpp(), but needed standalone
+    RNGScope scope;             
+
+    // clone a into b to leave a alone
+    IntegerVector b = clone(a);
+
+    std::random_shuffle(b.begin(), b.end(), randWrapper);
+
+    return b;
+}
+
+
+IntegerVector order_(NumericVector x) {
+  // Have removed the 3 lines below because checking should be done elsewhere
+  //  if (is_true(any(duplicated(x)))) {
+  //   Rf_warning("There are duplicates in 'x'; order not guaranteed to match that of R's base::order");
+  // }
+  NumericVector sorted = clone(x).sort();
+  return match(sorted, x);
+}
+
+double square( double x ){
+    return x*x;
+}
+
+
+NumericVector SdPerRow(NumericMatrix obj) {
+
+  NumericVector res(obj.nrow());
+  
+  for (int i=0; i<obj.nrow(); i++) {
+    res[i] = var(obj.row(i));
+  }
+ 
+  return(res);
+}
+
+
+//[[Rcpp::export]]
+NumericVector sra(IntegerMatrix rankMat, int maxlength, int B, IntegerVector cens) {
+
+  // The number of lists
+  int nLists = rankMat.ncol();
+
+  // Need a numeric copy of this for the sort function. Doesn't really make sense but
+  // keeping it for now 
+  NumericMatrix x(maxlength, nLists);
+  for (int l=0; l<nLists; l++) {
+    for (int i=0; i<std::min(maxlength, rankMat.nrow()); i++) {
+      x(i, l) = rankMat(i,l);
+    }
+  }
+  
+ 
+  // Keep the average result and the individual permutations
+  NumericVector returnVector(maxlength);
+  NumericMatrix result(maxlength, B);
+
+    // Input validation
+
+    // Check that censoring elements is at least one. Or possibly not?
+
+    // Check that the censoring is less than maxlength
+  if(max(cens)>rankMat.nrow()) {
+    stop("You cannot censor a list later than you have data");
+  }
+  if(max(cens)>maxlength) {
+    stop("You cannot censor a list later than the maximum of the data");
+  }
+
+
[…]

### 2015-03-20 R/Rcpp.R
RCpp

@@ -1,0 +1,10 @@
+library(Rcpp)
+
+
+sourceCpp(file="sra.cpp")
+
+
+nitems <- 20
+m <- cbind(1:nitems, c(1,3,4,2,5,6, sample(7:nitems)))
+
+sra(m, 6, B=1, cens=c(6,6))

### 2015-03-20 R/example1.R

@@ -81,21 +81,7 @@ nulleval <- sapply(1:2, function(i) {
 })
 
 nulleval1 <- sapply(1:2, function(i) {
-    mt.p <- sapply(index, function(i) { anova(lm(y ~ x[,i]))[1,5] } )
-    list1 <- order(mt.p)
-
-    #3
-    MIC <- sapply(index, function(i) { cc <- complete.cases(snps[,i],y) ; mine(x[cc,i], y[cc])$MIC})
-    list3 <- order(MIC, decreasing=TRUE)
-
-    #4
-    f1 <- rfsrc(y~., data=d, ntree=1000)
-    variables <- abs(f1$importance)
-    num.undecided <- sum(variables==0)
-    list4 <- order(variables, decreasing=TRUE)
-    list4[(length(variables)-num.undecided):length(variables)] <- 0
-
-    res <- sra(list(list1, list3))$agreement
-    res
+    nobs <- length(list1)
+    res <- sra(list(rep(0, nobs), rep(0, nobs), rep(0, nobs)), listlength=nobs)$agreement
 })
 

### 2015-04-07 latex/paper.Rnw
Diverse
asd

@@
 \documentclass[12pt,a4paper]{article}
+\usepackage[utf8]{inputenc}
 \usepackage[]{graphicx}
@@
 \usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
+\usepackage{subcaption}
+\usepackage{natbib}
+\usepackage{xspace}
+\usepackage[noae]{Sweave}
 
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-
-\usepackage{framed}
-\usepackage{subcaption}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+%
+% New commands
+%
+\newcommand{\ie}{\emph{i.e.,}\@\xspace}
+\newcommand{\eg}{\emph{e.g,}\@\xspace}
 
@@
 \end{abstract}
+
+Key words: sequential rank agreement, rank agreement, order statistic,
+methods comparison, variable selection
 
@@
 \begin{equation}
-\textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)A(p)^2}{(L-1)|S_d|}}.
+\textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in
+      S_d\}}(L-1)A(p)^2}{(L-1)|S_d|}}, \label{def:sra}
 \end{equation}
+and small values close to zero suggests that the lists agree on the
+ordering while larger values suggests disagreement. If the ranked
+lists are identical then the value of SRA will be zero for all depths
+$d$.  The sequential rank agreement can be interpreted as the average
+distance of the individual rankings of the lists from the average
+ranking for each of the items we have seen until depth $d$.
 
 \subsection{All lists fully observed}
-We shall start by the simplest case where all $L$ lists are fully observed.
+We shall start by the simplest case where all $L$ lists are fully
+observed, \ie, we have the rank of all $P$ items for all of the $L$
+lists. This situation occurs is common when we have the original
+dataset available and when we wish to, say, compare the results from
+different analysis methods.
+
+For the fully obseved list situation we can plot the sequantial
+rank agreement \eqref{def:sra} as a function of depth $d$. If there is
+a ... An example is seen in Figure~\ref{XX}
 
@@
 
+Noget med
+
 \section{Benchmarks}
+
+Three approaches
+
+\subsection{Independent lists}
+
+\subsection{Permuted outcomes + analyses}
+
+Do the following a large number of times
+\begin{enumerate}
+  \item Permute outcome vector
+  \item Redo analyses for all $L$ methods
+  \item Compute sequential rank agreement for 
+\end{enumerate}
+
+Lidt spøjs ting man matcher dem op imod
+
+\subsection{Changepoint analysis}
 
@@
 
-\subsection{Evaluating results from top-$k$ lists}
-
 \subsection{Comparing results across different method}
 
+In a classical paper by \citet{Golub1999} a dataset of 3051 gene
+expression values were measured on 38 tumor mRNA samples in order to
+improve the classification of acute leukemias between two types: acute
+lymphoblastic leukemia (ALL) or acute myeloid leukemia
+(AML). Preprocessing of the gene expression data was done as described
+in \citep{Dudoit2002}.
+
+We analyzed these gene expression data using five different
+approaches:
+
 \subsection{Stability of results}
+
+\subsection{Evaluating results from top-$k$ lists}
 
@@
 
+\bibliographystyle{apalike}
+\bibliography{paperref}
+
 \end{document}

### 2015-04-07 R/example1.R

@@ -3,85 +3,116 @@
 # Read data
 #
 #
-
+library(glmnet)
+library(minerva)
+library(changepoint)
+library(randomForestSRC)
 
 source("sra.R")
 
-
 # Read data
-snps <- read.table("http://people.umass.edu/foulkes/asg/data/snpSimCVDNum.txt", header=TRUE)
-traits <- read.table("http://people.umass.edu/foulkes/asg/data/ySimCVD.txt", header=TRUE)
+library(multtest)
+data(golub)
 
 
-y <- traits$yA  
-x <- snps
-index <- seq(1, NCOL(snps))
-d <- data.frame(y, x)
-#md <- mice(d)
-
-# Marginal t-tests
-
-mt.p <- sapply(index, function(i) { anova(lm(y ~ x[,i]))[1,5] } )
-list1 <- order(mt.p)
-
-# Elastic net
-
-library(glmnet)
-#library(mice)
-#res <- coef(glmnet(as.matrix(x), y, alpha=0), s=lambda)[-1]
-
-# MIC
-
-library(minerva)
-MIC <- sapply(index, function(i) { cc <- complete.cases(snps[,i],y) ; mine(x[cc,i], y[cc])$MIC})
-list3 <- order(MIC, decreasing=TRUE)
-
-## Forest
-library(randomForestSRC)
-f1 <- rfsrc(y~., data=d, ntree=1000)
-variables <- abs(f1$importance)
-num.undecided <- sum(variables==0)
-list4 <- order(variables, decreasing=TRUE)
-list4[(length(variables)-num.undecided):length(variables)] <- 0
+y <- golub.cl
+x <- golub 
 
 
-res <- sra(list(list1, list3, list4))
-plot(sqrt(res$agreement),xlim=c(0,20))
+vs.vimp.md <- function(obj, which.outcome, conservative = TRUE) {
+   msub <- max.subtree(obj, conservative = conservative)
+   vimp <- obj$importance
+   if (is.null(vimp)) {
+       vimp <- matrix(1, nrow(msub$order), ncol(obj$xvar))
+   }
+   names <- rownames(msub$order)
+   names[vimp[, which.outcome] > 0 & msub$order[, 1] <= msub$threshold]    
+} 
+
+producelists <- function(x, y) {
+    index <- seq(1, nrow(golub))
+                                        # d <- data.frame(y, x)
+
+                                        # Marginale t-tests
+    mt.p <- sapply(index, function(i) { t.test(x[i,] ~ y)$p.value } )
[…]

### 2015-04-07 R/sra.R

@@ -1,79 +1,0 @@
-sra <- function(object, metric=var, depth=NULL, listlength=NULL, B=10) {
-    metric <- match.fun(metric)
-    
-    # Make sure that the input object ends up as a matrix with integer columns all
-    # consisting of elements from 1 and up to listlength
-    if (is.matrix(object))
-        rankmat <- object
-    else
-        rankmat <- as.matrix(do.call("cbind",object))
-    
-    if (missing(depth)) depth <- NROW(rankmat)
-    
-    # Now do sanity check of input
-    # Check if there are dupliates in the columns
-    if (any(apply(rankmat, 2, duplicated, incomparables=c(0, NA)))) {
-      stop("an element must only occur exactly once in each list")
-    }
-    # Check and compute the maximum list length
-    ll <- max(c(NROW(rankmat), max(rankmat, na.rm=TRUE)))
-    if (is.null(listlength)) {
-      listlength <- ll
-    }    
-    if (listlength<ll) {
-      stop("You are requesting a list length that is smaller than the largest element found")
-    }    
-    # Should probably also check integers and such
-        
-    # Compute the full metric for each item in a full sized list
-    # The speed here could be increased (for example by converting everything to itemlist scale first)
-    full.list.metric <- function(m, metric) {
-      itemmatrix <- apply(m, 2, order)
-      apply(itemmatrix, 1, metric)
-    }
-    resample <- function(x, ...) x[sample.int(length(x), ...)]
-
-    
-    # General approach
-    # 1) Fill in missing items in each list
-    # 2) Compute the metric for the full lists
-    # 3) Compute agreement 
-    # 4) Finally average everything
-
-    nseq <- seq(listlength)
-    nlists <- NCOL(rankmat)
-  
-
-    # Compute a list of missing items for each list
-    missing.items <- lapply(as.data.frame(rankmat), function(x) { nseq[-x] })
-    
-    fullrankmat <- matrix(rep(0, nlists*listlength), ncol=nlists)
-    fullrankmat[1:NROW(rankmat),] <- rankmat
-    maxdepth <- min(nrow(fullrankmat), ifelse(is.null(depth), nrow(fullrankmat), depth))
-
-      
-    tmpres <- sapply(1:B, function(i) {
-      # Ad 1)
-      for (j in 1:nlists) {
-          if (length(missing.items[[j]])>0) {
[…]

### 2015-04-07 R/oldsra.R
asd

@@ -1,0 +1,83 @@
+#' @useDynLib SuperRanker
+#' @importFrom Rcpp sourceCpp
+
+
+sra <- function(object, metric=var, depth=NULL, listlength=NULL, B=10) {
+    metric <- match.fun(metric)
+    
+    # Make sure that the input object ends up as a matrix with integer columns all
+    # consisting of elements from 1 and up to listlength
+    if (is.matrix(object))
+        rankmat <- object
+    else
+        rankmat <- as.matrix(do.call("cbind",object))
+    
+    if (missing(depth)) depth <- NROW(rankmat)
+    
+    # Now do sanity check of input
+    # Check if there are dupliates in the columns
+    if (any(apply(rankmat, 2, duplicated, incomparables=c(0, NA)))) {
+      stop("an element must only occur exactly once in each list")
+    }
+    # Check and compute the maximum list length
+    ll <- max(c(NROW(rankmat), max(rankmat, na.rm=TRUE)))
+    if (is.null(listlength)) {
+      listlength <- ll
+    }    
+    if (listlength<ll) {
+      stop("You are requesting a list length that is smaller than the largest element found")
+    }    
+    # Should probably also check integers and such
+        
+    # Compute the full metric for each item in a full sized list
+    # The speed here could be increased (for example by converting everything to itemlist scale first)
+    full.list.metric <- function(m, metric) {
+      itemmatrix <- apply(m, 2, order)
+      apply(itemmatrix, 1, metric)
+    }
+    resample <- function(x, ...) x[sample.int(length(x), ...)]
+
+    
+    # General approach
+    # 1) Fill in missing items in each list
+    # 2) Compute the metric for the full lists
+    # 3) Compute agreement 
+    # 4) Finally average everything
+
+    nseq <- seq(listlength)
+    nlists <- NCOL(rankmat)
+  
+
+    # Compute a list of missing items for each list
+    missing.items <- lapply(as.data.frame(rankmat), function(x) { nseq[-x] })
+    
+    fullrankmat <- matrix(rep(0, nlists*listlength), ncol=nlists)
+    fullrankmat[1:NROW(rankmat),] <- rankmat
+    maxdepth <- min(nrow(fullrankmat), ifelse(is.null(depth), nrow(fullrankmat), depth))
+
+      
+    tmpres <- sapply(1:B, function(i) {
+      # Ad 1)
[…]

### 2015-04-10 latex/paper.Rnw
Text

@@
 
-We cannot directly compute $A(p)$ for all predictors because we only observe a censored version of $\tilde R$ for some of the lists.
-Instead we assume that the rank assigned to predictor $p$ in list $l$ is uniformly distributed among the ranks that have been unassigned for list $l$.
-The ranks are clearly not independent since each of the lists essentially contains full set of ranks
+We cannot directly compute $A(p)$ for all predictors because we only
+observe a censored version of $\tilde R$ for some of the
+lists. Instead we assume that the rank assigned to predictor $p$ in
+list $l$ is uniformly distributed among the ranks that have \emph{not}
+been assigned for list $l$.
+
+The rankings within a single list are clearly not independent since each of the lists
+essentially contains full set of ranks
 
@@
 
+<<>>=
+library(SuperRanker)
+@
+
 \subsection{Stability of results}
@@
 Bootstrap across a single method and compare results. Discuss collinearity
+
+\section{Alternatives}
 

### 2015-04-13 latex/paper.Rnw
Small text update

@@
 \title{Sequential rank agreement methods for comparison of ranked lists}
-\author{}
+\author{Claus Thorn Ekstrøm and Thomas Alexander Gerds and Kasper Brink-Jensen}
 
@@
 
-<<>>=
+<<fig=TRUE, echo=FALSE>>=
 library(SuperRanker)
+library(glmnet)
+library(minerva)
+library(changepoint)
+library(randomForestSRC)
+
+# Read data
+library(multtest)
+data(golub)
+
+y <- golub.cl
+x <- golub
+
+producelists <- function(x, y) {
+    index <- seq(1, nrow(golub))
+                                        # d <- data.frame(y, x)
+
+                                        # Marginale t-tests
+    mt.p <- sapply(index, function(i) { t.test(x[i,] ~ y)$p.value } )
+    list1 <- order(mt.p)
+
+                                        # Marginale logreg-tests
+    mlogreg.p <- sapply(index, function(i) { drop1(glm(y ~ x[i,], family=binomial), test="Chisq")[2,5] } )
+    list2 <- order(mlogreg.p)
+
+                                        # Elastic net
+    X <- scale(t(x))
+    enet <- glmnet(X, y, family="binomial", alpha=.8)
+    nyres <- cv.glmnet(X, y, family="binomial", alpha=.8)
+    coefficients <- coef(enet, s=nyres$lambda.1se)[-1]
+    nonzeros <- sum(coefficients!=0)
+    list3 <- order(abs(coefficients), decreasing=TRUE)
+
+                                        # MIC
+    MIC <- sapply(index, function(i) { mine(x[i,], y)$MIC})
+    list4 <- order(MIC, decreasing=TRUE)
+
+                                        # Random Forest
+###    dd <- data.frame(y=factor(y), t(x))
+                                        # dd <- data.frame(y=y, t(x))
+###    f1 <- rfsrc(y ~ ., data=dd, ntree=100)
+###    variables <- abs(f1$importance[,1])
+###    num.undecided <- sum(variables==0)
+###    list5 <- order(variables, decreasing=TRUE)
+###    list5[(length(variables)-num.undecided):length(variables)] <- 0
+
+    cbind(list1,list2,list3,list4)
+}
+
+inputmatrix <- producelists(x, y)
+colnames(inputmatrix) <- c("T", "LogReg", "ElasticNet", "MIC", "RF")[1:ncol(inputmatrix)]
+
+res <- sqrt(sra(inputmatrix))
+
+plot(res[1:1000], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement")
+
+## 
+null <- sqrt(random_list_sra(inputmatrix, B=1, n=100))
+matlines(null, col="gray", lty=1)
+@
+
+<<echo=FALSE, results=hide>>=
+xxx <- as.data.frame(inputmatrix[1:10,])
+xxx
+@
+
+<<label=tab1,echo=FALSE,results=tex>>=
+library(xtable)
+print(xtable(xxx))
+#print(xtable(xxx, caption = "Top 10 list of results", label = "tab:one",
+#digits = c(0, 0, 2, 0, 2, 3, 3)), table.placement = "tbp",
+#caption.placement = "top")
 @

### 2015-04-13 R/InformationRetrieval.R

@@ -1,10 +1,0 @@
-#' 
-#' 
-#' 
-#' 
-#' 
-
-PrecisionAtK <- function()
-{
-  
-}

### 2015-04-14 latex/paper.Rnw
Text

@@
 %
-\newcommand{\ie}{\emph{i.e.,}\@\xspace}
-\newcommand{\eg}{\emph{e.g,}\@\xspace}
+\newcommand{\ie}{\emph{i.e.}\@\xspace}
+\newcommand{\eg}{\emph{e.g.}\@\xspace}
 
@@
 difference in expression across samples. A common research question is
-where to stop, i.e., to decide the maximal significant rank.  In the
+where to stop, \ie, to decide the maximal significant rank.  In the
 three examples this would correspond to the number of significantly
@@
 agreement across a set of lists which soon will enter the state-of-the
-art. The methods should be useful whenever there are multiple rankings
-of the same list. The idea is to define agreement based on the ranks
-of the first k elements in each list.
+art (FIXME: denne sætning skal laves om). The methods should be useful
+whenever there are multiple rankings of the same list. The idea is to
+define agreement based on the ranks of the first $k$ elements in each
+list.
+
+Several other approaches have proposed to measure ... Of these, 
+Kendall's $\tau$ iis the most well-known but xxx and yyy have proposed
+methods that measure the ..
+
+The manuscript is ... as follows
+
+<<echo=FALSE>>=
+#note: always pass alpha on the 0-255 scale
+makeTransparent<-function(someColor, alpha=100)
+{
+  newColor<-col2rgb(someColor)
+  apply(newColor, 2, function(curcoldata){rgb(red=curcoldata[1], green=curcoldata[2],
+    blue=curcoldata[3],alpha=alpha, maxColorValue=255)})
+}
+@ %
 
@@
 observed, \ie, we have the rank of all $P$ items for all of the $L$
-lists. This situation occurs is common when we have the original
-dataset available and when we wish to, say, compare the results from
-different analysis methods.
+lists. Fully observed ranked lists are common and occur, for example,
+when we have a single dataset and apply different statistical analysis
+methods to produce lists of predictors ranked according to their
+importance, or if the same analysis method is applied to data from
+different populations.
 
@@
 rank agreement \eqref{def:sra} as a function of depth $d$. If there is
-a ... An example is seen in Figure~\ref{XX}
+a ... An example is seen in Figure~\ref{fig:example1}
 
@@
 
-<<fig=TRUE, echo=FALSE>>=
+<<example1, echo=FALSE>>=
 library(SuperRanker)
@@
 res <- sqrt(sra(inputmatrix))
+@
 
-plot(res[1:1000], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement")
+\begin{figure}[tb]
+\begin{center}
+<<fig=TRUE,echo=FALSE>>=
+ysize <- 1000
+plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement")
 
 ## 
+# col1 <- makeTransparent()
 null <- sqrt(random_list_sra(inputmatrix, B=1, n=100))
-matlines(null, col="gray", lty=1)
-@
+matlines(null[1:ysize,], col="gray", lty=1)
+
+##
+
+null2 <- sapply(1:40, function(i) {
+    sra(producelists(x, sample(y)), B=1)
+}
+                )
+matlines(null2[1:ysize,], col="lightblue", lty=1)
+
+@ %
+\end{center}
+ \label{fig:example1}
+ \caption{Sequential rank agreement for 4 different analysis methods
+   applied to the Golub data.}
+\end{figure}
 

### 2015-04-14 latex/paper.Rnw
Text

@@
 \begin{abstract}
-
+  The comparison of ranked lists has been studied for several decades
+  but since the advent of large data types it has received increased
+  interest. Computer science and genomics are just two areas that need
+  to compare ranked lists in order to in- vestigate similarities
+  between different analytical methods. Often the aim is to find a
+  measure of similarity for two or more ordered lists that comprise
+  the first k items from each list. Since ordered lists typically are
+  more similar at the top of the list another approach, which is taken
+  in this paper is to ask: for how long are the lists highly similar?
+  Using terminology from agreement analysis we derive the measure
+  depth of agreement using the standard deviation of the ranks, a
+  method that both provides an intuitive interpretation and can be
+  applied to any number of lists even if these are incomplete. The
+  finding are illustrated using simulations and applied to a publicly
+  available cancer dataset.
 \end{abstract}
@@
 of an event, and genetic studies rank genes according to their
-difference in expression across samples. A common research question is
-where to stop, \ie, to decide the maximal significant rank.  In the
-three examples this would correspond to the number of significantly
-associated predictor variables, the number of patients at high risk of
-an event, and the number of genes that are worth to pursue in further
-experiments, respectively.
+difference in expression across samples.
+
+A common research question is where to stop, \ie, to decide the
+maximal significant rank.  In the three examples this would correspond
+to the number of significantly associated predictor variables, the
+number of patients at high risk of an event, and the number of genes
+that are worth to pursue in further experiments, respectively.
 
@@
 
-The manuscript is ... as follows
+The manuscript is organized as follows:
 

### 2015-04-14 latex/paper.Rnw
Text

@@
 methods that measure the ..
+
+Classical pairwise similarity measures have been based on correlations as is the case for Kendall’s τ (Kendall, 1948) and Spearman’s ρ (Spearman, 1910).
+Both correlations measure the similarity of two ranked lists on a correlation scale from -1 to 1, where 1 implies that the lists are completely similar and -1 implies that they are completely dissimilar, in the case of two lists that translates to one being the reverse of the other, while 0 correspond to random lists.
+Correlation based measures treat discrepancies at different depths of the lists equally, so that a difference at the end contributes as much as one in the top of the lists.
+Several modified versions of these measures has been proposed, such as weighted τ by Shieh (1998) where each pair of rankings can be assigned different weights, and the τap (Yilmaz et al., 2008) ap refers to average precision, which places higher emphasis on the top of the lists.
+Spearman’s ρ use the ranks of the variables for calculation of the correlation and the use of ranks are also employed in the M measure of Bar-Ilan et al.
+where the reciprocal rank differences are used to calculate the similarity measure.
+Rather than looking at correlations between lists some recent papers use the inter- section of the lists as the basis for a similarity measure.
+It can be argued that intersection is a better measure when looking at non-conjoint lists, but the simple intersection places equal weights on all depths of the list.
+Thus, Fagin et al.
+(2003) and Webber et al.
+(2010) attempt to modify the simple intersection in a way that is top weighted by using the cu- mulative intersection divided by the depth as the basis for their similarity measures — in the following referred to as the average overlap (AO).Webber et al.
+(2010) define their rank-biased overlap (RBO) by weighting with a converging series thus ensuring the top is always weighted higher than the (potentially indefinite) bottom of the lists.
+The intersection metric by Fagin et al.
+(2003) uses the mean of the overlaps at all depths up to d.
+While it is possible to calculate measures that describe how similar lists are at a given depth with some of the methods described above, the interpretation of the measures is unclear and as we will demonstrate it is not trivial to identify a depth at which the lists are no longer similar.
+We will use some terminology from agreement analysis as our
+2
+method resembles this in some aspects.
+We will refer to the similar part of the lists, which is expected at the top if it exists, as having high agreement and the dissimilar part, where items are ranked randomly, as having low agreement.
+As a measure for agreement we will base this on the limits-of-agreement known from agreement between quantitative variables (Carstensen, 2011) .
+Since the limit-of-agreement essentially is a transformation of the pooled standard deviation of the ranked items we will use that, since it has a direct and simple interpretation: at given depth how far do the items deviate from the mean rank.
+Thus, twice the standard deviation is a measure of how far apart in ranks the methods classify the items.
+The rest of the paper is structured as follows, in the methods section we show how to calculate and apply the limits-of-agreement-based measure, sd, to a small example dataset and we perform a comparison to average overlap.
+We also present and discuss how changepoint methods can be applied to identify where a set of lists goes from high agreement to low agreement.
+The simulation section shows the performance of the mea- sure under a number of different scenarios while the application shows calculations of agreement and average overlap on a publicly available colorectal cancer dataset.
+Finally the findings are discussed along with possible extensions.
 

### 2015-04-20 latex/paper.Rnw
Revision a

@@
 
+Multiple lists. Weigh top over bottom. Censoring. Interpretation
+
 \section{Introduction}
@@
 Several other approaches have proposed to measure ... Of these, 
-Kendall's $\tau$ iis the most well-known but xxx and yyy have proposed
+Kendall's $\tau$ is the most well-known but xxx and yyy have proposed
 methods that measure the ..
 
-Classical pairwise similarity measures have been based on correlations as is the case for Kendall’s τ (Kendall, 1948) and Spearman’s ρ (Spearman, 1910).
-Both correlations measure the similarity of two ranked lists on a correlation scale from -1 to 1, where 1 implies that the lists are completely similar and -1 implies that they are completely dissimilar, in the case of two lists that translates to one being the reverse of the other, while 0 correspond to random lists.
-Correlation based measures treat discrepancies at different depths of the lists equally, so that a difference at the end contributes as much as one in the top of the lists.
-Several modified versions of these measures has been proposed, such as weighted τ by Shieh (1998) where each pair of rankings can be assigned different weights, and the τap (Yilmaz et al., 2008) ap refers to average precision, which places higher emphasis on the top of the lists.
-Spearman’s ρ use the ranks of the variables for calculation of the correlation and the use of ranks are also employed in the M measure of Bar-Ilan et al.
-where the reciprocal rank differences are used to calculate the similarity measure.
-Rather than looking at correlations between lists some recent papers use the inter- section of the lists as the basis for a similarity measure.
-It can be argued that intersection is a better measure when looking at non-conjoint lists, but the simple intersection places equal weights on all depths of the list.
-Thus, Fagin et al.
-(2003) and Webber et al.
-(2010) attempt to modify the simple intersection in a way that is top weighted by using the cu- mulative intersection divided by the depth as the basis for their similarity measures — in the following referred to as the average overlap (AO).Webber et al.
-(2010) define their rank-biased overlap (RBO) by weighting with a converging series thus ensuring the top is always weighted higher than the (potentially indefinite) bottom of the lists.
-The intersection metric by Fagin et al.
-(2003) uses the mean of the overlaps at all depths up to d.
-While it is possible to calculate measures that describe how similar lists are at a given depth with some of the methods described above, the interpretation of the measures is unclear and as we will demonstrate it is not trivial to identify a depth at which the lists are no longer similar.
-We will use some terminology from agreement analysis as our
-2
-method resembles this in some aspects.
-We will refer to the similar part of the lists, which is expected at the top if it exists, as having high agreement and the dissimilar part, where items are ranked randomly, as having low agreement.
-As a measure for agreement we will base this on the limits-of-agreement known from agreement between quantitative variables (Carstensen, 2011) .
-Since the limit-of-agreement essentially is a transformation of the pooled standard deviation of the ranked items we will use that, since it has a direct and simple interpretation: at given depth how far do the items deviate from the mean rank.
-Thus, twice the standard deviation is a measure of how far apart in ranks the methods classify the items.
-The rest of the paper is structured as follows, in the methods section we show how to calculate and apply the limits-of-agreement-based measure, sd, to a small example dataset and we perform a comparison to average overlap.
-We also present and discuss how changepoint methods can be applied to identify where a set of lists goes from high agreement to low agreement.
-The simulation section shows the performance of the mea- sure under a number of different scenarios while the application shows calculations of agreement and average overlap on a publicly available colorectal cancer dataset.
-Finally the findings are discussed along with possible extensions.
+Rank order statistics
+
+Classical pairwise similarity measures have been based on correlations
+as is the case for Kendall’s $\tau$ \citep{kendall1948} and Spearman’s
+$\rho$ \citep{Spearman1910}. Both correlations measure the similarity of two
+ranked lists on a correlation scale from -1 to 1, where 1 implies that
+the lists are completely similar and -1 implies that they are
+completely dissimilar, in the case of two lists that translates to one
+being the reverse of the other, while 0 correspond to random
+lists. Correlation based measures treat discrepancies at different
+depths of the lists equally, so that a difference at the end
+contributes as much as one in the top of the lists.
+
+Kendall's $\tau$ does not make any distinction between errors
+towards the top versus the errors towards the bottom
+
+Several modified versions of these measures has been proposed, such as
+weighted $\tau$ by Shieh (1998) where each pair of rankings can be
+assigned different weights, and the τap (Yilmaz et al., 2008) ap
+refers to average precision, which places higher emphasis on the top
+of the lists. Spearman’s ρ use the ranks of the variables for
+calculation of the correlation and the use of ranks are also employed
+in the M measure of Bar-Ilan et al. where the reciprocal rank
+differences are used to calculate the similarity measure.  Rather than
+looking at correlations between lists some recent papers use the
+inter- section of the lists as the basis for a similarity measure. It
+can be argued that intersection is a better measure when looking at
+non-conjoint lists, but the simple intersection places equal weights
+on all depths of the list. Thus, Fagin et al. (2003) and Webber et
+al. (2010) attempt to modify the simple intersection in a way that is
+top weighted by using the cu- mulative intersection divided by the
+depth as the basis for their similarity measures — in the following
+referred to as the average overlap (AO).Webber et al. (2010) define
+their rank-biased overlap (RBO) by weighting with a converging series
+thus ensuring the top is always weighted higher than the (potentially
+indefinite) bottom of the lists. The intersection metric by Fagin et
+al. (2003) uses the mean of the overlaps at all depths up to d.  While
+it is possible to calculate measures that describe how similar lists
+are at a given depth with some of the methods described above, the
+interpretation of the measures is unclear and as we will demonstrate
+it is not trivial to identify a depth at which the lists are no longer
+similar. We will use some terminology from agreement analysis as our 2
+method resembles this in some aspects. We will refer to the similar
+part of the lists, which is expected at the top if it exists, as
+having high agreement and the dissimilar part, where items are ranked
+randomly, as having low agreement. As a measure for agreement we will
+base this on the limits-of-agreement known from agreement between
+quantitative variables (Carstensen, 2011) . Since the
+limit-of-agreement essentially is a transformation of the pooled
+standard deviation of the ranked items we will use that, since it has
+a direct and simple interpretation: at given depth how far do the
+items deviate from the mean rank. Thus, twice the standard deviation
+is a measure of how far apart in ranks the methods classify the items.
+The rest of the paper is structured as follows, in the methods section
+we show how to calculate and apply the limits-of-agreement-based
+measure, sd, to a small example dataset and we perform a comparison to
+average overlap. We also present and discuss how changepoint methods
+can be applied to identify where a set of lists goes from high
+agreement to low agreement. The simulation section shows the
+performance of the measure under a number of different scenarios
+while the application shows calculations of agreement and average
+overlap on a publicly available colorectal cancer dataset. Finally the
+findings are discussed along with possible extensions.
 
@@
 
-We analyzed these gene expression data using five different
-approaches:
+We analyzed these gene expression data using four different
+approaches: marginal two-sample $t$ tests, marginal logistic
+regression analyses, logistic regression eleastic net, and marginal
+maximum information content correlations (MIC) \citep{}. For the first
+two methods, the genes were ranked according to $p$ value.
 
@@
 res <- sqrt(sra(inputmatrix))
+
 @
@@
 ysize <- 1000
-plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement")
+plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
 
@@
 # col1 <- makeTransparent()
-null <- sqrt(random_list_sra(inputmatrix, B=1, n=100))
-matlines(null[1:ysize,], col="gray", lty=1)
+null <- sqrt(random_list_sra(inputmatrix, B=1, n=400))
+## matlines(null[1:ysize,], col="gray", lty=1)
 
@@
 
-null2 <- sapply(1:40, function(i) {
-    sra(producelists(x, sample(y)), B=1)
-}
-                )
-matlines(null2[1:ysize,], col="lightblue", lty=1)
+##null2 <- sapply(1:40, function(i) {
+##    sra(producelists(x, sample(y)), B=1)
+##} )
+## matlines(null2[1:ysize,], col="lightblue", lty=1)
+
+load("R/fig1null.rda")
+
+bcolor <- makeTransparent("lightblue", alpha=80)
+bcolor2 <- makeTransparent("red", alpha=80)
+
+www <- smooth_sra(null2)
+polygon(c(1:length(www$lower), rev(1:length(www$lower))),
+        c(www$lower, rev(www$upper)),
+        col=bcolor, border=NA)
+
+www <- smooth_sra(null)
+polygon(c(1:length(www$lower), rev(1:length(www$lower))),
+        c(www$lower, rev(www$upper)),
+        col=bcolor2, border=NA)
 
@@
  \caption{Sequential rank agreement for 4 different analysis methods
-   applied to the Golub data.}
+   applied to the Golub data (thick red line). The red and blue shaded
+ areas represent the pointwise 95\% normal areas for the XXX and XXY benchmarks,
+respectively.}
 \end{figure}
@@
 xxx
+
 @
@@
 #caption.placement = "top")
+
 @

### 2015-04-20 R/Rcpp.R

@@ -1,10 +1,0 @@
-library(Rcpp)
-
-
-sourceCpp(file="sra.cpp")
-
-
-nitems <- 20
-m <- cbind(1:nitems, c(1,3,4,2,5,6, sample(7:nitems)))
-
-sra(m, 6, B=1, cens=c(6,6))

### 2015-04-20 R/SuperRanker.R

@@ -1,0 +1,14 @@
+#' Functions related to comparison of ranked lists
+#'
+#' SuperRanker allows you tRoxygen2 allows you to write documentation in comment blocks co-located
+#' with code.
+#'
+#' The only function you're likely to need from \pkg{roxygen2} is
+#' \code{\link{roxygenize}}. Otherwise refer to the vignettes to see
+#' how to format the documentation.
+#'
+#' @docType package
+#' @name SuperRanker
+#' @useDynLib SuperRanker
+#' @importFrom Rcpp sourceCpp
+NULL

### 2015-04-20 R/example1.R

@@ -1,118 +1,0 @@
-#
-#
-# Read data
-#
-#
-library(glmnet)
-library(minerva)
-library(changepoint)
-library(randomForestSRC)
-
-source("sra.R")
-
-# Read data
-library(multtest)
-data(golub)
-
-
-y <- golub.cl
-x <- golub 
-
-
-vs.vimp.md <- function(obj, which.outcome, conservative = TRUE) {
-   msub <- max.subtree(obj, conservative = conservative)
-   vimp <- obj$importance
-   if (is.null(vimp)) {
-       vimp <- matrix(1, nrow(msub$order), ncol(obj$xvar))
-   }
-   names <- rownames(msub$order)
-   names[vimp[, which.outcome] > 0 & msub$order[, 1] <= msub$threshold]    
-} 
-
-producelists <- function(x, y) {
-    index <- seq(1, nrow(golub))
-                                        # d <- data.frame(y, x)
-
-                                        # Marginale t-tests
-    mt.p <- sapply(index, function(i) { t.test(x[i,] ~ y)$p.value } )
-    list1 <- order(mt.p)
-    
-                                        # Marginale logreg-tests
-    mlogreg.p <- sapply(index, function(i) { drop1(glm(y ~ x[i,], family=binomial), test="Chisq")[2,5] } )
-    list2 <- order(mlogreg.p)
-    
-                                        # Elastic net
-    X <- scale(t(x))
-    enet <- glmnet(X, y, family="binomial", alpha=.8)
-    nyres <- cv.glmnet(X, y, family="binomial", alpha=.8)
-    coefficients <- coef(enet, s=nyres$lambda.1se)[-1]
-    nonzeros <- sum(coefficients!=0)
-    list3 <- order(abs(coefficients), decreasing=TRUE)
-    
-                                        # MIC
-    MIC <- sapply(index, function(i) { mine(x[i,], y)$MIC})
-    list4 <- order(MIC, decreasing=TRUE)
-    
-    
-                                        # Random Forest
-    dd <- data.frame(y=factor(y), t(x))
-                                        # dd <- data.frame(y=y, t(x))
-    f1 <- rfsrc(y ~ ., data=dd, ntree=100)
-    variables <- abs(f1$importance[,1])
-    num.undecided <- sum(variables==0)
-    list5 <- order(variables, decreasing=TRUE)
[…]

### 2015-04-20 R/oldsra.R

@@ -1,83 +1,0 @@
-#' @useDynLib SuperRanker
-#' @importFrom Rcpp sourceCpp
-
-
-sra <- function(object, metric=var, depth=NULL, listlength=NULL, B=10) {
-    metric <- match.fun(metric)
-    
-    # Make sure that the input object ends up as a matrix with integer columns all
-    # consisting of elements from 1 and up to listlength
-    if (is.matrix(object))
-        rankmat <- object
-    else
-        rankmat <- as.matrix(do.call("cbind",object))
-    
-    if (missing(depth)) depth <- NROW(rankmat)
-    
-    # Now do sanity check of input
-    # Check if there are dupliates in the columns
-    if (any(apply(rankmat, 2, duplicated, incomparables=c(0, NA)))) {
-      stop("an element must only occur exactly once in each list")
-    }
-    # Check and compute the maximum list length
-    ll <- max(c(NROW(rankmat), max(rankmat, na.rm=TRUE)))
-    if (is.null(listlength)) {
-      listlength <- ll
-    }    
-    if (listlength<ll) {
-      stop("You are requesting a list length that is smaller than the largest element found")
-    }    
-    # Should probably also check integers and such
-        
-    # Compute the full metric for each item in a full sized list
-    # The speed here could be increased (for example by converting everything to itemlist scale first)
-    full.list.metric <- function(m, metric) {
-      itemmatrix <- apply(m, 2, order)
-      apply(itemmatrix, 1, metric)
-    }
-    resample <- function(x, ...) x[sample.int(length(x), ...)]
-
-    
-    # General approach
-    # 1) Fill in missing items in each list
-    # 2) Compute the metric for the full lists
-    # 3) Compute agreement 
-    # 4) Finally average everything
-
-    nseq <- seq(listlength)
-    nlists <- NCOL(rankmat)
-  
-
-    # Compute a list of missing items for each list
-    missing.items <- lapply(as.data.frame(rankmat), function(x) { nseq[-x] })
-    
-    fullrankmat <- matrix(rep(0, nlists*listlength), ncol=nlists)
-    fullrankmat[1:NROW(rankmat),] <- rankmat
-    maxdepth <- min(nrow(fullrankmat), ifelse(is.null(depth), nrow(fullrankmat), depth))
-
-      
-    tmpres <- sapply(1:B, function(i) {
-      # Ad 1)
[…]

### 2015-04-20 R/sra.R

@@ -1,0 +1,110 @@
+#' Compute the sequential rank agreement
+#'
+#' @param obj Either matrix where each column is a ranked list of items or a list of ranked lists of items. Elements are integers between 1 and the length of the lists. The lists should have the same length but censoring can be used by setting the list to zero from a point onwards. See details for more information.
+#' @param B An integer giving the number of randomization to sample over in the case of censored observations
+#' @return A vector of the sequential rank agreement
+#' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
+#' @export
+sra <- function(object, B=1) {
+
+    # Make sure that the input object ends up as a matrix with integer columns all
+    # consisting of elements from 1 and up to listlength
+    if (is.matrix(object))
+        rankmat <- object
+    else
+        rankmat <- as.matrix(do.call("cbind",object))
+
+    listlength <- nrow(rankmat)
+    nlists <- NCOL(rankmat)
+    nseq <- seq(listlength)
+
+    # Special version of sample needed
+    resample <- function(x, ...) x[sample.int(length(x), ...)]
+    
+    # Compute a list of missing items for each list
+    missing.items <- lapply(as.data.frame(rankmat), function(x) { nseq[-x] })
+
+    ## Should make a sanity check that zeros are from a point onwards
+    if (!all(sapply(1:nlists, function(x) {  res <- TRUE
+                                             if (length(missing.items[[x]])>0) {
+                                                 if (any(rankmat[(nitems-missing.items[[x]]):nitems,x]) )
+                                                     { res <- FALSE }
+                                             }
+                                             res
+                                         }))) {
+        stop("Censored ranked lists should be coded 0 from a rank onwards")
+    }
+
+    ## If there is no censoring then we should set B to 1
+    if (max(sapply(missing.items, function(i) length(i) ))==0) {
+        B <- 1
+    }
+    
+    tmpres <- sapply(1:B, function(i) {
+        for (j in 1:nlists) {
+            if (length(missing.items[[j]])>0) {      
+                rankmat[(listlength-length(missing.items[[j]])+1):listlength,j] <- resample(missing.items[[j]])
+            }
+        }
+        res <- sracppfull(rankmat)
+        res
+    })
+
[…]

### 2015-04-20 src/sra.cpp

@@ -45,8 +45,18 @@ NumericVector SdPerRow(NumericMatrix obj) {
 }
 
 
+//' Compute the sequential rank agreement between k ranked lists
+//' 
+//' @description Computes the sequential rank agreement (number of items present in all k lists divided by the current rank) for each rank in the k lists
+//' @param rankMat A matrix with k columns corresponding to the k ranked lists. Elements of each column are integers between 1 and the length of the lists
+//' @param maxlength The maximum depth that are needed XXX
+//' @param B The number of resamples to use in the presence of censored lists
+//' @param cens A vector of integer values that 
+//' @return A vector of the same length as the rows in rankMat containing the sequential rank agreement between the lists for each depth
+//' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
+//' @export
 //[[Rcpp::export]]
-NumericVector sra(IntegerMatrix rankMat, int maxlength, int B, IntegerVector cens) {
+NumericVector sracpp(IntegerMatrix rankMat, int maxlength, int B, IntegerVector cens) {
 
   // The number of lists
   int nLists = rankMat.ncol();
@@ -153,4 +163,67 @@ NumericVector sra(IntegerMatrix rankMat, int maxlength, int B, IntegerVector cen
     return(returnVector);
 }
 
+
+//' Compute the sequential rank agreement between k ranked lists
+//' 
+//' @description Computes the sequential rank agreement (number of items present in all k lists divided by the current rank) for each rank in the k lists
+//' @param rankMat A matrix with k columns corresponding to the k ranked lists. Elements of each column are integers between 1 and the length of the lists
+//' @return A vector of the same length as the rows in rankMat containing the sequential rank agreement between the lists for each depth
+//' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
+//' @export
+//[[Rcpp::export]]
+NumericVector sracppfull(IntegerMatrix rankMat) {
+
+  // The number of lists
+  int nLists = rankMat.ncol();
+  int maxlength=rankMat.nrow();
+
+  // Need a numeric copy of this for the sort function. Doesn't really make sense but
+  // keeping it for now 
+  NumericMatrix x(maxlength, nLists);
+  for (int l=0; l<nLists; l++) {
+    for (int i=0; i<std::min(maxlength, rankMat.nrow()); i++) {
+      x(i, l) = rankMat(i,l);
+    }
+  }
[…]

### 2015-04-20 R/plot.sra.R
Revision a2

@@ -1,6 +1,6 @@
 ##' Plot the agreement between lists as a function of the list depth
 ##' 
-##' @title Plot agreement 
+##' @title Plot sequential rank agreement
 ##' @param x Agreement object
 ##' @param xlim x-axis limits
 ##' @param ylim y-axis limits
@@ -27,13 +27,13 @@
 ##' 
 #' @export
 ##' @author Thomas A. Gerds <tag@@biostat.ku.dk>
-plot.agreement <- function(x,xlim,ylim,xlab="List depth",ylab="Agreement",add=FALSE,...){
-    if (missing(xlim)) xlim=c(1,x$depth)
-    if (missing(ylim)) ylim=c(0,max(x$agreement))
+plot.sra <- function(x, xlim, ylim, xlab="List depth", ylab="Agreement", add=FALSE, ...){
+    if (missing(xlim)) xlim=c(1,length(x))
+    if (missing(ylim)) ylim=c(0,max(x))
     axis1.DefaultArgs <- list()
     axis2.DefaultArgs <- list(side=2)
     plot.DefaultArgs <- list(x=0,y=0,type = "n",ylim = ylim,xlim = xlim,xlab = xlab,ylab = ylab)
-    lines.DefaultArgs <- list(type="l",lwd=3,x=1:x$depth,y=x$agreement)
+    lines.DefaultArgs <- list(type="l",lwd=3,x=1:length(x),y=x)
     smartA <- prodlim::SmartControl(call=list(...),
                                     keys=c("plot","lines","axis1","axis2"),
                                     ignore=c("x","ylim","xlim","xlab","ylab"),

### 2015-04-20 R/sra.R

@@ -53,7 +53,8 @@ sra <- function(object, B=1) {
     agreement <- apply(tmpres, 1, mean)
 
     class(agreement) <- "sra"
-
+    attr(agreement, "B") <- B
+    
     agreement
 }
 

### 2015-04-20 src/RcppExports.cpp
Revision a3

@@ -1,0 +1,43 @@
+// This file was generated by Rcpp::compileAttributes
+// Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
+
+#include <Rcpp.h>
+
+using namespace Rcpp;
+
+// overlap
+NumericVector overlap(IntegerMatrix rankMat);
+RcppExport SEXP SuperRanker_overlap(SEXP rankMatSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject __result;
+    Rcpp::RNGScope __rngScope;
+    Rcpp::traits::input_parameter< IntegerMatrix >::type rankMat(rankMatSEXP);
+    __result = Rcpp::wrap(overlap(rankMat));
+    return __result;
+END_RCPP
+}
+// sracpp
+NumericVector sracpp(IntegerMatrix rankMat, int maxlength, int B, IntegerVector cens);
+RcppExport SEXP SuperRanker_sracpp(SEXP rankMatSEXP, SEXP maxlengthSEXP, SEXP BSEXP, SEXP censSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject __result;
+    Rcpp::RNGScope __rngScope;
+    Rcpp::traits::input_parameter< IntegerMatrix >::type rankMat(rankMatSEXP);
+    Rcpp::traits::input_parameter< int >::type maxlength(maxlengthSEXP);
+    Rcpp::traits::input_parameter< int >::type B(BSEXP);
+    Rcpp::traits::input_parameter< IntegerVector >::type cens(censSEXP);
+    __result = Rcpp::wrap(sracpp(rankMat, maxlength, B, cens));
+    return __result;
+END_RCPP
+}
+// sracppfull
+NumericVector sracppfull(IntegerMatrix rankMat);
+RcppExport SEXP SuperRanker_sracppfull(SEXP rankMatSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject __result;
+    Rcpp::RNGScope __rngScope;
+    Rcpp::traits::input_parameter< IntegerMatrix >::type rankMat(rankMatSEXP);
+    __result = Rcpp::wrap(sracppfull(rankMat));
+    return __result;
+END_RCPP
+}

### 2015-04-20 src/overlap.cpp

@@ -1,0 +1,56 @@
+#include <Rcpp.h>
+using namespace Rcpp;
+
+//' Compute the overlap between k ranked lists
+//' 
+//' @description Computes the overlap (number of items present in all k lists divided by the current rank) for each rank in the k lists
+//' @param rankMat A matrix with k columns corresponding to the k ranked lists. Elements of each column are integers between 1 and the length of the lists
+//' @return A vector of the same length as the rows in rankMat containing the overlap between the lists for each rank
+//' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
+//' @export
+// [[Rcpp::export]]
+NumericVector overlap(IntegerMatrix rankMat) {
+
+  // The number of lists
+  int nLists = rankMat.ncol();
+  int nItems = rankMat.nrow();
+  NumericVector res(nItems);
+
+  LogicalMatrix items(nItems, nLists);
+  IntegerVector pick(nLists);
+  IntegerVector lookat;
+  
+  int i, j, k;
+  double count=0;
+  bool keep;
+ 
+  for (i=0; i<nItems; i++) {
+
+    // Get the items in the ith row
+    for (j=0; j<nLists; j++) {
+      pick(j) = rankMat(i, j)-1;	    
+      items(pick(j), j) = TRUE;
+    }
+    // Now run through the unique items that were updated
+
+    lookat = unique(pick);
+
+    for (j=0; j<lookat.length(); j++) {
+      keep=TRUE;
+      for (k=0; k<nLists; k++) {
+	keep = (keep & items(lookat(j), k));
+	if (!keep)
+	  break;
+      }
+      if (keep) {
+      	count++;
+      }
+    }
+
+    res(i) = count / (i+1);    
+  } 
+    
+    return(res);
+}
+
+

### 2015-04-20 latex/paper.Rnw
Revision a4

@@
 \section{Introduction}
+Ranking of items or results is common in scientific research and
+ranked lists occur as the result of many applications in
+statistics. Regression methods rank predictor variables according to
+magnitude of association with outcome, prediction models rank subjects
+according to their risk of an event, and genetic studies rank genes
+according to their difference in gene expression levels across
+samples.
 
-Ranked lists occur in many applications of statistics. Regression
-methods rank predictor variables according to magnitude of association
-with outcome, prediction models rank subjects according to their risk
-of an event, and genetic studies rank genes according to their
-difference in expression across samples.
+When several ranked lists are available a common research question is
+to what extent the lists agree on the ranking. In other words: is it
+possible to decide on a optimal rank until which the lists agree on
+the items?  A typical situation would arise in high-dimensional
+genomics studies where several analysis methods had been applied to a
+dataset or where a literature search provided differing tables of,
+say, top 10 lists of the most important genes associated to an
+outcome.  Similarly, this would correspond to the number of associated
+predictor variables that are found consistently over the lists, the
+number of patients at high risk of an event in all lists, and the
+number of genes that are worth to pursue in further experiments in the
+three examples above, respectively.
 
-A common research question is where to stop, \ie, to decide the
-maximal significant rank.  In the three examples this would correspond
-to the number of significantly associated predictor variables, the
-number of patients at high risk of an event, and the number of genes
-that are worth to pursue in further experiments, respectively.
+In this article we propose a method to measure agreement among ranked
+lists. The proposed method allows for multiple lists simultaneously,
+provides a dynamic measure of agreement for each depth in the lists,
+places more weight on the top of the list, accommodates censored/incomplete
+lists, and has a natural interpretation that directly relates to the
+ranks. The general idea is to define agreement based on the ranks of
+the first $k$ elements in each list.
 
-In this article we describe some new breakthrough tools for measuring
-agreement across a set of lists which soon will enter the state-of-the
-art (FIXME: denne sætning skal laves om). The methods should be useful
-whenever there are multiple rankings of the same list. The idea is to
-define agreement based on the ranks of the first $k$ elements in each
-list.
+The analysis of rankings is hardly new, and several other approaches
+have proposed a measure of the ``distance'' between two ranked
+lists. Of these, Kendall's $\tau$ \citep{kendall1948} and Spearman’s
+$\rho$ \citep{Spearman1910} are the most well-known. Both correlations
+measure the similarity of two ranked lists on a correlation scale from
+-1 to 1, but neither of them make any distinction between errors
+towards the top versus the errors towards the bottom of the lists.
+Several modified versions of Kendall's $\tau$ and Spearman's $\rho$
+have been proposed. \citet{Shieh1998} proposed a weighted $\tau$ where
+each pair of rankings can be assigned different weights, and
+\citet{Yilmaz2008} proposed the $\tau_{ap}$ which places higher
+emphasis on the top of the lists.
+%
+Spearman’s $\rho$ use the ranks of the variables for calculation of
+the correlation and the use of ranks are also employed in the $M$
+measure of \citet{Bar-Ilan2006} where the reciprocal rank differences
+are used to calculate the similarity measure.
 
-Several other approaches have proposed to measure ... Of these, 
-Kendall's $\tau$ is the most well-known but xxx and yyy have proposed
-methods that measure the ..
+Recent approaches consider the intersection of the lists as the basis
+for a similarity measure
 
-Rank order statistics
+\citetCarterette2009}
 
-Classical pairwise similarity measures have been based on correlations
-as is the case for Kendall’s $\tau$ \citep{kendall1948} and Spearman’s
-$\rho$ \citep{Spearman1910}. Both correlations measure the similarity of two
-ranked lists on a correlation scale from -1 to 1, where 1 implies that
-the lists are completely similar and -1 implies that they are
-completely dissimilar, in the case of two lists that translates to one
-being the reverse of the other, while 0 correspond to random
-lists. Correlation based measures treat discrepancies at different
-depths of the lists equally, so that a difference at the end
-contributes as much as one in the top of the lists.
+Rather than looking at correlations between lists some recent papers
+use the intersection of the lists as the basis for a similarity
+measure. It can be argued that intersection is a better measure when
+looking at non-conjoint lists, but the simple intersection places
+equal weights on all depths of the list. Thus, \citet{Fagin2003} and
+Webber et al. (2010) attempt to modify the simple intersection in a
+way that is top weighted by using the cumulative intersection divided
+by the depth as the basis for their similarity measures — in the
+following referred to as the average overlap (AO). Webber et
+al. (2010) define their rank-biased overlap (RBO) by weighting with a
+converging series thus ensuring the top is always weighted higher than
+the (potentially indefinite) bottom of the lists. The intersection
+metric by Fagin et al. (2003) uses the mean of the overlaps at all
+depths up to d.  While it is possible to calculate measures that
+describe how similar lists are at a given depth with some of the
+methods described above, the interpretation of the measures is unclear
+and as we will demonstrate it is not trivial to identify a depth at
+which the lists are no longer similar. We will use some terminology
+from agreement analysis as our 2 method resembles this in some
+aspects. We will refer to the similar part of the lists, which is
+expected at the top if it exists, as having high agreement and the
+dissimilar part, where items are ranked randomly, as having low
+agreement. As a measure for agreement we will base this on the
+limits-of-agreement known from agreement between quantitative
+variables (Carstensen, 2011) . Since the limit-of-agreement
+essentially is a transformation of the pooled standard deviation of
+the ranked items we will use that, since it has a direct and simple
+interpretation: at given depth how far do the items deviate from the
+mean rank. Thus, twice the standard deviation is a measure of how far
+apart in ranks the methods classify the items.  The rest of the paper
+is structured as follows, in the methods section we show how to
+calculate and apply the limits-of-agreement-based measure, sd, to a
+small example dataset and we perform a comparison to average
+overlap. We also present and discuss how changepoint methods can be
+applied to identify where a set of lists goes from high agreement to
+low agreement. The simulation section shows the performance of the
+measure under a number of different scenarios while the application
+shows calculations of agreement and average overlap on a publicly
+available colorectal cancer dataset. Finally the findings are
+discussed along with possible extensions.
 
-Kendall's $\tau$ does not make any distinction between errors
-towards the top versus the errors towards the bottom
-
-Several modified versions of these measures has been proposed, such as
-weighted $\tau$ by Shieh (1998) where each pair of rankings can be
-assigned different weights, and the τap (Yilmaz et al., 2008) ap
-refers to average precision, which places higher emphasis on the top
-of the lists. Spearman’s ρ use the ranks of the variables for
-calculation of the correlation and the use of ranks are also employed
-in the M measure of Bar-Ilan et al. where the reciprocal rank
-differences are used to calculate the similarity measure.  Rather than
-looking at correlations between lists some recent papers use the
-inter- section of the lists as the basis for a similarity measure. It
-can be argued that intersection is a better measure when looking at
-non-conjoint lists, but the simple intersection places equal weights
-on all depths of the list. Thus, Fagin et al. (2003) and Webber et
-al. (2010) attempt to modify the simple intersection in a way that is
-top weighted by using the cu- mulative intersection divided by the
-depth as the basis for their similarity measures — in the following
-referred to as the average overlap (AO).Webber et al. (2010) define
-their rank-biased overlap (RBO) by weighting with a converging series
-thus ensuring the top is always weighted higher than the (potentially
-indefinite) bottom of the lists. The intersection metric by Fagin et
-al. (2003) uses the mean of the overlaps at all depths up to d.  While
-it is possible to calculate measures that describe how similar lists
-are at a given depth with some of the methods described above, the
-interpretation of the measures is unclear and as we will demonstrate
-it is not trivial to identify a depth at which the lists are no longer
-similar. We will use some terminology from agreement analysis as our 2
-method resembles this in some aspects. We will refer to the similar
-part of the lists, which is expected at the top if it exists, as
-having high agreement and the dissimilar part, where items are ranked
-randomly, as having low agreement. As a measure for agreement we will
-base this on the limits-of-agreement known from agreement between
-quantitative variables (Carstensen, 2011) . Since the
-limit-of-agreement essentially is a transformation of the pooled
-standard deviation of the ranked items we will use that, since it has
-a direct and simple interpretation: at given depth how far do the
-items deviate from the mean rank. Thus, twice the standard deviation
-is a measure of how far apart in ranks the methods classify the items.
-The rest of the paper is structured as follows, in the methods section
-we show how to calculate and apply the limits-of-agreement-based
-measure, sd, to a small example dataset and we perform a comparison to
-average overlap. We also present and discuss how changepoint methods
-can be applied to identify where a set of lists goes from high
-agreement to low agreement. The simulation section shows the
-performance of the measure under a number of different scenarios
-while the application shows calculations of agreement and average
-overlap on a publicly available colorectal cancer dataset. Finally the
-findings are discussed along with possible extensions.
-
-The manuscript is organized as follows:
+The manuscript is organized as follows: In the next section we will
+define ...
 

### 2015-04-21 latex/paper.Rnw
Intro draft

@@
 
-Key words: sequential rank agreement, rank agreement, order statistic,
+Key words: sequential rank agreement, top lists, rank agreement, order statistic,
 methods comparison, variable selection
@@
 
-In this article we propose a method to measure agreement among ranked
-lists. The proposed method allows for multiple lists simultaneously,
-provides a dynamic measure of agreement for each depth in the lists,
-places more weight on the top of the list, accommodates censored/incomplete
-lists, and has a natural interpretation that directly relates to the
-ranks. The general idea is to define agreement based on the ranks of
-the first $k$ elements in each list.
-
 The analysis of rankings is hardly new, and several other approaches
 have proposed a measure of the ``distance'' between two ranked
-lists. Of these, Kendall's $\tau$ \citep{kendall1948} and Spearman’s
-$\rho$ \citep{Spearman1910} are the most well-known. Both correlations
-measure the similarity of two ranked lists on a correlation scale from
--1 to 1, but neither of them make any distinction between errors
-towards the top versus the errors towards the bottom of the lists.
-Several modified versions of Kendall's $\tau$ and Spearman's $\rho$
-have been proposed. \citet{Shieh1998} proposed a weighted $\tau$ where
-each pair of rankings can be assigned different weights, and
-\citet{Yilmaz2008} proposed the $\tau_{ap}$ which places higher
-emphasis on the top of the lists.
+lists. Of these, Kendall's $\tau$ 
+%\citep{kendall1948} 
+and Spearman’s footrule
+%$\rho$ \citep{Spearman1910} 
+are the most well-known. Both measure the similarity of two ranked
+lists but neither of them make any distinction between errors towards
+the top versus the errors towards the bottom of the lists.  Several
+modified versions of Kendall's $\tau$ and Spearman's footrule have been
+proposed. \citet{Shieh1998} proposed a weighted $\tau$ where each pair
+of rankings can be assigned different weights, and \citet{Yilmaz2008}
+proposed the $\tau_{ap}$ which places higher emphasis on the top of
+the lists.
 %
-Spearman’s $\rho$ use the ranks of the variables for calculation of
-the correlation and the use of ranks are also employed in the $M$
+Spearman’s footrule use the ranks of the variables for calculation of
+the distance and the use of ranks are also employed in the $M$
 measure of \citet{Bar-Ilan2006} where the reciprocal rank differences
@@
 
-Recent approaches consider the intersection of the lists as the basis
-for a similarity measure
+More recent approaches consider the intersection of lists as the basis
+for a similarity measure. However, simple intersections also places
+equal weights on all depths of the list which led \citet{Fagin2003}
+and \citet{Webber2010} to modify the simple intersection to put more
+emphasis on the top of the lists by using the cumulative intersection
+divided by the depth as their similarity measures (denoted the average
+overlap (AO)). \citet{Webber2010} define their rank-biased overlap
+(RBO) by weighting with a converging series thus ensuring the top is
+always weighted higher than the (potentially indefinite) bottom of the
+lists.  While it is possible to calculate measures that describe how
+similar lists are at a given depth with some of the methods described
+above, the interpretation of the measures is unclear --- especially
+with multiple lists --- and it is difficult to identify a depth at
+which the lists are no longer similar.
 
-\citetCarterette2009}
-
-Rather than looking at correlations between lists some recent papers
-use the intersection of the lists as the basis for a similarity
-measure. It can be argued that intersection is a better measure when
-looking at non-conjoint lists, but the simple intersection places
-equal weights on all depths of the list. Thus, \citet{Fagin2003} and
-Webber et al. (2010) attempt to modify the simple intersection in a
-way that is top weighted by using the cumulative intersection divided
-by the depth as the basis for their similarity measures — in the
-following referred to as the average overlap (AO). Webber et
-al. (2010) define their rank-biased overlap (RBO) by weighting with a
-converging series thus ensuring the top is always weighted higher than
-the (potentially indefinite) bottom of the lists. The intersection
-metric by Fagin et al. (2003) uses the mean of the overlaps at all
-depths up to d.  While it is possible to calculate measures that
-describe how similar lists are at a given depth with some of the
-methods described above, the interpretation of the measures is unclear
-and as we will demonstrate it is not trivial to identify a depth at
-which the lists are no longer similar. We will use some terminology
-from agreement analysis as our 2 method resembles this in some
-aspects. We will refer to the similar part of the lists, which is
-expected at the top if it exists, as having high agreement and the
-dissimilar part, where items are ranked randomly, as having low
-agreement. As a measure for agreement we will base this on the
-limits-of-agreement known from agreement between quantitative
-variables (Carstensen, 2011) . Since the limit-of-agreement
-essentially is a transformation of the pooled standard deviation of
-the ranked items we will use that, since it has a direct and simple
-interpretation: at given depth how far do the items deviate from the
-mean rank. Thus, twice the standard deviation is a measure of how far
-apart in ranks the methods classify the items.  The rest of the paper
-is structured as follows, in the methods section we show how to
-calculate and apply the limits-of-agreement-based measure, sd, to a
-small example dataset and we perform a comparison to average
-overlap. We also present and discuss how changepoint methods can be
-applied to identify where a set of lists goes from high agreement to
-low agreement. The simulation section shows the performance of the
-measure under a number of different scenarios while the application
-shows calculations of agreement and average overlap on a publicly
-available colorectal cancer dataset. Finally the findings are
-discussed along with possible extensions.
+In this article we propose a method to measure similarity among ranked
+lists for each depth in the list. As a measure for similarity we will
+use the limits of agreement known from agreement between quantitative
+variables \citep{Carstensen2011} but will compute the agreement on the
+ranks of the items.  The proposed method allows for multiple lists
+simultaneously, provides a dynamic measure of agreement for each depth
+in the lists, places more weight on the top of the list, accommodates
+censored/incomplete lists of varying lengths, and has a natural
+interpretation that directly relates to the ranks. The general idea is
+to define agreement based on the sequence of ranks from the first $k$
+elements in each list. This approach allows us to estimate a potential
+change point in the agreement in order to determine an optimal $k$
+where there is a shift in the agreement among lists and enables us to
+use randomization approaches to evaluate the rank agreement. In this
+sense it is a combination and generalization of some of the ideas of
+\citet{Carterette2009} and \citet{Boulesteix2009}. The former compares
+two rankings based on the distance between them as measured by a
+multivariate Gaussian distribution and the latter presents an overview
+of approaches for aggregation of ranked lists including bootstrap and
+leave-one-out jackknife approaches.
 
 The manuscript is organized as follows: In the next section we will
-define ...
+define the sequential rank agreement among multiple ranked lists and
+discuss how to handle incomplete/censored lists. In section 3 we
+present and discuss approaches to evaluate the sequential rank
+agreement obtained from an experiment and to identify the optimal
+depth $k$ for which the ranked lists agree. Finally we present three
+quite different applications of the proposed sequential rank agreement
+approach to real data before we discuss the findings along with
+possible extensions.
 
@@
 mappings. Panels (a) and (b) of Table~\ref{tab:example} show a
-schematic example of these mappings.
+schematic example of these mappings. Thus if $\pi_l(1)=X_{34}$ then
+item $X_{34}$ is ranked first in list $l$.
 
@@
 
-\subsection{All lists fully observed}
+\subsection{Agreement among fully observed lists}
 We shall start by the simplest case where all $L$ lists are fully
@@
 
-\subsection{Analysis of top $k$ lists}
+\subsection{Analysis of incomplete/censored lists}
 Not uncommon for lists to be

### 2015-04-21 latex/paper.Rnw

@@
 use the limits of agreement known from agreement between quantitative
-variables \citep{Carstensen2011} but will compute the agreement on the
+variables \citep{Carstensen2010} but will compute the agreement on the
 ranks of the items.  The proposed method allows for multiple lists
@@
 schematic example of these mappings. Thus if $\pi_l(1)=X_{34}$ then
-item $X_{34}$ is ranked first in list $l$.
+item $X_{34}$ is ranked first in list $l$, and similarly $R_l(X_{34})=1$.
 
@@
 regression analyses, logistic regression eleastic net, and marginal
-maximum information content correlations (MIC) \citep{}. For the first
-two methods, the genes were ranked according to $p$ value.
+maximum information content correlations (MIC) \citep{Reshef2011}. For
+the first two methods, the genes were ranked according to $p$ value.
 

### 2015-04-21 latex/paper.Rnw
Fix for censoring

@@
 producelists <- function(x, y) {
+    nitems <- length(y)
     index <- seq(1, nrow(golub))
@@
     list3 <- order(abs(coefficients), decreasing=TRUE)
-
+    if (nonzeros==0) {
+        list3[(nonzeros+1):nitems] <- 0
+    } 
                                         # MIC
@@
 
-res <- sqrt(sra(inputmatrix))
+res <- sqrt(sra(inputmatrix, B=20))
+inputmatrix2 <- inputmatrix
+inputmatrix2[51:nrow(inputmatrix),] <- 0
+res2 <- sqrt(sra(inputmatrix2, B=20))
 
@@
 plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
-
+lines(res2[1:ysize], lwd=2, col="purple")
 ## 

### 2015-04-22 latex/paper.Rnw
Initial changepoint analysis

@@
 null <- sqrt(random_list_sra(inputmatrix, B=1, n=400))
+null3 <- sqrt(random_list_sra(inputmatrix, B=20, n=200))
 ## matlines(null[1:ysize,], col="gray", lty=1)
@@
 bcolor2 <- makeTransparent("red", alpha=80)
+bcolor3 <- makeTransparent("green", alpha=80)
 
@@
         col=bcolor2, border=NA)
+
+www <- smooth_sra(null3)
+polygon(c(1:length(www$lower), rev(1:length(www$lower))),
+        c(www$lower, rev(www$upper)),
+        col=bcolor3, border=NA)
 
@@
 
+<<>>=
+library(changepoint)
+cpt.mean(res)
+cpt.mean(res2)
+@
+
 \subsection{Stability of results}

### 2015-04-24 latex/paper.Rnw
Initial changepoint analysis
Full lists

@@
 
-For the fully obseved list situation we can plot the sequantial
-rank agreement \eqref{def:sra} as a function of depth $d$. If there is
-a ... An example is seen in Figure~\ref{fig:example1}
+For the fully obseved list situation we can plot the sequential rank
+agreement \eqref{def:sra} as a function of depth $d$.  An example is
+seen in Figure~\ref{fig:example1} where smaller values on the $y$ axis
+correspond to better agreement among the lists than larger values. The
+sequential rank agreement at depth $d$ can be interpreted as how many
+ranks apart do we on average expect the items found until depth $d$ to
+be.
+
+If the lists generally agree on all items then we expect the
+sequential rank agreement curve to be flat, but if there are changes
+in the rank agreement levels then this suggests that there are changes
+in the levels of agreement. In particular, we expect the agreement
+among the lists to be better towards the top of the lists and worse
+towards the end of the lists.
+
+\texttt{FIXME: Skal der et specifikt eksempel ind her eller kan vi
+  bare bruge det fra senere?}
 
@@
 Noget med
+
+FIXME: For censored lists we expect a change point around the depth
+where the lists are censored. This is an artefact stemming from the
+fact that
 
@@
 \begin{center}
-<<fig=TRUE,echo=FALSE>>=
+<<fig=TRUE,echo=FALSE,width=3>>=
 ysize <- 1000
@@
 @ %
+
+<<fig=TRUE,echo=FALSE,width=3>>=
+ysize <- 100
+plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
+lines(res2[1:ysize], lwd=2, col="purple")
+## 
+# col1 <- makeTransparent()
+null <- sqrt(random_list_sra(inputmatrix, B=1, n=400))
+null3 <- sqrt(random_list_sra(inputmatrix, B=20, n=200))
+## matlines(null[1:ysize,], col="gray", lty=1)
+
+##
+
+##null2 <- sapply(1:40, function(i) {
+##    sra(producelists(x, sample(y)), B=1)
+##} )
+## matlines(null2[1:ysize,], col="lightblue", lty=1)
+
+load("R/fig1null.rda")
+
+bcolor <- makeTransparent("lightblue", alpha=80)
+bcolor2 <- makeTransparent("red", alpha=80)
+bcolor3 <- makeTransparent("green", alpha=80)
+
+www <- smooth_sra(null2)
+polygon(c(1:length(www$lower), rev(1:length(www$lower))),
+        c(www$lower, rev(www$upper)),
+        col=bcolor, border=NA)
+
+www <- smooth_sra(null)
+polygon(c(1:length(www$lower), rev(1:length(www$lower))),
+        c(www$lower, rev(www$upper)),
+        col=bcolor2, border=NA)
+
+www <- smooth_sra(null3)
+polygon(c(1:length(www$lower), rev(1:length(www$lower))),
+        c(www$lower, rev(www$upper)),
+        col=bcolor3, border=NA)
+
+@ %
 \end{center}
- \label{fig:example1}
  \caption{Sequential rank agreement for 4 different analysis methods
@@
 respectively.}
+ \label{fig:example1}
 \end{figure}
@@
 library(changepoint)
-cpt.mean(res)
+cpt.mean(res, method="BinSeg")
 cpt.mean(res2)

### 2015-04-26 R/sra.R
fixed minor bugs

@@ -13,11 +13,11 @@ sra <- function(object, B=1) {
         rankmat <- object
     else
         rankmat <- as.matrix(do.call("cbind",object))
-
+    
     listlength <- nrow(rankmat)
     nlists <- NCOL(rankmat)
     nseq <- seq(listlength)
-
+    
     # Special version of sample needed
     resample <- function(x, ...) x[sample.int(length(x), ...)]
     
@@ -25,13 +25,14 @@ sra <- function(object, B=1) {
     missing.items <- lapply(as.data.frame(rankmat), function(x) { nseq[-x] })
 
     ## Should make a sanity check that zeros are from a point onwards
-    if (!all(sapply(1:nlists, function(x) {  res <- TRUE
-                                             if (length(missing.items[[x]])>0) {
-                                                 if (any(rankmat[(nitems-missing.items[[x]]):nitems,x]) )
-                                                     { res <- FALSE }
-                                             }
-                                             res
-                                         }))) {
+    if (!all(sapply(1:nlists, function(x) {
+                        res <- TRUE
+                        if (length(missing.items[[x]])>0) {
+                            if (any(rankmat[(listlength-missing.items[[x]]):listlength,x]) )
+                                { res <- FALSE }
+                        }
+                        res
+                    }))) {
         stop("Censored ranked lists should be coded 0 from a rank onwards")
     }
 
@@ -87,9 +88,6 @@ random_list_sra <- function(object, B=1, n=1) {
     res
     
 }
-
-
-
 
 
 

### 2015-04-26 R/sra.R
new interface for arbitrary lists

@@ -1,62 +1,120 @@
 #' Compute the sequential rank agreement
 #'
-#' @param obj Either matrix where each column is a ranked list of items or a list of ranked lists of items. Elements are integers between 1 and the length of the lists. The lists should have the same length but censoring can be used by setting the list to zero from a point onwards. See details for more information.
-#' @param B An integer giving the number of randomization to sample over in the case of censored observations
+#' @param object Either matrix where each column is a ranked list of
+#' items or a list of ranked lists of items. Elements are integers
+#' between 1 and the length of the lists. The lists should have the
+#' same length but censoring can be used by setting the list to zero
+#' from a point onwards. See details for more information.
+#' @param cens.code Code for missing items 
+#' @param B An integer giving the number of randomization to sample
+#' over in the case of censored observations
 #' @return A vector of the sequential rank agreement
-#' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
+##' @examples
+##' 
+##' alist <- list(a=1:8,b=sample(1:8),c=sample(1:8))
+##' sra(alist)
+##' 
+##' blist <- list(x1=letters,x2=sample(letters),x3=sample(letters))
+##' sra(blist)
+##' 
+##' ## censored lists are either too short
+##' clist <- list(x1=c("a","b","c","d","e","f","g","h"),
+##'               x2=c("h","c","f","g","b"),
+##'               x3=c("d","e","a"))
+##' set.seed(17)
+##' sra(clist,cens.code="z",B=10)
+##' 
+##' ## or use a special code for missing elements
+##' Clist <- list(x1=c("a","b","c","d","e","f","g","h"),
+##'               x2=c("h","c","f","g","b","z","z","z"),
+##'               x3=c("d","e","a","z","z","z","z","z"))
+##' set.seed(17)
+##' sra(Clist,cens.code="z",B=10)
+##' 
+#' @author Claus Ekstrøm <ekstrom@@sund.ku.dk> and Thomas A Gerds <tag@@biostat.ku.dk>
+#' 
 #' @export
-sra <- function(object, B=1) {
-
+sra <- function(object, cens.code=NA, B=1) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength
     if (is.matrix(object))
-        rankmat <- object
+        object <- apply(object,2,function(x)x)
     else
-        rankmat <- as.matrix(do.call("cbind",object))
[…]

### 2015-04-26 R/sra.R
rename cens.code to na.strings

@@ -5,7 +5,7 @@
 #' between 1 and the length of the lists. The lists should have the
 #' same length but censoring can be used by setting the list to zero
 #' from a point onwards. See details for more information.
-#' @param cens.code Code for missing items 
+#' @param na.strings Code for missing items 
 #' @param B An integer giving the number of randomization to sample
 #' over in the case of censored observations
 #' @return A vector of the sequential rank agreement
@@ -22,19 +22,19 @@
 ##'               x2=c("h","c","f","g","b"),
 ##'               x3=c("d","e","a"))
 ##' set.seed(17)
-##' sra(clist,cens.code="z",B=10)
+##' sra(clist,na.strings="z",B=10)
 ##' 
 ##' ## or use a special code for missing elements
 ##' Clist <- list(x1=c("a","b","c","d","e","f","g","h"),
 ##'               x2=c("h","c","f","g","b","z","z","z"),
 ##'               x3=c("d","e","a","z","z","z","z","z"))
 ##' set.seed(17)
-##' sra(Clist,cens.code="z",B=10)
+##' sra(Clist,na.strings="z",B=10)
 ##' 
 #' @author Claus Ekstrøm <ekstrom@@sund.ku.dk> and Thomas A Gerds <tag@@biostat.ku.dk>
 #' 
 #' @export
-sra <- function(object, cens.code=NA, B=1) {
+sra <- function(object, na.strings=NA, B=1) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength
     if (is.matrix(object))
@@ -47,9 +47,9 @@ sra <- function(object, cens.code=NA, B=1) {
     ## sanity checks
     object <- lapply(object,function(x){
                          out <- which(is.na(x))
-                         cens.code <- cens.code[!is.na(cens.code)]
-                         if (length(cens.code)>0)
-                             out <- c(out,grep(paste0("^",cens.code,"$"),x))
+                         na.strings <- na.strings[!is.na(na.strings)]
+                         if (length(na.strings)>0)
+                             out <- c(out,grep(paste0("^",na.strings,"$"),x))
                          ## remove censored items with side effect:
                          ## in case where all lists have trailing censored information
                          ## this is pruned

### 2015-04-27 latex/paper.Rnw
now with description of two benchmarks

@@
 \section{Introduction}
+
 Ranking of items or results is common in scientific research and
@@
 
-Three approaches
+To interpretate the sequential rank agreement scale we propose by
+several different benchmark values. Our data free benchmark value is
+obtained by computing the sequential rank agreement of $L$ lists each
+containing a permutation of $P$ items. Practically this benchmark
+value can be obtained by averaging the result of repeated computer
+experiments. In the uncensored case each experiment contains $L$ lists
+of random permutations of the numbers $1,\dots,P$. In the censored
+case only the uncensored part of the orginal lists is randomly
+permuted.
+
+In our experience XX repetitions are sufficient to reduce Monte Carlo
+error. Table Appendix A shows benchmark values obtained with
+
+Our no-information benchmark can be applied in situations where each
+list is an importance ranking of $P$ predictors resulting from
+regression analyses. The idea is to repeat the regression analyses in
+modified data where the outcome vector is randomly permuted. This
+artificially removes any potential information regarding the
+regression that the real data may contain.
+
+- third approach?
 

### 2015-04-27 R/sra.R

@@ -142,7 +142,7 @@ random_list_sra <- function(object, B=1, n=1) {
         for (j in 1:ncol(object)) {
             object[,j] <- c(sample(nitems, size=notmiss[j]), rep(0, nitems-notmiss[j]))
         }
-        sra(object, B=B) 
+        sra(object, na.strings="0", B=B) 
     })
     res
     

### 2015-04-27 latex/paper.Rnw
Manuskript

--- 2015-04-24 latex/paper.Rnw
@@
 
+\SweaveOpts{prefix.string=paper, keep.source=TRUE}
+
 \begin{document}
-%\SweaveOpts{concordance=TRUE}
 
@@
 
-\texttt{FIXME: Skal der et specifikt eksempel ind her eller kan vi
-  bare bruge det fra senere?}
+\begin{figure}[tb]
+\begin{center}
+\includegraphics[width=.45\textwidth]{paper-fig1a}
+\includegraphics[width=.45\textwidth]{paper-fig1b}
+\end{center}
+ \caption{Sequential rank agreement for 4 different analysis methods
+   applied to the Golub data (thick red line). The red and blue shaded
+ areas represent the pointwise 95\% normal areas for the XXX and XXY benchmarks,
+respectively.}
+ \label{fig:example1}
+\end{figure}
 
 \subsection{Analysis of incomplete/censored lists}
-Not uncommon for lists to be
+Incomplete lists are also a common occurrence that arise from, for
+example, comparison of top $k$ lists, when some methods only rank some
+of the items (e.g., penalized regression based on the Lasso provides a
+sparse set of predictors that have non-zero coefficients. There is no
+obvious ordering of the predictors that have been shrunk to zero), or
+if
+
+Top $k$, methods only provide information on some ranking (Lasso),
+only interested in significant ...predictors
 
@@
 
-\begin{figure}[tb]
-\begin{center}
-<<fig=TRUE,echo=FALSE,width=3>>=
+<<fig1a, fig=TRUE,echo=FALSE,include=FALSE>>=
+ysize <- 1000
+plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
+lines(res2[1:ysize], lwd=2, col="purple")
+
+@ %
+
+<<fig1b, fig=TRUE,echo=FALSE,include=FALSE >>=
+ysize <- 80
+plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
+lines(res2[1:ysize], lwd=2, col="purple")
+
+@ %
+
+<<fig2a, fig=TRUE,echo=FALSE,include=FALSE>>=
 ysize <- 1000
@@
 
-<<fig=TRUE,echo=FALSE,width=3>>=
-ysize <- 100
+<<fig2b, fig=TRUE,echo=FALSE,include=FALSE >>=
+ysize <- 80
 plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
@@
 @ %
+
+\begin{figure}[tb]
+\begin{center}
+\includegraphics[width=.45\textwidth]{paper-fig2a}
+\includegraphics[width=.45\textwidth]{paper-fig2b}
 \end{center}
@@
 library(changepoint)
-cpt.mean(res, method="BinSeg")
-cpt.mean(res2)
+cp1 <- cpt.mean(res, method="BinSeg")
+cp2 <- cpt.mean(res2, method="BinSeg")
+cp1@cpts
+cp1@param.est
+cp2@cpts
+cp2@param.est
 @

### 2015-05-05 latex/paper.Rnw
Figure 1

@@
 \usepackage[noae]{Sweave}
+\usepackage{authblk}
 
@@
 
-\title{Sequential rank agreement methods for comparison of ranked lists}
-\author{Claus Thorn Ekstrøm and Thomas Alexander Gerds and Kasper Brink-Jensen}
+\title{Sequential rank agreement methods for comparison of ranked
+  lists}
+\author{Claus Thorn Ekstrøm}
+\author{Thomas Alexander Gerds}
+\author{Kasper Brink-Jensen}
+\affil{Biostatistics, University of Copenhagen}
+
+% Use square brackets for numbering affiliations
 
@@
 \subsection{Agreement among fully observed lists}
-We shall start by the simplest case where all $L$ lists are fully
-observed, \ie, we have the rank of all $P$ items for all of the $L$
-lists. Fully observed ranked lists are common and occur, for example,
-when we have a single dataset and apply different statistical analysis
-methods to produce lists of predictors ranked according to their
+The simplest case occurs when all $L$ lists are fully
+observed, \ie, we have observed the rank of all $P$ items for all $L$
+lists. Fully observed ranked lists are common and arise, for example,
+when different statistical analysis methods are applied to a single
+dataset to produce lists of predictors ranked according to their
 importance, or if the same analysis method is applied to data from
@@
 
-For the fully obseved list situation we can plot the sequential rank
-agreement \eqref{def:sra} as a function of depth $d$.  An example is
-seen in Figure~\ref{fig:example1} where smaller values on the $y$ axis
-correspond to better agreement among the lists than larger values. The
-sequential rank agreement at depth $d$ can be interpreted as how many
-ranks apart do we on average expect the items found until depth $d$ to
-be.
+With fully observed lists we can plot the sequential rank agreement
+\eqref{def:sra} as a function of depth $d$.  An example is seen in the
+left panel of Figure~\ref{fig:example1} where four analysis methods
+were used to rank genes (for further detail see the application
+section below). The sequential rank agreement at depth $d$ can be
+interpreted as how many ranks apart do we on average expect the items
+found until depth $d$ to be so smaller values on the $y$ axis
+correspond to better agreement among the lists than larger values.
 
@@
 sequential rank agreement curve to be flat, but if there are changes
-in the rank agreement levels then this suggests that there are changes
-in the levels of agreement. In particular, we expect the agreement
-among the lists to be better towards the top of the lists and worse
-towards the end of the lists.
+in the levels of rank agreement then this suggests that there are changes
+in the levels of agreement. In particular, we generally expect the
+agreement among the lists to be better towards the top of the lists
+and worse towards the end of the lists, and in this case the sequential rank
+agreement curve will start at a low level, and then have an increase
+before it levels off.
 
@@
 \end{center}
- \caption{Sequential rank agreement for 4 different analysis methods
-   applied to the Golub data (thick red line). The red and blue shaded
- areas represent the pointwise 95\% normal areas for the XXX and XXY benchmarks,
-respectively.}
+ \caption{Left panel: Sequential rank agreement for 4 different analysis methods
+   applied to the 3051 genes in the Golub data (solid line). The dashed line shows the
+   corresponding estimated sequential rank agreement for censored
+   lists based purely on the top 50 ranked genes. Right panel contains
+   the same information but has zoomed in on the first part of the $x$-axis.}
  \label{fig:example1}
@@
 respectively.}
- \label{fig:example1}
+ \label{fig:case1}
 \end{figure}

### 2015-05-06 latex/paper.Rnw
Rettelser til tekst

@@
 Consider a set of $P$ different items $X=\{X_1,\dots,X_P\}$. An
-ordered list is a permutation function, $R: \{1,\dots,P\}\to
+ordered list is a permutation function, $R: \{X_1,\dots,X_P\}\to
 \{1,\dots,P\}$, such that $R(X_p)$ is the rank of item $X_p$ in the
@@
 
-The agreement of the lists regarding the rank given to an item can be
+The agreement of the lists regarding the rank given to an item $X_p$ can be
 measured by
 \begin{equation}
-  A(p) = f(R_1(p), \ldots, R_L(p)),
+  A(X_p) = f(R_1(X_p), \ldots, R_L(X_p)),
 \end{equation}
@@
 sample standard error as our function $f$ and hence use
-$$A(p) = \sqrt{\frac{\sum_{i=1}^L (R_i(p) - \bar{R}(p))^2}{L-1}},
+$$A(X_p) = \sqrt{\frac{\sum_{i=1}^L (R_i(X_p) - \bar{R}(X_p))^2}{L-1}},
 $$
@@
 \textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in
-      S_d\}}(L-1)A(p)^2}{(L-1)|S_d|}}, \label{def:sra}
+      S_d\}}(L-1)A(X_p)^2}{(L-1)|S_d|}}, \label{def:sra}
 \end{equation}
@@
 sequential rank agreement curve to be flat, but if there are changes
-in the levels of rank agreement then this suggests that there are changes
-in the levels of agreement. In particular, we generally expect the
-agreement among the lists to be better towards the top of the lists
-and worse towards the end of the lists, and in this case the sequential rank
+in the levels of rank agreement then this suggests that there are sets
+of items that the lists agree on and other sets where they rank the
+items differently. In particular, we generally expect the agreement
+among the lists to be better towards the top of the lists and worse
+towards the end of the lists, and in this case the sequential rank
 agreement curve will start at a low level, and then have an increase
@@
 Incomplete lists are also a common occurrence that arise from, for
-example, comparison of top $k$ lists, when some methods only rank some
-of the items (e.g., penalized regression based on the Lasso provides a
+example, comparison of top $k$ lists, or when some methods only rank a
+subset of the items (e.g., penalized regression based on the Lasso provides a
 sparse set of predictors that have non-zero coefficients. There is no
-obvious ordering of the predictors that have been shrunk to zero), or
-if
+obvious ordering of the predictors that have all been shrunk to zero).
 
-Top $k$, methods only provide information on some ranking (Lasso),
-only interested in significant ...predictors
-
-censored
-
-Let $\Lambda_l, l=1, \ldots, L$ be the set of items found in list $l$
-so $\Lambda_l$ is the top $k_l$ list of items from list $l$ where $k_l
-= |\Lambda_l|$. Note that we observe the top $k$ items for each of the
-$L$ lists if $k_1=\cdots=k_L=k$. For censored lists the rank function
-becomes
+Sequential rank agreement can be generalized to incomplete/censored
+lists in the following way.  Let $\Lambda_l, l=1, \ldots, L$ be the
+set of items found in list $l$ so $\Lambda_l$ is the top $k_l$ list of
+items from list $l$ where $k_l = |\Lambda_l|$. Note that if
+$k_1=\cdots=k_L=k$ then we observe the top $k$ items for each of the
+$L$ lists. For censored lists the rank function becomes
 \begin{equation}
-\tilde R_l(p) = \left\{\begin{array}{cl} \{\pi_l^{-1}(p)\} & \text{ for } p\in \Lambda_l \\ 
+\tilde R_l(X_p) = \left\{\begin{array}{cl} \{\pi_l^{-1}(p)\} & \text{ for } p\in \Lambda_l \\ 
 \{k_l+1,\dots,P\} & \text{ for } p \not\in \Lambda_l\end{array}\right.
 \end{equation}
-where we only know that the rank for the unobserved items in list $l$ must be larger than the largest rank observed in that list.
+where we only know that the rank for the unobserved items in list $l$
+must be larger than the largest rank observed in that list.
+Unless we have additional prior knowledge about the distribution of the
+unranked items then we can restrict attention of the sequential rank
+agreement to 
+\begin{equation}
+d \leq \max(k_1, \ldots, k_L),
+\end{equation}
+since we have no information about the ordering of the items after a
+depth $d$ where all lists are censored.
 
-In the case of censored lists it is sufficient (FIXME: requires
-argument) to look at depths where we have corresponding observations
-so the largest rank we should consider will be
+The agreement, $A(X_p)$, cannot be computed directly for all
+predictors because we only observe a censored version of $\tilde R$
+for some of the lists. Instead we assume that the rank assigned to
+item $X_p$ in list $l$ is uniformly distributed among the ranks that
+have \emph{not} been assigned for list $l$,
 \begin{equation}
-d \leq \max(k_1, \ldots, k_L).
+\tilde A(X_p) = \frac{\sum_{r_1; r_1\in \tilde R_1(X_p)}  \cdots \sum_{r_L; r_L\in \tilde R_L(X_p)} A(X_p)}{\prod_l |\tilde R_l(X_p)|}.
 \end{equation}
 
-We cannot directly compute $A(p)$ for all predictors because we only
-observe a censored version of $\tilde R$ for some of the
-lists. Instead we assume that the rank assigned to predictor $p$ in
-list $l$ is uniformly distributed among the ranks that have \emph{not}
-been assigned for list $l$.
+The rankings within a single list are clearly not independent since
+each of the lists essentially contains full set of ranks and by
 
-The rankings within a single list are clearly not independent since each of the lists
-essentially contains full set of ranks
-
-\begin{equation}
-\tilde A(p) = \frac{\sum_{r_1; r_1\in \tilde R_1(p)}  \cdots \sum_{r_L; r_L\in \tilde R_L(p)} A(p)}{\prod_l |\tilde R_l(p)|}
-\end{equation}
-
-FIXME: If instead of running through all elements of $\tilde R_1(p)$
-one would use the average rank in $\tilde R_1(p)$ we would end up with
+FIXME: If instead of running through all elements of $\tilde R_1(X_p)$
+one would use the average rank in $\tilde R_1(X_p)$ we would end up with
 a too small variance.
-
-Noget med
 

### 2015-05-06 latex/paper.Rnw
Rettelser til tekst

@@
 item $X_p$ in list $l$ is uniformly distributed among the ranks that
-have \emph{not} been assigned for list $l$,
+have \emph{not} been assigned for list $l$, so the augmented agreement
+becomes
 \begin{equation}
-\tilde A(X_p) = \frac{\sum_{r_1; r_1\in \tilde R_1(X_p)}  \cdots \sum_{r_L; r_L\in \tilde R_L(X_p)} A(X_p)}{\prod_l |\tilde R_l(X_p)|}.
+\tilde A(X_p) = \frac{\sum_{r_1; r_1\in \tilde R_1(X_p)}  \cdots
+  \sum_{r_L; r_L\in \tilde R_L(X_p)} A(X_p)}{\prod_l |\tilde
+  R_l(X_p)|}. \label{formula:censored-agreement}
+\end{equation}
+Similarly, the sequential rank agreement becomes
+\begin{equation}
+XXX \textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)\tilde{A}(X_p)^2}{(L-1)|S_d|}}, \label{def:sra-censored}
 \end{equation}
@@
 The rankings within a single list are clearly not independent since
-each of the lists essentially contains full set of ranks and by
+each rank must appear exactly in each list. Thus, by randomizing
+unobserved items as described in \eqref{formula:censored-agreement}
+
+ ess contains full set of ranks and by using
+the ... instead of applying the average ...
+
+the mean then we
 

### 2015-05-08 latex/paper.Rnw
Text update

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage[]{color}
@@
 \usepackage{xspace}
+\usepackage{calc}
 \usepackage[noae]{Sweave}
 \usepackage{authblk}
-
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 %
@@
 The agreement, $A(X_p)$, cannot be computed directly for all
-predictors because we only observe a censored version of $\tilde R$
+predictors because we only observe the censored version of $R(X_p)$
 for some of the lists. Instead we assume that the rank assigned to
 item $X_p$ in list $l$ is uniformly distributed among the ranks that
-have \emph{not} been assigned for list $l$, so the augmented agreement
-becomes
-\begin{equation}
-\tilde A(X_p) = \frac{\sum_{r_1; r_1\in \tilde R_1(X_p)}  \cdots
-  \sum_{r_L; r_L\in \tilde R_L(X_p)} A(X_p)}{\prod_l |\tilde
-  R_l(X_p)|}. \label{formula:censored-agreement}
-\end{equation}
-Similarly, the sequential rank agreement becomes
-\begin{equation}
-XXX \textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)\tilde{A}(X_p)^2}{(L-1)|S_d|}}, \label{def:sra-censored}
-\end{equation}
+have \emph{not} been assigned for list $l$. The algorithm is described
+in detail in Algorithm~\ref{sra-algorithm}.
+
+%, so the augmented agreement
+%becomes
+%\begin{equation}
+%\bar{A}(X_p) = \frac{\sum_{r_1; r_1\in \tilde R_1(X_p)}  \cdots
+%  \sum_{r_L; r_L\in \tilde R_L(X_p)} A(X_p)}{\prod_l |\tilde
+%  R_l(X_p)|}. \label{formula:censored-agreement}
+%\end{equation}
+%Similarly, the sequential rank agreement becomes
+%\begin{equation}
+%XXX \textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)\tilde{A}(X_p)^2}{(L-1)|S_d|}}, \label{def:sra-censored}
+%\end{equation}
 
@@
 unobserved items as described in \eqref{formula:censored-agreement}
+
+\begin{algorithm}
+\caption{Sequential rank agreement algorithm for censored lists}
+\label{sra-algorithm}
+\begin{algorithmic}[1]
+\Procedure{Censored rank agreement}{}
+\State Let $B$ be the number of permutations to use
+\For{each $b \in B$}
+\For{each censored list $l \in L$}
+\State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-2cm}{Permute the unassigned ranks, $k_l+1, \ldots, P$ and assign
+them randomly to the items \emph{not} found in the list, i.e.,
+$\Lambda^\complement$, in order to fill out the lists.}
+\EndFor
+\State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1cm}{Let sra($b$) be the sequential rank agreement computed from the
+filled out lists.}
+\EndFor
+\State Return element-wise averages across all $B$ permutations of sra($b$).
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
 

### 2015-05-23 latex/paper.Rnw
Style

@@
-\documentclass[12pt,a4paper]{article}
+\documentclass[12pt,a4paper]{statsoc}
 \usepackage[utf8]{inputenc}
@@
 \For{each censored list $l \in L$}
-\State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-2cm}{Permute the unassigned ranks, $k_l+1, \ldots, P$ and assign
+\State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1.7cm}{Permute the unassigned ranks, $k_l+1, \ldots, P$ and assign
 them randomly to the items \emph{not} found in the list, i.e.,
@@
 
-\bibliographystyle{apalike}
+\bibliographystyle{rss}
 \bibliography{paperref}

### 2015-05-23 R/sra.R
Fix of previously mistaken update

@@ -38,9 +38,9 @@ sra <- function(object, na.strings=NA, B=1) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength
     if (is.matrix(object))
-        object <- apply(object,2,function(x)x)
+        object <- split(object, c(col(object)))  # Convert matrix to list
     else
-        stopifnot(is.list(object)) # data.frame is a list
+        stopifnot(is.list(object)) # object is not a list
 
     nlists <- length(object)
 
@@ -58,7 +58,7 @@ sra <- function(object, na.strings=NA, B=1) {
                          # stop at duplicated items
                          if (any(duplicated(x)))
                              stop(paste0("Duplicated items found in list ",n))
-                         print(x)
+                         # print(x)  Shouldn't be in production software
                          x
                      })
     ## check class of elements, then coerce to integer

### 2015-05-23 R/sra.R
Fix for na.strings

@@ -34,7 +34,7 @@
 #' @author Claus Ekstrøm <ekstrom@@sund.ku.dk> and Thomas A Gerds <tag@@biostat.ku.dk>
 #' 
 #' @export
-sra <- function(object, na.strings=NA, B=1) {
+sra <- function(object, na.strings=c(NA, 0), B=1) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength
     if (is.matrix(object))

### 2015-05-23 R/sra.R
Fix for bug on missings

@@ -90,9 +90,9 @@ sra <- function(object, na.strings=c(NA, 0), B=1) {
     ## set B to 1, if there is no censoring
     ##             or if only one element is censored
     iscensored <- any(nmiss!=0)
-    if (B!=1 && (!iscensored || (max(nmiss)==1))) {B <- 1}
+    if (B!=1 && (!iscensored || (max(nmiss)==1))) {B <- 1 ; cat ("UHUHUH")}
 
-    # Special version of sample needed
+    ## Special version of sample needed
     resample <- function(x, ...) x[sample.int(length(x), ...)]
     tmpres <- sapply(1:B, function(b) {
                          obj.b <- lapply(1:nlists,function(j){

### 2015-05-23 latex/paper.Rnw
Bug fixes

@@
 producelists <- function(x, y) {
-    nitems <- length(y)
+    nitems <- ncol(x)
     index <- seq(1, nrow(golub))
@@
     list3 <- order(abs(coefficients), decreasing=TRUE)
-    if (nonzeros==0) {
+    if (nonzeros<nitems) {
         list3[(nonzeros+1):nitems] <- 0

### 2015-05-23 R/sra.R

@@ -11,6 +11,9 @@
 #' @return A vector of the sequential rank agreement
 ##' @examples
 ##' 
+##' mlist <- matrix(cbind(1:8,c(1,2,3,5,6,7,4,8),c(1,5,3,4,2,8,7,6)))
+##' sra(mlist)
+##' 
 ##' alist <- list(a=1:8,b=sample(1:8),c=sample(1:8))
 ##' sra(alist)
 ##' 
@@ -90,7 +93,7 @@ sra <- function(object, na.strings=c(NA, 0), B=1) {
     ## set B to 1, if there is no censoring
     ##             or if only one element is censored
     iscensored <- any(nmiss!=0)
-    if (B!=1 && (!iscensored || (max(nmiss)==1))) {B <- 1 ; cat ("UHUHUH")}
+    if (B!=1 && (!iscensored || (max(nmiss)==1))) {B <- 1}
 
     ## Special version of sample needed
     resample <- function(x, ...) x[sample.int(length(x), ...)]

### 2015-05-23 R/sra.R
Implementing generic function

@@ -1,43 +1,59 @@
-#' Compute the sequential rank agreement
-#'
-#' @param object Either matrix where each column is a ranked list of
-#' items or a list of ranked lists of items. Elements are integers
-#' between 1 and the length of the lists. The lists should have the
-#' same length but censoring can be used by setting the list to zero
-#' from a point onwards. See details for more information.
-#' @param na.strings Code for missing items 
-#' @param B An integer giving the number of randomization to sample
-#' over in the case of censored observations
-#' @return A vector of the sequential rank agreement
-##' @examples
-##' 
-##' mlist <- matrix(cbind(1:8,c(1,2,3,5,6,7,4,8),c(1,5,3,4,2,8,7,6)))
-##' sra(mlist)
-##' 
-##' alist <- list(a=1:8,b=sample(1:8),c=sample(1:8))
-##' sra(alist)
-##' 
-##' blist <- list(x1=letters,x2=sample(letters),x3=sample(letters))
-##' sra(blist)
-##' 
-##' ## censored lists are either too short
-##' clist <- list(x1=c("a","b","c","d","e","f","g","h"),
-##'               x2=c("h","c","f","g","b"),
-##'               x3=c("d","e","a"))
-##' set.seed(17)
-##' sra(clist,na.strings="z",B=10)
-##' 
-##' ## or use a special code for missing elements
-##' Clist <- list(x1=c("a","b","c","d","e","f","g","h"),
-##'               x2=c("h","c","f","g","b","z","z","z"),
-##'               x3=c("d","e","a","z","z","z","z","z"))
-##' set.seed(17)
-##' sra(Clist,na.strings="z",B=10)
-##' 
-#' @author Claus Ekstrøm <ekstrom@@sund.ku.dk> and Thomas A Gerds <tag@@biostat.ku.dk>
-#' 
-#' @export
-sra <- function(object, na.strings=c(NA, 0), B=1) {
+sra.default <- function(object, B=1, na.strings=c(0, NA)) {
+    # Make sure that the input object ends up as a matrix with integer columns all
+    # consisting of elements from 1 and up to listlength
+    if (is.matrix(object))
+        rankmat <- object
+    else
+        rankmat <- as.matrix(do.call("cbind",object))
+    
+    listlength <- nrow(rankmat)
+    nlists <- NCOL(rankmat)
+    nseq <- seq(listlength)
+    
+    # Special version of sample needed
+    resample <- function(x, ...) x[sample.int(length(x), ...)]
+    
[…]

### 2015-05-23 R/sra.R
Generic

@@ -1,4 +1,48 @@
-sra.default <- function(object, B=1, na.strings=c(0, NA)) {
+#' Compute the sequential rank agreement
+#'
+#' @param object Either matrix where each column is a ranked list of
+#' items or a list of ranked lists of items. Elements are integers
+#' between 1 and the length of the lists. The lists should have the
+#' same length but censoring can be used by setting the list to zero
+#' from a point onwards. See details for more information.
+#' @param na.strings Code for missing items 
+#' @param B An integer giving the number of randomization to sample
+#' over in the case of censored observations
+#' @return A vector of the sequential rank agreement
+##' @examples
+##' 
+##' mlist <- matrix(cbind(1:8,c(1,2,3,5,6,7,4,8),c(1,5,3,4,2,8,7,6)),ncol=3)
+##' sra(mlist)
+##' 
+##' alist <- list(a=1:8,b=sample(1:8),c=sample(1:8))
+##' sra(alist)
+##' 
+##' blist <- list(x1=letters,x2=sample(letters),x3=sample(letters))
+##' sra(blist)
+##' 
+##' ## censored lists are either too short
+##' clist <- list(x1=c("a","b","c","d","e","f","g","h"),
+##'               x2=c("h","c","f","g","b"),
+##'               x3=c("d","e","a"))
+##' set.seed(17)
+##' sra(clist,na.strings="z",B=10)
+##' 
+##' ## or use a special code for missing elements
+##' Clist <- list(x1=c("a","b","c","d","e","f","g","h"),
+##'               x2=c("h","c","f","g","b","z","z","z"),
+##'               x3=c("d","e","a","z","z","z","z","z"))
+##' set.seed(17)
+##' sra(Clist,na.strings="z",B=10)
+##' 
+#' @author Claus Ekstrøm <ekstrom@@sund.ku.dk> and Thomas A Gerds <tag@@biostat.ku.dk>
+#' 
+#' @export
+sra <- function(object, B=1, ...) {
+  UseMethod("sra")
+}
+
+
+sra.default <- function(object, B=1) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength
     if (is.matrix(object))
@@ -52,8 +96,7 @@ sra.default <- function(object, B=1, na.strings=c(0, NA)) {
 }
 
 
-
-sra.tagteam <- function(object, B=1, na.strings=c(NA, 0)) {
[…]

### 2015-05-23 R/sra.R
Generic

@@ -37,11 +37,13 @@
 #' @author Claus Ekstrøm <ekstrom@@sund.ku.dk> and Thomas A Gerds <tag@@biostat.ku.dk>
 #' 
 #' @export
-sra <- function(object, B=1, ...) {
+sra <- function(object, B=1) {
   UseMethod("sra")
 }
 
-
+#' @rdname sra
+#' @method sra default
+#' @S3method sra default
 sra.default <- function(object, B=1) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength
@@ -96,6 +98,9 @@ sra.default <- function(object, B=1) {
 }
 
 
+#' @rdname sra
+#' @method sra list
+#' @S3method sra list
 sra.list <- function(object, B=1, na.strings=c(NA, 0)) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength

### 2015-05-23 R/sra.R
Generic implementation

@@ -36,21 +36,21 @@
 ##' 
 #' @author Claus Ekstrøm <ekstrom@@sund.ku.dk> and Thomas A Gerds <tag@@biostat.ku.dk>
 #' 
+#' @rdname sra
 #' @export
 sra <- function(object, B=1) {
   UseMethod("sra")
 }
 
 #' @rdname sra
-#' @method sra default
-#' @S3method sra default
+#' @method sra
 sra.default <- function(object, B=1) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength
-    if (is.matrix(object))
-        rankmat <- object
-    else
-        rankmat <- as.matrix(do.call("cbind",object))
+    if (!is.matrix(object))
+       stop("Input must be a matrix")
+
+    rankmat <- object
     
     listlength <- nrow(rankmat)
     nlists <- NCOL(rankmat)
@@ -61,12 +61,17 @@ sra.default <- function(object, B=1) {
     
     # Compute a list of missing items for each list
     missing.items <- lapply(as.data.frame(rankmat), function(x) { nseq[-x] })
+    nmissing.items <- sapply(missing.items, length)
+   
+    ## Checking that all values are non-negative
+    if (min(rankmat)<0) 
+        stop("Negative items not allowed")
 
     ## Should make a sanity check that zeros are from a point onwards
     if (!all(sapply(1:nlists, function(x) {
                         res <- TRUE
-                        if (length(missing.items[[x]])>0) {
-                            if (any(rankmat[(listlength-missing.items[[x]]):listlength,x]) )
+                        if (nmissing.items[x]>0) {
+                            if (sum(rankmat[(listlength-nmissing.items[[x]]):listlength,x]) )
                                 { res <- FALSE }
                         }
                         res
@@ -99,8 +104,7 @@ sra.default <- function(object, B=1) {
 
 
 #' @rdname sra
-#' @method sra list
-#' @S3method sra list
+#' @method sra
 sra.list <- function(object, B=1, na.strings=c(NA, 0)) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength

### 2015-05-23 R/sra.R
Generic

@@ -43,7 +43,7 @@ sra <- function(object, B=1) {
 }
 
 #' @rdname sra
-#' @method sra
+#' @export sra
 sra.default <- function(object, B=1) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength
@@ -104,7 +104,7 @@ sra.default <- function(object, B=1) {
 
 
 #' @rdname sra
-#' @method sra
+#' @export sra
 sra.list <- function(object, B=1, na.strings=c(NA, 0)) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength

### 2015-05-23 R/sra.R
Generic

@@ -43,7 +43,7 @@ sra <- function(object, B=1) {
 }
 
 #' @rdname sra
-#' @export sra
+#' @export
 sra.default <- function(object, B=1) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength
@@ -104,7 +104,7 @@ sra.default <- function(object, B=1) {
 
 
 #' @rdname sra
-#' @export sra
+#' @export
 sra.list <- function(object, B=1, na.strings=c(NA, 0)) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength

### 2015-05-23 R/sra.R
Generic

@@ -71,7 +71,7 @@ sra.default <- function(object, B=1) {
     if (!all(sapply(1:nlists, function(x) {
                         res <- TRUE
                         if (nmissing.items[x]>0) {
-                            if (sum(rankmat[(listlength-nmissing.items[[x]]):listlength,x]) )
+                            if (sum(rankmat[(listlength-nmissing.items[[x]]+1):listlength,x])>0)
                                 { res <- FALSE }
                         }
                         res

### 2015-05-23 R/sra.R
Update to version 0.2

@@ -221,7 +221,7 @@ random_list_sra <- function(object, B=1, n=1) {
         for (j in 1:ncol(object)) {
             object[,j] <- c(sample(nitems, size=notmiss[j]), rep(0, nitems-notmiss[j]))
         }
-        sra(object, na.strings="0", B=B) 
+        sra(object, B=B) 
     })
     res
     
@@ -240,7 +240,5 @@ smooth_sra <- function(object, confidence=0.95) {
 
     alpha <- (1-confidence)/2
     limits <- apply(object, 1, function(x) { quantile(x, probs=c(alpha, 1-alpha)) })
-#    res <- apply(limits, 1, function(x) { loess(x)} )
-#    list(lower=res[[1]]$y, upper=res[[2]]$y)
     list(lower=limits[1,], upper=limits[2,])
 }

### 2015-05-23 latex/paper.Rnw
Implementing generic function

@@
 
+%%% RSS specific items
+
+\title[Sequential rank agreement]{Sequential rank agreement methods for comparison of ranked
+  lists}
+\author[Ekstrøm {\it et al.}]{Claus Thorn Ekstrøm}
+\address{Affiliation,
+City,
+Country.}
+\email{Author@emailaddress.com}
+\author{Thomas Alexander Gerds}
+\address{Affiliation,
+City,
+Country.}
+\author{Andreas Kryger Jensen}
+\address{Affiliation,
+City,
+Country.}
+\author{Kasper Brink-Jensen}
+\address{Affiliation,
+City,
+Country.}
+
 \begin{document}
 
-\title{Sequential rank agreement methods for comparison of ranked
+% \title{Sequential rank agreement methods for comparison of ranked
   lists}
-\author{Claus Thorn Ekstrøm}
-\author{Thomas Alexander Gerds}
-\author{Kasper Brink-Jensen}
-\affil{Biostatistics, University of Copenhagen}
+% \author{Claus Thorn Ekstrøm}
+% \author{Thomas Alexander Gerds}
+% \author{Kasper Brink-Jensen}
+% \affil{Biostatistics, University of Copenhagen}
 
@@
 producelists <- function(x, y) {
-    nitems <- ncol(x)
+    nitems <- nrow(x)
     index <- seq(1, nrow(golub))

### 2015-05-24 R/sra.R
fixed bugs in treatment of matrix and censored lists

@@ -38,29 +38,28 @@ sra <- function(object, na.strings=NA, B=1) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength
     if (is.matrix(object))
-        object <- apply(object,2,function(x)x)
+        object <- lapply(1:NCOL(object),function(j)object[,j]) # Convert matrix to list
     else
         stopifnot(is.list(object)) # data.frame is a list
-
     nlists <- length(object)
 
     ## sanity checks
-    object <- lapply(object,function(x){
-                         out <- which(is.na(x))
-                         na.strings <- na.strings[!is.na(na.strings)]
-                         if (length(na.strings)>0)
-                             out <- c(out,grep(paste0("^",na.strings,"$"),x))
-                         ## remove censored items with side effect:
-                         ## in case where all lists have trailing censored information
-                         ## this is pruned
-                         if (length(out)>0)
-                             x <- x[-out]
-                         # stop at duplicated items
-                         if (any(duplicated(x)))
-                             stop(paste0("Duplicated items found in list ",n))
-                         print(x)
-                         x
-                     })
+    object <- lapply(1:length(object),function(j){
+                                 x <- object[[j]]
+                                 out <- which(is.na(x))
+                                 na.strings <- na.strings[!is.na(na.strings)]
+                                 if (length(na.strings)>0)
+                                     out <- c(out,grep(paste0("^",na.strings,"$"),x))
+                                 ## remove censored items with side effect:
+                                 ## in case where all lists have trailing censored information
+                                 ## this is pruned
+                                 if (length(out)>0)
+                                     x <- x[-out]
+                                 # stop at duplicated items
+                                 if (any(duplicated(x)))
+                                     stop(paste0("Duplicated items found in list ",j))
+                                 x
+                             })
     ## check class of elements, then coerce to integer
     cc <- sapply(object,class)
     if (length(cc <- unique(cc))>1) stop(paste("All elements of object must have the same class. Found:",paste(cc,collapse=", ")))
@@ -83,9 +82,9 @@ sra <- function(object, na.strings=NA, B=1) {
     ## fill too short lists with 0 (code for missing)
[…]

### 2015-05-24 R/sra.R
sra.matrix added and improved

@@ -80,7 +80,7 @@ sra.default <- function(object, B=1) {
     }
 
     ## If there is no censoring then we should set B to 1
-    if (max(sapply(missing.items, function(i) length(i) ))==0) {
+    if (max(nmissing.items)==0) {
         B <- 1
     }
     
@@ -101,6 +101,29 @@ sra.default <- function(object, B=1) {
     
     agreement
 }
+
+
+#' @rdname sra
+#' @export
+sra.matrix <- function(object, B=1, na.strings=c(NA, 0)) {
+    if (!is.matrix(object))
+        stop("Input object must be a matrix")
+
+    ## Convert all missing types to zeros
+    object[object %in% na.strings] <- 0
+
+    ## Check that we dont have more unique values than rows
+    if (length(unique(object[object != 0]))>nrow(object)) {
+        warning("Found more unique items in the matrix than rows. Expanding the number of rows to match")
+        unique.items <- length(unique(object[object != 0]))
+        # Expand the columns in the matrix to have length unique.items
+        glue <- matrix(rep(0, ncol(object)*(unique.items - nrow(object))), ncol=ncol(object))
+        object <- rbind(object, glue)
+    }
+
+    sra.default(object, B=B)
+}
+
 
 
 #' @rdname sra
@@ -159,6 +182,11 @@ sra.list <- function(object, B=1, na.strings=c(NA, 0)) {
     if (tooshort)
         object <- lapply(object,function(x){c(x,rep(0,nitems-length(x)))})
 
+
+
+    print(matrix(unlist(z), ncol = 10, byrow = TRUE))
+
+
     ## set B to 1, if there is no censoring
     ##             or if only one element is censored
     iscensored <- any(nmiss!=0)
@@ -191,7 +219,10 @@ sra.list <- function(object, B=1, na.strings=c(NA, 0)) {
 }
 
 
+oldstuff <- function() {
 
+
+}
 
 
 

### 2015-05-24 R/sra.R
leaner version of sra.list

@@ -171,7 +171,7 @@ sra.list <- function(object, B=1, na.strings=c(NA, 0)) {
 
     ## Compute a list of missing items for each list
     missing.items <- lapply(object, function(x) {items[match(items,x,nomatch=0)==0]})
-    missing.items1 <- lapply(object, function(x) {items[-x]})
+    ### missing.items1 <- lapply(object, function(x) {items[-x]})
     nmiss <- sapply(missing.items,length)
     
 
@@ -182,51 +182,17 @@ sra.list <- function(object, B=1, na.strings=c(NA, 0)) {
     if (tooshort)
         object <- lapply(object,function(x){c(x,rep(0,nitems-length(x)))})
 
-
-
-    print(matrix(unlist(z), ncol = 10, byrow = TRUE))
-
-
     ## set B to 1, if there is no censoring
     ##             or if only one element is censored
     iscensored <- any(nmiss!=0)
     if (B!=1 && (!iscensored || (max(nmiss)==1))) {B <- 1}
 
-    ## Special version of sample needed
-    resample <- function(x, ...) x[sample.int(length(x), ...)]
-    tmpres <- sapply(1:B, function(b) {
-                         obj.b <- lapply(1:nlists,function(j){
-                                             list <- object[[j]]
-                                             if (nmiss[[j]]>0){
-                                                 if (nmiss[[j]]==1){ ## fill in missing item
-                                                     list[list==0] <- missing.items[[j]]
-                                                 } else{ ## replace censored items with random order
-                                                       list[list==0] <- resample(missing.items[[j]])
-                                                   }
-                                             }
-                                             list
-                                         })
-                         ## bind lists
-                         rankmat <- do.call("cbind",obj.b)
-                         res <- sracppfull(rankmat)
-                         res
-                     })
-    agreement <- rowMeans(tmpres)
-    ### names(agreement) <- labels
-    class(agreement) <- "sra"
-    attr(agreement, "B") <- B
-    agreement
+    ## Fast conversion from list to matrix
+    object <- matrix(unlist(object), ncol = nlists)
+    
+    ## Call the computation
+    sra.default(object, B=B)
 }
-
-
-oldstuff <- function() {
-
-
-}
-
-
-
-
 
 
 

### 2015-05-24 R/sra.R
leaner version of sra.list

@@ -38,7 +38,7 @@
 #' 
 #' @rdname sra
 #' @export
-sra <- function(object, B=1) {
+sra <- function(object, B=1, ...) {
   UseMethod("sra")
 }
 

### 2015-05-26 R/sra.R
nitems

@@ -105,19 +105,23 @@ sra.default <- function(object, B=1) {
 
 #' @rdname sra
 #' @export
-sra.matrix <- function(object, B=1, na.strings=c(NA, 0)) {
+sra.matrix <- function(object, B=1, na.strings=c(NA, 0), nitems=rows(object)) {
     if (!is.matrix(object))
         stop("Input object must be a matrix")
 
     ## Convert all missing types to zeros
     object[object %in% na.strings] <- 0
 
+    unique.items <- length(unique(object[object != 0]))
+
     ## Check that we dont have more unique values than rows
-    if (length(unique(object[object != 0]))>nrow(object)) {
-        warning("Found more unique items in the matrix than rows. Expanding the number of rows to match")
-        unique.items <- length(unique(object[object != 0]))
-        # Expand the columns in the matrix to have length unique.items
-        glue <- matrix(rep(0, ncol(object)*(unique.items - nrow(object))), ncol=ncol(object))
+    if (unique.items > nitems) {
+        stop("Found more unique items in the matrix than rows/nitems. Increase nitems to match")
+    }
+
+    ## Expand the columns in the matrix to have length unique.items
+    if (nitems>nrow(object)) {
+        glue <- matrix(rep(0, ncol(object)*(nitems - nrow(object))), ncol=ncol(object))
         object <- rbind(object, glue)
     }
 
@@ -128,13 +132,12 @@ sra.matrix <- function(object, B=1, na.strings=c(NA, 0)) {
 
 #' @rdname sra
 #' @export
-sra.list <- function(object, B=1, na.strings=c(NA, 0)) {
+sra.list <- function(object, B=1, na.strings=c(NA, 0), nitems=max(sapply(object, length))) {
+
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength
-    if (is.matrix(object))
-        object <- lapply(1:NCOL(object),function(j)object[,j]) # Convert matrix to list
-    else
-        stopifnot(is.list(object)) # data.frame is a list
+
+    stopifnot(is.list(object)) # data.frame is a list
     nlists <- length(object)
 
     ## sanity checks
@@ -163,7 +166,7 @@ sra.list <- function(object, B=1, na.strings=c(NA, 0)) {
     labels <- unique(unlist(object,recursive=TRUE,use.names=FALSE))
     nitems <- length(labels)
[…]

### 2015-05-29 latex/paper.Rnw
Small fix

--- 2015-05-23 latex/paper.Rnw
@@
-\documentclass[12pt,a4paper]{statsoc}
+\documentclass[12pt,a4paper]{article}
 \usepackage[utf8]{inputenc}
@@
 
-\begin{document}
+%%% RSS specific items
 
@@
 \affil{Biostatistics, University of Copenhagen}
+
+\begin{document}
 
@@
 
-\section{Benchmarks}
+\section{Evaluating sequential rank agreement}
 
@@
 
-\section{Alternatives}
+\section{Comparison to ...}
+
+The overlap among $L$ lists observed until depth $d$ is defined as the
+proportion of items that are found in all $L$ lists compared to the
+possible number of items found in all lists observed until depth $d$:
+\begin{equation}
+O(d) = \frac{| \cap_{l=1}^L \{\pi_l(r); r\leq d \} |}{d}.
+\end{equation}
+Typically, the overlap is only considered for pairwise comparisons
+
+The average overlap at depth $d$ is defined as the average of the
+overlaps until $d$,
+\begin{equation}
+AO(d) = \sum_{i=1}^d \frac{O(d)}{d}.
+\end{equation}
+This definition indirectly ensures that extra emphasis is included on
+the top of the list since the items
 

### 2015-05-29 R/sra.R

@@ -5,37 +5,37 @@
 #' between 1 and the length of the lists. The lists should have the
 #' same length but censoring can be used by setting the list to zero
 #' from a point onwards. See details for more information.
-#' @param na.strings Code for missing items 
+#' @param na.strings Code for missing items
 #' @param B An integer giving the number of randomization to sample
 #' over in the case of censored observations
 #' @return A vector of the sequential rank agreement
 ##' @examples
-##' 
+##'
 ##' mlist <- matrix(cbind(1:8,c(1,2,3,5,6,7,4,8),c(1,5,3,4,2,8,7,6)),ncol=3)
 ##' sra(mlist)
-##' 
+##'
 ##' alist <- list(a=1:8,b=sample(1:8),c=sample(1:8))
 ##' sra(alist)
-##' 
+##'
 ##' blist <- list(x1=letters,x2=sample(letters),x3=sample(letters))
 ##' sra(blist)
-##' 
+##'
 ##' ## censored lists are either too short
 ##' clist <- list(x1=c("a","b","c","d","e","f","g","h"),
 ##'               x2=c("h","c","f","g","b"),
 ##'               x3=c("d","e","a"))
 ##' set.seed(17)
 ##' sra(clist,na.strings="z",B=10)
-##' 
+##'
 ##' ## or use a special code for missing elements
 ##' Clist <- list(x1=c("a","b","c","d","e","f","g","h"),
 ##'               x2=c("h","c","f","g","b","z","z","z"),
 ##'               x3=c("d","e","a","z","z","z","z","z"))
 ##' set.seed(17)
 ##' sra(Clist,na.strings="z",B=10)
-##' 
+##'
 #' @author Claus Ekstrøm <ekstrom@@sund.ku.dk> and Thomas A Gerds <tag@@biostat.ku.dk>
-#' 
+#'
 #' @rdname sra
 #' @export
 sra <- function(object, B=1, ...) {
@@ -51,20 +51,20 @@ sra.default <- function(object, B=1) {
        stop("Input must be a matrix")
 
     rankmat <- object
-    
+
     listlength <- nrow(rankmat)
     nlists <- NCOL(rankmat)
     nseq <- seq(listlength)
-    
+
     # Special version of sample needed
     resample <- function(x, ...) x[sample.int(length(x), ...)]
-    
+
     # Compute a list of missing items for each list
     missing.items <- lapply(as.data.frame(rankmat), function(x) { nseq[-x] })
[…]

### 2015-05-30 R/sra.R
clean up confusion

@@ -45,69 +45,12 @@ sra <- function(object, B=1, ...) {
 #' @rdname sra
 #' @export
 sra.default <- function(object, B=1) {
-    # Make sure that the input object ends up as a matrix with integer columns all
-    # consisting of elements from 1 and up to listlength
-    if (!is.matrix(object))
-       stop("Input must be a matrix")
-
-    rankmat <- object
-    
-    listlength <- nrow(rankmat)
-    nlists <- NCOL(rankmat)
-    nseq <- seq(listlength)
-    
-    # Special version of sample needed
-    resample <- function(x, ...) x[sample.int(length(x), ...)]
-    
-    # Compute a list of missing items for each list
-    missing.items <- lapply(as.data.frame(rankmat), function(x) { nseq[-x] })
-    nmissing.items <- sapply(missing.items, length)
-   
-    ## Checking that all values are non-negative
-    if (min(rankmat)<0) 
-        stop("Negative items not allowed")
-
-    ## Should make a sanity check that zeros are from a point onwards
-    if (!all(sapply(1:nlists, function(x) {
-                        res <- TRUE
-                        if (nmissing.items[x]>0) {
-                            if (sum(rankmat[(listlength-nmissing.items[[x]]+1):listlength,x])>0)
-                                { res <- FALSE }
-                        }
-                        res
-                    }))) {
-        stop("Censored ranked lists should be coded 0 from a rank onwards")
-    }
-
-    ## If there is no censoring then we should set B to 1
-    if (max(nmissing.items)==0) {
-        B <- 1
-    }
-    
-    tmpres <- sapply(1:B, function(i) {
-        for (j in 1:nlists) {
-            if (length(missing.items[[j]])>0) {      
-                rankmat[(listlength-length(missing.items[[j]])+1):listlength,j] <- resample(missing.items[[j]])
-            }
-        }
-        res <- sracppfull(rankmat)
-        res
-    })
-
-    agreement <- apply(tmpres, 1, mean)
-
-    class(agreement) <- "sra"
-    attr(agreement, "B") <- B
-    
-    agreement
+    stop("Input must be either a matrix, a data.frame or a list.")
 }
-
 
 #' @rdname sra
 #' @export
 sra.matrix <- function(object, B=1, na.strings=c(NA, 0)) {
-    if (!is.matrix(object))
-        stop("Input object must be a matrix")
 
     ## Convert all missing types to zeros
[…]

### 2015-06-01 latex/paper.Rnw
Important implementation items. NOVEL

@@
 inputmatrix2 <- inputmatrix
-inputmatrix2[51:nrow(inputmatrix),] <- 0
+inputmatrix2[51:nrow(inputmatrix),] <- NA
 res2 <- sqrt(sra(inputmatrix2, B=20))
@@
 
-<<fig2a, fig=TRUE,echo=FALSE,include=FALSE>>=
+<<fig2a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE>>=
 ysize <- 1000
@@
 
-<<fig2b, fig=TRUE,echo=FALSE,include=FALSE >>=
+<<fig2b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE >>=
 ysize <- 80
@@
 
-<<echo=FALSE, results=hide>>=
-xxx <- as.data.frame(inputmatrix[1:10,])
-xxx
-
-@
-
 <<label=tab1,echo=FALSE,results=tex>>=
 library(xtable)
-print(xtable(xxx))
+xxx <- as.data.frame(inputmatrix[1:10,])
+xxx2 <- xtable(xxx)
+digits(xxx2) <- 0
+print(xxx2)
 #print(xtable(xxx, caption = "Top 10 list of results", label = "tab:one",
@@
 
+In this section we apply the method to the ColonData data from the R
+package MAMA (Ihnatova, 2013), which consists of gene expression
+measurements from a subset of 500 genes from 77 individuals from
+Denmark, 36 from Australia and 41 from Japan as well as information on
+microsattelite stability (stable MSS/unstable MSI) which is often used
+as an indicator for colorectal cancer. The genes were tested for
+difference in expression in the MSI/MSS groups using the limma package
+(Smyth, 2005). The genes were then ranked according to p–values giving
+three lists that we wish to compare, the top of these can be seen in
+Table 4.
+
+<<fig=TRUE>>=
+myres <- average_overlap(inputmatrix[,1:2])
+myres2 <- average_overlap(inputmatrix[-1,1:2])
+xvals <- 1:500
+plot(myres, xlim=c(1, 500), type="l")
+lines(xvals[-1], myres2, col="red")
+@
+
 \section{Discussion}

### 2015-06-01 R/sra.R
na.strings

@@ -55,7 +55,7 @@ sra.matrix <- function(object, B=1, na.strings=NULL, nitems=nrow(object)) {
         stop("Input object must be a matrix")
 
     ## Convert all missing types to zeros
-    if (!is.null(na.stsrings)) {
+    if (!is.null(na.strings)) {
         object[object %in% na.strings] <- NA
     }
 

### 2015-06-01 R/sra.R
na.strings

@@ -92,9 +92,10 @@ sra.list <- function(object, B=1, na.strings=NULL, nitems=max(sapply(object, len
                          x <- object[[j]]
                          ## Add the NA items to be removed
                          out <- which(is.na(x))
-                         na.strings <- na.strings[!is.na(na.strings)]
-                         if (!is.null(na.strings))
+                         if (!is.null(na.strings)) {
+                             na.strings <- na.strings[!is.na(na.strings)]
                              out <- c(out,grep(paste0("^",na.strings,"$"),x))
+                         }
                          ## remove censored items with side effect:
                          ## in case where all lists have trailing censored information
                          ## this is pruned

### 2015-06-01 latex/paper.Rnw
na.strings

@@
     if (nonzeros<nitems) {
-        list3[(nonzeros+1):nitems] <- 0
+        list3[(nonzeros+1):nitems] <- NA
     }

### 2015-06-01 R/sra.R

@@ -54,7 +54,7 @@ sra.matrix <- function(object, B=1, na.strings=NULL, nitems=nrow(object)) {
     if (!is.matrix(object))
         stop("Input object must be a matrix")
 
-    ## Convert all missing types to zeros
+    ## Convert all missing types to NAs
     if (!is.null(na.strings)) {
         object[object %in% na.strings] <- NA
     }
@@ -72,7 +72,8 @@ sra.matrix <- function(object, B=1, na.strings=NULL, nitems=nrow(object)) {
         object <- rbind(object, glue)
     }
     object <- lapply(1:NCOL(object),function(j)object[,j]) # Convert matrix to list
-    sra.list(object, B=B)
+    object
+##    sra.list(object, B=B)
 }
 
 

### 2015-06-01 R/sra.R
na.strings

@@ -72,8 +72,7 @@ sra.matrix <- function(object, B=1, na.strings=NULL, nitems=nrow(object)) {
         object <- rbind(object, glue)
     }
     object <- lapply(1:NCOL(object),function(j)object[,j]) # Convert matrix to list
-    object
-##    sra.list(object, B=B)
+    sra.list(object, B=B, nitems=nitems)
 }
 
 
@@ -81,14 +80,15 @@ sra.matrix <- function(object, B=1, na.strings=NULL, nitems=nrow(object)) {
 #' @rdname sra
 #' @export
 sra.list <- function(object, B=1, na.strings=NULL, nitems=max(sapply(object, length))) {
-
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength
 
-    stopifnot(is.list(object)) # data.frame is a list
+    stopifnot(is.list(object))
 
     nlists <- length(object)
-    ## sanity checks
+    nitems <- nitems ## force evaluation here
+
+    ## Sanity checks
     object <- lapply(1:length(object),function(j){
                          x <- object[[j]]
                          ## Add the NA items to be removed

### 2015-06-01 R/sra.R
Fixed labels

@@ -158,7 +158,7 @@ sra.list <- function(object, B=1, na.strings=NULL, nitems=max(sapply(object, len
                          res
                      })
     agreement <- rowMeans(tmpres)
-    names(agreement) <- labels
+    names(agreement) <- items
     class(agreement) <- "sra"
     attr(agreement, "B") <- B
     agreement

### 2015-06-01 R/sra.R
Fixed na.strings in random lists

@@ -5,7 +5,7 @@
 #' between 1 and the length of the lists. The lists should have the
 #' same length but censoring can be used by setting the list to zero
 #' from a point onwards. See details for more information.
-#' @param na.strings Code for missing items
+#' @param na.strings A vector of strings/values that represent missing values in addition to NA. Defaults to NULL which means only NA are censored values.
 #' @param B An integer giving the number of randomization to sample
 #' over in the case of censored observations
 #' @return A vector of the sequential rank agreement
@@ -168,25 +168,31 @@ sra.list <- function(object, B=1, na.strings=NULL, nitems=max(sapply(object, len
 
 #' Simulate sequential rank agreement for randomized unrelated lists
 #'
-#' @param obj Either a vector or matrix
+#' @param obj A matrix
 #' @param B Either a vector or matrix
 #' @param n the number of sequential rank agreement curves to produce
+#' @param na.strings A vector of character values that represent vensored observations
 #' @return A matrix with n columns each representing the sequential rank agreement obtained from
 #' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
 #' @export
-random_list_sra <- function(object, B=1, n=1) {
+random_list_sra <- function(object, B=1, n=1, na.strings=NULL) {
 
     ## Make sure that the input object ends up as a matrix with integer columns all
     ## consisting of elements from 1 and up to listlength
     if (!is.matrix(object))
         object <- as.matrix(do.call("cbind",object))
 
+    ## Convert all missing types to NAs
+    if (!is.null(na.strings)) {
+        object[object %in% na.strings] <- NA
+    }
+
     nitems <- nrow(object)
-    notmiss <- apply(object, 2, function(x) {sum(x>0)} )
+    notmiss <- apply(object, 2, function(x) {sum(!is.na(x))} )
     res <- sapply(1:n, function(i) {
         ## Do a permutation with the same number of missing
         for (j in 1:ncol(object)) {
-            object[,j] <- c(sample(nitems, size=notmiss[j]), rep(0, nitems-notmiss[j]))
+            object[,j] <- c(sample(nitems, size=notmiss[j]), rep(NA, nitems-notmiss[j]))
         }
         sra(object, B=B)
     })

### 2015-06-01 latex/paper.Rnw
Added examples

@@
 # col1 <- makeTransparent()
-null <- sqrt(random_list_sra(inputmatrix, B=1, n=400))
-null3 <- sqrt(random_list_sra(inputmatrix, B=20, n=200))
+null <- sqrt(random_list_sra(inputmatrix, B=1, n=40))
+null3 <- sqrt(random_list_sra(inputmatrix, B=20, n=20))
 ## matlines(null[1:ysize,], col="gray", lty=1)
@@
 
-<<fig=TRUE>>=
+<<fig-ao1, echo=FALSE, fig=TRUE,include=FALSE>>=
 myres <- average_overlap(inputmatrix[,1:2])
-myres2 <- average_overlap(inputmatrix[-1,1:2])
+myres2 <- average_overlap(inputmatrix[-c(1,2),1:2])
 xvals <- 1:500
-plot(myres, xlim=c(1, 500), type="l")
-lines(xvals[-1], myres2, col="red")
+plot(myres, xlim=c(1, 300), ylim=c(0,1), type="l")
+lines(xvals[-c(1,2)], myres2[1:498], col="red")
 @
+
+<<fig-ao2, echo=FALSE, fig=TRUE,include=FALSE>>=
+mmyres <- sqrt(sra(inputmatrix[,1:2]))
+mmyres2 <- sqrt(sra(inputmatrix[-c(1,2),1:2]))
+xvals <- 1:500
+plot(mmyres, xlim=c(1, 300), ylim=c(0,1), type="l")
+lines(xvals[-c(1,2)], mmyres2[1:498], col="red")
+@
+
+\begin{figure}[tb]
+\begin{center}
+\includegraphics[width=.45\textwidth]{paper-fig-ao1}
+\includegraphics[width=.45\textwidth]{paper-fig-ao2}
+\end{center}
+ \caption{Sequential rank agreement for 4 different analysis methods
+   applied to the Golub data (thick red line). The red and blue shaded
+ areas represent the pointwise 95\% normal areas for the XXX and XXY benchmarks,
+respectively.}
+ \label{fig:case1}
+\end{figure}
 

### 2015-06-01 R/sra.R

@@ -14,6 +14,9 @@
 ##' mlist <- matrix(cbind(1:8,c(1,2,3,5,6,7,4,8),c(1,5,3,4,2,8,7,6)),ncol=3)
 ##' sra(mlist)
 ##'
+##' mlist <- matrix(cbind(1:8,c(1,2,3,5,6,7,4,8),c(1,5,3,4,2,8,7,6)),ncol=3)
+##' sra(mlist, nitems=20, B=10)
+##'
 ##' alist <- list(a=1:8,b=sample(1:8),c=sample(1:8))
 ##' sra(alist)
 ##'

### 2015-06-03 latex/paper.Rnw
sqrt result and plot label

@@
 must be larger than the largest rank observed in that list.
-Unless we have additional prior knowledge about the distribution of the
-unranked items then we can restrict attention of the sequential rank
-agreement to
-\begin{equation}
-d \leq \max(k_1, \ldots, k_L),
-\end{equation}
-since we have no information about the ordering of the items after a
-depth $d$ where all lists are censored.
 
 The agreement, $A(X_p)$, cannot be computed directly for all
-predictors because we only observe the censored version of $R(X_p)$
-for some of the lists. Instead we assume that the rank assigned to
-item $X_p$ in list $l$ is uniformly distributed among the ranks that
-have \emph{not} been assigned for list $l$. The algorithm is described
-in detail in Algorithm~\ref{sra-algorithm}.
+predictors when we have censored lists because we only do not know the
+exact rank assigned to each item in all lists.  Instead we assume that
+the rank assigned to an item $X_p$ that is not present in list $l$
+(i.e., is censored) is uniformly distributed among the ranks that have
+\emph{not} been assigned for list $l$.  The rankings within a single
+list are clearly not independent since each rank must appear exactly
+in each list. Thus we cannot simply assign
 
-%, so the augmented agreement
-%becomes
-%\begin{equation}
-%\bar{A}(X_p) = \frac{\sum_{r_1; r_1\in \tilde R_1(X_p)}  \cdots
-%  \sum_{r_L; r_L\in \tilde R_L(X_p)} A(X_p)}{\prod_l |\tilde
-%  R_l(X_p)|}. \label{formula:censored-agreement}
-%\end{equation}
-%Similarly, the sequential rank agreement becomes
-%\begin{equation}
-%XXX \textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)\tilde{A}(X_p)^2}{(L-1)|S_d|}}, \label{def:sra-censored}
-%\end{equation}
+Thus, by randomizing unobserved items as described in
+\eqref{formula:censored-agreement}
 
-The rankings within a single list are clearly not independent since
-each rank must appear exactly in each list. Thus, by randomizing
-unobserved items as described in \eqref{formula:censored-agreement}
+The algorithm is described in
+detail in Algorithm~\ref{sra-algorithm}.
 
@@
 \end{algorithm}
+
+Unless we have additional prior knowledge about the distribution of the
+unranked items then we can restrict attention of the sequential rank
+agreement to
+\begin{equation}
+d \leq \max(k_1, \ldots, k_L),
+\end{equation}
+since we have no information about the ordering of the items after a
+depth $d$ where all lists are censored.
 
@@
 \end{equation}
-This definition indirectly ensures that extra emphasis is included on
-the top of the list since the items
+This definition directly ensures that extra emphasis is included on
+the top of the list since the overlap of the items towards the top are
+included in the majority of the calculations.
+
+In a classical paper by \citet{Golub1999} a dataset of 3051 gene
+expression values were measured on 38 tumor mRNA samples in order to
+improve the classification of acute leukemias between two types: acute
+lymphoblastic leukemia (ALL) or acute myeloid leukemia
+(AML). Preprocessing of the gene expression data was done as described
+in \citep{Dudoit2002}.
+
+We analyzed these gene expression data using four different
+approaches: marginal two-sample $t$ tests, marginal logistic
+regression analyses, logistic regression eleastic net, and marginal
+maximum information content correlations (MIC) \citep{Reshef2011}. For
+the first two methods, the genes were ranked according to $p$ value.
 
@@
 xvals <- 1:500
-plot(mmyres, xlim=c(1, 300), ylim=c(0,1), type="l")
+plot(mmyres, xlim=c(1, 300), type="l")
 lines(xvals[-c(1,2)], mmyres2[1:498], col="red")

### 2015-06-03 R/plot.sra.R

@@ -1,5 +1,5 @@
 ##' Plot the agreement between lists as a function of the list depth
-##' 
+##'
 ##' @title Plot sequential rank agreement
 ##' @param x Agreement object
 ##' @param xlim x-axis limits
@@ -17,17 +17,17 @@
 ##'     sra(list(sample(R1),sample(R1)))
 ##' })))
 ##' lines(1:length(R1),arand,col=2,lwd=3)
-##' 
+##'
 ##' l <- c(1,2,3,4,5,7,6,8,9,10,11,12,13)
 ##' l <- 1:100
 ##' aa <- sapply(1:20,function(i){
 ##'     sra(list(sample(l),sample(l),sample(l)))[i]
 ##' })
 ##' c(mean(aa),sd(aa))
-##' 
+##'
 #' @export
 ##' @author Thomas A. Gerds <tag@@biostat.ku.dk>
-plot.sra <- function(x, xlim, ylim, xlab="List depth", ylab="Agreement", add=FALSE, ...){
+plot.sra <- function(x, xlim, ylim, xlab="List depth", ylab="Sequential rank agreement", add=FALSE, ...){
     if (missing(xlim)) xlim=c(1,length(x))
     if (missing(ylim)) ylim=c(0,max(x))
     axis1.DefaultArgs <- list()

### 2015-06-03 R/sra.R

@@ -71,7 +71,7 @@ sra.matrix <- function(object, B=1, na.strings=NULL, nitems=nrow(object)) {
 
     ## Expand the columns in the matrix to have length unique.items
     if (nitems>nrow(object)) {
-        glue <- matrix(rep(NA, ncol(object)*(nitems - nrow(object))), ncol=ncol(object))
+        glue <- matrix(rep(NA, NCOL(object)*(nitems - nrow(object))), ncol=NCOL(object))
         object <- rbind(object, glue)
     }
     object <- lapply(1:NCOL(object),function(j)object[,j]) # Convert matrix to list
@@ -160,7 +160,7 @@ sra.list <- function(object, B=1, na.strings=NULL, nitems=max(sapply(object, len
                          res <- sracppfull(rankmat)
                          res
                      })
-    agreement <- rowMeans(tmpres)
+    agreement <- sqrt(rowMeans(tmpres))
     names(agreement) <- items
     class(agreement) <- "sra"
     attr(agreement, "B") <- B

### 2015-06-10 latex/paper.Rnw
Smaller update

@@
 
-If the lists generally agree on all items then we expect the
-sequential rank agreement curve to be flat, but if there are changes
-in the levels of rank agreement then this suggests that there are sets
-of items that the lists agree on and other sets where they rank the
-items differently. In particular, we generally expect the agreement
-among the lists to be better towards the top of the lists and worse
-towards the end of the lists, and in this case the sequential rank
-agreement curve will start at a low level, and then have an increase
-before it levels off.
-
 \begin{figure}[tb]
@@
 \end{figure}
+
+If the lists generally agree on all items then we expect the
+sequential rank agreement curve to be flat, but if there are changes
+in the levels of rank agreement then this suggests that there are sets
+of items that the lists agree on and other sets where they rank the
+items differently. In particular, we generally expect the agreement
+among the lists to be better towards the top of the lists and worse
+towards the end of the lists, and in this case the sequential rank
+agreement curve will start at a low level, and then have an increase
+before it levels off.
 
@@
 The agreement, $A(X_p)$, cannot be computed directly for all
-predictors when we have censored lists because we only do not know the
-exact rank assigned to each item in all lists.  Instead we assume that
-the rank assigned to an item $X_p$ that is not present in list $l$
-(i.e., is censored) is uniformly distributed among the ranks that have
-\emph{not} been assigned for list $l$.  The rankings within a single
-list are clearly not independent since each rank must appear exactly
-in each list. Thus we cannot simply assign
+predictors in the presence of censored lists because the exact rank
+for some item is unknown in some of the lists.  The rankings within a
+single list are clearly not independent since each rank must appear
+exactly once in each list. Thus we cannot simply assign the same
+number (e.g., the mean of the unassigned ranks) to the censored items
+since that would result in less variation of the agreement, and it
+would artificially introduce a (downward) bias of agreement for items
+that are censored in multiple lists.
 
-Thus, by randomizing unobserved items as described in
-\eqref{formula:censored-agreement}
-
-The algorithm is described in
+Instead we assume that the rank assigned to an item $X_p$ that is not
+present in list $l$ (i.e., is censored) is uniformly distributed among
+the ranks that have \emph{not} been assigned for list $l$. By
+randomizing unobserved items we attain one realization of the rankings
+where all lists are fully observed. By randomizing a large number of
+times we can compute \eqref{def:sra} for each realization, and then
+compute the sequential rank agreement as the pointwise (for each
+depth) average of the rank agreements. The algorithm is described in
 detail in Algorithm~\ref{sra-algorithm}.
@@
 \end{algorithm}
+
+Discuss assumption
+
+Present stuff that is long enough
+
+information
+
+in
+\eqref{formula:censored-agreement}
 

### 2015-06-15 latex/paper.Rnw
Tekst

@@
 \author{Thomas Alexander Gerds}
+\author{Andreas Jensen}
 \author{Kasper Brink-Jensen}
 \affil{Biostatistics, University of Copenhagen}
+
+\renewcommand{\topfraction}{.85}
+\renewcommand{\bottomfraction}{.7}
+\renewcommand{\textfraction}{.15}
+\renewcommand{\floatpagefraction}{.66}
+\renewcommand{\dbltopfraction}{.66}
+\renewcommand{\dblfloatpagefraction}{.66}
+\setcounter{topnumber}{9}
+\setcounter{bottomnumber}{9}
+\setcounter{totalnumber}{20}
+\setcounter{dbltopnumber}{9}
 
@@
 }
+
 @ %
@@
 distance of the individual rankings of the lists from the average
-ranking for each of the items we have seen until depth $d$.
+ranking for each of the items we have seen until depth $d$. Note that
+the terms that are part of \eqref{def:sra} are not independent since
+each list contains the ranks from 1 tp $P$ exactly once. Hence we
+simply use the pooled standard deviation as a measure of the average
+distance between rankings.
 
@@
 With fully observed lists we can plot the sequential rank agreement
-\eqref{def:sra} as a function of depth $d$.  An example is seen in the
-left panel of Figure~\ref{fig:example1} where four analysis methods
-were used to rank genes (for further detail see the application
-section below). The sequential rank agreement at depth $d$ can be
+\eqref{def:sra} as a function of depth $d$. An example is seen in top
+panels of Figure~\ref{fig:example1} where four analysis methods were
+used to rank 3051 gene expression values measured on 38 tumor mRNA
+samples in order to improve the classification of acute leukemias
+between two types: acute lymphoblastic leukemia (ALL) or acute myeloid
+leukemia (AML) \citep{Golub1999}. Preprocessing of the gene expression
+data was done as described in \citet{Dudoit2002} and the four
+different analysis approaces were: marginal two-sample $t$ tests,
+marginal logistic regression analyses, logistic regression eleastic
+net, and marginal maximum information content correlations (MIC)
+\citep{Reshef2011} which resulted in the rankings seen in
+Table~\ref{tab1}.  The sequential rank agreement at depth $d$ can be
 interpreted as how many ranks apart do we on average expect the items
 found until depth $d$ to be so smaller values on the $y$ axis
-correspond to better agreement among the lists than larger values.
+correspond to \emph{better} agreement among the lists than larger
+values.
+
+<<example1, echo=FALSE>>=
+library(SuperRanker)
+library(glmnet)
+library(minerva)
+library(changepoint)
+library(randomForestSRC)
+
+# Read data
+library(multtest)
+data(golub)
+
+y <- golub.cl
+x <- golub
+
+topk <- 20
+
+producelists <- function(x, y) {
+    nitems <- nrow(x)
+    index <- seq(1, nrow(golub))
+    ## d <- data.frame(y, x)
+
+    ## Marginale t-tests
+    mt.p <- sapply(index, function(i) { t.test(x[i,] ~ y)$p.value } )
+    list1 <- order(mt.p)
+
+    ## Marginale logreg-tests
+    mlogreg.p <- sapply(index, function(i) { drop1(glm(y ~ x[i,], family=binomial), test="Chisq")[2,5] } )
+    list2 <- order(mlogreg.p)
+
+    ## Elastic net
+    X <- scale(t(x))
+    enet <- glmnet(X, y, family="binomial", alpha=.8)
+    nyres <- cv.glmnet(X, y, family="binomial", alpha=.8)
+    coefficients <- coef(enet, s=nyres$lambda.1se)[-1]
+    nonzeros <- sum(coefficients!=0)
+    list3 <- order(abs(coefficients), decreasing=TRUE)
+    if (nonzeros<nitems) {
+        list3[(nonzeros+1):nitems] <- NA
+    }
+    ## MIC
+    MIC <- sapply(index, function(i) { mine(x[i,], y)$MIC})
+    list4 <- order(MIC, decreasing=TRUE)
+
+                                        # Random Forest
+###    dd <- data.frame(y=factor(y), t(x))
+                                        # dd <- data.frame(y=y, t(x))
+###    f1 <- rfsrc(y ~ ., data=dd, ntree=100)
+###    variables <- abs(f1$importance[,1])
+###    num.undecided <- sum(variables==0)
+###    list5 <- order(variables, decreasing=TRUE)
+###    list5[(length(variables)-num.undecided):length(variables)] <- 0
+
+    cbind(list1,list2,list3,list4)
+}
+
+## Original run
+B <- 20
+inputmatrix <- producelists(x, y)
+colnames(inputmatrix) <- c("T", "LogReg", "ElasticNet", "MIC", "RF")[1:ncol(inputmatrix)]
+res <- sra(inputmatrix, B=B)
+
+## Now make the same run but where we only have top 50 lists
+inputmatrix2 <- inputmatrix
+inputmatrix2[(topk+1):nrow(inputmatrix),] <- NA
+res2 <- sra(inputmatrix2, B=B)
+
+@
+
+<<fig1a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE>>=
+ysize <- 1000
+nreference <- 20
+## Plot the line
+plot(res[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
+
+## Compute random list backgrounds
+null <- random_list_sra(inputmatrix, B=B, n=nreference)
+
+bcolor <- makeTransparent("red", alpha=80)
+bcolor2 <- makeTransparent("blue", alpha=80)
+
+## Plot the random lists
+www <- smooth_sra(null)
+polygon(c(1:length(www$lower), rev(1:length(www$lower))),
+        c(www$lower, rev(www$upper)),
+        col=bcolor2, border=NA)
+
+## Load the data for the methods
+makereference <- function(x, y, n=100, B=20) {
+    sapply(1:n, function(i) { sra(producelists(x, sample(y)), B=B) }  )
+}
+## Uncomment the two lines below to redu the analysis
+###null2 <- makereference(x, y, n=100, B=B)
+###save(null2, file="R/fig1null2.rda")
+load("R/fig1null2.rda")
+
+## Plot the reference sras
+www2 <- smooth_sra(null2)
+polygon(c(1:length(www2$lower), rev(1:length(www2$lower))),
+        c(www2$lower, rev(www2$upper)),
+        col=bcolor, border=NA)
+
+@ %
+
+<<fig1b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE >>=
+## Same as above but zoomed in
+ysize <- 50
+plot(res[1:ysize], lwd=3, col="black", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
+
+## Plot the random lists
+polygon(c(1:length(www$lower), rev(1:length(www$lower))),
+        c(www$lower, rev(www$upper)),
+        col=bcolor2, border=NA)
+
+## Plot the reference sras
+www2 <- smooth_sra(null2)
+polygon(c(1:length(www2$lower), rev(1:length(www2$lower))),
+        c(www2$lower, rev(www2$upper)),
+        col=bcolor, border=NA)
+
+@ %
+
+<<fig2a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE>>=
+ysize <- 1000
+nreference <- 20
+## Plot the line
+plot(res2[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
+
+## Compute random list backgrounds
+null <- random_list_sra(inputmatrix2, B=B, n=nreference)
+
+## Plot the random lists
+www <- smooth_sra(null)
+polygon(c(1:length(www$lower), rev(1:length(www$lower))),
+        c(www$lower, rev(www$upper)),
+        col=bcolor2, border=NA)
+
+## Load the data for the methods
+makereference <- function(x, y, n=100, B=20) {
+    sapply(1:n, function(i) { mymat <- producelists(x, sample(y)) ; mymat[(topk+1):nrow(mymat),] <- NA ; sra(mymat, B=B) }  )
+}
+## Uncomment the two lines below to redu the analysis
+##null2 <- makereference(x, y, n=100, B=B)
+##save(null2, file="R/fig1null2c.rda")
+load("R/fig1null2c.rda")
+
+## Plot the reference sras
+www2 <- smooth_sra(null2)
+polygon(c(1:length(www2$lower), rev(1:length(www2$lower))),
+        c(www2$lower, rev(www2$upper)),
+        col=bcolor, border=NA)
+
+@ %
+
+<<fig2b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE >>=
+## Same as above but zoomed in
+ysize <- 50
+plot(res2[1:ysize], lwd=3, col="black", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
+
+## Plot the random lists
+polygon(c(1:length(www$lower), rev(1:length(www$lower))),
+        c(www$lower, rev(www$upper)),
+        col=bcolor2, border=NA)
+
+## Plot the reference sras
+www2 <- smooth_sra(null2)
+polygon(c(1:length(www2$lower), rev(1:length(www2$lower))),
+        c(www2$lower, rev(www2$upper)),
+        col=bcolor, border=NA)
+
+@ %
+
+<<label=tab1,echo=FALSE,results=tex>>=
+library(xtable)
+xxx <- as.data.frame(cbind(1:10, inputmatrix[1:10,]))
+colnames(xxx) <- c("Ranking", colnames(xxx)[2:5])
+xxx2 <- xtable(xxx, caption = "Top 10 list of ranked results from the Golub data. Numbers indicate the predictor/gene for the given ranking and method")
+digits(xxx2) <- 0
+print(xxx2, table.placement = "tb", caption.placement = "top", include.rownames=FALSE)
+#print(xtable(xxx,  label = "tab:one",
+
+@
 
@@
 \includegraphics[width=.45\textwidth]{paper-fig1a}
+\includegraphics[width=.45\textwidth]{paper-fig2a}
 \includegraphics[width=.45\textwidth]{paper-fig1b}
+\includegraphics[width=.45\textwidth]{paper-fig2b}
+
 \end{center}
- \caption{Left panel: Sequential rank agreement for 4 different analysis methods
-   applied to the 3051 genes in the Golub data (solid line). The dashed line shows the
-   corresponding estimated sequential rank agreement for censored
-   lists based purely on the top 50 ranked genes. Right panel contains
-   the same information but has zoomed in on the first part of the $x$-axis.}
+ \caption{Left panels: Sequential rank agreement for 4 different analysis methods
+   applied to the 3051 genes in the Golub data. The two plots on the
+   right show the sequential rank agreement for the same data but
+   where only the top-\Sexpr{topk} ranked items are available. The blue
+   and red areas correspond to the independent and randomized
+   reference hypothesis areas, respectively. The bottom plots are
+   identical to the corresponding top plots but have been zoomed in on
+   the first part of the $x$ axis.}
  \label{fig:example1}
@@
 
-In a classical paper by \citet{Golub1999} a dataset of 3051 gene
-expression values were measured on 38 tumor mRNA samples in order to
-improve the classification of acute leukemias between two types: acute
-lymphoblastic leukemia (ALL) or acute myeloid leukemia
-(AML). Preprocessing of the gene expression data was done as described
-in \citep{Dudoit2002}.
-
-We analyzed these gene expression data using four different
-approaches: marginal two-sample $t$ tests, marginal logistic
-regression analyses, logistic regression eleastic net, and marginal
-maximum information content correlations (MIC) \citep{Reshef2011}. For
-the first two methods, the genes were ranked according to $p$ value.
-
-<<example1, echo=FALSE>>=
-library(SuperRanker)
-library(glmnet)
-library(minerva)
-library(changepoint)
-library(randomForestSRC)
-
-# Read data
-library(multtest)
-data(golub)
-
-y <- golub.cl
-x <- golub
-
-producelists <- function(x, y) {
-    nitems <- nrow(x)
-    index <- seq(1, nrow(golub))
-                                        # d <- data.frame(y, x)
-
-                                        # Marginale t-tests
-    mt.p <- sapply(index, function(i) { t.test(x[i,] ~ y)$p.value } )
-    list1 <- order(mt.p)
-
-                                        # Marginale logreg-tests
-    mlogreg.p <- sapply(index, function(i) { drop1(glm(y ~ x[i,], family=binomial), test="Chisq")[2,5] } )
-    list2 <- order(mlogreg.p)
-
-                                        # Elastic net
-    X <- scale(t(x))
-    enet <- glmnet(X, y, family="binomial", alpha=.8)
-    nyres <- cv.glmnet(X, y, family="binomial", alpha=.8)
-    coefficients <- coef(enet, s=nyres$lambda.1se)[-1]
-    nonzeros <- sum(coefficients!=0)
-    list3 <- order(abs(coefficients), decreasing=TRUE)
-    if (nonzeros<nitems) {
-        list3[(nonzeros+1):nitems] <- NA
-    }
-                                        # MIC
-    MIC <- sapply(index, function(i) { mine(x[i,], y)$MIC})
-    list4 <- order(MIC, decreasing=TRUE)
-
-                                        # Random Forest
-###    dd <- data.frame(y=factor(y), t(x))
-                                        # dd <- data.frame(y=y, t(x))
-###    f1 <- rfsrc(y ~ ., data=dd, ntree=100)
-###    variables <- abs(f1$importance[,1])
-###    num.undecided <- sum(variables==0)
-###    list5 <- order(variables, decreasing=TRUE)
-###    list5[(length(variables)-num.undecided):length(variables)] <- 0
-
-    cbind(list1,list2,list3,list4)
-}
-
-inputmatrix <- producelists(x, y)
-colnames(inputmatrix) <- c("T", "LogReg", "ElasticNet", "MIC", "RF")[1:ncol(inputmatrix)]
-
-res <- sqrt(sra(inputmatrix, B=20))
-inputmatrix2 <- inputmatrix
-inputmatrix2[51:nrow(inputmatrix),] <- NA
-res2 <- sqrt(sra(inputmatrix2, B=20))
-
-@
-
-<<fig1a, fig=TRUE,echo=FALSE,include=FALSE>>=
-ysize <- 1000
-plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
-lines(res2[1:ysize], lwd=2, col="purple")
-
-@ %
-
-<<fig1b, fig=TRUE,echo=FALSE,include=FALSE >>=
-ysize <- 80
-plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
-lines(res2[1:ysize], lwd=2, col="purple")
-
-@ %
-
-<<fig2a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE>>=
-ysize <- 1000
-plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
-lines(res2[1:ysize], lwd=2, col="purple")
-##
-# col1 <- makeTransparent()
-null <- sqrt(random_list_sra(inputmatrix, B=1, n=40))
-null3 <- sqrt(random_list_sra(inputmatrix, B=20, n=20))
-## matlines(null[1:ysize,], col="gray", lty=1)
-
-##
-
-##null2 <- sapply(1:40, function(i) {
-##    sra(producelists(x, sample(y)), B=1)
-##} )
-## matlines(null2[1:ysize,], col="lightblue", lty=1)
-
-load("R/fig1null.rda")
-
-bcolor <- makeTransparent("lightblue", alpha=80)
-bcolor2 <- makeTransparent("red", alpha=80)
-bcolor3 <- makeTransparent("green", alpha=80)
-
-www <- smooth_sra(null2)
-polygon(c(1:length(www$lower), rev(1:length(www$lower))),
-        c(www$lower, rev(www$upper)),
-        col=bcolor, border=NA)
-
-www <- smooth_sra(null)
-polygon(c(1:length(www$lower), rev(1:length(www$lower))),
-        c(www$lower, rev(www$upper)),
-        col=bcolor2, border=NA)
-
-www <- smooth_sra(null3)
-polygon(c(1:length(www$lower), rev(1:length(www$lower))),
-        c(www$lower, rev(www$upper)),
-        col=bcolor3, border=NA)
-
-@ %
-
-<<fig2b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE >>=
-ysize <- 80
-plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
-lines(res2[1:ysize], lwd=2, col="purple")
-##
-# col1 <- makeTransparent()
-null <- sqrt(random_list_sra(inputmatrix, B=1, n=40))
-null3 <- sqrt(random_list_sra(inputmatrix, B=20, n=20))
-## matlines(null[1:ysize,], col="gray", lty=1)
-
-##
-
-##null2 <- sapply(1:40, function(i) {
-##    sra(producelists(x, sample(y)), B=1)
-##} )
-## matlines(null2[1:ysize,], col="lightblue", lty=1)
-
-load("R/fig1null.rda")
-
-bcolor <- makeTransparent("lightblue", alpha=80)
-bcolor2 <- makeTransparent("red", alpha=80)
-bcolor3 <- makeTransparent("green", alpha=80)
-
-www <- smooth_sra(null2)
-polygon(c(1:length(www$lower), rev(1:length(www$lower))),
-        c(www$lower, rev(www$upper)),
-        col=bcolor, border=NA)
-
-www <- smooth_sra(null)
-polygon(c(1:length(www$lower), rev(1:length(www$lower))),
-        c(www$lower, rev(www$upper)),
-        col=bcolor2, border=NA)
-
-www <- smooth_sra(null3)
-polygon(c(1:length(www$lower), rev(1:length(www$lower))),
-        c(www$lower, rev(www$upper)),
-        col=bcolor3, border=NA)
-
-@ %
-
-\begin{figure}[tb]
-\begin{center}
-\includegraphics[width=.45\textwidth]{paper-fig2a}
-\includegraphics[width=.45\textwidth]{paper-fig2b}
-\end{center}
- \caption{Sequential rank agreement for 4 different analysis methods
-   applied to the Golub data (thick red line). The red and blue shaded
- areas represent the pointwise 95\% normal areas for the XXX and XXY benchmarks,
-respectively.}
- \label{fig:case1}
-\end{figure}
-
-<<label=tab1,echo=FALSE,results=tex>>=
-library(xtable)
-xxx <- as.data.frame(inputmatrix[1:10,])
-xxx2 <- xtable(xxx)
-digits(xxx2) <- 0
-print(xxx2)
-#print(xtable(xxx, caption = "Top 10 list of results", label = "tab:one",
-#digits = c(0, 0, 2, 0, 2, 3, 3)), table.placement = "tbp",
-#caption.placement = "top")
-
-@
-
 <<>>=

### 2015-06-16 latex/paper.Rnw
Data og tekst

@@
  \caption{Left panels: Sequential rank agreement for 4 different analysis methods
-   applied to the 3051 genes in the Golub data. The two plots on the
-   right show the sequential rank agreement for the same data but
-   where only the top-\Sexpr{topk} ranked items are available. The blue
+   applied to the 3051 genes in the Golub data. Right panels:
+   Corresponding xssequential rank agreement for the same data but
+   where only the top \Sexpr{topk} ranked items are available. The blue
    and red areas correspond to the independent and randomized
@@
 
+The proposed approach for censored rankings is based on two
+assumptions: 1) that the items of interest will be ..., and that 2)
+that the censored rankings provide so little information that it is
+reasonable to assume that the missing rankings
+
+For censored data it is also imminent that ...
+
 Discuss assumption
@@
 
-in
-\eqref{formula:censored-agreement}
+For censored lists we expect a change point around the depth
+where the lists are censored. This is an artefact stemming from the
+fact that
 
@@
 \end{equation}
-since we have no information about the ordering of the items after a
+since we have no real information about the ordering of the items after a
 depth $d$ where all lists are censored.
@@
 the mean then we
-
-FIXME: If instead of running through all elements of $\tilde R_1(X_p)$
-one would use the average rank in $\tilde R_1(X_p)$ we would end up with
-a too small variance.
-
-FIXME: For censored lists we expect a change point around the depth
-where the lists are censored. This is an artefact stemming from the
-fact that
 

### 2015-06-17 latex/paper.Rnw
Tekst

--- 2015-06-15 latex/paper.Rnw
@@
 net, and marginal maximum information content correlations (MIC)
-\citep{Reshef2011} which resulted in the rankings seen in
+\citep{Reshef2011}. For the first two methods, the genes were ranked
+according to $p$ value, for logistic regression the genes were ordered
+by size of the corresponding coefficients (after standardization), and
+MIC was ordered by correlation which resulted in the rankings seen in
 Table~\ref{tab1}.  The sequential rank agreement at depth $d$ can be
@@
  \caption{Left panels: Sequential rank agreement for 4 different analysis methods
-   applied to the 3051 genes in the Golub data. The two plots on the
-   right show the sequential rank agreement for the same data but
-   where only the top-\Sexpr{topk} ranked items are available. The blue
+   applied to the 3051 genes in the Golub data. Right panels:
+   Corresponding sequential rank agreement for the same data but
+   where only the top \Sexpr{topk} ranked items are available. The blue
    and red areas correspond to the independent and randomized
@@
 towards the end of the lists, and in this case the sequential rank
-agreement curve will start at a low level, and then have an increase
-before it levels off.
+agreement curve will start at a low level and then increase until it
+levels off.
 
@@
 
-Discuss assumption
+The proposed approach is based on two assumptions: 1) that the items
+of interest will be found towards the top of the lists, and 2) that
+the censored rankings provide so little information that it is
+reasonable to assume that the missing rankings provide no information
+and might as well be considered random.  The first assumption is
+justifiable because we have already accepted that it is reasonable to
+rank the items in the first place. The second assumption is fair
+in the light of the first assumption provided that we have a
+``sufficiently large'' part of the top of the lists available.
 
-Present stuff that is long enough
+When the two assumptions are combined then it is clear that the
+interesting part of the sequential rank agreement curves must be for
+smaller depths. Unless we have additional prior knowledge about the
+distribution of the unranked items then we can restrict attention of
+the sequential rank agreement to
+\begin{equation}
+d' \leq \max(k_1, \ldots, k_L),
+\end{equation}
+since we have no real information about the ordering of the items after a
+depth $d'$ where all lists are censored.
 
-information
-
-in
-\eqref{formula:censored-agreement}
-
-Unless we have additional prior knowledge about the distribution of the
-unranked items then we can restrict attention of the sequential rank
-agreement to
-\begin{equation}
-d \leq \max(k_1, \ldots, k_L),
-\end{equation}
-since we have no information about the ordering of the items after a
-depth $d$ where all lists are censored.
-
- ess contains full set of ranks and by using
-the ... instead of applying the average ...
-
-the mean then we
-
-FIXME: If instead of running through all elements of $\tilde R_1(X_p)$
-one would use the average rank in $\tilde R_1(X_p)$ we would end up with
-a too small variance.
-
-FIXME: For censored lists we expect a change point around the depth
-where the lists are censored. This is an artefact stemming from the
-fact that
+Like for fully observed lists we generally expect the sequential rank
+agreement to start low and then increase unless the lists are
+completely unrelated (in which case the sequential rank agreement will
+be constant at a high level) or if the lists mostly agree on the
+ranking (in which case the sequential rank agreement will also be
+constant but at a low level). For censored lists we also expect a
+change point around the depth where the lists are censored. This is an
+artefact stemming from the fact that we assume that the remainder of
+the lists can be replaced by a simple permutation of the missing
+items.
 
@@
 
-\section{Applications}
-
-\subsection{Comparing results across different method}
-
-<<>>=
-library(changepoint)
-cp1 <- cpt.mean(res, method="BinSeg")
-cp2 <- cpt.mean(res2, method="BinSeg")
-cp1@cpts
-cp1@param.est
-cp2@cpts
-cp2@param.est
-@
-
-\subsection{Stability of results}
-
-\subsection{Evaluating results from top-$k$ lists}
-
-Bootstrap across a single method and compare results. Discuss collinearity
-
-\section{Comparison to ...}
+\section{Comparison to average overlap}
 
@@
 
-In a classical paper by \citet{Golub1999} a dataset of 3051 gene
-expression values were measured on 38 tumor mRNA samples in order to
-improve the classification of acute leukemias between two types: acute
-lymphoblastic leukemia (ALL) or acute myeloid leukemia
-(AML). Preprocessing of the gene expression data was done as described
-in \citep{Dudoit2002}.
-
-We analyzed these gene expression data using four different
-approaches: marginal two-sample $t$ tests, marginal logistic
-regression analyses, logistic regression eleastic net, and marginal
-maximum information content correlations (MIC) \citep{Reshef2011}. For
-the first two methods, the genes were ranked according to $p$ value.
-
 In this section we apply the method to the ColonData data from the R
@@
 xvals <- 1:500
-plot(myres, xlim=c(1, 300), ylim=c(0,1), type="l")
-lines(xvals[-c(1,2)], myres2[1:498], col="red")
+plot(xvals, myres[xvals], xlim=c(1, 150), ylim=c(0,1), type="l", lwd=2, ylab="Average overlap", xlab="List depth")
+lines(xvals[-c(1,2)], myres2[1:498], col="red", lwd=2)
 @
@@
 <<fig-ao2, echo=FALSE, fig=TRUE,include=FALSE>>=
-mmyres <- sqrt(sra(inputmatrix[,1:2]))
-mmyres2 <- sqrt(sra(inputmatrix[-c(1,2),1:2]))
+mmyres <- sra(inputmatrix[,1:2])
+mmyres2 <- sra(inputmatrix[-c(1,2),1:2])
 xvals <- 1:500
-plot(mmyres, xlim=c(1, 300), type="l")
-lines(xvals[-c(1,2)], mmyres2[1:498], col="red")
+plot(xvals, mmyres[xvals], xlim=c(1, 150), ylim=c(0, 110), type="l", lwd=2, ylab="Sequential rank agreement", xlab="List depth")
+lines(xvals[-c(1,2)], mmyres2[1:498], col="red", lwd=2)
 @
@@
 \end{center}
- \caption{Sequential rank agreement for 4 different analysis methods
+ \caption{Average overlap (left figure) and sequential rank agreement
+   (right figure) for two different analysis methods
    applied to the Golub data (thick red line). The red and blue shaded
@@
 \end{figure}
+
+\section{Applications}
+
+\subsection{Comparing results across different method}
+
+<<>>=
+#library(changepoint)
+#cp1 <- cpt.mean(res, method="BinSeg")
+#cp2 <- cpt.mean(res2, method="BinSeg")
+#cp1@cpts
+#cp1@param.est
+#cp2@cpts
+#cp2@param.est
+@
+
+\subsection{Stability of results}
+
+\subsection{Evaluating results from top-$k$ lists}
+
+Bootstrap across a single method and compare results. Discuss collinearity
 

### 2015-06-22 latex/paper.Rnw
...

@@
 
-To interpretate the sequential rank agreement scale we propose by
-several different benchmark values. Our data free benchmark value is
+To interpret the sequential rank agreement values we propose two
+different benchmark values corresponding to two different
+hypotheses. We wish to determine if we observe better agreement that
+what would be expected if there were no relevant information available
+in the data.
+
+The first reference hypothesis is 
+\begin{eqnarray*}
+H_0  & : &  \text{The list rankings correspond to complete randomly}\\
+       &  & \text{permuted lists} \\
+\end{eqnarray*}
+
+\begin{eqnarray*}
+H_1 & :&  \text{The list rankings correspond to ranked lists based on}\\
+& &   \text{data containing no information}
+\end{eqnarray*}
+
+It is important to stress 
+Note that
+
+Compute the sequential rank agreement for a large number of random
+permutations ... with the same number of lists and with the same amount of censoring.
+
+understood about the outcome
+
+Clearly hypothesis .... is ... no association and that the methods are
+also completely independent.
+
+Our data free benchmark value is
 obtained by computing the sequential rank agreement of $L$ lists each
@@
 Lidt spøjs ting man matcher dem op imod
-
-\subsection{Changepoint analysis}
 
@@
 
+\subsection{Changepoint analysis}
+
 \bibliographystyle{rss}

### 2015-06-26 latex/paper.Rnw
Draft

--- 2015-06-17 latex/paper.Rnw
@@
 
-To interpretate the sequential rank agreement scale we propose by
-several different benchmark values. Our data free benchmark value is
-obtained by computing the sequential rank agreement of $L$ lists each
-containing a permutation of $P$ items. Practically this benchmark
-value can be obtained by averaging the result of repeated computer
-experiments. In the uncensored case each experiment contains $L$ lists
-of random permutations of the numbers $1,\dots,P$. In the censored
-case only the uncensored part of the orginal lists is randomly
-permuted.
+To interpret the sequential rank agreement values we propose two
+different benchmark values corresponding to two different
+hypotheses. We wish to determine if we observe better agreement than
+what would be expected if there were no relevant information available
+in the data.
 
-In our experience XX repetitions are sufficient to reduce Monte Carlo
-error. Table Appendix A shows benchmark values obtained with
+The first reference hypothesis is 
+\begin{eqnarray*}
+H_0  & : &  \text{The list rankings correspond to complete randomly}\\
+       &  & \text{permuted lists} 
+\end{eqnarray*}
+which not only assumes that there is no information in the rankings
+but also that the methods are completely independent.
 
-Our no-information benchmark can be applied in situations where each
-list is an importance ranking of $P$ predictors resulting from
-regression analyses. The idea is to repeat the regression analyses in
-modified data where the outcome vector is randomly permuted. This
-artificially removes any potential information regarding the
-regression that the real data may contain.
+Alternatively, we can remove the restriction on the independence among
+the methods by only requiring that there is no information in the
+ranking but that the rankings are all based on applying the 
+method/approaches to the same data
+\begin{eqnarray*}
+H_1 & :&  \text{The list rankings are based on data containing}\\
+& &   \text{no association to the outcome.}
+\end{eqnarray*}
+$H_0$ is clearly the strictest (and also potentially quite
+unrealistic) null hypothesis but we can easily obtain realizations
+from that null hypothesis simply by permuting the items for each list
+and then computing the sequential rank agreement for the permuted
+lists. In the uncensored case each experiment contains $L$ lists of
+random permutations of the numbers $1,\dots,P$. For the censored case
+we permute the number as before but censor the same number of items
+for each list as were censored in the original dataset. The sequential
+rank agreement curve from the the original data can then be compared
+to the pointwise rank agreements obtained from under $H_0$.
 
-- third approach?
+To obtain a reference distribution under $H_1$ we use a randomization
+test approach on the original data to remove any associations to the
+outcome in the data that was used to generate the rankings before
+computing the ranking and subsequently the sequential rank agreement
+from the randomized data. The idea is to repeat the ranking procedures
+in the modified data where the outcome vector is randomly permuted. In
+practice we permute the outcomes but keep any predictors before
+ranking to preserve any structure in the predictors. This artificially
+removes any potential information regarding the regression that the
+real data may contain. The randomization approach requires that we
+have the original data available, so we can compute the ranked lists
+from the randomized data and not just the ranked lists and as such it
+may not be possible to evaluate $H_1$ in all situations.
 
-\subsection{Independent lists}
+If the sequential rank agreement for the original data lies
+substantially below the distribution of the sequential rank agreements
+obtained under either $H_0$ or $H_1$ then this suggests that the
+original ranked lists agree \emph{more} that we would expect if the
+ranked lists were based on data with no information.
 
-\subsection{Permuted outcomes + analyses}
+Figure~\ref{fig:example1} shows the sequential rank distributions
+based on $400$ samples under $H_0$ and $H_1$, respectively. Not
+surprisingly, the null distribution of $H_1$ has lower sequential rank
+agreement than the null distribution of $H_0$ because the four methods
+used to rank the data ($t$ test, logistic regression, elastic net, and
+MIC) generally identify and rank the same set of predictors that show
+spurious associations. The two bottom panels in
+Figure~\ref{fig:example1} also indicate that the observed sequential
+rank agreement is better than what would be expected for the full
+dataset but that the censored data (the bottom-right plot) suggests
+that there may be at most 1 or 2 ranked items that are towards the top
+of the lists that yield a result better than what would be expected.
 
-Do the following a large number of times
-\begin{enumerate}
-  \item Permute outcome vector
-  \item Redo analyses for all $L$ methods
-  \item Compute sequential rank agreement for
-\end{enumerate}
-
-Lidt spøjs ting man matcher dem op imod
-
-\subsection{Changepoint analysis}
+It is important to stress that neither $H_0$ nor $H_1$ are concerned
+with whether the $L$ rankings are relevant. Both hypotheses are purely
+considering how the rankings agree in a situation where there is no relevant
+information available in the data used for creating the rankings.
 
 \section{Comparison to average overlap}
+
+The average overlap is widely used in comparing (two) ranked list and
+the idea behind the average overlap closely resembles the sequential
+rank agreement \citep{Fagin2003,Webber2010}. Here we will present an
+extention of the average overlap to more that two lists and compare this
+to sequential rank agreement.
 
@@
 \begin{equation}
-O(d) = \frac{| \cap_{l=1}^L \{\pi_l(r); r\leq d \} |}{d}.
+O(d) = \frac{| \cap_{l=1}^L \{\pi_l(r); r\leq d \} |}{d}. \label{def:overlap}
 \end{equation}
-Typically, the overlap is only considered for pairwise comparisons
+Typically, the overlap is only considered for pairwise comparisons but in
+\eqref{def:overlap} we make no such distinction.
 
 The average overlap at depth $d$ is defined as the average of the
-overlaps until $d$,
+overlaps until depth $d$,
 \begin{equation}
@@
 
-In this section we apply the method to the ColonData data from the R
-package MAMA (Ihnatova, 2013), which consists of gene expression
-measurements from a subset of 500 genes from 77 individuals from
-Denmark, 36 from Australia and 41 from Japan as well as information on
-microsattelite stability (stable MSS/unstable MSI) which is often used
-as an indicator for colorectal cancer. The genes were tested for
-difference in expression in the MSI/MSS groups using the limma package
-(Smyth, 2005). The genes were then ranked according to p–values giving
-three lists that we wish to compare, the top of these can be seen in
-Table 4.
+There are three main drawbacks of the average overlap method: First it
+is highly sensitive to items at the top of the list. The average
+overlap profile changes dramatically depending on the occurrence of
+similar items in the lists. Secondly, when the number of lists, $L$,
+is high then it might be difficult to obtain a non-zero overlap (and
+hence a non-zero average overlap) at the beginning of the lists because
+the overlap does not increase until an item is present in all $L$
+lists. Finally, the average overlap has an interpretation in terms of
+moving averages of  percentages (of $L$ sets) which somewhat less
+intuitive than simply the average rank distance that we propose.
+
+In this section we compare the average overlap and the sequential rank
+agreement to the Golub data  \citep{Golub1999}. To make it more
+similar to the way the average ovelap is typically used we restrict
+attention to only two analysis methods: $t$ tests and logistic
+regression. Thus we have two marginal model approaches that are very
+similar and the first two columns of Table~\ref{tab1} show the actual top-10
+ranking of the predictors.
 
@@
  \caption{Average overlap (left figure) and sequential rank agreement
-   (right figure) for two different analysis methods
-   applied to the Golub data (thick red line). The red and blue shaded
- areas represent the pointwise 95\% normal areas for the XXX and XXY benchmarks,
-respectively.}
+   (right figure) for comparing two different analysis methods
+   (marginal $t$ test and logistic regression)
+   applied to the Golub data. The black lines are based on the full
+   set of predictors while the red lines have the two highest
+   associated predictors removed
+   from the data before computing the average overlap and sequential
+   rank agreement.}
  \label{fig:case1}
 \end{figure}
+
+The black lines Figure~\ref{fig:case1} show the average overlap and
+sequential rank agreement for these data. It is clear from both plots
+that there is perfect agreement towards the top of the ranked lists:
+the average overlap is 1 and the sequential rank agreement is zero.
+However, if we remove those two items from the data (i.e.,
+gene/predictor 2124 and 896) and redo the analyses then we get
+substantially different curves for the average overlap but not for
+sequential rank agreement (red lines in Figure~\ref{fig:case1}). For
+the sequential rank agreement we get roughly the same estimate of
+agreement from rank 3 as we did from the full dataset which is a nice
+property. For the average overlap the curves completely change when
+these two items are missing and give a visually different conclusion
+about the ranking of the items from rank 3 downwards.
 
@@
 
+Censor lists at significane level?
+
+\subsection{Changepoint analysis}
+
 \bibliographystyle{rss}

### 2015-07-08 latex/paper.Rnw
Rettelser

@@
 \author{Thomas Alexander Gerds}
-\author{Andreas Jensen}
+\author{Andreas Kryger Jensen}
 \author{Kasper Brink-Jensen}
@@
   interest. Computer science and genomics are just two areas that need
-  to compare ranked lists in order to in- vestigate similarities
+  to compare ranked lists in order to investigate similarities
   between different analytical methods. Often the aim is to find a
@@
 methods comparison, variable selection
-
-Multiple lists. Weigh top over bottom. Censoring. Interpretation
 

### 2015-07-08 latex/paper.Rnw
Andreas' application - part 1

@@
 
-\subsection{Comparing results across different method}
+We now consider an application of the sequential rank agreement to two
+data sets consisting of MALDI-TOF (Matrix-Assisted Laser
+Desorption/Ionization Time Of Flight) mass spectra obtained from blood
+samples from patients with either benign or malignant ovarian
+tumors. The data sets are subsamples of the Danish MALOVA and DACOVA
+study populations.
 
-<<>>=
-#library(changepoint)
-#cp1 <- cpt.mean(res, method="BinSeg")
-#cp2 <- cpt.mean(res2, method="BinSeg")
-#cp1@cpts
-#cp1@param.est
-#cp2@cpts
-#cp2@param.est
-@
+The MALOVA study is a multidisciplinary Danish study on ovarian cancer
+\cite{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
+with an ovarian tumor and referred for surgery from the participating
+departments of gynecology were enrolled continuously from December
+1994 to May 1999. We use a random subsample of $119$ patients with a
+total of $58$ patients with malignant ovarian cancers as cases and
+$61$ patients with benign ovarian tumors as controls. The DACOVA study
+is similarly a multidisciplinary Danish study on ovarian cancer
+covering about $66\%$ of the population in Denmark
+\cite{bertelsen1991protocol}. It was attempted to continuous enroll
+all patients referred for surgery with an ovarian tumor clinically
+suspected to be cancer during the period 1984 to 1990. We use a random
+subsample of $113$ patients with a total of $54$ malignant ovarian
+cancers and $59$ benign ovarian tumors/gynecologic disorders.
 
-\subsection{Stability of results}
+Proteomic analysis was performed by MALDI-TOF mass spectrometry with a
+Bruker Ultraflex in positive, linear ion mode in the mass-to-charge
+range $800$ to $20000$ Dalton by an average of $500$ laser
+shots. Calibration of the instrument was performed using a protein
+standard mixture. Each spectrum consists of $49642$ samples over the
+mass-to-charge range which we down-sample on an equidistant grid of
+5000 points by linear interpolation. We further preprocess each
+spectrum by first removing the slow-varying baseline intensities with
+the SNIP algorithm \cite{ryan1988snip} followed by a normalization
+with respect to the total ion counts. Finally, we standardize the 5000
+predictors to have column-wise zero mean and unit variance.
 
-\subsection{Evaluating results from top-$k$ lists}
+We use these two data sets to illustrate how the sequential rank
+agreement can be applied in two different scenarios: i) assessing the
+agreement of ranking the importance of the predictors using different
+statistical classification methods and ii) assessing the agreement of
+rankings the predictions (the probability of the classifying a tumor
+as malignant) subject to perturbations in data. In the latter case we
+further illustrate the discrepancy between the ranking stability of
+different statistical methods through a simulation study.
 
-Bootstrap across a single method and compare results. Discuss collinearity
+\subsection{Agreement of predictor ranking}
+
+#includegraphics marginalVsModel.pdf
+
+\subsection{Agreement of predictions}
+
+#includegraphics thomas.pdf
+
+#includegraphics simulationTest2.pdf
 
@@
 
-\subsection{Changepoint analysis}
+Bootstrap across a single method and compare results. Discuss collinearity
+
+Changepoint analysis
 

### 2015-07-09 latex/paper.Rnw
started editing the abstract

--- 2015-07-08 latex/paper.Rnw
@@
 
-\title{Sequential rank agreement methods for comparison of ranked
-  lists}
+\title{Sequential rank agreement methods for comparison of lists}
 \author{Claus Thorn Ekstrøm}
 \author{Thomas Alexander Gerds}
-\author{Andreas Kryger Jensen}
+\author{Andreas Jensen}
 \author{Kasper Brink-Jensen}
 \affil{Biostatistics, University of Copenhagen}
-
 \renewcommand{\topfraction}{.85}
@@
 \begin{abstract}
-  The comparison of ranked lists has been studied for several decades
-  but since the advent of large data types it has received increased
-  interest. Computer science and genomics are just two areas that need
-  to compare ranked lists in order to investigate similarities
+  The comparison of rankings has received increased interest in the
+  last decades. Computer science and genomics are just two areas that
+  need to compare ranked lists in order to investigate similarities
   between different analytical methods. Often the aim is to find a
@@
 methods comparison, variable selection
+
+Multiple lists. Weigh top over bottom. Censoring. Interpretation
 
@@
 }
-
 @ %

### 2015-07-09 latex/paper.Rnw

--- 2015-07-08 latex/paper.Rnw
@@
 
-\title{Sequential rank agreement methods for comparison of ranked
-  lists}
+\title{Sequential rank agreement methods for comparison of lists}
 \author{Claus Thorn Ekstrøm}
@@
 \affil{Biostatistics, University of Copenhagen}
-
 \renewcommand{\topfraction}{.85}

### 2015-07-11 latex/paper.Rnw
smoothed abstract, intro, etc.

@@
 \begin{abstract}
-  The comparison of rankings has received increased interest in the
-  last decades. Computer science and genomics are just two areas that
-  need to compare ranked lists in order to investigate similarities
-  between different analytical methods. Often the aim is to find a
-  measure of similarity for two or more ordered lists that comprise
-  the first k items from each list. Since ordered lists typically are
-  more similar at the top of the list another approach, which is taken
-  in this paper is to ask: for how long are the lists highly similar?
-  Using terminology from agreement analysis we derive the measure
-  depth of agreement using the standard deviation of the ranks, a
-  method that both provides an intuitive interpretation and can be
-  applied to any number of lists even if these are incomplete. The
-  finding are illustrated using simulations and applied to a publicly
-  available cancer dataset.
+  The comparison of alternative rankings of a set of items is a
+  general and prominent task in applied statistics. Predictor
+  variables are ranked according to magnitude of association with an
+  outcome, prediction models rank subjects according to the
+  personalized risk of an event, and genetic studies rank genes
+  according to their difference in gene expression levels. This
+  article constructs measures of the agreement of two or more ordered
+  lists. We use the standard deviation of the ranks to define a
+  measure of agreement that both provides an intuitive interpretation
+  and can be applied to any number of lists even if some or all are
+  incomplete or censored. The main idea is then a permutation based
+  and graphically approach to assess the sequential changes of
+  agreement as a function of the depth of the lists. This may help to
+  indentify change points in the agreement of the lists. The
+  usefulness of these tools are illustrated using gene rankings, and
+  using data from two Danish ovarian cancer studies on where we assess
+  the within and between agreement of different statistical classification
+  methods.
 \end{abstract}
 
-Key words: sequential rank agreement, top lists, rank agreement, order statistic,
-methods comparison, variable selection
+Key words: sequential rank agreement, order statistic, gene rankings,
+methods comparison, variable selection, permutation
 
@@
 
-Ranking of items or results is common in scientific research and
-ranked lists occur as the result of many applications in
+A common task of scientific research is to rank items obtained by some
+form of data analysis. Ranked lists occur naturally in applications of
 statistics. Regression methods rank predictor variables according to
-magnitude of association with outcome, prediction models rank subjects
-according to their risk of an event, and genetic studies rank genes
-according to their difference in gene expression levels across
+magnitude of their association with an outcome, prediction models rank
+subjects according to their risk of an event, and genetic studies rank
+genes according to their difference in gene expression levels across
 samples.
 
-When several ranked lists are available a common research question is
-to what extent the lists agree on the ranking. In other words: is it
-possible to decide on a optimal rank until which the lists agree on
-the items?  A typical situation would arise in high-dimensional
-genomics studies where several analysis methods had been applied to a
-dataset or where a literature search provided differing tables of,
-say, top 10 lists of the most important genes associated to an
-outcome.  Similarly, this would correspond to the number of associated
-predictor variables that are found consistently over the lists, the
-number of patients at high risk of an event in all lists, and the
-number of genes that are worth to pursue in further experiments in the
-three examples above, respectively.
+When several rankings of the same items are available, a common
+research question is to what extent do they agree. Once agreement is
+established a further question is regarding the depths of agreement
+starting from the top of the ranked lists. A typical situation arises
+in high-dimensional genomics studies when several analysis methods are
+applied to rank a list of genes according to their association with a
+phenotype, treatment effect or other outcome. A measure of agreement
+of gene rankings obtained by different methods could often help to
+identify which genes are worth to pursue in further experiments.
 
-The analysis of rankings is hardly new, and several other approaches
-have proposed a measure of the ``distance'' between two ranked
-lists. Of these, Kendall's $\tau$
-%\citep{kendall1948}
-and Spearman’s footrule
-%$\rho$ \citep{Spearman1910}
-are the most well-known. Both measure the similarity of two ranked
-lists but neither of them make any distinction between errors towards
-the top versus the errors towards the bottom of the lists.  Several
-modified versions of Kendall's $\tau$ and Spearman's footrule have been
-proposed. \citet{Shieh1998} proposed a weighted $\tau$ where each pair
-of rankings can be assigned different weights, and \citet{Yilmaz2008}
+Several approaches exist to measure the ``distance'' between two
+ranked lists. Of these, Kendall's $\tau$ \citep{kendall1948} and
+Spearman’s footrule $\rho$ \citep{Spearman1910} are among the most
+well-known. But these measures do not distinguish between agreement in
+the top versus towards the bottom of the lists. \citet{Shieh1998}
+proposed a weighted version of Kendall's $\tau$ where each pair of
+rankings can be assigned different weights, and \citet{Yilmaz2008}
 proposed the $\tau_{ap}$ which places higher emphasis on the top of
@@
 % 
-Spearman’s footrule use the ranks of the variables for calculation of
-the distance and the use of ranks are also employed in the $M$
-measure of \citet{Bar-Ilan2006} where the reciprocal rank differences
-are used to calculate the similarity measure.
+Spearman’s footrule uses the ranks of the variables for calculation of
+the distance and the use of ranks are also employed in the $M$ measure
+of \citet{Bar-Ilan2006} where the reciprocal rank differences are used
+to calculate the agreement measure.
 
 More recent approaches consider the intersection of lists as the basis
-for a similarity measure. However, simple intersections also places
-equal weights on all depths of the list which led \citet{Fagin2003}
-and \citet{Webber2010} to modify the simple intersection to put more
-emphasis on the top of the lists by using the cumulative intersection
-divided by the depth as their similarity measures (denoted the average
-overlap (AO)). \citet{Webber2010} define their rank-biased overlap
-(RBO) by weighting with a converging series thus ensuring the top is
-always weighted higher than the (potentially indefinite) bottom of the
-lists.  While it is possible to calculate measures that describe how
-similar lists are at a given depth with some of the methods described
-above, the interpretation of the measures is unclear --- especially
-with multiple lists --- and it is difficult to identify a depth at
-which the lists are no longer similar.
+for a similarity measure. However, simple intersection also places
+equal weights on all depths of the list. Thus, \citet{Fagin2003} and
+\citet{Webber2010} proposed modified intersections which put more
+emphasis on the top of the lists. One proposal is denoted the average
+overlap (AO) and used for comparison later on. Specifically,
+\citet{Webber2010} define their rank-biased overlap (RBO) by weighting
+with a converging series to ensure that the top is weighted higher
+than the potentially non-informative bottom of the lists.
 
-In this article we propose a method to measure similarity among ranked
-lists for each depth in the list. As a measure for similarity we will
-use the limits of agreement known from agreement between quantitative
-variables \citep{Carstensen2010} but will compute the agreement on the
-ranks of the items.  The proposed method allows for multiple lists
-simultaneously, provides a dynamic measure of agreement for each depth
-in the lists, places more weight on the top of the list, accommodates
-censored/incomplete lists of varying lengths, and has a natural
-interpretation that directly relates to the ranks. The general idea is
-to define agreement based on the sequence of ranks from the first $k$
-elements in each list. This approach allows us to estimate a potential
-change point in the agreement in order to determine an optimal $k$
-where there is a shift in the agreement among lists and enables us to
-use randomization approaches to evaluate the rank agreement. In this
-sense it is a combination and generalization of some of the ideas of
-\citet{Carterette2009} and \citet{Boulesteix2009}. The former compares
-two rankings based on the distance between them as measured by a
-multivariate Gaussian distribution and the latter presents an overview
-of approaches for aggregation of ranked lists including bootstrap and
-leave-one-out jackknife approaches.
+It is possible to use the existing methods to calculate agreement of
+lists until a given depth, i.e., limited to the $d$ items of each
+list. However, the interpretation may not be straightforward,
+especially in the case of more than two lists.
 
-The manuscript is organized as follows: In the next section we will
-define the sequential rank agreement among multiple ranked lists and
-discuss how to handle incomplete/censored lists. In section 3 we
-present and discuss approaches to evaluate the sequential rank
-agreement obtained from an experiment and to identify the optimal
-depth $k$ for which the ranked lists agree. Finally we present three
-quite different applications of the proposed sequential rank agreement
-approach to real data before we discuss the findings along with
-possible extensions.
+In this article we introduce sequential rank agreement for measuring
+agreement among ranked lists. We provide graphical tools to analyse
+the depth of agreement of a set of ranked lists. The general idea is
+to define agreement based on the sequence of ranks from the first $d$
+elements in each list. As agreement metric we adapt the limits of
+agreement known from agreement between quantitative variables
+\citep{Carstensen2010}. Our proposed approach allows to compare more
+than two lists simultaneously, provides a dynamic measure of agreement
+as a function of the depth in the lists, places high weight on the top
+of the list, accommodates censored/incomplete lists of varying
+lengths, and has a natural interpretation that directly relates to the
+ranks. The graphical illustration of sequential rank agreement
+potentially allows to infer a changepoint, i.e., a list depth where
+there is a change in the agreement of the lists.
+
+In settings where rankings are obtained from analysis of correlated
+variables it is then possible to compare the actual results to the
+expected agreement in non-informative data obtained by
+randomization. In this sense our approach is a combination and
+generalization of some of the ideas of \citet{Carterette2009} and
+\citet{Boulesteix2009}. The former compares two rankings based on the
+distance between them as measured by a multivariate Gaussian
+distribution and the latter presents an overview of approaches for
+aggregation of ranked lists including bootstrap and leave-one-out
+jackknife approaches.
+
+The manuscript is organized as follows: In the next section we define
+sequential rank agreement for multiple ranked lists and discuss how to
+handle incomplete/censored lists. In section 3 we present and discuss
+approaches to evaluate the sequential rank agreement obtained from an
+experiment and to identify the maximal depth until which the ranked
+lists agree to certain extent. Finally we present three different
+applications of the proposed sequential rank agreement approach to
+real data before we discuss the findings along with possible
+extensions.
 
@@
 schematic example of these mappings. Thus if $\pi_l(1)=X_{34}$ then
-item $X_{34}$ is ranked first in list $l$, and similarly $R_l(X_{34})=1$.
+item $X_{34}$ is ranked first in list $l$, and similarly
+$R_l(X_{34})=1$.
 
@@
   rank agreement}. For an integer $1\le d\le P$ we define the unique
-set of items found in the $L$ top $d$ parts of the lists, i.e., the
-set of items ranked less than or equal to $d$ in any of the lists:
+set of items found by merging the first $d$ elements of each of the
+$L$ lists, i.e., the set of items ranked less than or equal to $d$ in
+any of the lists:
 \begin{equation}
@@
 \textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in
-      S_d\}}(L-1)A(X_p)^2}{(L-1)|S_d|}}, \label{def:sra}
+      S_d\}}(L-1)A(X_p)^2}{(L-1)|S_d|}}. \label{def:sra}
 \end{equation}
-and small values close to zero suggests that the lists agree on the
-ordering while larger values suggests disagreement. If the ranked
-lists are identical then the value of SRA will be zero for all depths
-$d$.  The sequential rank agreement can be interpreted as the average
-distance of the individual rankings of the lists from the average
-ranking for each of the items we have seen until depth $d$. Note that
-the terms that are part of \eqref{def:sra} are not independent since
-each list contains the ranks from 1 tp $P$ exactly once. Hence we
-simply use the pooled standard deviation as a measure of the average
-distance between rankings.
+Values of \textrm{SRA} close to zero suggests that the lists agree
+while larger values suggests disagreement. If the ranked lists are
+identical then the value of SRA will be zero for all depths $d$. The
+sequential rank agreement can be interpreted as the average distance
+of the individual rankings of the lists from the average ranking for
+each of the items we have seen until depth $d$. Note that the terms
+that are part of \eqref{def:sra} are not independent since each list
+contains the ranks from 1 to $P$ exactly once. Hence we simply use the
+pooled standard deviation as a measure of the average distance between
+rankings.
 
 \subsection{Agreement among fully observed lists}
-The simplest case occurs when all $L$ lists are fully
-observed, \ie, we have observed the rank of all $P$ items for all $L$
-lists. Fully observed ranked lists are common and arise, for example,
-when different statistical analysis methods are applied to a single
-dataset to produce lists of predictors ranked according to their
-importance, or if the same analysis method is applied to data from
-different populations.
+\label{sec:amfol}
+
+The simplest case occurs when all $L$ lists are fully observed, \ie,
+we have observed the rank of all $P$ items for all $L$ lists. Fully
+observed ranked lists are common and arise, for example, when
+different statistical analysis methods are applied to a single dataset
+to produce lists of predictors ranked according to their importance,
+or if the same analysis method is applied to data from different
+populations.
 
@@
 used to rank 3051 gene expression values measured on 38 tumor mRNA
-samples in order to improve the classification of acute leukemias
-between two types: acute lymphoblastic leukemia (ALL) or acute myeloid
-leukemia (AML) \citep{Golub1999}. Preprocessing of the gene expression
-data was done as described in \citet{Dudoit2002} and the four
-different analysis approaces were: marginal two-sample $t$ tests,
-marginal logistic regression analyses, logistic regression eleastic
-net, and marginal maximum information content correlations (MIC)
-\citep{Reshef2011}. For the first two methods, the genes were ranked
-according to $p$ value, for logistic regression the genes were ordered
-by size of the corresponding coefficients (after standardization), and
-MIC was ordered by correlation which resulted in the rankings seen in
-Table~\ref{tab1}.  The sequential rank agreement at depth $d$ can be
-interpreted as how many ranks apart do we on average expect the items
-found until depth $d$ to be so smaller values on the $y$ axis
-correspond to \emph{better} agreement among the lists than larger
-values.
+samples in order to classify between acute lymphoblastic leukemia
+(ALL) and acute myeloid leukemia (AML)
+\citep{Golub1999}. Preprocessing of the gene expression data was done
+as described in \citet{Dudoit2002} and the four different analysis
+approaces were: marginal two-sample $t$ tests, marginal logistic
+regression analyses, logistic regression eleastic net
+\citep{friedman2010regularization}, and marginal maximum information
+content correlations (MIC) \citep{Reshef2011}. For the first two
+methods, the genes were ranked according to $p$ value, for logistic
+regression the genes were ordered by size of the corresponding
+coefficients (after standardization), and MIC was ordered by
+correlation which resulted in the rankings seen in Table~\ref{tab1}.
+The sequential rank agreement shown in Figure~\ref{fig:example1} shows
+for each depth $d$ how many ranks apart do we on average expect the
+genes found in the top $d$ part of the lists. It can be seen that
+agreement is higher in the top of the lists (smaller values on the $y$
+axis) correspond to \emph{better} than in the tails of the lists.
 
-<<example1, echo=FALSE>>=
+<<example1, echo=FALSE, eval=FALSE>>=
 library(SuperRanker)
@@
 library(randomForestSRC)
-
 # Read data
@@
 
-<<fig1a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE>>=
+<<fig1a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE,eval=FALSE>>=
 ysize <- 1000
@@
 
-<<fig1b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE >>=
+<<fig1b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=FALSE>>=
 ## Same as above but zoomed in
@@
 
-<<fig2a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE>>=
+<<fig2a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=FALSE>>=
 ysize <- 1000
@@
 
-<<fig2b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE >>=
+<<fig2b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=FALSE>>=
 ## Same as above but zoomed in
@@
 
-<<label=tab1,echo=FALSE,results=tex>>=
+<<label=tab1,echo=FALSE,results=tex, eval=FALSE>>=
 library(xtable)
@@
  \includegraphics[width=.45\textwidth]{paper-fig2b}
-
  \end{center}
- \caption{Left panels: Sequential rank agreement for 4 different analysis methods
-   applied to the 3051 genes in the Golub data. Right panels:
-   Corresponding sequential rank agreement for the same data but
-   where only the top \Sexpr{topk} ranked items are available. The blue
-   and red areas correspond to the independent and randomized
-   reference hypothesis areas, respectively. The bottom plots are
-   identical to the corresponding top plots but have been zoomed in on
-   the first part of the $x$ axis.}
+ \caption{Left panels: Sequential rank agreement for 4 different
+   analysis methods applied to the 3051 genes in the Golub data. Right
+   panels: Corresponding sequential rank agreement for the same data
+   but where only the top 20 ranked items are available. The blue and
+   red areas correspond to the independent and randomized reference
+   hypothesis areas, respectively. The bottom plots are identical to
+   the corresponding top plots but have been zoomed in on the first
+   part of the $x$ axis.}
  \label{fig:example1}
@@
 
-If the lists generally agree on all items then we expect the
-sequential rank agreement curve to be flat, but if there are changes
-in the levels of rank agreement then this suggests that there are sets
-of items that the lists agree on and other sets where they rank the
-items differently. In particular, we generally expect the agreement
-among the lists to be better towards the top of the lists and worse
-towards the end of the lists, and in this case the sequential rank
-agreement curve will start at a low level and then increase until it
-levels off.
+If the lists would agree on all items then we would expect the
+sequential rank agreement curve to be flat. And if there are changes
+in the levels of rank agreement then this would suggest that there are
+sets of items for which the lists are in high agreement and sets of
+other items where the lists rank the items differently. In gene
+association studies we generally expect the agreement among the lists
+to be better towards the top of the lists and worse towards the end of
+the lists. In this case the sequential rank agreement curve will start
+at a low level and then increase until it levels off as in
+Figure~\ref{fig:example1}.
 
 \subsection{Analysis of incomplete/censored lists}
-Incomplete lists are also a common occurrence that arise from, for
-example, comparison of top $k$ lists, or when some methods only rank a
-subset of the items (e.g., penalized regression based on the Lasso provides a
-sparse set of predictors that have non-zero coefficients. There is no
-obvious ordering of the predictors that have all been shrunk to zero).
+
+Incomplete lists are also a common occurrence. They arises for example
+in case of missing data (items) or when some methods only rank a
+subset of the items. For example, penalized regression based on the
+Lasso provides a sparse set of predictors that have non-zero
+coefficients. There is no obvious ordering of the set of predictors
+whose coefficient has been shrunk to zero. Incomplete lists also occur
+when the analyst decides to censor all lists at depth $d$ such that
+the measurement of agreement is restricted to the first $d$ elements
+of each list.
 
 Sequential rank agreement can be generalized to incomplete/censored
-lists in the following way.  Let $\Lambda_l, l=1, \ldots, L$ be the
-set of items found in list $l$ so $\Lambda_l$ is the top $k_l$ list of
-items from list $l$ where $k_l = |\Lambda_l|$. Note that if
-$k_1=\cdots=k_L=k$ then we observe the top $k$ items for each of the
-$L$ lists. For censored lists the rank function becomes
+lists in the following way. Let $\Lambda_P$ be a set of $P$ items and
+$\Lambda_l\subset \Lambda_P$ the subset of $d_l$ items that occur in
+list $l$.  The case where all lists are censored at depth $d$
+corresponds to $d_1=\cdots=d_L=d$. For incomplete/censored lists the
+rank function becomes
 \begin{equation}
   \tilde R_l(X_p) = \left\{\begin{array}{cl} \{\pi_l^{-1}(p)\} & \text{ for } p\in \Lambda_l \\
-\{k_l+1,\dots,P\} & \text{ for } p \not\in \Lambda_l\end{array}\right.
+      \{d_l+1,\dots,P\} & \text{ for } p \not\in \Lambda_l\end{array}\right.
 \end{equation}
@@
 predictors in the presence of censored lists because the exact rank
-for some item is unknown in some of the lists.  The rankings within a
-single list are clearly not independent since each rank must appear
-exactly once in each list. Thus we cannot simply assign the same
-number (e.g., the mean of the unassigned ranks) to the censored items
-since that would result in less variation of the agreement, and it
-would artificially introduce a (downward) bias of agreement for items
-that are censored in multiple lists.
+for some items will be unknown. Also, the rankings within a single
+list are not independent since each rank must appear exactly once in
+each list. Thus, we cannot simply assign the same number (e.g., the
+mean of the unassigned ranks) to the censored items since that would
+% @Claus: less variation of the ranks? or of the agreement?
+result in less variation of the agreement, and it would artificially
+introduce a (downward) bias of agreement for items that are censored
+in multiple lists.
 
-Instead we assume that the rank assigned to an item $X_p$ that is not
-present in list $l$ (i.e., is censored) is uniformly distributed among
-the ranks that have \emph{not} been assigned for list $l$. By
-randomizing unobserved items we attain one realization of the rankings
-where all lists are fully observed. By randomizing a large number of
-times we can compute \eqref{def:sra} for each realization, and then
-compute the sequential rank agreement as the pointwise (for each
-depth) average of the rank agreements. The algorithm is described in
-detail in Algorithm~\ref{sra-algorithm}.
+Instead we randomize the ranks $\{d_{l}+1,\dots,P\}$ to the items that
+do not occur in list $\Lambda_l$. By randomizing the missing items of
+each list once we attain one realization of the $L$ rankings of the
+set $\Lambda_P$. By randomizing a large number of times we can compute
+\eqref{def:sra} for each realization, and then compute the sequential
+rank agreement as the pointwise (for each depth) average of the rank
+agreements. The algorithm is described in detail in
+Algorithm~\ref{sra-algorithm}.
 
@@
 \begin{algorithmic}[1]
-\Procedure{Censored rank agreement}{}
-\State Let $B$ be the number of permutations to use
-\For{each $b \in B$}
-\For{each censored list $l \in L$}
-\State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1.7cm}{Permute the unassigned ranks, $k_l+1, \ldots, P$ and assign
-them randomly to the items \emph{not} found in the list, i.e.,
-$\Lambda^\complement$, in order to fill out the lists.}
+  \Procedure{Censored rank agreement}{} \State Let $B$ be the number
+  of permutations to use \For{each $b \in B$} \For{each censored list
+    $l \in L$}
+  \State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1.7cm}{Permute
+    the unassigned ranks, $\{d_l+1, \ldots, P\}$, and assign them
+    randomly to the items \emph{not} found in the list, i.e.,
+    $\Lambda^\complement=\Lambda_P\setminus\Lambda_l$. Combine the
+    result with $\Lambda_l$ to fill out the list.}
 \EndFor
@@
 
-The proposed approach is based on two assumptions: 1) that the items
-of interest will be found towards the top of the lists, and 2) that
-the censored rankings provide so little information that it is
-reasonable to assume that the missing rankings provide no information
-and might as well be considered random.  The first assumption is
-justifiable because we have already accepted that it is reasonable to
-rank the items in the first place. The second assumption is fair
-in the light of the first assumption provided that we have a
+The proposed approach is based on two assumptions: 1) that the most
+interesting items are found in the top of the lists, and 2) that the
+censored rankings provide so little information that it is reasonable
+to assume that they can be represented by a random order. The first
+assumption is justifiable because we have already accepted that it is
+reasonable to rank the items in the first place. The second assumption
+is fair in the light of the first assumption provided that we have a
 ``sufficiently large'' part of the top of the lists available.
 
-When the two assumptions are combined then it is clear that the
-interesting part of the sequential rank agreement curves must be for
-smaller depths. Unless we have additional prior knowledge about the
-distribution of the unranked items then we can restrict attention of
-the sequential rank agreement to
-\begin{equation}
-d' \leq \max(k_1, \ldots, k_L),
-\end{equation}
-since we have no real information about the ordering of the items after a
-depth $d'$ where all lists are censored.
+When the two assumptions are satisfied then it is clear that the
+interesting part of the sequential rank agreement curves is restricted
+to the depths where the percentage of censored items is
+low. Generally, without additional prior knowledge about the
+distribution of the censored items it seems reasonable to restrict the
+attention of the sequential rank agreement to depths smaller than
+$\max(d_1, \ldots, d_L)$.
 
@@
 
-\section{Evaluating sequential rank agreement}
+\section{No information agreement}
 
-To interpret the sequential rank agreement values we propose two
+We now consider the special case where the rankings are obtained by
+applying different methods to the same data set which contains one
+outcome variable and a list of candidate predictors. To interpret the
+sequential rank agreement values in this situation we propose two
 different benchmark values corresponding to two different
@@
 H_0  & : &  \text{The list rankings correspond to complete randomly}\\
-       &  & \text{permuted lists}
+       &  & \text{permuted lists}.
 \end{eqnarray*}
-which not only assumes that there is no information in the rankings
-but also that the methods are completely independent.
+This does not only assume that there is no information data on which
+in the rankings are based but also that the methods are completely
+independent.
 
@@
 method/approaches to the same data
+% @Claus: maybe H_1 is not a good name as people would expect H_1 to be the complement of H_0
 \begin{eqnarray*}
-H_1 & :&  \text{The list rankings are based on data containing}\\
+\tilde H_0 & :&  \text{The list rankings are based on data containing}\\
 & &   \text{no association to the outcome.}
 \end{eqnarray*}
-$H_0$ is clearly the strictest (and also potentially quite
-unrealistic) null hypothesis but we can easily obtain realizations
-from that null hypothesis simply by permuting the items for each list
-and then computing the sequential rank agreement for the permuted
-lists. In the uncensored case each experiment contains $L$ lists of
-random permutations of the numbers $1,\dots,P$. For the censored case
-we permute the number as before but censor the same number of items
-for each list as were censored in the original dataset. The sequential
-rank agreement curve from the the original data can then be compared
-to the pointwise rank agreements obtained from under $H_0$.
+$H_0$ is a quite unrealistic null hypothesis but we can easily obtain
+realizations from that null hypothesis simply by permuting the items
+within each list and then computing the sequential rank agreement for
+the permuted lists. In the uncensored case each experiment contains
+$L$ lists of random permutations of the numbers $1,\dots,P$. For the
+censored/incomplete case we first permute the numbers $1,\dots,P$ and
+then censor list $l$ at $d_l$. The sequential rank agreement curve
+from the the original lists can then be compared to the pointwise rank
+agreements obtained under $H_0$.
 
-To obtain a reference distribution under $H_1$ we use a randomization
-test approach on the original data to remove any associations to the
-outcome in the data that was used to generate the rankings before
-computing the ranking and subsequently the sequential rank agreement
-from the randomized data. The idea is to repeat the ranking procedures
-in the modified data where the outcome vector is randomly permuted. In
-practice we permute the outcomes but keep any predictors before
-ranking to preserve any structure in the predictors. This artificially
-removes any potential information regarding the regression that the
-real data may contain. The randomization approach requires that we
-have the original data available, so we can compute the ranked lists
-from the randomized data and not just the ranked lists and as such it
-may not be possible to evaluate $H_1$ in all situations.
+To obtain the distribution under $\tilde H_0$ the idea is to repeat
+many times the ranking procedures in uncorrelated data.  Thus, we
+first permute the outcome variable of the data set on which the
+rankings are based. This removes any association between the predictor
+variables and the outcome. Then we apply the methods to the permuted
+data set to generate $L$ new rankings and subsequently compute the
+sequential rank agreement. Note that we only permute the outcomes and
+preserve the structure of the candidate predictors. The randomization
+approach requires that we have the original data available and as such
+it may not be possible to evaluate $\tilde H_0$ in all situations.
 
@@
 substantially below the distribution of the sequential rank agreements
-obtained under either $H_0$ or $H_1$ then this suggests that the
-original ranked lists agree \emph{more} that we would expect if the
-ranked lists were based on data with no information.
+obtained under either $H_0$ or $\tilde H_0$ then this suggests that
+the original ranked lists agree \emph{more} than what we would expect
+in data with no information.
 
-Figure~\ref{fig:example1} shows the sequential rank distributions
-based on $400$ samples under $H_0$ and $H_1$, respectively. Not
-surprisingly, the null distribution of $H_1$ has lower sequential rank
-agreement than the null distribution of $H_0$ because the four methods
-used to rank the data ($t$ test, logistic regression, elastic net, and
-MIC) generally identify and rank the same set of predictors that show
-spurious associations. The two bottom panels in
+For the analysis of the Golub data described in Section \ref{sec:amfol},
+Figure~\ref{fig:example1} shows the empirical distributions of
+sequential rank agreement under $H_0$ and $\tilde H_0$ each based on
+$400$ permutations. Not surprisingly, the sequential rank agreement
+under $\tilde H_0$ is lower than the sequential rank agreement under
+$H_0$ because the four methods used to rank the data ($t$ test,
+logistic regression, elastic net, and MIC) generally tend to identify
+(and rank high) similar predictors even if there are only spurious
+associations. The two bottom panels in
+% @Claus jeg forstaar ikke den den naste lange saetning. mener du full lists i stedet for full dataset?
 Figure~\ref{fig:example1} also indicate that the observed sequential
@@
 
-It is important to stress that neither $H_0$ nor $H_1$ are concerned
-with whether the $L$ rankings are relevant. Both hypotheses are purely
-considering how the rankings agree in a situation where there is no relevant
+It is important to stress that neither $H_0$ nor $\tilde H_0$ are
+related to questions regarding the association between the outcome and
+the predictors in the data set. Both hypotheses are purely considering
+how the rankings agree in a situation where there is no relevant
 information available in the data used for creating the rankings.
@@
 
-The average overlap is widely used in comparing (two) ranked list and
-the idea behind the average overlap closely resembles the sequential
-rank agreement \citep{Fagin2003,Webber2010}. Here we will present an
-extention of the average overlap to more that two lists and compare this
-to sequential rank agreement.
+The average overlap is widely used in comparisons of (two) ranked
+lists \citep{Fagin2003,Webber2010} and the idea behind the average
+overlap closely resembles the sequential rank agreement. In this
+section we present an extention of the average overlap to more that
+two lists and compare this to sequential rank agreement.
 
@@
 \end{equation}
-Typically, the overlap is only considered for pairwise comparisons but in
-\eqref{def:overlap} we make no such distinction.
-
+Here $|A|$ is the cardinality of the set $A$.
+Originally the overlap was only considered for pairwise comparisons
+but formula \eqref{def:overlap} can be applied when $L>2$.
+% @Claus: har du en reference hvor OA bliver brugt paa to lister?
 The average overlap at depth $d$ is defined as the average of the
@@
 \end{equation}
-This definition directly ensures that extra emphasis is included on
-the top of the list since the overlap of the items towards the top are
-included in the majority of the calculations.
+This definition directly ensures that further emphasis is put on items
+in the top of the lists since their item specific overlap contributes
+to the calculation of average overlap at higher depths.
 
 There are three main drawbacks of the average overlap method: First it
-is highly sensitive to items at the top of the list. The average
-overlap profile changes dramatically depending on the occurrence of
-similar items in the lists. Secondly, when the number of lists, $L$,
-is high then it might be difficult to obtain a non-zero overlap (and
-hence a non-zero average overlap) at the beginning of the lists because
-the overlap does not increase until an item is present in all $L$
-lists. Finally, the average overlap has an interpretation in terms of
-moving averages of  percentages (of $L$ sets) which somewhat less
-intuitive than simply the average rank distance that we propose.
+is highly sensitive to single items in the top of one the lists.
+% @Claus: jeg forstaar ikke den naeste saetning.
+The average overlap profile may change dramatically depending on the
+occurrence of similar items in the lists. Secondly, when the number of
+lists, $L$, is high then it might be difficult to obtain a non-zero
+overlap (and hence a non-zero average overlap) at the beginning of the
+lists because the overlap will be zero until an item is present in all
+$L$ lists. Finally, the average overlap has an interpretation in terms
+of moving averages of percentages (of $L$ sets) which is somewhat less
+intuitive than our proposed average rank distance.
 
-In this section we compare the average overlap and the sequential rank
-agreement to the Golub data  \citep{Golub1999}. To make it more
-similar to the way the average ovelap is typically used we restrict
-attention to only two analysis methods: $t$ tests and logistic
-regression. Thus we have two marginal model approaches that are very
-similar and the first two columns of Table~\ref{tab1} show the actual top-10
-ranking of the predictors.
+In the following we compare the average overlap and the sequential
+rank agreement using the analysis of the Golub data \citep{Golub1999}
+of Section \ref{sec:amfol}. We consider a typical application of the
+average overlap and thus restrict attention to only two gene rankings
+obtained by the $t$ tests and logistic regression. These two marginal
+modelling approaches should result in very similar gene rankings. The
+first two columns of Table~\ref{tab1} show the actual top-10 ranking
+of the predictors.
 
-<<fig-ao1, echo=FALSE, fig=TRUE,include=FALSE>>=
+<<fig-ao1, echo=FALSE, fig=TRUE,include=FALSE, eval=FALSE>>=
 myres <- average_overlap(inputmatrix[,1:2])
@@
 
-<<fig-ao2, echo=FALSE, fig=TRUE,include=FALSE>>=
+<<fig-ao2, echo=FALSE, fig=TRUE,include=FALSE, eval=FALSE>>=
 mmyres <- sra(inputmatrix[,1:2])
@@
 
-\begin{figure}[tb]
-\begin{center}
-\includegraphics[width=.45\textwidth]{paper-fig-ao1}
-\includegraphics[width=.45\textwidth]{paper-fig-ao2}
-\end{center}
- \caption{Average overlap (left figure) and sequential rank agreement
-   (right figure) for comparing two different analysis methods
-   (marginal $t$ test and logistic regression)
-   applied to the Golub data. The black lines are based on the full
-   set of predictors while the red lines have the two highest
-   associated predictors removed
-   from the data before computing the average overlap and sequential
-   rank agreement.}
- \label{fig:case1}
-\end{figure}
+%% \begin{figure}[tb]
+%% \begin{center}
+%% \includegraphics[width=.45\textwidth]{paper-fig-ao1}
+%% \includegraphics[width=.45\textwidth]{paper-fig-ao2}
+%% \end{center}
+ %% \caption{Average overlap (left figure) and sequential rank agreement
+   %% (right figure) for comparing two different analysis methods
+   %% (marginal $t$ test and logistic regression)
+   %% applied to the Golub data. The black lines are based on the full
+   %% set of predictors while the red lines have the two highest
+   %% associated predictors removed
+   %% from the data before computing the average overlap and sequential
+   %% rank agreement.}
+ %% \label{fig:case1}
+%% \end{figure}
 
@@
 the sequential rank agreement we get roughly the same estimate of
-agreement from rank 3 as we did from the full dataset which is a nice
-property. For the average overlap the curves completely change when
-these two items are missing and give a visually different conclusion
-about the ranking of the items from rank 3 downwards.
+agreement from rank 3 as we did from the full dataset. This indicates
+that sequential rank is more robust against small pertubations of the
+data. For the average overlap the curves completely change when these
+two items are removed and which would lead to a different conclusion
+about the agreement among the items from rank 3 downwards.
 
@@
 departments of gynecology were enrolled continuously from December
-1994 to May 1999. We use a random subsample of $119$ patients with a
-total of $58$ patients with malignant ovarian cancers as cases and
-$61$ patients with benign ovarian tumors as controls. The DACOVA study
-is similarly a multidisciplinary Danish study on ovarian cancer
-covering about $66\%$ of the population in Denmark
-\cite{bertelsen1991protocol}. It was attempted to continuous enroll
-all patients referred for surgery with an ovarian tumor clinically
-suspected to be cancer during the period 1984 to 1990. We use a random
-subsample of $113$ patients with a total of $54$ malignant ovarian
-cancers and $59$ benign ovarian tumors/gynecologic disorders.
+1994 to May 1999.
+% @Andreas: hvorfor random subsample?
+We use a random subsample of $119$ patients with a total of $58$
+patients with malignant ovarian cancers as cases and $61$ patients
+with benign ovarian tumors as controls. The DACOVA study is another
+multidisciplinary Danish study on ovarian cancer which included about
+$66\%$ of the female population of Denmark
+\cite{bertelsen1991protocol}. It was aimed to continuously enroll all
+patients that were referred to surgery of an ovarian tumor clinically
+suspected to be cancer during the period 1984 to 1990.
+% @Andreas: hvorfor random subsample?
+For the purpose of illustration use a random subsample of $113$
+patients with a total of $54$ malignant ovarian cancers and $59$
+benign ovarian tumors/gynecologic disorders.
 
@@
 We use these two data sets to illustrate how the sequential rank
-agreement can be applied in two different scenarios: i) assessing the
-agreement of ranking the importance of the predictors using different
-statistical classification methods and ii) assessing the agreement of
-rankings the predictions (the probability of the classifying a tumor
-as malignant) subject to perturbations in data. In the latter case we
-further illustrate the discrepancy between the ranking stability of
-different statistical methods through a simulation study.
+agreement can be applied in two different scenarios. In the first
+scenario we assess the agreement of different statistical
+classification methods in how they rank predictors according to
+importance for classification accuracy. In the second scenario we
+assess the agreement among rankings of the same patients according to
+the predicted probabilities of having a malignant tumor. Here we apply
+the same statistical classification method to XX bootstrap samples of
+the MALOVA data to obtain XX prediction models. From each model we
+predict the risks of a malignant patients in the DACOVA study.
+Finally we estimate the agreement of the resulting XX rankings of the
+DACOVA patients to illustrate the sensitivity of the corresponding
+statistical classification method to pertubations of the learning
+data. In the latter case we further investigate the behaviour of
+different statistical methods in a simulation study.
 
@@
 
-#includegraphics marginalVsModel.pdf
+%% #includegraphics marginalVsModel.pdf
 
@@
 
-#includegraphics thomas.pdf
+%% #includegraphics thomas.pdf
 
-#includegraphics simulationTest2.pdf
+%% #includegraphics simulationTest2.pdf
 

### 2015-07-13 R/sra.R
Tweaks

@@ -207,7 +207,7 @@ random_list_sra <- function(object, B=1, n=1, na.strings=NULL) {
 
 #' Smooth quantiles of a matrix of sequential ranked agreements
 #'
-#' @param obj A matrix
+#' @param object A matrix
 #' @param confidence the limits to compute
 #' @return A list containing two vectors for the smoothed lower and upper limits
 #' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
@@ -217,4 +217,34 @@ smooth_sra <- function(object, confidence=0.95) {
     alpha <- (1-confidence)/2
     limits <- apply(object, 1, function(x) { quantile(x, probs=c(alpha, 1-alpha)) })
     list(lower=limits[1,], upper=limits[2,])
+}
+
+
+
+#' Compute a Kolmogorov-Smirnoff-like test for Smooth quantiles of a matrix of sequential ranked agreements
+#'
+#' @param object The results from an sra
+#' @param nullobject the limits to compute
+#' @param weights Either a single value or a vector of the same length as the number of item with the weight that should be given to specific depths
+#' @return A single value corresponding to the p-value
+#' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
+#' @export
+test_sra <- function(object, nullobject, weights=1) {
+    ## Sanity checks
+    if (! (length(weights) %in% c(1, length(object))))
+        stop("the vector of weights must have the same length as the number of items")
+    
+    ## Test statistic
+    T <- max(weights*abs(object - apply(nullobject, 1, mean)))
+
+    ## Now compute the individual jackknife variations from the null object
+    B <- ncol(nullobject)
+    nullres <- sapply(1:B, function(i) {
+        max(weights*abs(nullobject[,i] - apply(nullobject[,-i], 1, mean)))
+    })
+
+    res <- sum(nullres>=T)/(B+1)
+    attr(res, "B") <- B
+    return(res)
+    
 }

### 2015-07-13 latex/paper.Rnw
Tweaks
Answers/rewrites to address Thomas' questions

@@
 mean of the unassigned ranks) to the censored items since that would
-% @Claus: less variation of the ranks? or of the agreement?
-result in less variation of the agreement, and it would artificially
-introduce a (downward) bias of agreement for items that are censored
-in multiple lists.
+result in less variation of the ranks and hence less variation of the
+agreement, and it would artificially introduce a (downward) bias of
+agreement for items that are censored in multiple lists.
 
@@
 method/approaches to the same data
-% @Claus: maybe H_1 is not a good name as people would expect H_1 to be the complement of H_0
 \begin{eqnarray*}
@@
 
-For the analysis of the Golub data described in Section \ref{sec:amfol},
-Figure~\ref{fig:example1} shows the empirical distributions of
-sequential rank agreement under $H_0$ and $\tilde H_0$ each based on
-$400$ permutations. Not surprisingly, the sequential rank agreement
-under $\tilde H_0$ is lower than the sequential rank agreement under
-$H_0$ because the four methods used to rank the data ($t$ test,
-logistic regression, elastic net, and MIC) generally tend to identify
-(and rank high) similar predictors even if there are only spurious
-associations. The two bottom panels in
-% @Claus jeg forstaar ikke den den naste lange saetning. mener du full lists i stedet for full dataset?
+For the analysis of the Golub data described in Section
+\ref{sec:amfol}, Figure~\ref{fig:example1} shows the empirical
+distributions of sequential rank agreement under $H_0$ and
+$\tilde H_0$ each based on $400$ permutations. Not surprisingly, the
+sequential rank agreement under $\tilde H_0$ is lower than the
+sequential rank agreement under $H_0$ because the four methods used to
+rank the data ($t$ test, logistic regression, elastic net, and MIC)
+generally tend to identify (and rank high) similar predictors even if
+there are only spurious associations. The two bottom panels in
 Figure~\ref{fig:example1} also indicate that the observed sequential
-rank agreement is better than what would be expected for the full
-dataset but that the censored data (the bottom-right plot) suggests
-that there may be at most 1 or 2 ranked items that are towards the top
-of the lists that yield a result better than what would be expected.
+rank agreement (the black line) is better than what would be expected
+by chance for data that contain no information since it lies below the
+reference areas. However, the censored data also suggests that there
+may be at most 1 or 2 ranked items towards the top of the lists that
+yield a result better than what would be expected (the bottom-right
+plot).
 
@@
 \end{equation}
-Here $|A|$ is the cardinality of the set $A$.
-Originally the overlap was only considered for pairwise comparisons
-but formula \eqref{def:overlap} can be applied when $L>2$.
-% @Claus: har du en reference hvor OA bliver brugt paa to lister?
+Here $|A|$ is the cardinality of the set $A$.  Typically the overlap
+is only considered for pairwise comparisons but formula
+\eqref{def:overlap} can be applied even when $L>2$
+\citep{Bar-Ilan2006,Boulesteix2009}.
+
 The average overlap at depth $d$ is defined as the average of the
@@
 There are three main drawbacks of the average overlap method: First it
-is highly sensitive to single items in the top of one the lists.
-% @Claus: jeg forstaar ikke den naeste saetning.
-The average overlap profile may change dramatically depending on the
-occurrence of similar items in the lists. Secondly, when the number of
-lists, $L$, is high then it might be difficult to obtain a non-zero
-overlap (and hence a non-zero average overlap) at the beginning of the
-lists because the overlap will be zero until an item is present in all
-$L$ lists. Finally, the average overlap has an interpretation in terms
-of moving averages of percentages (of $L$ sets) which is somewhat less
-intuitive than our proposed average rank distance.
+is highly sensitive to single items in the top of one the lists.  The
+average overlap profile may change dramatically depending on the
+presence or absence of similar items in the beginning of the
+lists. For example, if all lists have the same item at the top then
+the average overlap attains its maximum at 1, but if just one of the
+lists differ then the overage overlap starts at its minimum of
+0. Secondly, when the number of lists, $L$, is high then it might be
+difficult to obtain a non-zero overlap (and hence a non-zero average
+overlap) at the beginning of the lists because the overlap will be
+zero until an item is present in all $L$ lists. Finally, the average
+overlap has an interpretation in terms of moving averages of
+percentages (of $L$ sets) which is somewhat less intuitive than our
+proposed average rank distance.
 
@@
 
-%% #includegraphics marginalVsModel.pdf
+\begin{figure}[tb]
+\begin{center}
+\includegraphics[width=.5\textwidth]{pics/marginalVsModel.pdf}
+\end{center}
+ \caption{Marginal}
+ \label{fig:app1}
+\end{figure}
 
@@
 
-%% #includegraphics thomas.pdf
+\begin{figure}[tb]
+\begin{center}
+\includegraphics[width=.5\textwidth]{pics/thomas.pdf}
+\end{center}
+ \caption{Thomas}
+ \label{fig:app2}
+\end{figure}
 
-%% #includegraphics simulationTest2.pdf
+\begin{figure}[tb]
+\begin{center}
+\includegraphics[width=.5\textwidth]{pics/simulationTest2.pdf}
+\end{center}
+ \caption{SimulationTest2}
+ \label{fig:app3}
+\end{figure}
 

### 2015-07-14 latex/paper.Rnw
Discussion

@@
 real data before we discuss the findings along with possible
-extensions.
+extensions. The approaches presented in this manuscript are available
+in the R package \texttt{SuperRanker}.
 
@@
 
-<<example1, echo=FALSE, eval=FALSE>>=
+<<example1, echo=FALSE, eval=TRUE>>=
 library(SuperRanker)
@@
 
-<<fig1a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE,eval=FALSE>>=
+<<fig1a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE,eval=TRUE>>=
 ysize <- 1000
@@
 
-<<fig1b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=FALSE>>=
+<<fig1b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=TRUE>>=
 ## Same as above but zoomed in
@@
 
-<<label=tab1,echo=FALSE,results=tex, eval=FALSE>>=
+<<label=tab1,echo=FALSE,results=tex, eval=TRUE>>=
 library(xtable)
@@
 colnames(xxx) <- c("Ranking", colnames(xxx)[2:5])
-xxx2 <- xtable(xxx, caption = "Top 10 list of ranked results from the Golub data. Numbers indicate the predictor/gene for the given ranking and method")
+xxx2 <- xtable(xxx, caption = "Top 10 list of ranked results from the Golub data. Numbers indicate the predictor/gene for the given ranking and method", label="tab1")
 digits(xxx2) <- 0
@@
 
-%% \begin{figure}[tb]
-%% \begin{center}
-%% \includegraphics[width=.45\textwidth]{paper-fig-ao1}
-%% \includegraphics[width=.45\textwidth]{paper-fig-ao2}
-%% \end{center}
- %% \caption{Average overlap (left figure) and sequential rank agreement
-   %% (right figure) for comparing two different analysis methods
-   %% (marginal $t$ test and logistic regression)
-   %% applied to the Golub data. The black lines are based on the full
-   %% set of predictors while the red lines have the two highest
-   %% associated predictors removed
-   %% from the data before computing the average overlap and sequential
-   %% rank agreement.}
- %% \label{fig:case1}
-%% \end{figure}
+\begin{figure}[tb]
+\begin{center}
+\includegraphics[width=.45\textwidth]{paper-fig-ao1}
+\includegraphics[width=.45\textwidth]{paper-fig-ao2}
+\end{center}
+ \caption{Average overlap (left figure) and sequential rank agreement
+ (right figure) for comparing two different analysis methods
+ (marginal $t$ test and logistic regression)
+    applied to the Golub data. The black lines are based on the full
+    set of predictors while the red lines have the two highest
+    associated predictors removed
+    from the data before computing the average overlap and sequential
+    rank agreement.}
+  \label{fig:case1}
+ \end{figure}
 
-The black lines Figure~\ref{fig:case1} show the average overlap and
+The black lines in Figure~\ref{fig:case1} show the average overlap and
 sequential rank agreement for these data. It is clear from both plots
@@
 \section{Discussion}
+In this article we address the problem of comparing ranked lists of
+the same items. Our proposed method can handle both the situation
+where the underlying data to generate the ranked lists are available
+and the situation where the only available data is the actual ranked
+lists. In addition, censored ranked lists where only the top $k$
+ranked items are known can be accommodated as well. The proposed
+agreement measure can be interpreted as the average distance between
+an item's rank and the average rank assigned to that item and thus
+determine how well the lists agree on the ranks of a specific
+item and it circumvents the problem with other existing similarity measures.
 
-Mention/discuss different measures.
+The sequential rank agreement is extremely versatile. We have shown
+that it can be used not only to compare ranked lists of items produced
+from different samples/populations but that it also can be used to
+study the ranks obtained from different analysis methods on the same
+data as well as to evalaute the stability of the ranks from a single
+method by bootstrapping the data repeatedly and comparing the ranks
+obtained from the bootstrapped data.
 
-Censor lists at significane level?
+Our applications show that XXX
+Through simulations we showed that XXX --- afventer Andreas.
 
-Bootstrap across a single method and compare results. Discuss collinearity
+The sequential rank agreement can be used to determine the depth at
+which the rank agreement becomes too large to be desirable based on
+prior requirements or acceptable differences, or it can be used to
+visually determine when the change in agreement becomes too large.  In
+that regard the investigator can have prior limits on the level of
+agreement that is acceptable.
 
-Changepoint analysis
+While the sequential rank agreement is primarily an exploratory tool
+we have suggested two null hypotheses that can be used to evaluate the
+sequential rank agreement obtained. Note that none of the two null
+hypotheses are concerned with the actual ``true ranking'' but are
+purely concerned with consistency/stability of the rankings among the
+lists. As such we cannot determine if the rankings are good but only
+whether they agree. The sequential rank agreemenet curve can be
+compared visually to the curves obtained under either of the null
+distributions and simple point-wide $p$-values can be obtained for
+each depth by counting the number of sequential rank agreements under
+the null hypothesis that is less than or equal to the observed rank
+agreement. Two simple extensions can be pursued to make more formal
+uniform tests: one would be to use a Kolmogorov-Smirnov-like test and
+use the largest difference between the variance-weighted sequential
+rank agreement curve and the mean null rank agreement curve as a test
+statistic. Alternatively, a change point analysis could be made on the
+sequential rank agreement in order to determine the depths at which
+there are ``jumps'' in the rank agreement. These jumps would
+correspond to depths for which the agreement among the lists was
+substantially worse and could serve as indicators for when the lists
+no longer agree sufficiently satisfactory.
 
+Finally, we have --- whenever possible --- used as much information
+from the lists as possible, i.e., we have used all available ranks. If
+we have information on the rankings as well as the corresponding
+evidence for significance that we could choose to censor the rankings
+of all the items that were deemed not to be significant. That would
+ensure that there would be put less emphasis on the agreement of the
+non-significant items and it would be easier to identify a change in
+agreement among the items that were deemed to relevant.
+
+We note that the sequential rank agreement is still marred by problems
+that generally apply to ranking of items and/or
+indiviuals. Collinearity in particular can be a huge problem when
+bootstrapping data or when comparing different analysis methods. For
+example, marginal analyses where each item is analyzed seperately will
+assign similar ranks to two highly correlated predictors while methods
+that provide a sparse solution such as the lasso will just rank one of
+the two predictors high while the other might have a very low
+rank. When the rankings from these two analysis methods are compared
+then the agreement will differ simply because of the way correlated
+predictors are handled. This is not a shortcoming of the sequential
+rank agreement but is a problem general to all ranked lists.
+
+In conclusion we have introduced a method for evaluation of ranked
+(censored) lists that can be easily interpreted and that can be
+applied to a large number of situations.  The method can be adapted
+even further by using it to identify and cluster statistical analysis
+methods that provide more or less the same agreement, or by using the
+rank agreement as a hyper-parameter in, say, elastic net regularized
+regression where the rank agreement is used to determine the mixing
+proportion between the $L_1$ and the $L_2$ penalty. Finally, the
+proposed method may be adapted (with some additional assumptions) to
+the situation where the lists are subsets of each other. We will be
+investigating these extensions further in the future.
+
+%\bibliographystyle{rss}
 \bibliographystyle{rss}

### 2015-07-23 latex/paper.Rnw
Updated manuscript

@@
 \usepackage{algpseudocode}
+%\usepackage{geometry}
+%\geometry{a4paper, portrait, margin=1in}
 %
@@
   agreement as a function of the depth of the lists. This may help to
-  indentify change points in the agreement of the lists. The
+  indentify change-points in the agreement of the lists. The
   usefulness of these tools are illustrated using gene rankings, and
@@
 
-Key words: sequential rank agreement, order statistic, gene rankings,
+Key words: sequential rank agreement, partial ranking, order statistic, gene rankings,
 methods comparison, variable selection, permutation
@@
 Spearman’s footrule $\rho$ \citep{Spearman1910} are among the most
-well-known. But these measures do not distinguish between agreement in
+well-known, but these measures do not distinguish between agreement in
 the top versus towards the bottom of the lists. \citet{Shieh1998}
@@
 ranks. The graphical illustration of sequential rank agreement
-potentially allows to infer a changepoint, i.e., a list depth where
+potentially allows to infer a change-point, i.e., a list depth where
 there is a change in the agreement of the lists.
@@
 
-Consider a set of $P$ different items $X=\{X_1,\dots,X_P\}$. An
-ordered list is a permutation function, $R: \{X_1,\dots,X_P\}\to
-\{1,\dots,P\}$, such that $R(X_p)$ is the rank of item $X_p$ in the
-list. The inverse mapping $\pi=R^{-1}$ assigns to rank
-$r\in\{1,\dots,P\}$ the item $\pi(r)$ found at that rank. The methods
-described below work for a set of $L$ lists $R_1,\dots,R_L$,
-$L\geq2$. We denote $\pi_l=R_l^{-1}$ for the corresponding inverse
-mappings. Panels (a) and (b) of Table~\ref{tab:example} show a
-schematic example of these mappings. Thus if $\pi_l(1)=X_{34}$ then
-item $X_{34}$ is ranked first in list $l$, and similarly
-$R_l(X_{34})=1$.
+Consider a set of $P$ different items $X=\{X_1,\dots,X_P\}$ and a
+ranking function $R: \{X_1,\dots,X_P\}\to \{1,\dots,P\}$, such that
+$R(X_p)$ is the rank of item $X_p$. The inverse mapping $R^{-1}$
+assigns to rank $r\in\{1,\dots,P\}$ the item $R^{-1}(r)$ that was
+given rank $r$. An ordered list is the realization of a ranking
+function $R$ applied to a data sample $X$ and we let
+$R_1(X),\dots,R_L(X)$, $L\geq2$, denote ordered lists obtained from
+$L$ different ranking functions applied to the same data $X$. We will
+use the same notation for the equivalent scenario where we have
+ordered lists obtained from applying the same ranking function to $L$
+samples from the same population, i.e., for ranked lists
+$R(X^1), \ldots, R(X^L)$.  Panels (a) and (b) of
+Table~\ref{tab:example} show a schematic example of these
+mappings. Thus if $R^{-1}(1)=X_{34}$ then item $X_{34}$ is ranked
+first in list $l$, and similarly $R_l(X_{34})=1$.
 
 \begin{table}[tb]
-  \caption{Example set of ranked lists.
-(a) shows the ranked list of items for each of three lists, (b) presents the ranks obtained by each item in each of the three lists and (c) shows the cumulative set of items up to a given depth in the three lists.}
+  \caption{Example set of ranked lists. (a) shows the ranked list of
+    items for each of three lists, (b) presents the ranks obtained by
+    each item in each of the three lists and (c) shows the cumulative
+    set of items up to a given depth in the three lists.}
 \begin{center}
-  \begin{subtable}{4cm}%
+\footnotesize{
+  \begin{subtable}{4.5cm}%
     \caption{}
@@
         \hline\hline % & \multicolumn{3}{c}{List} \\
-        Rank & $\pi_1$ & $\pi_2$ & $\pi_3$ \\ \hline
+        Rank & $R^{-1}_1$ & $R^{-1}_2$ & $R^{-1}_3$ \\ \hline
         1 & A & A & B \\
@@
 \end{subtable}
+}
 \end{center}
@@
 \end{equation}
-for a distance function $f$. Throughout this paper we will use the
+for a distance function $f.$ Throughout this paper we will use the
 sample standard error as our function $f$ and hence use
@@
 $$
-but other choices could be made (see the discussion). The sample
-standard error has an interpretation as the average distance of the
-individual rankings of the lists from the average ranking.
+where $\bar{R}(X_p)$ is the average rank assigned to item $X_p$, but
+other choices could be made. The sample standard error has an
+interpretation as the average distance of the individual rankings of
+the lists from the average ranking.
 
@@
 \begin{equation}
-S_d = \{\pi_l(r) ; r \leq d, l = 1, \ldots, L \}.
+S_d = \{R_l^{-1}(r) ; r \leq d, l = 1, \ldots, L \}.
 \end{equation}
@@
 \textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in
-      S_d\}}(L-1)A(X_p)^2}{(L-1)|S_d|}}. \label{def:sra}
+      S_d\}}(L-1)A(X_p)^2}{(L-1)|S_d|}}, \label{def:sra}
 \end{equation}
-Values of \textrm{SRA} close to zero suggests that the lists agree
-while larger values suggests disagreement. If the ranked lists are
-identical then the value of SRA will be zero for all depths $d$. The
-sequential rank agreement can be interpreted as the average distance
-of the individual rankings of the lists from the average ranking for
-each of the items we have seen until depth $d$. Note that the terms
-that are part of \eqref{def:sra} are not independent since each list
-contains the ranks from 1 to $P$ exactly once. Hence we simply use the
-pooled standard deviation as a measure of the average distance between
-rankings.
+where $|S_d|$ refers to the cardinaly of the set $S_d$. Values of
+\textrm{SRA} close to zero suggests that the lists agree while larger
+values suggests disagreement. If the ranked lists are identical then
+the value of SRA will be zero for all depths $d$. The sequential rank
+agreement can be interpreted as the average distance of the individual
+rankings of the lists from the average ranking for each of the items
+we have seen until depth $d$. Note that the terms that are part of
+\eqref{def:sra} are not independent since each list contains the ranks
+from 1 to $P$ exactly once. Hence we simply use the pooled standard
+deviation as a measure of the average distance between rankings.
 
@@
 coefficients (after standardization), and MIC was ordered by
-correlation which resulted in the rankings seen in Table~\ref{tab1}.
-The sequential rank agreement shown in Figure~\ref{fig:example1} shows
-for each depth $d$ how many ranks apart do we on average expect the
-genes found in the top $d$ part of the lists. It can be seen that
-agreement is higher in the top of the lists (smaller values on the $y$
-axis) correspond to \emph{better} than in the tails of the lists.
+correlation which resulted in the top rankings seen in Table~\ref{tab1}.
+
+The sequential rank agreement plotted in Figure~\ref{fig:example1}
+shows how many ranks apart we on average expect the genes found in the
+top $d$ part of the lists to be for each depth $d$.  It can be seen
+that agreement is higher in the top of the lists (smaller values on
+the $y$ axis) corresponding to \emph{better} agreement than in the
+tails of the lists.
 
@@
 
-If the lists would agree on all items then we would expect the
-sequential rank agreement curve to be flat. And if there are changes
-in the levels of rank agreement then this would suggest that there are
-sets of items for which the lists are in high agreement and sets of
-other items where the lists rank the items differently. In gene
-association studies we generally expect the agreement among the lists
-to be better towards the top of the lists and worse towards the end of
-the lists. In this case the sequential rank agreement curve will start
-at a low level and then increase until it levels off as in
+If the lists agree on all items then the sequential rank agreement is
+ zero for all depths. Generally,  if there are changes in the level of rank
+ agreement then this suggests that there are sets of items that are ranked
+ similarly in all lists and other sets of items that have been assigned vastly
+ different ranks in the lists. In gene association studies we
+ generally expect the agreement among the lists to be better towards
+ the top of the lists and worse towards the end of the lists. In this
+ case the sequential rank agreement curve will start at a low level
+ and then increase until it levels off as in
  Figure~\ref{fig:example1}.
@@
 
-Incomplete lists are also a common occurrence. They arises for example
-in case of missing data (items) or when some methods only rank a
-subset of the items. For example, penalized regression based on the
-Lasso provides a sparse set of predictors that have non-zero
-coefficients. There is no obvious ordering of the set of predictors
-whose coefficient has been shrunk to zero. Incomplete lists also occur
-when the analyst decides to censor all lists at depth $d$ such that
-the measurement of agreement is restricted to the first $d$ elements
-of each list.
+Incomplete lists are also a common occurrence. They arise for example
+in case of missing data (items), when comparing top $d$ list results
+from publications, or when some methods only rank a subset of the
+items. For example, penalized regression based on the Lasso provides a
+sparse set of predictors that have non-zero coefficients. There is no
+obvious ordering of the set of predictors whose coefficient has been
+shrunk to zero. Incomplete lists also occur when the analyst decides
+to censor all lists at depth $d$ such that the measurement of
+agreement is restricted to the first $d$ elements of each list.
 
@@
 \begin{equation}
-  \tilde R_l(X_p) = \left\{\begin{array}{cl} \{\pi_l^{-1}(p)\} & \text{ for } p\in \Lambda_l \\
-      \{d_l+1,\dots,P\} & \text{ for } p \not\in \Lambda_l\end{array}\right.
+  \tilde R_l(X_p) = \left\{\begin{array}{cl} \{R_l(X_p)\} & \text{ for } X_p\in \Lambda_l \\
+      \{d_l+1,\dots,P\} & \text{ for } X_p \not\in \Lambda_l\end{array}\right.
 \end{equation}
@@
 constant but at a low level). For censored lists we also expect a
-change point around the depth where the lists are censored. This is an
+change-point around the depth where the lists are censored. This is an
 artefact stemming from the fact that we assume that the remainder of
@@
 \end{eqnarray*}
-This does not only assume that there is no information data on which
-in the rankings are based but also that the methods are completely
+This does not only assume that there is no information in data on
+which the rankings are based but also that the methods are completely
 independent.
@@
 \begin{equation}
-  O(d) = \frac{| \cap_{l=1}^L \{\pi_l(r); r\leq d \} |}{d}. \label{def:overlap}
+  O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}. \label{def:overlap}
 \end{equation}
-Here $|A|$ is the cardinality of the set $A$.  Typically the overlap
-is only considered for pairwise comparisons but formula
-\eqref{def:overlap} can be applied even when $L>2$
+Typically the overlap is only considered for pairwise comparisons but
+formula \eqref{def:overlap} can be applied even when $L>2$
 \citep{Bar-Ilan2006,Boulesteix2009}.
@@
 The MALOVA study is a multidisciplinary Danish study on ovarian cancer
-\cite{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
+\citep{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
 with an ovarian tumor and referred for surgery from the participating
 departments of gynecology were enrolled continuously from December
-1994 to May 1999.
-% @Andreas: hvorfor random subsample?
-We use a random subsample of $119$ patients with a total of $58$
-patients with malignant ovarian cancers as cases and $61$ patients
-with benign ovarian tumors as controls. The DACOVA study is another
+1994 to May 1999. For the purpose of illustration we use a random
+subsample of $119$ patients with a total of $58$ patients with
+malignant ovarian cancers as cases and $61$ patients with benign
+ovarian tumors as controls. The DACOVA study is another
 multidisciplinary Danish study on ovarian cancer which included about
 $66\%$ of the female population of Denmark
-\cite{bertelsen1991protocol}. It was aimed to continuously enroll all
+\citep{bertelsen1991protocol}. It was aimed to continuously enroll all
 patients that were referred to surgery of an ovarian tumor clinically
@@
 % @Andreas: hvorfor random subsample?
-For the purpose of illustration use a random subsample of $113$
+Similarly, we use a random subsample of $113$
 patients with a total of $54$ malignant ovarian cancers and $59$
@@
 spectrum by first removing the slow-varying baseline intensities with
-the SNIP algorithm \cite{ryan1988snip} followed by a normalization
+the SNIP algorithm \citep{ryan1988snip} followed by a normalization
 with respect to the total ion counts. Finally, we standardize the 5000
@@
 scenario we assess the agreement of different statistical
-classification methods in how they rank predictors according to
-importance for classification accuracy. In the second scenario we
-assess the agreement among rankings of the same patients according to
-the predicted probabilities of having a malignant tumor. Here we apply
-the same statistical classification method to XX bootstrap samples of
-the MALOVA data to obtain XX prediction models. From each model we
-predict the risks of a malignant patients in the DACOVA study.
-Finally we estimate the agreement of the resulting XX rankings of the
-DACOVA patients to illustrate the sensitivity of the corresponding
-statistical classification method to pertubations of the learning
-data. In the latter case we further investigate the behaviour of
-different statistical methods in a simulation study.
+classification methods in how they rank predictors by different
+statistical classification methods for the two study populations. In
+the second scenario we assess the agreement among ranking of the
+predicted individual risk probabilities of having malignant tumor
+using four different prediction methods.
+For this second example we furthermore
+investigate the behaviour of different statistical methods in a
+simulation study.
 
@@
 \subsection{Agreement of predictions}
+
+Here we apply the same statistical classification
+method to XX bootstrap samples of the MALOVA data to obtain four
+prediction models. From each model we predict the risks of a malignant
+patients in the DACOVA study.  Finally we estimate the agreement of
+the resulting XX rankings of the DACOVA patients to illustrate the
+sensitivity of the corresponding statistical classification method to
+pertubations of the learning data.
 
@@
 compared visually to the curves obtained under either of the null
-distributions and simple point-wide $p$-values can be obtained for
+distributions and simple point-wise $p$-values can be obtained for
 each depth by counting the number of sequential rank agreements under
@@
 rank agreement curve and the mean null rank agreement curve as a test
-statistic. Alternatively, a change point analysis could be made on the
+statistic. Alternatively, a change-point analysis could be made on the
 sequential rank agreement in order to determine the depths at which
@@
 Finally, we have --- whenever possible --- used as much information
-from the lists as possible, i.e., we have used all available ranks. If
-we have information on the rankings as well as the corresponding
-evidence for significance that we could choose to censor the rankings
-of all the items that were deemed not to be significant. That would
+from the lists as possible, i.e., we have used all available ranks.
+We could choose to censor the rankings of all the items that were
+deemed not to be significant if information on both rankings and the
+corresponding evidence for significance was available. That would
 ensure that there would be put less emphasis on the agreement of the
@@
 methods that provide more or less the same agreement, or by using the
-rank agreement as a hyper-parameter in, say, elastic net regularized
+rank agreement to optimize a hyper-parameter in, say, elastic net regularized
 regression where the rank agreement is used to determine the mixing

### 2015-07-25 latex/paper.Rnw
Applicaitons

@@
 \subsection{Agreement of predictor ranking}
+We apply 10 different classification methods to each of the two data
+sets. All methods are capable of producing a ranking of the 5000
+predictors according to their importance with respect to the
+association with the outcome class label. The methods under
+consideration can be separated into two groups: marginal methods and
+joint methods. The marginal methods treat each of the of 5000
+predictors separately in succession while the joint models use all the
+predictors simultaneously. The marginal methods are t-test statistics,
+univariate logistic regressions, AUC-based ROC analyses and maximum
+information content correlations (MIC) statistics. For the t-test
+statistics and the logistic regressions we rank the predictors
+according to decreasing p-value while for the ROC analysis and MIC
+statistics we simply rank according to increasing numerical value. The
+joint methods are a Random Forest classifier
+\citep{breiman2001random}, logistic Lasso regression
+\citep{tibshirani1996regression}, logistic Ridge regression
+\cite{segerstedt1992ordinary}, Partial Least Squares Discriminant
+Analysis (PLS-DA) \cite{boulesteix2004pls}, Shrinkage Discriminant
+Analysis with local false discovery rate control (SDA)
+\cite{ahdesmaki2010} and nearest shrunken centroid (PAM)
+\cite{tibshirani2003class}. For the Random Forest classifier the
+importance of the predictors are ranked according to the Gini index
+while for the logistic Lasso and Ridge regressions we order by
+absolute magnitude of the estimated coefficients. For the PLS-DA
+method the importance of the predictors is based on a weighted sum of
+the absolute coefficients where the weights are proportional to the
+reduction in the sums of squares across the components, and for PAM
+the importance is given by the difference between the class centroids
+and the overall centroid. Finally, for the SDA method the ranking of
+the predictors is performed using the correlation-adjusted t-scores,
+the local false discovery rate (lFDR) is estimated and only the
+predictors with $lFDR < 0.8$ are retained. Together with the logistic
+Lasso regression and the nearest shrunken centroid methods this leads
+to a censored list according to our previous discussion. All tuning
+parameters are estimated by repeated 10-fold cross-validation using 20
+random splits. We use the implementations in the R packages
+\texttt{randomForest} \citep{liaw2002classification}, \texttt{glmnet}
+\citep{Friedman:2010aa}, \texttt{caret} \citep{Jed-Wing:2014aa},
+\texttt{sda} \citep{Ahdesmaki:2014aa} and \texttt{minerva}
+\citep{Filosi:2014aa}.
+
+The agreement of the ranking of the predictors given by the ten
+different methods can be compared in different ways using the
+sequential rank agreement. One possibility is to consider all $55$
+pair-wise agreement functions. This will be useful for identifying
+statistical methods that perform similarly on a given data set with
+respect to how the order of predictor importance is determined. This
+could for example be done by applying a statistical clustering method
+to the pair-wise sequential rank agreement curves. Here we take a
+simpler approach and consider the agreement of the importance
+determined by the different marginal methods and the agreement of the
+importance determined by the different joint methods. This leads to
+two agreement curves for each data set which we show in figure
+\ref{fig:app1}.
 
@@
 
-\subsection{Agreement of predictions}
-
-Here we apply the same statistical classification
-method to XX bootstrap samples of the MALOVA data to obtain four
-prediction models. From each model we predict the risks of a malignant
-patients in the DACOVA study.  Finally we estimate the agreement of
-the resulting XX rankings of the DACOVA patients to illustrate the
-sensitivity of the corresponding statistical classification method to
-pertubations of the learning data.
+From the figure it is seen that generally the agreement between
+ranking the predictors using the different methods is better for the
+MALOVA data than for the DACOVA data. The shapes of the agreement
+curves start at a low level followed by an increase after which they
+stabilize at some level. This corresponds to the expected behavior
+that the methods agree better with respect to the most important
+predictors than with the not so important ones. Also, it is clear that
+ranking the predictors by the marginal methods show better overall
+agreement than using the joint methods. This is not surprising due to
+the high correlation of the spectral intensities across neighboring
+mass-to-charge ratios. This leads to co-linearity problems that
+especially affects the classification methods that use a sparse
+representation of the predictors.
+
+\subsection{Agreement of predicted risk probabilities}
+We now consider the case where we predict the patient-specific risks
+of having a malignant tumor. We consider this as an example of the
+clinical decision problem of deciding the order in which a group of
+patients must be treated. This order is given by ranking the patients
+according to increasing predicted probability of malignant tumor. We
+assess the agreement of estimating the rankings of the same patients
+by four different statistical prediction methods using the sequential
+rank agreement. The four prediction models are: a Random Forest
+classifier, logistic Lasso regression, logistic Ridge regression and
+Partial Least Squares Discriminant Analysis (PLSDA).
+
+For each prediction method we estimate a model based on the MALOVA
+data and predict the risks of a malignant tumor for the patients in
+the DACOVA study. We repeat this procedure on 100 bootstrap samples of
+the MALOVA data used to construct the model and then estimate the
+agreement of the resulting 100 rankings of the same DACOVA
+patients. This serves as an assessment of the sensitivity of the
+different statistical prediction methods to perturbations of the
+learning data.
+
+For the Random Forest method we grow 500 tree using 22 randomly
+selected predictors at each split. The penalty parameter for the Lasso
+and Ridge regression models is optimized with respect to the deviance
+loss function using internal repeated 10-fold cross-validation on the
+learning data using 20 random splits. The number of components for the
+PLSDA method is optimized in a similar manner using classification
+accuracy as criterion.
+
+Figure \ref{fig:app2} shows the sequential rank agreement for the four
+methods. Generally, the ranking of the predictions made by the Random
+Forest method appears to be in better agreement across resampled
+training data than for any of the other method. The only exception is
+for the patient with the overall largest risk of a malignant tumor for
+which the logistic Lasso regression method provides a better
+agreement. Something about Ridge is surprising etc. etc. etc. that we
+now look at in a simulation study.
 
@@
 
+To further investigate the apparent lack of agreement for the ridge
+regression method we consider a small simulation study. We simulate
+two data sets each with $120$ observations and $500$ predictors. Each
+vector of predictors is drawn independently from a $500$-dimensional
+Gaussian distribution with zero mean, unit variance and a correlation
+coefficient of $0.9$ in a compound symmetry structure. The 120
+observations is split into two groups where the coefficient vector for
+the first group is zero and similarly for the second group except that
+the first $100$ values are set to one. We then predict the risk
+probabilities in a similar manner as before where we repeatedly fit
+models on bootstrapped versions on the first data set and use the
+models to predict the risk probability of the same observations from
+the second data set. Here we only consider the logistic Ridge
+regression model but we consider three different ways of optimizing
+the penalty parameter. These consists of optimizing with respect to
+the total deviance, the prediction accuracy and the area under the ROC
+curve. The resulting sequential rank agreement curves from 100
+bootstrap iterations are seen in figure \ref{fig:app3}
+
 \begin{figure}[tb]
@@
 \end{figure}
+
+Redo the figure and comment on the result. Note that none of the loss
+functions directly considers the agreement of predicted
+probabilities. The AUC is indirectly related in a complicated manner.
 

### 2015-08-05 latex/paper.Rnw
Application afsnit opdateret

--- 2015-07-23 latex/paper.Rnw
@@
 
-\section{Applications}
+\section{Application}
+We now consider an application of the sequential rank agreement to two data sets consisting of MALDI-TOF (Matrix-Assisted Laser Desorption/Ionization Time Of Flight) mass spectra obtained from blood samples from patients with either benign or malignant ovarian tumors.
+The data sets are subsamples of the Danish MALOVA and DACOVA study populations.
 
-We now consider an application of the sequential rank agreement to two
-data sets consisting of MALDI-TOF (Matrix-Assisted Laser
-Desorption/Ionization Time Of Flight) mass spectra obtained from blood
-samples from patients with either benign or malignant ovarian
-tumors. The data sets are subsamples of the Danish MALOVA and DACOVA
-study populations.
+The MALOVA study is a multidisciplinary Danish study on ovarian cancer \citep{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed with an ovarian tumor and referred for surgery from the participating departments of gynecology were enrolled continuously from December 1994 to May 1999.
+For the purpose of illustration we use a random subsample of $119$ patients with a total of $58$ patients with malignant ovarian cancers as cases and $61$ patients with benign
+ovarian tumors as controls.
+The DACOVA study is another multidisciplinary Danish study on ovarian cancer which included about $66\%$ of the female population of Denmark \citep{bertelsen1991protocol}.
+It was aimed to continuously enroll all patients that were referred to surgery of an ovarian tumor clinically suspected to be cancer during the period 1984 to 1990.
+Similarly, we use a random subsample of $113$ patients with a total of $54$ malignant ovarian cancers and $59$ benign ovarian tumors/gynecologic disorders.
 
-The MALOVA study is a multidisciplinary Danish study on ovarian cancer
-\citep{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
-with an ovarian tumor and referred for surgery from the participating
-departments of gynecology were enrolled continuously from December
-1994 to May 1999. For the purpose of illustration we use a random
-subsample of $119$ patients with a total of $58$ patients with
-malignant ovarian cancers as cases and $61$ patients with benign
-ovarian tumors as controls. The DACOVA study is another
-multidisciplinary Danish study on ovarian cancer which included about
-$66\%$ of the female population of Denmark
-\citep{bertelsen1991protocol}. It was aimed to continuously enroll all
-patients that were referred to surgery of an ovarian tumor clinically
-suspected to be cancer during the period 1984 to 1990.
-% @Andreas: hvorfor random subsample?
-Similarly, we use a random subsample of $113$
-patients with a total of $54$ malignant ovarian cancers and $59$
-benign ovarian tumors/gynecologic disorders.
+Each spectrum consists of $49642$ samples over a mass-to-charge range of $800$ to $20000$ which we down-sample on an equidistant grid of 5000 points by linear interpolation.
+We preprocess each spectrum by first removing the slow-varying baseline intensity with the SNIP algorithm \citep{ryan1988snip} followed by a normalization with respect to the total ion count.
+We then standardize the 5000 predictors to have column-wise zero mean and unit variance in each data set.
 
-Proteomic analysis was performed by MALDI-TOF mass spectrometry with a
-Bruker Ultraflex in positive, linear ion mode in the mass-to-charge
-range $800$ to $20000$ Dalton by an average of $500$ laser
-shots. Calibration of the instrument was performed using a protein
-standard mixture. Each spectrum consists of $49642$ samples over the
-mass-to-charge range which we down-sample on an equidistant grid of
-5000 points by linear interpolation. We further preprocess each
-spectrum by first removing the slow-varying baseline intensities with
-the SNIP algorithm \citep{ryan1988snip} followed by a normalization
-with respect to the total ion counts. Finally, we standardize the 5000
-predictors to have column-wise zero mean and unit variance.
+We use the two data sets to illustrate how the sequential rank agreement can be applied in two different scenarios.
+In the first scenario we assess the agreement of four different statistical classification methods in how they rank the predictors according to their importance with respect to the binary outcome.
+In the second scenario we assess the agreement among rankings of individual predicted risk probabilities of having a malignant tumor using the four methods.
+The first scenario is relevant in the context of biomarker discovery and the latter is important when e.g., ranking patients according to immediacy of treatment.
 
-We use these two data sets to illustrate how the sequential rank
-agreement can be applied in two different scenarios. In the first
-scenario we assess the agreement of different statistical
-classification methods in how they rank predictors by different
-statistical classification methods for the two study populations. In
-the second scenario we assess the agreement among ranking of the
-predicted individual risk probabilities of having malignant tumor
-using four different prediction methods.
-For this second example we furthermore
-investigate the behaviour of different statistical methods in a
-simulation study.
+We use the following four classification methods: Random Forest \citep{breiman2001random} implemented in the R package \texttt{randomForest} \citep{liaw2002classification}, logistic Lasso \citep{tibshirani1996regression} and Ridge regression \citep{segerstedt1992ordinary} both implemented in the R package \texttt{glmnet} \citep{Friedman:2010aa}, and Partial Least Squares Discriminant Analysis (PLS-DA) \citep{boulesteix2004pls} implemented in the R package \texttt{caret} \citep{Jed-Wing:2014aa}.
 
-\subsection{Agreement of predictor ranking}
+In both scenarios we use the MALOVA data to train the statistical models, and the two types of agreements are assessed with respect to perturbations of the training data in the following manner.
+We repeatedly draw a sub-sample (without replication) consisting of 90\% of the MALOVA observations and train the four models on the sub-sample.
+Each of these models produces a ranking of the 5000 predictors according to their importance with respect to discriminating between  the tumor types, and the agreement of these rankings across the different sub-samples are used to assess the stability of the importance of the predictors.
+For the Random Forest classifier the predictors are ranked according to the Gini index, while for the logistic Lasso and Ridge regression models we order the predictors by absolute magnitude of the estimated regression coefficients.
+For the PLS-DA model the importance of the predictors is based on a weighted sum of the absolute coefficients where the weights are proportional to the reduction in the sums of squares across the components.
+To assess the stability of the individual risk predictions we apply the predictors from the DACOVA data set to each of the models.
+These predicted probabilities are then ranked in decreasing order, and the agreement of the risk predictions are similarly obtained by the sequential rank agreement across the different sub-samples of the training data.
+We use 1000 iterations for the sub-sampling procedure.
 
-\begin{figure}[tb]
+All four models depend on tuning parameters.
+We estimate these by a 20-times repeated 5-fold internal cross-validation on the training data in each iteration.
+The implementation of Lasso and Ridge regression in the \texttt{glmnet} package offers three different cross-validated optimization criteria for the penalty parameter: total deviance, classification accuracy and Area Under ROC.
+We apply all three criteria to our data to investigate their effect on the agreement.
+
+\begin{figure}[tbp]
 \begin{center}
-\includegraphics[width=.5\textwidth]{pics/marginalVsModel.pdf}
+\includegraphics[width=.5\textwidth]{pics/riskAgreementPlot}%
+\includegraphics[width=.5\textwidth]{pics/predictorAgreementPlot}
 \end{center}
- \caption{Marginal}
+ \caption{Sequential rank agreement of predicted risk probabilities for the DACOVA data (left) and estimated predictor importance (right) based on 1000 sub-samples of the MALOVA training data.}
  \label{fig:app1}
@@
 
-\subsection{Agreement of predictions}
+Figure \ref{fig:app1} (left) shows the sequential rank agreement applied to the estimated risk probabilities of the $113$ patients in the DACOVA data set obtained from the $1000$ sub-samples of the MALOVA data, and Figure \ref{fig:app1} (right) shows the sequential rank agreement of the estimated importance of the 5000 predictors.
+For clarity we only show the agreement for the latter up to list depth 600 where all the curves have approximately become stable.
+We note here that the predictor importance estimated with the logistic Lasso regression model is capable of producing censored lists depending on the value of the penalty parameter.
+With the deviance criterion the median number of non-zero coefficients was 33 (range 18 to 52).
+In order to calculate the sequential rank agreement we therefore applied Algorithm \ref{sra-algorithm} with $B = 1000$ and $P=5000$.
+The effect of the censoring caused by the sparsity of the Lasso is clearly visible in the agreement curve for the importance of the predictors.
 
-Here we apply the same statistical classification
-method to XX bootstrap samples of the MALOVA data to obtain four
-prediction models. From each model we predict the risks of a malignant
-patients in the DACOVA study.  Finally we estimate the agreement of
-the resulting XX rankings of the DACOVA patients to illustrate the
-sensitivity of the corresponding statistical classification method to
-pertubations of the learning data.
+For most of the sequential rank agreement curves we see as expected that they start low followed by an increase until they approximately become constant.
+This has the interpretation that the classification methods show higher agreement with respect to the highest ranked items (the most high-risk patients and the most important predictors respectively) then for the items deeper in the lists across the different sub-samples.
+The change-point where the curves become approximately constant is the list depth where the ranks of the remaining items become almost random.
+The only exceptions are the predictor importance agreements using the logistic Ridge regression models and the risk probability agreement using the logistic Lasso regression with the AUC criterion which show higher disagreement in the top of the lists.
+The reason for this in the Lasso case is that the AUC criterion tends to favor a very high value of the penalty parameter causing only a single predictor to be selected in all of the 1000 iterations.
+This results in a lack of generalizability to the DACOVA data  giving rise to the higher disagreement in the predicted risk probabilities.
+In the extreme case where the penalty becomes so high that none of the predictors are selected by the Lasso, the sequential rank agreement for the predicted probabilities becomes undefined since all the ranks will be ties.
 
-\begin{figure}[tb]
-\begin{center}
-\includegraphics[width=.5\textwidth]{pics/thomas.pdf}
-\end{center}
- \caption{Thomas}
- \label{fig:app2}
-\end{figure}
+From Figure \ref{fig:app1} it can also be seen that some of the methods appear to be better at agreeing with respect to the predicted probabilities than at agreeing with respect to the ranking of the importance of the predictors and vice versa.
+Ridge regression performs very well for the risk predictions but is outperformed by PLS-DA for the predictor importances.
+On the other hand, PLS-DA appears to be similarly worse for the risk predictions compared to Ridge regression.
+The Random Forest method seems to perform equally well in both scenarios.
+Lasso regression (except with the AUC criterion as discussed previously) performs very well for the high-risk patients but very bad for the importance of the predictors.
+This reason for this is the high auto-correlation between the intensities in the mass spectra which leads to collinearity in the regression model.
+It is well-known that variable selection with the Lasso penalty does not preform very well when the predictors are highly correlated.
+The collinearity does, however, not affect the risk predictions, since it is not so important what specific variable that get selected from a group of highly correlated predictors when the purpose is risk predictions.
 
-\begin{figure}[tb]
-\begin{center}
-\includegraphics[width=.5\textwidth]{pics/simulationTest2.pdf}
-\end{center}
- \caption{SimulationTest2}
- \label{fig:app3}
-\end{figure}
+Finally, it would appear that Ridge regression with the AUC criterion for the penalty parameter achieves the best performance with respect to the stability of ranking the individual predicted risk probabilities.
+It must, however, be stressed that sequential rank agreement is only concerned with whether similar orderings of the patients is obtained under perturbations of the training data.
+It is not concerned with whether such an ordering is the correct one.
+This is similarly to the usual distinction between accuracy and precision.
+Looking at the distributions of the AUCs across the $1000$ iterations for each of the models with respect to the known class labels of the DACOVA data we see that the PLS-DA model attained the best AUC with a median value of $0.71$ (IQR = 0.07) while the Ridge regression with AUC criterion has a median AUC of $0.49$ (IQR = 0.02).
+This means that while Ridge regression with the AUC criterion achieves the highest rank agreement it performs similar to a random coin toss with respect to classification.
+Both concerns are of course important in practice and for this data set it appears that PLS-DA is the best method with respect to both concerns.
 

### 2015-08-05 latex/paper.Rnw
Fixing application section and correcting minor spelling mistakes

@@
   agreement as a function of the depth of the lists. This may help to
-  indentify change-points in the agreement of the lists. The
+  identify change-points in the agreement of the lists. The
   usefulness of these tools are illustrated using gene rankings, and
@@
 \end{equation}
-where $|S_d|$ refers to the cardinaly of the set $S_d$. Values of
+where $|S_d|$ refers to the cardinality of the set $S_d$. Values of
 \textrm{SRA} close to zero suggests that the lists agree while larger
@@
 as described in \citet{Dudoit2002} and the four different analysis
-approaces were: marginal two-sample $t$ tests, marginal logistic
-regression analyses, logistic regression eleastic net
+approaches were: marginal two-sample $t$ tests, marginal logistic
+regression analyses, elastic net logistic regression 
 \citep{friedman2010regularization}, and marginal maximum information
@@
 overlap closely resembles the sequential rank agreement. In this
-section we present an extention of the average overlap to more that
+section we present an extension of the average overlap to more that
 two lists and compare this to sequential rank agreement.
@@
 agreement from rank 3 as we did from the full dataset. This indicates
-that sequential rank is more robust against small pertubations of the
+that sequential rank is more robust against small perturbations of the
 data. For the average overlap the curves completely change when these
@@
 We repeatedly draw a sub-sample (without replication) consisting of 90\% of the MALOVA observations and train the four models on the sub-sample.
-Each of these models produces a ranking of the 5000 predictors according to their importance with respect to discriminating between  the tumor types, and the agreement of these rankings across the different sub-samples are used to assess the stability of the importance of the predictors.
+Each of the models produces a ranking of the 5000 predictors according to their importance with respect to discriminating between  the tumor types, and the agreement of these rankings across the different sub-samples are used to assess the stability of ranking the importance of the predictors.
 For the Random Forest classifier the predictors are ranked according to the Gini index, while for the logistic Lasso and Ridge regression models we order the predictors by absolute magnitude of the estimated regression coefficients.
@@
 To assess the stability of the individual risk predictions we apply the predictors from the DACOVA data set to each of the models.
-These predicted probabilities are then ranked in decreasing order, and the agreement of the risk predictions are similarly obtained by the sequential rank agreement across the different sub-samples of the training data.
+These predicted probabilities are then ranked in decreasing order, and the agreement of the risk predictions are similarly obtained by the sequential rank agreement applied across the different sub-samples of the training data.
 We use 1000 iterations for the sub-sampling procedure.
@@
 All four models depend on tuning parameters.
-We estimate these by a 20-times repeated 5-fold internal cross-validation on the training data in each iteration.
+These are estimated by 20 times repeated 5-fold cross-validation on the training data in each iteration.
 The implementation of Lasso and Ridge regression in the \texttt{glmnet} package offers three different cross-validated optimization criteria for the penalty parameter: total deviance, classification accuracy and Area Under ROC.
@@
 For clarity we only show the agreement for the latter up to list depth 600 where all the curves have approximately become stable.
-We note here that the predictor importance estimated with the logistic Lasso regression model is capable of producing censored lists depending on the value of the penalty parameter.
+We note here that the predictor importance estimated with the Lasso model is capable of producing censored lists depending on the value of the penalty parameter.
 With the deviance criterion the median number of non-zero coefficients was 33 (range 18 to 52).
 In order to calculate the sequential rank agreement we therefore applied Algorithm \ref{sra-algorithm} with $B = 1000$ and $P=5000$.
-The effect of the censoring caused by the sparsity of the Lasso is clearly visible in the agreement curve for the importance of the predictors.
+The effect of the censoring caused by the sparsity of the Lasso is clearly visible in the agreement curves.
 
@@
 The change-point where the curves become approximately constant is the list depth where the ranks of the remaining items become almost random.
-The only exceptions are the predictor importance agreements using the logistic Ridge regression models and the risk probability agreement using the logistic Lasso regression with the AUC criterion which show higher disagreement in the top of the lists.
-The reason for this in the Lasso case is that the AUC criterion tends to favor a very high value of the penalty parameter causing only a single predictor to be selected in all of the 1000 iterations.
-This results in a lack of generalizability to the DACOVA data  giving rise to the higher disagreement in the predicted risk probabilities.
+The only exceptions are the predictor importance agreements using the Ridge models and the risk probability agreement using the Lasso model with the AUC criterion which show higher disagreement in the top of the lists.
+The reason for this in the Lasso case is that optimizing with respect to the AUC criterion tends to favor a very high value of the penalty parameter causing only a single predictor to be selected in each of the 1000 iterations.
+This results in a lack of generalizability to the DACOVA data which gives rise to the higher disagreement in the predicted risk probabilities.
 In the extreme case where the penalty becomes so high that none of the predictors are selected by the Lasso, the sequential rank agreement for the predicted probabilities becomes undefined since all the ranks will be ties.
 
-From Figure \ref{fig:app1} it can also be seen that some of the methods appear to be better at agreeing with respect to the predicted probabilities than at agreeing with respect to the ranking of the importance of the predictors and vice versa.
+From Figure \ref{fig:app1} it can also be seen that some of the methods show better agreement with respect to the predicted probabilities than for ranking  the importance of the predictors and vice versa.
 Ridge regression performs very well for the risk predictions but is outperformed by PLS-DA for the predictor importances.
@@
 The Random Forest method seems to perform equally well in both scenarios.
-Lasso regression (except with the AUC criterion as discussed previously) performs very well for the high-risk patients but very bad for the importance of the predictors.
-This reason for this is the high auto-correlation between the intensities in the mass spectra which leads to collinearity in the regression model.
+Lasso regression (except with the AUC criterion as discussed previously) performs very well at ranking the high-risk patients but very poorly for the importance of the predictors.
+This reason for this is the high auto-correlation between the intensities in the mass spectra which leads to collinearity issues in the regression models.
 It is well-known that variable selection with the Lasso penalty does not preform very well when the predictors are highly correlated.
-The collinearity does, however, not affect the risk predictions, since it is not so important what specific variable that get selected from a group of highly correlated predictors when the purpose is risk predictions.
+This collinearity does, however, not affect the risk predictions, as seen on the left part of the figure, since it is not so important which specific variable that get selected from a group of highly correlated predictors when the purpose is risk predictions.
 
 Finally, it would appear that Ridge regression with the AUC criterion for the penalty parameter achieves the best performance with respect to the stability of ranking the individual predicted risk probabilities.
-It must, however, be stressed that sequential rank agreement is only concerned with whether similar orderings of the patients is obtained under perturbations of the training data.
+It must, however, be stressed that the sequential rank agreement is only concerned with whether similar orderings of the patients is obtained under perturbations of the training data.
 It is not concerned with whether such an ordering is the correct one.
-This is similarly to the usual distinction between accuracy and precision.
+This is similar to the usual distinction between accuracy and precision.
 Looking at the distributions of the AUCs across the $1000$ iterations for each of the models with respect to the known class labels of the DACOVA data we see that the PLS-DA model attained the best AUC with a median value of $0.71$ (IQR = 0.07) while the Ridge regression with AUC criterion has a median AUC of $0.49$ (IQR = 0.02).
-This means that while Ridge regression with the AUC criterion achieves the highest rank agreement it performs similar to a random coin toss with respect to classification.
-Both concerns are of course important in practice and for this data set it appears that PLS-DA is the best method with respect to both concerns.
+This means that while Ridge regression optimized with respect to the AUC criterion achieves the best rank agreement it performs similar to a random coin toss with respect to classification on the DACOVA data.
+Both concerns are of course important in practice and for this data set it appears that PLS-DA is the overall best method taking both agreement and classification into account.
 
@@
 study the ranks obtained from different analysis methods on the same
-data as well as to evalaute the stability of the ranks from a single
+data as well as to evaluate the stability of the ranks from a single
 method by bootstrapping the data repeatedly and comparing the ranks
@@
 lists. As such we cannot determine if the rankings are good but only
-whether they agree. The sequential rank agreemenet curve can be
+whether they agree. The sequential rank agreement curve can be
 compared visually to the curves obtained under either of the null
@@
 that generally apply to ranking of items and/or
-indiviuals. Collinearity in particular can be a huge problem when
+individuals. Collinearity in particular can be a huge problem when
 bootstrapping data or when comparing different analysis methods. For
-example, marginal analyses where each item is analyzed seperately will
+example, marginal analyses where each item is analyzed separately will
 assign similar ranks to two highly correlated predictors while methods

### 2015-08-05 R/RcppExports.R
MAD addition to code

@@ -18,22 +18,24 @@ overlap <- function(rankMat) {
 #' @param rankMat A matrix with k columns corresponding to the k ranked lists. Elements of each column are integers between 1 and the length of the lists
 #' @param maxlength The maximum depth that are needed XXX
 #' @param B The number of resamples to use in the presence of censored lists
-#' @param cens A vector of integer values that 
-#' @return A vector of the same length as the rows in rankMat containing the sequential rank agreement between the lists for each depth
+#' @param cens A vector of integer values that
+#' @param type The type of distance measure to use: 0 (the default) is the variance while 1 is MAD (mean absolute deviation)
+#' @return A vector of the same length as the rows in rankMat containing the squared (!) sequential rank agreement between the lists for each depth. If the MAD type was chosen then the sequential MAD values are returned
 #' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
 #' @export
-sracpp <- function(rankMat, maxlength, B, cens) {
-    .Call('SuperRanker_sracpp', PACKAGE = 'SuperRanker', rankMat, maxlength, B, cens)
+sracpp <- function(rankMat, maxlength, B, cens, type = 0L) {
+    .Call('SuperRanker_sracpp', PACKAGE = 'SuperRanker', rankMat, maxlength, B, cens, type)
 }
 
 #' Compute the sequential rank agreement between k ranked lists
 #' 
 #' @description Computes the sequential rank agreement (number of items present in all k lists divided by the current rank) for each rank in the k lists
 #' @param rankMat A matrix with k columns corresponding to the k ranked lists. Elements of each column are integers between 1 and the length of the lists
-#' @return A vector of the same length as the rows in rankMat containing the sequential rank agreement between the lists for each depth
+#' @param type The type of distance measure to use: 0 (the default) is the variance while 1 is MAD (mean absolute deviation)
+#' @return A vector of the same length as the rows in rankMat containing the sequential rank agreement between the lists for each depth (squared for type=0)
 #' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
 #' @export
-sracppfull <- function(rankMat) {
-    .Call('SuperRanker_sracppfull', PACKAGE = 'SuperRanker', rankMat)
+sracppfull <- function(rankMat, type = 0L) {
+    .Call('SuperRanker_sracppfull', PACKAGE = 'SuperRanker', rankMat, type)
 }
 

### 2015-08-05 src/RcppExports.cpp

@@ -17,8 +17,8 @@ BEGIN_RCPP
 END_RCPP
 }
 // sracpp
-NumericVector sracpp(IntegerMatrix rankMat, int maxlength, int B, IntegerVector cens);
-RcppExport SEXP SuperRanker_sracpp(SEXP rankMatSEXP, SEXP maxlengthSEXP, SEXP BSEXP, SEXP censSEXP) {
+NumericVector sracpp(IntegerMatrix rankMat, int maxlength, int B, IntegerVector cens, int type);
+RcppExport SEXP SuperRanker_sracpp(SEXP rankMatSEXP, SEXP maxlengthSEXP, SEXP BSEXP, SEXP censSEXP, SEXP typeSEXP) {
 BEGIN_RCPP
     Rcpp::RObject __result;
     Rcpp::RNGScope __rngScope;
@@ -26,18 +26,20 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< int >::type maxlength(maxlengthSEXP);
     Rcpp::traits::input_parameter< int >::type B(BSEXP);
     Rcpp::traits::input_parameter< IntegerVector >::type cens(censSEXP);
-    __result = Rcpp::wrap(sracpp(rankMat, maxlength, B, cens));
+    Rcpp::traits::input_parameter< int >::type type(typeSEXP);
+    __result = Rcpp::wrap(sracpp(rankMat, maxlength, B, cens, type));
     return __result;
 END_RCPP
 }
 // sracppfull
-NumericVector sracppfull(IntegerMatrix rankMat);
-RcppExport SEXP SuperRanker_sracppfull(SEXP rankMatSEXP) {
+NumericVector sracppfull(IntegerMatrix rankMat, int type);
+RcppExport SEXP SuperRanker_sracppfull(SEXP rankMatSEXP, SEXP typeSEXP) {
 BEGIN_RCPP
     Rcpp::RObject __result;
     Rcpp::RNGScope __rngScope;
     Rcpp::traits::input_parameter< IntegerMatrix >::type rankMat(rankMatSEXP);
-    __result = Rcpp::wrap(sracppfull(rankMat));
+    Rcpp::traits::input_parameter< int >::type type(typeSEXP);
+    __result = Rcpp::wrap(sracppfull(rankMat, type));
     return __result;
 END_RCPP
 }

### 2015-08-05 src/sra.cpp

@@ -33,12 +33,18 @@ double square( double x ){
 }
 
 
-NumericVector SdPerRow(NumericMatrix obj) {
+NumericVector SdPerRow(NumericMatrix obj, int type) {
 
   NumericVector res(obj.nrow());
   
   for (int i=0; i<obj.nrow(); i++) {
-    res[i] = var(obj.row(i));
+    if (type == 1) {
+      // MAD
+      res[i] = mean(abs(obj.row(i) - mean(obj.row(i))));
+    }
+    else {
+      res[i] = var(obj.row(i));
+    }
   }
  
   return(res);
@@ -51,15 +57,20 @@ NumericVector SdPerRow(NumericMatrix obj) {
 //' @param rankMat A matrix with k columns corresponding to the k ranked lists. Elements of each column are integers between 1 and the length of the lists
 //' @param maxlength The maximum depth that are needed XXX
 //' @param B The number of resamples to use in the presence of censored lists
-//' @param cens A vector of integer values that 
-//' @return A vector of the same length as the rows in rankMat containing the sequential rank agreement between the lists for each depth
+//' @param cens A vector of integer values that
+//' @param type The type of distance measure to use: 0 (the default) is the variance while 1 is MAD (mean absolute deviation)
+//' @return A vector of the same length as the rows in rankMat containing the squared (!) sequential rank agreement between the lists for each depth. If the MAD type was chosen then the sequential MAD values are returned
 //' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
 //' @export
 //[[Rcpp::export]]
-NumericVector sracpp(IntegerMatrix rankMat, int maxlength, int B, IntegerVector cens) {
+NumericVector sracpp(IntegerMatrix rankMat, int maxlength, int B, IntegerVector cens, int type=0) {
 
   // The number of lists
   int nLists = rankMat.ncol();
+
+  // Sanity check of type
+  if (type != 1)
+    type = 0;
 
   // Need a numeric copy of this for the sort function. Doesn't really make sense but
   // keeping it for now 
@@ -138,7 +149,7 @@ NumericVector sracpp(IntegerMatrix rankMat, int maxlength, int B, IntegerVector 
       for (int l = 0; l < nLists; l++) {
 	itemRank(_, l) = order_( x( _ , l) );
       }
-      itemMetric = SdPerRow(itemRank);
+      itemMetric = SdPerRow(itemRank, type);
 
       // Now compute the agreement for each depth
[…]

### 2015-08-05 latex/paper.Rnw
Updating application section

@@
 
-We use the following four classification methods: Random Forest \citep{breiman2001random} implemented in the R package \texttt{randomForest} \citep{liaw2002classification}, logistic Lasso \citep{tibshirani1996regression} and Ridge regression \citep{segerstedt1992ordinary} both implemented in the R package \texttt{glmnet} \citep{Friedman:2010aa}, and Partial Least Squares Discriminant Analysis (PLS-DA) \citep{boulesteix2004pls} implemented in the R package \texttt{caret} \citep{Jed-Wing:2014aa}.
+We use the following four classification methods: Random Forest \citep{breiman2001random} implemented in the R package \texttt{randomForest} \citep{liaw2002classification}, logistic Lasso \citep{tibshirani1996regression} and Ridge regression \citep{segerstedt1992ordinary} both implemented in the R package \texttt{glmnet} \citep{friedman2010regularization}, and Partial Least Squares Discriminant Analysis (PLS-DA) \citep{boulesteix2004pls} implemented in the R package \texttt{caret} \citep{Jed-Wing:2014aa}.
 
@@
 All four models depend on tuning parameters.
-These are estimated by 20 times repeated 5-fold cross-validation on the training data in each iteration.
+We estimate these by a 20 times repeated 5-fold cross-validation on the training data in each iteration.
 The implementation of Lasso and Ridge regression in the \texttt{glmnet} package offers three different cross-validated optimization criteria for the penalty parameter: total deviance, classification accuracy and Area Under ROC.
@@
 
-\begin{figure}[tbp]
+\begin{figure}[htbp]
 \begin{center}
@@
 \end{center}
- \caption{Sequential rank agreement of predicted risk probabilities for the DACOVA data (left) and estimated predictor importance (right) based on 1000 sub-samples of the MALOVA training data.}
+ \caption{Sequential rank agreement of predicted risk probabilities for the DACOVA data (left) and predictor importance (right) based on 1000 sub-samples of the MALOVA training data.}
  \label{fig:app1}
@@
 For clarity we only show the agreement for the latter up to list depth 600 where all the curves have approximately become stable.
-We note here that the predictor importance estimated with the Lasso model is capable of producing censored lists depending on the value of the penalty parameter.
-With the deviance criterion the median number of non-zero coefficients was 33 (range 18 to 52).
+We note that the predictor importances estimated with the Lasso model are capable of producing censored lists depending on the value of the penalty parameter.
+Under the deviance optimization criterion the median number of non-zero coefficients was 33 (range 18 to 52).
 In order to calculate the sequential rank agreement we therefore applied Algorithm \ref{sra-algorithm} with $B = 1000$ and $P=5000$.
@@
 
-For most of the sequential rank agreement curves we see as expected that they start low followed by an increase until they approximately become constant.
-This has the interpretation that the classification methods show higher agreement with respect to the highest ranked items (the most high-risk patients and the most important predictors respectively) then for the items deeper in the lists across the different sub-samples.
-The change-point where the curves become approximately constant is the list depth where the ranks of the remaining items become almost random.
-The only exceptions are the predictor importance agreements using the Ridge models and the risk probability agreement using the Lasso model with the AUC criterion which show higher disagreement in the top of the lists.
-The reason for this in the Lasso case is that optimizing with respect to the AUC criterion tends to favor a very high value of the penalty parameter causing only a single predictor to be selected in each of the 1000 iterations.
+For most of the sequential rank agreement curves we see, as expected, that they start low followed by an increase until they approximately become constant.
+This has the interpretation that the classification methods show higher agreement with respect to the highest ranked items (the most high-risk patients and the most important predictors respectively) than for the items deeper in the lists across the different sub-samples.
+The change-points where the curves become approximately constant are the list depths where the ranks of the remaining items become close to random.
+
+The only exceptions to the expected shape of the agreement curves are the predictor importance agreements from the Ridge models and the risk prediction agreement from the Lasso model with the AUC criterion.
+They all show higher disagreement in the top of the lists followed by a decrease.
+The explanations for their behavior is different for the two methods.
+For the Lasso model the reason is that optimizing the penalty parameter with respect to the AUC criterion tends to favor a very high penalty value causing only a single predictor to be selected in each of the 1000 iterations.
 This results in a lack of generalizability to the DACOVA data which gives rise to the higher disagreement in the predicted risk probabilities.
@@
 
-From Figure \ref{fig:app1} it can also be seen that some of the methods show better agreement with respect to the predicted probabilities than for ranking  the importance of the predictors and vice versa.
-Ridge regression performs very well for the risk predictions but is outperformed by PLS-DA for the predictor importances.
-On the other hand, PLS-DA appears to be similarly worse for the risk predictions compared to Ridge regression.
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=.5\textwidth]{pics/ridgeCensoring}%
+\includegraphics[width=.5\textwidth]{pics/aucPlot}
+\end{center}
+ \caption{Sequential rank agreement for Ridge regression with an artificial censoring on the absolute value of the regression coefficients (left) and box plots of AUC values across the $1000$ sub-samples with respect to the known class labels of the DACOVA data (right).}
+ \label{fig:app2}
+\end{figure}
+
+The reason behind the behavior of the sequential rank agreement of predictor importances estimated by the Ridge models is more subtle.
+Looking at the distribution of the regression coefficients it is seen that a large proportion of them are numerically very close to zero and of almost equal absolute value.
+This implies that when they are ranked according to their magnitude, their actual order becomes more uncertain and more close to a random permutation.
+This problem can be alleviate by introducing an artificial censoring depending on the absolute value of the coefficients.
+Figure \ref{fig:app2} (left) shows the sequential rank agreement curve of the rankings of the predictor importances from the Ridge model using the deviance criterion (same curve as in figure \ref{fig:app1} (left)) as well as an agreement curve where the absolute value of the regression coefficients have been censored at the $0.1\%$ quantile.
+In our application the median number of censored predictors across the 1000 sub-samples is 5 (range 1 to 11).
+Even though the number of censored predictors is very small compared to the total number of predictors, the effect on the sequential rank agreement is substantial as it has now flipped to show higher agreement in the top of the lists as was expected.
+
+Returning to Figure \ref{fig:app1}, it can further be seen that some of the methods show better agreement with respect to the predicted probabilities than for ranking the importance of the predictors and vice versa.
+Ridge regression performs very well for the risk predictions, but PLS-DA performs better for the predictor importances and similarly worse for the risk predictions.
 The Random Forest method seems to perform equally well in both scenarios.
-Lasso regression (except with the AUC criterion as discussed previously) performs very well at ranking the high-risk patients but very poorly for the importance of the predictors.
+Lasso regression (except with the AUC criterion) performs very well at ranking the high-risk patients but very poorly at ranking the importance of the predictors.
 This reason for this is the high auto-correlation between the intensities in the mass spectra which leads to collinearity issues in the regression models.
 It is well-known that variable selection with the Lasso penalty does not preform very well when the predictors are highly correlated.
-This collinearity does, however, not affect the risk predictions, as seen on the left part of the figure, since it is not so important which specific variable that get selected from a group of highly correlated predictors when the purpose is risk predictions.
+This collinearity does, however, not affect the agreement of the risk predictions, as seen on the left part of the figure, since it is not so important which specific variable that gets selected from a group of highly correlated predictors when the purpose is risk predictions.
 
 Finally, it would appear that Ridge regression with the AUC criterion for the penalty parameter achieves the best performance with respect to the stability of ranking the individual predicted risk probabilities.
-It must, however, be stressed that the sequential rank agreement is only concerned with whether similar orderings of the patients is obtained under perturbations of the training data.
+It must, however, be stressed that the sequential rank agreement in this application is only concerned with whether similar orderings of the patients is obtained under perturbations of the training data.
 It is not concerned with whether such an ordering is the correct one.
 This is similar to the usual distinction between accuracy and precision.
-Looking at the distributions of the AUCs across the $1000$ iterations for each of the models with respect to the known class labels of the DACOVA data we see that the PLS-DA model attained the best AUC with a median value of $0.71$ (IQR = 0.07) while the Ridge regression with AUC criterion has a median AUC of $0.49$ (IQR = 0.02).
-This means that while Ridge regression optimized with respect to the AUC criterion achieves the best rank agreement it performs similar to a random coin toss with respect to classification on the DACOVA data.
-Both concerns are of course important in practice and for this data set it appears that PLS-DA is the overall best method taking both agreement and classification into account.
+Looking at the distributions of the AUCs illustrated in Figure \ref{fig:app2} (left) across the $1000$ sub-samples for a selection of the models with respect to the known class labels of the DACOVA data, we see that PLS-DA attains the highest AUC with a median value of $0.71$ while the Ridge model with the AUC criterion attains a median AUC of $0.49$.
+This implies that while Ridge regression optimized with respect to the AUC criterion achieves the best sequential rank agreement it performs similar to a random coin toss with respect to classifying the DACOVA patients.
+Both concerns are of course important in practice and for this data set it appears that PLS-DA is the overall best method taking both agreement and classification accuracy into account.
 

### 2015-08-06 R/sra.R
Initial MAD implemcdentation

@@ -8,6 +8,7 @@
 #' @param na.strings A vector of strings/values that represent missing values in addition to NA. Defaults to NULL which means only NA are censored values.
 #' @param B An integer giving the number of randomization to sample
 #' over in the case of censored observations
+#' @param type The type of measure to use. Either sd (standard deviation - the default) or mad (mean absolute deviance)
 #' @return A vector of the sequential rank agreement
 ##' @examples
 ##'
@@ -41,19 +42,19 @@
 #'
 #' @rdname sra
 #' @export
-sra <- function(object, B=1, ...) {
+sra <- function(object, B=1, type=c("sd", "mad"), ...) {
   UseMethod("sra")
 }
 
 #' @rdname sra
 #' @export
-sra.default <- function(object, B=1) {
+sra.default <- function(object, B=1, type=c("sd", "mad")) {
     stop("Input must be either a matrix, a data.frame or a list.")
 }
 
 #' @rdname sra
 #' @export
-sra.matrix <- function(object, B=1, na.strings=NULL, nitems=nrow(object)) {
+sra.matrix <- function(object, B=1, na.strings=NULL, nitems=nrow(object), type=c("sd", "mad")) {
     if (!is.matrix(object))
         stop("Input object must be a matrix")
 
@@ -75,14 +76,14 @@ sra.matrix <- function(object, B=1, na.strings=NULL, nitems=nrow(object)) {
         object <- rbind(object, glue)
     }
     object <- lapply(1:NCOL(object),function(j)object[,j]) # Convert matrix to list
-    sra.list(object, B=B, nitems=nitems)
+    sra.list(object, B=B, nitems=nitems, type=type)
 }
 
 
 
 #' @rdname sra
 #' @export
-sra.list <- function(object, B=1, na.strings=NULL, nitems=max(sapply(object, length))) {
+sra.list <- function(object, B=1, na.strings=NULL, nitems=max(sapply(object, length)), type=c("sd", "mad")) {
     # Make sure that the input object ends up as a matrix with integer columns all
     # consisting of elements from 1 and up to listlength
 
@@ -90,6 +91,8 @@ sra.list <- function(object, B=1, na.strings=NULL, nitems=max(sapply(object, len
 
     nlists <- length(object)
     nitems <- nitems ## force evaluation here
+
+    type <- match.arg(type)
 
     ## Sanity checks
     object <- lapply(1:length(object),function(j){
[…]

### 2015-08-06 latex/paper.Rnw
Minor tweaks

@@
 \section{Application}
-We now consider an application of the sequential rank agreement to two data sets consisting of MALDI-TOF (Matrix-Assisted Laser Desorption/Ionization Time Of Flight) mass spectra obtained from blood samples from patients with either benign or malignant ovarian tumors.
-The data sets are subsamples of the Danish MALOVA and DACOVA study populations.
-
-The MALOVA study is a multidisciplinary Danish study on ovarian cancer \citep{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed with an ovarian tumor and referred for surgery from the participating departments of gynecology were enrolled continuously from December 1994 to May 1999.
-For the purpose of illustration we use a random subsample of $119$ patients with a total of $58$ patients with malignant ovarian cancers as cases and $61$ patients with benign
-ovarian tumors as controls.
-The DACOVA study is another multidisciplinary Danish study on ovarian cancer which included about $66\%$ of the female population of Denmark \citep{bertelsen1991protocol}.
-It was aimed to continuously enroll all patients that were referred to surgery of an ovarian tumor clinically suspected to be cancer during the period 1984 to 1990.
-Similarly, we use a random subsample of $113$ patients with a total of $54$ malignant ovarian cancers and $59$ benign ovarian tumors/gynecologic disorders.
+We now consider an application of the sequential rank agreement to two
+data sets consisting of MALDI-TOF (Matrix-Assisted Laser
+Desorption/Ionization Time Of Flight) mass spectra obtained from blood
+samples from patients with either benign or malignant ovarian
+tumors. The data sets are subsamples of the Danish MALOVA and DACOVA
+study populations.
+
+The MALOVA study is a multidisciplinary Danish study on ovarian cancer
+\citep{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
+with an ovarian tumor and referred for surgery from the participating
+departments of gynecology were enrolled continuously from December
+1994 to May 1999. For the purpose of illustration we use a random
+subsample of $119$ patients with a total of $58$ patients with
+malignant ovarian cancers as cases and $61$ patients with benign
+ovarian tumors as controls. The DACOVA study is another
+multidisciplinary Danish study on ovarian cancer which included about
+$66\%$ of the female population of Denmark
+\citep{bertelsen1991protocol}. The study aimed to continuously enroll
+all patients that were referred to surgery of an ovarian tumor
+clinically suspected to be cancer during the period from 1984 to
+1990. Similarly, we use a random subsample from the DACOVA study of
+$113$ patients with a total of $54$ malignant ovarian cancers and $59$
+benign ovarian tumors/gynecologic disorders.
 
@@
 
-We use the following four classification methods: Random Forest \citep{breiman2001random} implemented in the R package \texttt{randomForest} \citep{liaw2002classification}, logistic Lasso \citep{tibshirani1996regression} and Ridge regression \citep{segerstedt1992ordinary} both implemented in the R package \texttt{glmnet} \citep{friedman2010regularization}, and Partial Least Squares Discriminant Analysis (PLS-DA) \citep{boulesteix2004pls} implemented in the R package \texttt{caret} \citep{Jed-Wing:2014aa}.
-
-In both scenarios we use the MALOVA data to train the statistical models, and the two types of agreements are assessed with respect to perturbations of the training data in the following manner.
-We repeatedly draw a sub-sample (without replication) consisting of 90\% of the MALOVA observations and train the four models on the sub-sample.
-Each of the models produces a ranking of the 5000 predictors according to their importance with respect to discriminating between  the tumor types, and the agreement of these rankings across the different sub-samples are used to assess the stability of ranking the importance of the predictors.
-For the Random Forest classifier the predictors are ranked according to the Gini index, while for the logistic Lasso and Ridge regression models we order the predictors by absolute magnitude of the estimated regression coefficients.
-For the PLS-DA model the importance of the predictors is based on a weighted sum of the absolute coefficients where the weights are proportional to the reduction in the sums of squares across the components.
-To assess the stability of the individual risk predictions we apply the predictors from the DACOVA data set to each of the models.
-These predicted probabilities are then ranked in decreasing order, and the agreement of the risk predictions are similarly obtained by the sequential rank agreement applied across the different sub-samples of the training data.
-We use 1000 iterations for the sub-sampling procedure.
-
-All four models depend on tuning parameters.
-We estimate these by a 20 times repeated 5-fold cross-validation on the training data in each iteration.
-The implementation of Lasso and Ridge regression in the \texttt{glmnet} package offers three different cross-validated optimization criteria for the penalty parameter: total deviance, classification accuracy and Area Under ROC.
-We apply all three criteria to our data to investigate their effect on the agreement.
+Four classification methods are considered: Random Forest
+\citep{breiman2001random} implemented in the R package
+\texttt{randomForest} \citep{liaw2002classification}, logistic Lasso
+\citep{tibshirani1996regression} and Ridge regression
+\citep{segerstedt1992ordinary} both implemented in the R package
+\texttt{glmnet} \citep{friedman2010regularization}, and Partial Least
+Squares Discriminant Analysis (PLS-DA) \citep{boulesteix2004pls}
+implemented in the R package \texttt{caret} \citep{Jed-Wing:2014aa}.
+
+In both scenarios we use the MALOVA data to train the statistical
+models, and 
+%the two types of 
+in both situations the
+agreements are assessed with respect to
+perturbations of the training data in the following manner. We
+repeatedly draw a sub-sample (without replication) consisting of 90\%
+of the MALOVA observations and train the four models on the
+sub-sample. Each of the models produces a ranking of the 5000
+predictors according to their importance with respect to
+discriminating between the tumor types, and the agreement of these
+rankings across the different sub-samples are used to assess the
+stability of ranking the importance of the predictors. For the Random
+Forest classifier the predictors are ranked according to the Gini
+index, while for the logistic Lasso and Ridge regression models we
+order the predictors by absolute magnitude of the estimated regression
+coefficients. For the PLS-DA model the importance of the predictors is
+based on a weighted sum of the absolute coefficients where the weights
+are proportional to the reduction in the sums of squares across the
+components. To assess the stability of the individual risk predictions
+we apply the predictors from the DACOVA data set to each of the
+models. These predicted probabilities are then ranked in decreasing
+order, and the agreement of the risk predictions are similarly
+obtained by the sequential rank agreement applied across the different
+sub-samples of the training data. We use 1000 iterations for the
+sub-sampling procedure.
+
+All four models depend on tuning parameters. We estimate these by a 20
+times repeated 5-fold cross-validation on the training data in each
+iteration. The implementation of Lasso and Ridge regression in the
+\texttt{glmnet} package offers three different cross-validated
+optimization criteria for the penalty parameter: total deviance,
+classification accuracy and area under ROC. We apply all three
+criteria to our data to investigate their effect on the agreement.
 
@@
 
-Figure \ref{fig:app1} (left) shows the sequential rank agreement applied to the estimated risk probabilities of the $113$ patients in the DACOVA data set obtained from the $1000$ sub-samples of the MALOVA data, and Figure \ref{fig:app1} (right) shows the sequential rank agreement of the estimated importance of the 5000 predictors.
-For clarity we only show the agreement for the latter up to list depth 600 where all the curves have approximately become stable.
-We note that the predictor importances estimated with the Lasso model are capable of producing censored lists depending on the value of the penalty parameter.
-Under the deviance optimization criterion the median number of non-zero coefficients was 33 (range 18 to 52).
-In order to calculate the sequential rank agreement we therefore applied Algorithm \ref{sra-algorithm} with $B = 1000$ and $P=5000$.
-The effect of the censoring caused by the sparsity of the Lasso is clearly visible in the agreement curves.
+Figure \ref{fig:app1} (left) shows the sequential rank agreement
+applied to the estimated risk probabilities of the $113$ patients in
+the DACOVA data set obtained from the $1000$ sub-samples of the MALOVA
+data, and Figure \ref{fig:app1} (right) shows the sequential rank
+agreement of the estimated importance of the 5000 predictors. For
+clarity we only show the agreement for the latter up to list depth 600
+where all the curves have approximately become stable.  The Lasso
+models produce censored lists depending on the value of the penalty
+parameter. Under the deviance optimization criterion the median number
+of non-zero coefficients was 33 (range 18 to 52). In order to
+calculate the sequential rank agreement we therefore applied Algorithm
+\ref{sra-algorithm} with $B = 1000$ and $P=5000$. The effect of the
+censoring caused by the sparsity of the Lasso is clearly visible in
+the agreement curves.
 

### 2015-08-06 latex/paper.Rnw
Final draft

@@
 
+Another caveat with the way the sequential rank agreement is defined
+is the use of the standard deviation to measure agreement. The
+standard deviation is an integral part of the limits-of-agreement as
+discussed by \citet{alt:bland:1983}. However, the standard deviation
+can also be unstable when the number of observations is low and
+alternative measures such as the mean absolute deviance may prove more
+stable in some situations. However, the current definition using the
+standard deviation is analogous to the approach used for agreement in
+method comparisong studies.
+
 In conclusion we have introduced a method for evaluation of ranked

### 2015-08-07 src/sra.cpp
Median function added to cpp

@@ -32,6 +32,27 @@ double square( double x ){
     return x*x;
 }
 
+// [ [ Rcpp::export]]
+int median(NumericVector x) {
+  int xlen=x.size();
+  int n;
+  
+  if (xlen % 2) { /* x has odd length */
+    n = (xlen+1)/2;
+  } else {
+    n = xlen/2+1;
+  }
+  
+  NumericVector y = clone(x);
+  std::nth_element(y.begin(), y.begin()+n-1, y.end());
+
+  if (xlen % 2) { /* x has odd length */
+    return (y[n]);
+  } else {
+    return((y[n-1] + y[n])/2.0);
+  }
+}
+
 
 NumericVector SdPerRow(NumericMatrix obj, int type) {
 

### 2015-08-07 latex/paper.Rnw
fine tuning of application and discussion sections

@@
 from publications, or when some methods only rank a subset of the
-items. For example, penalized regression based on the Lasso provides a
+items. For example, penalized regression based on the LASSO provides a
 sparse set of predictors that have non-zero coefficients. There is no
@@
 samples from patients with either benign or malignant ovarian
-tumors. The data sets are subsamples of the Danish MALOVA and DACOVA
+tumors. The data sets are sub-samples of the Danish MALOVA and DACOVA
 study populations.
@@
 1994 to May 1999. For the purpose of illustration we use a random
-subsample of $119$ patients with a total of $58$ patients with
+sub-sample of $119$ patients with a total of $58$ patients with
 malignant ovarian cancers as cases and $61$ patients with benign
@@
 clinically suspected to be cancer during the period from 1984 to
-1990. Similarly, we use a random subsample from the DACOVA study of
+1990. Similarly, we use a random sub-sample from the DACOVA study of
 $113$ patients with a total of $54$ malignant ovarian cancers and $59$
@@
 
-Each spectrum consists of $49642$ samples over a mass-to-charge range of $800$ to $20000$ which we down-sample on an equidistant grid of 5000 points by linear interpolation.
-We preprocess each spectrum by first removing the slow-varying baseline intensity with the SNIP algorithm \citep{ryan1988snip} followed by a normalization with respect to the total ion count.
-We then standardize the 5000 predictors to have column-wise zero mean and unit variance in each data set.
-
-We use the two data sets to illustrate how the sequential rank agreement can be applied in two different scenarios.
-In the first scenario we assess the agreement of four different statistical classification methods in how they rank the predictors according to their importance with respect to the binary outcome.
-In the second scenario we assess the agreement among rankings of individual predicted risk probabilities of having a malignant tumor using the four methods.
-The first scenario is relevant in the context of biomarker discovery and the latter is important when e.g., ranking patients according to immediacy of treatment.
+Each spectrum consists of $49642$ samples over a mass-to-charge range
+of $800$ to $20000$ which we down-sample on an equidistant grid of
+5000 points by linear interpolation. After this dimension reduction
+the mass spectrum is represented by 5000 predictors. We preprocess
+%% ANDREAS: det er uklar om preprocessing forgaar i de oprindelig spectra eller i de down-samplede 
+each spectrum by first removing the slow-varying baseline intensity
+with the SNIP algorithm \citep{ryan1988snip} followed by a
+normalization with respect to the total ion count. We then standardize
+the 5000 predictors to have column-wise zero mean and unit variance in
+each data set.
+
+We use the two data sets to illustrate how the sequential rank
+agreement can be applied in two different scenarios. In the first
+scenario we assess the agreement of four different statistical
+classification methods in how they rank the predictors according to
+their importance for distinguishing benign and malignant tumors. In
+the second scenario we assess the agreement among rankings of
+individual predicted risks of having a malignant tumor. The first
+scenario is relevant in the context of biomarker discovery and the
+latter is important, e.g., to rank patients according to immediacy of
+treatment.
 
@@
 \citep{breiman2001random} implemented in the R package
-\texttt{randomForest} \citep{liaw2002classification}, logistic Lasso
+\texttt{randomForest} \citep{liaw2002classification}, logistic LASSO
 \citep{tibshirani1996regression} and Ridge regression
@@
 In both scenarios we use the MALOVA data to train the statistical
-models, and
-%the two types of
-in both situations the
-agreements are assessed with respect to
-perturbations of the training data in the following manner. We
-repeatedly draw a sub-sample (without replication) consisting of 90\%
-of the MALOVA observations and train the four models on the
-sub-sample. Each of the models produces a ranking of the 5000
-predictors according to their importance with respect to
-discriminating between the tumor types, and the agreement of these
-rankings across the different sub-samples are used to assess the
-stability of ranking the importance of the predictors. For the Random
-Forest classifier the predictors are ranked according to the Gini
-index, while for the logistic Lasso and Ridge regression models we
-order the predictors by absolute magnitude of the estimated regression
-coefficients. For the PLS-DA model the importance of the predictors is
-based on a weighted sum of the absolute coefficients where the weights
-are proportional to the reduction in the sums of squares across the
-components. To assess the stability of the individual risk predictions
-we apply the predictors from the DACOVA data set to each of the
-models. These predicted probabilities are then ranked in decreasing
-order, and the agreement of the risk predictions are similarly
-obtained by the sequential rank agreement applied across the different
-sub-samples of the training data. We use 1000 iterations for the
-sub-sampling procedure.
-
-All four models depend on tuning parameters. We estimate these by a 20
-times repeated 5-fold cross-validation on the training data in each
-iteration. The implementation of Lasso and Ridge regression in the
+models, and in both situations the agreements are assessed with
+respect to perturbations of the training data in the following
+manner. We repeatedly draw a random sub-sample (without replication)
+consisting of 90\% of the MALOVA observations and train the four
+models on each sub-sample. We use 1000 iterations for the sub-sampling
+procedure.
+
+\subsection{Agreement of predictor rankings}
+
+For each of the four methods, each of the 1000 models trained in on of
+the 1000 sub-samples of the MALOVA data produces a ranking of the 5000
+predictors according to their importance for discriminating between
+the tumor types.
+
+All four methods depend on tuning parameters. We select these
+separately for each sub-sample by a 20 times repeated 5-fold
+cross-validation.
+%% ANDREAS: hvilke tuning parameters har du optimeret for Random forest og PLS-DA?
+The implementation of LASSO and Ridge regression in the
 \texttt{glmnet} package offers three different cross-validated
@@
 criteria to our data to investigate their effect on the agreement.
+Note also that the LASSO models produce censored lists depending on
+the value of the penalty parameter. Under the deviance optimization
+criterion the median number of non-zero coefficients was 33 (range 18
+to 52). In order to calculate the sequential rank agreement for the
+LASSO we therefore applied our proposed Algorithm \ref{sra-algorithm}
+with $B = 1000$ and $P=5000$.
+
+For the Random Forest classifier the predictors are ranked according
+to the Gini index, while for the logistic LASSO and Ridge regression
+models we order the predictors by absolute magnitude of the estimated
+regression coefficients. For the PLS-DA model the importance of the
+predictors is based on a weighted sum of the absolute coefficients
+where the weights are proportional to the reduction in the sums of
+squares across the components.
+
+The right panel of Figure \ref{fig:app1} shows the sequential rank
+agreement of the estimated importance of the 5000 predictors.  For
+clarity of presentation we zoom in on the agreement up to list depth
+120. At deeper list depths all agreement curves appeared stable.
 
@@
 \end{center}
- \caption{Sequential rank agreement of predicted risk probabilities for the DACOVA data (left) and predictor importance (right) based on 1000 sub-samples of the MALOVA training data.}
+%% ANDREAS: kan du aendre titel for begge grafer:
+%% "Risk probabilities" -> "Predicted risks of malignant tumor in 113 patients (DACOVA)"
+%% "Predictor importance" -> "Discrimination importance of 5000 predictors"
+\caption{Left panel: Sequential rank agreement of 1000 rankings of the
+  predicted risks of malignant tumor. For each method the different
+  rankings were obtained by first training models in 1000 random
+  sub-samples of the MALOVA data and then predicting the risk of
+  malignant tumor in the 113 MALOVA patients. Right panel: Sequential
+  rank agreement of 1000 rankings of the 5000 predictors. The rankings
+  were obtained from the same 1000 trained models.}
  \label{fig:app1}
@@
 
-Figure \ref{fig:app1} (left) shows the sequential rank agreement
-applied to the estimated risk probabilities of the $113$ patients in
-the DACOVA data set obtained from the $1000$ sub-samples of the MALOVA
-data, and Figure \ref{fig:app1} (right) shows the sequential rank
-agreement of the estimated importance of the 5000 predictors. For
-clarity we only show the agreement for the latter up to list depth 600
-where all the curves have approximately become stable.  The Lasso
-models produce censored lists depending on the value of the penalty
-parameter. Under the deviance optimization criterion the median number
-of non-zero coefficients was 33 (range 18 to 52). In order to
-calculate the sequential rank agreement we therefore applied Algorithm
-\ref{sra-algorithm} with $B = 1000$ and $P=5000$. The effect of the
-censoring caused by the sparsity of the Lasso is clearly visible in
-the agreement curves.
-
-For most of the sequential rank agreement curves we see, as expected, that they start low followed by an increase until they approximately become constant.
-This has the interpretation that the classification methods show higher agreement with respect to the highest ranked items (the most high-risk patients and the most important predictors respectively) than for the items deeper in the lists across the different sub-samples.
-The change-points where the curves become approximately constant are the list depths where the ranks of the remaining items become close to random.
-
-The only exceptions to the expected shape of the agreement curves are the predictor importance agreements from the Ridge models and the risk prediction agreement from the Lasso model with the AUC criterion.
-They all show higher disagreement in the top of the lists followed by a decrease.
-The explanations for their behavior is different for the two methods.
-For the Lasso model the reason is that optimizing the penalty parameter with respect to the AUC criterion tends to favor a very high penalty value causing only a single predictor to be selected in each of the 1000 iterations.
-This results in a lack of generalizability to the DACOVA data which gives rise to the higher disagreement in the predicted risk probabilities.
-In the extreme case where the penalty becomes so high that none of the predictors are selected by the Lasso, the sequential rank agreement for the predicted probabilities becomes undefined since all the ranks will be ties.
+For most of the sequential rank agreement curves we see, as expected,
+that they start low, indicating good agreement, followed by an
+increase until they approximately become constant. This has the
+interpretation that the agreement across the different sub-samples is
+higher in the top as compared to the tail of the lists for all these
+classification methods. The change-points where the curves become
+approximately constant are the list depths where the ranks of the
+remaining items become close to random.
+%% ANDREAS: boer "close to random" ikke vaere ens for alle metoder?
+
+A not expected shape of the agreement curves is seen for the Ridge
+models for all three tuning criteria. They all show higher
+disagreement in the top of the lists followed by a decrease. The
+reason behind this behavior is rather subtle. Looking at the
+distribution of the regression coefficients we see that a large
+proportion of them are numerically very close to zero and have almost
+equal absolute value. This is a general feature of the Ridge models in
+this data set and seen for all the 1000 trained models.
+%% ANDREAS: er det sidste ok?
+This implies that when predictors are ranked according to the
+magnitude of their coefficients, their actual order becomes more
+uncertain and more close to a random permutation. This problem can be
+alleviated by artificially censoring all predictors with absolute
+coefficient values below a given threshold. For the Rigde models tuned
+with the deviance criterion, Figure \ref{fig:app2} (left) shows the
+the sequential rank agreement where for each of the 1000 trained
+models predictors were artificially censored when their absolute
+coefficient value was lower than the $0.1\%$ quantile of the 5000
+absolute coefficient values. The corresponding curve from Figure
+\ref{fig:app1} (left panel) is shown for comparison.
+%% ANDREAS: godkender du min aendring af de sidste 2 saetninger?
+Thus, in our application the median number of censored predictors
+across the 1000 sub-samples was 5 (range 1 to 11). Even though the
+number of censored predictors is very small compared to the total
+number of predictors, the effect on the sequential rank agreement is
+substantial and with the artificial censoring the shape of the curves
+is as expected, starting low and then increasing. Also, the effect of
+the censoring caused by the sparsity of the LASSO is clearly visible
+in the agreement curves.
+%% ANDREAS: man kan desvaerre ikke se hvor LASSO agreement forsvinder
+%% (ved list-depth 5?) og om den er laver for list depth 1-5 end rigde
+%% eller om den er kun laver for list depth 1-2. maaske kan du
+%% beskrive dette i teksten?
+
+\subsection{Agreement of individual risk predictions}
+\label{sec:airp}
+
+To assess the stability of the individual risk predictions we apply
+the predictors from the DACOVA data set to each of the models. The
+predicted probabilities are then ranked in decreasing order such the
+patients with the highest risk of a malignant tumor appears on the top
+of the list. Figure \ref{fig:app1} (left) shows the sequential rank
+agreement separately for each method, based on 1000 risk predictions
+obtained from the models trained in the same $1000$ random sub-samples
+of the MALOVA data. For clarity of presentation we zoom in on the
+agreement up to list depth 600. At deeper list depths all agreement
+curves appeared stable.
+
+Most curves start low and then increase indicating higher agreement
+among high risk patients.
+%% ANDREAS og CLAUS: er det ogsaa forvented eller skyldes det at vi kumulerer fra toppen?
+%% agreement boer jo igen vaere hoejt bland low risk patienter ...
+An exception is the risk prediction agreement for the LASSO tuned with
+the AUC criterion which shows very low agreement among the high values
+of the predicted risks. The reason is that optimizing the penalty
+parameter with respect to the AUC criterion tends to favor a very high
+penalty value causing only a single predictor to be selected in each
+of the 1000 iterations. This results in a lack of generalizability to
+the DACOVA data which gives rise to the higher disagreement in the
+predicted risk probabilities. In the extreme case where the penalty
+becomes so high that none of the predictors are selected by the LASSO,
+the sequential rank agreement for the predicted probabilities becomes
+undefined since all the ranks will be ties.
 
@@
 \end{center}
- \caption{Sequential rank agreement for Ridge regression with an artificial censoring on the absolute value of the regression coefficients (left) and box plots of AUC values across the $1000$ sub-samples with respect to the known class labels of the DACOVA data (right).}
+\caption{Left panel: Sequential rank agreement for Ridge regression
+  obtained by artificially censoring predictors when their absolute
+  coefficient values are lower than the 0.1\% quantile. Right panel:
+  Box plots of AUC values across the $1000$ sub-samples with respect
+  to the known class labels of the DACOVA data.}
  \label{fig:app2}
@@
 
-The reason behind the behavior of the sequential rank agreement of predictor importances estimated by the Ridge models is more subtle.
-Looking at the distribution of the regression coefficients it is seen that a large proportion of them are numerically very close to zero and of almost equal absolute value.
-This implies that when they are ranked according to their magnitude, their actual order becomes more uncertain and more close to a random permutation.
-This problem can be alleviate by introducing an artificial censoring depending on the absolute value of the coefficients.
-Figure \ref{fig:app2} (left) shows the sequential rank agreement curve of the rankings of the predictor importances from the Ridge model using the deviance criterion (same curve as in figure \ref{fig:app1} (left)) as well as an agreement curve where the absolute value of the regression coefficients have been censored at the $0.1\%$ quantile.
-In our application the median number of censored predictors across the 1000 sub-samples is 5 (range 1 to 11).
-Even though the number of censored predictors is very small compared to the total number of predictors, the effect on the sequential rank agreement is substantial as it has now flipped to show higher agreement in the top of the lists as was expected.
-
-Returning to Figure \ref{fig:app1}, it can further be seen that some of the methods show better agreement with respect to the predicted probabilities than for ranking the importance of the predictors and vice versa.
-Ridge regression performs very well for the risk predictions, but PLS-DA performs better for the predictor importances and similarly worse for the risk predictions.
-The Random Forest method seems to perform equally well in both scenarios.
-Lasso regression (except with the AUC criterion) performs very well at ranking the high-risk patients but very poorly at ranking the importance of the predictors.
-This reason for this is the high auto-correlation between the intensities in the mass spectra which leads to collinearity issues in the regression models.
-It is well-known that variable selection with the Lasso penalty does not preform very well when the predictors are highly correlated.
-This collinearity does, however, not affect the agreement of the risk predictions, as seen on the left part of the figure, since it is not so important which specific variable that gets selected from a group of highly correlated predictors when the purpose is risk predictions.
-
-Finally, it would appear that Ridge regression with the AUC criterion for the penalty parameter achieves the best performance with respect to the stability of ranking the individual predicted risk probabilities.
-It must, however, be stressed that the sequential rank agreement in this application is only concerned with whether similar orderings of the patients is obtained under perturbations of the training data.
-It is not concerned with whether such an ordering is the correct one.
-This is similar to the usual distinction between accuracy and precision.
-Looking at the distributions of the AUCs illustrated in Figure \ref{fig:app2} (left) across the $1000$ sub-samples for a selection of the models with respect to the known class labels of the DACOVA data, we see that PLS-DA attains the highest AUC with a median value of $0.71$ while the Ridge model with the AUC criterion attains a median AUC of $0.49$.
-This implies that while Ridge regression optimized with respect to the AUC criterion achieves the best sequential rank agreement it performs similar to a random coin toss with respect to classifying the DACOVA patients.
-Both concerns are of course important in practice and for this data set it appears that PLS-DA is the overall best method taking both agreement and classification accuracy into account.
+In Figure \ref{fig:app1} it can further be seen that some of the
+methods show better agreement with respect to the predicted
+probabilities than for ranking the importance of the predictors and
+vice versa. Ridge regression shows higher agreement across training
+sets for the risk predictions than PLS-DA, and PLS-DA shows higher
+agreement for predictor importance than Ridge regression. The Random
+Forest method seems to perform equally well in both scenarios.
+%% Andreas: jeg forstaar ikke den sidste saetning. kan man sammenligne de to rank agreements (left vs right panel)?
+LASSO shows similar agreement for ranking the risk predictions than
+PLS-DA (unless with the AUC criterion), and poor agreement for ranking
+predictors. This reason for the latter is the high auto-correlation
+between the intensities in the mass spectra which leads to
+collinearity issues in the regression models. It is well-known that
+variable selection with the LASSO penalty does not preform very well
+when the predictors are highly correlated. The collinearity does,
+however, not affect the agreement of the risk predictions, as seen on
+the left panel of Figure \ref{fig:app1}, since it is not so important
+which specific variable is selected from a group of highly correlated
+predictors when the purpose is risk predictions.
+
+It appears that Ridge regression tuned with the AUC criterion achieves
+the best performance with respect to the stability of ranking the
+individual predicted risk probabilities. It must, however, be stressed
+that the sequential rank agreement in this application is only
+concerned with the agreement of the risk predictions across
+sub-samples and not with the actual accuracy of the risk
+predictions. Thus, we also computed the AUC for the different models
+based on the DACOVA data. The distributions across the $1000$
+sub-samples for a selection of the models is shown in the right panel
+of Figure \ref{fig:app2}.  Here we see that PLS-DA attains the highest
+AUC values with a median value of $0.71$ while the Ridge model with
+the AUC criterion attains a median AUC of $0.49$. This implies that
+while Ridge regression optimized with respect to the AUC criterion
+achieves the best sequential rank agreement it performs similar to a
+random coin toss with respect to classifying the DACOVA patients.
+%% and for this data set it appears that PLS-DA is the overall best
+%% method taking both agreement and discrimination ability into account.
+%% Andreas: vi boer ogsaa se paa Brier score (calibration) inden vi konkluderer at PLS-DA er den bedste
 
 \section{Discussion}
+
 In this article we address the problem of comparing ranked lists of
@@
 and the situation where the only available data is the actual ranked
-lists. In addition, censored ranked lists where only the top $k$
-ranked items are known can be accommodated as well. The proposed
-agreement measure can be interpreted as the average distance between
-an item's rank and the average rank assigned to that item and thus
-determine how well the lists agree on the ranks of a specific
-item and it circumvents the problem with other existing similarity measures.
+lists. In addition, censored ranked lists where only the ranks of the
+top $k$ ranked items are known can be accommodated as well. The
+proposed agreement measure can be interpreted as the average distance
+between an item's rank and the average rank assigned to that item
+across lists. Thus the measure determines how well the lists agree on
+the ranks of a specific item.
 
@@
 data as well as to evaluate the stability of the ranks from a single
-method by bootstrapping the data repeatedly and comparing the ranks
-obtained from the bootstrapped data.
+method by bootstrapping (or sub-sampling) the data repeatedly and
+comparing the ranks obtained from training the models in the
+bootstrapped data.
 
@@
 from the lists as possible, i.e., we have used all available ranks.
+%% Claus: hvad vil du sige med den sidste saetning?
 We could choose to censor the rankings of all the items that were
@@
 non-significant items and it would be easier to identify a change in
-agreement among the items that were deemed to relevant.
+agreement among the items that were deemed to relevant. In our
+application section we have successfully introduced such an artificial
+censoring for the predictor rankings obtained with ridge regression.
 
@@
 assign similar ranks to two highly correlated predictors while methods
-that provide a sparse solution such as the lasso will just rank one of
-the two predictors high while the other might have a very low
-rank. When the rankings from these two analysis methods are compared
-then the agreement will differ simply because of the way correlated
+that provide a sparse solution such as the LASSO will just rank one of
+the two predictors high while the other might have a very low rank.
+Thus in such a scenario we would expect low agreement of the rankings
+from LASSO and marginal analyses simply because of the way correlated
 predictors are handled. This is not a shortcoming of the sequential
@@
 standard deviation is analogous to the approach used for agreement in
-method comparisong studies.
-
-In conclusion we have introduced a method for evaluation of ranked
-(censored) lists that can be easily interpreted and that can be
-applied to a large number of situations.  The method can be adapted
-even further by using it to identify and cluster statistical analysis
-methods that provide more or less the same agreement, or by using the
-rank agreement to optimize a hyper-parameter in, say, elastic net regularized
-regression where the rank agreement is used to determine the mixing
-proportion between the $L_1$ and the $L_2$ penalty. Finally, the
-proposed method may be adapted (with some additional assumptions) to
-the situation where the lists are subsets of each other. We will be
-investigating these extensions further in the future.
+method comparison studies.
+
+%% In conclusion we have introduced a method for evaluation of ranked
+%% (censored) lists that can be easily interpreted and that can be
+%% applied to a large number of situations.
+%% CLAUS: den sidste saetning er bar en omformulering af Diskussionen's foerste saetning.
+
+%% The method here proposed can be adapted further by using it to
+%% identify and cluster statistical analysis methods that provide more or
+%% less the same agreement,
+%% Claus: det er enten for kort eller for lang
+
+%% or by using the rank agreement to optimize a
+%% hyper-parameter in, say, elastic net regularized regression where the
+%% rank agreement is used to determine the mixing proportion between the
+%% $L_1$ and the $L_2$ penalty.
+%% Claus: det er enten for kort eller for lang
+
+%% Finally, the proposed method may be
+%% adapted (with some additional assumptions) to the situation where the
+%% lists are subsets of each other.
+%% Claus: det er enten for kort eller for lang
+
+%% We will be investigating these
+%% extensions further in the future.
 

### 2015-08-07 latex/paper.Rnw
Fixing a comment
Answering some comments

@@
 
-Each spectrum consists of $49642$ samples over a mass-to-charge range
-of $800$ to $20000$ which we down-sample on an equidistant grid of
-5000 points by linear interpolation. After this dimension reduction
-the mass spectrum is represented by 5000 predictors. We preprocess
-%% ANDREAS: det er uklar om preprocessing forgaar i de oprindelig spectra eller i de down-samplede 
-each spectrum by first removing the slow-varying baseline intensity
+Each spectrum consists of $49642$ samples over a range of mass-to-charge ratios 
+between $800$ to $20000$ Dalton which we downsample on an equidistant grid of
+5000 points by linear interpolation. We then preprocess the downsampled 
+spectra individually by first removing the slow-varying baseline intensity
 with the SNIP algorithm \citep{ryan1988snip} followed by a
@@
 
-All four methods depend on tuning parameters. We select these
+All four methods depend on a tuning parameter. We select these
 separately for each sub-sample by a 20 times repeated 5-fold
-cross-validation.
-%% ANDREAS: hvilke tuning parameters har du optimeret for Random forest og PLS-DA?
+cross-validation procedure. For Lasso and Ridge regression the tuning parameter
+is the degree of penalization, for PLS-DA it is the number of components 
+(the dimensionality of the subspace), and for Random Forest it is the number
+ of predictors randomly sampled at each split.
+
 The implementation of LASSO and Ridge regression in the
@@
 \end{center}
-%% ANDREAS: kan du aendre titel for begge grafer:
-%% "Risk probabilities" -> "Predicted risks of malignant tumor in 113 patients (DACOVA)"
-%% "Predictor importance" -> "Discrimination importance of 5000 predictors"
 \caption{Left panel: Sequential rank agreement of 1000 rankings of the
@@
 alleviated by artificially censoring all predictors with absolute
-coefficient values below a given threshold. For the Rigde models tuned
+coefficient values below a given threshold. For the Ridge models tuned
 with the deviance criterion, Figure \ref{fig:app2} (left) shows the
@@
 \section{Discussion}
-
 In this article we address the problem of comparing ranked lists of

### 2015-08-08 latex/applicationscripts/mergeFiles.R
Adding R scripts for application

@@ -1,0 +1,12 @@
+res <- NULL
+for(i in 1:1500) {
+  filename <- paste('SuperFinal-', i, '.RData', sep="")
+  if(file.exists(filename)) {
+    load(file=filename)
+    res <- c(res, list(out))
+  } else {
+    cat(filename, " doesn't exist\n")
+  }
+}
+
+save(res, file="joinedOutput.RData")

### 2015-08-08 latex/applicationscripts/plots.R
Adding more application scripts

@@ -1,0 +1,113 @@
+library(SuperRanker)
+
+rm(list=ls())
+load("joinedOutput.RData")
+load("DACOVAlabels.RData")
+
+B <- 2 #number of randomization for censored lists
+
+
+########## PREPARE SRA OBJECTS ##########
+rfRisk <- sra(lapply(res, function(a) a$rf$riskOrder))
+plsRisk <- sra(lapply(res, function(a) a$pls$riskOrder))
+
+lassoDevRisk <- sra(lapply(res, function(a) a$lassoDev$riskOrder))
+lassoClassRisk <- sra(lapply(res, function(a) a$lassoClass$riskOrder))
+lassoAucRisk <- sra(lapply(res, function(a) a$lassoAUC$riskOrder))
+
+ridgeDevRisk <- sra(lapply(res, function(a) a$ridgeDev$riskOrder))
+ridgeClassRisk <- sra(lapply(res, function(a) a$ridgeClass$riskOrder))
+ridgeAucRisk <- sra(lapply(res, function(a) a$ridgeAUC$riskOrder))
+
+
+rfImpSRA <- sra(lapply(res, function(a) a$rf$importance), nitems=5000, B=B)
+plsImpSRA <- sra(lapply(res, function(a) a$pls$importance))
+
+lassoDevImpSRA <- sra(lapply(res, function(a) a$lassoDev$importance), nitems=5000, B=B)
+lassoClassImpSRA <- sra(lapply(res, function(a) a$lassoClass$importance), nitems=5000, B=B)
+lassoAucImpSRA <- sra(lapply(res, function(a) a$lassoAUC$importance), nitems=5000, B=B)
+
+ridgeDevImpSRA <- sra(lapply(res, function(a) a$ridgeDev$importance))
+ridgeClassImpSRA <- sra(lapply(res, function(a) a$ridgeClass$importance))
+ridgeAucImpSRA <- sra(lapply(res, function(a) a$ridgeAUC$importance))
+
+
+########## RISK PLOT ##########
+plot(rfRisk, ylim=c(0,30), xlim=c(0,120))
+title("Predicted risks of malignant tumor in 113 patients (DACOVA)")
+
+plot(lassoDevRisk, add=TRUE, lines.col="red")
+plot(lassoClassRisk, add=TRUE, lines.col="red", lines.lty=2)
+plot(lassoAucRisk, add=TRUE, lines.col="red", lines.lty=3)
+
+plot(plsRisk, add=TRUE, lines.col="blue")
+
[…]

### 2015-08-08 latex/paper.Rnw
Fixing comments

@@
 from publications, or when some methods only rank a subset of the
-items. For example, penalized regression based on the LASSO provides a
+items. For example, penalized regression based on the Lasso provides a
 sparse set of predictors that have non-zero coefficients. There is no
@@
 \citep{breiman2001random} implemented in the R package
-\texttt{randomForest} \citep{liaw2002classification}, logistic LASSO
+\texttt{randomForest} \citep{liaw2002classification}, logistic Lasso
 \citep{tibshirani1996regression} and Ridge regression
@@
 
+All four methods depend on a tuning parameter. We select these
+separately for each sub-sample by a 20 times repeated 5-fold
+cross-validation procedure. The tuning parameter for Lasso and Ridge regression 
+is the degree of penalization, for PLS-DA it is the number of components 
+(the dimensionality of the subspace), and for Random Forest it is the number
+ of predictors randomly sampled at each split.
+The implementation of Lasso and Ridge regression in the
+\texttt{glmnet} package offers three different cross-validated
+optimization criteria for the penalty parameter: total deviance,
+classification accuracy and area under ROC. We apply all three
+criteria to our data to investigate their effect on the agreements.
+Note also that the Lasso models produce censored lists depending on
+the value of the penalty parameter.
+
 \subsection{Agreement of predictor rankings}
-
 For each of the four methods, each of the 1000 models trained in on of
@@
 predictors according to their importance for discriminating between
-the tumor types.
-
-All four methods depend on a tuning parameter. We select these
-separately for each sub-sample by a 20 times repeated 5-fold
-cross-validation procedure. For Lasso and Ridge regression the tuning parameter
-is the degree of penalization, for PLS-DA it is the number of components 
-(the dimensionality of the subspace), and for Random Forest it is the number
- of predictors randomly sampled at each split.
-
-The implementation of LASSO and Ridge regression in the
-\texttt{glmnet} package offers three different cross-validated
-optimization criteria for the penalty parameter: total deviance,
-classification accuracy and area under ROC. We apply all three
-criteria to our data to investigate their effect on the agreement.
-Note also that the LASSO models produce censored lists depending on
-the value of the penalty parameter. Under the deviance optimization
-criterion the median number of non-zero coefficients was 33 (range 18
-to 52). In order to calculate the sequential rank agreement for the
-LASSO we therefore applied our proposed Algorithm \ref{sra-algorithm}
-with $B = 1000$ and $P=5000$.
-
-For the Random Forest classifier the predictors are ranked according
-to the Gini index, while for the logistic LASSO and Ridge regression
+the tumor types. For the Random Forest classifier the predictors are ranked according
+to the Gini index, while for the logistic Lasso and Ridge regression
 models we order the predictors by absolute magnitude of the estimated
@@
 clarity of presentation we zoom in on the agreement up to list depth
-120. At deeper list depths all agreement curves appeared stable.
+600. At deeper list depths all agreement curves become approximately constant.
 
@@
 approximately constant are the list depths where the ranks of the
-remaining items become close to random.
-%% ANDREAS: boer "close to random" ikke vaere ens for alle metoder?
+remaining items become close to uniformly random.
 
@@
 this data set and seen for all the 1000 trained models.
-%% ANDREAS: er det sidste ok?
 This implies that when predictors are ranked according to the
@@
 \ref{fig:app1} (left panel) is shown for comparison.
-%% ANDREAS: godkender du min aendring af de sidste 2 saetninger?
+%%%% CHECK UP: CENSOR ON POOLED DATA OR CENSOR BY INVIDIAL THRESHOLD
+
+%%%%%% DET HER ER NOGET ROD. DER ER FORSVUNDET NOGET LASSO TEKST
+%%%%%% GÅ TILBAGE OG FIND DET IGEN
+Under the deviance optimization criterion the median number of non-zero 
+coefficients was 33 (range 18 to 52). In order to calculate the sequential 
+rank agreement for the Lasso we therefore applied our proposed Algorithm \ref{sra-algorithm}
+with $B = 1000$ and $P=5000$.
+
 Thus, in our application the median number of censored predictors
@@
 is as expected, starting low and then increasing. Also, the effect of
-the censoring caused by the sparsity of the LASSO is clearly visible
+the censoring caused by the sparsity of the Lasso is clearly visible
 in the agreement curves.
-%% ANDREAS: man kan desvaerre ikke se hvor LASSO agreement forsvinder
-%% (ved list-depth 5?) og om den er laver for list depth 1-5 end rigde
-%% eller om den er kun laver for list depth 1-2. maaske kan du
-%% beskrive dette i teksten?
+\textbf{ANDREAS: man kan desvaerre ikke se hvor Lasso agreement forsvinder
+  (ved list-depth 5?) og om den er laver for list depth 1-5 end rigde
+  eller om den er kun laver for list depth 1-2. maaske kan du
+  beskrive dette i teksten?}
 
@@
 \label{sec:airp}
-
 To assess the stability of the individual risk predictions we apply
 the predictors from the DACOVA data set to each of the models. The
-predicted probabilities are then ranked in decreasing order such the
-patients with the highest risk of a malignant tumor appears on the top
+predicted probabilities are then ranked in decreasing order such that the
+patients with the highest risk of a malignant tumor appears in the top
 of the list. Figure \ref{fig:app1} (left) shows the sequential rank
-agreement separately for each method, based on 1000 risk predictions
+agreement separately for each method, based on the 1000 risk predictions
 obtained from the models trained in the same $1000$ random sub-samples
-of the MALOVA data. For clarity of presentation we zoom in on the
-agreement up to list depth 600. At deeper list depths all agreement
-curves appeared stable.
+of the MALOVA data.
 
@@
 among high risk patients.
-%% ANDREAS og CLAUS: er det ogsaa forvented eller skyldes det at vi kumulerer fra toppen?
-%% agreement boer jo igen vaere hoejt bland low risk patienter ...
-An exception is the risk prediction agreement for the LASSO tuned with
+\textbf{ANDREAS og CLAUS: er det ogsaa forvented eller skyldes det at vi kumulerer fra toppen? agreement boer jo igen vaere hoejt bland low risk patienter ... Claus indsætter måske noget her!}
+An exception is the risk prediction agreement for the Lasso tuned with
 the AUC criterion which shows very low agreement among the high values
@@
 predicted risk probabilities. In the extreme case where the penalty
-becomes so high that none of the predictors are selected by the LASSO,
+becomes so high that none of the predictors are selected by the Lasso,
 the sequential rank agreement for the predicted probabilities becomes
@@
 Forest method seems to perform equally well in both scenarios.
-%% Andreas: jeg forstaar ikke den sidste saetning. kan man sammenligne de to rank agreements (left vs right panel)?
-LASSO shows similar agreement for ranking the risk predictions than
+\textbf{Andreas: jeg forstaar ikke den sidste saetning. kan man sammenligne de to rank agreements (left vs right panel)?}
+Lasso shows similar agreement for ranking the risk predictions than
 PLS-DA (unless with the AUC criterion), and poor agreement for ranking
@@
 collinearity issues in the regression models. It is well-known that
-variable selection with the LASSO penalty does not preform very well
+variable selection with the Lasso penalty does not preform very well
 when the predictors are highly correlated. The collinearity does,
@@
 random coin toss with respect to classifying the DACOVA patients.
-%% and for this data set it appears that PLS-DA is the overall best
-%% method taking both agreement and discrimination ability into account.
-%% Andreas: vi boer ogsaa se paa Brier score (calibration) inden vi konkluderer at PLS-DA er den bedste
+\textbf{and for this data set it appears that PLS-DA is the overall best
+  method taking both agreement and discrimination ability into account.
+  Andreas: vi boer ogsaa se paa Brier score (calibration) inden vi konkluderer at PLS-DA er den bedste
+}
 
@@
 assign similar ranks to two highly correlated predictors while methods
-that provide a sparse solution such as the LASSO will just rank one of
+that provide a sparse solution such as the Lasso will just rank one of
 the two predictors high while the other might have a very low rank.
 Thus in such a scenario we would expect low agreement of the rankings
-from LASSO and marginal analyses simply because of the way correlated
+from Lasso and marginal analyses simply because of the way correlated
 predictors are handled. This is not a shortcoming of the sequential

### 2015-08-08 latex/paper.Rnw
More comment fixing

@@
 \ref{fig:app1} (left panel) is shown for comparison.
-%%%% CHECK UP: CENSOR ON POOLED DATA OR CENSOR BY INVIDIAL THRESHOLD
-
-%%%%%% DET HER ER NOGET ROD. DER ER FORSVUNDET NOGET LASSO TEKST
-%%%%%% GÅ TILBAGE OG FIND DET IGEN
+\textbf{CHECK UP: CENSOR ON POOLED DATA OR CENSOR BY INDIVIDUAL THRESHOLD}
+
+\textbf{DET HER ER NOGET VÆRRE ROD. DER ER FORSVUNDET NOGET LASSO TEKST. GÅ TILBAGE OG FIND DET IGEN}
 Under the deviance optimization criterion the median number of non-zero 

### 2015-08-09 latex/paper.Rnw
Comments

--- 2015-08-07 latex/paper.Rnw
@@
 
-Each spectrum consists of $49642$ samples over a range of mass-to-charge ratios 
-between $800$ to $20000$ Dalton which we downsample on an equidistant grid of
-5000 points by linear interpolation. We then preprocess the downsampled 
-spectra individually by first removing the slow-varying baseline intensity
+Each spectrum consists of $49642$ samples over a mass-to-charge range
+of $800$ to $20000$ which we down-sample on an equidistant grid of
+5000 points by linear interpolation. After this dimension reduction
+the mass spectrum is represented by 5000 predictors. We preprocess
+%% ANDREAS: det er uklar om preprocessing forgaar i de oprindelig spectra eller i de down-samplede 
+each spectrum by first removing the slow-varying baseline intensity
 with the SNIP algorithm \citep{ryan1988snip} followed by a
@@
 
-All four methods depend on a tuning parameter. We select these
+All four methods depend on tuning parameters. We select these
 separately for each sub-sample by a 20 times repeated 5-fold
-cross-validation procedure. For Lasso and Ridge regression the tuning parameter
-is the degree of penalization, for PLS-DA it is the number of components 
-(the dimensionality of the subspace), and for Random Forest it is the number
- of predictors randomly sampled at each split.
-
+cross-validation.
+%% ANDREAS: hvilke tuning parameters har du optimeret for Random forest og PLS-DA?
 The implementation of LASSO and Ridge regression in the
@@
 \end{center}
+%% ANDREAS: kan du aendre titel for begge grafer:
+%% "Risk probabilities" -> "Predicted risks of malignant tumor in 113 patients (DACOVA)"
+%% "Predictor importance" -> "Discrimination importance of 5000 predictors"
 \caption{Left panel: Sequential rank agreement of 1000 rankings of the
@@
 alleviated by artificially censoring all predictors with absolute
-coefficient values below a given threshold. For the Ridge models tuned
+coefficient values below a given threshold. For the Rigde models tuned
 with the deviance criterion, Figure \ref{fig:app2} (left) shows the
@@
 %% agreement boer jo igen vaere hoejt bland low risk patienter ...
+
 An exception is the risk prediction agreement for the LASSO tuned with
@@
 \section{Discussion}
+
 In this article we address the problem of comparing ranked lists of

### 2015-08-09 latex/paper.Rnw
Comments

--- 2015-08-08 latex/paper.Rnw
@@
 Most curves start low and then increase indicating higher agreement
-among high risk patients.
-\textbf{ANDREAS og CLAUS: er det ogsaa forvented eller skyldes det at vi kumulerer fra toppen? agreement boer jo igen vaere hoejt bland low risk patienter ... Claus indsætter måske noget her!}
-An exception is the risk prediction agreement for the Lasso tuned with
-the AUC criterion which shows very low agreement among the high values
-of the predicted risks. The reason is that optimizing the penalty
-parameter with respect to the AUC criterion tends to favor a very high
-penalty value causing only a single predictor to be selected in each
-of the 1000 iterations. This results in a lack of generalizability to
-the DACOVA data which gives rise to the higher disagreement in the
-predicted risk probabilities. In the extreme case where the penalty
-becomes so high that none of the predictors are selected by the Lasso,
-the sequential rank agreement for the predicted probabilities becomes
-undefined since all the ranks will be ties.
+among high risk patients. This is expected if we rank the individuals
+according to highest risk of disease. However, it is also expected
+that individuals with very low risk also show high agreement. In this
+case we order the patients according to (high) risk prediction but we
+could essentially also have reversed the order to identify the
+patients that have low risk prediction.  An exception is the risk
+prediction agreement for the Lasso tuned with the AUC criterion which
+shows very low agreement among the high values of the predicted
+risks. The reason is that optimizing the penalty parameter with
+respect to the AUC criterion tends to favor a very high penalty value
+causing only a single predictor to be selected in each of the 1000
+iterations. This results in a lack of generalizability to the DACOVA
+data which gives rise to the higher disagreement in the predicted risk
+probabilities. In the extreme case where the penalty becomes so high
+that none of the predictors are selected by the Lasso, the sequential
+rank agreement for the predicted probabilities becomes undefined since
+all the ranks will be ties.
 

### 2015-08-09 latex/paper.Rnw
Kommentarer

@@
 
-Finally, we have --- whenever possible --- used as much information
-from the lists as possible, i.e., we have used all available ranks.
-%% Claus: hvad vil du sige med den sidste saetning?
-We could choose to censor the rankings of all the items that were
-deemed not to be significant if information on both rankings and the
-corresponding evidence for significance was available. That would
-ensure that there would be put less emphasis on the agreement of the
-non-significant items and it would be easier to identify a change in
-agreement among the items that were deemed to relevant. In our
-application section we have successfully introduced such an artificial
-censoring for the predictor rankings obtained with ridge regression.
+Finally, we have --- whenever possible --- used all available ranks
+from the lists.  We could choose to censor the rankings of all the
+items that were deemed not to be significant if information on both
+rankings and the corresponding evidence for significance was
+available. That would ensure that there would be put less emphasis on
+the agreement of the non-significant items and it would be easier to
+identify a change in agreement among the items that were deemed to
+relevant. In our application section we have successfully introduced
+such an artificial censoring for the predictor rankings obtained with
+ridge regression.
 
@@
 
-%% In conclusion we have introduced a method for evaluation of ranked
-%% (censored) lists that can be easily interpreted and that can be
-%% applied to a large number of situations.
-%% CLAUS: den sidste saetning er bar en omformulering af Diskussionen's foerste saetning.
-
-%% The method here proposed can be adapted further by using it to
-%% identify and cluster statistical analysis methods that provide more or
-%% less the same agreement,
-%% Claus: det er enten for kort eller for lang
-
-%% or by using the rank agreement to optimize a
-%% hyper-parameter in, say, elastic net regularized regression where the
-%% rank agreement is used to determine the mixing proportion between the
-%% $L_1$ and the $L_2$ penalty.
-%% Claus: det er enten for kort eller for lang
-
-%% Finally, the proposed method may be
-%% adapted (with some additional assumptions) to the situation where the
-%% lists are subsets of each other.
-%% Claus: det er enten for kort eller for lang
-
-%% We will be investigating these
-%% extensions further in the future.
+In conclusion we have introduced a method for evaluation of ranked
+(censored) lists that can be easily interpreted and that can be
+applied to a large number of situations.  The method presented here
+can be adapted further by using it to compare and classify statistical
+analysis methods that agree on the rankings they provide or by using
+the rank agreement to optimize a hyper-parameter in, say, elastic net
+regularized regression where the rank agreement is used to determine
+the mixing proportion between the $L_1$ and the $L_2$ penalty.
+%Finally, the proposed method may be adapted (with some additional
+%assumptions) to the situation where the lists are subsets of each
+%other.  
+We will be investigating these extensions further in the future.
 

### 2015-08-09 latex/paper.Rnw
More comment fixing

@@
 with the SNIP algorithm \citep{ryan1988snip} followed by a
-normalization with respect to the total ion count. We then standardize
-the 5000 predictors to have column-wise zero mean and unit variance in
+normalization with respect to the total ion count. Finally, we standardize
+the $5000$ predictors to have column-wise zero mean and unit variance in
 each data set.
@@
 scenario is relevant in the context of biomarker discovery and the
-latter is important, e.g., to rank patients according to immediacy of
+latter is important e.g., when ranking patients according to immediacy of
 treatment.
@@
 
-All four methods depend on a tuning parameter. We select these
-separately for each sub-sample by a 20 times repeated 5-fold
-cross-validation procedure. The tuning parameter for Lasso and Ridge regression
-is the degree of penalization, for PLS-DA it is the number of components
-(the dimensionality of the subspace), and for Random Forest it is the number
- of predictors randomly sampled at each split.
+All four methods depend on a tuning parameter. The tuning parameter for 
+Lasso and Ridge regression is the degree of penalization, and for PLS-DA 
+it is the number of components (the dimensionality of the subspace). 
+We estimate these separately for each sub-sample by a 20 times repeated 
+5-fold cross-validation procedure. For the Random Forest we grow a fixed
+number of $5000$ trees and let the tuning parameter be the number of 
+predictors randomly sampled at each split. We estimate this by a binary
+search with respect to minimizing the Out-of-Bag classification error estimate.
+
 The implementation of Lasso and Ridge regression in the
@@
 \subsection{Agreement of predictor rankings}
-For each of the four methods, each of the 1000 models trained in on of
+For each of the four methods, each of the 1000 models trained on 
 the 1000 sub-samples of the MALOVA data produces a ranking of the 5000
@@
 to the Gini index, while for the logistic Lasso and Ridge regression
-models we order the predictors by absolute magnitude of the estimated
+models we order by absolute magnitude of the estimated
 regression coefficients. For the PLS-DA model the importance of the
@@
 clarity of presentation we zoom in on the agreement up to list depth
-600. At deeper list depths all agreement curves become approximately constant.
+600. At deeper list depths all agreement curves are approximately constant.
 
@@
   sub-samples of the MALOVA data and then predicting the risk of
-  malignant tumor in the 113 MALOVA patients. Right panel: Sequential
+  malignant tumor in the 113 DACOVA patients. Right panel: Sequential
   rank agreement of 1000 rankings of the 5000 predictors. The rankings
@@
 iterations. This results in a lack of generalizability to the DACOVA
-data which gives rise to the higher disagreement in the predicted risk
-probabilities. In the extreme case where the penalty becomes so high
+data which gives rise to the higher disagreement in the predicted risks. 
+In the extreme case where the penalty becomes so high
 that none of the predictors are selected by the Lasso, the sequential
@@
 
-In Figure \ref{fig:app1} it can further be seen that some of the
-methods show better agreement with respect to the predicted
-probabilities than for ranking the importance of the predictors and
-vice versa. Ridge regression shows higher agreement across training
+Comparing the left and right panels of Figure \ref{fig:app1} it can 
+further be seen that some of the methods show better agreement with 
+respect to the predicted probabilities than for ranking the importance 
+of the predictors and vice versa. Ridge regression shows higher agreement across training
 sets for the risk predictions than PLS-DA, and PLS-DA shows higher
-agreement for predictor importance than Ridge regression. The Random
-Forest method seems to perform equally well in both scenarios.
-\textbf{Andreas: jeg forstaar ikke den sidste saetning. kan man sammenligne de to rank agreements (left vs right panel)?}
+agreement for predictor importance than Ridge regression.
+
 Lasso shows similar agreement for ranking the risk predictions than
@@
 sub-samples and not with the actual accuracy of the risk
-predictions. Thus, we also computed the AUC for the different models
+predictions. Thus, we also computed the AUC and Brier scores for the different models
 based on the DACOVA data. The distributions across the $1000$
@@
 random coin toss with respect to classifying the DACOVA patients.
-\textbf{and for this data set it appears that PLS-DA is the overall best
-  method taking both agreement and discrimination ability into account.
-  Andreas: vi boer ogsaa se paa Brier score (calibration) inden vi konkluderer at PLS-DA er den bedste
-}
+\textbf{Comment on Brier scores}
+Both concerns are of course important in practice and for this data set 
+it appears thatPLS-DA is the overall best method taking both agreement 
+and classification accuracy into account.
 

### 2015-08-10 latex/applicationscripts/mergeFiles.R
Revising application yet again

@@ -1,3 +1,5 @@
+rm(list=ls())
+
 res <- NULL
 for(i in 1:1500) {
   filename <- paste('SuperFinal-', i, '.RData', sep="")

### 2015-08-10 latex/applicationscripts/plots.R

@@ -3,11 +3,11 @@ library(SuperRanker)
 rm(list=ls())
 load("joinedOutput.RData")
 load("DACOVAlabels.RData")
+res <- res[1:1000]
 
-B <- 2 #number of randomization for censored lists
+B <- 1000 #number of randomization for censored lists
 
-
-########## PREPARE SRA OBJECTS ##########
+########## PREPARE OBJECTS ##########
 rfRisk <- sra(lapply(res, function(a) a$rf$riskOrder))
 plsRisk <- sra(lapply(res, function(a) a$pls$riskOrder))
 
@@ -19,7 +19,7 @@ ridgeDevRisk <- sra(lapply(res, function(a) a$ridgeDev$riskOrder))
 ridgeClassRisk <- sra(lapply(res, function(a) a$ridgeClass$riskOrder))
 ridgeAucRisk <- sra(lapply(res, function(a) a$ridgeAUC$riskOrder))
 
-
+##
 rfImpSRA <- sra(lapply(res, function(a) a$rf$importance), nitems=5000, B=B)
 plsImpSRA <- sra(lapply(res, function(a) a$pls$importance))
 
@@ -30,6 +30,24 @@ lassoAucImpSRA <- sra(lapply(res, function(a) a$lassoAUC$importance), nitems=500
 ridgeDevImpSRA <- sra(lapply(res, function(a) a$ridgeDev$importance))
 ridgeClassImpSRA <- sra(lapply(res, function(a) a$ridgeClass$importance))
 ridgeAucImpSRA <- sra(lapply(res, function(a) a$ridgeAUC$importance))
+
+##
+aucMat <- cbind(sapply(res, function(a) a$rf$auc),
+                sapply(res, function(a) a$pls$auc),
+                sapply(res, function(a) a$lassoDev$auc),
+                sapply(res, function(a) a$ridgeDev$auc),
+                sapply(res, function(a) a$ridgeAUC$auc))
+colnames(aucMat) <- c("Random Forest", "PLS-DA", "Lasso deviance", "Ridge deviance", "Ridge AUC")
+
+brierMat <- cbind(sapply(res, function(a) mean((a$rf$risk - dacovaLabels)^2)),
+                  sapply(res, function(a) mean((a$pls$risk - dacovaLabels)^2)),
[…]

### 2015-08-10 latex/applicationscripts/plots.R
More application revisions

@@ -92,62 +92,18 @@ abline(h=0.5, lty=3)
 
 
 ########## RIDGE CENSORING ##########
-cutOff <- quantile(abs(unlist(out)), 0.001)
-outOrderCens <- lapply(out, function(a) {
-  b <- abs(a)
-  b <- b[b > cutOff]
+ridgeDevCens <- lapply(res, function(a) {
+  b <- abs(a$ridgeDev$coef)
+  b <- b[b > quantile(b, 0.001)]
   order(b, decreasing=TRUE)	
 })
+ridgeDevCencRisk <- sra(ridgeDevCens, nitems=5000, B=B)
 
-median(sapply(outOrderCens, function(a) 5000 - length(a))) #median number of censorings
-range(sapply(outOrderCens, function(a) 5000 - length(a)))
+plot(ridgeDevImpSRA, ylim=c(0,1500), xlim=c(0,600))
+plot(ridgeDevCencRisk, add=TRUE, lines.col="red")
+title("Discrimination importance for Ridge with thresholding")
+legend("topright", c("No censoring", "Censored at 0.1% quantile"), col=c(1,2), lwd=3, lty=1, bty="n")
 
-
-
-
-
-
-
-
-
-boxplot(brierMat)
-
-boxplot(cbind(aucMat[,1], brierMat[,1],
-              aucMat[,2], brierMat[,2], 
-              aucMat[,3], brierMat[,3],
-              aucMat[,4], brierMat[,4],
-              aucMat[,5], brierMat[,5]),
-              at=c(1,1.5, 4,4.5, 7,7.5, 10,10.5, 13,13.5), ylim=c(0.4,0.8))
-abline(h=0.5, lty=3)
-
-
-
-boxplot(aucMat, ylab="AUC", ylim=c(0.4,0.8), axes=FALSE, main="AUC distribution of DACOVA predictions")
-axis(2, at=seq(0.4,0.8,0.1))
-axis(1, at=1:5, labels=colnames(aucMat), cex.axis=0.8)
-abline(h=0.5, lty=3)
-
-
-########## RIDGE CENSORING ##########
-cutOff <- quantile(abs(unlist(out)), 0.001)
-outOrderCens <- lapply(out, function(a) {
-  b <- abs(a)
-  b <- b[b > cutOff]
-  order(b, decreasing=TRUE)	
-})
-
-median(sapply(outOrderCens, function(a) 5000 - length(a))) #median number of censorings
[…]

### 2015-08-10 latex/paper.Rnw
final application revision

@@
 \begin{equation}
-  \tilde R_l(X_p) = \left\{\begin{array}{cl} \{R_l(X_p)\} & \text{ for } X_p\in \Lambda_l \\
+  \widetilde R_l(X_p) = \left\{\begin{array}{cl} \{R_l(X_p)\} & \text{ for } X_p\in \Lambda_l \\
       \{d_l+1,\dots,P\} & \text{ for } X_p \not\in \Lambda_l\end{array}\right.
@@
 \begin{eqnarray*}
-\tilde H_0 & :&  \text{The list rankings are based on data containing}\\
+\widetilde H_0 & :&  \text{The list rankings are based on data containing}\\
 & &   \text{no association to the outcome.}
@@
 
-To obtain the distribution under $\tilde H_0$ the idea is to repeat
+To obtain the distribution under $\widetilde H_0$ the idea is to repeat
 many times the ranking procedures in uncorrelated data.  Thus, we
@@
 approach requires that we have the original data available and as such
-it may not be possible to evaluate $\tilde H_0$ in all situations.
+it may not be possible to evaluate $\widetilde H_0$ in all situations.
 
@@
 substantially below the distribution of the sequential rank agreements
-obtained under either $H_0$ or $\tilde H_0$ then this suggests that
+obtained under either $H_0$ or $\widetilde H_0$ then this suggests that
 the original ranked lists agree \emph{more} than what we would expect
@@
 distributions of sequential rank agreement under $H_0$ and
-$\tilde H_0$ each based on $400$ permutations. Not surprisingly, the
-sequential rank agreement under $\tilde H_0$ is lower than the
+$\wildetilde H_0$ each based on $400$ permutations. Not surprisingly, the
+sequential rank agreement under $\widetilde H_0$ is lower than the
 sequential rank agreement under $H_0$ because the four methods used to
@@
 
-It is important to stress that neither $H_0$ nor $\tilde H_0$ are
+It is important to stress that neither $H_0$ nor $\widetilde H_0$ are
 related to questions regarding the association between the outcome and
@@
 reason behind this behavior is rather subtle. Looking at the
-distribution of the regression coefficients we see that a large
+distribution of the absolute value of the regression coefficients we see that a large
 proportion of them are numerically very close to zero and have almost
@@
 uncertain and more close to a random permutation. This problem can be
-alleviated by artificially censoring all predictors with absolute
-coefficient values below a given threshold. For the Ridge models tuned
-with the deviance criterion, Figure \ref{fig:app2} (left) shows the
+alleviated by truncating all predictors with absolute coefficient values 
+below a given threshold thereby introducing an artificial censoring of the lists. 
+For the Ridge models tuned with the deviance criterion, Figure \ref{fig:app2} (left) shows the
 the sequential rank agreement where for each of the 1000 trained
-models predictors were artificially censored when their absolute
+models the predictors were artificially censored when their absolute
 coefficient value was lower than the $0.1\%$ quantile of the 5000
-absolute coefficient values. The corresponding curve from Figure
-\ref{fig:app1} (left panel) is shown for comparison.
-\textbf{CHECK UP: CENSOR ON POOLED DATA OR CENSOR BY INDIVIDUAL THRESHOLD}
-
-\textbf{DET HER ER NOGET VÆRRE ROD. DER ER FORSVUNDET NOGET LASSO TEKST. GÅ TILBAGE OG FIND DET IGEN}
-Under the deviance optimization criterion the median number of non-zero
-coefficients was 33 (range 18 to 52). In order to calculate the sequential
-rank agreement for the Lasso we therefore applied our proposed Algorithm \ref{sra-algorithm}
-with $B = 1000$ and $P=5000$.
-
-Thus, in our application the median number of censored predictors
-across the 1000 sub-samples was 5 (range 1 to 11). Even though the
-number of censored predictors is very small compared to the total
+absolute coefficient values. The curve was calculated
+using Algorithm \ref{sra-algorithm} with $B=1000$ and $P=5000$. The corresponding 
+curve from Figure \ref{fig:app1} (right panel) is shown for comparison.
+Even though the number of censored predictors is very small compared to the total
 number of predictors, the effect on the sequential rank agreement is
 substantial and with the artificial censoring the shape of the curves
-is as expected, starting low and then increasing. Also, the effect of
-the censoring caused by the sparsity of the Lasso is clearly visible
-in the agreement curves.
-\textbf{ANDREAS: man kan desvaerre ikke se hvor Lasso agreement forsvinder
-  (ved list-depth 5?) og om den er laver for list depth 1-5 end rigde
-  eller om den er kun laver for list depth 1-2. maaske kan du
-  beskrive dette i teksten?}
+is as expected, starting low and then increasing.
+
+Looking at the agreement curves for the Lasso models in Figure \ref{fig:app1} 
+(right) we clearly see the effect of the sparsity inducing penalization giving 
+rise to censored lists. These curves were similarly calculated using \ref{sra-algorithm} 
+and $1000$ random permutations. Under the deviance optimization criterion the 
+median number of non-zero coefficients was 33 (range 16 to 50) and for the 
+class accuracy criterion 14 (range 4 to 56). These values correspond to
+the list depths where the agreement curves become constant as a result of
+the subsequent censoring.
 
@@
 could essentially also have reversed the order to identify the
-patients that have low risk prediction.  An exception is the risk
-prediction agreement for the Lasso tuned with the AUC criterion which
+patients that have low risk prediction.
+
+An exception is the risk prediction agreement for the Lasso tuned with the AUC criterion which
 shows very low agreement among the high values of the predicted
@@
 Lasso shows similar agreement for ranking the risk predictions than
-PLS-DA (unless with the AUC criterion), and poor agreement for ranking
+PLS-DA (unless with the AUC criterion), and poorer agreement for ranking
 predictors. This reason for the latter is the high auto-correlation
@@
 the left panel of Figure \ref{fig:app1}, since it is not so important
-which specific variable is selected from a group of highly correlated
+which specific variable that gets selected from a group of highly correlated
 predictors when the purpose is risk predictions.
@@
 sub-samples and not with the actual accuracy of the risk
-predictions. Thus, we also computed the AUC and Brier scores for the different models
+predictions. Thus, we also computed the AUC values for the different models
 based on the DACOVA data. The distributions across the $1000$
@@
 of Figure \ref{fig:app2}.  Here we see that PLS-DA attains the highest
-AUC values with a median value of $0.71$ while the Ridge model with
+AUC values with a median value of $0.70$ while the Ridge model with
 the AUC criterion attains a median AUC of $0.49$. This implies that
 while Ridge regression optimized with respect to the AUC criterion
-achieves the best sequential rank agreement it performs similar to a
+achieves the best sequential rank agreement, it performs similar to a
 random coin toss with respect to classifying the DACOVA patients.
-\textbf{Comment on Brier scores}
-Both concerns are of course important in practice and for this data set 
-it appears thatPLS-DA is the overall best method taking both agreement 
-and classification accuracy into account.
+In practice both concerns are of importance.
 
@@
 
-Our applications show that XXX
-Through simulations we showed that XXX --- afventer Andreas.
+\textbf{Our applications show that XXX
+Through simulations we showed that XXX --- afventer Andreas.}
 

### 2015-08-10 latex/applicationscripts/plots.R
final application revision

@@ -105,7 +105,6 @@ title("Discrimination importance for Ridge with thresholding")
 legend("topright", c("No censoring", "Censored at 0.1% quantile"), col=c(1,2), lwd=3, lty=1, bty="n")
 
 
-
 ########## LASSO SUMMARY ##########
 median(sapply(res, function(a) length(a$lassoDev$importance)))
 range(sapply(res, function(a) length(a$lassoDev$importance)))
@@ -115,6 +114,11 @@ range(sapply(res, function(a) length(a$lassoClass$importance)))
 
 median(sapply(res, function(a) length(a$lassoAUC$importance)))
 range(sapply(res, function(a) length(a$lassoAUC$importance)))
+
+########## AUC SUMMARY ##########
+median(sapply(res, function(a) a$pls$auc))
+median(sapply(res, function(a) a$ridgeAUC$auc))
+
 
 
 save(list=ls()[-which(ls() == "res")], file="plots.RData")

### 2015-08-10 latex/applicationscripts/plots.R
final application revision

@@ -46,8 +46,6 @@ brierMat <- cbind(sapply(res, function(a) mean((a$rf$risk - dacovaLabels)^2)),
                   sapply(res, function(a) mean((a$ridgeAUC$risk - dacovaLabels)^2)))
 colnames(brierMat) <- c("Random Forest", "PLS-DA", "Lasso deviance", "Ridge deviance", "Ridge AUC")
 
-##
-save(list=ls()[-which(ls() == "res")], file="plots.RData")
 
 
 ########## RISK PLOT ##########

### 2015-08-11 latex/paper.Rnw
Finilizing pt 1

@@
   and can be applied to any number of lists even if some or all are
-  incomplete or censored. The main idea is then a permutation based
-  and graphically approach to assess the sequential changes of
-  agreement as a function of the depth of the lists. This may help to
-  identify change-points in the agreement of the lists. The
-  usefulness of these tools are illustrated using gene rankings, and
-  using data from two Danish ovarian cancer studies on where we assess
-  the within and between agreement of different statistical classification
-  methods.
+  incomplete or censored. The approach can identify change-points in
+  the agreement of the lists and the sequential changes of agreement
+  as a function of the depth of the lists can be compared graphically
+  to a permutation based reference set.  The usefulness of these tools
+  are illustrated using gene rankings, and using data from two Danish
+  ovarian cancer studies where we assess the within and between
+  agreement of different statistical classification methods.
 \end{abstract}
 
-Key words: sequential rank agreement, partial ranking, order statistic, gene rankings,
-methods comparison, variable selection, permutation
+Key words: sequential rank agreement, partial ranking, order
+statistic, gene rankings, methods comparison, variable selection,
+permutation
 
@@
 
-A common task of scientific research is to rank items obtained by some
-form of data analysis. Ranked lists occur naturally in applications of
-statistics. Regression methods rank predictor variables according to
+Ranking of items or results is common in scientific research and
+ranked lists occur naturally as the result of many statistical
+applications. Regression methods rank predictor variables according to
 magnitude of their association with an outcome, prediction models rank
@@
 When several rankings of the same items are available, a common
-research question is to what extent do they agree. Once agreement is
-established a further question is regarding the depths of agreement
-starting from the top of the ranked lists. A typical situation arises
-in high-dimensional genomics studies when several analysis methods are
-applied to rank a list of genes according to their association with a
-phenotype, treatment effect or other outcome. A measure of agreement
-of gene rankings obtained by different methods could often help to
-identify which genes are worth to pursue in further experiments.
+research question is to what extent they agree. In particular, is it
+possible to identify an optimal rank until which the lists agree on
+the items?  A typical situation arises in high-dimensional genomics
+studies when several analysis methods are applied to rank a list of
+genes according to their association with a phenotype, treatment
+effect or other outcome. A measure of agreement of gene rankings
+obtained by different methods could often help to identify which genes
+are worth to pursue in further experiments.
 
@@
 well-known, but these measures do not distinguish between agreement in
-the top versus towards the bottom of the lists. \citet{Shieh1998}
-proposed a weighted version of Kendall's $\tau$ where each pair of
-rankings can be assigned different weights, and \citet{Yilmaz2008}
-proposed the $\tau_{ap}$ which places higher emphasis on the top of
-the lists.
-%
+the top versus towards the bottom of the lists and the provide a
+single measure of the overall distance. \citet{Shieh1998} proposed a
+weighted version of Kendall's $\tau$ where each pair of rankings can
+be assigned different weights, and \citet{Yilmaz2008} proposed the
+$\tau_{ap}$ which places higher emphasis on the top of the lists.
 Spearman’s footrule uses the ranks of the variables for calculation of
@@
 of \citet{Bar-Ilan2006} where the reciprocal rank differences are used
-to calculate the agreement measure.
-
-More recent approaches consider the intersection of lists as the basis
+to calculate the similarity measure.
+
+Other recent approaches consider the intersection of lists as the basis
 for a similarity measure. However, simple intersection also places
-equal weights on all depths of the list. Thus, \citet{Fagin2003} and
-\citet{Webber2010} proposed modified intersections which put more
+equal weights on all depths of the list and therefore \citet{Fagin2003} and
+\citet{Webber2010} proposed weighted intersections which put more
 emphasis on the top of the lists. One proposal is denoted the average
@@
 than the potentially non-informative bottom of the lists.
-
 It is possible to use the existing methods to calculate agreement of
@@
 list. However, the interpretation may not be straightforward,
-especially in the case of more than two lists.
+especially in the case of more than two lists, and they may not
+accommodate partial rankings.
 
 In this article we introduce sequential rank agreement for measuring
-agreement among ranked lists. We provide graphical tools to analyse
-the depth of agreement of a set of ranked lists. The general idea is
-to define agreement based on the sequence of ranks from the first $d$
+agreement among (partially) ranked lists.  The general idea is to
+define agreement based on the sequence of ranks from the first $d$
 elements in each list. As agreement metric we adapt the limits of
 agreement known from agreement between quantitative variables
-\citep{Carstensen2010}. Our proposed approach allows to compare more
-than two lists simultaneously, provides a dynamic measure of agreement
-as a function of the depth in the lists, places high weight on the top
-of the list, accommodates censored/incomplete lists of varying
-lengths, and has a natural interpretation that directly relates to the
-ranks. The graphical illustration of sequential rank agreement
-potentially allows to infer a change-point, i.e., a list depth where
-there is a change in the agreement of the lists.
-
-In settings where rankings are obtained from analysis of correlated
-variables it is then possible to compare the actual results to the
-expected agreement in non-informative data obtained by
-randomization. In this sense our approach is a combination and
+\citep{alt:bland:1983,Carstensen2010}. Our proposed approach allows us
+to compare more than two lists simultaneously, it provides a dynamic
+measure of agreement as a function of the depth in the lists, places
+high weight on the top of the list, accommodates censored/incomplete
+lists of varying lengths, and has a natural interpretation that
+directly relates to the ranks. Graphical illustration of sequential
+rank agreement potentially allows us to infer a change-point, i.e., a
+list depth where a change in the agreement of the lists occur but we
+also provide randomization-based graphical tools to compare the
+observed rank agreement to the expected agreement found in
+non-informative data.  In this sense our approach is a combination and
 generalization of some of the ideas of \citet{Carterette2009} and
@@
 sequential rank agreement for multiple ranked lists and discuss how to
-handle incomplete/censored lists. In section 3 we present and discuss
-approaches to evaluate the sequential rank agreement obtained from an
-experiment and to identify the maximal depth until which the ranked
-lists agree to certain extent. Finally we present three different
-applications of the proposed sequential rank agreement approach to
-real data before we discuss the findings along with possible
-extensions. The approaches presented in this manuscript are available
-in the R package \texttt{SuperRanker}.
+handle incomplete/censored/partial lists. In section 3 we present and
+discuss approaches to evaluate the results obtained from sequential
+rank agreement. Finally we apply the sequential rank agreement to two
+Danish ovarian cancer studies before we discuss the findings along
+with possible extensions. The approaches presented in this manuscript
+are available in the R package \texttt{SuperRanker}.
 
@@
 ranking function $R: \{X_1,\dots,X_P\}\to \{1,\dots,P\}$, such that
-$R(X_p)$ is the rank of item $X_p$. The inverse mapping $R^{-1}$
-assigns to rank $r\in\{1,\dots,P\}$ the item $R^{-1}(r)$ that was
-given rank $r$. An ordered list is the realization of a ranking
-function $R$ applied to a data sample $X$ and we let
-$R_1(X),\dots,R_L(X)$, $L\geq2$, denote ordered lists obtained from
-$L$ different ranking functions applied to the same data $X$. We will
-use the same notation for the equivalent scenario where we have
-ordered lists obtained from applying the same ranking function to $L$
-samples from the same population, i.e., for ranked lists
-$R(X^1), \ldots, R(X^L)$.  Panels (a) and (b) of
+$R(X_p)$ is the rank of item $X_p$. The inverse mapping $R^{-1}$ gives
+the item $R^{-1}(r)$ that was assigned to rank $r\in\{1,\dots,P\}$. An
+ordered list is the realization of a ranking function $R$ applied to a
+data sample $X$ and we let $R_1(X),\dots,R_L(X)$, $L\geq2$, denote
+ordered lists obtained from $L$ different ranking functions applied to
+the same data $X$. We will use the same notation for the equivalent
+scenario where we have ordered lists obtained from applying the same
+ranking function to $L$ samples from the same population, i.e., for
+ranked lists $R(X^1), \ldots, R(X^L)$.  Panels (a) and (b) of
 Table~\ref{tab:example} show a schematic example of these
-mappings. Thus if $R^{-1}(1)=X_{34}$ then item $X_{34}$ is ranked
-first in list $l$, and similarly $R_l(X_{34})=1$.
+mappings. Thus if $R_l^{-1}(1)=X_{34}$ then item $X_{34}$ is ranked
+first in list $l$ and similarly $R_l(X_{34})=1$.
 
@@
 \end{equation}
-for a distance function $f.$ Throughout this paper we will use the
-sample standard error as our function $f$ and hence use
+for a non-negative real-valued distance function $f.$ Throughout this
+paper we will use the sample standard error as our function $f$ and
+hence use
 $$A(X_p) = \sqrt{\frac{\sum_{i=1}^L (R_i(X_p) - \bar{R}(X_p))^2}{L-1}},
@@
 where $\bar{R}(X_p)$ is the average rank assigned to item $X_p$, but
-other choices could be made. The sample standard error has an
-interpretation as the average distance of the individual rankings of
-the lists from the average ranking.
-
-We now describe what is exemplified in Panel (c) of Table
-\ref{tab:example} and how it can be used to define \emph{sequential
-  rank agreement}. For an integer $1\le d\le P$ we define the unique
-set of items found by merging the first $d$ elements of each of the
-$L$ lists, i.e., the set of items ranked less than or equal to $d$ in
-any of the lists:
+other possibilities will be discussed below. For each item, the sample
+standard error has an interpretation as the average distance of the
+individual rankings from the average ranking over all the lists, and
+it corresponds to the same measure that is used in method comparison
+studies to compute the limits of agreement \citep{alt:bland:1983}.
+
+For an integer $1\le d\le P$ we define the unique set of items found
+by merging the first $d$ elements of each of the $L$ lists, i.e., the
+set of items ranked less than or equal to $d$ in any of the lists:
 \begin{equation}
-S_d = \{R_l^{-1}(r) ; r \leq d, l = 1, \ldots, L \}.
+S_d = \{R_l^{-1}(r) ; r \leq d, l = 1, \ldots, L \},
 \end{equation}
-The \emph{sequential rank agreement} is the pooled standard deviation
-of the items found in the set $S_d$:
+which is exemplified in Panel (c) of Table
+\ref{tab:example}.
+
+The \emph{sequential rank agreement} is the pooled standard
+deviation of the items found in the set $S_d$:
 \begin{equation}
-\textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in
+\textrm{sra}(d)= \sqrt{\frac{\sum_{\{p \in
       S_d\}}(L-1)A(X_p)^2}{(L-1)|S_d|}}, \label{def:sra}
@@
 where $|S_d|$ refers to the cardinality of the set $S_d$. Values of
-\textrm{SRA} close to zero suggests that the lists agree while larger
-values suggests disagreement. If the ranked lists are identical then
-the value of SRA will be zero for all depths $d$. The sequential rank
-agreement can be interpreted as the average distance of the individual
-rankings of the lists from the average ranking for each of the items
-we have seen until depth $d$. Note that the terms that are part of
-\eqref{def:sra} are not independent since each list contains the ranks
-from 1 to $P$ exactly once. Hence we simply use the pooled standard
-deviation as a measure of the average distance between rankings.
+\textrm{sra} close to zero suggests that the lists agree while larger
+values suggest disagreement. Just like in method comparison studies we
+prefer smaller limits of agreement because that means that the
+differences between the methods is small. If the ranked lists are
+identical then the value of sequential rank agreement will be zero for
+all depths $d$. The sequential rank agreement can be interpreted as
+the average distance of the individual rankings of the lists from the
+average ranking for each of the items we have seen until depth $d$.
+
+Note that the terms that are part of \eqref{def:sra} are not
+independent since every list contains the ranks from 1 to $P$ exactly
+once. As a consequence we simply use the pooled standard deviation as
+a measure of the average distance between rankings and we refrain from
+using traditional distributional results for this measure.
 
@@
 content correlations (MIC) \citep{Reshef2011}. For the first two
-methods, the genes were ranked according to $p$ value, for logistic
-regression the genes were ordered by size of the corresponding
-coefficients (after standardization), and MIC was ordered by
-correlation which resulted in the top rankings seen in Table~\ref{tab1}.
+methods, the genes were ranked according to minimum $p$ value, for
+logistic regression the genes were ordered by size of the
+corresponding coefficients (after standardization), and MIC was
+ordered by absolute correlation which resulted in the top rankings
+seen in Table~\ref{tab1}.
 
@@
 top $d$ part of the lists to be for each depth $d$.  It can be seen
-that agreement is higher in the top of the lists (smaller values on
-the $y$ axis) corresponding to \emph{better} agreement than in the
-tails of the lists.
+that the sequential rank agreement limits are better towards the top
+of the lists (smaller values on the $y$ axis) corresponding to
+\emph{better} agreement than in the tails of the lists.
 
@@
 topk <- 20
+
+options(SweaveHooks=list(fig=function() par(cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)))
 
@@
 ## Plot the line
-plot(res[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
+plot(res[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth", cex=1.6)
 
@@
  \caption{Left panels: Sequential rank agreement for 4 different
-   analysis methods applied to the 3051 genes in the Golub data. Right
+   analysis methods applied to the 3051 genes in the Golub data /black
+   line). Right
    panels: Corresponding sequential rank agreement for the same data
@@
  If the lists agree on all items then the sequential rank agreement is
- zero for all depths. Generally,  if there are changes in the level of rank
- agreement then this suggests that there are sets of items that are ranked
- similarly in all lists and other sets of items that have been assigned vastly
- different ranks in the lists. In gene association studies we
- generally expect the agreement among the lists to be better towards
- the top of the lists and worse towards the end of the lists. In this
- case the sequential rank agreement curve will start at a low level
- and then increase until it levels off as in
+ zero for all depths. Generally, if there are changes in the level of
+ rank agreement then this suggests that there are sets of items that
+ are ranked similarly in all lists and other sets of items that have
+ been assigned vastly different ranks in the lists. In gene
+ association studies we generally expect the agreement among the lists
+ to be better towards the top of the lists and worse towards the end
+ of the lists. In this case the sequential rank agreement curve will
+ start at a low level and then increase until it levels off as in
  Figure~\ref{fig:example1}.
 
-\subsection{Analysis of incomplete/censored lists}
-
-Incomplete lists are also a common occurrence. They arise for example
-in case of missing data (items), when comparing top $d$ list results
-from publications, or when some methods only rank a subset of the
-items. For example, penalized regression based on the Lasso provides a
-sparse set of predictors that have non-zero coefficients. There is no
-obvious ordering of the set of predictors whose coefficient has been
-shrunk to zero. Incomplete lists also occur when the analyst decides
+\subsection{Analysis of partial/censored lists}
+
+Incomplete or partial lists are also a common occurrence. They arise
+for example in case of missing data (items), when comparing top $d$
+list results from publications, or when some methods only rank a
+subset of the items. For example, penalized regression based on the
+Lasso provides a sparse set of predictors that have non-zero
+coefficients. There is no obvious ordering of the set of predictors
+whose coefficient has been shrunk to zero and thus we end up with a
+partial ordering. Incomplete lists also occur when the analyst decides
 to censor all lists at depth $d$ such that the measurement of
-agreement is restricted to the first $d$ elements of each list.
-
-Sequential rank agreement can be generalized to incomplete/censored
-lists in the following way. Let $\Lambda_P$ be a set of $P$ items and
-$\Lambda_l\subset \Lambda_P$ the subset of $d_l$ items that occur in
-list $l$.  The case where all lists are censored at depth $d$
-corresponds to $d_1=\cdots=d_L=d$. For incomplete/censored lists the
-rank function becomes
+agreement is restricted to the first $d$ elements of each list, or
+when the analyst only wants to consider the ranking of items that have
+been found to be statistically significant.
+
+Sequential rank agreement can be generalized to partial lists in the
+following way. Let $\Lambda_l\subset X$ be the subset of $d_l$ items
+that have been ranked highest in list $l$.  The case where all lists are
+censored at the same depth $d$ corresponds to $d_1=\cdots=d_L=d$. For
+incomplete/censored lists the rank function becomes
 \begin{equation}
@@
 predictors in the presence of censored lists because the exact rank
-for some items will be unknown. Also, the rankings within a single
-list are not independent since each rank must appear exactly once in
-each list. Thus, we cannot simply assign the same number (e.g., the
-mean of the unassigned ranks) to the censored items since that would
-result in less variation of the ranks and hence less variation of the
-agreement, and it would artificially introduce a (downward) bias of
-agreement for items that are censored in multiple lists.
+for some items will be unknown. Also, recall that the rankings within
+a single list are not independent since each rank must appear exactly
+once in each list. Thus, we cannot simply assign the same number
+(e.g., the mean of the unassigned ranks) to the censored items since
+that would result in less variation of the ranks and hence less
+variation of the agreement, and it would artificially introduce a
+(downward) bias of agreement for items that are censored in multiple
+lists.
 
 Instead we randomize the ranks $\{d_{l}+1,\dots,P\}$ to the items that
-do not occur in list $\Lambda_l$. By randomizing the missing items of
-each list once we attain one realization of the $L$ rankings of the
-set $\Lambda_P$. By randomizing a large number of times we can compute
+do not occur in list $\Lambda_l$. One realization of the $L$ rankings
+of the set $X$ is obtained by randomizing the missing items of each
+list. By randomizing a large number of times we can compute
 \eqref{def:sra} for each realization, and then compute the sequential
@@
     randomly to the items \emph{not} found in the list, i.e.,
-    $\Lambda^\complement=\Lambda_P\setminus\Lambda_l$. Combine the
+    $\Lambda^\complement=X\setminus\Lambda_l$. Combine the
     result with $\Lambda_l$ to fill out the list.}
@@
 
-\section{No information agreement}
+\section{Evaluating sequential rank agreement}
 
@@
 distributions of sequential rank agreement under $H_0$ and
-$\wildetilde H_0$ each based on $400$ permutations. Not surprisingly, the
+$\widetilde H_0$ each based on $400$ permutations. Not surprisingly, the
 sequential rank agreement under $\widetilde H_0$ is lower than the

### 2015-08-11 latex/paper.Rnw
Code update

@@
 ## Plot the line
-plot(res[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth", cex=1.6)
+plot(res[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
 
@@
 
-<<fig2a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=FALSE>>=
+<<fig2a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=TRUE>>=
 ysize <- 1000
@@
 
-<<fig2b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=FALSE>>=
+<<fig2b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=TRUE>>=
 ## Same as above but zoomed in
@@
 interesting part of the sequential rank agreement curves is restricted
-to the depths where the percentage of censored items is
+to the depths where the number of censored items is
 low. Generally, without additional prior knowledge about the
@@
 attention of the sequential rank agreement to depths smaller than
-$\max(d_1, \ldots, d_L)$.
+$\max(d_1, \ldots, d_L)$ where at least one list provides an
+uncensored item.
 
@@
 ranking (in which case the sequential rank agreement will also be
-constant but at a low level). For censored lists we also expect a
-change-point around the depth where the lists are censored. This is an
-artefact stemming from the fact that we assume that the remainder of
-the lists can be replaced by a simple permutation of the missing
-items.
+constant but at a low level). For partially ranked lists we also
+expect a change-point around the depth where the lists are
+censored. This is an artefact stemming from the fact that we assume
+that the remainder of the lists can be replaced by a simple
+permutation of the missing items.
 
@@
 
-We now consider the special case where the rankings are obtained by
-applying different methods to the same data set which contains one
-outcome variable and a list of candidate predictors. To interpret the
-sequential rank agreement values in this situation we propose two
+To interpret the sequential rank agreement values we propose two
 different benchmark values corresponding to two different
@@
 H_0  & : &  \text{The list rankings correspond to complete randomly}\\
-       &  & \text{permuted lists}.
+       &  & \text{permuted lists},
 \end{eqnarray*}
-This does not only assume that there is no information in data on
+which not only assumes that there is no information in the data on
 which the rankings are based but also that the methods are completely
@@
 within each list and then computing the sequential rank agreement for
-the permuted lists. In the uncensored case each experiment contains
-$L$ lists of random permutations of the numbers $1,\dots,P$. For the
-censored/incomplete case we first permute the numbers $1,\dots,P$ and
-then censor list $l$ at $d_l$. The sequential rank agreement curve
-from the the original lists can then be compared to the pointwise rank
-agreements obtained under $H_0$.
-
-To obtain the distribution under $\widetilde H_0$ the idea is to repeat
-many times the ranking procedures in uncorrelated data.  Thus, we
-first permute the outcome variable of the data set on which the
+the permuted lists. In the fully observed case each experiment
+contains $L$ lists of random permutations of the items in $X$. For the
+censored/partially observed case we first permute the items
+$X_1,\dots,X_P$ and then censor list $l$ at $d_l$. The sequential rank
+agreement curve from the the original lists can then be compared to,
+say, the pointwise 95\% quantiles of the observed rank agreements
+obtained under $H_0$.
+
+To obtain the distribution under $\widetilde H_0$ the idea is to
+repeat the ranking procedures for unassociated data many times.  Thus,
+we first permute the outcome variable of the data set on which the
 rankings are based. This removes any association between the predictor
@@
 the original ranked lists agree \emph{more} than what we would expect
-in data with no information.
-
-For the analysis of the Golub data described in Section
-\ref{sec:amfol}, Figure~\ref{fig:example1} shows the empirical
-distributions of sequential rank agreement under $H_0$ and
-$\widetilde H_0$ each based on $400$ permutations. Not surprisingly, the
-sequential rank agreement under $\widetilde H_0$ is lower than the
-sequential rank agreement under $H_0$ because the four methods used to
-rank the data ($t$ test, logistic regression, elastic net, and MIC)
-generally tend to identify (and rank high) similar predictors even if
-there are only spurious associations. The two bottom panels in
-Figure~\ref{fig:example1} also indicate that the observed sequential
-rank agreement (the black line) is better than what would be expected
-by chance for data that contain no information since it lies below the
-reference areas. However, the censored data also suggests that there
-may be at most 1 or 2 ranked items towards the top of the lists that
-yield a result better than what would be expected (the bottom-right
-plot).
+in data with no information, and therefore that the information in the
+lists is significantly more in agreement than what would be expected.
+
+Figure~\ref{fig:example1} shows the empirical distributions of
+sequential rank agreement under $H_0$ and $\widetilde H_0$ each based
+on $400$ permutations of the Golub data described in Section
+\ref{sec:amfol}. Not surprisingly, the sequential rank agreement under
+$\widetilde H_0$ is lower than the sequential rank agreement under
+$H_0$ because the four methods used to rank the data ($t$ test,
+logistic regression, elastic net, and MIC) generally tend to identify
+(and rank) similar predictors even if there are only spurious
+associations. The two bottom panels in Figure~\ref{fig:example1} also
+indicate that the observed sequential rank agreement (the black line)
+is better than what would be expected by chance for data that contain
+no information since it lies below the reference areas. However, the
+censored data also suggests that there may be at most 1 or 2 ranked
+items towards the top of the lists that yield a result better than
+what would be expected (the bottom-right plot).
 
@@
 
-The average overlap is widely used in comparisons of (two) ranked
-lists \citep{Fagin2003,Webber2010} and the idea behind the average
-overlap closely resembles the sequential rank agreement. In this
-section we present an extension of the average overlap to more that
-two lists and compare this to sequential rank agreement.
+The average overlap is a measure that is widely used in comparisons of
+two fully observed ranked lists \citep{Fagin2003,Webber2010} and the
+idea behind the average overlap closely resembles the sequential rank
+agreement. In this section we present an extension of the average
+overlap to more that two lists and compare this to sequential rank
+agreement.
 
@@
 \end{equation}
-Typically the overlap is only considered for pairwise comparisons but
-formula \eqref{def:overlap} can be applied even when $L>2$
-\citep{Bar-Ilan2006,Boulesteix2009}.
+Typically the overlap is only considered for pairwise comparisons
+\citep{Bar-Ilan2006,Boulesteix2009} but formula \eqref{def:overlap}
+is easily extended to the situation where $L>2$ as shown above.
 
@@
 \begin{equation}
-AO(d) = \sum_{i=1}^d \frac{O(d)}{d}.
+AO(d) = \frac1d\sum_{i=1}^d O(d).
 \end{equation}
@@
 percentages (of $L$ sets) which is somewhat less intuitive than our
-proposed average rank distance.
+proposed sequential rank distance.
 
 In the following we compare the average overlap and the sequential
-rank agreement using the analysis of the Golub data \citep{Golub1999}
-of Section \ref{sec:amfol}. We consider a typical application of the
-average overlap and thus restrict attention to only two gene rankings
-obtained by the $t$ tests and logistic regression. These two marginal
-modelling approaches should result in very similar gene rankings. The
-first two columns of Table~\ref{tab1} show the actual top-10 ranking
-of the predictors.
-
-<<fig-ao1, echo=FALSE, fig=TRUE,include=FALSE, eval=FALSE>>=
+rank agreement using the analysis of the Golub data
+\citep{Golub1999}. We consider a typical application of the average
+overlap and thus restrict attention to only two gene rankings obtained
+by the $t$ tests and logistic regression. These two marginal modeling
+approaches should result in very similar gene rankings. The first two
+columns of Table~\ref{tab1} show the actual top-10 ranking of the
+predictors.
+
+<<fig-ao1, echo=FALSE, fig=TRUE,include=FALSE, eval=TRUE>>=
 myres <- average_overlap(inputmatrix[,1:2])
@@
 
-<<fig-ao2, echo=FALSE, fig=TRUE,include=FALSE, eval=FALSE>>=
+<<fig-ao2, echo=FALSE, fig=TRUE,include=FALSE, eval=TRUE>>=
 mmyres <- sra(inputmatrix[,1:2])
@@
 
-The black lines in Figure~\ref{fig:case1} show the average overlap and
-sequential rank agreement for these data. It is clear from both plots
-that there is perfect agreement towards the top of the ranked lists:
-the average overlap is 1 and the sequential rank agreement is zero.
-However, if we remove those two items from the data (i.e.,
+ The black lines in Figure~\ref{fig:case1} show the average overlap
+ and sequential rank agreement for these data. It is clear from both
+ plots that there is perfect agreement towards the top of the ranked
+ lists: the average overlap is 1 and the sequential rank agreement is
+ zero.  However, if we remove those two items from the data (i.e.,
  gene/predictor 2124 and 896) and redo the analyses then we get
@@
  agreement from rank 3 as we did from the full dataset. This indicates
-that sequential rank is more robust against small perturbations of the
-data. For the average overlap the curves completely change when these
-two items are removed and which would lead to a different conclusion
-about the agreement among the items from rank 3 downwards.
+ that sequential rank is more robust against small perturbations of
+ the data. For the average overlap the curves completely change when
+ these two items are removed and which would lead to a different
+ conclusion about the agreement among the items from rank 3 downwards.
 

### 2015-08-11 R/sra.R

@@ -167,7 +167,11 @@ sra.list <- function(object, B=1, na.strings=NULL, nitems=max(sapply(object, len
                          res <- sracppfull(rankmat, type=itype)
                          res
                      })
-    agreement <- ifelse(itype==0, sqrt(rowMeans(tmpres)), rowMeans(tmpres))
+    if (itype==0) {
+        agreement <- sqrt(rowMeans(tmpres))
+    } else {
+        agreement <- rowMeans(tmpres)
+    }
     names(agreement) <- items
     class(agreement) <- "sra"
     attr(agreement, "B") <- B

### 2015-08-11 latex/paper.Rnw
Submitted version

@@
 of the lists (smaller values on the $y$ axis) corresponding to
-\emph{better} agreement than in the tails of the lists.
+\emph{better} agreement than in the tails of the lists. It is clear
+from the curve in Figure~\ref{fig:example1} that there is a
+substantial drop in agreement even after the first depth. Thus, if we
+were to restrict attention to a small set of predictors then we would
+focus on predictors 2124, 829, and 378 as seen in Table~\ref{tab1}.
 
@@
 
-\section{Application}
+\section{Application to ovarian cancer data}
 We now consider an application of the sequential rank agreement to two
@@
 
-\textbf{Our applications show that XXX
-Through simulations we showed that XXX --- afventer Andreas.}
-
 The sequential rank agreement can be used to determine the depth at
@@
 identify a change in agreement among the items that were deemed to
-relevant. In our application section we have successfully introduced
+be relevant. In our application section we have successfully introduced
 such an artificial censoring for the predictor rankings obtained with
@@
 can also be unstable when the number of observations is low and
-alternative measures such as the mean absolute deviance may prove more
+alternative measures such as the median absolute deviance may prove more
 stable in some situations. However, the current definition using the
 standard deviation is analogous to the approach used for agreement in
-method comparison studies.
+method comparison studies so we have used that.
 

### 2015-08-11 R/RcppExports.R
Implementation of median absolute difference

@@ -12,6 +12,14 @@ overlap <- function(rankMat) {
     .Call('SuperRanker_overlap', PACKAGE = 'SuperRanker', rankMat)
 }
 
+median_rcpp <- function(x) {
+    .Call('SuperRanker_median_rcpp', PACKAGE = 'SuperRanker', x)
+}
+
+mad_rcpp <- function(x, scale_factor = 1.4826) {
+    .Call('SuperRanker_mad_rcpp', PACKAGE = 'SuperRanker', x, scale_factor)
+}
+
 #' Compute the sequential rank agreement between k ranked lists
 #' 
 #' @description Computes the sequential rank agreement (number of items present in all k lists divided by the current rank) for each rank in the k lists
@@ -19,7 +27,7 @@ overlap <- function(rankMat) {
 #' @param maxlength The maximum depth that are needed XXX
 #' @param B The number of resamples to use in the presence of censored lists
 #' @param cens A vector of integer values that
-#' @param type The type of distance measure to use: 0 (the default) is the variance while 1 is MAD (mean absolute deviation)
+#' @param type The type of distance measure to use: 0 (the default) is the variance while 1 is MAD (median absolute deviation)
 #' @return A vector of the same length as the rows in rankMat containing the squared (!) sequential rank agreement between the lists for each depth. If the MAD type was chosen then the sequential MAD values are returned
 #' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
 #' @export
@@ -37,5 +45,9 @@ sracpp <- function(rankMat, maxlength, B, cens, type = 0L) {
 #' @export
 sracppfull <- function(rankMat, type = 0L) {
     .Call('SuperRanker_sracppfull', PACKAGE = 'SuperRanker', rankMat, type)
+}
+
+Median <- function(x) {
+    .Call('SuperRanker_Median', PACKAGE = 'SuperRanker', x)
 }
 

### 2015-08-11 R/sra.R

@@ -8,7 +8,7 @@
 #' @param na.strings A vector of strings/values that represent missing values in addition to NA. Defaults to NULL which means only NA are censored values.
 #' @param B An integer giving the number of randomization to sample
 #' over in the case of censored observations
-#' @param type The type of measure to use. Either sd (standard deviation - the default) or mad (mean absolute deviance)
+#' @param type The type of measure to use. Either sd (standard deviation - the default) or mad (median absolute deviance)
 #' @return A vector of the sequential rank agreement
 ##' @examples
 ##'
@@ -187,7 +187,7 @@ sra.list <- function(object, B=1, na.strings=NULL, nitems=max(sapply(object, len
 #' @param B Either a vector or matrix
 #' @param n the number of sequential rank agreement curves to produce
 #' @param na.strings A vector of character values that represent vensored observations
-#' @param type The type of measure to use. Either sd (standard deviation - the default) or mad (mean absolute deviance)
+#' @param type The type of measure to use. Either sd (standard deviation - the default) or mad (median absolute deviance)
 #' @return A matrix with n columns each representing the sequential rank agreement obtained from
 #' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
 #' @export

### 2015-08-11 src/RcppExports.cpp

@@ -16,6 +16,29 @@ BEGIN_RCPP
     return __result;
 END_RCPP
 }
+// median_rcpp
+double median_rcpp(NumericVector x);
+RcppExport SEXP SuperRanker_median_rcpp(SEXP xSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject __result;
+    Rcpp::RNGScope __rngScope;
+    Rcpp::traits::input_parameter< NumericVector >::type x(xSEXP);
+    __result = Rcpp::wrap(median_rcpp(x));
+    return __result;
+END_RCPP
+}
+// mad_rcpp
+double mad_rcpp(NumericVector x, double scale_factor);
+RcppExport SEXP SuperRanker_mad_rcpp(SEXP xSEXP, SEXP scale_factorSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject __result;
+    Rcpp::RNGScope __rngScope;
+    Rcpp::traits::input_parameter< NumericVector >::type x(xSEXP);
+    Rcpp::traits::input_parameter< double >::type scale_factor(scale_factorSEXP);
+    __result = Rcpp::wrap(mad_rcpp(x, scale_factor));
+    return __result;
+END_RCPP
+}
 // sracpp
 NumericVector sracpp(IntegerMatrix rankMat, int maxlength, int B, IntegerVector cens, int type);
 RcppExport SEXP SuperRanker_sracpp(SEXP rankMatSEXP, SEXP maxlengthSEXP, SEXP BSEXP, SEXP censSEXP, SEXP typeSEXP) {
@@ -40,6 +63,17 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< IntegerMatrix >::type rankMat(rankMatSEXP);
     Rcpp::traits::input_parameter< int >::type type(typeSEXP);
     __result = Rcpp::wrap(sracppfull(rankMat, type));
+    return __result;
+END_RCPP
+}
+// Median
+double Median(NumericVector x);
+RcppExport SEXP SuperRanker_Median(SEXP xSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject __result;
+    Rcpp::RNGScope __rngScope;
+    Rcpp::traits::input_parameter< NumericVector >::type x(xSEXP);
+    __result = Rcpp::wrap(Median(x));
     return __result;
 END_RCPP
 }

### 2015-08-11 src/sra.cpp

@@ -32,25 +32,31 @@ double square( double x ){
     return x*x;
 }
 
-// [ [ Rcpp::export]]
-int median(NumericVector x) {
-  int xlen=x.size();
-  int n;
-  
-  if (xlen % 2) { /* x has odd length */
-    n = (xlen+1)/2;
-  } else {
-    n = xlen/2+1;
-  }
-  
+// [[Rcpp::export]]
+double median_rcpp(NumericVector x) {
   NumericVector y = clone(x);
-  std::nth_element(y.begin(), y.begin()+n-1, y.end());
-
-  if (xlen % 2) { /* x has odd length */
-    return (y[n]);
+  int n, half;
+  double y1, y2;
+  n = y.size();
+  half = n / 2;
+  if(n % 2 == 1) {
+    // median for odd length vector
+    std::nth_element(y.begin(), y.begin()+half, y.end());
+    return y[half];
   } else {
-    return((y[n-1] + y[n])/2.0);
+    // median for even length vector
+    std::nth_element(y.begin(), y.begin()+half, y.end());
+    y1 = y[half];
+    std::nth_element(y.begin(), y.begin()+half-1, y.begin()+half);
+    y2 = y[half-1];
+    return (y1 + y2) / 2.0;
   }
+}
+
+// [[Rcpp::export]]
+double mad_rcpp(NumericVector x, double scale_factor = 1.4826) {
+  // scale_factor = 1.4826; default for normal distribution consistent with R
+  return median_rcpp(abs(x - median_rcpp(x))) * scale_factor;
 }
 
 
@@ -61,7 +67,8 @@ NumericVector SdPerRow(NumericMatrix obj, int type) {
   for (int i=0; i<obj.nrow(); i++) {
     if (type == 1) {
       // MAD
-      res[i] = mean(abs(obj.row(i) - mean(obj.row(i))));
+      // res[i] = mean(abs(obj.row(i) - mean(obj.row(i))));
+      res[i] = mad_rcpp(obj.row(i), 1);
     }
     else {
       res[i] = var(obj.row(i));
@@ -79,7 +86,7 @@ NumericVector SdPerRow(NumericMatrix obj, int type) {
 //' @param maxlength The maximum depth that are needed XXX
 //' @param B The number of resamples to use in the presence of censored lists
 //' @param cens A vector of integer values that
-//' @param type The type of distance measure to use: 0 (the default) is the variance while 1 is MAD (mean absolute deviation)
[…]

### 2015-08-13 R/RcppExports.R
Code fix

@@ -47,7 +47,3 @@ sracppfull <- function(rankMat, type = 0L) {
     .Call('SuperRanker_sracppfull', PACKAGE = 'SuperRanker', rankMat, type)
 }
 
-Median <- function(x) {
-    .Call('SuperRanker_Median', PACKAGE = 'SuperRanker', x)
-}
-

### 2015-08-13 src/RcppExports.cpp

@@ -66,14 +66,3 @@ BEGIN_RCPP
     return __result;
 END_RCPP
 }
-// Median
-double Median(NumericVector x);
-RcppExport SEXP SuperRanker_Median(SEXP xSEXP) {
-BEGIN_RCPP
-    Rcpp::RObject __result;
-    Rcpp::RNGScope __rngScope;
-    Rcpp::traits::input_parameter< NumericVector >::type x(xSEXP);
-    __result = Rcpp::wrap(Median(x));
-    return __result;
-END_RCPP
-}

### 2015-08-18 latex/paper.Rnw
Code fix
Adjusting figure margins
Fixing figure margins
homogenizing graphics

@@
 
-\title{Sequential rank agreement methods for comparison of lists}
+\title{Sequential rank agreement methods for comparison of ranked lists}
 \author{Claus Thorn Ekstrøm}
@@
 
-options(SweaveHooks=list(fig=function() par(cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)))
+options(SweaveHooks=list(fig=function() par(cex.lab=1.4, cex.axis=1.5, cex.main=1.2, cex.sub=1.5)))
 
@@
 ## Plot the line
-plot(res[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
+plot(res[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="List depth", bty="n")
 
@@
 ysize <- 50
-plot(res[1:ysize], lwd=3, col="black", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
+plot(res[1:ysize], lwd=3, col="black", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="List depth", bty="n")
 
@@
 ## Plot the line
-plot(res2[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
+plot(res2[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="List depth", bty="n")
 
@@
 }
-## Uncomment the two lines below to redu the analysis
+## Uncomment the two lines below to redo the analysis
 ##null2 <- makereference(x, y, n=100, B=B)
@@
 ysize <- 50
-plot(res2[1:ysize], lwd=3, col="black", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
+plot(res2[1:ysize], lwd=3, col="black", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="List depth", bty="n")
 
@@
 
-\begin{figure}[tb]
+\begin{figure}[htbp]
    \begin{center}
- \includegraphics[width=.45\textwidth]{paper-fig1a}
- \includegraphics[width=.45\textwidth]{paper-fig2a}
- \includegraphics[width=.45\textwidth]{paper-fig1b}
- \includegraphics[width=.45\textwidth]{paper-fig2b}
+ \includegraphics[width=.49\textwidth]{paper-fig1a}
+ \includegraphics[width=.49\textwidth]{paper-fig2a}
+ \includegraphics[width=.49\textwidth]{paper-fig1b}
+ \includegraphics[width=.49\textwidth]{paper-fig2b}
  \end{center}
@@
 xvals <- 1:500
-plot(xvals, myres[xvals], xlim=c(1, 150), ylim=c(0,1), type="l", lwd=2, ylab="Average overlap", xlab="List depth")
+plot(xvals, myres[xvals], xlim=c(1, 150), ylim=c(0,1), type="l", lwd=2, ylab="Average overlap", xlab="List depth", bty="n")
 lines(xvals[-c(1,2)], myres2[1:498], col="red", lwd=2)
@@
 xvals <- 1:500
-plot(xvals, mmyres[xvals], xlim=c(1, 150), ylim=c(0, 110), type="l", lwd=2, ylab="Sequential rank agreement", xlab="List depth")
+plot(xvals, mmyres[xvals], xlim=c(1, 150), ylim=c(0, 110), type="l", lwd=2, ylab="Sequential rank agreement", xlab="List depth", bty="n")
 lines(xvals[-c(1,2)], mmyres2[1:498], col="red", lwd=2)
@@
 
-\begin{figure}[tb]
+\begin{figure}[htbp]
 \begin{center}
-\includegraphics[width=.45\textwidth]{paper-fig-ao1}
-\includegraphics[width=.45\textwidth]{paper-fig-ao2}
+\includegraphics[width=.49\textwidth]{paper-fig-ao1}
+\includegraphics[width=.49\textwidth]{paper-fig-ao2}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=.5\textwidth]{pics/riskAgreementPlot}%
-\includegraphics[width=.5\textwidth]{pics/predictorAgreementPlot}
+\includegraphics[width=.49\textwidth]{pics/riskAgreementPlot}%
+\includegraphics[width=.49\textwidth]{pics/predictorAgreementPlot}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=.5\textwidth]{pics/ridgeCensoring}%
-\includegraphics[width=.5\textwidth]{pics/aucPlot}
+\includegraphics[width=.49\textwidth]{pics/ridgeCensoring}%
+\includegraphics[width=.49\textwidth]{pics/aucPlot}
 \end{center}

### 2015-08-18 latex/applicationscripts/plots.R

@@ -37,19 +37,20 @@ aucMat <- cbind(sapply(res, function(a) a$rf$auc),
                 sapply(res, function(a) a$lassoDev$auc),
                 sapply(res, function(a) a$ridgeDev$auc),
                 sapply(res, function(a) a$ridgeAUC$auc))
-colnames(aucMat) <- c("Random Forest", "PLS-DA", "Lasso deviance", "Ridge deviance", "Ridge AUC")
+colnames(aucMat) <- c("RF", "PLS-DA", "Lasso dev", "Ridge dev", "Ridge AUC")
 
 brierMat <- cbind(sapply(res, function(a) mean((a$rf$risk - dacovaLabels)^2)),
                   sapply(res, function(a) mean((a$pls$risk - dacovaLabels)^2)),
                   sapply(res, function(a) mean((a$lassoDev$risk - dacovaLabels)^2)),
                   sapply(res, function(a) mean((a$ridgeDev$risk - dacovaLabels)^2)),
                   sapply(res, function(a) mean((a$ridgeAUC$risk - dacovaLabels)^2)))
-colnames(brierMat) <- c("Random Forest", "PLS-DA", "Lasso deviance", "Ridge deviance", "Ridge AUC")
+colnames(brierMat) <- c("RF", "PLS-DA", "Lasso dev", "Ridge dev", "Ridge AUC")
 
 
 
 ########## RISK PLOT ##########
-plot(rfRisk, ylim=c(0,30), xlim=c(0,120))
+par(cex.lab=1.4, cex.axis=1.5, cex.main=1.2, cex.sub=1.5)
+plot(rfRisk[1:120], ylim=c(0,30), xlim=c(0,120), type="l", lwd=3, bty="n", xlab="List depth", ylab="Sequential rank agreement")
 title("Predicted risks of malignant tumor in 113 patients (DACOVA)")
 
 plot(lassoDevRisk, add=TRUE, lines.col="red")
@@ -64,11 +65,12 @@ plot(ridgeAucRisk, add=TRUE, lines.col="forestgreen", lines.lty=3)
 
 legend("topright", c("Random Forest", "PLS-DA", "Lasso deviance", "Lasso class", "Lasso AUC", "Ridge deviance", "Ridge class", "Ridge AUC"),
[…]

### 2015-08-19 latex/paper.Rnw
homogenizing graphics

@@
 xvals <- 1:500
-plot(xvals, myres[xvals], xlim=c(1, 150), ylim=c(0,1), type="l", lwd=2, ylab="Average overlap", xlab="List depth", bty="n")
-lines(xvals[-c(1,2)], myres2[1:498], col="red", lwd=2)
+plot(xvals, myres[xvals], xlim=c(1, 150), ylim=c(0,1), type="l", lwd=3, ylab="Average overlap", xlab="List depth", bty="n")
+lines(xvals[-c(1,2)], myres2[1:498], col="red", lwd=3)
 @
@@
 xvals <- 1:500
-plot(xvals, mmyres[xvals], xlim=c(1, 150), ylim=c(0, 110), type="l", lwd=2, ylab="Sequential rank agreement", xlab="List depth", bty="n")
-lines(xvals[-c(1,2)], mmyres2[1:498], col="red", lwd=2)
+plot(xvals, mmyres[xvals], xlim=c(1, 150), ylim=c(0, 110), type="l", lwd=3, ylab="Sequential rank agreement", xlab="List depth", bty="n")
+lines(xvals[-c(1,2)], mmyres2[1:498], col="red", lwd=3)
 @

### 2016-02-24 latex/paper.Rnw
start of new section for review: change point analysis

@@
 
+\section{Change-point analysis}
+
+In many applications it is of interest to estimate a list depths which
+satisfies a change point criterion. In this section we investigate the
+theoretical properties of our proposed method for this specific
+task. As Hall and Schimek, we work with an infinite set of lists
+\mathcal{L} and study the asymptotic behaviour for
+$L\to\infty$. However, we do not allow the length of the lists to
+change with $L$ as in most applications the length of the lists is
+fixed [FIXME: for finite $L$ are there only finitely many forskelige
+lists (aber dass stoert keinen grossen geist, siger karlsson)]. First,
+we define the parameter of interest. For any given threshold value $q$
+a change-point $d_0(q)$ in the list agreement is defined by
+\begin{equation*}
+  d_0=\textrm{arginf}_{d\in\{1,\dots,L\}}\textrm{sra}(d)>q.
+\end{equation*}
+The value of $q$ could for example be the agreement of randomly
+permuted lists [FIXME: there is only one SRA value for infinitely many
+random permutations, no skygge!].
+
+The theorem shows consistency of the plug-in estimate of
+\begin{equation*}
+\hat{d_0} = \textrm{arginf}_{d\in\{1,\dots,L\}}\widehat{\textrm{sra}}(d)>q
+\end{equation*}
+[FIXME: $\textrm{sra}$ -> $\widehat{\textrm{sra}}$ in rest of article?]
+\begin{theorem}
+  $\hat{d_0}\overset{L\to\infty}{\longrightarrow}d_0$
+\end{theorem}
+
+Proof: We rewrite the square of the sequential rank agreement \eqref{def:sra} as
+\begin{equation}
+(\textrm{sra}(d))^2= \sum_{p=1}^P \sum_{l=1}^L 1\{R_l(X_p)\le d\}\sum_{p=1}^P(1-\prod_{l=1}^L 1\{R_l(X_p)>d\}) A(X_p)^2
+\end{equation}
+[FIXME: Andreas details the proof with Cauchy-Schwarz.]
+[FIXME: Det er lidt kedelig for saadan en fin en som Andreas: Skal theoremet og bevis ikke vaere for en general agreement metric, ikke nodvendigvis den i equation (2), og kan vi charakterisere hvad den skal opfylde?]
+
 \section{Comparison to average overlap}

### 2016-03-30 latex/paper.Rnw
fixed typo in formula
Old changes
Fixing Tex + new compilation

@@
 
+\usepackage[amsthm,thmmarks]{ntheorem}
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
+\newtheorem{proposition}{Proposition}
+\newtheorem*{proposition2}{Proposition}
+\newtheorem{corollary}{Corollary}
+\newtheorem{lemma}{Lemma}
+\newtheorem{assumption}{Assumption}
+\newtheorem{example}{Example}
+\renewcommand\proofSymbol{\ensuremath{\blacksquare}}
+
 \SweaveOpts{prefix.string=paper, keep.source=TRUE}
@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 
@@
 task. As Hall and Schimek, we work with an infinite set of lists
-\mathcal{L} and study the asymptotic behaviour for
+$\mathcal{L}$ and study the asymptotic behaviour for
 $L\to\infty$. However, we do not allow the length of the lists to
@@
 [FIXME: $\textrm{sra}$ -> $\widehat{\textrm{sra}}$ in rest of article?]
+
 \begin{theorem}
   $\hat{d_0}\overset{L\to\infty}{\longrightarrow}d_0$
-\end{theorem}
-
+
+\begin{proof}
 Proof: We rewrite the square of the sequential rank agreement \eqref{def:sra} as
 \begin{equation}
-(\textrm{sra}(d))^2= \sum_{p=1}^P \sum_{l=1}^L 1\{R_l(X_p)\le d\}\sum_{p=1}^P(1-\prod_{l=1}^L 1\{R_l(X_p)>d\}) A(X_p)^2
+(\textrm{sra}(d))^2= \frac 1{\sum_{p=1}^P \sum_{l=1}^L 1\{R_l(X_p)\le d\}}\sum_{p=1}^P(1-\prod_{l=1}^L 1\{R_l(X_p)>d\}) A(X_p)^2
 \end{equation}
@@
 [FIXME: Det er lidt kedelig for saadan en fin en som Andreas: Skal theoremet og bevis ikke vaere for en general agreement metric, ikke nodvendigvis den i equation (2), og kan vi charakterisere hvad den skal opfylde?]
+\end{proof}
+\end{theorem}
 

### 2016-03-31 latex/paper.tex
Introducing population-level definition

@@
-\documentclass[12pt,a4paper]{article}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
+\documentclass[12pt,a4paper]{article}
+\usepackage[utf8]{inputenc}
+\usepackage[]{graphicx}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage[]{color}
+\usepackage{subcaption}
+\usepackage{natbib}
+\usepackage{xspace}
+\usepackage{calc}
+\usepackage[noae]{Sweave}
+\usepackage{authblk}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+%\usepackage{geometry}
+%\geometry{a4paper, portrait, margin=1in}
+%
+% New commands
+%
+\newcommand{\ie}{\emph{i.e.}\@\xspace}
+\newcommand{\eg}{\emph{e.g.}\@\xspace}
 
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\DeclareMathOperator{\E}{\mathbb{E}}
+\DeclareMathOperator{\Var}{Var}
+\DeclareMathOperator{\Cov}{Cov}
 
-\usepackage{framed}
-\usepackage{subcaption}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
+\usepackage[amsthm,thmmarks]{ntheorem}
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
+\newtheorem{proposition}{Proposition}
+\newtheorem*{proposition2}{Proposition}
+\newtheorem{corollary}{Corollary}
+\newtheorem{lemma}{Lemma}
+\newtheorem{assumption}{Assumption}
+\newtheorem{example}{Example}
+\renewcommand\proofSymbol{\ensuremath{\blacksquare}}
 
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
-\usepackage{alltt}
-\usepackage{amsmath}
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\begin{document}
-%\SweaveOpts{concordance=TRUE}
+%%% RSS specific items
 
 \title{Sequential rank agreement methods for comparison of ranked lists}
-\author{}
+\author{Claus Thorn Ekstrøm}
+\author{Thomas Alexander Gerds}
+\author{Andreas Kryger Jensen}
+\author{Kasper Brink-Jensen}
+\affil{Biostatistics, University of Copenhagen}
+\renewcommand{\topfraction}{.85}
+\renewcommand{\bottomfraction}{.7}
+\renewcommand{\textfraction}{.15}
+\renewcommand{\floatpagefraction}{.66}
+\renewcommand{\dbltopfraction}{.66}
+\renewcommand{\dblfloatpagefraction}{.66}
+\setcounter{topnumber}{9}
+\setcounter{bottomnumber}{9}
+\setcounter{totalnumber}{20}
+\setcounter{dbltopnumber}{9}
+
+\begin{document}
+\input{paper-concordance}
+
+% Use square brackets for numbering affiliations
 
@@
 \begin{abstract}
+  The comparison of alternative rankings of a set of items is a
+  general and prominent task in applied statistics. Predictor
+  variables are ranked according to magnitude of association with an
+  outcome, prediction models rank subjects according to the
+  personalized risk of an event, and genetic studies rank genes
+  according to their difference in gene expression levels. This
+  article constructs measures of the agreement of two or more ordered
+  lists. We use the standard deviation of the ranks to define a
+  measure of agreement that both provides an intuitive interpretation
+  and can be applied to any number of lists even if some or all are
+  incomplete or censored. The approach can identify change-points in
+  the agreement of the lists and the sequential changes of agreement
+  as a function of the depth of the lists can be compared graphically
+  to a permutation based reference set.  The usefulness of these tools
+  are illustrated using gene rankings, and using data from two Danish
+  ovarian cancer studies where we assess the within and between
+  agreement of different statistical classification methods.
+\end{abstract}
 
-\end{abstract}
+Key words: sequential rank agreement, partial ranking, order
+statistic, gene rankings, methods comparison, variable selection,
+permutation
 
@@
 
-Ranked lists occur in many applications of statistics. Regression
-methods rank predictor variables according to magnitude of association
-with outcome, prediction models rank subjects according to their risk
-of an event, and genetic studies rank genes according to their
-difference in expression across samples. A common research question is
-where to stop, i.e., to decide the maximal significant rank.  In the
-three examples this would correspond to the number of significantly
-associated predictor variables, the number of patients at high risk of
-an event, and the number of genes that are worth to pursue in further
-experiments, respectively.
+Ranking of items or results is common in scientific research and
+ranked lists occur naturally as the result of many statistical
+applications. Regression methods rank predictor variables according to
+magnitude of their association with an outcome, prediction models rank
+subjects according to their risk of an event, and genetic studies rank
+genes according to their difference in gene expression levels across
+samples.
 
-In this article we describe some new breakthrough tools for measuring
-agreement across a set of lists which soon will enter the state-of-the
-art. The methods should be useful whenever there are multiple rankings
-of the same list. The idea is to define agreement based on the ranks
-of the first k elements in each list.
+When several rankings of the same items are available, a common
+research question is to what extent they agree. In particular, is it
+possible to identify an optimal rank until which the lists agree on
+the items?  A typical situation arises in high-dimensional genomics
+studies when several analysis methods are applied to rank a list of
+genes according to their association with a phenotype, treatment
+effect or other outcome. A measure of agreement of gene rankings
+obtained by different methods could often help to identify which genes
+are worth to pursue in further experiments.
+
+Several approaches exist to measure the ``distance'' between two
+ranked lists. Of these, Kendall's $\tau$ \citep{kendall1948} and
+Spearman’s footrule $\rho$ \citep{Spearman1910} are among the most
+well-known, but these measures do not distinguish between agreement in
+the top versus towards the bottom of the lists and the provide a
+single measure of the overall distance. \citet{Shieh1998} proposed a
+weighted version of Kendall's $\tau$ where each pair of rankings can
+be assigned different weights, and \citet{Yilmaz2008} proposed the
+$\tau_{ap}$ which places higher emphasis on the top of the lists.
+Spearman’s footrule uses the ranks of the variables for calculation of
+the distance and the use of ranks are also employed in the $M$ measure
+of \citet{Bar-Ilan2006} where the reciprocal rank differences are used
+to calculate the similarity measure.
+
+Other recent approaches consider the intersection of lists as the basis
+for a similarity measure. However, simple intersection also places
+equal weights on all depths of the list and therefore \citet{Fagin2003} and
+\citet{Webber2010} proposed weighted intersections which put more
+emphasis on the top of the lists. One proposal is denoted the average
+overlap (AO) and used for comparison later on. Specifically,
+\citet{Webber2010} define their rank-biased overlap (RBO) by weighting
+with a converging series to ensure that the top is weighted higher
+than the potentially non-informative bottom of the lists.
+It is possible to use the existing methods to calculate agreement of
+lists until a given depth, i.e., limited to the $d$ items of each
+list. However, the interpretation may not be straightforward,
+especially in the case of more than two lists, and they may not
+accommodate partial rankings.
+
+In this article we introduce sequential rank agreement for measuring
+agreement among (partially) ranked lists.  The general idea is to
+define agreement based on the sequence of ranks from the first $d$
+elements in each list. As agreement metric we adapt the limits of
+agreement known from agreement between quantitative variables
+\citep{alt:bland:1983,Carstensen2010}. Our proposed approach allows us
+to compare more than two lists simultaneously, it provides a dynamic
+measure of agreement as a function of the depth in the lists, places
+high weight on the top of the list, accommodates censored/incomplete
+lists of varying lengths, and has a natural interpretation that
+directly relates to the ranks. Graphical illustration of sequential
+rank agreement potentially allows us to infer a change-point, i.e., a
+list depth where a change in the agreement of the lists occur but we
+also provide randomization-based graphical tools to compare the
+observed rank agreement to the expected agreement found in
+non-informative data.  In this sense our approach is a combination and
+generalization of some of the ideas of \citet{Carterette2009} and
+\citet{Boulesteix2009}. The former compares two rankings based on the
+distance between them as measured by a multivariate Gaussian
+distribution and the latter presents an overview of approaches for
+aggregation of ranked lists including bootstrap and leave-one-out
+jackknife approaches.
+
+The manuscript is organized as follows: In the next section we define
+sequential rank agreement for multiple ranked lists and discuss how to
+handle incomplete/censored/partial lists. In section 3 we present and
+discuss approaches to evaluate the results obtained from sequential
+rank agreement. Finally we apply the sequential rank agreement to two
+Danish ovarian cancer studies before we discuss the findings along
+with possible extensions. The approaches presented in this manuscript
+are available in the R package \texttt{SuperRanker}.
 
@@
 
-Consider a set of $P$ different items $X=\{X_1,\dots,X_P\}$. An
-ordered list is a permutation function, $R: \{1,\dots,P\}\to
-\{1,\dots,P\}$, such that $R(X_p)$ is the rank of item $X_p$ in the
-list. The inverse mapping $\pi=R^{-1}$ assigns to rank
-$r\in\{1,\dots,P\}$ the item $\pi(r)$ found at that rank. The methods
-described below work for a set of $L$ lists $R_1,\dots,R_L$,
-$L\geq2$. We denote $\pi_l=R_l^{-1}$ for the corresponding inverse
-mappings. Panels (a) and (b) of Table~\ref{tab:example} show a
-schematic example of these mappings.
+Consider a set of $P$ different items $X=\{X_1,\dots,X_P\}$ and a
+ranking function $R: \{X_1,\dots,X_P\}\to \{1,\dots,P\}$, such that
+$R(X_p)$ is the rank of item $X_p$. The inverse mapping $R^{-1}$ gives
+the item $R^{-1}(r)$ that was assigned to rank $r\in\{1,\dots,P\}$. An
+ordered list is the realization of a ranking function $R$ applied to a
+data sample $X$ and we let $R_1(X),\dots,R_L(X)$, $L\geq2$, denote
+ordered lists obtained from $L$ different ranking functions applied to
+the same data $X$. We will use the same notation for the equivalent
+scenario where we have ordered lists obtained from applying the same
+ranking function to $L$ samples from the same population, i.e., for
+ranked lists $R(X^1), \ldots, R(X^L)$.  Panels (a) and (b) of
+Table~\ref{tab:example} show a schematic example of these
+mappings. Thus if $R_l^{-1}(1)=X_{34}$ then item $X_{34}$ is ranked
+first in list $l$ and similarly $R_l(X_{34})=1$.
 
 \begin{table}[tb]
-  \caption{Example set of ranked lists.
-(a) shows the ranked list of items for each of three lists, (b) presents the ranks obtained by each item in each of the three lists and (c) shows the cumulative set of items up to a given depth in the three lists.}
+  \caption{Example set of ranked lists. (a) shows the ranked list of
+    items for each of three lists, (b) presents the ranks obtained by
+    each item in each of the three lists and (c) shows the cumulative
+    set of items up to a given depth in the three lists.}
 \begin{center}
-  \begin{subtable}{4cm}%
+\footnotesize{
+  \begin{subtable}{4.5cm}%
     \caption{}
@@
 
-The agreement of the lists regarding the rank given to an item can be
+The agreement of the lists regarding the rank given to an item $X_p$ can be
 measured by
 \begin{equation}
-  A(p) = f(R_1(p), \ldots, R_L(p)),
+  A(X_p) = f(R_1(X_p), \ldots, R_L(X_p))
 \end{equation}
-for a distance function $f$. Throughout this paper we will use the
-sample standard error as our function $f$ and hence use
-$$A(p) = \sqrt{\frac{\sum_{i=1}^L (R_i(p) - \bar{R}(p))^2}{L-1}},
-$$
-but other choices could be made (see the discussion). The sample
-standard error has an interpretation as the average distance of the
-individual rankings of the lists from the average ranking.
-
-We now describe what is exemplified in Panel (c) of Table
-\ref{tab:example} and how it can be used to define \emph{sequential
-  rank agreement}. For an integer $1\le d\le P$ we define the unique
-set of items found in the $L$ top $d$ parts of the lists, i.e., the
+for a measurable non-negative real-valued distance function $f$.
+For an integer $1\le d\le P$ we define the unique set of items found
+by merging the first $d$ elements of each of the $L$ lists, i.e., the
 set of items ranked less than or equal to $d$ in any of the lists:
 \begin{equation}
-S_d = \{\pi_l(r) ; r \leq d, l = 1, \ldots, L \}.
+S_d = \{R_l^{-1}(r) ; r \leq d, l = 1, \ldots, L \},
 \end{equation}
-The \emph{sequential rank agreement} is the pooled standard deviation
-of the items found in the set $S_d$:
+which is exemplified in Panel (c) of Table \ref{tab:example}.
+We then define the \emph{sequential rank agreement} as pooled expected agreement of the items
+found in the set $S_d$:
 \begin{equation}
-\textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)A(p)^2}{(L-1)|S_d|}}.
+\textrm{sra}(d)= \frac{\sum_{\{p \in S_d\}}(L-1)\E[A(X_p)]}{(L-1)|S_d|}\label{def:sra}
 \end{equation}
+where $|S_d|$ refers to the cardinality of the set $S_d$, and the expectation
+is taken with respect to the distribution of the rankings of $X_p$ accross the $L$ lists.
+Values of \textrm{sra} close to zero suggests that the lists agree while larger
+values suggest disagreement. Just like in method comparison studies we
+prefer smaller limits of agreement because that means that the
+differences between the methods is small. If the ranked lists are
+identical then the value of sequential rank agreement will be zero for
+all depths $d$.
 
-\subsection{All lists fully observed}
-We shall start by the simplest case where all $L$ lists are fully observed.
+%Note that the terms that are part of \eqref{def:sra} are not
+%independent since every list contains the ranks from 1 to $P$ exactly
+%once. As a consequence we simply use the pooled standard deviation as
+%a measure of the average distance between rankings and we refrain from
+%using traditional distributional results for this measure.
 
-\subsection{Analysis of top $k$ lists}
-Not uncommon for lists to be
+Throughout the rest of this paper we will use the variance to quantify the
+agreement of the rankings and therefore define
+\begin{align}
+  A(X_p) = (R_i(X_p) -  \E[R_i(X_p)])^2
+\end{align}
+For each item, the function $A$ has an interpretation as the expected
+euclidean distance of the individual rankings from the expected ranking over all the lists, and
+it corresponds to the same measure that is used in method comparison studies to compute the
+limits of agreement \citep{alt:bland:1983}. The sequential rank agreement then becomes equivalent
+to the pooled variance of the items found in $S_d$, and it can be interpreted as
+the average distance of the individual rankings of the lists from the
+average ranking for each of the items we have seen until depth $d$.
 
-censored
+Given a sample of $1,\ldots,L$ lists and estimator of estimated sequatial rank agreement is given by
+\begin{equation}
+\widehat{\textrm{sra}}_L(d)= \frac{\sum_{\{p \in S_d\}}(L-1)\widehat{A_L}(X_p)}{(L-1)|S_d|}\label{def:sra}
+\end{equation}
+where
+\begin{align}
+\widehat{A}_L(X_p) = \frac{1}{L-1}\sum_{i=1}^L (R_i(X_p) - \overline{R}(X_p))^2, \quad \overline{R}_L(X_p) = \frac{1}{L}\sum_{i=1}^L R_i(X_p)
+\end{align}
+is the sample variance of the rank of item $X_p$. For an easier interpretation we suggest to plot $\widehat{\textrm{sra}}_L(d)^{1/2}$ corresponding to the pooled standard deviation.
 
-Let $\Lambda_l, l=1, \ldots, L$ be the set of items found in list $l$
-so $\Lambda_l$ is the top $k_l$ list of items from list $l$ where $k_l
-= |\Lambda_l|$. Note that we observe the top $k$ items for each of the
-$L$ lists if $k_1=\cdots=k_L=k$. For censored lists the rank function
-becomes
+\subsection{Agreement among fully observed lists}
+\label{sec:amfol}
+
+The simplest case occurs when all $L$ lists are fully observed, \ie,
+we have observed the rank of all $P$ items for all $L$ lists. Fully
+observed ranked lists are common and arise, for example, when
+different statistical analysis methods are applied to a single dataset
+to produce lists of predictors ranked according to their importance,
+or if the same analysis method is applied to data from different
+populations.
+
+With fully observed lists we can plot the sequential rank agreement
+\eqref{def:sra} as a function of depth $d$. An example is seen in top
+panels of Figure~\ref{fig:example1} where four analysis methods were
+used to rank 3051 gene expression values measured on 38 tumor mRNA
+samples in order to classify between acute lymphoblastic leukemia
+(ALL) and acute myeloid leukemia (AML)
+\citep{Golub1999}. Preprocessing of the gene expression data was done
+as described in \citet{Dudoit2002} and the four different analysis
+approaches were: marginal two-sample $t$ tests, marginal logistic
+regression analyses, elastic net logistic regression
+\citep{friedman2010regularization}, and marginal maximum information
+content correlations (MIC) \citep{Reshef2011}. For the first two
+methods, the genes were ranked according to minimum $p$ value, for
+logistic regression the genes were ordered by size of the
+corresponding coefficients (after standardization), and MIC was
+ordered by absolute correlation which resulted in the top rankings
+seen in Table~\ref{tab1}.
+
+The sequential rank agreement plotted in Figure~\ref{fig:example1}
+shows how many ranks apart we on average expect the genes found in the
+top $d$ part of the lists to be for each depth $d$.  It can be seen
+that the sequential rank agreement limits are better towards the top
+of the lists (smaller values on the $y$ axis) corresponding to
+\emph{better} agreement than in the tails of the lists. It is clear
+from the curve in Figure~\ref{fig:example1} that there is a
+substantial drop in agreement even after the first depth. Thus, if we
+were to restrict attention to a small set of predictors then we would
+focus on predictors 2124, 829, and 378 as seen in Table~\ref{tab1}.
+
+% latex table generated in R 3.2.2 by xtable 1.7-4 package
+% Thu Mar 31 16:50:36 2016
+\begin{table}[tb]
+\centering
+\caption{Top 10 list of ranked results from the Golub data. Numbers indicate the predictor/gene for the given ranking and method} 
+\label{tab1}
+\begin{tabular}{rrrrr}
+  \hline
+Ranking & T & LogReg & ElasticNet & MIC \\ 
+  \hline
+1 & 2124 & 2124 & 829 & 378 \\ 
+  2 & 896 & 896 & 2124 & 829 \\ 
+  3 & 2600 & 829 & 2198 & 896 \\ 
+  4 & 766 & 394 & 1995 & 1037 \\ 
+  5 & 829 & 766 & 808 & 2124 \\ 
+  6 & 2851 & 2670 & 523 & 808 \\ 
+  7 & 703 & 2939 & 1042 & 108 \\ 
+  8 & 2386 & 2386 & 2489 & 515 \\ 
+  9 & 2645 & 1834 & 1665 & 2670 \\ 
+  10 & 2002 & 378 & 849 & 2600 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{figure}[htbp]
+   \begin{center}
+ \includegraphics[width=.49\textwidth]{paper-fig1a}
+ \includegraphics[width=.49\textwidth]{paper-fig2a}
+ \includegraphics[width=.49\textwidth]{paper-fig1b}
+ \includegraphics[width=.49\textwidth]{paper-fig2b}
+ \end{center}
+ \caption{Left panels: Sequential rank agreement for 4 different
+   analysis methods applied to the 3051 genes in the Golub data /black
+   line). Right
+   panels: Corresponding sequential rank agreement for the same data
+   but where only the top 20 ranked items are available. The blue and
+   red areas correspond to the independent and randomized reference
+   hypothesis areas, respectively. The bottom plots are identical to
+   the corresponding top plots but have been zoomed in on the first
+   part of the $x$ axis.}
+ \label{fig:example1}
+ \end{figure}
+
+ If the lists agree on all items then the sequential rank agreement is
+ zero for all depths. Generally, if there are changes in the level of
+ rank agreement then this suggests that there are sets of items that
+ are ranked similarly in all lists and other sets of items that have
+ been assigned vastly different ranks in the lists. In gene
+ association studies we generally expect the agreement among the lists
+ to be better towards the top of the lists and worse towards the end
+ of the lists. In this case the sequential rank agreement curve will
+ start at a low level and then increase until it levels off as in
+ Figure~\ref{fig:example1}.
+
+\subsection{Analysis of partial/censored lists}
+
+Incomplete or partial lists are also a common occurrence. They arise
+for example in case of missing data (items), when comparing top $d$
+list results from publications, or when some methods only rank a
+subset of the items. For example, penalized regression based on the
+Lasso provides a sparse set of predictors that have non-zero
+coefficients. There is no obvious ordering of the set of predictors
+whose coefficient has been shrunk to zero and thus we end up with a
+partial ordering. Incomplete lists also occur when the analyst decides
+to censor all lists at depth $d$ such that the measurement of
+agreement is restricted to the first $d$ elements of each list, or
+when the analyst only wants to consider the ranking of items that have
+been found to be statistically significant.
+
+Sequential rank agreement can be generalized to partial lists in the
+following way. Let $\Lambda_l\subset X$ be the subset of $d_l$ items
+that have been ranked highest in list $l$.  The case where all lists are
+censored at the same depth $d$ corresponds to $d_1=\cdots=d_L=d$. For
+incomplete/censored lists the rank function becomes
 \begin{equation}
-\tilde R_l(p) = \left\{\begin{array}{cl} \{\pi_l^{-1}(p)\} & \text{ for } p\in \Lambda_l \\ 
-\{k_l+1,\dots,P\} & \text{ for } p \not\in \Lambda_l\end{array}\right.
+  \widetilde R_l(X_p) = \left\{\begin{array}{cl} \{R_l(X_p)\} & \text{ for } X_p\in \Lambda_l \\
+      \{d_l+1,\dots,P\} & \text{ for } X_p \not\in \Lambda_l\end{array}\right.
 \end{equation}
-where we only know that the rank for the unobserved items in list $l$ must be larger than the largest rank observed in that list.
+where we only know that the rank for the unobserved items in list $l$
+must be larger than the largest rank observed in that list.
 
-In the case of censored lists it is sufficient (FIXME: requires
-argument) to look at depths where we have corresponding observations
-so the largest rank we should consider will be
+The agreement, $A(X_p)$, cannot be computed directly for all
+predictors in the presence of censored lists because the exact rank
+for some items will be unknown. Also, recall that the rankings within
+a single list are not independent since each rank must appear exactly
+once in each list. Thus, we cannot simply assign the same number
+(e.g., the mean of the unassigned ranks) to the censored items since
+that would result in less variation of the ranks and hence less
+variation of the agreement, and it would artificially introduce a
+(downward) bias of agreement for items that are censored in multiple
+lists.
+
+Instead we randomize the ranks $\{d_{l}+1,\dots,P\}$ to the items that
+do not occur in list $\Lambda_l$. One realization of the $L$ rankings
+of the set $X$ is obtained by randomizing the missing items of each
+list. By randomizing a large number of times we can compute
+\eqref{def:sra} for each realization, and then compute the sequential
+rank agreement as the pointwise (for each depth) average of the rank
+agreements. The algorithm is described in detail in
+Algorithm~\ref{sra-algorithm}.
+
+\begin{algorithm}
+\caption{Sequential rank agreement algorithm for censored lists}
+\label{sra-algorithm}
+\begin{algorithmic}[1]
+  \Procedure{Censored rank agreement}{} \State Let $B$ be the number
+  of permutations to use \For{each $b \in B$} \For{each censored list
+    $l \in L$}
+  \State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1.7cm}{Permute
+    the unassigned ranks, $\{d_l+1, \ldots, P\}$, and assign them
+    randomly to the items \emph{not} found in the list, i.e.,
+    $\Lambda^\complement=X\setminus\Lambda_l$. Combine the
+    result with $\Lambda_l$ to fill out the list.}
+\EndFor
+\State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1cm}{Let sra($b$) be the sequential rank agreement computed from the
+filled out lists.}
+\EndFor
+\State Return element-wise averages across all $B$ permutations of sra($b$).
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+The proposed approach is based on two assumptions: 1) that the most
+interesting items are found in the top of the lists, and 2) that the
+censored rankings provide so little information that it is reasonable
+to assume that they can be represented by a random order. The first
+assumption is justifiable because we have already accepted that it is
+reasonable to rank the items in the first place. The second assumption
+is fair in the light of the first assumption provided that we have a
+``sufficiently large'' part of the top of the lists available.
+
+When the two assumptions are satisfied then it is clear that the
+interesting part of the sequential rank agreement curves is restricted
+to the depths where the number of censored items is
+low. Generally, without additional prior knowledge about the
+distribution of the censored items it seems reasonable to restrict the
+attention of the sequential rank agreement to depths smaller than
+$\max(d_1, \ldots, d_L)$ where at least one list provides an
+uncensored item.
+
+Like for fully observed lists we generally expect the sequential rank
+agreement to start low and then increase unless the lists are
+completely unrelated (in which case the sequential rank agreement will
+be constant at a high level) or if the lists mostly agree on the
+ranking (in which case the sequential rank agreement will also be
+constant but at a low level). For partially ranked lists we also
+expect a change-point around the depth where the lists are
+censored. This is an artefact stemming from the fact that we assume
+that the remainder of the lists can be replaced by a simple
+permutation of the missing items.
+
+\section{Evaluating sequential rank agreement}
+
+To interpret the sequential rank agreement values we propose two
+different benchmark values corresponding to two different
+hypotheses. We wish to determine if we observe better agreement than
+what would be expected if there were no relevant information available
+in the data.
+
+The first reference hypothesis is
+\begin{eqnarray*}
+H_0  & : &  \text{The list rankings correspond to complete randomly}\\
+       &  & \text{permuted lists},
+\end{eqnarray*}
+which not only assumes that there is no information in the data on
+which the rankings are based but also that the methods are completely
+independent.
+
+Alternatively, we can remove the restriction on the independence among
+the methods by only requiring that there is no information in the
+ranking but that the rankings are all based on applying the
+method/approaches to the same data
+\begin{eqnarray*}
+\widetilde H_0 & :&  \text{The list rankings are based on data containing}\\
+& &   \text{no association to the outcome.}
+\end{eqnarray*}
+$H_0$ is a quite unrealistic null hypothesis but we can easily obtain
+realizations from that null hypothesis simply by permuting the items
+within each list and then computing the sequential rank agreement for
+the permuted lists. In the fully observed case each experiment
+contains $L$ lists of random permutations of the items in $X$. For the
+censored/partially observed case we first permute the items
+$X_1,\dots,X_P$ and then censor list $l$ at $d_l$. The sequential rank
+agreement curve from the the original lists can then be compared to,
+say, the pointwise 95\% quantiles of the observed rank agreements
+obtained under $H_0$.
+
+To obtain the distribution under $\widetilde H_0$ the idea is to
+repeat the ranking procedures for unassociated data many times.  Thus,
+we first permute the outcome variable of the data set on which the
+rankings are based. This removes any association between the predictor
+variables and the outcome. Then we apply the methods to the permuted
+data set to generate $L$ new rankings and subsequently compute the
+sequential rank agreement. Note that we only permute the outcomes and
+preserve the structure of the candidate predictors. The randomization
+approach requires that we have the original data available and as such
+it may not be possible to evaluate $\widetilde H_0$ in all situations.
+
+If the sequential rank agreement for the original data lies
+substantially below the distribution of the sequential rank agreements
+obtained under either $H_0$ or $\widetilde H_0$ then this suggests that
+the original ranked lists agree \emph{more} than what we would expect
+in data with no information, and therefore that the information in the
+lists is significantly more in agreement than what would be expected.
+
+Figure~\ref{fig:example1} shows the empirical distributions of
+sequential rank agreement under $H_0$ and $\widetilde H_0$ each based
+on $400$ permutations of the Golub data described in Section
+\ref{sec:amfol}. Not surprisingly, the sequential rank agreement under
+$\widetilde H_0$ is lower than the sequential rank agreement under
+$H_0$ because the four methods used to rank the data ($t$ test,
+logistic regression, elastic net, and MIC) generally tend to identify
+(and rank) similar predictors even if there are only spurious
+associations. The two bottom panels in Figure~\ref{fig:example1} also
+indicate that the observed sequential rank agreement (the black line)
+is better than what would be expected by chance for data that contain
+no information since it lies below the reference areas. However, the
+censored data also suggests that there may be at most 1 or 2 ranked
+items towards the top of the lists that yield a result better than
+what would be expected (the bottom-right plot).
+
+It is important to stress that neither $H_0$ nor $\widetilde H_0$ are
+related to questions regarding the association between the outcome and
+the predictors in the data set. Both hypotheses are purely considering
+how the rankings agree in a situation where there is no relevant
+information available in the data used for creating the rankings.
+
+\section{Change-point analysis}
+
+In many applications it is of interest to estimate a list depths which
+satisfies a change point criterion. In this section we investigate the
+theoretical properties of our proposed method for this specific
+task. As Hall and Schimek, we work with an infinite set of lists
+$\mathcal{L}$ and study the asymptotic behaviour for
+$L\to\infty$. However, we do not allow the length of the lists to
+change with $L$ as in most applications the length of the lists is
+fixed [FIXME: for finite $L$ are there only finitely many forskelige
+lists (aber dass stoert keinen grossen geist, siger karlsson)]. First,
+we define the parameter of interest. For any given threshold value $q$
+a change-point $d_0(q)$ in the list agreement is defined by
+\begin{equation*}
+  d_0=\textrm{arginf}_{d\in\{1,\dots,L\}}\textrm{sra}(d)>q.
+\end{equation*}
+The value of $q$ could for example be the agreement of randomly
+permuted lists [FIXME: there is only one SRA value for infinitely many
+random permutations, no skygge!].
+
+The theorem shows consistency of the plug-in estimate of
+\begin{equation*}
+\hat{d_0} = \textrm{arginf}_{d\in\{1,\dots,L\}}\widehat{\textrm{sra}}(d)>q
+\end{equation*}
+[FIXME: $\textrm{sra}$ -> $\widehat{\textrm{sra}}$ in rest of article?]
+
+\begin{theorem}
+  $\hat{d_0}\overset{L\to\infty}{\longrightarrow}d_0$
+
+\begin{proof}
+Proof: We rewrite the square of the sequential rank agreement \eqref{def:sra} as
 \begin{equation}
-d \leq \max(k_1, \ldots, k_L).
+(\textrm{sra}(d))^2= \frac 1{\sum_{p=1}^P \sum_{l=1}^L 1\{R_l(X_p)\le d\}}\sum_{p=1}^P(1-\prod_{l=1}^L 1\{R_l(X_p)>d\}) A(X_p)^2
 \end{equation}
+[FIXME: Andreas details the proof with Cauchy-Schwarz.]
+[FIXME: Det er lidt kedelig for saadan en fin en som Andreas: Skal theoremet og bevis ikke vaere for en general agreement metric, ikke nodvendigvis den i equation (2), og kan vi charakterisere hvad den skal opfylde?]
+\end{proof}
+\end{theorem}
 
-We cannot directly compute $A(p)$ for all predictors because we only observe a censored version of $\tilde R$ for some of the lists.
-Instead we assume that the rank assigned to predictor $p$ in list $l$ is uniformly distributed among the ranks that have been unassigned for list $l$.
-The ranks are clearly not independent since each of the lists essentially contains full set of ranks
+\section{Comparison to average overlap}
 
+The average overlap is a measure that is widely used in comparisons of
+two fully observed ranked lists \citep{Fagin2003,Webber2010} and the
+idea behind the average overlap closely resembles the sequential rank
+agreement. In this section we present an extension of the average
+overlap to more that two lists and compare this to sequential rank
+agreement.
+
+The overlap among $L$ lists observed until depth $d$ is defined as the
+proportion of items that are found in all $L$ lists compared to the
+possible number of items found in all lists observed until depth $d$:
 \begin{equation}
-\tilde A(p) = \frac{\sum_{r_1; r_1\in \tilde R_1(p)}  \cdots \sum_{r_L; r_L\in \tilde R_L(p)} A(p)}{\prod_l |\tilde R_l(p)|}
+  O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}. \label{def:overlap}
 \end{equation}
+Typically the overlap is only considered for pairwise comparisons
+\citep{Bar-Ilan2006,Boulesteix2009} but formula \eqref{def:overlap}
+is easily extended to the situation where $L>2$ as shown above.
 
-FIXME: If instead of running through all elements of $\tilde R_1(p)$
-one would use the average rank in $\tilde R_1(p)$ we would end up with
-a too small variance.
+The average overlap at depth $d$ is defined as the average of the
+overlaps until depth $d$,
+\begin{equation}
+AO(d) = \frac1d\sum_{i=1}^d O(d).
+\end{equation}
+This definition directly ensures that further emphasis is put on items
+in the top of the lists since their item specific overlap contributes
+to the calculation of average overlap at higher depths.
 
-\section{Benchmarks}
+There are three main drawbacks of the average overlap method: First it
+is highly sensitive to single items in the top of one the lists.  The
+average overlap profile may change dramatically depending on the
+presence or absence of similar items in the beginning of the
+lists. For example, if all lists have the same item at the top then
+the average overlap attains its maximum at 1, but if just one of the
+lists differ then the overage overlap starts at its minimum of
+0. Secondly, when the number of lists, $L$, is high then it might be
+difficult to obtain a non-zero overlap (and hence a non-zero average
+overlap) at the beginning of the lists because the overlap will be
+zero until an item is present in all $L$ lists. Finally, the average
+overlap has an interpretation in terms of moving averages of
+percentages (of $L$ sets) which is somewhat less intuitive than our
+proposed sequential rank distance.
 
-\section{Applications}
+In the following we compare the average overlap and the sequential
+rank agreement using the analysis of the Golub data
+\citep{Golub1999}. We consider a typical application of the average
+overlap and thus restrict attention to only two gene rankings obtained
+by the $t$ tests and logistic regression. These two marginal modeling
+approaches should result in very similar gene rankings. The first two
+columns of Table~\ref{tab1} show the actual top-10 ranking of the
+predictors.
 
-\subsection{Evaluating results from top-$k$ lists}
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=.49\textwidth]{paper-fig-ao1}
+\includegraphics[width=.49\textwidth]{paper-fig-ao2}
+\end{center}
+ \caption{Average overlap (left figure) and sequential rank agreement
+ (right figure) for comparing two different analysis methods
+ (marginal $t$ test and logistic regression)
+    applied to the Golub data. The black lines are based on the full
+    set of predictors while the red lines have the two highest
+    associated predictors removed
+    from the data before computing the average overlap and sequential
+    rank agreement.}
+  \label{fig:case1}
+ \end{figure}
 
-\subsection{Comparing results across different method}
+ The black lines in Figure~\ref{fig:case1} show the average overlap
+ and sequential rank agreement for these data. It is clear from both
+ plots that there is perfect agreement towards the top of the ranked
+ lists: the average overlap is 1 and the sequential rank agreement is
+ zero.  However, if we remove those two items from the data (i.e.,
+ gene/predictor 2124 and 896) and redo the analyses then we get
+ substantially different curves for the average overlap but not for
+ sequential rank agreement (red lines in Figure~\ref{fig:case1}). For
+ the sequential rank agreement we get roughly the same estimate of
+ agreement from rank 3 as we did from the full dataset. This indicates
+ that sequential rank is more robust against small perturbations of
+ the data. For the average overlap the curves completely change when
+ these two items are removed and which would lead to a different
+ conclusion about the agreement among the items from rank 3 downwards.
 
-\subsection{Stability of results}
+\section{Application to ovarian cancer data}
+We now consider an application of the sequential rank agreement to two
+data sets consisting of MALDI-TOF (Matrix-Assisted Laser
+Desorption/Ionization Time Of Flight) mass spectra obtained from blood
+samples from patients with either benign or malignant ovarian
+tumors. The data sets are sub-samples of the Danish MALOVA and DACOVA
+study populations.
 
-Bootstrap across a single method and compare results. Discuss collinearity
+The MALOVA study is a multidisciplinary Danish study on ovarian cancer
+\citep{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
+with an ovarian tumor and referred for surgery from the participating
+departments of gynecology were enrolled continuously from December
+1994 to May 1999. For the purpose of illustration we use a random
+sub-sample of $119$ patients with a total of $58$ patients with
+malignant ovarian cancers as cases and $61$ patients with benign
+ovarian tumors as controls. The DACOVA study is another
+multidisciplinary Danish study on ovarian cancer which included about
+$66\%$ of the female population of Denmark
+\citep{bertelsen1991protocol}. The study aimed to continuously enroll
+all patients that were referred to surgery of an ovarian tumor
+clinically suspected to be cancer during the period from 1984 to
+1990. Similarly, we use a random sub-sample from the DACOVA study of
+$113$ patients with a total of $54$ malignant ovarian cancers and $59$
+benign ovarian tumors/gynecologic disorders.
+
+Each spectrum consists of $49642$ samples over a range of mass-to-charge ratios
+between $800$ to $20000$ Dalton which we downsample on an equidistant grid of
+5000 points by linear interpolation. We then preprocess the downsampled
+spectra individually by first removing the slow-varying baseline intensity
+with the SNIP algorithm \citep{ryan1988snip} followed by a
+normalization with respect to the total ion count. Finally, we standardize
+the $5000$ predictors to have column-wise zero mean and unit variance in
+each data set.
+
+We use the two data sets to illustrate how the sequential rank
+agreement can be applied in two different scenarios. In the first
+scenario we assess the agreement of four different statistical
+classification methods in how they rank the predictors according to
+their importance for distinguishing benign and malignant tumors. In
+the second scenario we assess the agreement among rankings of
+individual predicted risks of having a malignant tumor. The first
+scenario is relevant in the context of biomarker discovery and the
+latter is important e.g., when ranking patients according to immediacy of
+treatment.
+
+Four classification methods are considered: Random Forest
+\citep{breiman2001random} implemented in the R package
+\texttt{randomForest} \citep{liaw2002classification}, logistic Lasso
+\citep{tibshirani1996regression} and Ridge regression
+\citep{segerstedt1992ordinary} both implemented in the R package
+\texttt{glmnet} \citep{friedman2010regularization}, and Partial Least
+Squares Discriminant Analysis (PLS-DA) \citep{boulesteix2004pls}
+implemented in the R package \texttt{caret} \citep{Jed-Wing:2014aa}.
+
+In both scenarios we use the MALOVA data to train the statistical
+models, and in both situations the agreements are assessed with
+respect to perturbations of the training data in the following
+manner. We repeatedly draw a random sub-sample (without replication)
+consisting of 90\% of the MALOVA observations and train the four
+models on each sub-sample. We use 1000 iterations for the sub-sampling
+procedure.
+
+All four methods depend on a tuning parameter. The tuning parameter for
+Lasso and Ridge regression is the degree of penalization, and for PLS-DA
+it is the number of components (the dimensionality of the subspace).
+We estimate these separately for each sub-sample by a 20 times repeated
+5-fold cross-validation procedure. For the Random Forest we grow a fixed
+number of $5000$ trees and let the tuning parameter be the number of
+predictors randomly sampled at each split. We estimate this by a binary
+search with respect to minimizing the Out-of-Bag classification error estimate.
+
+The implementation of Lasso and Ridge regression in the
+\texttt{glmnet} package offers three different cross-validated
+optimization criteria for the penalty parameter: total deviance,
+classification accuracy and area under ROC. We apply all three
+criteria to our data to investigate their effect on the agreements.
+Note also that the Lasso models produce censored lists depending on
+the value of the penalty parameter.
+
+\subsection{Agreement of predictor rankings}
+For each of the four methods, each of the 1000 models trained on
+the 1000 sub-samples of the MALOVA data produces a ranking of the 5000
+predictors according to their importance for discriminating between
+the tumor types. For the Random Forest classifier the predictors are ranked according
+to the Gini index, while for the logistic Lasso and Ridge regression
+models we order by absolute magnitude of the estimated
+regression coefficients. For the PLS-DA model the importance of the
+predictors is based on a weighted sum of the absolute coefficients
+where the weights are proportional to the reduction in the sums of
+squares across the components.
+
+The right panel of Figure \ref{fig:app1} shows the sequential rank
+agreement of the estimated importance of the 5000 predictors.  For
+clarity of presentation we zoom in on the agreement up to list depth
+600. At deeper list depths all agreement curves are approximately constant.
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=.49\textwidth]{pics/riskAgreementPlot}%
+\includegraphics[width=.49\textwidth]{pics/predictorAgreementPlot}
+\end{center}
+\caption{Left panel: Sequential rank agreement of 1000 rankings of the
+  predicted risks of malignant tumor. For each method the different
+  rankings were obtained by first training models in 1000 random
+  sub-samples of the MALOVA data and then predicting the risk of
+  malignant tumor in the 113 DACOVA patients. Right panel: Sequential
+  rank agreement of 1000 rankings of the 5000 predictors. The rankings
+  were obtained from the same 1000 trained models.}
+ \label{fig:app1}
+\end{figure}
+
+For most of the sequential rank agreement curves we see, as expected,
+that they start low, indicating good agreement, followed by an
+increase until they approximately become constant. This has the
+interpretation that the agreement across the different sub-samples is
+higher in the top as compared to the tail of the lists for all these
+classification methods. The change-points where the curves become
+approximately constant are the list depths where the ranks of the
+remaining items become close to uniformly random.
+
+A not expected shape of the agreement curves is seen for the Ridge
+models for all three tuning criteria. They all show higher
+disagreement in the top of the lists followed by a decrease. The
+reason behind this behavior is rather subtle. Looking at the
+distribution of the absolute value of the regression coefficients we see that a large
+proportion of them are numerically very close to zero and have almost
+equal absolute value. This is a general feature of the Ridge models in
+this data set and seen for all the 1000 trained models.
+This implies that when predictors are ranked according to the
+magnitude of their coefficients, their actual order becomes more
+uncertain and more close to a random permutation. This problem can be
+alleviated by truncating all predictors with absolute coefficient values
+below a given threshold thereby introducing an artificial censoring of the lists.
+For the Ridge models tuned with the deviance criterion, Figure \ref{fig:app2} (left) shows the
+the sequential rank agreement where for each of the 1000 trained
+models the predictors were artificially censored when their absolute
+coefficient value was lower than the $0.1\%$ quantile of the 5000
+absolute coefficient values. The curve was calculated
+using Algorithm \ref{sra-algorithm} with $B=1000$ and $P=5000$. The corresponding
+curve from Figure \ref{fig:app1} (right panel) is shown for comparison.
+Even though the number of censored predictors is very small compared to the total
+number of predictors, the effect on the sequential rank agreement is
+substantial and with the artificial censoring the shape of the curves
+is as expected, starting low and then increasing.
+
+Looking at the agreement curves for the Lasso models in Figure \ref{fig:app1}
+(right) we clearly see the effect of the sparsity inducing penalization giving
+rise to censored lists. These curves were similarly calculated using \ref{sra-algorithm}
+and $1000$ random permutations. Under the deviance optimization criterion the
+median number of non-zero coefficients was 33 (range 16 to 50) and for the
+class accuracy criterion 14 (range 4 to 56). These values correspond to
+the list depths where the agreement curves become constant as a result of
+the subsequent censoring.
+
+\subsection{Agreement of individual risk predictions}
+\label{sec:airp}
+To assess the stability of the individual risk predictions we apply
+the predictors from the DACOVA data set to each of the models. The
+predicted probabilities are then ranked in decreasing order such that the
+patients with the highest risk of a malignant tumor appears in the top
+of the list. Figure \ref{fig:app1} (left) shows the sequential rank
+agreement separately for each method, based on the 1000 risk predictions
+obtained from the models trained in the same $1000$ random sub-samples
+of the MALOVA data.
+
+Most curves start low and then increase indicating higher agreement
+among high risk patients. This is expected if we rank the individuals
+according to highest risk of disease. However, it is also expected
+that individuals with very low risk also show high agreement. In this
+case we order the patients according to (high) risk prediction but we
+could essentially also have reversed the order to identify the
+patients that have low risk prediction.
+
+An exception is the risk prediction agreement for the Lasso tuned with the AUC criterion which
+shows very low agreement among the high values of the predicted
+risks. The reason is that optimizing the penalty parameter with
+respect to the AUC criterion tends to favor a very high penalty value
+causing only a single predictor to be selected in each of the 1000
+iterations. This results in a lack of generalizability to the DACOVA
+data which gives rise to the higher disagreement in the predicted risks.
+In the extreme case where the penalty becomes so high
+that none of the predictors are selected by the Lasso, the sequential
+rank agreement for the predicted probabilities becomes undefined since
+all the ranks will be ties.
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=.49\textwidth]{pics/ridgeCensoring}%
+\includegraphics[width=.49\textwidth]{pics/aucPlot}
+\end{center}
+\caption{Left panel: Sequential rank agreement for Ridge regression
+  obtained by artificially censoring predictors when their absolute
+  coefficient values are lower than the 0.1\% quantile. Right panel:
+  Box plots of AUC values across the $1000$ sub-samples with respect
+  to the known class labels of the DACOVA data.}
+ \label{fig:app2}
+\end{figure}
+
+Comparing the left and right panels of Figure \ref{fig:app1} it can
+further be seen that some of the methods show better agreement with
+respect to the predicted probabilities than for ranking the importance
+of the predictors and vice versa. Ridge regression shows higher agreement across training
+sets for the risk predictions than PLS-DA, and PLS-DA shows higher
+agreement for predictor importance than Ridge regression.
+
+Lasso shows similar agreement for ranking the risk predictions than
+PLS-DA (unless with the AUC criterion), and poorer agreement for ranking
+predictors. This reason for the latter is the high auto-correlation
+between the intensities in the mass spectra which leads to
+collinearity issues in the regression models. It is well-known that
+variable selection with the Lasso penalty does not preform very well
+when the predictors are highly correlated. The collinearity does,
+however, not affect the agreement of the risk predictions, as seen on
+the left panel of Figure \ref{fig:app1}, since it is not so important
+which specific variable that gets selected from a group of highly correlated
+predictors when the purpose is risk predictions.
+
+It appears that Ridge regression tuned with the AUC criterion achieves
+the best performance with respect to the stability of ranking the
+individual predicted risk probabilities. It must, however, be stressed
+that the sequential rank agreement in this application is only
+concerned with the agreement of the risk predictions across
+sub-samples and not with the actual accuracy of the risk
+predictions. Thus, we also computed the AUC values for the different models
+based on the DACOVA data. The distributions across the $1000$
+sub-samples for a selection of the models is shown in the right panel
+of Figure \ref{fig:app2}.  Here we see that PLS-DA attains the highest
+AUC values with a median value of $0.70$ while the Ridge model with
+the AUC criterion attains a median AUC of $0.49$. This implies that
+while Ridge regression optimized with respect to the AUC criterion
+achieves the best sequential rank agreement, it performs similar to a
+random coin toss with respect to classifying the DACOVA patients.
+In practice both concerns are of importance.
 
 \section{Discussion}
+In this article we address the problem of comparing ranked lists of
+the same items. Our proposed method can handle both the situation
+where the underlying data to generate the ranked lists are available
+and the situation where the only available data is the actual ranked
+lists. In addition, censored ranked lists where only the ranks of the
+top $k$ ranked items are known can be accommodated as well. The
+proposed agreement measure can be interpreted as the average distance
+between an item's rank and the average rank assigned to that item
+across lists. Thus the measure determines how well the lists agree on
+the ranks of a specific item.
 
-Mention/discuss different measures.
+The sequential rank agreement is extremely versatile. We have shown
+that it can be used not only to compare ranked lists of items produced
+from different samples/populations but that it also can be used to
+study the ranks obtained from different analysis methods on the same
+data as well as to evaluate the stability of the ranks from a single
+method by bootstrapping (or sub-sampling) the data repeatedly and
+comparing the ranks obtained from training the models in the
+bootstrapped data.
+
+The sequential rank agreement can be used to determine the depth at
+which the rank agreement becomes too large to be desirable based on
+prior requirements or acceptable differences, or it can be used to
+visually determine when the change in agreement becomes too large.  In
+that regard the investigator can have prior limits on the level of
+agreement that is acceptable.
+
+While the sequential rank agreement is primarily an exploratory tool
+we have suggested two null hypotheses that can be used to evaluate the
+sequential rank agreement obtained. Note that none of the two null
+hypotheses are concerned with the actual ``true ranking'' but are
+purely concerned with consistency/stability of the rankings among the
+lists. As such we cannot determine if the rankings are good but only
+whether they agree. The sequential rank agreement curve can be
+compared visually to the curves obtained under either of the null
+distributions and simple point-wise $p$-values can be obtained for
+each depth by counting the number of sequential rank agreements under
+the null hypothesis that is less than or equal to the observed rank
+agreement. Two simple extensions can be pursued to make more formal
+uniform tests: one would be to use a Kolmogorov-Smirnov-like test and
+use the largest difference between the variance-weighted sequential
+rank agreement curve and the mean null rank agreement curve as a test
+statistic. Alternatively, a change-point analysis could be made on the
+sequential rank agreement in order to determine the depths at which
+there are ``jumps'' in the rank agreement. These jumps would
+correspond to depths for which the agreement among the lists was
+substantially worse and could serve as indicators for when the lists
+no longer agree sufficiently satisfactory.
+
+Finally, we have --- whenever possible --- used all available ranks
+from the lists.  We could choose to censor the rankings of all the
+items that were deemed not to be significant if information on both
+rankings and the corresponding evidence for significance was
+available. That would ensure that there would be put less emphasis on
+the agreement of the non-significant items and it would be easier to
+identify a change in agreement among the items that were deemed to
+be relevant. In our application section we have successfully introduced
+such an artificial censoring for the predictor rankings obtained with
+ridge regression.
+
+We note that the sequential rank agreement is still marred by problems
+that generally apply to ranking of items and/or
+individuals. Collinearity in particular can be a huge problem when
+bootstrapping data or when comparing different analysis methods. For
+example, marginal analyses where each item is analyzed separately will
+assign similar ranks to two highly correlated predictors while methods
+that provide a sparse solution such as the Lasso will just rank one of
+the two predictors high while the other might have a very low rank.
+Thus in such a scenario we would expect low agreement of the rankings
+from Lasso and marginal analyses simply because of the way correlated
+predictors are handled. This is not a shortcoming of the sequential
+rank agreement but is a problem general to all ranked lists.
+
+Another caveat with the way the sequential rank agreement is defined
+is the use of the standard deviation to measure agreement. The
+standard deviation is an integral part of the limits-of-agreement as
+discussed by \citet{alt:bland:1983}. However, the standard deviation
+can also be unstable when the number of observations is low and
+alternative measures such as the median absolute deviance may prove more
+stable in some situations. However, the current definition using the
+standard deviation is analogous to the approach used for agreement in
+method comparison studies so we have used that.
+
+In conclusion we have introduced a method for evaluation of ranked
+(censored) lists that can be easily interpreted and that can be
+applied to a large number of situations.  The method presented here
+can be adapted further by using it to compare and classify statistical
+analysis methods that agree on the rankings they provide or by using
+the rank agreement to optimize a hyper-parameter in, say, elastic net
+regularized regression where the rank agreement is used to determine
+the mixing proportion between the $L_1$ and the $L_2$ penalty.
+%Finally, the proposed method may be adapted (with some additional
+%assumptions) to the situation where the lists are subsets of each
+%other.
+We will be investigating these extensions further in the future.
+
+%\bibliographystyle{rss}
+\bibliographystyle{rss}
+\bibliography{paperref}
 

### 2016-03-31 latex/paper.Rnw
Starting on the asymptotic

--- 2015-08-19 latex/paper.Rnw
@@
 
+\DeclareMathOperator{\E}{\mathbb{E}}
+\DeclareMathOperator{\Var}{Var}
+\DeclareMathOperator{\Cov}{Cov}
+\DeclareMathOperator{\arginf}{arginf}
+
+\usepackage[amsthm,thmmarks]{ntheorem}
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
+\newtheorem{proposition}{Proposition}
+\newtheorem{corollary}{Corollary}
+\newtheorem{lemma}{Lemma}
+\newtheorem{assumption}{Assumption}
+\newtheorem{example}{Example}
+\renewcommand\proofSymbol{\ensuremath{\blacksquare}}
+
 \SweaveOpts{prefix.string=paper, keep.source=TRUE}
@@
 \begin{equation}
-  A(X_p) = f(R_1(X_p), \ldots, R_L(X_p)),
+  A(X_p) = f(R_1(X_p), \ldots, R_L(X_p))
 \end{equation}
-for a non-negative real-valued distance function $f.$ Throughout this
-paper we will use the sample standard error as our function $f$ and
-hence use
-$$A(X_p) = \sqrt{\frac{\sum_{i=1}^L (R_i(X_p) - \bar{R}(X_p))^2}{L-1}},
-$$
-where $\bar{R}(X_p)$ is the average rank assigned to item $X_p$, but
-other possibilities will be discussed below. For each item, the sample
-standard error has an interpretation as the average distance of the
-individual rankings from the average ranking over all the lists, and
-it corresponds to the same measure that is used in method comparison
-studies to compute the limits of agreement \citep{alt:bland:1983}.
-
+for a measurable non-negative real-valued distance function $f$.
 For an integer $1\le d\le P$ we define the unique set of items found
@@
 \end{equation}
-which is exemplified in Panel (c) of Table
-\ref{tab:example}.
-
-The \emph{sequential rank agreement} is the pooled standard
-deviation of the items found in the set $S_d$:
+which is exemplified in Panel (c) of Table \ref{tab:example}.
+We then define the \emph{sequential rank agreement} as pooled expected agreement of the items
+found in the set $S_d$:
 \begin{equation}
-\textrm{sra}(d)= \sqrt{\frac{\sum_{\{p \in
-      S_d\}}(L-1)A(X_p)^2}{(L-1)|S_d|}}, \label{def:sra}
+\textrm{sra}(d)= \frac{\sum_{\{p \in S_d\}}(L-1)\E[A(X_p)]}{(L-1)|S_d|}\label{def:sra}
 \end{equation}
-where $|S_d|$ refers to the cardinality of the set $S_d$. Values of
-\textrm{sra} close to zero suggests that the lists agree while larger
+where $|S_d|$ refers to the cardinality of the set $S_d$, and the expectation
+is taken with respect to the distribution of the rankings of $X_p$ accross the $L$ lists.
+Values of \textrm{sra} close to zero suggests that the lists agree while larger
 values suggest disagreement. Just like in method comparison studies we
@@
 identical then the value of sequential rank agreement will be zero for
-all depths $d$. The sequential rank agreement can be interpreted as
+all depths $d$.
+
+%Note that the terms that are part of \eqref{def:sra} are not
+%independent since every list contains the ranks from 1 to $P$ exactly
+%once. As a consequence we simply use the pooled standard deviation as
+%a measure of the average distance between rankings and we refrain from
+%using traditional distributional results for this measure.
+
+Throughout the rest of this paper we will use the variance to quantify the
+agreement of the rankings and therefore define
+\begin{align}
+  A(X_p) = (R_i(X_p) -  \E[R_i(X_p)])^2
+\end{align}
+For each item, the function $A$ has an interpretation as the expected
+euclidean distance of the individual rankings from the expected ranking over all the lists, and
+it corresponds to the same measure that is used in method comparison studies to compute the
+limits of agreement \citep{alt:bland:1983}. The sequential rank agreement then becomes equivalent
+to the pooled variance of the items found in $S_d$, and it can be interpreted as
 the average distance of the individual rankings of the lists from the
@@
 
-Note that the terms that are part of \eqref{def:sra} are not
-independent since every list contains the ranks from 1 to $P$ exactly
-once. As a consequence we simply use the pooled standard deviation as
-a measure of the average distance between rankings and we refrain from
-using traditional distributional results for this measure.
+Given a sample of $1,\ldots,L$ lists and estimator of estimated sequatial rank agreement is given by
+\begin{equation}
+\widehat{\textrm{sra}}_L(d)= \frac{\sum_{\{p \in S_d\}}(L-1)\widehat{A_L}(X_p)}{(L-1)|S_d|}\label{def:sra}
+\end{equation}
+where
+\begin{align}
+\widehat{A}_L(X_p) = \frac{1}{L-1}\sum_{i=1}^L (R_i(X_p) - \overline{R}(X_p))^2, \quad \overline{R}_L(X_p) = \frac{1}{L}\sum_{i=1}^L R_i(X_p)
+\end{align}
+is the sample variance of the rank of item $X_p$. For an easier interpretation we suggest to plot $\widehat{\textrm{sra}}_L(d)^{1/2}$ corresponding to the pooled standard deviation.
 
@@
 information available in the data used for creating the rankings.
+
+\section{Change-point analysis}
+In many applications it is of interest to estimate a list depths which
+satisfies a change point criterion. In this section we investigate the
+theoretical properties of our proposed method for this specific
+task. As Hall and Schimek, we work with an infinite set of lists
+$\mathcal{L}$ and study the asymptotic behaviour for
+$L\to\infty$. However, we do not allow the length of the lists to
+change with $L$ as in most applications the length of the lists is
+fixed [FIXME: for finite $L$ are there only finitely many forskelige
+lists (aber dass stoert keinen grossen geist, siger karlsson)]. First,
+we define the parameter of interest. For any given threshold value $q$
+a change-point $d^\ast(q)$ in the list agreement is defined by
+\begin{equation}
+  d^\ast(q) = \arginf_{d\in\{1,\dots,P\}}\textrm{sra}(d)>q.
+\end{equation}
+The value of $q$ could for example be the agreement of randomly
+permuted lists [FIXME: there is only one SRA value for infinitely many
+random permutations, no skygge!].
+The corresponding estimated change-point is similarly defined by
+\begin{equation}
+\widehat{d^\ast_L}(q) = \arginf_{d\in\{1,\dots,P\}}\widehat{\textrm{sra}}_L(d)>q
+\end{equation}
+
+We start by showing that $\widehat{\textrm{sra}}_L(d)$ is consistent for $\textrm{sra}(d)$ at each
+$d \in \left\{1,\ldots,P\right\}$ for $L \rightarrow \infty$. The consistency of $\widehat{d^\ast_L}(q)$
+then follows by a simple corollary.
+
+\begin{theorem}
+hej
+\begin{proof}
+We start by noting that we can rewrite the sequential rank agreement in equation \ref{def:sra} as
+\begin{equation}
+\textrm{sra}(d)= \frac 1{\sum_{p=1}^P \sum_{l=1}^L 1\{R_l(X_p)\le d\}}\sum_{p=1}^P(1-\prod_{l=1}^L 1\{R_l(X_p)>d\}) \E[A(X_p)]
+\end{equation}
+
+[FIXME: Andreas details the proof with Cauchy-Schwarz.]
+[FIXME: Det er lidt kedelig for saadan en fin en som Andreas: Skal theoremet og bevis ikke vaere for en general agreement metric, ikke nodvendigvis den i equation (2), og kan vi charakterisere hvad den skal opfylde?]
+\end{proof}
+\label{thm:consistency}
+\end{theorem}
+
+\begin{corollary}
+For fixed $q$ it follows from theorem \ref{thm:consistency}
+  $\widehat{d^\ast_L}(q) \overset{P}{\longrightarrow} d^\ast(q)$
+\begin{proof}
+ hej
+\end{proof}
+\end{corollary}
 

### 2016-03-31 latex/paper.tex
Starting on the asymptotic

@@
 \DeclareMathOperator{\Cov}{Cov}
+\DeclareMathOperator{\arginf}{arginf}
 
@@
 \newtheorem{proposition}{Proposition}
-\newtheorem*{proposition2}{Proposition}
 \newtheorem{corollary}{Corollary}
@@
 % latex table generated in R 3.2.2 by xtable 1.7-4 package
-% Thu Mar 31 16:50:36 2016
+% Thu Mar 31 17:59:03 2016
 \begin{table}[tb]
@@
   3 & 2600 & 829 & 2198 & 896 \\ 
-  4 & 766 & 394 & 1995 & 1037 \\ 
-  5 & 829 & 766 & 808 & 2124 \\ 
-  6 & 2851 & 2670 & 523 & 808 \\ 
-  7 & 703 & 2939 & 1042 & 108 \\ 
-  8 & 2386 & 2386 & 2489 & 515 \\ 
-  9 & 2645 & 1834 & 1665 & 2670 \\ 
-  10 & 2002 & 378 & 849 & 2600 \\ 
+  4 & 766 & 394 & 1665 & 1037 \\ 
+  5 & 829 & 766 & 1920 & 2124 \\ 
+  6 & 2851 & 2670 & 1042 & 808 \\ 
+  7 & 703 & 2939 & 808 & 108 \\ 
+  8 & 2386 & 2386 & 849 & 515 \\ 
+  9 & 2645 & 1834 & 937 & 2670 \\ 
+  10 & 2002 & 378 & 1524 & 2600 \\ 
    \hline
@@
 \section{Change-point analysis}
-
 In many applications it is of interest to estimate a list depths which
@@
 we define the parameter of interest. For any given threshold value $q$
-a change-point $d_0(q)$ in the list agreement is defined by
-\begin{equation*}
-  d_0=\textrm{arginf}_{d\in\{1,\dots,L\}}\textrm{sra}(d)>q.
-\end{equation*}
+a change-point $d^\ast(q)$ in the list agreement is defined by
+\begin{equation}
+  d^\ast(q) = \arginf_{d\in\{1,\dots,P\}}\textrm{sra}(d)>q.
+\end{equation}
 The value of $q$ could for example be the agreement of randomly
@@
 random permutations, no skygge!].
-
-The theorem shows consistency of the plug-in estimate of
-\begin{equation*}
-\hat{d_0} = \textrm{arginf}_{d\in\{1,\dots,L\}}\widehat{\textrm{sra}}(d)>q
-\end{equation*}
-[FIXME: $\textrm{sra}$ -> $\widehat{\textrm{sra}}$ in rest of article?]
+The corresponding estimated change-point is similarly defined by
+\begin{equation}
+\widehat{d^\ast_L}(q) = \arginf_{d\in\{1,\dots,P\}}\widehat{\textrm{sra}}_L(d)>q
+\end{equation}
+
+We start by showing that $\widehat{\textrm{sra}}_L(d)$ is consistent for $\textrm{sra}(d)$ at each
+$d \in \left\{1,\ldots,P\right\}$ for $L \rightarrow \infty$. The consistency of $\widehat{d^\ast_L}(q)$
+then follows by a simple corollary.
 
 \begin{theorem}
-  $\hat{d_0}\overset{L\to\infty}{\longrightarrow}d_0$
-
+hej
 \begin{proof}
-Proof: We rewrite the square of the sequential rank agreement \eqref{def:sra} as
+We start by noting that we can rewrite the sequential rank agreement in equation \ref{def:sra} as
 \begin{equation}
-(\textrm{sra}(d))^2= \frac 1{\sum_{p=1}^P \sum_{l=1}^L 1\{R_l(X_p)\le d\}}\sum_{p=1}^P(1-\prod_{l=1}^L 1\{R_l(X_p)>d\}) A(X_p)^2
+\textrm{sra}(d)= \frac 1{\sum_{p=1}^P \sum_{l=1}^L 1\{R_l(X_p)\le d\}}\sum_{p=1}^P(1-\prod_{l=1}^L 1\{R_l(X_p)>d\}) \E[A(X_p)]
 \end{equation}
+
 [FIXME: Andreas details the proof with Cauchy-Schwarz.]
@@
 \end{proof}
+\label{thm:consistency}
 \end{theorem}
+
+\begin{corollary}
+For fixed $q$ it follows from theorem \ref{thm:consistency}
+  $\widehat{d^\ast_L}(q) \overset{P}{\longrightarrow} d^\ast(q)$
+\begin{proof}
+ hej
+\end{proof}
+\end{corollary}
 

### 2016-03-31 latex/paper.Rnw
more asymptotic

@@
 which is exemplified in Panel (c) of Table \ref{tab:example}.
-We then define the \emph{sequential rank agreement} as pooled expected agreement of the items
-found in the set $S_d$:
-\begin{equation}
-\textrm{sra}(d)= \frac{\sum_{\{p \in S_d\}}(L-1)\E[A(X_p)]}{(L-1)|S_d|}\label{def:sra}
-\end{equation}
+We then define the \emph{sequential rank agreement} as the weighted expected agreement of the items
+found in the set $S_d$: [FIX: we have talked about this before, but does that redundant (L-1) factor help wrt interpretation?]
+\begin{align}
+\textrm{sra}(d)&= \frac{\sum_{\{p \in S_d\}}(L-1)\E[A(X_p)]}{(L-1)|S_d|}\label{def:sra}\\
+   &= \sum_{\{p \in S_d\}}|S_d|^{-1} \E[A(X_p)]
+\end{align}
 where $|S_d|$ refers to the cardinality of the set $S_d$, and the expectation
@@
 \begin{corollary}
-For fixed $q$ it follows from theorem \ref{thm:consistency}
+For fixed $q$ it follows from theorem \ref{thm:consistency} that
 $\widehat{d^\ast_L}(q) \overset{P}{\longrightarrow} d^\ast(q)$
+for $L \rightarrow \infty$.
 \begin{proof}
@@
 Typically the overlap is only considered for pairwise comparisons
-\citep{Bar-Ilan2006,Boulesteix2009} but formula \eqref{def:overlap}
+\citep{Bar-Ilan2006,Boulesteix2009} but formula \ref{def:overlap}
 is easily extended to the situation where $L>2$ as shown above.

### 2016-03-31 latex/paper.tex
more asymptotic

@@
 % latex table generated in R 3.2.2 by xtable 1.7-4 package
-% Thu Mar 31 17:59:03 2016
+% Thu Mar 31 18:58:05 2016
 \begin{table}[tb]
@@
   4 & 766 & 394 & 1665 & 1037 \\ 
-  5 & 829 & 766 & 1920 & 2124 \\ 
-  6 & 2851 & 2670 & 1042 & 808 \\ 
+  5 & 829 & 766 & 1042 & 2124 \\ 
+  6 & 2851 & 2670 & 1920 & 808 \\ 
   7 & 703 & 2939 & 808 & 108 \\ 
@@
   9 & 2645 & 1834 & 937 & 2670 \\ 
-  10 & 2002 & 378 & 1524 & 2600 \\ 
+  10 & 2002 & 378 & 523 & 2600 \\ 
    \hline

### 2016-04-07 latex/paper.Rnw
more asymptotic

@@
 \usepackage{algpseudocode}
+\usepackage{enumitem}
 %\usepackage{geometry}
@@
 We then define the \emph{sequential rank agreement} as the weighted expected agreement of the items
-found in the set $S_d$: [FIX: we have talked about this before, but does that redundant (L-1) factor help wrt interpretation?]
+found in the set $S_d$:
 \begin{align}
-\textrm{sra}(d)&= \frac{\sum_{\{p \in S_d\}}(L-1)\E[A(X_p)]}{(L-1)|S_d|}\label{def:sra}\\
-   &= \sum_{\{p \in S_d\}}|S_d|^{-1} \E[A(X_p)]
+\textrm{sra}(d) &= \sum_{\{p \in S_d\}}|S_d|^{-1} \E[A(X_p)]
 \end{align}
@@
 
-Given a sample of $1,\ldots,L$ lists and estimator of estimated sequatial rank agreement is given by
+Given a sample of $1,\ldots,L$ lists an estimator of the sequatial rank agreement is given by
 \begin{equation}
@@
 \end{align}
-is the sample variance of the rank of item $X_p$. For an easier interpretation we suggest to plot $\widehat{\textrm{sra}}_L(d)^{1/2}$ corresponding to the pooled standard deviation.
+is the sample variance of the rank of item $X_p$.
+For an easier interpretation we suggest to plot $\widehat{\textrm{sra}}_L(d)^{1/2}$ corresponding to the pooled standard deviation.
+The consistency of $\widehat{\textrm{sra}}_L$ as $L \rightarrow \infty$ will be discussed further in section \ref{sec:chgpoint}.
 
@@
 \subsection{Analysis of partial/censored lists}
-
 Incomplete or partial lists are also a common occurrence. They arise
@@
 
-\section{Change-point analysis}
+\section{Change-point analysis}\label{sec:chgpoint}
 In many applications it is of interest to estimate a list depths which
@@
 task. As Hall and Schimek, we work with an infinite set of lists
-$\mathcal{L}$ and study the asymptotic behaviour for
-$L\to\infty$. However, we do not allow the length of the lists to
-change with $L$ as in most applications the length of the lists is
-fixed [FIXME: for finite $L$ are there only finitely many forskelige
-lists (aber dass stoert keinen grossen geist, siger karlsson)]. First,
-we define the parameter of interest. For any given threshold value $q$
+$\mathcal{L}$ and study the asymptotic behaviour for $L\to\infty$.
+%However, we do not allow the length of the lists to
+%change with $L$ as in most applications the length of the lists is
+%fixed [FIXME: for finite $L$ are there only finitely many forskelige
+%lists (aber dass stoert keinen grossen geist, siger karlsson)]. 
+First, we define the parameter of interest. For any given threshold value $q$
 a change-point $d^\ast(q)$ in the list agreement is defined by
@@
 
-We start by showing that $\widehat{\textrm{sra}}_L(d)$ is consistent for $\textrm{sra}(d)$ at each
-$d \in \left\{1,\ldots,P\right\}$ for $L \rightarrow \infty$. The consistency of $\widehat{d^\ast_L}(q)$
-then follows by a simple corollary.
+We start by showing that $\widehat{\textrm{sra}}_L$ is consistent for $\textrm{sra}$ for $P,L \rightarrow \infty$ and $\frac{P}{L} \rightarrow 0$. 
+The consistency of $\widehat{d^\ast_L}(q)$ then follows by a simple corollary. The proof requires the following assumptions.
+
+\begin{assumption}
+We assume that:
+\begin{enumerate}[label=\Alph*)]
+\item{hej}
+\item{med}
+\item{dig}
+\end{enumerate}
+\end{assumption}
 
 \begin{theorem}
-hej
+Under the assumptions it follows that $\widehat{\textrm{sra}}_L \overset{P}{\longrightarrow} \textrm{sra}$ for $L,P\rightarrow \infty$ and $\frac{P}{L} \rightarrow 0$.
 \begin{proof}
 We start by noting that we can rewrite the sequential rank agreement in equation \ref{def:sra} as
-\begin{equation}
-\textrm{sra}(d)= \frac 1{\sum_{p=1}^P \sum_{l=1}^L 1\{R_l(X_p)\le d\}}\sum_{p=1}^P(1-\prod_{l=1}^L 1\{R_l(X_p)>d\}) \E[A(X_p)]
-\end{equation}
-
-[FIXME: Andreas details the proof with Cauchy-Schwarz.]
-[FIXME: Det er lidt kedelig for saadan en fin en som Andreas: Skal theoremet og bevis ikke vaere for en general agreement metric, ikke nodvendigvis den i equation (2), og kan vi charakterisere hvad den skal opfylde?]
+\begin{align}
+\textrm{sra}(d) &= \frac 1{\sum_{p=1}^P \sum_{l=1}^L 1\{R_l(X_p)\le d\}}\sum_{p=1}^P(1-\prod_{l=1}^L 1\{R_l(X_p)>d\}) \E[A(X_p)]\\
+    &= \frac{1}{L} \sum_{l=1}^L \sum_{p=1}^P S_l(p)
+\end{align}
+where
+\begin{align}
+  S_l(p) &= ?
+\end{align}
+Maybe we can use the LLN for triangular arrays?
 \end{proof}
@@
 \begin{equation}
-  O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}. \label{def:overlap}
+  O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}\label{def:overlap}
 \end{equation}

### 2016-04-07 latex/paper.tex

@@
 % latex table generated in R 3.2.2 by xtable 1.7-4 package
-% Thu Mar 31 18:58:05 2016
+% Thu Apr  7 20:04:21 2016
 \begin{table}[tb]
@@
   5 & 829 & 766 & 1042 & 2124 \\ 
-  6 & 2851 & 2670 & 1920 & 808 \\ 
-  7 & 703 & 2939 & 808 & 108 \\ 
+  6 & 2851 & 2670 & 808 & 808 \\ 
+  7 & 703 & 2939 & 1920 & 108 \\ 
   8 & 2386 & 2386 & 849 & 515 \\ 
-  9 & 2645 & 1834 & 937 & 2670 \\ 
-  10 & 2002 & 378 & 523 & 2600 \\ 
+  9 & 2645 & 1834 & 523 & 2670 \\ 
+  10 & 2002 & 378 & 937 & 2600 \\ 
    \hline

### 2016-04-08 latex/paper.tex
Intro, forfatterliste, bibtex

@@
 \documentclass[12pt,a4paper]{article}
+\pdfoutput=1
 \usepackage[utf8]{inputenc}
@@
 \author{Andreas Kryger Jensen}
-\author{Kasper Brink-Jensen}
+%\author{Kasper Brink-Jensen}
 \affil{Biostatistics, University of Copenhagen}
@@
 
+Very recently, \citet{hall:schi:2012} proposed a method for comparing
+pairwise rankings and derived the asymptotic distribution of the
+endpoint where the two ranked lists no longer are in agreement. Their
+approach was based on anchoring one of the two lists and subsequently
+generating a sequence of 0s and 1s depending on whether the ranks in
+the second list was close to the rank from the anchored
+list. \citet{samp:verd:2013} followed up on this idea for pairwise
+comparison of lists but used penalties based on a truncated geometric
+probability instead of a 0-1 process and they evaluated the
+distribution of the endpoint of agreement by computational approaches.
+The asymptotic distribution in the \citet{hall:schi:2012} paper is
+based on letting the number of \emph{lists} increase to infinity which
+is a situation that is only relevant in special cases, and the
+simulation-based null distribution approach of \citet{samp:verd:2013}
+is guaranteed to work also for small samples.
+
 In this article we introduce sequential rank agreement for measuring
@@
 define agreement based on the sequence of ranks from the first $d$
-elements in each list. As agreement metric we adapt the limits of
+items in each list. As agreement metric we adapt the limits of
 agreement known from agreement between quantitative variables
-\citep{alt:bland:1983,Carstensen2010}. Our proposed approach allows us
-to compare more than two lists simultaneously, it provides a dynamic
+\citep{alt:bland:1983,Carstensen2010} but any measure for agreement
+could be essentially be used. Our proposed approach allows us to
+compare more than two lists simultaneously, it provides a dynamic
 measure of agreement as a function of the depth in the lists, places
-high weight on the top of the list, accommodates censored/incomplete
-lists of varying lengths, and has a natural interpretation that
-directly relates to the ranks. Graphical illustration of sequential
-rank agreement potentially allows us to infer a change-point, i.e., a
-list depth where a change in the agreement of the lists occur but we
-also provide randomization-based graphical tools to compare the
-observed rank agreement to the expected agreement found in
-non-informative data.  In this sense our approach is a combination and
-generalization of some of the ideas of \citet{Carterette2009} and
-\citet{Boulesteix2009}. The former compares two rankings based on the
-distance between them as measured by a multivariate Gaussian
-distribution and the latter presents an overview of approaches for
-aggregation of ranked lists including bootstrap and leave-one-out
-jackknife approaches.
+high weight on items at the top of the list, accommodates partially
+observed lists of varying lengths, and has a natural interpretation
+that directly relates to the ranks.  Graphical illustration of
+sequential rank agreement potentially allows us to infer a
+change-point, i.e., a list depth where a change in the agreement of
+the lists occur but we also provide randomization-based graphical
+tools to compare the observed rank agreement to the expected agreement
+found in non-informative data.  In this sense our approach is a
+combination and generalization of some of the ideas of
+\citet{Carterette2009} and \citet{Boulesteix2009}. The former compares
+two rankings based on the distance between them as measured by a
+multivariate Gaussian distribution and the latter presents an overview
+of approaches for aggregation of ranked lists including bootstrap and
+leave-one-out jackknife approaches. We show the asymptotic
+distribution of the endpoint of agreement and discuss how to infer the
+distribution in small-sample situations using computational
+methods. This enables us to make inferences about the endpoint of
+agreement even for situations where there is no actual changepoint
+present but just a gradual decline in agreement.
 
@@
 with possible extensions. The approaches presented in this manuscript
-are available in the R package \texttt{SuperRanker}.
+are available in the R package \texttt{SuperRanker}. [XXX check dette
+passer, submit pakke]
 
@@
 
-% latex table generated in R 3.2.2 by xtable 1.7-4 package
-% Thu Apr  7 20:04:21 2016
+% latex table generated in R 3.2.3 by xtable 1.8-2 package
+% Fri Apr  8 10:22:58 2016
 \begin{table}[tb]
@@
   4 & 766 & 394 & 1665 & 1037 \\ 
-  5 & 829 & 766 & 1042 & 2124 \\ 
-  6 & 2851 & 2670 & 808 & 808 \\ 
-  7 & 703 & 2939 & 1920 & 108 \\ 
+  5 & 829 & 766 & 1920 & 2124 \\ 
+  6 & 2851 & 2670 & 1042 & 808 \\ 
+  7 & 703 & 2939 & 808 & 108 \\ 
   8 & 2386 & 2386 & 849 & 515 \\ 
-  9 & 2645 & 1834 & 523 & 2670 \\ 
-  10 & 2002 & 378 & 937 & 2600 \\ 
+  9 & 2645 & 1834 & 937 & 2670 \\ 
+  10 & 2002 & 378 & 1995 & 2600 \\ 
    \hline

### 2016-04-19 latex/superranker-manuscript.tex
new version of section 2 in new file superranker-manuscript.tex

--- 2016-04-08 latex/paper.tex
@@
 \begin{document}
-\input{paper-concordance}
 
@@
 
+% chunk 1
+
 \section{Methods}
@@
 the item $R^{-1}(r)$ that was assigned to rank $r\in\{1,\dots,P\}$. An
-ordered list is the realization of a ranking function $R$ applied to a
-data sample $X$ and we let $R_1(X),\dots,R_L(X)$, $L\geq2$, denote
-ordered lists obtained from $L$ different ranking functions applied to
-the same data $X$. We will use the same notation for the equivalent
-scenario where we have ordered lists obtained from applying the same
-ranking function to $L$ samples from the same population, i.e., for
-ranked lists $R(X^1), \ldots, R(X^L)$.  Panels (a) and (b) of
+ordered list is the realization of a ranking function $R$ applied to
+the set of items $X$. In all what follows we condition on the set of
+items and consider the ranking function to be a random variable. Thus,
+let $R_1(X),\dots,R_L(X)$, $L\geq2$, be a sample of $L$ independent
+identically distributed draws from an unknown probability distribution
+function $Q$. One aim is then to test how much $Q$ resembles the
+uniform distribution which assigns probability $(1/P)^P$ to each of
+the $P^P$ different rankings.
+
+\bigskip\hrule HOV: er det virkelig equivalent ogsaa mht vores nye theoretiske
+ideer? Og bruger vi det nogen steder?  enten skal det forklares med et
+konkret eksempel eller ud?  We will use the same notation for the
+equivalent scenario where we have ordered lists obtained from applying
+the same ranking function to $L$ samples from the same population,
+i.e., for ranked lists $R(X^1), \ldots, R(X^L)$.  
+\hrule\bigskip
+
+Panels (a) and (b) of
 Table~\ref{tab:example} show a schematic example of these
@@
 
-The agreement of the lists regarding the rank given to an item $X_p$ can be
-measured by
+The agreement of the lists regarding the rank given to an item $X_p$
+can be measured by the expected variance across the lists:
 \begin{equation}
-  A(X_p) = f(R_1(X_p), \ldots, R_L(X_p))
+  {A}(X_p) = \sum_{r=1}^{P^P} \{r(X_P)-\E_Q R(X_P)\}^2 Q(r)
 \end{equation}
-for a measurable non-negative real-valued distance function $f$.
-For an integer $1\le d\le P$ we define the unique set of items found
-by merging the first $d$ elements of each of the $L$ lists, i.e., the
-set of items ranked less than or equal to $d$ in any of the lists:
+where $\E_Q R(X_p)=\sum_{r=1}^{P^P} r(X_p)Q(r)$. The empirical
+counterpart is
+
+\begin{align}
+\widehat{A}_L(X_p) = \frac{1}{L-1}\sum_{i=1}^L (R_i(X_p) - \overline{R}(X_p))^2, \quad \overline{R}_L(X_p) = \frac{1}{L}\sum_{i=1}^L R_i(X_p).
+\end{align}
+For each item, the function $\widehat{A}_L$ has an interpretation as the expected
+Euclidean distance of the individual rankings from the expected
+ranking over the $L$ lists, and it corresponds to the same measure
+that is used in method comparison studies to compute the limits of
+agreement \citep{alt:bland:1983}.
+
+\bigskip\hrule HOV
+Vi kunne sige her: Alternative distance functions (f non-negativ real valued) are discussed in section ??
+\hrule\bigskip 
+% \begin{equation}
+  % \widehat{A_L}(X_p) = f(R_1(X_p), \ldots, R_L(X_p))
+% \end{equation}
+% for a measurable non-negative real-valued distance function $f$.
+For an integer $1\le d\le P$ we define the expected set of unique
+items found by merging the first $d$ elements across the possible
+lists:
 \begin{equation}
-S_d = \{R_l^{-1}(r) ; r \leq d, l = 1, \ldots, L \},
+S_d = \{X_p; \left(\sum_{r=1}^{P^P} I\{r(X_p)\le d\} Q(r)\right) >0\}.
+\end{equation}
+The empirical counterpart the set of unique items ranked less than or
+equal to $d$ in any of the $L$ lists:
+\begin{equation}
+\widehat{S}_{d,L} = \{X_p; \left(\sum_{l=1}^L I\{R_l(X_p)\le d\}\right)>0\},
 \end{equation}
 which is exemplified in Panel (c) of Table \ref{tab:example}.
-We then define the \emph{sequential rank agreement} as the weighted expected agreement of the items
-found in the set $S_d$:
-\begin{align}
-\textrm{sra}(d) &= \sum_{\{p \in S_d\}}|S_d|^{-1} \E[A(X_p)]
-\end{align}
-where $|S_d|$ refers to the cardinality of the set $S_d$, and the expectation
-is taken with respect to the distribution of the rankings of $X_p$ accross the $L$ lists.
-Values of \textrm{sra} close to zero suggests that the lists agree while larger
-values suggest disagreement. Just like in method comparison studies we
-prefer smaller limits of agreement because that means that the
-differences between the methods is small. If the ranked lists are
-identical then the value of sequential rank agreement will be zero for
-all depths $d$.
-
+
+We then define the \emph{sequential rank agreement} as the weighted
+expected agreement of the items found in the set $S_d$:
+\begin{equation}
+  \textrm{sra}(d) = \frac{\sum_{\{p \in S_d\}}A(X_p)}{|S_d|}\label{def:sra}
+  % & =\sum_{p=1}^P \sum_{l=1}^L I\{R_l(X_p)\le d\}>0\}\notag
+\end{equation}
+ where $|S_d|$ is the cardinality of the set $S_d$.
+The empirical counterpart is given by
+\begin{equation}
+  \widehat{\textrm{sra}}_L(d) = \frac{\sum_{\{p \in \widehat{S}_{d,L}\}}(L-1)\widehat{A_L}(X_p)}{(L-1)|\widehat {S}_{d,L}|}.\label{def:sra}
+                                % \intertext{where}
+                                % |S_d| & =\sum_{p=1}^P \sum_{l=1}^L I\{R_l(X_p)\le d\}>0\}\notag
+\end{equation}
+% where is the cardinality of the set $\widehat{S}_{d,L}$.
+
+Values of \textrm{sra} close to zero suggests that the lists agree
+while larger values suggest disagreement.
+
+Just like in method comparison studies we prefer smaller limits of
+agreement because that means that the differences between the methods
+is small. If the ranked lists are identical then the value of
+sequential rank agreement will be zero for all depths $d$.
 %Note that the terms that are part of \eqref{def:sra} are not
@@
 
-Throughout the rest of this paper we will use the variance to quantify the
-agreement of the rankings and therefore define
-\begin{align}
-  A(X_p) = (R_i(X_p) -  \E[R_i(X_p)])^2
-\end{align}
-For each item, the function $A$ has an interpretation as the expected
-euclidean distance of the individual rankings from the expected ranking over all the lists, and
-it corresponds to the same measure that is used in method comparison studies to compute the
-limits of agreement \citep{alt:bland:1983}. The sequential rank agreement then becomes equivalent
-to the pooled variance of the items found in $S_d$, and it can be interpreted as
-the average distance of the individual rankings of the lists from the
+The sequential rank agreement then becomes equivalent to the pooled
+variance of the items found in $S_d$, and it can be interpreted as the
+average distance of the individual rankings of the lists from the
 average ranking for each of the items we have seen until depth $d$.
 
-Given a sample of $1,\ldots,L$ lists an estimator of the sequatial rank agreement is given by
+\subsection{Population parameter}
+
+Given a sample of $1,\ldots,L$ lists an estimator of the sequential
+rank agreement is given by
 \begin{equation}
@@
 
-% latex table generated in R 3.2.3 by xtable 1.8-2 package
-% Fri Apr  8 10:22:58 2016
-\begin{table}[tb]
-\centering
-\caption{Top 10 list of ranked results from the Golub data. Numbers indicate the predictor/gene for the given ranking and method} 
-\label{tab1}
-\begin{tabular}{rrrrr}
-  \hline
-Ranking & T & LogReg & ElasticNet & MIC \\ 
-  \hline
-1 & 2124 & 2124 & 829 & 378 \\ 
-  2 & 896 & 896 & 2124 & 829 \\ 
-  3 & 2600 & 829 & 2198 & 896 \\ 
-  4 & 766 & 394 & 1665 & 1037 \\ 
-  5 & 829 & 766 & 1920 & 2124 \\ 
-  6 & 2851 & 2670 & 1042 & 808 \\ 
-  7 & 703 & 2939 & 808 & 108 \\ 
-  8 & 2386 & 2386 & 849 & 515 \\ 
-  9 & 2645 & 1834 & 937 & 2670 \\ 
-  10 & 2002 & 378 & 1995 & 2600 \\ 
-   \hline
-\end{tabular}
-\end{table}
+% chunks 2,3,4,5,6,7
 
@@
 predictors.
+
+% chunks 8,9
 

### 2016-05-20 latex/superranker-manuscript.tex
closing in on asymptotic

@@
 \usepackage{calc}
-\usepackage[noae]{Sweave}
+%\usepackage[noae]{Sweave}
 \usepackage{authblk}
@@
 
+\newcommand{\nn}{\nonumber}
+
+\newcommand{\mcomment}[1]{\textcolor{red}{\textbf{#1}}}
+
 \DeclareMathOperator{\E}{\mathbb{E}}
@@
 \DeclareMathOperator{\Cov}{Cov}
-\DeclareMathOperator{\arginf}{arginf}
+\newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
 
@@
 function $Q$. One aim is then to test how much $Q$ resembles the
-uniform distribution which assigns probability $(1/P)^P$ to each of
-the $P^P$ different rankings.
+uniform distribution which assigns probability $(P!)^{-1}$ to each of
+the $P!$ different rankings.
+Panels (a) and (b) of
+Table~\ref{tab:example} show a schematic example of these
+mappings. Thus if $R_l^{-1}(1)=X_{34}$ then item $X_{34}$ is ranked
+first in list $l$ and similarly $R_l(X_{34})=1$.
 
@@
 \hrule\bigskip
-
-Panels (a) and (b) of
-Table~\ref{tab:example} show a schematic example of these
-mappings. Thus if $R_l^{-1}(1)=X_{34}$ then item $X_{34}$ is ranked
-first in list $l$ and similarly $R_l(X_{34})=1$.
 
@@
 The agreement of the lists regarding the rank given to an item $X_p$
-can be measured by the expected variance across the lists:
+can be measured by the variance across the lists
 \begin{equation}
-  {A}(X_p) = \sum_{r=1}^{P^P} \{r(X_P)-\E_Q R(X_P)\}^2 Q(r)
+  {A}(X_p) = \sum_{r=1}^{P!} \{r(X_P)-\E_Q R(X_P)\}^2 Q(r)
 \end{equation}
-where $\E_Q R(X_p)=\sum_{r=1}^{P^P} r(X_p)Q(r)$. The empirical
-counterpart is
-
+where $\E_Q R(X_p)=\sum_{r=1}^{P!} r(X_p)Q(r)$ and $Q$ is a probability mass function
+on the space of all permutations of lists with $P$ elements.
+The empirical counterpart is
 \begin{align}
@@
 
-\bigskip\hrule HOV
-Vi kunne sige her: Alternative distance functions (f non-negativ real valued) are discussed in section ??
-\hrule\bigskip 
-% \begin{equation}
-  % \widehat{A_L}(X_p) = f(R_1(X_p), \ldots, R_L(X_p))
-% \end{equation}
-% for a measurable non-negative real-valued distance function $f$.
 For an integer $1\le d\le P$ we define the expected set of unique
@@
 lists:
-\begin{equation}
-S_d = \{X_p; \left(\sum_{r=1}^{P^P} I\{r(X_p)\le d\} Q(r)\right) >0\}.
-\end{equation}
+\begin{align}
+S(d) &= \left\{X_p; \left(\sum_{r=1}^{P!} 1\left(r(X_p)\le d\right) Q(r)\right) >0\right\}\label{eq:sumset}\\
+     &= \left\{X_p; Q(R(X_p) \leq d) > 0 \right\}\nn
+\end{align}
+where $1(\cdot)$ denotes the indicator function.
 The empirical counterpart the set of unique items ranked less than or
@@
 \begin{equation}
-\widehat{S}_{d,L} = \{X_p; \left(\sum_{l=1}^L I\{R_l(X_p)\le d\}\right)>0\},
+\widehat{S}_{L}(d) = \left\{X_p; \left(\frac{1}{L}\sum_{l=1}^L 1\left(R_l(X_p)\le d\right)\right)>0\right\},
 \end{equation}
 which is exemplified in Panel (c) of Table \ref{tab:example}.
-
 We then define the \emph{sequential rank agreement} as the weighted
-expected agreement of the items found in the set $S_d$:
+expected agreement of the items found in the set $S(d)$:
+\begin{align}
+  \textrm{sra}(d) = \frac{1}{|S(d)|}\sum_{p \in S(d)}A(X_p)\label{def:sra}
+\end{align}
+where $|S(d)|$ is the cardinality of the set $S(d)$.
+The empirical counterpart is equivalently given by
 \begin{equation}
-  \textrm{sra}(d) = \frac{\sum_{\{p \in S_d\}}A(X_p)}{|S_d|}\label{def:sra}
-  % & =\sum_{p=1}^P \sum_{l=1}^L I\{R_l(X_p)\le d\}>0\}\notag
+  \widehat{\textrm{sra}}_L(d) = \frac{\sum_{\{p \in \widehat{S}_{L}(d)\}}(L-1)\widehat{A_L}(X_p)}{(L-1)|\widehat {S}_{L}(d)|}.\label{def:sraEst}
 \end{equation}
- where $|S_d|$ is the cardinality of the set $S_d$.
-The empirical counterpart is given by
-\begin{equation}
-  \widehat{\textrm{sra}}_L(d) = \frac{\sum_{\{p \in \widehat{S}_{d,L}\}}(L-1)\widehat{A_L}(X_p)}{(L-1)|\widehat {S}_{d,L}|}.\label{def:sra}
-                                % \intertext{where}
-                                % |S_d| & =\sum_{p=1}^P \sum_{l=1}^L I\{R_l(X_p)\le d\}>0\}\notag
-\end{equation}
-% where is the cardinality of the set $\widehat{S}_{d,L}$.
 
@@
 while larger values suggest disagreement.
-
 Just like in method comparison studies we prefer smaller limits of
@@
 sequential rank agreement will be zero for all depths $d$.
-%Note that the terms that are part of \eqref{def:sra} are not
-%independent since every list contains the ranks from 1 to $P$ exactly
-%once. As a consequence we simply use the pooled standard deviation as
-%a measure of the average distance between rankings and we refrain from
-%using traditional distributional results for this measure.
-
-The sequential rank agreement then becomes equivalent to the pooled
-variance of the items found in $S_d$, and it can be interpreted as the
+The sequential rank agreement is  equivalent to the pooled
+variance of the items found in $S(d)$, and it can be interpreted as the
 average distance of the individual rankings of the lists from the
 average ranking for each of the items we have seen until depth $d$.
-
-\subsection{Population parameter}
-
-Given a sample of $1,\ldots,L$ lists an estimator of the sequential
-rank agreement is given by
-\begin{equation}
-\widehat{\textrm{sra}}_L(d)= \frac{\sum_{\{p \in S_d\}}(L-1)\widehat{A_L}(X_p)}{(L-1)|S_d|}\label{def:sra}
-\end{equation}
-where
-\begin{align}
-\widehat{A}_L(X_p) = \frac{1}{L-1}\sum_{i=1}^L (R_i(X_p) - \overline{R}(X_p))^2, \quad \overline{R}_L(X_p) = \frac{1}{L}\sum_{i=1}^L R_i(X_p)
-\end{align}
-is the sample variance of the rank of item $X_p$.
-For an easier interpretation we suggest to plot $\widehat{\textrm{sra}}_L(d)^{1/2}$ corresponding to the pooled standard deviation.
-The consistency of $\widehat{\textrm{sra}}_L$ as $L \rightarrow \infty$ will be discussed further in section \ref{sec:chgpoint}.
+For an easier interpretation we suggest to plot $\widehat{\textrm{sra}}_L(d)^{1/2}$ 
+corresponding to the pooled standard deviation.
 
@@
 With fully observed lists we can plot the sequential rank agreement
-\eqref{def:sra} as a function of depth $d$. An example is seen in top
+\ref{def:sra} as a function of depth $d$. An example is seen in top
 panels of Figure~\ref{fig:example1} where four analysis methods were
@@
 
+\subsection{Comparison to average overlap}
+The average overlap is a measure that is widely used in comparisons of
+two fully observed ranked lists \citep{Fagin2003,Webber2010} and the
+idea behind the average overlap closely resembles the sequential rank
+agreement. In this section we present an extension of the average
+overlap to more that two lists and compare this to sequential rank
+agreement.
+
+The overlap among $L$ lists observed until depth $d$ is defined as the
+proportion of items that are found in all $L$ lists compared to the
+possible number of items found in all lists observed until depth $d$:
+\begin{align}
+  O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}\label{def:overlap}
+\end{align}
+Typically the overlap is only considered for pairwise comparisons
+\citep{Bar-Ilan2006,Boulesteix2009} but formula \ref{def:overlap}
+is easily extended to the situation where $L>2$ as shown above.
+
+The average overlap at depth $d$ is defined as the average of the
+overlaps until depth $d$,
+\begin{equation}
+AO(d) = \frac1d\sum_{i=1}^d O(d).
+\end{equation}
+This definition directly ensures that further emphasis is put on items
+in the top of the lists since their item specific overlap contributes
+to the calculation of average overlap at higher depths.
+
+There are three main drawbacks of the average overlap method: First it
+is highly sensitive to single items in the top of one the lists.  The
+average overlap profile may change dramatically depending on the
+presence or absence of similar items in the beginning of the
+lists. For example, if all lists have the same item at the top then
+the average overlap attains its maximum at 1, but if just one of the
+lists differ then the overage overlap starts at its minimum of
+0. Secondly, when the number of lists, $L$, is high then it might be
+difficult to obtain a non-zero overlap (and hence a non-zero average
+overlap) at the beginning of the lists because the overlap will be
+zero until an item is present in all $L$ lists. Finally, the average
+overlap has an interpretation in terms of moving averages of
+percentages (of $L$ sets) which is somewhat less intuitive than our
+proposed sequential rank distance.
+
+In the following we compare the average overlap and the sequential
+rank agreement using the analysis of the Golub data
+\citep{Golub1999}. We consider a typical application of the average
+overlap and thus restrict attention to only two gene rankings obtained
+by the $t$ tests and logistic regression. These two marginal modeling
+approaches should result in very similar gene rankings. The first two
+columns of Table~\ref{tab1} show the actual top-10 ranking of the
+predictors.
+
+% chunks 8,9
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=.49\textwidth]{paper-fig-ao1}
+\includegraphics[width=.49\textwidth]{paper-fig-ao2}
+\end{center}
+ \caption{Average overlap (left figure) and sequential rank agreement
+ (right figure) for comparing two different analysis methods
+ (marginal $t$ test and logistic regression)
+    applied to the Golub data. The black lines are based on the full
+    set of predictors while the red lines have the two highest
+    associated predictors removed
+    from the data before computing the average overlap and sequential
+    rank agreement.}
+  \label{fig:case1}
+ \end{figure}
+
+ The black lines in Figure~\ref{fig:case1} show the average overlap
+ and sequential rank agreement for these data. It is clear from both
+ plots that there is perfect agreement towards the top of the ranked
+ lists: the average overlap is 1 and the sequential rank agreement is
+ zero.  However, if we remove those two items from the data (i.e.,
+ gene/predictor 2124 and 896) and redo the analyses then we get
+ substantially different curves for the average overlap but not for
+ sequential rank agreement (red lines in Figure~\ref{fig:case1}). For
+ the sequential rank agreement we get roughly the same estimate of
+ agreement from rank 3 as we did from the full dataset. This indicates
+ that sequential rank is more robust against small perturbations of
+ the data. For the average overlap the curves completely change when
+ these two items are removed and which would lead to a different
+ conclusion about the agreement among the items from rank 3 downwards.
+
 \section{Evaluating sequential rank agreement}
-
 To interpret the sequential rank agreement values we propose two
@@
 
-\section{Change-point analysis}\label{sec:chgpoint}
+\section{Asymptotics}\label{sec:chgpoint}
 In many applications it is of interest to estimate a list depths which
@@
 theoretical properties of our proposed method for this specific
-task. As Hall and Schimek, we work with an infinite set of lists
-$\mathcal{L}$ and study the asymptotic behaviour for $L\to\infty$.
-%However, we do not allow the length of the lists to
-%change with $L$ as in most applications the length of the lists is
-%fixed [FIXME: for finite $L$ are there only finitely many forskelige
-%lists (aber dass stoert keinen grossen geist, siger karlsson)].
-First, we define the parameter of interest. For any given threshold value $q$
+task. As Hall and Schimek \mcomment{REF}, we work with an infinite set of lists
+and study the asymptotic behaviour for $L\to\infty$.
+However, we do not allow the length of the lists to
+change with $L$ as in most applications the length of the lists is fixed.
+
+First, we define the parameter of interest. For any given threshold value $q$, 
 a change-point $d^\ast(q)$ in the list agreement is defined by
 \begin{equation}
-  d^\ast(q) = \arginf_{d\in\{1,\dots,P\}}\textrm{sra}(d)>q.
+  d^\ast(q) = \argmin_{d\in\{1,\dots,P\}}\textrm{sra}(d)>q.
 \end{equation}
 The value of $q$ could for example be the agreement of randomly
-permuted lists [FIXME: there is only one SRA value for infinitely many
-random permutations, no skygge!].
-The corresponding estimated change-point is similarly defined by
+permuted lists. The corresponding estimated change-point is similarly defined by
 \begin{equation}
-\widehat{d^\ast_L}(q) = \arginf_{d\in\{1,\dots,P\}}\widehat{\textrm{sra}}_L(d)>q
+\widehat{d^\ast_L}(q) = \argmin_{d\in\{1,\dots,P\}}\widehat{\textrm{sra}}_L(d)>q
 \end{equation}
 
-We start by showing that $\widehat{\textrm{sra}}_L$ is consistent for $\textrm{sra}$ for $P,L \rightarrow \infty$ and $\frac{P}{L} \rightarrow 0$.
+We start by showing that $\widehat{\textrm{sra}}_L$ is a consistent estimator of $\textrm{sra}$ for $L \rightarrow \infty$.
 The consistency of $\widehat{d^\ast_L}(q)$ then follows by a simple corollary. The proof requires the following assumptions.
@@
 \begin{enumerate}[label=\Alph*)]
-\item{hej}
-\item{med}
-\item{dig}
+\item{$\E_Q[R(X_p)^4]$ is finite for all $p \in 1,\ldots,P$}.
+\item{$|S(d)| > 0$ uniformly in $d$}.
 \end{enumerate}
 \end{assumption}
+We note that assumption $B$ implies that for every list depth $d$ there must be a non-zero probability of ... under the probability measure $Q$.
 
 \begin{theorem}
-Under the assumptions it follows that $\widehat{\textrm{sra}}_L \overset{P}{\longrightarrow} \textrm{sra}$ for $L,P\rightarrow \infty$ and $\frac{P}{L} \rightarrow 0$.
-\begin{proof}
-We start by noting that we can rewrite the sequential rank agreement in equation \ref{def:sra} as
-\begin{align}
-\textrm{sra}(d) &= \frac 1{\sum_{p=1}^P \sum_{l=1}^L 1\{R_l(X_p)\le d\}}\sum_{p=1}^P(1-\prod_{l=1}^L 1\{R_l(X_p)>d\}) \E[A(X_p)]\\
-    &= \frac{1}{L} \sum_{l=1}^L \sum_{p=1}^P S_l(p)
-\end{align}
-where
-\begin{align}
-  S_l(p) &= ?
-\end{align}
-Maybe we can use the LLN for triangular arrays?
-\end{proof}
+Under the assumptions it follows that $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty = o_P(1)$.
 \label{thm:consistency}
@@
 \begin{corollary}
-For fixed $q$ it follows from theorem \ref{thm:consistency} that
+For fixed $q$ it follows from Theorem \ref{thm:consistency} that
 $\widehat{d^\ast_L}(q) \overset{P}{\longrightarrow} d^\ast(q)$
 for $L \rightarrow \infty$.
-\begin{proof}
- hej
-\end{proof}
+\label{cor:consistency}
 \end{corollary}
-
-\section{Comparison to average overlap}
-
-The average overlap is a measure that is widely used in comparisons of
-two fully observed ranked lists \citep{Fagin2003,Webber2010} and the
-idea behind the average overlap closely resembles the sequential rank
-agreement. In this section we present an extension of the average
-overlap to more that two lists and compare this to sequential rank
-agreement.
-
-The overlap among $L$ lists observed until depth $d$ is defined as the
-proportion of items that are found in all $L$ lists compared to the
-possible number of items found in all lists observed until depth $d$:
-\begin{equation}
-  O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}\label{def:overlap}
-\end{equation}
-Typically the overlap is only considered for pairwise comparisons
-\citep{Bar-Ilan2006,Boulesteix2009} but formula \ref{def:overlap}
-is easily extended to the situation where $L>2$ as shown above.
-
-The average overlap at depth $d$ is defined as the average of the
-overlaps until depth $d$,
-\begin{equation}
-AO(d) = \frac1d\sum_{i=1}^d O(d).
-\end{equation}
-This definition directly ensures that further emphasis is put on items
-in the top of the lists since their item specific overlap contributes
-to the calculation of average overlap at higher depths.
-
-There are three main drawbacks of the average overlap method: First it
-is highly sensitive to single items in the top of one the lists.  The
-average overlap profile may change dramatically depending on the
-presence or absence of similar items in the beginning of the
-lists. For example, if all lists have the same item at the top then
-the average overlap attains its maximum at 1, but if just one of the
-lists differ then the overage overlap starts at its minimum of
-0. Secondly, when the number of lists, $L$, is high then it might be
-difficult to obtain a non-zero overlap (and hence a non-zero average
-overlap) at the beginning of the lists because the overlap will be
-zero until an item is present in all $L$ lists. Finally, the average
-overlap has an interpretation in terms of moving averages of
-percentages (of $L$ sets) which is somewhat less intuitive than our
-proposed sequential rank distance.
-
-In the following we compare the average overlap and the sequential
-rank agreement using the analysis of the Golub data
-\citep{Golub1999}. We consider a typical application of the average
-overlap and thus restrict attention to only two gene rankings obtained
-by the $t$ tests and logistic regression. These two marginal modeling
-approaches should result in very similar gene rankings. The first two
-columns of Table~\ref{tab1} show the actual top-10 ranking of the
-predictors.
-
-% chunks 8,9
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.49\textwidth]{paper-fig-ao1}
-\includegraphics[width=.49\textwidth]{paper-fig-ao2}
-\end{center}
- \caption{Average overlap (left figure) and sequential rank agreement
- (right figure) for comparing two different analysis methods
- (marginal $t$ test and logistic regression)
-    applied to the Golub data. The black lines are based on the full
-    set of predictors while the red lines have the two highest
-    associated predictors removed
-    from the data before computing the average overlap and sequential
-    rank agreement.}
-  \label{fig:case1}
- \end{figure}
-
- The black lines in Figure~\ref{fig:case1} show the average overlap
- and sequential rank agreement for these data. It is clear from both
- plots that there is perfect agreement towards the top of the ranked
- lists: the average overlap is 1 and the sequential rank agreement is
- zero.  However, if we remove those two items from the data (i.e.,
- gene/predictor 2124 and 896) and redo the analyses then we get
- substantially different curves for the average overlap but not for
- sequential rank agreement (red lines in Figure~\ref{fig:case1}). For
- the sequential rank agreement we get roughly the same estimate of
- agreement from rank 3 as we did from the full dataset. This indicates
- that sequential rank is more robust against small perturbations of
- the data. For the average overlap the curves completely change when
- these two items are removed and which would lead to a different
- conclusion about the agreement among the items from rank 3 downwards.
 
@@
 
+\appendix
+
+\newpage
+\section{Proof of Theorem \ref{thm:consistency}}
+We start by defining the following cadlag function
+\begin{align}
+  N(p; d) = \sum_{s=1}^p 1\left(Q(R(X_s) \leq d) > 0\right), \quad p = 1,\ldots, P
+\end{align}
+which jumps at a time $p$ if the probability that the rank of $X_p$ is less than or equal to $d$ is bounded away from zero under the probability measure $Q$.
+Similarly we have the empirical counterpart
+\begin{align}
+  \widehat{N}_L(p; d) &= \sum_{s=1}^p 1\left( \frac{1}{L} \sum_{l=1}^L 1(R_l(X_s) \leq d) > 0\right)\\
+        &= \sum_{s=1}^p 1\left(\widehat{Q}_L(R(X_s) \leq d) > 0\right)\nn
+\end{align}
+Since the joint distribution of $\left\{\widehat{N}_L(p;d) : p=1,\ldots,P\right\}$ is completely determined by its finite set of jump times and $\widehat{Q}_L(R(X_p) \leq d) \overset{P}{\longrightarrow} Q(R(X_p) \leq d)$ for every $p$ and $d$ for $L \rightarrow \infty$ by the law of large numbers, it follows that
+\begin{align}
+  \sup_{d \in 1,\ldots,P}\sup_{p \in 1,\ldots, P} \left|\widehat{N}_L(p; d) - N(p; d)\right| = o_P(1)\label{eq:countingConv}
+\end{align}
+
+Recalling the definition of the sequential rank agreement in equation (\ref{def:sra}), it may be rewritten as the following integral with respect to the counting measure induced by $N$
+\begin{align}
+  \textrm{sra}(d) = \int_{1}^P \frac{A(X_p)}{N(P; d)}dN(p; d)
+\end{align}
+where we note that the cardinality of the set $S(d)$ in equation (\ref{eq:sumset}) is equal to the counting process evaluated at $P$.
+The empirical sequential rank agreement is similarly given by
+\begin{align}
+  \widehat{\textrm{sra}}_L(d) &= \int_1^P \frac{\widehat{A_L}(X_p)}{\widehat{N}_L(P; d)}d\widehat{N}_L(p; d) 
+\end{align}
+and by elementary algebra it follows that
+\begin{align}
+  \sup_{d \in 1,\ldots,P}\left|\widehat{\textrm{sra}}_L(d) - \textrm{sra}(d)\right| \leq \sup_{d \in 1,\ldots,P}\left|U_L(d)\right| - \sup_{d \in 1,\ldots,P}\left|V_L(d)\right|
+\end{align}
+where
+\begin{align}
+  U_L(d) &= \int_1^P\frac{A(X_p)}{N(P;d)} d\left(\widehat{N}_L(p; d) - N(p; d)\right)\\
+  V_L(d) &=  \int_1^P \left(\frac{\widehat{A_L}(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right)d\widehat{N}_L(p; d)
+\end{align}
+The conclusion of the proof therefore follows if each of these two terms are uniformly of order $o_P(1)$ in $d$.
+
+Starting with $V_L(d)$, since $\widehat{N}_L$ is of bounded variation it follows that
+\begin{align}
+ V_L(d) &\leq \sup_{p \in 1,\ldots,P}\left|\frac{\widehat{A_L}(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right| \int_1^P \left|d\widehat{N}_L(p; d)\right|\label{eq:secondTerm}
+\end{align}
+Looking at $\widehat{A_L}(X_p)$ we have that
+\begin{align}
+  ??
+\end{align}
+Therefore, under assumption A) we have that ...
+and therefore $\widehat{A_L}(X_p) = A(X_p) + o_P(1)$ uniformly in $p$.
+A consequence of equation (\ref{eq:countingConv}) is that $\widehat{N}_L(P; d) = N(P;d) + o_P(1)$ uniformly in $d$ and therefore by the continuous mapping theorem and assumption B) the first factor in equation (\ref{eq:secondTerm}) is of order $o_P(1)$ uniformly in $p$ and $d$.
+The result then follows by noting that the second factor is bounded by $P$ uniformly in $d$.
+
+Continuing with $U_l(d)$ we have that
+\begin{align}
+	...
+\end{align}
+using integration by parts.
+
+\begin{align}
+S_d &= \left\{X_p; \left(\sum_{r=1}^{P!} I\{r(X_p)\le d\} Q(r)\right) >0\right\}\\
+    &= \left\{X_p; \E[I\{R(X_p)\le d\}] >0\right\}\\
+    &= \left\{X_p; P(R(X_p)\le d) > 0\right\}\\
+\widehat{S}_{d,L} &= \left\{X_p; \left(\sum_{l=1}^L I\{R_l(X_p)\le d\}\right)>0\right\}
+\end{align}
+This completes the proof.
+
+\section{Proof of corollary \ref{cor:consistency}}
+
 \end{document}

### 2016-05-20 latex/superranker-manuscript.tex
proof of thm. 1

@@
 with possible extensions. The approaches presented in this manuscript
-are available in the R package \texttt{SuperRanker}. [XXX check dette
-passer, submit pakke]
+are available in the R package \texttt{SuperRanker}. \mcomment{XXX check dette
+passer, submit pakke}
 
@@
 can be measured by the variance across the lists
-\begin{equation}
-  {A}(X_p) = \sum_{r=1}^{P!} \{r(X_P)-\E_Q R(X_P)\}^2 Q(r)
-\end{equation}
+\begin{align}
+  A(X_p) &= \E_Q\left[\left(R(X_P)-\E_Q R(X_P)\right)^2\right]\\
+   &= \sum_{r=1}^{P!} \left(r(X_P)-\E_Q R(X_P)\right)^2 Q(r)\nn
+\end{align}
 where $\E_Q R(X_p)=\sum_{r=1}^{P!} r(X_p)Q(r)$ and $Q$ is a probability mass function
@@
 \begin{align}
-\widehat{A}_L(X_p) = \frac{1}{L-1}\sum_{i=1}^L (R_i(X_p) - \overline{R}(X_p))^2, \quad \overline{R}_L(X_p) = \frac{1}{L}\sum_{i=1}^L R_i(X_p).
+\widehat{A}_L(X_p) = \frac{1}{L-1}\sum_{i=1}^L (R_i(X_p) - \overline{R}_L(X_p))^2, \quad \overline{R}_L(X_p) = \frac{1}{L}\sum_{i=1}^L R_i(X_p)\label{eq:empDistance}
 \end{align}
@@
 \begin{equation}
-  \widehat{\textrm{sra}}_L(d) = \frac{\sum_{\{p \in \widehat{S}_{L}(d)\}}(L-1)\widehat{A_L}(X_p)}{(L-1)|\widehat {S}_{L}(d)|}.\label{def:sraEst}
+  \widehat{\textrm{sra}}_L(d) = \frac{\sum_{\{p \in \widehat{S}_{L}(d)\}}(L-1)\widehat{A}_L(X_p)}{(L-1)|\widehat {S}_{L}(d)|}\label{def:sraEst}
 \end{equation}
@@
 
-First, we define the parameter of interest. For any given threshold value $q$, 
+We start by showing that $\widehat{\textrm{sra}}_L$ is a consistent estimator of $\textrm{sra}$ for $L \rightarrow \infty$.
+The proof is given in Theorem \ref{thm:consistency} below and requires the following assumptions.
+\begin{assumption}
+We assume that:
+\begin{enumerate}[label=\Alph*)]
+\item{The $L$ lists are independent and identically distributed samples from $Q$.}
+\item{$\E_Q[R(X_p)^2] < \infty$ uniformly in $p$.}
+\item{$|S(d)| > 0$ uniformly in $d$}.
+\end{enumerate}
+\end{assumption}
+We note that assumption $C$ implies that for every list depth $d$ there must be a non-zero probability of ... under the probability measure $Q$.
+
+\begin{theorem}
+Under the assumptions it follows that $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty = o_P(1)$.
+\label{thm:consistency}
+\end{theorem}
+
+We now define the parameter of interest. For any given threshold value $q$, 
 a change-point $d^\ast(q)$ in the list agreement is defined by
@@
 \end{equation}
-
-We start by showing that $\widehat{\textrm{sra}}_L$ is a consistent estimator of $\textrm{sra}$ for $L \rightarrow \infty$.
-The consistency of $\widehat{d^\ast_L}(q)$ then follows by a simple corollary. The proof requires the following assumptions.
-
-\begin{assumption}
-We assume that:
-\begin{enumerate}[label=\Alph*)]
-\item{$\E_Q[R(X_p)^4]$ is finite for all $p \in 1,\ldots,P$}.
-\item{$|S(d)| > 0$ uniformly in $d$}.
-\end{enumerate}
-\end{assumption}
-We note that assumption $B$ implies that for every list depth $d$ there must be a non-zero probability of ... under the probability measure $Q$.
-
-\begin{theorem}
-Under the assumptions it follows that $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty = o_P(1)$.
-\label{thm:consistency}
-\end{theorem}
+The consistency of $\widehat{d^\ast_L}(q)$ then follows by a simple corollary.
 
@@
 \end{corollary}
+
+\mcomment{Comment on how to apply this in practice (finite sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style.}
 
@@
 \end{align}
-which jumps at a time $p$ if the probability that the rank of $X_p$ is less than or equal to $d$ is bounded away from zero under the probability measure $Q$.
-Similarly we have the empirical counterpart
+which jumps with an increment of one at time $p$ if the probability that the rank of $X_p$ is less than or equal to $d$ is bounded away from zero under the probability measure $Q$.
+Similarly, we have the empirical counterpart
 \begin{align}
@@
 \end{align}
-Since the joint distribution of $\left\{\widehat{N}_L(p;d) : p=1,\ldots,P\right\}$ is completely determined by its finite set of jump times and $\widehat{Q}_L(R(X_p) \leq d) \overset{P}{\longrightarrow} Q(R(X_p) \leq d)$ for every $p$ and $d$ for $L \rightarrow \infty$ by the law of large numbers, it follows that
+The joint law of $\{\widehat{N}_L(p;d)\}_{1\leq p \leq P}$ is completely determined by its finite set of jump times, and because
+$1(R_l(X_s) \leq d)$ are independent and identically Bernoulli distributed over $l$ with expectation $Q(R(X_s) \leq d)$ 
+it follows from the law of large numbers that $\widehat{Q}_L(R(X_p) \leq d) \overset{P}{\longrightarrow} Q(R(X_p) \leq d)$ for every $p$ and $d$ 
+for $L \rightarrow \infty$, and therefore
 \begin{align}
@@
 
-Recalling the definition of the sequential rank agreement in equation (\ref{def:sra}), it may be rewritten as the following integral with respect to the counting measure induced by $N$
+The sequential rank agreement given in equation (\ref{def:sra}) may therefore be rewritten as the following integral with respect to the counting measure induced by $N$
 \begin{align}
@@
 \end{align}
-where we note that the cardinality of the set $S(d)$ in equation (\ref{eq:sumset}) is equal to the counting process evaluated at $P$.
+where we note that the cardinality of the set $S(d)$ in equation (\ref{eq:sumset}) is equal to the counting process evaluated at time $P$.
 The empirical sequential rank agreement is similarly given by
 \begin{align}
-  \widehat{\textrm{sra}}_L(d) &= \int_1^P \frac{\widehat{A_L}(X_p)}{\widehat{N}_L(P; d)}d\widehat{N}_L(p; d) 
+  \widehat{\textrm{sra}}_L(d) &= \int_1^P \frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)}d\widehat{N}_L(p; d) 
 \end{align}
@@
   U_L(d) &= \int_1^P\frac{A(X_p)}{N(P;d)} d\left(\widehat{N}_L(p; d) - N(p; d)\right)\\
-  V_L(d) &=  \int_1^P \left(\frac{\widehat{A_L}(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right)d\widehat{N}_L(p; d)
+  V_L(d) &=  \int_1^P \left(\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right)d\widehat{N}_L(p; d)
 \end{align}
@@
 
-Starting with $V_L(d)$, since $\widehat{N}_L$ is of bounded variation it follows that
-\begin{align}
- V_L(d) &\leq \sup_{p \in 1,\ldots,P}\left|\frac{\widehat{A_L}(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right| \int_1^P \left|d\widehat{N}_L(p; d)\right|\label{eq:secondTerm}
-\end{align}
-Looking at $\widehat{A_L}(X_p)$ we have that
-\begin{align}
-  ??
-\end{align}
-Therefore, under assumption A) we have that ...
-and therefore $\widehat{A_L}(X_p) = A(X_p) + o_P(1)$ uniformly in $p$.
-A consequence of equation (\ref{eq:countingConv}) is that $\widehat{N}_L(P; d) = N(P;d) + o_P(1)$ uniformly in $d$ and therefore by the continuous mapping theorem and assumption B) the first factor in equation (\ref{eq:secondTerm}) is of order $o_P(1)$ uniformly in $p$ and $d$.
+Starting with the first term, we have that
+\begin{align}
+  \left|U_L(d)\right| \leq \left(\sup_{p \in 1,\ldots,P}\left|\frac{A(X_p)}{N(P;d)}\right|\right)\left(\widehat{N}_L(p; d) - N(p; d)\right)\biggr\rvert_1^P
+\end{align}
+where the second factor is $o_P(1)$ uniformly in $d$ by equation (\ref{eq:countingConv}).
+The first factor is $O(1)$ since $N(P;d)$ uniformly bounded away from zero in d by assumption C and $A(X_p)$ is uniformly bounded in $p$ by assumption B.
+It thus follows that $\left|U_L(d)\right| = o_P(1)$ uniformly in $d$.
+
+Continuing with $V_L(d)$ we may similarly bound as
+\begin{align}
+ V_L(d) &\leq \sup_{p \in 1,\ldots,P}\left|\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right| \int_1^P \left|d\widehat{N}_L(p; d)\right|\label{eq:secondTerm}
+\end{align}
+It follows directly from the law of large numbers that $\widehat{A_L}(X_p) = A(X_p) + o_P(1)$ uniformly in $p$ since $R_i(X_p)$ in equation (\ref{eq:empDistance}) for $i=1,\ldots,L$ are independent and identically distributed random variables from $Q$ with finite second moment by assumption B).
+Again by equation (\ref{eq:countingConv}) it follows that $\widehat{N}_L(P; d) = N(P;d) + o_P(1)$ uniformly in $d$ and therefore by the continuous mapping theorem and assumption C) the first factor in equation (\ref{eq:secondTerm}) is of order $o_P(1)$ uniformly in $p$ and $d$.
 The result then follows by noting that the second factor is bounded by $P$ uniformly in $d$.
 
-Continuing with $U_l(d)$ we have that
-\begin{align}
-	...
-\end{align}
-using integration by parts.
-
-\begin{align}
-S_d &= \left\{X_p; \left(\sum_{r=1}^{P!} I\{r(X_p)\le d\} Q(r)\right) >0\right\}\\
-    &= \left\{X_p; \E[I\{R(X_p)\le d\}] >0\right\}\\
-    &= \left\{X_p; P(R(X_p)\le d) > 0\right\}\\
-\widehat{S}_{d,L} &= \left\{X_p; \left(\sum_{l=1}^L I\{R_l(X_p)\le d\}\right)>0\right\}
-\end{align}
 This completes the proof.

### 2016-05-23 latex/superranker-manuscript.tex
proof read proof

@@
 
+% latex table generated in R 3.2.3 by xtable 1.8-2 package
+% Fri Apr  8 10:22:58 2016
+\begin{table}[tb]
+\centering
+\caption{Top 10 list of ranked results from the Golub data. Numbers indicate the predictor/gene for the given ranking and method} 
+\label{tab1}
+\begin{tabular}{rrrrr}
+  \hline
+Ranking & T & LogReg & ElasticNet & MIC \\ 
+  \hline
+1 & 2124 & 2124 & 829 & 378 \\ 
+  2 & 896 & 896 & 2124 & 829 \\ 
+  3 & 2600 & 829 & 2198 & 896 \\ 
+  4 & 766 & 394 & 1665 & 1037 \\ 
+  5 & 829 & 766 & 1920 & 2124 \\ 
+  6 & 2851 & 2670 & 1042 & 808 \\ 
+  7 & 703 & 2939 & 808 & 108 \\ 
+  8 & 2386 & 2386 & 849 & 515 \\ 
+  9 & 2645 & 1834 & 937 & 2670 \\ 
+  10 & 2002 & 378 & 1995 & 2600 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
 % chunks 2,3,4,5,6,7
@@
 theoretical properties of our proposed method for this specific
-task. As Hall and Schimek \mcomment{REF}, we work with an infinite set of lists
-and study the asymptotic behaviour for $L\to\infty$.
-However, we do not allow the length of the lists to
-change with $L$ as in most applications the length of the lists is fixed.
-
-We start by showing that $\widehat{\textrm{sra}}_L$ is a consistent estimator of $\textrm{sra}$ for $L \rightarrow \infty$.
-The proof is given in Theorem \ref{thm:consistency} below and requires the following assumptions.
-\begin{assumption}
-We assume that:
-\begin{enumerate}[label=\Alph*)]
-\item{The $L$ lists are independent and identically distributed samples from $Q$.}
-\item{$\E_Q[R(X_p)^2] < \infty$ uniformly in $p$.}
-\item{$|S(d)| > 0$ uniformly in $d$}.
-\end{enumerate}
-\end{assumption}
-We note that assumption $C$ implies that for every list depth $d$ there must be a non-zero probability of ... under the probability measure $Q$.
+task. As Hall and Schimek \mcomment{REF}, we work with an infinite set
+of lists and study the asymptotic behaviour for $L\to\infty$.
+However, we do not allow the length of the lists to change with $L$
+because in most applications the lengths of the lists are fixed.
+
+We start by showing that $\widehat{\textrm{sra}}_L$ is a consistent
+estimator of $\textrm{sra}$ for $L \rightarrow \infty$. The result is
+stated in Theorem \ref{thm:consistency}.
+
+% below and requires the
+% following assumptions.
+% \begin{assumption}
+% We assume that:
+% \begin{enumerate}[label=\Alph*)]
+% \item{The $L$ lists are independent and identically distributed
+    % samples from a probability distribution $Q$ on the set of lists.}
+% \item{$\E_Q[R(X_p)^2] < \infty$ uniformly in $p$.}
+% \item{$|S(d)| > 0$ uniformly in $d$}.
+% \end{enumerate}
+% \end{assumption}
+% We note that assumption $C$ implies that for every list depth $d$
+% there must be a non-zero probability of ... under the probability
+% measure $Q$.
 
 \begin{theorem}
-Under the assumptions it follows that $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty = o_P(1)$.
+  Assume that $\{R_l(X)\}_{l=1}^L$ are independent draws from the a
+  probability distribution $Q$ on the set of lists.  Then,
+  $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty =
+  o_P(1)$.
 \label{thm:consistency}
@@
 \section{Proof of Theorem \ref{thm:consistency}}
-We start by defining the following cadlag function
+
+We start by defining the following counting process
 \begin{align}
@@
 \end{align}
-which jumps with an increment of one at time $p$ if the probability that the rank of $X_p$ is less than or equal to $d$ is bounded away from zero under the probability measure $Q$.
-Similarly, we have the empirical counterpart
+which runs through the list elements in an arbitrary order and counts
+how many list elements have a strictly positive probability to have
+rank less than or equal to $d$ under the probability measure $Q$.
+Note that the cardinality of the set $S(d)$ in equation
+(\ref{eq:sumset}) is equal to the counting process evaluated at the last element:
+\begin{equation*}
+  |S(d)| = N(P;d).
+\end{equation*}
+
+The empirical counterpart of the counting process is given by
 \begin{align}
   \widehat{N}_L(p; d) &= \sum_{s=1}^p 1\left( \frac{1}{L} \sum_{l=1}^L 1(R_l(X_s) \leq d) > 0\right)\\
-        &= \sum_{s=1}^p 1\left(\widehat{Q}_L(R(X_s) \leq d) > 0\right)\nn
-\end{align}
-The joint law of $\{\widehat{N}_L(p;d)\}_{1\leq p \leq P}$ is completely determined by its finite set of jump times, and because
-$1(R_l(X_s) \leq d)$ are independent and identically Bernoulli distributed over $l$ with expectation $Q(R(X_s) \leq d)$ 
-it follows from the law of large numbers that $\widehat{Q}_L(R(X_p) \leq d) \overset{P}{\longrightarrow} Q(R(X_p) \leq d)$ for every $p$ and $d$ 
-for $L \rightarrow \infty$, and therefore
-\begin{align}
-  \sup_{d \in 1,\ldots,P}\sup_{p \in 1,\ldots, P} \left|\widehat{N}_L(p; d) - N(p; d)\right| = o_P(1)\label{eq:countingConv}
-\end{align}
-
-The sequential rank agreement given in equation (\ref{def:sra}) may therefore be rewritten as the following integral with respect to the counting measure induced by $N$
-\begin{align}
-  \textrm{sra}(d) = \int_{1}^P \frac{A(X_p)}{N(P; d)}dN(p; d)
-\end{align}
-where we note that the cardinality of the set $S(d)$ in equation (\ref{eq:sumset}) is equal to the counting process evaluated at time $P$.
+        &= \sum_{s=1}^p 1\left(\widehat{Q}_L(R(X_s) \leq d) > 0\right)\nn.
+\end{align}
+The joint law of the sequence
+$\{\widehat{N}_L(p;d)\}_{1\leq p \leq P}$ is completely determined by
+the finite set of the jump times of $N(p;d)$. Since
+$\{1(R_l(X_s) \leq d); l=1,\dots, L\}$ consists of independent and
+identically Bernoulli distributed variables with expectation
+$Q(R(X_s) \leq d)$ it follows from the law of large numbers that
+$\widehat{Q}_L(R(X_p) \leq d) \overset{P}{\longrightarrow} Q(R(X_p)
+\leq d)$ for every $p$ and $d$ as $L \rightarrow \infty$. 
+Therefore, we have
+\begin{align}
+  \sup_{d \in 1,\ldots,P}\sup_{p \in 1,\ldots, P} \left|\widehat{N}_L(p; d) - N(p; d)\right| = o_P(1).\label{eq:countingConv}
+\end{align}
+
+The sequential rank agreement given in equation
+(\ref{def:sra}) may therefore be rewritten as the following integral
+with respect to $N$
+\begin{align}
+  \textrm{sra}(d) = \int_{1}^P \frac{A(X_p)}{N(P; d)}\mathrm dN(p; d).
+\end{align}
 The empirical sequential rank agreement is similarly given by
 \begin{align}
-  \widehat{\textrm{sra}}_L(d) &= \int_1^P \frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)}d\widehat{N}_L(p; d) 
-\end{align}
-and by elementary algebra it follows that
+  \widehat{\textrm{sra}}_L(d) &= \int_1^P \frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)}\mathrm d \widehat{N}_L(p; d) 
+\end{align}
+and it follows that
 \begin{align}
@@
 \begin{align}
-  U_L(d) &= \int_1^P\frac{A(X_p)}{N(P;d)} d\left(\widehat{N}_L(p; d) - N(p; d)\right)\\
-  V_L(d) &=  \int_1^P \left(\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right)d\widehat{N}_L(p; d)
-\end{align}
-The conclusion of the proof therefore follows if each of these two terms are uniformly of order $o_P(1)$ in $d$.
-
-Starting with the first term, we have that
+  U_L(d) &= \int_1^P\frac{A(X_p)}{N(P;d)} \mathrm d\left(\widehat{N}_L(p; d) - N(p; d)\right)\\
+  V_L(d) &=  \int_1^P \left(\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right)\mathrm d\widehat{N}_L(p; d)
+\end{align}
+The conclusion of the proof therefore follows if each of these two
+terms are uniformly of order $o_P(1)$ in $d$.  For the first term we
+have that
 \begin{align}
@@
 \end{align}
-where the second factor is $o_P(1)$ uniformly in $d$ by equation (\ref{eq:countingConv}).
-The first factor is $O(1)$ since $N(P;d)$ uniformly bounded away from zero in d by assumption C and $A(X_p)$ is uniformly bounded in $p$ by assumption B.
-It thus follows that $\left|U_L(d)\right| = o_P(1)$ uniformly in $d$.
-
-Continuing with $V_L(d)$ we may similarly bound as
-\begin{align}
- V_L(d) &\leq \sup_{p \in 1,\ldots,P}\left|\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right| \int_1^P \left|d\widehat{N}_L(p; d)\right|\label{eq:secondTerm}
-\end{align}
-It follows directly from the law of large numbers that $\widehat{A_L}(X_p) = A(X_p) + o_P(1)$ uniformly in $p$ since $R_i(X_p)$ in equation (\ref{eq:empDistance}) for $i=1,\ldots,L$ are independent and identically distributed random variables from $Q$ with finite second moment by assumption B).
-Again by equation (\ref{eq:countingConv}) it follows that $\widehat{N}_L(P; d) = N(P;d) + o_P(1)$ uniformly in $d$ and therefore by the continuous mapping theorem and assumption C) the first factor in equation (\ref{eq:secondTerm}) is of order $o_P(1)$ uniformly in $p$ and $d$.
-The result then follows by noting that the second factor is bounded by $P$ uniformly in $d$.
-
-This completes the proof.
+where the second factor is $o_P(1)$ uniformly in $d$ by equation
+(\ref{eq:countingConv}). The first factor is $O(1)$ since $N(P;d)>0$
+uniformly in $d$ and because $A(X_p)\le 2L^2$ u uniformly in $p$. . It
+thus follows that $\left|U_L(d)\right| = o_P(1)$ uniformly in $d$.
+
+Similarly we derive an upper bound for $V_L(d)$ by
+\begin{align}
+ V_L(d) &\leq \sup_{p \in 1,\ldots,P}\left|\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right| \int_1^P \left|\mathrm d\widehat{N}_L(p; d)\right|\label{eq:secondTerm}
+\end{align}
+It follows directly from the law of large numbers that
+$\widehat{A_L}(X_p) = A(X_p) + o_P(1)$ uniformly in $p$ since
+$R_i(X_p)$ in equation (\ref{eq:empDistance}) for $i=1,\ldots,L$ are
+independent and $Q$-identically distributed random variables with
+finite second moment since $A(X_p)\le 2L^2$). Again by equation
+(\ref{eq:countingConv}) it follows that
+$\widehat{N}_L(P; d) = N(P;d) + o_P(1)$ uniformly in $d$ and therefore
+by the continuous mapping theorem and since $N(P;d)>0$) the first factor
+in equation (\ref{eq:secondTerm}) is of order $o_P(1)$ uniformly in
+$p$ and $d$. The result then follows by noting that the second factor
+is bounded by $P$ uniformly in $d$.
+% This completes the proof.
 

### 2016-05-25 latex/superranker-manuscript.tex
minor adjustments

@@
 \begin{equation}
-  d^\ast(q) = \argmin_{d\in\{1,\dots,P\}}\textrm{sra}(d)>q.
+  d^\ast(q) = \min\left\{\argmin_{d\in\{1,\dots,P\}}\textrm{sra}(d)>q, P\right\}.
 \end{equation}
@@
 \begin{equation}
-\widehat{d^\ast_L}(q) = \argmin_{d\in\{1,\dots,P\}}\widehat{\textrm{sra}}_L(d)>q
+\widehat{d^\ast_L}(q) = \min\left\{\argmin_{d\in\{1,\dots,P\}}\widehat{\textrm{sra}}_L(d)>q, P \right\}
 \end{equation}
@@
 
-We start by defining the following counting process
+We start by defining the following cadlag function
 \begin{align}
@@
 (\ref{eq:countingConv}). The first factor is $O(1)$ since $N(P;d)>0$
-uniformly in $d$ and because $A(X_p)\le 2L^2$ u uniformly in $p$. . It
+uniformly in $d$ and because $A(X_p)\le 2P^2$ u uniformly in $p$ \mcomment{double check bound}. It
 thus follows that $\left|U_L(d)\right| = o_P(1)$ uniformly in $d$.

### 2016-05-25 latex/superranker-manuscript.tex
minor adjustments

@@
 \begin{align}
-S(d) &= \left\{X_p; \left(\sum_{r=1}^{P!} 1\left(r(X_p)\le d\right) Q(r)\right) >0\right\}\label{eq:sumset}\\
-     &= \left\{X_p; Q(R(X_p) \leq d) > 0 \right\}\nn
+S(d) &= \left\{X_p; \left(\sum_{r=1}^{P!} 1\left(r(X_p)\le d\right) Q(r)\right) >0\right\}\label{eq:sumset}
 \end{align}
 where $1(\cdot)$ denotes the indicator function.
-The empirical counterpart the set of unique items ranked less than or
+The empirical counterpart is the set of unique items ranked less than or
 equal to $d$ in any of the $L$ lists:
@@
 (\ref{eq:countingConv}). The first factor is $O(1)$ since $N(P;d)>0$
-uniformly in $d$ and because $A(X_p)\le 2P^2$ u uniformly in $p$ \mcomment{double check bound}. It
+uniformly in $d$ and because $A(X_p)\le 2P^2$ uniformly in $p$ \mcomment{double check bound}. It
 thus follows that $\left|U_L(d)\right| = o_P(1)$ uniformly in $d$.

### 2016-05-26 latex/superranker-manuscript.tex
minor updates

@@
 \begin{equation}
-  \widetilde R_l(X_p) = \left\{\begin{array}{cl} \{R_l(X_p)\} & \text{ for } X_p\in \Lambda_l \\
-      \{d_l+1,\dots,P\} & \text{ for } X_p \not\in \Lambda_l\end{array}\right.
+	\widetilde R_l(X_p) = \begin{cases}
+	     \{R_l(X_p)\} & \text{for } X_p\in \Lambda_l\\
+	     \{d_l+1,\dots,P\} & \text{for} X_p \not\in \Lambda_l
+        \end{cases}
 \end{equation}
@@
 The first reference hypothesis is
-\begin{eqnarray*}
-H_0  & : &  \text{The list rankings correspond to complete randomly}\\
-       &  & \text{permuted lists},
-\end{eqnarray*}
+\begin{eqnarray}
+H_0  & : &  \text{The list rankings correspond to complete randomly}\label{eq:permutationHypothesis}\\
+       &  & \text{permuted lists}\nn
+\end{eqnarray}
 which not only assumes that there is no information in the data on
@@
 theoretical properties of our proposed method for this specific
-task. As Hall and Schimek \mcomment{REF}, we work with an infinite set
+task. As in \cite{hall:schi:2012} we work with an infinite set
 of lists and study the asymptotic behaviour for $L\to\infty$.
@@
 
-We now define the parameter of interest. For any given threshold value $q$, 
-a change-point $d^\ast(q)$ in the list agreement is defined by
-\begin{equation}
-  d^\ast(q) = \min\left\{\argmin_{d\in\{1,\dots,P\}}\textrm{sra}(d)>q, P\right\}.
-\end{equation}
-The value of $q$ could for example be the agreement of randomly
-permuted lists. The corresponding estimated change-point is similarly defined by
-\begin{equation}
-\widehat{d^\ast_L}(q) = \min\left\{\argmin_{d\in\{1,\dots,P\}}\widehat{\textrm{sra}}_L(d)>q, P \right\}
-\end{equation}
-The consistency of $\widehat{d^\ast_L}(q)$ then follows by a simple corollary.
-
+We now define the parameter of interest. For any given deterministic threshold function $q\colon\, 1,\ldots,P \mapsto \mathbb{R}_{\geq 0}$
+we define the population superlevel set of the sequential rank agreement as
+\begin{align}
+  \mathcal{L}(q) = \left\{d : \textrm{sra}(d) \geq q(d)\right\}
+\end{align}
+The values of $q$ could for example be the limits-of-agreement for randomly
+permuted lists as in the null-hypothesis in equation (\ref{eq:permutationHypothesis}).
+A change-point $d^\ast(q)$ in the list agreement is then defined by the position
+\begin{align}
+  d^\ast(q) &= 1\left(|\mathcal{L}(q)| > 0\right)\inf \mathcal{L}(q) + 1\left(|\mathcal{L}(q)| = 0\right)P
+\end{align}
+corresponding to the first list depth where the sequential rank agreement exceeds the threshold 
+if such a position exists. Otherwise, the full list is in agreement according to $q$ and the change-point
+is defined as the length of the lists. The empirical superlevel set is similarly defined as
+\begin{align}
+ \widehat{\mathcal{L}}_L(\widehat{q}_L) = \left\{d : \widehat{\textrm{sra}}_L(d) \geq \widehat{q}_L(d)\right\}
+\end{align}
+where we allow that the threshold function can depend on the sample size as well. The estimated change-point is therefore
+\begin{align}
+  \widehat{d^\ast_L}(\widehat{q}_L) &= 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| > 0)\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) + 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0)P
+\end{align}
+
+The consistency of the estimated change-point, $\widehat{d^\ast_L}(\widehat{q}_L)$, follows from Theorem \ref{thm:consistency} by the following corollary. 
 \begin{corollary}
-For fixed $q$ it follows from Theorem \ref{thm:consistency} that
-$\widehat{d^\ast_L}(q) \overset{P}{\longrightarrow} d^\ast(q)$
-for $L \rightarrow \infty$.
+Let $\widehat{q}_L$ be a threshold function such that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ for some limiting
+function $q$.  Then, $\widehat{d^\ast_L}(\widehat{q}_L) \overset{P}{\longrightarrow} d^\ast(q)$ for $L \rightarrow \infty$.
 \label{cor:consistency}
@@
 
-\mcomment{Comment on how to apply this in practice (finite sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style.}
+\mcomment{Elaborate more on how to apply this in practice (finite sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style?}
 
@@
 Note that the cardinality of the set $S(d)$ in equation
-(\ref{eq:sumset}) is equal to the counting process evaluated at the last element:
-\begin{equation*}
-  |S(d)| = N(P;d).
-\end{equation*}
-
+(\ref{eq:sumset}) is equal to the counting process evaluated at the last element: $|S(d)| = N(P;d)$.
 The empirical counterpart of the counting process is given by
@@
 (\ref{eq:countingConv}). The first factor is $O(1)$ since $N(P;d)>0$
-uniformly in $d$ and because $A(X_p)\le 2P^2$ uniformly in $p$ \mcomment{double check bound}. It
+uniformly in $d$ and because $A(X_p) = O(P^2)$ uniformly in $p$ by the Cauchy-Ekstr{\o}m inequality. It
 thus follows that $\left|U_L(d)\right| = o_P(1)$ uniformly in $d$.
@@
 \end{align}
-It follows directly from the law of large numbers that
+and it follows directly from the law of large numbers that
 $\widehat{A_L}(X_p) = A(X_p) + o_P(1)$ uniformly in $p$ since
@@
 independent and $Q$-identically distributed random variables with
-finite second moment since $A(X_p)\le 2L^2$). Again by equation
-(\ref{eq:countingConv}) it follows that
-$\widehat{N}_L(P; d) = N(P;d) + o_P(1)$ uniformly in $d$ and therefore
-by the continuous mapping theorem and since $N(P;d)>0$) the first factor
+finite second moment. Again, $\widehat{N}_L(P; d) = N(P;d) + o_P(1)$ uniformly in $d$ by equation
+(\ref{eq:countingConv}) and $N(P;d)>0$ so by the continuous mapping theorem the first factor
 in equation (\ref{eq:secondTerm}) is of order $o_P(1)$ uniformly in

### 2016-05-26 latex/superranker-manuscript.tex
minor additions

@@
 
-\section{Proof of corollary \ref{cor:consistency}}
+\section{Proof of Corollary \ref{cor:consistency}}
+We have that 	
+\begin{align}
+  |\widehat{d^\ast_L}(\widehat{q}_L) - d^\ast(q)| &\leq  |\widehat{d^\ast_L}(q)  - d^\ast(q)| + |\widehat{d^\ast_L}(\widehat{q}_L)  - \widehat{d^\ast_L}(q_L)|
+\end{align}
+The first term is of order $o_P(1)$ by the continuous mapping theorem if a) $|\inf \widehat{\mathcal{L}}_L(q) - \inf  \mathcal{L}(q)| = o_P(1)$ and b) $|1(|\widehat{\mathcal{L}}_L(q)| = 0) - 1(|\mathcal{L}(q)| = 0)| = o_P(1)$ since $1(|\mathcal{L}(q)| > 0) = 1 - 1(|\mathcal{L}(q)| = 0)$.
+Similarly, the second term is of order $o_P(1)$ by the same reasoning if c) $|\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) - \inf \widehat{\mathcal{L}}_L(q)| = o_P(1)$ and d) $|1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0) - 1(|\widehat{\mathcal{L}}_L(q)| = 0)| = o_P(1)$.
+
+\begin{align}
+\inf \left\{d : \widehat{\textrm{sra}}_L(d) \geq \widehat{q}_L(d)\right\} = \inf \left\{d : \widehat{\textrm{sra}}_L(d) \geq q(d) + o_P(1)\right\}\\
+\inf \left\{d : \widehat{\textrm{sra}}_L(d) \geq q(d)\right\}
+\end{align}
+This completes the proof.
 

### 2016-05-28 latex/superranker-manuscript.tex
fiddling with the corollary

--- 2016-05-26 latex/superranker-manuscript.tex
@@
 is bounded by $P$ uniformly in $d$.
-% This completes the proof.
-
-\section{Proof of corollary \ref{cor:consistency}}
+
+\section{Proof of Corollary \ref{cor:consistency}}
+Let $A \triangle B = (A \setminus B) \cup (B \setminus A)$ is the symmetric difference between sets $A$ and $B$. It is then sufficient to show that
+$|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q)|$ converges to zero in probability as $L \rightarrow \infty$. By the triangle inequality
+we have that
+\begin{align}
+ \left|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q)\right| &= \left|\left(\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \widehat{\mathcal{L}}_L(q)\right) \cup \left(\widehat{\mathcal{L}}_L(q) \triangle \mathcal{L}(q)\right)\right|\label{eq:corSetDiff}\\
+   &\leq \left|\widehat{\mathcal{L}}_L(q) \triangle \mathcal{L}(q)\right| + \left|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \widehat{\mathcal{L}}_L(q)\right|\nn
+\end{align}
+Looking at the first term we have
+\begin{align}
+  \left|\widehat{\mathcal{L}}_L(q) \triangle \mathcal{L}(q)\right| &= \sum_{d=1}^P \left|1\left(\widehat{\textrm{sra}}_L(d) \geq q(d)\right)  - 1\left(\textrm{sra}(d)  \geq q(d)\right)\right|\\
+   &=
+\end{align}
+Similarly we have
+\begin{align}
+\left|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \widehat{\mathcal{L}}_L(q)\right| &= \sum_{d=1}^P \left|1\left(\widehat{\textrm{sra}}_L(d) \geq \widehat{q}_L(d)\right) - 1\left(\widehat{\textrm{sra}}_L(d) \geq q(d)\right)\right|\\
+   &=
+\end{align}
+
+%We have that 	
+%\begin{align}
+%  |\widehat{d^\ast_L}(\widehat{q}_L) - d^\ast(q)| &\leq  |\widehat{d^\ast_L}(q)  - d^\ast(q)| + |\widehat{d^\ast_L}(\widehat{q}_L)  - \widehat{d^\ast_L}(q_L)|
+%\end{align}
+%The first term is of order $o_P(1)$ by the continuous mapping theorem if a) $|\inf \widehat{\mathcal{L}}_L(q) - \inf  \mathcal{L}(q)| = o_P(1)$ and b) $|1(|\widehat{\mathcal{L}}_L(q)| = 0) - 1(|\mathcal{L}(q)| = 0)| = o_P(1)$ since $1(|\mathcal{L}(q)| > 0) = 1 - 1(|\mathcal{L}(q)| = 0)$.
+Similarly, the second term is of order $o_P(1)$ by the same reasoning if c) $|\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) - \inf \widehat{\mathcal{L}}_L(q)| = o_P(1)$ and d) $|1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0) - 1(|\widehat{\mathcal{L}}_L(q)| = 0)| = o_P(1)$.
+
+This completes the proof.
 

### 2016-05-29 R/plot.sra.R
cleaned for first CRAN submission. now runs through all devel checks

@@ -6,7 +6,8 @@
 ##' @param ylim y-axis limits
 ##' @param xlab x-axis lab
 ##' @param ylab y-axis lab
-##' @param ... Processed by \code{\link{prodlim::SmartControl}}.
+##' @param add Logical. If \code{TRUE} add graph to existing plot.
+##' @param ... Processed by function \code{prodlim::SmartControl}.
 ##' @return Graph
 ##' @examples
 ##' R1=c(1,2,3,4,5,7,6,8,9,10,11,12,13)

### 2016-05-29 R/sra.R
cleaned for first CRAN submission. now runs through all devel checks

@@ -1,14 +1,19 @@
 #' Compute the sequential rank agreement
 #'
 #' @param object Either matrix where each column is a ranked list of
-#' items or a list of ranked lists of items. Elements are integers
-#' between 1 and the length of the lists. The lists should have the
-#' same length but censoring can be used by setting the list to zero
-#' from a point onwards. See details for more information.
-#' @param na.strings A vector of strings/values that represent missing values in addition to NA. Defaults to NULL which means only NA are censored values.
+#'     items or a list of ranked lists of items. Elements are integers
+#'     between 1 and the length of the lists. The lists should have
+#'     the same length but censoring can be used by setting the list
+#'     to zero from a point onwards. See details for more information.
 #' @param B An integer giving the number of randomization to sample
-#' over in the case of censored observations
-#' @param type The type of measure to use. Either sd (standard deviation - the default) or mad (median absolute deviance)
+#'     over in the case of censored observations
+#' @param na.strings A vector of strings/values that represent missing
+#'     values in addition to NA. Defaults to NULL which means only NA
+#'     are censored values.
+#' @param nitems Claus: please describe this argument
+#' @param type The type of measure to use. Either sd (standard
+#'     deviation - the default) or mad (median absolute deviance)
+#' @param ... Arguments passed to methods.
 #' @return A vector of the sequential rank agreement
 ##' @examples
 ##'
@@ -42,19 +47,19 @@
 #'
 #' @rdname sra
 #' @export
-sra <- function(object, B=1, type=c("sd", "mad"), ...) {
+sra <- function(object,B,na.strings,nitems,type,...) {
   UseMethod("sra")
 }
 
 #' @rdname sra
 #' @export
-sra.default <- function(object, B=1, type=c("sd", "mad")) {
+sra.default <- function(object,B,na.strings,nitems,type,...) {
     stop("Input must be either a matrix, a data.frame or a list.")
 }
 
 #' @rdname sra
 #' @export
-sra.matrix <- function(object, B=1, na.strings=NULL, nitems=nrow(object), type=c("sd", "mad")) {
+sra.matrix <- function(object, B=1, na.strings=NULL, nitems=nrow(object), type=c("sd", "mad"),...) {
     if (!is.matrix(object))
         stop("Input object must be a matrix")
 
[…]

### 2016-05-30 R/SuperRanker.R
Updated version

@@ -1,14 +1,0 @@
-#' Functions related to comparison of ranked lists
-#'
-#' SuperRanker allows you tRoxygen2 allows you to write documentation in comment blocks co-located
-#' with code.
-#'
-#' The only function you're likely to need from \pkg{roxygen2} is
-#' \code{\link{roxygenize}}. Otherwise refer to the vignettes to see
-#' how to format the documentation.
-#'
-#' @docType package
-#' @name SuperRanker
-#' @useDynLib SuperRanker
-#' @importFrom Rcpp sourceCpp
-NULL

### 2016-05-30 R/averageOverlap.R

@@ -1,5 +1,6 @@
 #' Compute the average overlap
 #'
+#'
 #' @param obj Either a vector or matrix
 #' @return A vector of the average overlap
 #' @export

### 2016-06-14 latex/superranker-manuscript.tex
Part I of corrections

@@
   The comparison of alternative rankings of a set of items is a
-  general and prominent task in applied statistics. Predictor
-  variables are ranked according to magnitude of association with an
-  outcome, prediction models rank subjects according to the
-  personalized risk of an event, and genetic studies rank genes
-  according to their difference in gene expression levels. This
-  article constructs measures of the agreement of two or more ordered
-  lists. We use the standard deviation of the ranks to define a
-  measure of agreement that both provides an intuitive interpretation
-  and can be applied to any number of lists even if some or all are
-  incomplete or censored. The approach can identify change-points in
-  the agreement of the lists and the sequential changes of agreement
-  as a function of the depth of the lists can be compared graphically
-  to a permutation based reference set.  The usefulness of these tools
-  are illustrated using gene rankings, and using data from two Danish
-  ovarian cancer studies where we assess the within and between
-  agreement of different statistical classification methods.
+  general and common task in applied statistics. Predictor variables
+  are ranked according to magnitude of association with an outcome,
+  prediction models rank subjects according to the personalized risk
+  of an event, and genetic studies rank genes according to their
+  difference in gene expression levels. This article constructs
+  measures of agreement among two or more ordered lists. We use the
+  standard deviation of the ranks to define a measure of agreement
+  that both provides an intuitive interpretation and can be applied to
+  any number of lists even if some or all are incomplete or
+  censored. The approach can identify change-points in the agreement
+  of the lists, and the sequential changes of agreement as a function
+  of the depth of the lists can be compared graphically to a
+  permutation based reference set or evaluated using asymptotic
+  results.  The usefulness of these tools are illustrated using gene
+  rankings, and using data from two Danish ovarian cancer studies
+  where we assess the within and between agreement of different
+  statistical classification methods.
 \end{abstract}
@@
 well-known, but these measures do not distinguish between agreement in
-the top versus towards the bottom of the lists and the provide a
+the top versus towards the bottom of the lists and they both provide a
 single measure of the overall distance. \citet{Shieh1998} proposed a
@@
 
-Other recent approaches consider the intersection of lists as the basis
+Other more recent approaches consider the intersection of lists as the basis
 for a similarity measure. However, simple intersection also places
@@
 \citet{Webber2010} proposed weighted intersections which put more
-emphasis on the top of the lists. One proposal is denoted the average
-overlap (AO) and used for comparison later on. Specifically,
-\citet{Webber2010} define their rank-biased overlap (RBO) by weighting
-with a converging series to ensure that the top is weighted higher
-than the potentially non-informative bottom of the lists.
-It is possible to use the existing methods to calculate agreement of
-lists until a given depth, i.e., limited to the $d$ items of each
-list. However, the interpretation may not be straightforward,
+emphasis on the top of the lists. 
+% One proposal is denoted the average overlap (AO) and used for
+% comparison later on. 
+Specifically, \citet{Webber2010} define their rank-biased overlap
+(RBO) by weighting with a converging series to ensure that the top is
+weighted higher than the potentially non-informative bottom of the
+lists.  It is possible to use the existing methods to calculate
+agreement of lists until a given depth, i.e., limited to the $d$ items
+of each list. However, the interpretation may not be straightforward,
 especially in the case of more than two lists, and they may not
@@
 based on letting the number of \emph{lists} increase to infinity which
-is a situation that is only relevant in special cases, and the
+is a situation that is only relevant in special cases, whereas the
 simulation-based null distribution approach of \citet{samp:verd:2013}
@@
 sequential rank agreement potentially allows us to infer a
-change-point, i.e., a list depth where a change in the agreement of
+change-point, i.e., a list depth where a substantial change in the agreement of
 the lists occur but we also provide randomization-based graphical
@@
 with possible extensions. The approaches presented in this manuscript
-are available in the R package \texttt{SuperRanker}. \mcomment{XXX check dette
-passer, submit pakke}
-
+are available in the R package \texttt{SuperRanker} which can be found
+on CRAN.
 % chunk 1
@@
 ordered list is the realization of a ranking function $R$ applied to
-the set of items $X$. In all what follows we condition on the set of
-items and consider the ranking function to be a random variable. Thus,
-let $R_1(X),\dots,R_L(X)$, $L\geq2$, be a sample of $L$ independent
+the set of items $X$.  Panels (a) and (b) of Table~\ref{tab:example}
+show a schematic example of these mappings. Thus if
+$R_l^{-1}(1)=X_{34}$ then item $X_{34}$ is ranked first in list $l$
+and similarly $R_l(X_{34})=1$.
+
+In all what follows we consider a fixed set of items and consider the
+ranking function to be a random variable. Thus, let
+$R_1(X),\dots,R_L(X)$, $L\geq2$, be a sample of $L$ independent
 identically distributed draws from an unknown probability distribution
 function $Q$. One aim is then to test how much $Q$ resembles the
-uniform distribution which assigns probability $(P!)^{-1}$ to each of
-the $P!$ different rankings.
-Panels (a) and (b) of
-Table~\ref{tab:example} show a schematic example of these
-mappings. Thus if $R_l^{-1}(1)=X_{34}$ then item $X_{34}$ is ranked
-first in list $l$ and similarly $R_l(X_{34})=1$.
-
-\bigskip\hrule HOV: er det virkelig equivalent ogsaa mht vores nye theoretiske
-ideer? Og bruger vi det nogen steder?  enten skal det forklares med et
-konkret eksempel eller ud?  We will use the same notation for the
-equivalent scenario where we have ordered lists obtained from applying
-the same ranking function to $L$ samples from the same population,
-i.e., for ranked lists $R(X^1), \ldots, R(X^L)$.  
-\hrule\bigskip
+uniform distribution which assigns probability $1/P!$ to each of
+the $P!$ different possible rankings.
 
 \begin{table}[tb]
-  \caption{Example set of ranked lists. (a) shows the ranked list of
+  \caption{Example set of ranked lists. (a) shows the ranked lists of
     items for each of three lists, (b) presents the ranks obtained by
@@
   A(X_p) &= \E_Q\left[\left(R(X_P)-\E_Q R(X_P)\right)^2\right]\\
-   &= \sum_{r=1}^{P!} \left(r(X_P)-\E_Q R(X_P)\right)^2 Q(r)\nn
-\end{align}
+   &= \sum_{r=1}^{P!} \left(r(X_P)-\E_Q R(X_P)\right)^2 Q(r),\nn
+\end{align}
+\textsf{Ikke så vild med notationen her ovenfor}
 where $\E_Q R(X_p)=\sum_{r=1}^{P!} r(X_p)Q(r)$ and $Q$ is a probability mass function
@@
 \begin{align}
-\widehat{A}_L(X_p) = \frac{1}{L-1}\sum_{i=1}^L (R_i(X_p) - \overline{R}_L(X_p))^2, \quad \overline{R}_L(X_p) = \frac{1}{L}\sum_{i=1}^L R_i(X_p)\label{eq:empDistance}
+\widehat{A}_L(X_p) = \frac{1}{L-1}\sum_{i=1}^L (R_i(X_p) - \overline{R}_L(X_p))^2, \quad \overline{R}_L(X_p) = \frac{1}{L}\sum_{i=1}^L R_i(X_p).\label{eq:empDistance}
 \end{align}
@@
 which is exemplified in Panel (c) of Table \ref{tab:example}.
-We then define the \emph{sequential rank agreement} as the weighted
+
+We define the \emph{sequential rank agreement} as the weighted
 expected agreement of the items found in the set $S(d)$:
@@
 
-Values of \textrm{sra} close to zero suggests that the lists agree
-while larger values suggest disagreement.
-Just like in method comparison studies we prefer smaller limits of
-agreement because that means that the differences between the methods
-is small. If the ranked lists are identical then the value of
-sequential rank agreement will be zero for all depths $d$.
-The sequential rank agreement is  equivalent to the pooled
-variance of the items found in $S(d)$, and it can be interpreted as the
-average distance of the individual rankings of the lists from the
-average ranking for each of the items we have seen until depth $d$.
-For an easier interpretation we suggest to plot $\widehat{\textrm{sra}}_L(d)^{1/2}$ 
-corresponding to the pooled standard deviation.
+Values of \textrm{sra} close to zero suggests that the lists agree on
+the rankings while larger values suggest disagreement.  Just like in
+method comparison studies we prefer smaller limits of agreement
+because that means that the differences between the methods is
+small. If the ranked lists are identical then the value of sequential
+rank agreement will be zero for all depths $d$.  The sequential rank
+agreement is equivalent to the pooled variance of the items found in
+$S(d)$, and it can be interpreted as the average distance of the
+individual rankings of the lists from the average ranking for each of
+the items we have seen until depth $d$.  For an easier interpretation
+we suggest to plot $\sqrt{\widehat{\textrm{sra}}_L(d)}$ corresponding
+to the pooled standard deviation.
 

### 2016-06-16 latex/superranker-manuscript.tex
Changes

@@
 With fully observed lists we can plot the sequential rank agreement
-\ref{def:sra} as a function of depth $d$. An example is seen in top
+\eqref{def:sra} as a function of depth $d$. An example is seen in top
 panels of Figure~\ref{fig:example1} where four analysis methods were
@@
 as described in \citet{Dudoit2002} and the four different analysis
-approaches were: marginal two-sample $t$ tests, marginal logistic
-regression analyses, elastic net logistic regression
+approaches were: marginal unequal variances two-sample $t$ tests,
+marginal logistic regression analyses, elastic net logistic regression
 \citep{friedman2010regularization}, and marginal maximum information
@@
 
-The sequential rank agreement plotted in Figure~\ref{fig:example1}
-shows how many ranks apart we on average expect the genes found in the
+The sequential rank agreement curve plotted in Figure~\ref{fig:example1}
+roughly shows how many ranks apart we on average expect the genes found in the
 top $d$ part of the lists to be for each depth $d$.  It can be seen
@@
 of the lists (smaller values on the $y$ axis) corresponding to
-\emph{better} agreement than in the tails of the lists. It is clear
+\emph{better} agreement than towards the tails of the lists. It is clear
 from the curve in Figure~\ref{fig:example1} that there is a
@@
    panels: Corresponding sequential rank agreement for the same data
-   but where only the top 20 ranked items are available. The blue and
+   but where only the top 20 ranked items are available and the rest
+   of the items have been censored. The blue and
    red areas correspond to the independent and randomized reference
@@
  of the lists. In this case the sequential rank agreement curve will
- start at a low level and then increase until it levels off as in
+ start at a low level and then increase until it levels off as seen in
  Figure~\ref{fig:example1}.
@@
 \subsection{Analysis of partial/censored lists}
-Incomplete or partial lists are also a common occurrence. They arise
-for example in case of missing data (items), when comparing top $d$
-list results from publications, or when some methods only rank a
-subset of the items. For example, penalized regression based on the
-Lasso provides a sparse set of predictors that have non-zero
+Incomplete or partial lists are a common occurrence that arise, for
+example, in case of missing data (items), when comparing top $d$ list
+results from publications, or when some methods only rank a subset of
+the items. For example, penalized regression based on the Lasso
+provides a sparse set of predictors that have non-zero
 coefficients. There is no obvious ordering of the set of predictors
@@
 partial ordering. Incomplete lists also occur when the analyst decides
-to censor all lists at depth $d$ such that the measurement of
-agreement is restricted to the first $d$ elements of each list, or
-when the analyst only wants to consider the ranking of items that have
-been found to be statistically significant.
+to censor all lists at depth $d$ (\ie, top $d$ lists) such that the
+measurement of agreement is restricted to the first $d$ elements of
+each list, or when the analyst only wants to consider the ranking of
+items that have been found to be statistically significant.
 
@@
 (e.g., the mean of the unassigned ranks) to the censored items since
-that would result in less variation of the ranks and hence less
+that would result in \emph{less} variation of the ranks and hence less
 variation of the agreement, and it would artificially introduce a
@@
 
-\subsection{Comparison to average overlap}
-The average overlap is a measure that is widely used in comparisons of
-two fully observed ranked lists \citep{Fagin2003,Webber2010} and the
-idea behind the average overlap closely resembles the sequential rank
-agreement. In this section we present an extension of the average
-overlap to more that two lists and compare this to sequential rank
-agreement.
-
-The overlap among $L$ lists observed until depth $d$ is defined as the
-proportion of items that are found in all $L$ lists compared to the
-possible number of items found in all lists observed until depth $d$:
-\begin{align}
-  O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}\label{def:overlap}
-\end{align}
-Typically the overlap is only considered for pairwise comparisons
-\citep{Bar-Ilan2006,Boulesteix2009} but formula \ref{def:overlap}
-is easily extended to the situation where $L>2$ as shown above.
-
-The average overlap at depth $d$ is defined as the average of the
-overlaps until depth $d$,
-\begin{equation}
-AO(d) = \frac1d\sum_{i=1}^d O(d).
-\end{equation}
-This definition directly ensures that further emphasis is put on items
-in the top of the lists since their item specific overlap contributes
-to the calculation of average overlap at higher depths.
-
-There are three main drawbacks of the average overlap method: First it
-is highly sensitive to single items in the top of one the lists.  The
-average overlap profile may change dramatically depending on the
-presence or absence of similar items in the beginning of the
-lists. For example, if all lists have the same item at the top then
-the average overlap attains its maximum at 1, but if just one of the
-lists differ then the overage overlap starts at its minimum of
-0. Secondly, when the number of lists, $L$, is high then it might be
-difficult to obtain a non-zero overlap (and hence a non-zero average
-overlap) at the beginning of the lists because the overlap will be
-zero until an item is present in all $L$ lists. Finally, the average
-overlap has an interpretation in terms of moving averages of
-percentages (of $L$ sets) which is somewhat less intuitive than our
-proposed sequential rank distance.
-
-In the following we compare the average overlap and the sequential
-rank agreement using the analysis of the Golub data
-\citep{Golub1999}. We consider a typical application of the average
-overlap and thus restrict attention to only two gene rankings obtained
-by the $t$ tests and logistic regression. These two marginal modeling
-approaches should result in very similar gene rankings. The first two
-columns of Table~\ref{tab1} show the actual top-10 ranking of the
-predictors.
-
-% chunks 8,9
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.49\textwidth]{paper-fig-ao1}
-\includegraphics[width=.49\textwidth]{paper-fig-ao2}
-\end{center}
- \caption{Average overlap (left figure) and sequential rank agreement
- (right figure) for comparing two different analysis methods
- (marginal $t$ test and logistic regression)
-    applied to the Golub data. The black lines are based on the full
-    set of predictors while the red lines have the two highest
-    associated predictors removed
-    from the data before computing the average overlap and sequential
-    rank agreement.}
-  \label{fig:case1}
- \end{figure}
-
- The black lines in Figure~\ref{fig:case1} show the average overlap
- and sequential rank agreement for these data. It is clear from both
- plots that there is perfect agreement towards the top of the ranked
- lists: the average overlap is 1 and the sequential rank agreement is
- zero.  However, if we remove those two items from the data (i.e.,
- gene/predictor 2124 and 896) and redo the analyses then we get
- substantially different curves for the average overlap but not for
- sequential rank agreement (red lines in Figure~\ref{fig:case1}). For
- the sequential rank agreement we get roughly the same estimate of
- agreement from rank 3 as we did from the full dataset. This indicates
- that sequential rank is more robust against small perturbations of
- the data. For the average overlap the curves completely change when
- these two items are removed and which would lead to a different
- conclusion about the agreement among the items from rank 3 downwards.
+% \subsection{Comparison to average overlap}
+% The average overlap is a measure that is widely used in comparisons of
+% two fully observed ranked lists \citep{Fagin2003,Webber2010} and the
+% idea behind the average overlap closely resembles the sequential rank
+% agreement. In this section we present an extension of the average
+% overlap to more that two lists and compare this to sequential rank
+% agreement.
+
+% The overlap among $L$ lists observed until depth $d$ is defined as the
+% proportion of items that are found in all $L$ lists compared to the
+% possible number of items found in all lists observed until depth $d$:
+% \begin{align}
+%   O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}\label{def:overlap}
+% \end{align}
+% Typically the overlap is only considered for pairwise comparisons
+% \citep{Bar-Ilan2006,Boulesteix2009} but formula \ref{def:overlap}
+% is easily extended to the situation where $L>2$ as shown above.
+
+% The average overlap at depth $d$ is defined as the average of the
+% overlaps until depth $d$,
+% \begin{equation}
+% AO(d) = \frac1d\sum_{i=1}^d O(d).
+% \end{equation}
+% This definition directly ensures that further emphasis is put on items
+% in the top of the lists since their item specific overlap contributes
+% to the calculation of average overlap at higher depths.
+
+% There are three main drawbacks of the average overlap method: First it
+% is highly sensitive to single items in the top of one the lists.  The
+% average overlap profile may change dramatically depending on the
+% presence or absence of similar items in the beginning of the
+% lists. For example, if all lists have the same item at the top then
+% the average overlap attains its maximum at 1, but if just one of the
+% lists differ then the overage overlap starts at its minimum of
+% 0. Secondly, when the number of lists, $L$, is high then it might be
+% difficult to obtain a non-zero overlap (and hence a non-zero average
+% overlap) at the beginning of the lists because the overlap will be
+% zero until an item is present in all $L$ lists. Finally, the average
+% overlap has an interpretation in terms of moving averages of
+% percentages (of $L$ sets) which is somewhat less intuitive than our
+% proposed sequential rank distance.
+
+% In the following we compare the average overlap and the sequential
+% rank agreement using the analysis of the Golub data
+% \citep{Golub1999}. We consider a typical application of the average
+% overlap and thus restrict attention to only two gene rankings obtained
+% by the $t$ tests and logistic regression. These two marginal modeling
+% approaches should result in very similar gene rankings. The first two
+% columns of Table~\ref{tab1} show the actual top-10 ranking of the
+% predictors.
+
+% % chunks 8,9
+
+% \begin{figure}[htbp]
+% \begin{center}
+% \includegraphics[width=.49\textwidth]{paper-fig-ao1}
+% \includegraphics[width=.49\textwidth]{paper-fig-ao2}
+% \end{center}
+%  \caption{Average overlap (left figure) and sequential rank agreement
+%  (right figure) for comparing two different analysis methods
+%  (marginal $t$ test and logistic regression)
+%     applied to the Golub data. The black lines are based on the full
+%     set of predictors while the red lines have the two highest
+%     associated predictors removed
+%     from the data before computing the average overlap and sequential
+%     rank agreement.}
+%   \label{fig:case1}
+%  \end{figure}
+
+%  The black lines in Figure~\ref{fig:case1} show the average overlap
+%  and sequential rank agreement for these data. It is clear from both
+%  plots that there is perfect agreement towards the top of the ranked
+%  lists: the average overlap is 1 and the sequential rank agreement is
+%  zero.  However, if we remove those two items from the data (i.e.,
+%  gene/predictor 2124 and 896) and redo the analyses then we get
+%  substantially different curves for the average overlap but not for
+%  sequential rank agreement (red lines in Figure~\ref{fig:case1}). For
+%  the sequential rank agreement we get roughly the same estimate of
+%  agreement from rank 3 as we did from the full dataset. This indicates
+%  that sequential rank is more robust against small perturbations of
+%  the data. For the average overlap the curves completely change when
+%  these two items are removed and which would lead to a different
+%  conclusion about the agreement among the items from rank 3 downwards.
 
@@
 \begin{eqnarray}
-H_0  & : &  \text{The list rankings correspond to complete randomly}\label{eq:permutationHypothesis}\\
+H_0  & : &  \text{The list rankings correspond to completely randomly}\label{eq:permutationHypothesis}\\
        &  & \text{permuted lists}\nn
@@
 \end{eqnarray*}
+
+\subsection{Permutation-based inference}
+
 $H_0$ is a quite unrealistic null hypothesis but we can easily obtain
@@
 censored/partially observed case we first permute the items
-$X_1,\dots,X_P$ and then censor list $l$ at $d_l$. The sequential rank
-agreement curve from the the original lists can then be compared to,
-say, the pointwise 95\% quantiles of the observed rank agreements
-obtained under $H_0$.
+$X_1,\dots,X_P$ and then censor list $l$ at $d_l$ (\ie, each list so
+it has the same number of observed rankings as was observed for list
+$l$ in the original dataset). The sequential rank agreement curve from
+the the original lists can then be compared to, say, the pointwise
+95\% quantiles of the observed rank agreements obtained under $H_0$.
 
@@
 
-\section{Asymptotics}\label{sec:chgpoint}
-In many applications it is of interest to estimate a list depths which
-satisfies a change point criterion. In this section we investigate the
+\subsection{Asymptotic inference of change in agreement}\label{sec:chgpoint}
+In many applications it is of interest to estimate a list depth which
+satisfies a change point criterion since that corresponds to a change
+in agreement among the list ranks. In this section we investigate the
 theoretical properties of our proposed method for this specific
-task. As in \cite{hall:schi:2012} we work with an infinite set
+task. As in \cite{hall:schi:2012} we consider an infinite set
 of lists and study the asymptotic behaviour for $L\to\infty$.
@@
 \begin{theorem}
-  Assume that $\{R_l(X)\}_{l=1}^L$ are independent draws from the a
+  Assume that $\{R_l(X)\}_{l=1}^L$ are independent draws from a
   probability distribution $Q$ on the set of lists.  Then,
@@
 \begin{align}
-  \mathcal{L}(q) = \left\{d : \textrm{sra}(d) \geq q(d)\right\}
+  \mathcal{L}(q) = \left\{d : \textrm{sra}(d) \geq q(d)\right\}.
 \end{align}
@@
 \end{align}
-where we allow that the threshold function can depend on the sample size as well. The estimated change-point is therefore
+where we allow the threshold function to depend on the sample size as well. The estimated change-point is therefore
 \begin{align}
@@
 
-\mcomment{Elaborate more on how to apply this in practice (finite sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style?}
+\mcomment{Elaborate more on how to apply this in practice (finite
+  sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style? Show
+some result althrough }
 
@@
 (\ref{eq:countingConv}). The first factor is $O(1)$ since $N(P;d)>0$
-uniformly in $d$ and because $A(X_p) = O(P^2)$ uniformly in $p$ by the Cauchy-Ekstr{\o}m inequality. It
+uniformly in $d$ and because $A(X_p) = O(P^2)$ uniformly in $p$ by the Cauchy-Schwarz inequality. It
 thus follows that $\left|U_L(d)\right| = o_P(1)$ uniformly in $d$.

### 2016-06-17 latex/superranker-manuscript.tex
Fixed part II

@@
 \begin{align}
-  \widehat{d^\ast_L}(\widehat{q}_L) &= 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| > 0)\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) + 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0)P
+  \widehat{d^\ast_L}(\widehat{q}_L) &= 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| > 0)\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) + 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0)P.
 \end{align}
@@
 The consistency of the estimated change-point, $\widehat{d^\ast_L}(\widehat{q}_L)$, follows from Theorem \ref{thm:consistency} by the following corollary. 
-\begin{corollary}
+\begin{corollary}\label{thecor}
 Let $\widehat{q}_L$ be a threshold function such that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ for some limiting
@@
 
+Corollary \ref{thecor} tells us that we can use the a threshold
+function $\widehat{q}_L$ estimated under the null hypothesis as
+discussed in the previous section as a limiting threshold function for
+inferring the depth $d$, where the observed sequential rank agreement
+first crosses the threshold of the null threshold, \ie, the depth
+until which the observed ranked lists are in better agreement than
+expected under the null hypothesis. In that sense the threshold
+function serves the same role as the limits of agreement in method
+comparison studies, except that the threshold funciton is not constant
+but can accommodate the changing nature of the number of items used
+for the computation of the equential rank agreement for a given depth.
+
+In practice we can compute an estimate of the threshold function under
+the null using the permutation approach sketched in the previous
+section which makes it relevant even in small sample settings.
+
 \mcomment{Elaborate more on how to apply this in practice (finite
-  sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style? Show
-some result althrough }
+  sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style?  }
 
@@
 In conclusion we have introduced a method for evaluation of ranked
-(censored) lists that can be easily interpreted and that can be
+(partial/censored) lists that can be easily interpreted and that can be
 applied to a large number of situations.  The method presented here

### 2016-07-11 R/RcppExports.R
fixed examples and improved help texts

@@ -12,3 +12,14 @@ overlap <- function(rankMat) {
     .Call('SuperRanker_overlap', PACKAGE = 'SuperRanker', rankMat)
 }
 
+#' Compute the sequential rank agreement between k ranked lists
+#' 
+#' @description Computes the sequential rank agreement (number of items present in all k lists divided by the current rank) for each rank in the k lists
+#' @param rankMat A matrix with k columns corresponding to the k ranked lists. Elements of each column are integers between 1 and the length of the lists
+#' @param type The type of distance measure to use: 0 (the default) is the variance while 1 is MAD (mean absolute deviation)
+#' @return A vector of the same length as the rows in rankMat containing the sequential rank agreement between the lists for each depth (squared for type=0)
+#' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
+sracppfull <- function(rankMat, type = 0L) {
+    .Call('SuperRanker_sracppfull', PACKAGE = 'SuperRanker', rankMat, type)
+}
+

### 2016-07-11 R/sra.R

@@ -160,18 +160,18 @@ sra.list <- function(object, B=1, na.strings=NULL, nitems=max(sapply(object, len
     # Special version of sample needed
     resample <- function(x, ...) x[sample.int(length(x), ...)]
     tmpres <- sapply(1:B, function(b) {
-                         obj.b <- lapply(1:nlists,function(j){
-                                             list <- object[[j]]
-                                             if (nmiss[[j]]>0){
-                                                 list[list==0] <- resample(missing.items[[j]])
-                                             }
-                                             list
-                                         })
-                         ## bind lists
-                         rankmat <- do.call("cbind",obj.b)
-                         res <- sracppfull(rankmat, type=itype)
-                         res
-                     })
+        obj.b <- lapply(1:nlists,function(j){
+            list <- object[[j]]
+            if (nmiss[[j]]>0){
+                list[list==0] <- resample(missing.items[[j]])
+            }
+            list
+        })
+        ## bind lists
+        rankmat <- do.call("cbind",obj.b)
+        res <- sracppfull(rankmat, type=itype)
+        res
+    })
     if (itype==0) {
         agreement <- sqrt(rowMeans(tmpres))
     } else {
@@ -188,25 +188,27 @@ sra.list <- function(object, B=1, na.strings=NULL, nitems=max(sapply(object, len
 
 #' Simulate sequential rank agreement for randomized unrelated lists
 #'
-#' Simulate sequential rank agreement from completely uninformative lists (ie., raw permutations of items) and compute the corresponding sequential rank agreement curve.
-#' The number of lists, items and amount of censoring is identical to the input object.
+#' Simulate sequential rank agreement from completely uninformative lists (ie., raw permutations of items) and compute the corresponding sequential rank agreement curves.
+#' The following attributes are copied from the input object: number of lists, number of items and amount of censoring.
 #'
 #' @param object A matrix or list of vectors representing ranked lists.
-#' @param B An integer giving the number of randomization to sample
+#' @param B An integer giving the number of randomizations to sample
 #'     over in the case of censored observations
-#' @param n the number of sequential rank agreement curves to simulate
[…]

### 2016-07-11 src/sra.cpp

@@ -204,14 +204,14 @@ NumericVector sracpp(IntegerMatrix rankMat, int maxlength, int B, IntegerVector 
 }
 
 
-// ' Compute the sequential rank agreement between k ranked lists
-// ' 
-// ' @description Computes the sequential rank agreement (number of items present in all k lists divided by the current rank) for each rank in the k lists
-// ' @param rankMat A matrix with k columns corresponding to the k ranked lists. Elements of each column are integers between 1 and the length of the lists
-// ' @param type The type of distance measure to use: 0 (the default) is the variance while 1 is MAD (mean absolute deviation)
-// ' @return A vector of the same length as the rows in rankMat containing the sequential rank agreement between the lists for each depth (squared for type=0)
-// ' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
-// ' [[Rcpp::export]]
+//' Compute the sequential rank agreement between k ranked lists
+//' 
+//' @description Computes the sequential rank agreement (number of items present in all k lists divided by the current rank) for each rank in the k lists
+//' @param rankMat A matrix with k columns corresponding to the k ranked lists. Elements of each column are integers between 1 and the length of the lists
+//' @param type The type of distance measure to use: 0 (the default) is the variance while 1 is MAD (mean absolute deviation)
+//' @return A vector of the same length as the rows in rankMat containing the sequential rank agreement between the lists for each depth (squared for type=0)
+//' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
+// [[Rcpp::export]]
 NumericVector sracppfull(IntegerMatrix rankMat, int type=0) {
 
   // The number of lists
@@ -268,4 +268,6 @@ NumericVector sracppfull(IntegerMatrix rankMat, int type=0) {
     
     return(returnVector);
 }
+
+
 

### 2016-07-11 latex/superranker-manuscript.tex

@@
 function $Q$. One aim is then to test how much $Q$ resembles the
-uniform distribution which assigns probability $1/P!$ to each of
-the $P!$ different possible rankings.
+uniform distribution which assigns probability $1/P!$ to each of the
+$P!$ different possible rankings.
 
@@
 \begin{align}
-  A(X_p) &= \E_Q\left[\left(R(X_P)-\E_Q R(X_P)\right)^2\right]\\
-   &= \sum_{r=1}^{P!} \left(r(X_P)-\E_Q R(X_P)\right)^2 Q(r),\nn
-\end{align}
-\textsf{Ikke så vild med notationen her ovenfor}
-where $\E_Q R(X_p)=\sum_{r=1}^{P!} r(X_p)Q(r)$ and $Q$ is a probability mass function
-on the space of all permutations of lists with $P$ elements.
-The empirical counterpart is
+  A(X_p) &= \E_Q\left[\left(R(X_p)-\E_Q R(X_p)\right)^2\right]\\
+   &= \sum_{r\in\Pi} \left(r(X_p)-\E_Q R(X_p)\right)^2 Q(r),\nn
+\end{align}
+where $\Pi$ is the set of all permutations of $X$, $Q$ is a
+probability mass function on $\Pi$, and
+$\E_Q R(X_p)=\sum_{r\in\Pi} r(X_p)Q(r)$. The empirical counterpart is
 \begin{align}
@@
 \begin{align}
-S(d) &= \left\{X_p; \left(\sum_{r=1}^{P!} 1\left(r(X_p)\le d\right) Q(r)\right) >0\right\}\label{eq:sumset}
+S(d) &= \left\{X_p; \left(\sum_{r\in\Pi} 1\left(r(X_p)\le d\right) Q(r)\right) >0\right\}\label{eq:sumset}
 \end{align}
@@
   Assume that $\{R_l(X)\}_{l=1}^L$ are independent draws from a
-  probability distribution $Q$ on the set of lists.  Then,
+  probability distribution $Q$ on the set of lists $\Pi$. Then,
   $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty =
@@
 \end{theorem}
-
-We now define the parameter of interest. For any given deterministic threshold function $q\colon\, 1,\ldots,P \mapsto \mathbb{R}_{\geq 0}$
-we define the population superlevel set of the sequential rank agreement as
+\proof See Appendix \ref{sec:appA}.
+
+We now define the change point as the first crossing point of the
+sequential rank agreement and a threshold function
+$q\colon\,\{1,\ldots,P\} \mapsto \mathbb{R}_{\geq 0}$. The values of
+$q$ could be a deterministic constant, or in the experimental setting,
+for example, the limits-of-agreement obtained in randomly permuted
+lists corresponding to the null-hypothesis in equation
+(\ref{eq:permutationHypothesis}). We define the superlevel set of the
+sequential rank agreement with respect to $q$ as
 \begin{align}
@@
 \end{align}
-The values of $q$ could for example be the limits-of-agreement for randomly
-permuted lists as in the null-hypothesis in equation (\ref{eq:permutationHypothesis}).
 A change-point $d^\ast(q)$ in the list agreement is then defined by the position
-\begin{align}
-  d^\ast(q) &= 1\left(|\mathcal{L}(q)| > 0\right)\inf \mathcal{L}(q) + 1\left(|\mathcal{L}(q)| = 0\right)P
-\end{align}
-corresponding to the first list depth where the sequential rank agreement exceeds the threshold 
-if such a position exists. Otherwise, the full list is in agreement according to $q$ and the change-point
-is defined as the length of the lists. The empirical superlevel set is similarly defined as
+\begin{equation}
+  d^\ast(q) =
+  \begin{cases}
+    \inf(\mathcal{L}(q)) & |\mathcal{L}(q)| > 0\\
+  P & |\mathcal{L}(q)| = 0
+  \end{cases}
+\end{equation}
+corresponding to the first list depth where the sequential rank
+agreement exceeds the threshold if such a position exists. Otherwise,
+the full list is in agreement according to $q$ and the change-point is
+set to the length of the lists. The empirical superlevel set is
+similarly defined as
 \begin{align}
@@
 \end{align}
-where we allow the threshold function to depend on the sample size as well. The estimated change-point is therefore
+where we allow the threshold function to depend on the sample size as
+well. The estimated change-point is therefore
 \begin{align}
@@
 \end{corollary}
-
-Corollary \ref{thecor} tells us that we can use the a threshold
+\proof See Appendix \ref{sec:appB}.
+
+Corollary \ref{thecor} indicates that we can use the a threshold
 function $\widehat{q}_L$ estimated under the null hypothesis as
@@
 
-\mcomment{Elaborate more on how to apply this in practice (finite
-  sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style?  }
+% \mcomment{Elaborate more on how to apply this in practice (finite
+  % sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style?  }
 
@@
 \section{Proof of Theorem \ref{thm:consistency}}
+\label{sec:appA}
 
@@
 \section{Proof of Corollary \ref{cor:consistency}}
-Let $A \triangle B = (A \setminus B) \cup (B \setminus A)$ is the symmetric difference between sets $A$ and $B$. It is then sufficient to show that
-$|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q)|$ converges to zero in probability as $L \rightarrow \infty$. By the triangle inequality
-we have that
+\label{sec:appB}
+
+Recall that $\mathcal{L}$ is the superlevel set of list positions
+where the sequential rank agreement exceeds the threshold function
+$q$. Let $A \triangle B = (A \setminus B) \cup (B \setminus A)$ is the
+symmetric difference between sets $A$ and $B$. It is then sufficient
+to show that
+$|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q)|$
+converges to zero in probability as $L \rightarrow \infty$. By the
+triangle inequality we have that
 \begin{align}

### 2016-07-19 R/plot.sra.R
CRAN version (plus spelling check)

@@ -26,6 +26,7 @@
 ##' })
 ##' c(mean(aa),sd(aa))
 ##'
+##' @importFrom graphics plot axis lines
 #' @export
 ##' @author Thomas A. Gerds <tag@@biostat.ku.dk>
 plot.sra <- function(x, xlim, ylim, xlab="List depth", ylab="Sequential rank agreement", add=FALSE, ...){

### 2016-07-19 latex/superranker-manuscript.tex
small fixes

@@
 \newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
+\newcommand{\plim}{\operatornamewithlimits{plim}}
 
@@
 \label{thm:consistency}
+\begin{proof}
+See Appendix \ref{sec:appA}.
+\end{proof}
 \end{theorem}
-\proof See Appendix \ref{sec:appA}.
 
@@
 \begin{corollary}\label{thecor}
-Let $\widehat{q}_L$ be a threshold function such that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ for some limiting
+Let $\widehat{q}_L$ be a positive threshold function such that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ for some limiting
 function $q$.  Then, $\widehat{d^\ast_L}(\widehat{q}_L) \overset{P}{\longrightarrow} d^\ast(q)$ for $L \rightarrow \infty$.
 \label{cor:consistency}
+\begin{proof}
+See Appendix \ref{sec:appB}.
+\end{proof}
 \end{corollary}
-\proof See Appendix \ref{sec:appB}.
 
@@
 identically Bernoulli distributed variables with expectation
-$Q(R(X_s) \leq d)$ it follows from the law of large numbers that
+$Q(R(X_s) \leq d)$ and finite variance it follows from the law of large numbers that
 $\widehat{Q}_L(R(X_p) \leq d) \overset{P}{\longrightarrow} Q(R(X_p)
@@
 \label{sec:appB}
-
 Recall that $\mathcal{L}$ is the superlevel set of list positions
 where the sequential rank agreement exceeds the threshold function
-$q$. Let $A \triangle B = (A \setminus B) \cup (B \setminus A)$ is the
+$q$. Let $A \triangle B = (A \setminus B) \cup (B \setminus A)$ be the
 symmetric difference between sets $A$ and $B$. It is then sufficient
-to show that
-$|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q)|$
-converges to zero in probability as $L \rightarrow \infty$. By the
-triangle inequality we have that
-\begin{align}
- \left|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q)\right| &= \left|\left(\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \widehat{\mathcal{L}}_L(q)\right) \cup \left(\widehat{\mathcal{L}}_L(q) \triangle \mathcal{L}(q)\right)\right|\label{eq:corSetDiff}\\
-   &\leq \left|\widehat{\mathcal{L}}_L(q) \triangle \mathcal{L}(q)\right| + \left|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \widehat{\mathcal{L}}_L(q)\right|\nn
-\end{align}
-Looking at the first term we have
-\begin{align}
-  \left|\widehat{\mathcal{L}}_L(q) \triangle \mathcal{L}(q)\right| &= \sum_{d=1}^P \left|1\left(\widehat{\textrm{sra}}_L(d) \geq q(d)\right)  - 1\left(\textrm{sra}(d)  \geq q(d)\right)\right|\\
-   &=
-\end{align}
-Similarly we have
-\begin{align}
-\left|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \widehat{\mathcal{L}}_L(q)\right| &= \sum_{d=1}^P \left|1\left(\widehat{\textrm{sra}}_L(d) \geq \widehat{q}_L(d)\right) - 1\left(\widehat{\textrm{sra}}_L(d) \geq q(d)\right)\right|\\
-   &=
-\end{align}
-
-%We have that 	
-%\begin{align}
-%  |\widehat{d^\ast_L}(\widehat{q}_L) - d^\ast(q)| &\leq  |\widehat{d^\ast_L}(q)  - d^\ast(q)| + |\widehat{d^\ast_L}(\widehat{q}_L)  - \widehat{d^\ast_L}(q_L)|
-%\end{align}
-%The first term is of order $o_P(1)$ by the continuous mapping theorem if a) $|\inf \widehat{\mathcal{L}}_L(q) - \inf  \mathcal{L}(q)| = o_P(1)$ and b) $|1(|\widehat{\mathcal{L}}_L(q)| = 0) - 1(|\mathcal{L}(q)| = 0)| = o_P(1)$ since $1(|\mathcal{L}(q)| > 0) = 1 - 1(|\mathcal{L}(q)| = 0)$.
-Similarly, the second term is of order $o_P(1)$ by the same reasoning if c) $|\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) - \inf \widehat{\mathcal{L}}_L(q)| = o_P(1)$ and d) $|1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0) - 1(|\widehat{\mathcal{L}}_L(q)| = 0)| = o_P(1)$.
-
-This completes the proof.
+to show that $\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q) \overset{P}{\longrightarrow} \emptyset$
+for $L \rightarrow \infty$ where $\emptyset$ denotes the empty set.
+We have that
+\begin{align}
+  \plim_{L \rightarrow \infty}\left(\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q)\right) &=   \left(\plim_{L \rightarrow \infty}\widehat{\mathcal{L}}_L(\widehat{q}_L)\right) \triangle \mathcal{L}(q)\\
+  &= \left\{d : \plim_{L\rightarrow \infty} \left(\widehat{\textrm{sra}}_L(d) - \widehat{q}_L(d)\right)\geq 0\right\} \triangle \mathcal{L}(q)\nn\\
+  &= \left\{d : \textrm{sra}(d) - q(d) \geq 0\right\}\triangle \mathcal{L}(q)\nn\\
+  &= \mathcal{L}(q)\triangle \mathcal{L}(q)\nn\\
+  &= \emptyset\nn
+\end{align}
+as a consequence of Theorem \ref{thm:consistency}, the assumption that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ and the continuous mapping theorem. This completes the proof.
 

### 2016-08-02 latex/superranker-manuscript.tex
Smårettelser

@@
   of the depth of the lists can be compared graphically to a
-  permutation based reference set or evaluated using asymptotic
+  permutation based reference set or evaluated using analytical asymptotic
   results.  The usefulness of these tools are illustrated using gene
@@
 
-Other more recent approaches consider the intersection of lists as the basis
+Other recent approaches consider the intersection of lists as the basis
 for a similarity measure. However, simple intersection also places
@@
 \citep{alt:bland:1983,Carstensen2010} but any measure for agreement
-could be essentially be used. Our proposed approach allows us to
-compare more than two lists simultaneously, it provides a dynamic
-measure of agreement as a function of the depth in the lists, places
-high weight on items at the top of the list, accommodates partially
+could essentially be used. Our proposed approach allows us to
+compare multiple lists simultaneously, it provides a dynamic
+measure of agreement as a function of the depth in the lists, it places
+higher weight on items at the top of the list, it accommodates partially
 observed lists of varying lengths, and has a natural interpretation
@@
 change-point, i.e., a list depth where a substantial change in the agreement of
-the lists occur but we also provide randomization-based graphical
+the lists occur but we also provide asymptotical and randomization-based graphical
 tools to compare the observed rank agreement to the expected agreement
@@
 sequential rank agreement for multiple ranked lists and discuss how to
-handle incomplete/censored/partial lists. In section 3 we present and
+handle incomplete lists. In section 3 we present and
 discuss approaches to evaluate the results obtained from sequential
@@
 
-The agreement of the lists regarding the rank given to an item $X_p$
+The agreement among the lists regarding the rank given to an item $X_p$
 can be measured by the variance across the lists
@@
 method comparison studies we prefer smaller limits of agreement
-because that means that the differences between the methods is
+because that means that the differences between the rankings is
 small. If the ranked lists are identical then the value of sequential

### 2016-08-04 latex/superranker-manuscript.tex
Small changes 2

@@
   any number of lists even if some or all are incomplete or
-  censored. The approach can identify change-points in the agreement
+  censored. The approach can identify changepoints in the agreement
   of the lists, and the sequential changes of agreement as a function
@@
 sequential rank agreement potentially allows us to infer a
-change-point, i.e., a list depth where a substantial change in the agreement of
+changepoint, i.e., a list depth where a substantial change in the agreement of
 the lists occur but we also provide asymptotical and randomization-based graphical
@@
 
-The simplest case occurs when all $L$ lists are fully observed, \ie,
-we have observed the rank of all $P$ items for all $L$ lists. Fully
+The simplest case occurs when all $L$ lists are fully observed, \ie, when
+we observe the rank of all $P$ items for all $L$ lists. Fully
 observed ranked lists are common and arise, for example, when
@@
 With fully observed lists we can plot the sequential rank agreement
-\eqref{def:sra} as a function of depth $d$. An example is seen in top
+\eqref{def:sra} as a function of depth $d$. An example is seen in the top
 panels of Figure~\ref{fig:example1} where four analysis methods were
@@
 that the sequential rank agreement limits are better towards the top
-of the lists (smaller values on the $y$ axis) corresponding to
-\emph{better} agreement than towards the tails of the lists. It is clear
+of the lists (smaller values on the $y$ axis corresponds to
+\emph{better} agreement) than towards the tails of the lists. It is clear
 from the curve in Figure~\ref{fig:example1} that there is a
@@
   \hline
-Ranking & T & LogReg & ElasticNet & MIC \\ 
+Ranking & Welsh's $t$ & LogReg & ElasticNet & MIC \\ 
   \hline
@@
  \caption{Left panels: Sequential rank agreement for 4 different
-   analysis methods applied to the 3051 genes in the Golub data /black
+   analysis methods applied to the 3051 genes in the Golub data (black
    line). Right
@@
    but where only the top 20 ranked items are available and the rest
-   of the items have been censored. The blue and
+   of the items have been censored in each list. The blue and
    red areas correspond to the independent and randomized reference
@@
 
- If the lists agree on all items then the sequential rank agreement is
- zero for all depths. Generally, if there are changes in the level of
+% If the lists agree on all items then the sequential rank agreement is
+% zero for all depths. 
+ Generally, if there are changes in the level of
  rank agreement then this suggests that there are sets of items that
  are ranked similarly in all lists and other sets of items that have
- been assigned vastly different ranks in the lists. In gene
+ been assigned vastly different ranks in the lists. For example, in gene
  association studies we generally expect the agreement among the lists
@@
 whose coefficient has been shrunk to zero and thus we end up with a
-partial ordering. Incomplete lists also occur when the analyst decides
+partial ordering. Incomplete lists also occur if the analyst decides
 to censor all lists at depth $d$ (\ie, top $d$ lists) such that the
@@
 once in each list. Thus, we cannot simply assign the same number
-(e.g., the mean of the unassigned ranks) to the censored items since
+(\eg, the mean of the unassigned ranks) to the censored items since
 that would result in \emph{less} variation of the ranks and hence less
@@
 constant but at a low level). For partially ranked lists we also
-expect a change-point around the depth where the lists are
+expect a changepoint around the depth where the lists are
 censored. This is an artefact stemming from the fact that we assume
@@
 \begin{eqnarray*}
-\widetilde H_0 & :&  \text{The list rankings are based on data containing}\\
+\widetilde H_0 & :&  \text{The list rankings are based on data that contain}\\
 & &   \text{no association to the outcome.}
@@
 repeat the ranking procedures for unassociated data many times.  Thus,
-we first permute the outcome variable of the data set on which the
+we first permute the outcome variable of the dataset on which the
 rankings are based. This removes any association between the predictor
 variables and the outcome. Then we apply the methods to the permuted
-data set to generate $L$ new rankings and subsequently compute the
+dataset to generate $L$ new rankings and subsequently compute the
 sequential rank agreement. Note that we only permute the outcomes and
-preserve the structure of the candidate predictors. The randomization
+preserve the structure of the candidate predictors. This randomization
 approach requires that we have the original data available and as such
@@
 is better than what would be expected by chance for data that contain
-no information since it lies below the reference areas. However, the
-censored data also suggests that there may be at most 1 or 2 ranked
-items towards the top of the lists that yield a result better than
-what would be expected (the bottom-right plot).
+no information since it lies below the reference areas. The censored
+data also suggests that there may be at most 1 or 2 ranked items
+towards the top of the lists that yield a result better than what
+would be expected (the bottom-right plot).
 
@@
 related to questions regarding the association between the outcome and
-the predictors in the data set. Both hypotheses are purely considering
+the predictors in the dataset. Both hypotheses are purely considering
 how the rankings agree in a situation where there is no relevant
@@
 In many applications it is of interest to estimate a list depth which
-satisfies a change point criterion since that corresponds to a change
+satisfies a changepoint criterion since that corresponds to a change
 in agreement among the list ranks. In this section we investigate the
@@
 
-We now define the change point as the first crossing point of the
+We now define the changepoint as the first crossing point of the
 sequential rank agreement and a threshold function
 $q\colon\,\{1,\ldots,P\} \mapsto \mathbb{R}_{\geq 0}$. The values of
-$q$ could be a deterministic constant, or in the experimental setting,
+$q$ could be a deterministic constant or, % in the experimental setting,
 for example, the limits-of-agreement obtained in randomly permuted
@@
 \end{align}
-A change-point $d^\ast(q)$ in the list agreement is then defined by the position
+A changepoint $d^\ast(q)$ in the list agreement is then defined by the position
 \begin{equation}
@@
 agreement exceeds the threshold if such a position exists. Otherwise,
-the full list is in agreement according to $q$ and the change-point is
-set to the length of the lists. The empirical superlevel set is
+the full list is in agreement according to $q$ and the changepoint is
+set to the full length of the lists. The empirical superlevel set is
 similarly defined as
@@
 where we allow the threshold function to depend on the sample size as
-well. The estimated change-point is therefore
+well. The estimated changepoint is therefore
 \begin{align}
@@
 
-The consistency of the estimated change-point, $\widehat{d^\ast_L}(\widehat{q}_L)$, follows from Theorem \ref{thm:consistency} by the following corollary. 
+The consistency of the estimated changepoint, $\widehat{d^\ast_L}(\widehat{q}_L)$, follows from Theorem \ref{thm:consistency} by the following corollary. 
 \begin{corollary}\label{thecor}
@@
 We now consider an application of the sequential rank agreement to two
-data sets consisting of MALDI-TOF (Matrix-Assisted Laser
+datasets consisting of MALDI-TOF (Matrix-Assisted Laser
 Desorption/Ionization Time Of Flight) mass spectra obtained from blood
 samples from patients with either benign or malignant ovarian
-tumors. The data sets are sub-samples of the Danish MALOVA and DACOVA
+tumors. The datasets are sub-samples of the Danish MALOVA and DACOVA
 study populations.
@@
 the $5000$ predictors to have column-wise zero mean and unit variance in
-each data set.
-
-We use the two data sets to illustrate how the sequential rank
+each dataset.
+
+We use the two datasets to illustrate how the sequential rank
 agreement can be applied in two different scenarios. In the first
@@
 higher in the top as compared to the tail of the lists for all these
-classification methods. The change-points where the curves become
+classification methods. The changepoints where the curves become
 approximately constant are the list depths where the ranks of the
@@
 equal absolute value. This is a general feature of the Ridge models in
-this data set and seen for all the 1000 trained models.
+this dataset and seen for all the 1000 trained models.
 This implies that when predictors are ranked according to the
@@
 To assess the stability of the individual risk predictions we apply
-the predictors from the DACOVA data set to each of the models. The
+the predictors from the DACOVA dataset to each of the models. The
 predicted probabilities are then ranked in decreasing order such that the
@@
 rank agreement curve and the mean null rank agreement curve as a test
-statistic. Alternatively, a change-point analysis could be made on the
+statistic. Alternatively, a changepoint analysis could be made on the
 sequential rank agreement in order to determine the depths at which

### 2016-08-04 latex/superranker-manuscript.tex
Smårettelser

@@
 which runs through the list elements in an arbitrary order and counts
-how many list elements have a strictly positive probability to have
+how many list elements that have a strictly positive probability to have
 rank less than or equal to $d$ under the probability measure $Q$.

### 2017-03-31 latex/superranker-manuscript.tex
Slowpoke simulation draft done

@@
 sequential rank agreement for multiple ranked lists and discuss how to
-handle incomplete lists. In section 3 we present and
-discuss approaches to evaluate the results obtained from sequential
-rank agreement. Finally we apply the sequential rank agreement to two
-Danish ovarian cancer studies before we discuss the findings along
-with possible extensions. The approaches presented in this manuscript
-are available in the R package \texttt{SuperRanker} which can be found
-on CRAN.
+handle incomplete lists. In section 3 we present and discuss
+approaches to evaluate the results obtained from sequential rank
+agreement. Finally we apply the sequential rank agreement to two
+Danish ovarian cancer studies and compare our method to the method of
+\citet{hall:schi:2012} in a small sample simulation study before we
+discuss the findings along with possible extensions. The approaches
+presented in this manuscript are available in the R package
+\texttt{SuperRanker} which can be found on CRAN.
 % chunk 1
@@
 
+\section{Simulation study --- comparison of list agreements}
+
+We conducted a simulation study to investigate the small sample list
+depth agreement properties for our proposed method and to compare with
+the method propsed by \citet{hall:schi:2012}. For each of the
+simulations we assumed an set of 1000 predictors and a total of 400
+observations. The predictors were independent and drawn from a
+standard Gaussian distribution. The outcome was also drawn from a
+Gaussian distribution with variance 1 but with a mean that depends
+additively on the first 15 predictors, i.e.,
+
+$$y_i = \sum_{j=1}^{15}x_{ij} + \epsilon_i,$$
+
+where $y_i$ is the $i$th response, and $x_{ij}$ is the $j$th predictor
+for the $i$th measurement, and $\epsilon_i\sim N(0, 1)$. We produce
+$L$ ranked lists from a simulated dataset by first generating
+non-parametric bootstrap copies of the data, and then for each of the
+$L$ bootstrap sample ranking the 1000 predictors according to the size
+of their marginal $t$ test statistics. Thus, the simulation results
+compare lists that are analysed with the same technique on slightly
+different versions of the same data.
+
+In the simulations we focused on two main cases:
+\begin{enumerate}
+\item The choice of threshold, $q$. Throughout the simulations we
+  assume a constant, prespecified threshold function, $q_L$, and try
+  different values for the constant. The approach by
+  \citet{hall:schi:2012} uses a window size parameter, $K$, to
+  represent the maximum difference in rankings before they consider an
+  item in two lists to be in agreement.  Because the standard
+  deviation --- which is used in the sequential rank agreement
+  definition of $\hat{A}_L$--- squares its differences, it tends to
+  give more weight to larger rank differences and less weight to
+  smaller differences compared to the mean absolute difference.  To
+  ease comparison between the sequential rank agreement and the
+  \citet{hall:schi:2012} approach we would like to use the same
+  threshold for both methods, and in order to do this we use the
+  median absolute deviance throughout the simulations in lieu of the
+  pooled standard deviation.
+ \item The impact of the number of available lists, $L$. Asymptotic
+   results are available for both methods as the number of lists
+   increases, so this will investigate how much the results might
+   change as $L$ increases. For a fixed constant threshold of $q_L=5$
+   we increase the number of available lists to investigate the
+   sensitivity of the results to $L$.
+ \end{enumerate}
+
+ For all combinations we used 1000 simulations to see the distribution
+ of the estimated depth of agreement and we also computed the average
+ number of different predictors found for the given depth.
+
+ The results for varying threshold is shown in the left-hand plot of
+ Figure~\ref{fig:example99}, where the threshold/window size is varied
+ from 3 to 10 for a fixed number of lists $L=8$. The double histograms
+ show the distribution of the estimated depth of agreement, while the
+ bold numbers show the average number of unique predictors found given
+ the estimated depth of agreement. We see that the sequential rank
+ agrement method results in a substantially lower depth of agreement
+ than the \citet{hall:schi:2012} method (denoted topK) and also that
+ the average number of unique predictors is markedly smaller. Even
+ larger differences are found if we used the Euclidian distance
+ instead of the median absolute distance for the sequential rank
+ agreement (output not shown).  Not surprisingly, the depth of
+ agreement increases when the acceptable threshold for
+ agreement/window increases.
+
+ \begin{figure}[htbp]
+   \begin{center}
+ \includegraphics[width=.49\textwidth]{Ksim.pdf}
+ \includegraphics[width=.49\textwidth]{Lsim}
+ \end{center}
+ \caption{Left panels: Simulation study showing distribution of
+   estimated rank agreements for sequential rank agreement and the
+   topK for varying thresholds and fixed number of lists $L=8$. The bold numbers are the average
+   number of unique predictors in the set found to be in agreement.  Right
+   panels: Simulation results for varying number of lists, and with
+   fixed threshold of $q=5$.}
+ \label{fig:example99}
+ \end{figure}
+
+ Figure~\ref{fig:example99} also shows the results when the threshold
+ is fixed at $q=5$ and the number of lists increases. The difference
+ in agreement between the two methods is still noticeable with the
+ sequential rank agreement being more conservative than topK which
+ consequently results in fewer predictors among the part of the lists,
+ where they are in agreement.  However, the number of lists, $L$,
+ seems to have little impact on the distribution of the depth of
+ agreement and consequently on the number of different predictors
+ among the parts of the lists in agreement.
+
+ For fixed $q$, the two methods are clearly different in the depth
+ they find to be in agreement with the sequential rank agreement
+ method being the more conservative.
+
 \section{Discussion}

### 2017-03-31 R/sra.R

@@ -12,7 +12,7 @@
 #'     are censored values.
 #' @param nitems The total number of items in the original lists if we only have partial lists available.
 #' @param type The type of measure to use. Either sd (standard
-#'     deviation - the default) or mad (median absolute deviance)
+#'     deviation - the default) or mad (median absolute deviance around the median)
 #' @param ... Arguments passed to methods.
 #' @return A vector of the sequential rank agreement
 ##' @examples

### 2017-04-07 latex/superranker-manuscript.tex
forkortelse af simu sektion

--- 2016-08-04 latex/superranker-manuscript.tex
@@
 Lasso shows similar agreement for ranking the risk predictions than
-PLS-DA (unless with the AUC criterion), and poorer agreement for ranking
-predictors. This reason for the latter is the high auto-correlation
-between the intensities in the mass spectra which leads to
-collinearity issues in the regression models. It is well-known that
-variable selection with the Lasso penalty does not preform very well
-when the predictors are highly correlated. The collinearity does,
-however, not affect the agreement of the risk predictions, as seen on
-the left panel of Figure \ref{fig:app1}, since it is not so important
-which specific variable that gets selected from a group of highly correlated
-predictors when the purpose is risk predictions.
+PLS-DA (except for the AUC criterion), and poorer agreement for
+ranking predictors. This reason for the latter is the high
+auto-correlation between the intensities in the mass spectra which
+leads to collinearity issues in the regression models. It is
+well-known that variable selection with the Lasso penalty does not
+preform very well when the predictors are highly correlated. The
+collinearity does, however, not affect the agreement of the risk
+predictions, as seen on the left panel of Figure \ref{fig:app1}, since
+it is not so important which specific variable that gets selected from
+a group of highly correlated predictors when the purpose is risk
+predictions.
 
@@
 In practice both concerns are of importance.
+
+\section{Simulation study --- comparison of list agreements}
+
+We present results of a simulation study where we investigated the
+small sample properties of our proposed method (sra) and compared it
+with the topK method proposed by \citet{hall:schi:2012} with respect
+to list depth agreement. In each simulation run, we generated a
+dataset with 1000 predictors and 400 observations. The predictors were
+drawn independently from a standard Gaussian distribution. The outcome
+was drawn from a Gaussian distribution with variance 1 and mean that
+depended additively on the first 15 predictors, i.e.,
+
+$$y_i = \sum_{j=1}^{15}x_{ij} + \epsilon_i,$$
+
+where $y_i$ is the $i$th response, and $x_{ij}$ is the $j$th predictor
+for the $i$th measurement, and $\epsilon_i\sim N(0, 1)$. In each
+simulation run we obtained \(L\) ranked lists of the 1000 predictors.
+This was done by first drawing \(L\) bootstrap samples (with
+replacement) and then in each bootstrap sample ranking the 1000
+predictors according to their marginal $t$ test statistics. Thus, the
+simulation results compare lists that are obtained with the same
+technique on bootstrap versions of the same dataset. We report results
+from two sets of simulations (each based on 1000 simulated datasets):
+\begin{enumerate}
+\item[] Scenario I: \(L=8\) and $q\in\{3,4,5,6,7,8,9,10\}$.
+\item[] Scenario II: \(q=5\) and $L\in\{3,5,10,50\}$.
+\end{enumerate}
+In both scenarios, we summarized the distribution of the estimated
+depth of agreement and the average number of different predictors
+found at a given depth.
+ % Asymptotic results are available for both methods as the number of
+ % lists increases, so this will investigate how much the results might
+ % change as $L$ increases. For a fixed constant threshold of $q_L=5$ we
+ % increase the number of available lists to investigate the sensitivity
+ % of the results to $L$.
+  % Throughout the simulations we
+  % assume a constant, prespecified threshold function, $q_L$, and try
+  % different values for the constant. The approach by
+  % \citet{hall:schi:2012} uses a window size parameter, $K$, to
+  % represent the maximum difference in rankings before they consider an
+  % item in two lists to be in agreement.  Because the standard
+  % deviation --- which is used in the sequential rank agreement
+  % definition of $\hat{A}_L$--- squares its differences, it tends to
+  % give more weight to larger rank differences and less weight to
+  % smaller differences compared to the mean absolute difference.  To
+  % ease comparison between the sequential rank agreement and the
+  % \citet{hall:schi:2012} approach we would like to use the same
+  % threshold for both methods, and in order to do this we use the
+  % median absolute deviance throughout the simulations in lieu of the
+  % pooled standard deviation.
+ % For all parameter combinations we ran 1000 simulations and report the
+ % distribution of the estimated depth of agreement and the average
+ % number of different predictors found at a given depth.
+The results from Scenario I are shown in the left panel of
+Figure~\ref{fig:example99}. The double histograms show the
+distribution of the estimated depth of agreement. We see that sra
+results in a substantially lower depth of agreement than the topK
+method. Also the average numbers of unique predictors (bold numbers in
+the double histograms) are markedly smaller for sra. Even larger
+differences are found if we used the Euclidean distance instead of the
+median absolute distance for the sequential rank agreement (results
+not shown). As expected the depth of agreement increases when the
+threshold for agreement/window increases.
+\begin{figure}[htbp]
+   \begin{center}
+ \includegraphics[width=.49\textwidth]{Ksim.pdf}
+ \includegraphics[width=.49\textwidth]{Lsim}
+ \end{center}
+ \caption{Left panel: Simulation study showing distribution of
+   estimated rank agreements for sra and topK for varying thresholds
+   and fixed number of lists $L=8$. The bold numbers are the average
+   number of unique predictors in the set found to be in
+   agreement. Right panel: Simulation results for varying number of
+   lists, and with fixed threshold of $q=5$.}
+ \label{fig:example99}
+ \end{figure}
+ The right panel of Figure~\ref{fig:example99} shows the results from
+ Scenario II. The number of lists has little impact but at all depths
+ sra is more conservative than topK. As a consequence we also see that
+ sra systematically selects fewer predictors in the parts of
+ agreement.
 

### 2017-04-09 latex/superranker-manuscript.tex
revised simulation section

@@
 We present results of a simulation study where we investigated the
-small sample properties of our proposed method (sra) and compared it
-with the topK method proposed by \citet{hall:schi:2012} with respect
-to list depth agreement. In each simulation run, we generated a
-dataset with 1000 predictors and 400 observations. The predictors were
-drawn independently from a standard Gaussian distribution. The outcome
-was drawn from a Gaussian distribution with variance 1 and mean that
-depended additively on the first 15 predictors, i.e.,
+small sample properties of the sequential rank agreement and compared
+it with the topK method \citep{hall:schi:2012} with respect to list
+depth agreement. For the sequential rank agreement we use the function
+\texttt{sra} of the \texttt{R} package \texttt{SuperRanker} with the
+\texttt{"mad"} option (median absolute distance). To define the depth
+of agreement we set a constant threshold function to an integer \(q\)
+and report the first crossing point, i.e., the smallest list depth
+where sra exceeds \(q\). For topK we use the function
+\texttt{j0.multi} which is implemented in the \texttt{R} package
+\texttt{TopKLists}. Specifically, we set the tuning parameter
+\texttt{v} of \texttt{j0.multi} to the value 6 and the window
+parameter \texttt{d} to \(q\) and report the output parameter
+\texttt{maxK} as the depth of agreement. Thus, to make this comparison
+we assume that the first crossing point of sra and the result of the
+topK method measure the same underlying feature.
+
+In each simulation run, we generated a dataset with 1000 predictors
+and 400 observations. The predictors were drawn independently from a
+standard Gaussian distribution. The outcome was drawn from a Gaussian
+distribution with variance 1 and mean that depended additively on the
+first 15 predictors, i.e.,
 
@@
 replacement) and then in each bootstrap sample ranking the 1000
-predictors according to their marginal $t$ test statistics. Thus, the
-simulation results compare lists that are obtained with the same
-technique on bootstrap versions of the same dataset. We report results
-from two sets of simulations (each based on 1000 simulated datasets):
+predictors according to their marginal $t$ test statistics. Thus, we
+assessed the depth of agreement among lists that are ranked with the
+same statistical method on bootstrap versions of the same dataset. We
+report simulation results from two scenarios where each is based on
+1000 simulated datasets:
 \begin{enumerate}
-\item[] Scenario I: \(L=8\) and $q\in\{3,4,5,6,7,8,9,10\}$.
-\item[] Scenario II: \(q=5\) and $L\in\{3,5,10,50\}$.
+\item[] Scenario I: Fix \(L=8\) and vary $q\in\{3,4,5,6,7,8,9,10\}$.
+\item[] Scenario II: Fix \(q=5\) and vary $L\in\{3,5,10,50\}$.
 \end{enumerate}
-In both scenarios, we summarized the distribution of the estimated
-depth of agreement and the average number of different predictors
-found at a given depth.
+In both scenarios we summarized the distribution of the estimated
+depth of agreement and the average number of unique predictors found
+in the set of predictors which is selected by the estimated depth of
+agreement.
  % Asymptotic results are available for both methods as the number of
@@
 The results from Scenario I are shown in the left panel of
-Figure~\ref{fig:example99}. The double histograms show the
-distribution of the estimated depth of agreement. We see that sra
-results in a substantially lower depth of agreement than the topK
-method. Also the average numbers of unique predictors (bold numbers in
-the double histograms) are markedly smaller for sra. Even larger
-differences are found if we used the Euclidean distance instead of the
-median absolute distance for the sequential rank agreement (results
-not shown). As expected the depth of agreement increases when the
-threshold for agreement/window increases.
+Figure~\ref{fig:example99}. The violin plots (with rectangular kernel)
+show the distributions of the estimated depths of agreement for both
+methods. We see that sra results in a substantially lower depth of
+agreement than the topK method. Also the average numbers of unique
+predictors (bold numbers inside the plots) are markedly smaller for
+sra. Even larger differences were found when we used the Euclidean
+distance instead of the median absolute distance for the sequential
+rank agreement (results not shown). As expected the depth of agreement
+increased when the threshold for agreement/window increased.
 \begin{figure}[htbp]
@@
    and fixed number of lists $L=8$. The bold numbers are the average
-   number of unique predictors in the set found to be in
-   agreement. Right panel: Simulation results for varying number of
-   lists, and with fixed threshold of $q=5$.}
+   number of unique predictors included in the set where the lists
+   agree. Right panel: Simulation results for varying number of lists
+   and with fixed threshold of $q=5$.}
  \label{fig:example99}
 \end{figure}
+
 The right panel of Figure~\ref{fig:example99} shows the results from
- Scenario II. The number of lists has little impact but at all depths
- sra is more conservative than topK. As a consequence we also see that
- sra systematically selects fewer predictors in the parts of
- agreement.
+Scenario II. The number of lists has little impact on the
+results. Generally, sra is more conservative than topK and as a
+consequence sra includes fewer predictors in the selected set where
+the lists agree.
 

### 2017-05-23 latex/superranker-manuscript.tex
Minor tweaks

@@
   that both provides an intuitive interpretation and can be applied to
-  any number of lists even if some or all are incomplete or
-  censored. The approach can identify changepoints in the agreement
-  of the lists, and the sequential changes of agreement as a function
-  of the depth of the lists can be compared graphically to a
-  permutation based reference set or evaluated using analytical asymptotic
-  results.  The usefulness of these tools are illustrated using gene
-  rankings, and using data from two Danish ovarian cancer studies
-  where we assess the within and between agreement of different
-  statistical classification methods.
+  any number of lists even if some or all are incomplete. The approach
+  can identify changepoints in the agreement of the lists, and the
+  sequential changes of agreement as a function of the depth of the
+  lists can be compared graphically to a permutation based reference
+  set or evaluated using analytical asymptotic results.  The
+  usefulness of these tools are illustrated using gene rankings, and
+  using data from two Danish ovarian cancer studies where we assess
+  the within and between agreement of different statistical
+  classification methods.
 \end{abstract}
@@
 small. If the ranked lists are identical then the value of sequential
-rank agreement will be zero for all depths $d$.  The sequential rank
-agreement is equivalent to the pooled variance of the items found in
-$S(d)$, and it can be interpreted as the average distance of the
-individual rankings of the lists from the average ranking for each of
-the items we have seen until depth $d$.  For an easier interpretation
-we suggest to plot $\sqrt{\widehat{\textrm{sra}}_L(d)}$ corresponding
-to the pooled standard deviation.
+rank agreement will be zero for all position or depths $d$ down the
+lists.  The sequential rank agreement is equivalent to the pooled
+variance of the items found in $S(d)$, and it can be interpreted as
+the average distance of the individual rankings of the lists from the
+average ranking for each of the items we have seen until depth $d$.
+For an easier interpretation we suggest to plot
+$\sqrt{\widehat{\textrm{sra}}_L(d)}$ corresponding to the pooled
+standard deviation.
 
@@
 
-The sequential rank agreement curve plotted in Figure~\ref{fig:example1}
-roughly shows how many ranks apart we on average expect the genes found in the
-top $d$ part of the lists to be for each depth $d$.  It can be seen
-that the sequential rank agreement limits are better towards the top
-of the lists (smaller values on the $y$ axis corresponds to
-\emph{better} agreement) than towards the tails of the lists. It is clear
-from the curve in Figure~\ref{fig:example1} that there is a
-substantial drop in agreement even after the first depth. Thus, if we
-were to restrict attention to a small set of predictors then we would
-focus on predictors 2124, 829, and 378 as seen in Table~\ref{tab1}.
+The sequential rank agreement curve shown in the two left plots of
+Figure~\ref{fig:example1} roughly indicates the average distance in
+ranks for the genes found among the first $d$ positions in the $4$
+lists.  It can be seen that the sequential rank agreement limits are
+better towards the top of the lists (smaller values on the $y$ axis
+corresponds to \emph{better} agreement) than towards the bottom of the
+lists. It is clear from the curve in Figure~\ref{fig:example1} that
+there is a substantial drop in agreement (higher sra) even after the
+first depth. Thus, if we were to restrict attention to a small set of
+predictors then we would focus on predictors 2124, 829, and 378 as
+seen in Table~\ref{tab1}.
 
@@
 \centering
-\caption{Top 10 list of ranked results from the Golub data. Numbers indicate the predictor/gene for the given ranking and method} 
+\caption{List of ranked results from the Golub data. Numbers indicate the predictor/gene for the given ranking and method. Only the top 10 ranks are shown in the table.} 
 \label{tab1}
@@
    analysis methods applied to the 3051 genes in the Golub data (black
-   line). Right
-   panels: Corresponding sequential rank agreement for the same data
-   but where only the top 20 ranked items are available and the rest
-   of the items have been censored in each list. The blue and
+   line). Right panels: Corresponding sequential rank agreement for
+   the same data but where only the top 20 ranked items are available
+   and the rank of the remaining items are not available. The blue and
    red areas correspond to the independent and randomized reference
@@
 
-\subsection{Analysis of partial/censored lists}
+\subsection{Analysis of incomplete lists}
 Incomplete or partial lists are a common occurrence that arise, for
@@
 whose coefficient has been shrunk to zero and thus we end up with a
-partial ordering. Incomplete lists also occur if the analyst decides
-to censor all lists at depth $d$ (\ie, top $d$ lists) such that the
-measurement of agreement is restricted to the first $d$ elements of
-each list, or when the analyst only wants to consider the ranking of
-items that have been found to be statistically significant.
-
-Sequential rank agreement can be generalized to partial lists in the
-following way. Let $\Lambda_l\subset X$ be the subset of $d_l$ items
-that have been ranked highest in list $l$.  The case where all lists are
-censored at the same depth $d$ corresponds to $d_1=\cdots=d_L=d$. For
-incomplete/censored lists the rank function becomes
+partial ordering.  Incomplete lists also occur if the analyst
+restricts attention to the ranks of items that have been found to be
+statistically significant.
+
+Sequential rank agreement can be generalized to incomplete lists in
+the following way. Let $\Lambda_l\subset X$ be the subset of $d_l$
+items that have been ranked highest in list $l$.  The case where all
+lists are incomplete at the same depth $d$ corresponds to
+$d_1=\cdots=d_L=d$. For incomplete lists the rank function becomes
 \begin{equation}
@@
 The agreement, $A(X_p)$, cannot be computed directly for all
-predictors in the presence of censored lists because the exact rank
+predictors in the presence of incomplete lists because the exact rank
 for some items will be unknown. Also, recall that the rankings within
@@
 once in each list. Thus, we cannot simply assign the same number
-(\eg, the mean of the unassigned ranks) to the censored items since
+(\eg, the mean of the unassigned ranks) to the missing items since
 that would result in \emph{less} variation of the ranks and hence less
 variation of the agreement, and it would artificially introduce a
-(downward) bias of agreement for items that are censored in multiple
+(downward) bias of agreement for items that are missing in multiple
 lists.
@@
 \begin{algorithm}
-\caption{Sequential rank agreement algorithm for censored lists}
+\caption{Sequential rank agreement algorithm for incomplete lists}
 \label{sra-algorithm}
 \begin{algorithmic}[1]
-  \Procedure{Censored rank agreement}{} \State Let $B$ be the number
-  of permutations to use \For{each $b \in B$} \For{each censored list
+  \Procedure{Incomplete list rank agreement}{} \State Let $B$ be the number
+  of permutations to use \For{each $b \in B$} \For{each list
     $l \in L$}
@@
 interesting items are found in the top of the lists, and 2) that the
-censored rankings provide so little information that it is reasonable
+ranks that are missing from the lists provide so little information that it is reasonable
 to assume that they can be represented by a random order. The first
@@
 interesting part of the sequential rank agreement curves is restricted
-to the depths where the number of censored items is
+to the depths where the number of items without ranks available is
 low. Generally, without additional prior knowledge about the
-distribution of the censored items it seems reasonable to restrict the
-attention of the sequential rank agreement to depths smaller than
-$\max(d_1, \ldots, d_L)$ where at least one list provides an
-uncensored item.
+underlying distribution of the ranks it seems reasonable to restrict
+attention of the sequential rank agreement to lists no longer than the
+longest of the $L$ lists examined.
 
@@
 ranking (in which case the sequential rank agreement will also be
-constant but at a low level). For partially ranked lists we also
+constant but at a low level). For incomplete ranked lists we also
 expect a changepoint around the depth where the lists are
-censored. This is an artefact stemming from the fact that we assume
+become incomplete. This is an artefact stemming from the fact that we assume
 that the remainder of the lists can be replaced by a simple
 permutation of the missing items.
+
+The two right-most plots of Figure~\ref{fig:example1} show the impact
+of restricting the Golub data such that only top-20 lists are
+available instead of full lists of length 1000 (20 was chosen to
+resemble the list lengths that might be published in a
+manuscript). The sequential rank agreement increases much quicker
+because the incomplete lists introduce more noise in the estimation of
+the agreement, but it is still possible to see that the top of the
+list has a sequantial rank agreement that is not substantially
+different from the full lists.
 
@@
 which not only assumes that there is no information in the data on
-which the rankings are based but also that the methods are completely
-independent.
+which the rankings are based but also that the methods used to provide
+the rankings are completely independent.
 
 Alternatively, we can remove the restriction on the independence among
-the methods by only requiring that there is no information in the
-ranking but that the rankings are all based on applying the
-method/approaches to the same data
+the methods used to generate the individuals rankings in the $L$ lists
+by only requiring that there is no information contained in the ranking but that
+the rankings are all based on applying the method/approaches to the
+same data
 \begin{eqnarray*}
@@
 \end{eqnarray*}
+
+This alternative null hypothesis addresses the fact that some ranking
+methods are more likely to provide similar rankings of the same data
+because the ranking methods focus on the same features of the data
+rather than because of any information contained in the data.
 
@@
 contains $L$ lists of random permutations of the items in $X$. For the
-censored/partially observed case we first permute the items
-$X_1,\dots,X_P$ and then censor list $l$ at $d_l$ (\ie, each list so
-it has the same number of observed rankings as was observed for list
-$l$ in the original dataset). The sequential rank agreement curve from
-the the original lists can then be compared to, say, the pointwise
-95\% quantiles of the observed rank agreements obtained under $H_0$.
+incomplete case we first permute the items $X_1,\dots,X_P$ and then assign missing ranks
+for list $l$ from  $d_l$ to $P$ (\ie, each list so it has the same number of
+observed rankings as was observed for list $l$ in the original
+dataset). The sequential rank agreement curve from the original
+lists can then be compared to, say, the pointwise 95\% quantiles of
+the observed rank agreements obtained under $H_0$.
 
@@
 is better than what would be expected by chance for data that contain
-no information since it lies below the reference areas. The censored
+no information since it lies below the reference areas. The incomplete
 data also suggests that there may be at most 1 or 2 ranked items
@@
 
-Corollary \ref{thecor} indicates that we can use the a threshold
+Corollary \ref{thecor} indicates that we can use the threshold
 function $\widehat{q}_L$ estimated under the null hypothesis as
@@
 function serves the same role as the limits of agreement in method
-comparison studies, except that the threshold funciton is not constant
+comparison studies, except that the threshold function is not constant
 but can accommodate the changing nature of the number of items used
-for the computation of the equential rank agreement for a given depth.
+for the computation of the sequential rank agreement for a given depth.
 
@@
 criteria to our data to investigate their effect on the agreements.
-Note also that the Lasso models produce censored lists depending on
+Note also that the Lasso models produce incomplete lists depending on
 the value of the penalty parameter.
@@
 reason behind this behavior is rather subtle. Looking at the
-distribution of the absolute value of the regression coefficients we see that a large
-proportion of them are numerically very close to zero and have almost
-equal absolute value. This is a general feature of the Ridge models in
-this dataset and seen for all the 1000 trained models.
+distribution of the absolute value of the regression coefficients we
+see that a large proportion of them are numerically very close to zero
+and have almost equal absolute value. This is a general feature of the
+Ridge models in this dataset and seen for all the 1000 trained models.
 This implies that when predictors are ranked according to the
@@
 uncertain and more close to a random permutation. This problem can be
-alleviated by truncating all predictors with absolute coefficient values
-below a given threshold thereby introducing an artificial censoring of the lists.
-For the Ridge models tuned with the deviance criterion, Figure \ref{fig:app2} (left) shows the
-the sequential rank agreement where for each of the 1000 trained
-models the predictors were artificially censored when their absolute
+alleviated by truncating all predictors with absolute coefficient
+values below a given threshold thereby introducing an artificial
+incompletion of the lists.  For the Ridge models tuned with the
+deviance criterion, Figure \ref{fig:app2} (left) shows the the
+sequential rank agreement where for each of the 1000 trained models
+the predictors were artificially censored when their absolute
 coefficient value was lower than the $0.1\%$ quantile of the 5000
-absolute coefficient values. The curve was calculated
-using Algorithm \ref{sra-algorithm} with $B=1000$ and $P=5000$. The corresponding
-curve from Figure \ref{fig:app1} (right panel) is shown for comparison.
-Even though the number of censored predictors is very small compared to the total
-number of predictors, the effect on the sequential rank agreement is
-substantial and with the artificial censoring the shape of the curves
-is as expected, starting low and then increasing.
-
-Looking at the agreement curves for the Lasso models in Figure \ref{fig:app1}
-(right) we clearly see the effect of the sparsity inducing penalization giving
-rise to censored lists. These curves were similarly calculated using \ref{sra-algorithm}
-and $1000$ random permutations. Under the deviance optimization criterion the
-median number of non-zero coefficients was 33 (range 16 to 50) and for the
-class accuracy criterion 14 (range 4 to 56). These values correspond to
-the list depths where the agreement curves become constant as a result of
-the subsequent censoring.
+absolute coefficient values. The curve was calculated using Algorithm
+\ref{sra-algorithm} with $B=1000$ and $P=5000$. The corresponding
+curve from Figure \ref{fig:app1} (right panel) is shown for
+comparison.  Even though the number of predictors with missing ranks is very
+small compared to the total number of predictors, the effect on the
+sequential rank agreement is substantial and with the artificial
+censoring the shape of the curves is as expected, starting low and
+then increasing.
+
+Looking at the agreement curves for the Lasso models in Figure
+\ref{fig:app1} (right) we clearly see the effect of the sparsity
+inducing penalization giving rise to incomplete lists. These curves
+were similarly calculated using \ref{sra-algorithm} and $1000$ random
+permutations. Under the deviance optimization criterion the median
+number of non-zero coefficients was 33 (range 16 to 50) and for the
+class accuracy criterion 14 (range 4 to 56). These values correspond
+to the list depths where the agreement curves become constant as a
+result of the subsequent censoring.
 
@@
 \caption{Left panel: Sequential rank agreement for Ridge regression
-  obtained by artificially censoring predictors when their absolute
+  obtained by artificially censoring predictor ranks when their absolute
   coefficient values are lower than the 0.1\% quantile. Right panel:
@@
 well-known that variable selection with the Lasso penalty does not
-preform very well when the predictors are highly correlated. The
+perform very well when the predictors are highly correlated. The
 collinearity does, however, not affect the agreement of the risk
@@
 
-We present results of a simulation study where we investigated the
+We present results from a simulation study where we investigated the
 small sample properties of the sequential rank agreement and compared
-it with the topK method \citep{hall:schi:2012} with respect to list
+it to the topK method \citep{hall:schi:2012} with respect to list
 depth agreement. For the sequential rank agreement we use the function
@@
 \texttt{"mad"} option (median absolute distance). To define the depth
-of agreement we set a constant threshold function to an integer \(q\)
+of agreement we set a constant threshold function to an integer $q$
 and report the first crossing point, i.e., the smallest list depth
-where sra exceeds \(q\). For topK we use the function
+where sra exceeds $q$. For topK we use the function
 \texttt{j0.multi} which is implemented in the \texttt{R} package
@@
 \texttt{v} of \texttt{j0.multi} to the value 6 and the window
-parameter \texttt{d} to \(q\) and report the output parameter
+parameter \texttt{d} to $q$ and report the output parameter
 \texttt{maxK} as the depth of agreement. Thus, to make this comparison
@@
 for the $i$th measurement, and $\epsilon_i\sim N(0, 1)$. In each
-simulation run we obtained \(L\) ranked lists of the 1000 predictors.
-This was done by first drawing \(L\) bootstrap samples (with
+simulation run we obtained $L$ ranked lists of the 1000 predictors.
+This was done by first drawing $L$ bootstrap samples (with
 replacement) and then in each bootstrap sample ranking the 1000
@@
 and the situation where the only available data is the actual ranked
-lists. In addition, censored ranked lists where only the ranks of the
+lists. In addition, incomplete ranked lists where only the ranks of the
 top $k$ ranked items are known can be accommodated as well. The
@@
 Finally, we have --- whenever possible --- used all available ranks
-from the lists.  We could choose to censor the rankings of all the
-items that were deemed not to be significant if information on both
-rankings and the corresponding evidence for significance was
-available. That would ensure that there would be put less emphasis on
-the agreement of the non-significant items and it would be easier to
-identify a change in agreement among the items that were deemed to
-be relevant. In our application section we have successfully introduced
-such an artificial censoring for the predictor rankings obtained with
-ridge regression.
+from the lists.  We could choose to restrict attention to the rank of
+items which show evidence for significance in their models. That would
+ensure that there would be put less emphasis on the agreement of the
+non-significant items and it would be easier to identify a change in
+agreement among the items that were deemed to be relevant. In our
+application section we have successfully introduced such an artificial
+censoring for the predictor rankings obtained with ridge regression.
 

### 2018-01-24 latex/biosts3-supp.tex
Updatedd version

--- 2017-04-07 latex/superranker-manuscript.tex
@@
-\documentclass[12pt,a4paper]{article}
+\documentclass[oupdraft]{bio}
+%\usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, citecolor=citecolor]{hyperref}
+% \documentclass[12pt,a4paper]{article}
 \pdfoutput=1
 \usepackage[utf8]{inputenc}
-\usepackage[]{graphicx}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage[]{color}
-\usepackage{subcaption}
-\usepackage{natbib}
-\usepackage{xspace}
-\usepackage{calc}
-%\usepackage[noae]{Sweave}
-\usepackage{authblk}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{enumitem}
+%\usepackage[]{graphicx}
+%\usepackage{amsmath}
+%\usepackage{amssymb}
+%\usepackage[]{color}
+%\usepackage{subcaption}
+% \usepackage{natbib}
+%\usepackage{xspace}
+%\usepackage{calc}
+% \usepackage[noae]{Sweave}
+%\usepackage{algorithm}
+%\usepackage{algpseudocode}
+% \usepackage{enumitem}
+
+% \usepackage{authblk}
+
+% \usepackage[nolists,nomarkers]{endfloat}
+
 %\usepackage{geometry}
@@
 %
-\newcommand{\ie}{\emph{i.e.}\@\xspace}
-\newcommand{\eg}{\emph{e.g.}\@\xspace}
+%\newcommand{\ie}{\emph{i.e.}\@\xspace}
+%\newcommand{\eg}{\emph{e.g.}\@\xspace}
 
@@
 
-\newcommand{\mcomment}[1]{\textcolor{red}{\textbf{#1}}}
+% \newcommand{\mcomment}[1]{\textcolor{red}{\textbf{#1}}}
 
@@
 
-\usepackage[amsthm,thmmarks]{ntheorem}
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}
-\newtheorem{definition}{Definition}
-\newtheorem{proposition}{Proposition}
-\newtheorem{corollary}{Corollary}
-\newtheorem{lemma}{Lemma}
-\newtheorem{assumption}{Assumption}
-\newtheorem{example}{Example}
-\renewcommand\proofSymbol{\ensuremath{\blacksquare}}
+%\usepackage[amsthm,thmmarks]{ntheorem}
+% \theoremstyle{plain}
+% \newtheorem{theorem}{Theorem}
+%\newtheorem{definition}{Definition}
+%\newtheorem{proposition}{Proposition}
+\newtheorem{corollary}{Corollary}[section]
+%\newtheorem{lemma}{Lemma}
+%\newtheorem{assumption}{Assumption}
+%\newtheorem{example}{Example}
+%\renewcommand\proofSymbol{\ensuremath{\blacksquare}}
 
@@
 
-\title{Sequential rank agreement methods for comparison of ranked lists}
-\author{Claus Thorn Ekstrøm}
-\author{Thomas Alexander Gerds}
-\author{Andreas Kryger Jensen}
+%\title{Sequential rank agreement methods for comparison of ranked lists}
+%\author{Claus Thorn Ekstrøm}
+%\author{Thomas Alexander Gerds}
+%\author{Andreas Kryger Jensen}
 %\author{Kasper Brink-Jensen}
-\affil{Biostatistics, University of Copenhagen}
+%\affil{Biostatistics, University of Copenhagen}
+%\renewcommand{\topfraction}{.85}
+%\renewcommand{\bottomfraction}{.7}
+%\renewcommand{\textfraction}{.15}
+%\renewcommand{\floatpagefraction}{.66}
+%\renewcommand{\dbltopfraction}{.66}
+%\renewcommand{\dblfloatpagefraction}{.66}
+%\setcounter{topnumber}{9}
+%\setcounter{bottomnumber}{9}
+%\setcounter{totalnumber}{20}
+%\setcounter{dbltopnumber}{9}
+
 \renewcommand{\topfraction}{.85}
@@
 
+\title{Supplementary material for Sequential rank agreement methods for comparison of ranked lists}
+
+\author{CLAUS THORN EKSTRØM$^\ast$, THOMAS ALEXANDER GERDS, ANDREAS KRYGER JENSEN\\[4pt]
+%
+\textit{%
+Biostatistics, Department of Public Health,
+University of Copenhagen,
+Øster Farimagsgade 5 B, P.O.B. 2099,
+DK-1014 Copenhagen K, Denmark}
+\\[2pt]
+% E-mail address for correspondence
+{ekstrom@sund.ku.dk}}
+
 % Use square brackets for numbering affiliations
+
+% Running headers of paper:
+\markboth%
+% First field is the short list of authors
+{C. T. Ekstrøm and others}
+% Second field is the short title of the paper
+{Supplementary material for ``Sequential rank agreement for comparison of ranked lists''}
 
@@
 
-\begin{abstract}
-  The comparison of alternative rankings of a set of items is a
-  general and common task in applied statistics. Predictor variables
-  are ranked according to magnitude of association with an outcome,
-  prediction models rank subjects according to the personalized risk
-  of an event, and genetic studies rank genes according to their
-  difference in gene expression levels. This article constructs
-  measures of agreement among two or more ordered lists. We use the
-  standard deviation of the ranks to define a measure of agreement
-  that both provides an intuitive interpretation and can be applied to
-  any number of lists even if some or all are incomplete or
-  censored. The approach can identify changepoints in the agreement
-  of the lists, and the sequential changes of agreement as a function
-  of the depth of the lists can be compared graphically to a
-  permutation based reference set or evaluated using analytical asymptotic
-  results.  The usefulness of these tools are illustrated using gene
-  rankings, and using data from two Danish ovarian cancer studies
-  where we assess the within and between agreement of different
-  statistical classification methods.
-\end{abstract}
-
-Key words: sequential rank agreement, partial ranking, order
-statistic, gene rankings, methods comparison, variable selection,
-permutation
-
-\section{Introduction}
-
-Ranking of items or results is common in scientific research and
-ranked lists occur naturally as the result of many statistical
-applications. Regression methods rank predictor variables according to
-magnitude of their association with an outcome, prediction models rank
-subjects according to their risk of an event, and genetic studies rank
-genes according to their difference in gene expression levels across
-samples.
-
-When several rankings of the same items are available, a common
-research question is to what extent they agree. In particular, is it
-possible to identify an optimal rank until which the lists agree on
-the items?  A typical situation arises in high-dimensional genomics
-studies when several analysis methods are applied to rank a list of
-genes according to their association with a phenotype, treatment
-effect or other outcome. A measure of agreement of gene rankings
-obtained by different methods could often help to identify which genes
-are worth to pursue in further experiments.
-
-Several approaches exist to measure the ``distance'' between two
-ranked lists. Of these, Kendall's $\tau$ \citep{kendall1948} and
-Spearman’s footrule $\rho$ \citep{Spearman1910} are among the most
-well-known, but these measures do not distinguish between agreement in
-the top versus towards the bottom of the lists and they both provide a
-single measure of the overall distance. \citet{Shieh1998} proposed a
-weighted version of Kendall's $\tau$ where each pair of rankings can
-be assigned different weights, and \citet{Yilmaz2008} proposed the
-$\tau_{ap}$ which places higher emphasis on the top of the lists.
-Spearman’s footrule uses the ranks of the variables for calculation of
-the distance and the use of ranks are also employed in the $M$ measure
-of \citet{Bar-Ilan2006} where the reciprocal rank differences are used
-to calculate the similarity measure.
-
-Other recent approaches consider the intersection of lists as the basis
-for a similarity measure. However, simple intersection also places
-equal weights on all depths of the list and therefore \citet{Fagin2003} and
-\citet{Webber2010} proposed weighted intersections which put more
-emphasis on the top of the lists. 
-% One proposal is denoted the average overlap (AO) and used for
-% comparison later on. 
-Specifically, \citet{Webber2010} define their rank-biased overlap
-(RBO) by weighting with a converging series to ensure that the top is
-weighted higher than the potentially non-informative bottom of the
-lists.  It is possible to use the existing methods to calculate
-agreement of lists until a given depth, i.e., limited to the $d$ items
-of each list. However, the interpretation may not be straightforward,
-especially in the case of more than two lists, and they may not
-accommodate partial rankings.
-
-Very recently, \citet{hall:schi:2012} proposed a method for comparing
-pairwise rankings and derived the asymptotic distribution of the
-endpoint where the two ranked lists no longer are in agreement. Their
-approach was based on anchoring one of the two lists and subsequently
-generating a sequence of 0s and 1s depending on whether the ranks in
-the second list was close to the rank from the anchored
-list. \citet{samp:verd:2013} followed up on this idea for pairwise
-comparison of lists but used penalties based on a truncated geometric
-probability instead of a 0-1 process and they evaluated the
-distribution of the endpoint of agreement by computational approaches.
-The asymptotic distribution in the \citet{hall:schi:2012} paper is
-based on letting the number of \emph{lists} increase to infinity which
-is a situation that is only relevant in special cases, whereas the
-simulation-based null distribution approach of \citet{samp:verd:2013}
-is guaranteed to work also for small samples.
-
-In this article we introduce sequential rank agreement for measuring
-agreement among (partially) ranked lists.  The general idea is to
-define agreement based on the sequence of ranks from the first $d$
-items in each list. As agreement metric we adapt the limits of
-agreement known from agreement between quantitative variables
-\citep{alt:bland:1983,Carstensen2010} but any measure for agreement
-could essentially be used. Our proposed approach allows us to
-compare multiple lists simultaneously, it provides a dynamic
-measure of agreement as a function of the depth in the lists, it places
-higher weight on items at the top of the list, it accommodates partially
-observed lists of varying lengths, and has a natural interpretation
-that directly relates to the ranks.  Graphical illustration of
-sequential rank agreement potentially allows us to infer a
-changepoint, i.e., a list depth where a substantial change in the agreement of
-the lists occur but we also provide asymptotical and randomization-based graphical
-tools to compare the observed rank agreement to the expected agreement
-found in non-informative data.  In this sense our approach is a
-combination and generalization of some of the ideas of
-\citet{Carterette2009} and \citet{Boulesteix2009}. The former compares
-two rankings based on the distance between them as measured by a
-multivariate Gaussian distribution and the latter presents an overview
-of approaches for aggregation of ranked lists including bootstrap and
-leave-one-out jackknife approaches. We show the asymptotic
-distribution of the endpoint of agreement and discuss how to infer the
-distribution in small-sample situations using computational
-methods. This enables us to make inferences about the endpoint of
-agreement even for situations where there is no actual changepoint
-present but just a gradual decline in agreement.
-
-The manuscript is organized as follows: In the next section we define
-sequential rank agreement for multiple ranked lists and discuss how to
-handle incomplete lists. In section 3 we present and discuss
-approaches to evaluate the results obtained from sequential rank
-agreement. Finally we apply the sequential rank agreement to two
-Danish ovarian cancer studies and compare our method to the method of
-\citet{hall:schi:2012} in a small sample simulation study before we
-discuss the findings along with possible extensions. The approaches
-presented in this manuscript are available in the R package
-\texttt{SuperRanker} which can be found on CRAN.
-% chunk 1
-
-\section{Methods}
-
-Consider a set of $P$ different items $X=\{X_1,\dots,X_P\}$ and a
-ranking function $R: \{X_1,\dots,X_P\}\to \{1,\dots,P\}$, such that
-$R(X_p)$ is the rank of item $X_p$. The inverse mapping $R^{-1}$ gives
-the item $R^{-1}(r)$ that was assigned to rank $r\in\{1,\dots,P\}$. An
-ordered list is the realization of a ranking function $R$ applied to
-the set of items $X$.  Panels (a) and (b) of Table~\ref{tab:example}
-show a schematic example of these mappings. Thus if
-$R_l^{-1}(1)=X_{34}$ then item $X_{34}$ is ranked first in list $l$
-and similarly $R_l(X_{34})=1$.
-
-In all what follows we consider a fixed set of items and consider the
-ranking function to be a random variable. Thus, let
-$R_1(X),\dots,R_L(X)$, $L\geq2$, be a sample of $L$ independent
-identically distributed draws from an unknown probability distribution
-function $Q$. One aim is then to test how much $Q$ resembles the
-uniform distribution which assigns probability $1/P!$ to each of the
-$P!$ different possible rankings.
-
-\begin{table}[tb]
-  \caption{Example set of ranked lists. (a) shows the ranked lists of
-    items for each of three lists, (b) presents the ranks obtained by
-    each item in each of the three lists and (c) shows the cumulative
-    set of items up to a given depth in the three lists.}
-\begin{center}
-\footnotesize{
-  \begin{subtable}{4.5cm}%
-    \caption{}
-      \begin{tabular}{cccc}
-        \hline\hline % & \multicolumn{3}{c}{List} \\
-        Rank & $R^{-1}_1$ & $R^{-1}_2$ & $R^{-1}_3$ \\ \hline
-        1 & A & A & B \\
-        2 & B & C & A \\
-        3 & C & D & E \\
-        4 & D & B & C \\
-        5 & E & E & D \\ \hline
-    \end{tabular}
-  \end{subtable}
-% \hfill
-\hspace{1em}
-%%%
-  \begin{subtable}{4cm}%
-    \caption{}
-    \begin{tabular}{cccc} \hline\hline
-  % & \multicolumn{3}{c}{List} \\
-    Item & $R_1$ & $R_2$ & $R_3$ \\ \hline
-    A & 1 & 1 & 2 \\
-    B & 2 & 4 & 1 \\
-    C & 3 & 2 & 4 \\
-    D & 4 & 3 & 5 \\
-    E & 5 & 5 & 3 \\ \hline
-  \end{tabular}
-\end{subtable}
-% \hfill
-\hspace{1em}
-%%%
-\begin{subtable}{4cm}%
-  \caption{}
-\begin{tabular}{cc} \hline\hline
-% Depth   &     \\
-Depth  &  $S_d$ \\ \hline
-1 & $\{$A, B$\}$\\
-2 & $\{$A, B, C$\}$ \\
-3 & $\{$A, B, C, D, E$\}$ \\
-4 & $\{$A, B, C, D, E$\}$ \\
-5 & $\{$A, B, C, D, E$\}$ \\ \hline
-\end{tabular}
-\end{subtable}
-}
-\end{center}
-\label{tab:example}
-\end{table}
-
-The agreement among the lists regarding the rank given to an item $X_p$
-can be measured by the variance across the lists
-\begin{align}
-  A(X_p) &= \E_Q\left[\left(R(X_p)-\E_Q R(X_p)\right)^2\right]\\
-   &= \sum_{r\in\Pi} \left(r(X_p)-\E_Q R(X_p)\right)^2 Q(r),\nn
-\end{align}
-where $\Pi$ is the set of all permutations of $X$, $Q$ is a
-probability mass function on $\Pi$, and
-$\E_Q R(X_p)=\sum_{r\in\Pi} r(X_p)Q(r)$. The empirical counterpart is
-\begin{align}
-\widehat{A}_L(X_p) = \frac{1}{L-1}\sum_{i=1}^L (R_i(X_p) - \overline{R}_L(X_p))^2, \quad \overline{R}_L(X_p) = \frac{1}{L}\sum_{i=1}^L R_i(X_p).\label{eq:empDistance}
-\end{align}
-For each item, the function $\widehat{A}_L$ has an interpretation as the expected
-Euclidean distance of the individual rankings from the expected
-ranking over the $L$ lists, and it corresponds to the same measure
-that is used in method comparison studies to compute the limits of
-agreement \citep{alt:bland:1983}.
-
-For an integer $1\le d\le P$ we define the expected set of unique
-items found by merging the first $d$ elements across the possible
-lists:
-\begin{align}
-S(d) &= \left\{X_p; \left(\sum_{r\in\Pi} 1\left(r(X_p)\le d\right) Q(r)\right) >0\right\}\label{eq:sumset}
-\end{align}
-where $1(\cdot)$ denotes the indicator function.
-The empirical counterpart is the set of unique items ranked less than or
-equal to $d$ in any of the $L$ lists:
-\begin{equation}
-\widehat{S}_{L}(d) = \left\{X_p; \left(\frac{1}{L}\sum_{l=1}^L 1\left(R_l(X_p)\le d\right)\right)>0\right\},
-\end{equation}
-which is exemplified in Panel (c) of Table \ref{tab:example}.
-
-We define the \emph{sequential rank agreement} as the weighted
-expected agreement of the items found in the set $S(d)$:
-\begin{align}
-  \textrm{sra}(d) = \frac{1}{|S(d)|}\sum_{p \in S(d)}A(X_p)\label{def:sra}
-\end{align}
-where $|S(d)|$ is the cardinality of the set $S(d)$.
-The empirical counterpart is equivalently given by
-\begin{equation}
-  \widehat{\textrm{sra}}_L(d) = \frac{\sum_{\{p \in \widehat{S}_{L}(d)\}}(L-1)\widehat{A}_L(X_p)}{(L-1)|\widehat {S}_{L}(d)|}\label{def:sraEst}
-\end{equation}
-
-Values of \textrm{sra} close to zero suggests that the lists agree on
-the rankings while larger values suggest disagreement.  Just like in
-method comparison studies we prefer smaller limits of agreement
-because that means that the differences between the rankings is
-small. If the ranked lists are identical then the value of sequential
-rank agreement will be zero for all depths $d$.  The sequential rank
-agreement is equivalent to the pooled variance of the items found in
-$S(d)$, and it can be interpreted as the average distance of the
-individual rankings of the lists from the average ranking for each of
-the items we have seen until depth $d$.  For an easier interpretation
-we suggest to plot $\sqrt{\widehat{\textrm{sra}}_L(d)}$ corresponding
-to the pooled standard deviation.
-
-\subsection{Agreement among fully observed lists}
-\label{sec:amfol}
-
-The simplest case occurs when all $L$ lists are fully observed, \ie, when
-we observe the rank of all $P$ items for all $L$ lists. Fully
-observed ranked lists are common and arise, for example, when
-different statistical analysis methods are applied to a single dataset
-to produce lists of predictors ranked according to their importance,
-or if the same analysis method is applied to data from different
-populations.
-
-With fully observed lists we can plot the sequential rank agreement
-\eqref{def:sra} as a function of depth $d$. An example is seen in the top
-panels of Figure~\ref{fig:example1} where four analysis methods were
-used to rank 3051 gene expression values measured on 38 tumor mRNA
-samples in order to classify between acute lymphoblastic leukemia
-(ALL) and acute myeloid leukemia (AML)
-\citep{Golub1999}. Preprocessing of the gene expression data was done
-as described in \citet{Dudoit2002} and the four different analysis
-approaches were: marginal unequal variances two-sample $t$ tests,
-marginal logistic regression analyses, elastic net logistic regression
-\citep{friedman2010regularization}, and marginal maximum information
-content correlations (MIC) \citep{Reshef2011}. For the first two
-methods, the genes were ranked according to minimum $p$ value, for
-logistic regression the genes were ordered by size of the
-corresponding coefficients (after standardization), and MIC was
-ordered by absolute correlation which resulted in the top rankings
-seen in Table~\ref{tab1}.
-
-The sequential rank agreement curve plotted in Figure~\ref{fig:example1}
-roughly shows how many ranks apart we on average expect the genes found in the
-top $d$ part of the lists to be for each depth $d$.  It can be seen
-that the sequential rank agreement limits are better towards the top
-of the lists (smaller values on the $y$ axis corresponds to
-\emph{better} agreement) than towards the tails of the lists. It is clear
-from the curve in Figure~\ref{fig:example1} that there is a
-substantial drop in agreement even after the first depth. Thus, if we
-were to restrict attention to a small set of predictors then we would
-focus on predictors 2124, 829, and 378 as seen in Table~\ref{tab1}.
-
-% latex table generated in R 3.2.3 by xtable 1.8-2 package
-% Fri Apr  8 10:22:58 2016
-\begin{table}[tb]
-\centering
-\caption{Top 10 list of ranked results from the Golub data. Numbers indicate the predictor/gene for the given ranking and method} 
-\label{tab1}
-\begin{tabular}{rrrrr}
-  \hline
-Ranking & Welsh's $t$ & LogReg & ElasticNet & MIC \\ 
-  \hline
-1 & 2124 & 2124 & 829 & 378 \\ 
-  2 & 896 & 896 & 2124 & 829 \\ 
-  3 & 2600 & 829 & 2198 & 896 \\ 
-  4 & 766 & 394 & 1665 & 1037 \\ 
-  5 & 829 & 766 & 1920 & 2124 \\ 
-  6 & 2851 & 2670 & 1042 & 808 \\ 
-  7 & 703 & 2939 & 808 & 108 \\ 
-  8 & 2386 & 2386 & 849 & 515 \\ 
-  9 & 2645 & 1834 & 937 & 2670 \\ 
-  10 & 2002 & 378 & 1995 & 2600 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-% chunks 2,3,4,5,6,7
-
-\begin{figure}[htbp]
-   \begin{center}
- \includegraphics[width=.49\textwidth]{paper-fig1a}
- \includegraphics[width=.49\textwidth]{paper-fig2a}
- \includegraphics[width=.49\textwidth]{paper-fig1b}
- \includegraphics[width=.49\textwidth]{paper-fig2b}
- \end{center}
- \caption{Left panels: Sequential rank agreement for 4 different
-   analysis methods applied to the 3051 genes in the Golub data (black
-   line). Right
-   panels: Corresponding sequential rank agreement for the same data
-   but where only the top 20 ranked items are available and the rest
-   of the items have been censored in each list. The blue and
-   red areas correspond to the independent and randomized reference
-   hypothesis areas, respectively. The bottom plots are identical to
-   the corresponding top plots but have been zoomed in on the first
-   part of the $x$ axis.}
- \label{fig:example1}
- \end{figure}
-
-% If the lists agree on all items then the sequential rank agreement is
-% zero for all depths. 
- Generally, if there are changes in the level of
- rank agreement then this suggests that there are sets of items that
- are ranked similarly in all lists and other sets of items that have
- been assigned vastly different ranks in the lists. For example, in gene
- association studies we generally expect the agreement among the lists
- to be better towards the top of the lists and worse towards the end
- of the lists. In this case the sequential rank agreement curve will
- start at a low level and then increase until it levels off as seen in
- Figure~\ref{fig:example1}.
-
-\subsection{Analysis of partial/censored lists}
-Incomplete or partial lists are a common occurrence that arise, for
-example, in case of missing data (items), when comparing top $d$ list
-results from publications, or when some methods only rank a subset of
-the items. For example, penalized regression based on the Lasso
-provides a sparse set of predictors that have non-zero
-coefficients. There is no obvious ordering of the set of predictors
-whose coefficient has been shrunk to zero and thus we end up with a
-partial ordering. Incomplete lists also occur if the analyst decides
-to censor all lists at depth $d$ (\ie, top $d$ lists) such that the
-measurement of agreement is restricted to the first $d$ elements of
-each list, or when the analyst only wants to consider the ranking of
-items that have been found to be statistically significant.
-
-Sequential rank agreement can be generalized to partial lists in the
-following way. Let $\Lambda_l\subset X$ be the subset of $d_l$ items
-that have been ranked highest in list $l$.  The case where all lists are
-censored at the same depth $d$ corresponds to $d_1=\cdots=d_L=d$. For
-incomplete/censored lists the rank function becomes
-\begin{equation}
-	\widetilde R_l(X_p) = \begin{cases}
-	     \{R_l(X_p)\} & \text{for } X_p\in \Lambda_l\\
-	     \{d_l+1,\dots,P\} & \text{for} X_p \not\in \Lambda_l
-        \end{cases}
-\end{equation}
-where we only know that the rank for the unobserved items in list $l$
-must be larger than the largest rank observed in that list.
-
-The agreement, $A(X_p)$, cannot be computed directly for all
-predictors in the presence of censored lists because the exact rank
-for some items will be unknown. Also, recall that the rankings within
-a single list are not independent since each rank must appear exactly
-once in each list. Thus, we cannot simply assign the same number
-(\eg, the mean of the unassigned ranks) to the censored items since
-that would result in \emph{less} variation of the ranks and hence less
-variation of the agreement, and it would artificially introduce a
-(downward) bias of agreement for items that are censored in multiple
-lists.
-
-Instead we randomize the ranks $\{d_{l}+1,\dots,P\}$ to the items that
-do not occur in list $\Lambda_l$. One realization of the $L$ rankings
-of the set $X$ is obtained by randomizing the missing items of each
-list. By randomizing a large number of times we can compute
-\eqref{def:sra} for each realization, and then compute the sequential
-rank agreement as the pointwise (for each depth) average of the rank
-agreements. The algorithm is described in detail in
-Algorithm~\ref{sra-algorithm}.
-
-\begin{algorithm}
-\caption{Sequential rank agreement algorithm for censored lists}
-\label{sra-algorithm}
-\begin{algorithmic}[1]
-  \Procedure{Censored rank agreement}{} \State Let $B$ be the number
-  of permutations to use \For{each $b \in B$} \For{each censored list
-    $l \in L$}
-  \State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1.7cm}{Permute
-    the unassigned ranks, $\{d_l+1, \ldots, P\}$, and assign them
-    randomly to the items \emph{not} found in the list, i.e.,
-    $\Lambda^\complement=X\setminus\Lambda_l$. Combine the
-    result with $\Lambda_l$ to fill out the list.}
-\EndFor
-\State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1cm}{Let sra($b$) be the sequential rank agreement computed from the
-filled out lists.}
-\EndFor
-\State Return element-wise averages across all $B$ permutations of sra($b$).
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-The proposed approach is based on two assumptions: 1) that the most
-interesting items are found in the top of the lists, and 2) that the
-censored rankings provide so little information that it is reasonable
-to assume that they can be represented by a random order. The first
-assumption is justifiable because we have already accepted that it is
-reasonable to rank the items in the first place. The second assumption
-is fair in the light of the first assumption provided that we have a
-``sufficiently large'' part of the top of the lists available.
-
-When the two assumptions are satisfied then it is clear that the
-interesting part of the sequential rank agreement curves is restricted
-to the depths where the number of censored items is
-low. Generally, without additional prior knowledge about the
-distribution of the censored items it seems reasonable to restrict the
-attention of the sequential rank agreement to depths smaller than
-$\max(d_1, \ldots, d_L)$ where at least one list provides an
-uncensored item.
-
-Like for fully observed lists we generally expect the sequential rank
-agreement to start low and then increase unless the lists are
-completely unrelated (in which case the sequential rank agreement will
-be constant at a high level) or if the lists mostly agree on the
-ranking (in which case the sequential rank agreement will also be
-constant but at a low level). For partially ranked lists we also
-expect a changepoint around the depth where the lists are
-censored. This is an artefact stemming from the fact that we assume
-that the remainder of the lists can be replaced by a simple
-permutation of the missing items.
-
-% \subsection{Comparison to average overlap}
-% The average overlap is a measure that is widely used in comparisons of
-% two fully observed ranked lists \citep{Fagin2003,Webber2010} and the
-% idea behind the average overlap closely resembles the sequential rank
-% agreement. In this section we present an extension of the average
-% overlap to more that two lists and compare this to sequential rank
-% agreement.
-
-% The overlap among $L$ lists observed until depth $d$ is defined as the
-% proportion of items that are found in all $L$ lists compared to the
-% possible number of items found in all lists observed until depth $d$:
-% \begin{align}
-%   O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}\label{def:overlap}
-% \end{align}
-% Typically the overlap is only considered for pairwise comparisons
-% \citep{Bar-Ilan2006,Boulesteix2009} but formula \ref{def:overlap}
-% is easily extended to the situation where $L>2$ as shown above.
-
-% The average overlap at depth $d$ is defined as the average of the
-% overlaps until depth $d$,
-% \begin{equation}
-% AO(d) = \frac1d\sum_{i=1}^d O(d).
-% \end{equation}
-% This definition directly ensures that further emphasis is put on items
-% in the top of the lists since their item specific overlap contributes
-% to the calculation of average overlap at higher depths.
-
-% There are three main drawbacks of the average overlap method: First it
-% is highly sensitive to single items in the top of one the lists.  The
-% average overlap profile may change dramatically depending on the
-% presence or absence of similar items in the beginning of the
-% lists. For example, if all lists have the same item at the top then
-% the average overlap attains its maximum at 1, but if just one of the
-% lists differ then the overage overlap starts at its minimum of
-% 0. Secondly, when the number of lists, $L$, is high then it might be
-% difficult to obtain a non-zero overlap (and hence a non-zero average
-% overlap) at the beginning of the lists because the overlap will be
-% zero until an item is present in all $L$ lists. Finally, the average
-% overlap has an interpretation in terms of moving averages of
-% percentages (of $L$ sets) which is somewhat less intuitive than our
-% proposed sequential rank distance.
-
-% In the following we compare the average overlap and the sequential
-% rank agreement using the analysis of the Golub data
-% \citep{Golub1999}. We consider a typical application of the average
-% overlap and thus restrict attention to only two gene rankings obtained
-% by the $t$ tests and logistic regression. These two marginal modeling
-% approaches should result in very similar gene rankings. The first two
-% columns of Table~\ref{tab1} show the actual top-10 ranking of the
-% predictors.
-
-% % chunks 8,9
-
-% \begin{figure}[htbp]
-% \begin{center}
-% \includegraphics[width=.49\textwidth]{paper-fig-ao1}
-% \includegraphics[width=.49\textwidth]{paper-fig-ao2}
-% \end{center}
-%  \caption{Average overlap (left figure) and sequential rank agreement
-%  (right figure) for comparing two different analysis methods
-%  (marginal $t$ test and logistic regression)
-%     applied to the Golub data. The black lines are based on the full
-%     set of predictors while the red lines have the two highest
-%     associated predictors removed
-%     from the data before computing the average overlap and sequential
-%     rank agreement.}
-%   \label{fig:case1}
-%  \end{figure}
-
-%  The black lines in Figure~\ref{fig:case1} show the average overlap
-%  and sequential rank agreement for these data. It is clear from both
-%  plots that there is perfect agreement towards the top of the ranked
-%  lists: the average overlap is 1 and the sequential rank agreement is
-%  zero.  However, if we remove those two items from the data (i.e.,
-%  gene/predictor 2124 and 896) and redo the analyses then we get
-%  substantially different curves for the average overlap but not for
-%  sequential rank agreement (red lines in Figure~\ref{fig:case1}). For
-%  the sequential rank agreement we get roughly the same estimate of
-%  agreement from rank 3 as we did from the full dataset. This indicates
-%  that sequential rank is more robust against small perturbations of
-%  the data. For the average overlap the curves completely change when
-%  these two items are removed and which would lead to a different
-%  conclusion about the agreement among the items from rank 3 downwards.
-
-\section{Evaluating sequential rank agreement}
-To interpret the sequential rank agreement values we propose two
-different benchmark values corresponding to two different
-hypotheses. We wish to determine if we observe better agreement than
-what would be expected if there were no relevant information available
-in the data.
-
-The first reference hypothesis is
-\begin{eqnarray}
-H_0  & : &  \text{The list rankings correspond to completely randomly}\label{eq:permutationHypothesis}\\
-       &  & \text{permuted lists}\nn
-\end{eqnarray}
-which not only assumes that there is no information in the data on
-which the rankings are based but also that the methods are completely
-independent.
-
-Alternatively, we can remove the restriction on the independence among
-the methods by only requiring that there is no information in the
-ranking but that the rankings are all based on applying the
-method/approaches to the same data
-\begin{eqnarray*}
-\widetilde H_0 & :&  \text{The list rankings are based on data that contain}\\
-& &   \text{no association to the outcome.}
-\end{eqnarray*}
-
-\subsection{Permutation-based inference}
-
-$H_0$ is a quite unrealistic null hypothesis but we can easily obtain
-realizations from that null hypothesis simply by permuting the items
-within each list and then computing the sequential rank agreement for
-the permuted lists. In the fully observed case each experiment
-contains $L$ lists of random permutations of the items in $X$. For the
-censored/partially observed case we first permute the items
-$X_1,\dots,X_P$ and then censor list $l$ at $d_l$ (\ie, each list so
-it has the same number of observed rankings as was observed for list
-$l$ in the original dataset). The sequential rank agreement curve from
-the the original lists can then be compared to, say, the pointwise
-95\% quantiles of the observed rank agreements obtained under $H_0$.
-
-To obtain the distribution under $\widetilde H_0$ the idea is to
-repeat the ranking procedures for unassociated data many times.  Thus,
-we first permute the outcome variable of the dataset on which the
-rankings are based. This removes any association between the predictor
-variables and the outcome. Then we apply the methods to the permuted
-dataset to generate $L$ new rankings and subsequently compute the
-sequential rank agreement. Note that we only permute the outcomes and
-preserve the structure of the candidate predictors. This randomization
-approach requires that we have the original data available and as such
-it may not be possible to evaluate $\widetilde H_0$ in all situations.
-
-If the sequential rank agreement for the original data lies
-substantially below the distribution of the sequential rank agreements
-obtained under either $H_0$ or $\widetilde H_0$ then this suggests that
-the original ranked lists agree \emph{more} than what we would expect
-in data with no information, and therefore that the information in the
-lists is significantly more in agreement than what would be expected.
-
-Figure~\ref{fig:example1} shows the empirical distributions of
-sequential rank agreement under $H_0$ and $\widetilde H_0$ each based
-on $400$ permutations of the Golub data described in Section
-\ref{sec:amfol}. Not surprisingly, the sequential rank agreement under
-$\widetilde H_0$ is lower than the sequential rank agreement under
-$H_0$ because the four methods used to rank the data ($t$ test,
-logistic regression, elastic net, and MIC) generally tend to identify
-(and rank) similar predictors even if there are only spurious
-associations. The two bottom panels in Figure~\ref{fig:example1} also
-indicate that the observed sequential rank agreement (the black line)
-is better than what would be expected by chance for data that contain
-no information since it lies below the reference areas. The censored
-data also suggests that there may be at most 1 or 2 ranked items
-towards the top of the lists that yield a result better than what
-would be expected (the bottom-right plot).
-
-It is important to stress that neither $H_0$ nor $\widetilde H_0$ are
-related to questions regarding the association between the outcome and
-the predictors in the dataset. Both hypotheses are purely considering
-how the rankings agree in a situation where there is no relevant
-information available in the data used for creating the rankings.
-
-\subsection{Asymptotic inference of change in agreement}\label{sec:chgpoint}
-In many applications it is of interest to estimate a list depth which
-satisfies a changepoint criterion since that corresponds to a change
-in agreement among the list ranks. In this section we investigate the
-theoretical properties of our proposed method for this specific
-task. As in \cite{hall:schi:2012} we consider an infinite set
-of lists and study the asymptotic behaviour for $L\to\infty$.
-However, we do not allow the length of the lists to change with $L$
-because in most applications the lengths of the lists are fixed.
-
-We start by showing that $\widehat{\textrm{sra}}_L$ is a consistent
-estimator of $\textrm{sra}$ for $L \rightarrow \infty$. The result is
-stated in Theorem \ref{thm:consistency}.
-
-% below and requires the
-% following assumptions.
-% \begin{assumption}
-% We assume that:
-% \begin{enumerate}[label=\Alph*)]
-% \item{The $L$ lists are independent and identically distributed
-    % samples from a probability distribution $Q$ on the set of lists.}
-% \item{$\E_Q[R(X_p)^2] < \infty$ uniformly in $p$.}
-% \item{$|S(d)| > 0$ uniformly in $d$}.
-% \end{enumerate}
-% \end{assumption}
-% We note that assumption $C$ implies that for every list depth $d$
-% there must be a non-zero probability of ... under the probability
-% measure $Q$.
-
-\begin{theorem}
-  Assume that $\{R_l(X)\}_{l=1}^L$ are independent draws from a
-  probability distribution $Q$ on the set of lists $\Pi$. Then,
-  $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty =
-  o_P(1)$.
-\label{thm:consistency}
-\begin{proof}
-See Appendix \ref{sec:appA}.
-\end{proof}
-\end{theorem}
-
-We now define the changepoint as the first crossing point of the
-sequential rank agreement and a threshold function
-$q\colon\,\{1,\ldots,P\} \mapsto \mathbb{R}_{\geq 0}$. The values of
-$q$ could be a deterministic constant or, % in the experimental setting,
-for example, the limits-of-agreement obtained in randomly permuted
-lists corresponding to the null-hypothesis in equation
-(\ref{eq:permutationHypothesis}). We define the superlevel set of the
-sequential rank agreement with respect to $q$ as
-\begin{align}
-  \mathcal{L}(q) = \left\{d : \textrm{sra}(d) \geq q(d)\right\}.
-\end{align}
-A changepoint $d^\ast(q)$ in the list agreement is then defined by the position
-\begin{equation}
-  d^\ast(q) =
-  \begin{cases}
-    \inf(\mathcal{L}(q)) & |\mathcal{L}(q)| > 0\\
-  P & |\mathcal{L}(q)| = 0
-  \end{cases}
-\end{equation}
-corresponding to the first list depth where the sequential rank
-agreement exceeds the threshold if such a position exists. Otherwise,
-the full list is in agreement according to $q$ and the changepoint is
-set to the full length of the lists. The empirical superlevel set is
-similarly defined as
-\begin{align}
- \widehat{\mathcal{L}}_L(\widehat{q}_L) = \left\{d : \widehat{\textrm{sra}}_L(d) \geq \widehat{q}_L(d)\right\}
-\end{align}
-where we allow the threshold function to depend on the sample size as
-well. The estimated changepoint is therefore
-\begin{align}
-  \widehat{d^\ast_L}(\widehat{q}_L) &= 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| > 0)\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) + 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0)P.
-\end{align}
-
-The consistency of the estimated changepoint, $\widehat{d^\ast_L}(\widehat{q}_L)$, follows from Theorem \ref{thm:consistency} by the following corollary. 
-\begin{corollary}\label{thecor}
-Let $\widehat{q}_L$ be a positive threshold function such that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ for some limiting
-function $q$.  Then, $\widehat{d^\ast_L}(\widehat{q}_L) \overset{P}{\longrightarrow} d^\ast(q)$ for $L \rightarrow \infty$.
-\label{cor:consistency}
-\begin{proof}
-See Appendix \ref{sec:appB}.
-\end{proof}
-\end{corollary}
-
-Corollary \ref{thecor} indicates that we can use the a threshold
-function $\widehat{q}_L$ estimated under the null hypothesis as
-discussed in the previous section as a limiting threshold function for
-inferring the depth $d$, where the observed sequential rank agreement
-first crosses the threshold of the null threshold, \ie, the depth
-until which the observed ranked lists are in better agreement than
-expected under the null hypothesis. In that sense the threshold
-function serves the same role as the limits of agreement in method
-comparison studies, except that the threshold funciton is not constant
-but can accommodate the changing nature of the number of items used
-for the computation of the equential rank agreement for a given depth.
-
-In practice we can compute an estimate of the threshold function under
-the null using the permutation approach sketched in the previous
-section which makes it relevant even in small sample settings.
-
-% \mcomment{Elaborate more on how to apply this in practice (finite
-  % sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style?  }
-
-\section{Application to ovarian cancer data}
-We now consider an application of the sequential rank agreement to two
-datasets consisting of MALDI-TOF (Matrix-Assisted Laser
-Desorption/Ionization Time Of Flight) mass spectra obtained from blood
-samples from patients with either benign or malignant ovarian
-tumors. The datasets are sub-samples of the Danish MALOVA and DACOVA
-study populations.
-
-The MALOVA study is a multidisciplinary Danish study on ovarian cancer
-\citep{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
-with an ovarian tumor and referred for surgery from the participating
-departments of gynecology were enrolled continuously from December
-1994 to May 1999. For the purpose of illustration we use a random
-sub-sample of $119$ patients with a total of $58$ patients with
-malignant ovarian cancers as cases and $61$ patients with benign
-ovarian tumors as controls. The DACOVA study is another
-multidisciplinary Danish study on ovarian cancer which included about
-$66\%$ of the female population of Denmark
-\citep{bertelsen1991protocol}. The study aimed to continuously enroll
-all patients that were referred to surgery of an ovarian tumor
-clinically suspected to be cancer during the period from 1984 to
-1990. Similarly, we use a random sub-sample from the DACOVA study of
-$113$ patients with a total of $54$ malignant ovarian cancers and $59$
-benign ovarian tumors/gynecologic disorders.
-
-Each spectrum consists of $49642$ samples over a range of mass-to-charge ratios
-between $800$ to $20000$ Dalton which we downsample on an equidistant grid of
-5000 points by linear interpolation. We then preprocess the downsampled
-spectra individually by first removing the slow-varying baseline intensity
-with the SNIP algorithm \citep{ryan1988snip} followed by a
-normalization with respect to the total ion count. Finally, we standardize
-the $5000$ predictors to have column-wise zero mean and unit variance in
-each dataset.
-
-We use the two datasets to illustrate how the sequential rank
-agreement can be applied in two different scenarios. In the first
-scenario we assess the agreement of four different statistical
-classification methods in how they rank the predictors according to
-their importance for distinguishing benign and malignant tumors. In
-the second scenario we assess the agreement among rankings of
-individual predicted risks of having a malignant tumor. The first
-scenario is relevant in the context of biomarker discovery and the
-latter is important e.g., when ranking patients according to immediacy of
-treatment.
-
-Four classification methods are considered: Random Forest
-\citep{breiman2001random} implemented in the R package
-\texttt{randomForest} \citep{liaw2002classification}, logistic Lasso
-\citep{tibshirani1996regression} and Ridge regression
-\citep{segerstedt1992ordinary} both implemented in the R package
-\texttt{glmnet} \citep{friedman2010regularization}, and Partial Least
-Squares Discriminant Analysis (PLS-DA) \citep{boulesteix2004pls}
-implemented in the R package \texttt{caret} \citep{Jed-Wing:2014aa}.
-
-In both scenarios we use the MALOVA data to train the statistical
-models, and in both situations the agreements are assessed with
-respect to perturbations of the training data in the following
-manner. We repeatedly draw a random sub-sample (without replication)
-consisting of 90\% of the MALOVA observations and train the four
-models on each sub-sample. We use 1000 iterations for the sub-sampling
-procedure.
-
-All four methods depend on a tuning parameter. The tuning parameter for
-Lasso and Ridge regression is the degree of penalization, and for PLS-DA
-it is the number of components (the dimensionality of the subspace).
-We estimate these separately for each sub-sample by a 20 times repeated
-5-fold cross-validation procedure. For the Random Forest we grow a fixed
-number of $5000$ trees and let the tuning parameter be the number of
-predictors randomly sampled at each split. We estimate this by a binary
-search with respect to minimizing the Out-of-Bag classification error estimate.
-
-The implementation of Lasso and Ridge regression in the
-\texttt{glmnet} package offers three different cross-validated
-optimization criteria for the penalty parameter: total deviance,
-classification accuracy and area under ROC. We apply all three
-criteria to our data to investigate their effect on the agreements.
-Note also that the Lasso models produce censored lists depending on
-the value of the penalty parameter.
-
-\subsection{Agreement of predictor rankings}
-For each of the four methods, each of the 1000 models trained on
-the 1000 sub-samples of the MALOVA data produces a ranking of the 5000
-predictors according to their importance for discriminating between
-the tumor types. For the Random Forest classifier the predictors are ranked according
-to the Gini index, while for the logistic Lasso and Ridge regression
-models we order by absolute magnitude of the estimated
-regression coefficients. For the PLS-DA model the importance of the
-predictors is based on a weighted sum of the absolute coefficients
-where the weights are proportional to the reduction in the sums of
-squares across the components.
-
-The right panel of Figure \ref{fig:app1} shows the sequential rank
-agreement of the estimated importance of the 5000 predictors.  For
-clarity of presentation we zoom in on the agreement up to list depth
-600. At deeper list depths all agreement curves are approximately constant.
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.49\textwidth]{pics/riskAgreementPlot}%
-\includegraphics[width=.49\textwidth]{pics/predictorAgreementPlot}
-\end{center}
-\caption{Left panel: Sequential rank agreement of 1000 rankings of the
-  predicted risks of malignant tumor. For each method the different
-  rankings were obtained by first training models in 1000 random
-  sub-samples of the MALOVA data and then predicting the risk of
-  malignant tumor in the 113 DACOVA patients. Right panel: Sequential
-  rank agreement of 1000 rankings of the 5000 predictors. The rankings
-  were obtained from the same 1000 trained models.}
- \label{fig:app1}
-\end{figure}
-
-For most of the sequential rank agreement curves we see, as expected,
-that they start low, indicating good agreement, followed by an
-increase until they approximately become constant. This has the
-interpretation that the agreement across the different sub-samples is
-higher in the top as compared to the tail of the lists for all these
-classification methods. The changepoints where the curves become
-approximately constant are the list depths where the ranks of the
-remaining items become close to uniformly random.
-
-A not expected shape of the agreement curves is seen for the Ridge
-models for all three tuning criteria. They all show higher
-disagreement in the top of the lists followed by a decrease. The
-reason behind this behavior is rather subtle. Looking at the
-distribution of the absolute value of the regression coefficients we see that a large
-proportion of them are numerically very close to zero and have almost
-equal absolute value. This is a general feature of the Ridge models in
-this dataset and seen for all the 1000 trained models.
-This implies that when predictors are ranked according to the
-magnitude of their coefficients, their actual order becomes more
-uncertain and more close to a random permutation. This problem can be
-alleviated by truncating all predictors with absolute coefficient values
-below a given threshold thereby introducing an artificial censoring of the lists.
-For the Ridge models tuned with the deviance criterion, Figure \ref{fig:app2} (left) shows the
-the sequential rank agreement where for each of the 1000 trained
-models the predictors were artificially censored when their absolute
-coefficient value was lower than the $0.1\%$ quantile of the 5000
-absolute coefficient values. The curve was calculated
-using Algorithm \ref{sra-algorithm} with $B=1000$ and $P=5000$. The corresponding
-curve from Figure \ref{fig:app1} (right panel) is shown for comparison.
-Even though the number of censored predictors is very small compared to the total
-number of predictors, the effect on the sequential rank agreement is
-substantial and with the artificial censoring the shape of the curves
-is as expected, starting low and then increasing.
-
-Looking at the agreement curves for the Lasso models in Figure \ref{fig:app1}
-(right) we clearly see the effect of the sparsity inducing penalization giving
-rise to censored lists. These curves were similarly calculated using \ref{sra-algorithm}
-and $1000$ random permutations. Under the deviance optimization criterion the
-median number of non-zero coefficients was 33 (range 16 to 50) and for the
-class accuracy criterion 14 (range 4 to 56). These values correspond to
-the list depths where the agreement curves become constant as a result of
-the subsequent censoring.
-
-\subsection{Agreement of individual risk predictions}
-\label{sec:airp}
-To assess the stability of the individual risk predictions we apply
-the predictors from the DACOVA dataset to each of the models. The
-predicted probabilities are then ranked in decreasing order such that the
-patients with the highest risk of a malignant tumor appears in the top
-of the list. Figure \ref{fig:app1} (left) shows the sequential rank
-agreement separately for each method, based on the 1000 risk predictions
-obtained from the models trained in the same $1000$ random sub-samples
-of the MALOVA data.
-
-Most curves start low and then increase indicating higher agreement
-among high risk patients. This is expected if we rank the individuals
-according to highest risk of disease. However, it is also expected
-that individuals with very low risk also show high agreement. In this
-case we order the patients according to (high) risk prediction but we
-could essentially also have reversed the order to identify the
-patients that have low risk prediction.
-
-An exception is the risk prediction agreement for the Lasso tuned with the AUC criterion which
-shows very low agreement among the high values of the predicted
-risks. The reason is that optimizing the penalty parameter with
-respect to the AUC criterion tends to favor a very high penalty value
-causing only a single predictor to be selected in each of the 1000
-iterations. This results in a lack of generalizability to the DACOVA
-data which gives rise to the higher disagreement in the predicted risks.
-In the extreme case where the penalty becomes so high
-that none of the predictors are selected by the Lasso, the sequential
-rank agreement for the predicted probabilities becomes undefined since
-all the ranks will be ties.
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.49\textwidth]{pics/ridgeCensoring}%
-\includegraphics[width=.49\textwidth]{pics/aucPlot}
-\end{center}
-\caption{Left panel: Sequential rank agreement for Ridge regression
-  obtained by artificially censoring predictors when their absolute
-  coefficient values are lower than the 0.1\% quantile. Right panel:
-  Box plots of AUC values across the $1000$ sub-samples with respect
-  to the known class labels of the DACOVA data.}
- \label{fig:app2}
-\end{figure}
-
-Comparing the left and right panels of Figure \ref{fig:app1} it can
-further be seen that some of the methods show better agreement with
-respect to the predicted probabilities than for ranking the importance
-of the predictors and vice versa. Ridge regression shows higher agreement across training
-sets for the risk predictions than PLS-DA, and PLS-DA shows higher
-agreement for predictor importance than Ridge regression.
-
-Lasso shows similar agreement for ranking the risk predictions than
-PLS-DA (except for the AUC criterion), and poorer agreement for
-ranking predictors. This reason for the latter is the high
-auto-correlation between the intensities in the mass spectra which
-leads to collinearity issues in the regression models. It is
-well-known that variable selection with the Lasso penalty does not
-preform very well when the predictors are highly correlated. The
-collinearity does, however, not affect the agreement of the risk
-predictions, as seen on the left panel of Figure \ref{fig:app1}, since
-it is not so important which specific variable that gets selected from
-a group of highly correlated predictors when the purpose is risk
-predictions.
-
-It appears that Ridge regression tuned with the AUC criterion achieves
-the best performance with respect to the stability of ranking the
-individual predicted risk probabilities. It must, however, be stressed
-that the sequential rank agreement in this application is only
-concerned with the agreement of the risk predictions across
-sub-samples and not with the actual accuracy of the risk
-predictions. Thus, we also computed the AUC values for the different models
-based on the DACOVA data. The distributions across the $1000$
-sub-samples for a selection of the models is shown in the right panel
-of Figure \ref{fig:app2}.  Here we see that PLS-DA attains the highest
-AUC values with a median value of $0.70$ while the Ridge model with
-the AUC criterion attains a median AUC of $0.49$. This implies that
-while Ridge regression optimized with respect to the AUC criterion
-achieves the best sequential rank agreement, it performs similar to a
-random coin toss with respect to classifying the DACOVA patients.
-In practice both concerns are of importance.
-
-\section{Simulation study --- comparison of list agreements}
-
-We present results of a simulation study where we investigated the
-small sample properties of our proposed method (sra) and compared it
-with the topK method proposed by \citet{hall:schi:2012} with respect
-to list depth agreement. In each simulation run, we generated a
-dataset with 1000 predictors and 400 observations. The predictors were
-drawn independently from a standard Gaussian distribution. The outcome
-was drawn from a Gaussian distribution with variance 1 and mean that
-depended additively on the first 15 predictors, i.e.,
-
-$$y_i = \sum_{j=1}^{15}x_{ij} + \epsilon_i,$$
-
-where $y_i$ is the $i$th response, and $x_{ij}$ is the $j$th predictor
-for the $i$th measurement, and $\epsilon_i\sim N(0, 1)$. In each
-simulation run we obtained \(L\) ranked lists of the 1000 predictors.
-This was done by first drawing \(L\) bootstrap samples (with
-replacement) and then in each bootstrap sample ranking the 1000
-predictors according to their marginal $t$ test statistics. Thus, the
-simulation results compare lists that are obtained with the same
-technique on bootstrap versions of the same dataset. We report results
-from two sets of simulations (each based on 1000 simulated datasets):
-\begin{enumerate}
-\item[] Scenario I: \(L=8\) and $q\in\{3,4,5,6,7,8,9,10\}$.
-\item[] Scenario II: \(q=5\) and $L\in\{3,5,10,50\}$.
-\end{enumerate}
-In both scenarios, we summarized the distribution of the estimated
-depth of agreement and the average number of different predictors
-found at a given depth.
- % Asymptotic results are available for both methods as the number of
- % lists increases, so this will investigate how much the results might
- % change as $L$ increases. For a fixed constant threshold of $q_L=5$ we
- % increase the number of available lists to investigate the sensitivity
- % of the results to $L$.
-  % Throughout the simulations we
-  % assume a constant, prespecified threshold function, $q_L$, and try
-  % different values for the constant. The approach by
-  % \citet{hall:schi:2012} uses a window size parameter, $K$, to
-  % represent the maximum difference in rankings before they consider an
-  % item in two lists to be in agreement.  Because the standard
-  % deviation --- which is used in the sequential rank agreement
-  % definition of $\hat{A}_L$--- squares its differences, it tends to
-  % give more weight to larger rank differences and less weight to
-  % smaller differences compared to the mean absolute difference.  To
-  % ease comparison between the sequential rank agreement and the
-  % \citet{hall:schi:2012} approach we would like to use the same
-  % threshold for both methods, and in order to do this we use the
-  % median absolute deviance throughout the simulations in lieu of the
-  % pooled standard deviation.
- % For all parameter combinations we ran 1000 simulations and report the
- % distribution of the estimated depth of agreement and the average
- % number of different predictors found at a given depth.
-The results from Scenario I are shown in the left panel of
-Figure~\ref{fig:example99}. The double histograms show the
-distribution of the estimated depth of agreement. We see that sra
-results in a substantially lower depth of agreement than the topK
-method. Also the average numbers of unique predictors (bold numbers in
-the double histograms) are markedly smaller for sra. Even larger
-differences are found if we used the Euclidean distance instead of the
-median absolute distance for the sequential rank agreement (results
-not shown). As expected the depth of agreement increases when the
-threshold for agreement/window increases.
-\begin{figure}[htbp]
-   \begin{center}
- \includegraphics[width=.49\textwidth]{Ksim.pdf}
- \includegraphics[width=.49\textwidth]{Lsim}
- \end{center}
- \caption{Left panel: Simulation study showing distribution of
-   estimated rank agreements for sra and topK for varying thresholds
-   and fixed number of lists $L=8$. The bold numbers are the average
-   number of unique predictors in the set found to be in
-   agreement. Right panel: Simulation results for varying number of
-   lists, and with fixed threshold of $q=5$.}
- \label{fig:example99}
- \end{figure}
- The right panel of Figure~\ref{fig:example99} shows the results from
- Scenario II. The number of lists has little impact but at all depths
- sra is more conservative than topK. As a consequence we also see that
- sra systematically selects fewer predictors in the parts of
- agreement.
-
-\section{Discussion}
-In this article we address the problem of comparing ranked lists of
-the same items. Our proposed method can handle both the situation
-where the underlying data to generate the ranked lists are available
-and the situation where the only available data is the actual ranked
-lists. In addition, censored ranked lists where only the ranks of the
-top $k$ ranked items are known can be accommodated as well. The
-proposed agreement measure can be interpreted as the average distance
-between an item's rank and the average rank assigned to that item
-across lists. Thus the measure determines how well the lists agree on
-the ranks of a specific item.
-
-The sequential rank agreement is extremely versatile. We have shown
-that it can be used not only to compare ranked lists of items produced
-from different samples/populations but that it also can be used to
-study the ranks obtained from different analysis methods on the same
-data as well as to evaluate the stability of the ranks from a single
-method by bootstrapping (or sub-sampling) the data repeatedly and
-comparing the ranks obtained from training the models in the
-bootstrapped data.
-
-The sequential rank agreement can be used to determine the depth at
-which the rank agreement becomes too large to be desirable based on
-prior requirements or acceptable differences, or it can be used to
-visually determine when the change in agreement becomes too large.  In
-that regard the investigator can have prior limits on the level of
-agreement that is acceptable.
-
-While the sequential rank agreement is primarily an exploratory tool
-we have suggested two null hypotheses that can be used to evaluate the
-sequential rank agreement obtained. Note that none of the two null
-hypotheses are concerned with the actual ``true ranking'' but are
-purely concerned with consistency/stability of the rankings among the
-lists. As such we cannot determine if the rankings are good but only
-whether they agree. The sequential rank agreement curve can be
-compared visually to the curves obtained under either of the null
-distributions and simple point-wise $p$-values can be obtained for
-each depth by counting the number of sequential rank agreements under
-the null hypothesis that is less than or equal to the observed rank
-agreement. Two simple extensions can be pursued to make more formal
-uniform tests: one would be to use a Kolmogorov-Smirnov-like test and
-use the largest difference between the variance-weighted sequential
-rank agreement curve and the mean null rank agreement curve as a test
-statistic. Alternatively, a changepoint analysis could be made on the
-sequential rank agreement in order to determine the depths at which
-there are ``jumps'' in the rank agreement. These jumps would
-correspond to depths for which the agreement among the lists was
-substantially worse and could serve as indicators for when the lists
-no longer agree sufficiently satisfactory.
-
-Finally, we have --- whenever possible --- used all available ranks
-from the lists.  We could choose to censor the rankings of all the
-items that were deemed not to be significant if information on both
-rankings and the corresponding evidence for significance was
-available. That would ensure that there would be put less emphasis on
-the agreement of the non-significant items and it would be easier to
-identify a change in agreement among the items that were deemed to
-be relevant. In our application section we have successfully introduced
-such an artificial censoring for the predictor rankings obtained with
-ridge regression.
-
-We note that the sequential rank agreement is still marred by problems
-that generally apply to ranking of items and/or
-individuals. Collinearity in particular can be a huge problem when
-bootstrapping data or when comparing different analysis methods. For
-example, marginal analyses where each item is analyzed separately will
-assign similar ranks to two highly correlated predictors while methods
-that provide a sparse solution such as the Lasso will just rank one of
-the two predictors high while the other might have a very low rank.
-Thus in such a scenario we would expect low agreement of the rankings
-from Lasso and marginal analyses simply because of the way correlated
-predictors are handled. This is not a shortcoming of the sequential
-rank agreement but is a problem general to all ranked lists.
-
-Another caveat with the way the sequential rank agreement is defined
-is the use of the standard deviation to measure agreement. The
-standard deviation is an integral part of the limits-of-agreement as
-discussed by \citet{alt:bland:1983}. However, the standard deviation
-can also be unstable when the number of observations is low and
-alternative measures such as the median absolute deviance may prove more
-stable in some situations. However, the current definition using the
-standard deviation is analogous to the approach used for agreement in
-method comparison studies so we have used that.
-
-In conclusion we have introduced a method for evaluation of ranked
-(partial/censored) lists that can be easily interpreted and that can be
-applied to a large number of situations.  The method presented here
-can be adapted further by using it to compare and classify statistical
-analysis methods that agree on the rankings they provide or by using
-the rank agreement to optimize a hyper-parameter in, say, elastic net
-regularized regression where the rank agreement is used to determine
-the mixing proportion between the $L_1$ and the $L_2$ penalty.
-%Finally, the proposed method may be adapted (with some additional
-%assumptions) to the situation where the lists are subsets of each
-%other.
-We will be investigating these extensions further in the future.
-
-%\bibliographystyle{rss}
-\bibliographystyle{rss}
-\bibliography{paperref}
+This document provides proofs of the theorem and corollary shown in the main manuscript.
 
@@
 
-\newpage
-\section{Proof of Theorem \ref{thm:consistency}}
+% \newpage
+\section{Proof of Theorem 3.1}
 \label{sec:appA}
@@
 \begin{align}
-  N(p; d) = \sum_{s=1}^p 1\left(Q(R(X_s) \leq d) > 0\right), \quad p = 1,\ldots, P
+  N(p; d) = \sum_{s=1}^p 1\left(Q(R(X_s) \leq d) > \varepsilon\right), \quad p = 1,\ldots, P
 \end{align}
 which runs through the list elements in an arbitrary order and counts
-how many list elements that have a strictly positive probability to have
-rank less than or equal to $d$ under the probability measure $Q$.
+how many list elements that have a positive probability $>\varepsilon$ to have
+rank less than or equal to $d$ under the probability measure $Q$ for some prespecified constant $\varepsilon\in[0, 1)$.
 Note that the cardinality of the set $S(d)$ in equation
-(\ref{eq:sumset}) is equal to the counting process evaluated at the last element: $|S(d)| = N(P;d)$.
+(2.3) is equal to the counting process evaluated at the last element: $|S(d)| = N(P;d)$.
 The empirical counterpart of the counting process is given by
 \begin{align}
-  \widehat{N}_L(p; d) &= \sum_{s=1}^p 1\left( \frac{1}{L} \sum_{l=1}^L 1(R_l(X_s) \leq d) > 0\right)\\
-        &= \sum_{s=1}^p 1\left(\widehat{Q}_L(R(X_s) \leq d) > 0\right)\nn.
+  \widehat{N}_L(p; d) &= \sum_{s=1}^p 1\left( \frac{1}{L} \sum_{l=1}^L 1(R_l(X_s) \leq d) > \varepsilon\right)\\
+        &= \sum_{s=1}^p 1\left(\widehat{Q}_L(R(X_s) \leq d) > \varepsilon\right)\nn.
 \end{align}
@@
 The sequential rank agreement given in equation
-(\ref{def:sra}) may therefore be rewritten as the following integral
+(2.5) may therefore be rewritten as the following integral
 with respect to $N$
 \begin{align}
-  \textrm{sra}(d) = \int_{1}^P \frac{A(X_p)}{N(P; d)}\mathrm dN(p; d).
+  \textrm{sra}(d) = \int_{1}^P \frac{A(X_p)}{N(P; d)} 1(N(P;d)>0)\mathrm dN(p; d).
 \end{align}
@@
 \begin{align}
-  \widehat{\textrm{sra}}_L(d) &= \int_1^P \frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)}\mathrm d \widehat{N}_L(p; d) 
+  \widehat{\textrm{sra}}_L(d) &= \int_1^P \frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} 1(N(P;d)>0) \mathrm d \widehat{N}_L(p; d) 
 \end{align}
@@
 \begin{align}
-  U_L(d) &= \int_1^P\frac{A(X_p)}{N(P;d)} \mathrm d\left(\widehat{N}_L(p; d) - N(p; d)\right)\\
-  V_L(d) &=  \int_1^P \left(\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right)\mathrm d\widehat{N}_L(p; d)
+  U_L(d) &= \int_1^P\frac{A(X_p)}{N(P;d)} 1(N(P;d)>0) \mathrm d\left(\widehat{N}_L(p; d) - N(p; d)\right)\\
+  V_L(d) &=  \int_1^P \left(\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right) 1(N(P;d)>0) \mathrm d\widehat{N}_L(p; d)
 \end{align}
@@
 \begin{align}
-  \left|U_L(d)\right| \leq \left(\sup_{p \in 1,\ldots,P}\left|\frac{A(X_p)}{N(P;d)}\right|\right)\left(\widehat{N}_L(p; d) - N(p; d)\right)\biggr\rvert_1^P
+  \left|U_L(d)\right| \leq \left(\sup_{p \in 1,\ldots,P}\left|\frac{A(X_p)}{N(P;d)} 1(N(P;d)>0) \right|\right)\left(\widehat{N}_L(p; d) - N(p; d)\right)\biggr\rvert_1^P
 \end{align}
@@
 \begin{align}
- V_L(d) &\leq \sup_{p \in 1,\ldots,P}\left|\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right| \int_1^P \left|\mathrm d\widehat{N}_L(p; d)\right|\label{eq:secondTerm}
+ V_L(d) &\leq \sup_{p \in 1,\ldots,P}\left|\left(\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right) 1(N(P;d)>0) \right| \int_1^P \left|\mathrm d\widehat{N}_L(p; d)\right|\label{eq:secondTerm}
 \end{align}
@@
 $\widehat{A_L}(X_p) = A(X_p) + o_P(1)$ uniformly in $p$ since
-$R_i(X_p)$ in equation (\ref{eq:empDistance}) for $i=1,\ldots,L$ are
+$R_i(X_p)$ in equation (2.2) for $i=1,\ldots,L$ are
 independent and $Q$-identically distributed random variables with
@@
 
-\section{Proof of Corollary \ref{cor:consistency}}
+\section{Proof of Corollary 3.1}
 \label{sec:appB}
@@
 \end{align}
-as a consequence of Theorem \ref{thm:consistency}, the assumption that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ and the continuous mapping theorem. This completes the proof.
+as a consequence of Theorem 3.1, the assumption that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ and the continuous mapping theorem. This completes the proof.
 

### 2018-01-24 latex/biosts3.tex
Updatedd version

--- 2017-05-23 latex/superranker-manuscript.tex
@@
-\documentclass[12pt,a4paper]{article}
+\documentclass[oupdraft]{bio}
+%\usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, citecolor=citecolor]{hyperref}
+% \documentclass[12pt,a4paper]{article}
 \pdfoutput=1
@@
 \usepackage{subcaption}
-\usepackage{natbib}
+% \usepackage{natbib}
 \usepackage{xspace}
 \usepackage{calc}
-%\usepackage[noae]{Sweave}
-\usepackage{authblk}
+% \usepackage[noae]{Sweave}
 \usepackage{algorithm}
 \usepackage{algpseudocode}
-\usepackage{enumitem}
+\usepackage{xcolor}
+
+% \usepackage{authblk}
+
+\usepackage[nolists,nomarkers]{endfloat}
+
 %\usepackage{geometry}
@@
 \newcommand{\eg}{\emph{e.g.}\@\xspace}
-
 \newcommand{\nn}{\nonumber}
 
-\newcommand{\mcomment}[1]{\textcolor{red}{\textbf{#1}}}
+\providecolor{added}{rgb}{1,0,0}
+\newcommand{\added}[1]{{\color{added}{}#1}}
+% \newcommand{\added}[1]{#1}   % Use after correction
+
+% \newcommand{\mcomment}[1]{\textcolor{red}{\textbf{#1}}}
 
@@
 
-\usepackage[amsthm,thmmarks]{ntheorem}
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}
-\newtheorem{definition}{Definition}
-\newtheorem{proposition}{Proposition}
-\newtheorem{corollary}{Corollary}
-\newtheorem{lemma}{Lemma}
-\newtheorem{assumption}{Assumption}
-\newtheorem{example}{Example}
-\renewcommand\proofSymbol{\ensuremath{\blacksquare}}
+%\usepackage[amsthm,thmmarks]{ntheorem}
+% \theoremstyle{plain}
+% \newtheorem{theorem}{Theorem}
+%\newtheorem{definition}{Definition}
+%\newtheorem{proposition}{Proposition}
+\newtheorem{corollary}{\normalfont\scshape Corollary}[section]
+%\newtheorem{lemma}{Lemma}
+%\newtheorem{assumption}{Assumption}
+%\newtheorem{example}{Example}
+%\renewcommand\proofSymbol{\ensuremath{\blacksquare}}
 
@@
 
+\title{Sequential rank agreement methods for comparison of ranked lists}
+
+\author{CLAUS THORN EKSTRØM$^\ast$, THOMAS ALEXANDER GERDS, ANDREAS
+  KRYGER JENSEN\\[4pt]
+%
+\textit{%
+Biostatistics, Department of Public Health,
+University of Copenhagen,
+Øster Farimagsgade 5 B, P.O.B. 2099,
+DK-1014 Copenhagen K, Denmark}
+\\[2pt]
+% E-mail address for correspondence
+{ekstrom@sund.ku.dk}}
+
 % Use square brackets for numbering affiliations
 
+% Running headers of paper:
+\markboth%
+% First field is the short list of authors
+{C. T. Ekstrøm and others}
+% Second field is the short title of the paper
+{Sequential rank agreement for comparison of ranked
+  lists}
+
 \maketitle
 
-\begin{abstract}
-  The comparison of alternative rankings of a set of items is a
-  general and common task in applied statistics. Predictor variables
-  are ranked according to magnitude of association with an outcome,
-  prediction models rank subjects according to the personalized risk
-  of an event, and genetic studies rank genes according to their
-  difference in gene expression levels. This article constructs
-  measures of agreement among two or more ordered lists. We use the
-  standard deviation of the ranks to define a measure of agreement
-  that both provides an intuitive interpretation and can be applied to
-  any number of lists even if some or all are incomplete. The approach
-  can identify changepoints in the agreement of the lists, and the
-  sequential changes of agreement as a function of the depth of the
-  lists can be compared graphically to a permutation based reference
-  set or evaluated using analytical asymptotic results.  The
-  usefulness of these tools are illustrated using gene rankings, and
-  using data from two Danish ovarian cancer studies where we assess
-  the within and between agreement of different statistical
-  classification methods.
+\footnotetext{To whom correspondence should be addressed.}
+
+\begin{abstract} { The comparison of alternative rankings of a set of
+    items is a general and common task in applied
+    statistics. Predictor variables are ranked according to magnitude
+    of association with an outcome, prediction models rank subjects
+    according to the personalized risk of an event, and genetic
+    studies rank genes according to their difference in gene
+    expression levels. \added{We propose a sequential rank agreement
+      measure to quantify the rank agreement among two or more ordered
+      lists. This measure has an intuitive interpretation, it can be
+      applied to any number of lists even if some are partially
+      incomplete, and it provides information about the agreement
+      along the lists. The sequential rank agreement can be evaluated
+      analytically or be compared graphically to a permutation based
+      reference set in order to identify changes in the list
+      agreements.}  The usefulness of this measure is illustrated
+    using gene rankings, and using data from two Danish ovarian cancer
+    studies where we assess the within and between agreement of
+    different statistical classification methods.}  {rank agreement;
+    incomplete ranking; order statistic; methods comparison;
+    permutation}
 \end{abstract}
 
-Key words: sequential rank agreement, partial ranking, order
-statistic, gene rankings, methods comparison, variable selection,
-permutation
+%Key words: rank agreement, partial ranking, order
+%statistic, gene rankings, methods comparison, variable selection,
+%permutation
 
@@
 genes according to their difference in gene expression levels across
-samples.
-
-When several rankings of the same items are available, a common
-research question is to what extent they agree. In particular, is it
-possible to identify an optimal rank until which the lists agree on
-the items?  A typical situation arises in high-dimensional genomics
-studies when several analysis methods are applied to rank a list of
-genes according to their association with a phenotype, treatment
-effect or other outcome. A measure of agreement of gene rankings
-obtained by different methods could often help to identify which genes
-are worth to pursue in further experiments.
-
-Several approaches exist to measure the ``distance'' between two
-ranked lists. Of these, Kendall's $\tau$ \citep{kendall1948} and
-Spearman’s footrule $\rho$ \citep{Spearman1910} are among the most
-well-known, but these measures do not distinguish between agreement in
-the top versus towards the bottom of the lists and they both provide a
-single measure of the overall distance. \citet{Shieh1998} proposed a
-weighted version of Kendall's $\tau$ where each pair of rankings can
-be assigned different weights, and \citet{Yilmaz2008} proposed the
-$\tau_{ap}$ which places higher emphasis on the top of the lists.
-Spearman’s footrule uses the ranks of the variables for calculation of
-the distance and the use of ranks are also employed in the $M$ measure
-of \citet{Bar-Ilan2006} where the reciprocal rank differences are used
-to calculate the similarity measure.
+samples.  \added{Two common research questions are of interest when
+  several rankings of the same items are available: 1) To what extent
+  do the lists agree on the rankings and how will that change as we go
+  through the lists, and 2) Is it possible to identify an optimal rank
+  until which the lists agree on the items?
+
+  A typical situation where
+  these questions arise is in high-dimensional genomics studies such
+  as genome-wide association studies where several analysis methods
+  (e.g., regression methods, lasso, random forest) can be used
+  identify and rank millions of gene variants according to their
+  association with the outcome. The importance (i.e., ranking) of each
+  gene variant may vary from method to method and a consensus summary
+  of agreement of the findings is needed to determine which gene
+  variants to investigate more closely in subsequent validation
+  studies. To minimiize expenses it is only of interest to consider
+  gene variants that have high ranking across the different methods.
+  Another situation where multiple ranked lists appear is in machine
+  learning where the stability of the ranks produced by a
+  ``black-box''-technique can be evaluted by bootstrapping the data
+  repeatedly and comparing the ranks obtained from training the models
+  on the bootstrapped data. Assessing which items are stable (i.e.,
+  consistant across bootstrap samples) will help to weed out possible
+  fluke findings.}
+
+In this article we introduce sequential rank agreement for measuring
+agreement among  ranked lists.  The general idea is to
+define agreement based on the sequence of ranks from a subset of the first $d$
+items in each list. As agreement metric we adapt the limits of
+agreement known from agreement between quantitative variables
+\citep{alt:bland:1983,Carstensen2010} but any measure for agreement
+could essentially be used. Our proposed approach allows us to
+compare multiple lists simultaneously, it provides a dynamic
+measure of agreement as a function of the depth in the lists, it places
+higher weight on items at the top of the list, it accommodates partially
+observed lists of varying lengths, and has a natural interpretation
+that directly relates to the ranks.  Graphical illustration of
+sequential rank agreement allows us to infer any
+changepoints, i.e., a list depth where a substantial change in the agreement of
+the lists occur but we also provide asymptotical and randomization-based graphical
+tools to compare the observed rank agreement to the expected agreement
+found in non-informative data.  In this sense our approach is a
+combination and generalization of some of the ideas of
+\citet{Carterette2009} and \citet{Boulesteix2009}. The former compares
+two rankings based on the distance between them as measured by a
+multivariate Gaussian distribution and the latter presents an overview
+of approaches for aggregation of ranked lists including bootstrap and
+leave-one-out jackknife approaches. We show the asymptotic
+distribution of the endpoint of agreement and discuss how to infer the
+distribution in small-sample situations using computational
+methods. This enables us to make inferences about the endpoint of
+agreement even for situations where there is no actual changepoint
+present but just a gradual decline in agreement.
+
+% Several approaches exist to measure the ``distance'' between two
+% ranked lists. Of these, Kendall's $\tau$ \citep{kendall1948} and
+% Spearman’s footrule $\rho$ \citep{Spearman1910} are among the most
+% well-known, but these measures do not distinguish between agreement in
+% the top versus towards the bottom of the lists and they both provide a
+% single measure of the overall distance. \citet{Shieh1998} proposed a
+% weighted version of Kendall's $\tau$ where each pair of rankings can
+% be assigned different weights, and \citet{Yilmaz2008} proposed the
+% $\tau_{ap}$ which places higher emphasis on the top of the lists.
+% Spearman’s footrule uses the ranks of the variables for calculation of
+% the distance and the use of ranks are also employed in the $M$ measure
+% of \citet{Bar-Ilan2006} where the reciprocal rank differences are used
+% to calculate the similarity measure.
 
@@
 is a situation that is only relevant in special cases, whereas the
-simulation-based null distribution approach of \citet{samp:verd:2013}
-is guaranteed to work also for small samples.
-
-In this article we introduce sequential rank agreement for measuring
-agreement among (partially) ranked lists.  The general idea is to
-define agreement based on the sequence of ranks from the first $d$
-items in each list. As agreement metric we adapt the limits of
-agreement known from agreement between quantitative variables
-\citep{alt:bland:1983,Carstensen2010} but any measure for agreement
-could essentially be used. Our proposed approach allows us to
-compare multiple lists simultaneously, it provides a dynamic
-measure of agreement as a function of the depth in the lists, it places
-higher weight on items at the top of the list, it accommodates partially
-observed lists of varying lengths, and has a natural interpretation
-that directly relates to the ranks.  Graphical illustration of
-sequential rank agreement potentially allows us to infer a
-changepoint, i.e., a list depth where a substantial change in the agreement of
-the lists occur but we also provide asymptotical and randomization-based graphical
-tools to compare the observed rank agreement to the expected agreement
-found in non-informative data.  In this sense our approach is a
-combination and generalization of some of the ideas of
-\citet{Carterette2009} and \citet{Boulesteix2009}. The former compares
-two rankings based on the distance between them as measured by a
-multivariate Gaussian distribution and the latter presents an overview
-of approaches for aggregation of ranked lists including bootstrap and
-leave-one-out jackknife approaches. We show the asymptotic
-distribution of the endpoint of agreement and discuss how to infer the
-distribution in small-sample situations using computational
-methods. This enables us to make inferences about the endpoint of
-agreement even for situations where there is no actual changepoint
-present but just a gradual decline in agreement.
-
-The manuscript is organized as follows: In the next section we define
+\added{simulation-based null distribution approach of
+  \citet{samp:verd:2013} does not rely on asymptotic results to
+  evaluate their pairwise findings.}
+
+The manuscript is organized as follows: The next section defines
 sequential rank agreement for multiple ranked lists and discuss how to
-handle incomplete lists. In section 3 we present and discuss
-approaches to evaluate the results obtained from sequential rank
-agreement. Finally we apply the sequential rank agreement to two
-Danish ovarian cancer studies and compare our method to the method of
-\citet{hall:schi:2012} in a small sample simulation study before we
-discuss the findings along with possible extensions. The approaches
-presented in this manuscript are available in the R package
-\texttt{SuperRanker} which can be found on CRAN.
+handle incomplete lists. In section 3 we discuss approaches to
+evaluate the results obtained from sequential rank agreement. Finally
+we apply the sequential rank agreement to two Danish ovarian cancer
+studies and compare our method to the method of \citet{hall:schi:2012}
+in a small sample simulation study before we discuss the findings
+along with possible extensions. The approaches presented in this
+manuscript are available in the R package \texttt{SuperRanker} which
+can be found on CRAN.
 % chunk 1
@@
 
+The agreement among the lists regarding the rank given to an item $X_p$
+can be measured by the variance across the lists
+\begin{align}
+  A(X_p) &= \E_Q\left[\left(R(X_p)-\E_Q R(X_p)\right)^2\right]\\
+   &= \sum_{r\in\Pi} \left(r(X_p)-\E_Q R(X_p)\right)^2 Q(r),\nn
+\end{align}
+where $\Pi$ is the set of all permutations of $X$, $Q$ is a
+probability mass function on $\Pi$, and
+$\E_Q R(X_p)=\sum_{r\in\Pi} r(X_p)Q(r)$. The empirical counterpart is
+\begin{align}
+\widehat{A}_L(X_p) = \frac{1}{L-1}\sum_{i=1}^L (R_i(X_p) - \overline{R}_L(X_p))^2, \quad \overline{R}_L(X_p) = \frac{1}{L}\sum_{i=1}^L R_i(X_p).\label{eq:empDistance}
+\end{align}
+For each item, the function $\widehat{A}_L$ has an interpretation as the expected
+Euclidean distance of the individual rankings from the expected
+ranking over the $L$ lists, and it corresponds to the same measure
+that is used in method comparison studies to compute the limits of
+agreement \citep{alt:bland:1983}.
+
+For an integer $1\le d\le P$ we define the expected set of unique
+items found by merging the first $d$ elements across the possible
+lists:
+\begin{align}
+S(d) &= \left\{X_p; \left(\sum_{r\in\Pi} 1\left(r(X_p)\le d\right) Q(r)\right) > \added{\varepsilon}\right\}\label{eq:sumset}
+\end{align}
+where $1(\cdot)$ denotes the indicator function\added{, and where $\varepsilon\in[0,1)$ is a pre-specified constant that sets the minimum proportion of lists that an item must be present in before it is added to $S(d)$.
+When $\varepsilon=0$ then an item is included as soon as it is present in just one list.}
+The empirical counterpart is the set of unique items ranked less than or
+equal to $d$ in any of the $L$ lists:
+\begin{equation}
+\widehat{S}_{L}(d) = \left\{X_p; \left(\frac{1}{L}\sum_{l=1}^L 1\left(R_l(X_p)\le d\right)\right)>\added{\varepsilon}\right\},
+\end{equation}
+which is exemplified in Panel (c) of Table \ref{tab:example}.
+
+We define the \emph{sequential rank agreement} as the weighted
+expected agreement of the items found in the set $S(d)$:
+\begin{equation}
+\added{  \textrm{sra}(d) = \left\{\begin{array}{cl}\frac{1}{|S(d)|}\sum_{p \in S(d)}A(X_p)  & \text{when $|S(d)|>0$} \\ 0 & \text{otherwise}\end{array}  \right.}\label{def:sra}
+\end{equation}
+where $|S(d)|$ is the cardinality of the set $S(d)$. \added{As stated, we are only interested in sra($d$) when $|S(d)|>0$ so the value 0 above is somewhat arbitrary.}
+The empirical counterpart when $|S(d)|>0$ is equivalently given by
+\begin{equation}
+  \widehat{\textrm{sra}}_L(d) = \frac{\sum_{\{p \in \widehat{S}_{L}(d)\}}(L-1)\widehat{A}_L(X_p)}{(L-1)|\widehat {S}_{L}(d)|}.\label{def:sraEst}
+\end{equation}
+
+Values of \textrm{sra} close to zero when $|S(d)|>0$ suggest that the
+lists agree on the rankings while larger values suggest
+disagreement. \added{If $|S(d)|=0$ then no items were sufficiently
+  frequent among the observed lists and we can conclude that the lists
+  do not agree above the threshold $\varepsilon$.  } If the ranked
+lists are identical then the value of sequential rank agreement will
+be zero for all values of $d$. %position or depths $d$ down the lists.
+
+\subsection{Interpreting and applying sequential rank agreement}
+\label{sec:amfol}
+
+The sequential rank agreement is equivalent to the pooled variance of
+the items found in $S(d)$. Thus, the square root of the sequential
+rank agreement measures the average of the average difference in
+rankings among the lists for the items we have included until depth
+$d$.
+
+In method comparison studies the observed agreement is compared to a
+pre-specified acceptable limit and we can do similarly. For easy
+visualization of the rank agreement we suggest to plot
+$\sqrt{\widehat{\textrm{sra}}_L(d)}$ corresponding to the pooled SD
+against $d$.
+
+As an example consider the data by \citet{Golub1999} (also found in
+\citet{Dudoit2002}) where 3051 gene expression values measured on 38
+tumor mRNA samples were used to classify between acute lymphoblastic
+leukemia (ALL) and acute myeloid leukemia (AML). Several possible
+analysis methods can be used for this for example marginal unequal
+variances two-sample $t$ tests, marginal logistic regression analyses,
+elastic net logistic regression \citep{friedman2010regularization},
+and marginal maximum information content correlations (MIC)
+\citep{Reshef2011}, and we would like to identify a set of genes that
+consistently are most likely to be associated to leukemia.
+
+\added{
+For the first two methods, the genes were ranked according to minimum
+$p$ value, for logistic regression the genes were ordered by size of
+the corresponding coefficients (after standardization), and MIC was
+ordered by absolute correlation which resulted in the top rankings
+seen in Table~\ref{tab1}. The ranked lists appear to agree that genes
+2124 and 829 are among the most interesting while the best ranked
+gene from MIC, gene 378, is not found in the top 10 for two of the
+other methods.
+
+% The simplest case occurs when all $L$ lists are fully observed, \ie, when
+% we observe the rank of all $P$ items for all $L$ lists. Fully
+% observed ranked lists are common and arise, for example, when
+% different statistical methods are applied to a dataset
+% to produce lists of predictors ranked according to their importance,
+% or if the same analysis method is applied to data from different
+% populations.
+
+%With fully observed lists we can plot the sequential rank agreement
+%\eqref{def:sra} as a function of depth $d$. An example is seen in the top
+%panels of Figure~\ref{fig:example1} where four analysis methods were
+%used to rank 3051 gene expression values measured on 38 tumor mRNA
+%samples in order to classify between acute lymphoblastic leukemia
+%(ALL) and acute myeloid leukemia (AML)
+%\citep{Golub1999}. Preprocessing of the gene expression data was done
+% as described in \citet{Dudoit2002} and the four different analysis
+% approaches were:
+
+The sequential rank agreement curve (using $\varepsilon=0$ to decide
+that an item should be present in just a single list before it is
+included) shown in the left plot of Figure~\ref{fig:example1} show the
+average distance in ranks for the genes considered among the first $d$
+positions. Not surprisingly, the sequential rank agreement is better
+towards the top of the lists (smaller values on the $y$ axis
+corresponds to \emph{better} agreement) than towards the bottom of the
+lists. It is clear from the curve in Figure~\ref{fig:example1} that
+there is a substantial deterioration in agreement (higher sra) after
+depth 5. Thus, if we were to restrict attention to a small set of
+predictors then our prime focus would be on the items found among the
+top-5 lists from Table~\ref{tab1}). The choice of which depth until
+which we think the lists agree can be chosen either from a
+pre-specified threshold for an acceptable difference in rankings or
+from a pre-specified item set size. A changepoint analysis on the
+sequential rank agreement would be able to identify depths where a
+substantial increase/change in rank agreement occurs and would be
+another way to identify sets of items that share agreement among the
+lists if a pre-specified acceptable rank agreement threshold is not
+given.}
+
+% latex table generated in R 3.2.3 by xtable 1.8-2 package
+% Fri Apr  8 10:22:58 2016
+
+% chunks 2,3,4,5,6,7
+
+% If the lists agree on all items then the sequential rank agreement is
+% zero for all depths. 
+Generally, changes in the level of rank agreement suggest that there
+are sets of items that are ranked similarly in all lists while other
+items constitute set(s) that have been assigned vastly different
+ranks. When the lists are likely to agree on a few top ranked items
+then the sequential rank agreement curve will start at a low level and
+then increase until it levels off exactly as is seen in
+Figure~\ref{fig:example1}.
+
+\subsection{Analysis of incomplete lists}
+Incomplete or partial lists are a common occurrence that arise, for
+example, in case of missing data (items), when comparing top $d$ list
+results from publications, or when some methods only rank a subset of
+the items. For example, penalized regression such as the Lasso
+provides a sparse set of predictors that have non-zero
+coefficients. There is no obvious ordering of the set of predictors
+whose coefficient has been shrunk to zero and thus we end up with a
+partial ordering.  Incomplete lists also occur if for example the analyst
+restricts attention to the ranks of items that have been found to be
+statistically significant.
+
+Sequential rank agreement can be generalized to incomplete lists in
+the following way. Let $\Lambda_l\subset X$ be the subset of $d_l$
+items that have been ranked highest in list $l$.  The case where all
+lists are incomplete at the same depth $d$ corresponds to
+$d_1=\cdots=d_L=d$. For incomplete lists the rank function becomes
+\begin{equation}
+	\widetilde R_l(X_p) = \begin{cases}
+	     \{R_l(X_p)\} & \text{for } X_p\in \Lambda_l\\
+	     \{d_l+1,\dots,P\} & \text{for} X_p \not\in \Lambda_l
+        \end{cases}
+\end{equation}
+where we only know that the rank for the unobserved items in list $l$
+must be larger than the largest rank observed in that list.
+
+The agreement, $A(X_p)$, cannot be computed directly for all
+predictors in the presence of incomplete lists because the exact rank
+for some items will be unknown. Also, recall that the rankings within
+a single list are not independent since each rank must appear exactly
+once in each list. Thus, we cannot simply assign the same number
+(\eg, the mean of the unassigned ranks) to the missing items since
+that would result in \emph{less} variation of the ranks and hence less
+variation of the agreement, and it would artificially introduce a
+(downward) bias of agreement for items that are missing in multiple
+lists.
+
+Instead we randomize the ranks $\{d_{l}+1,\dots,P\}$ to the items that
+do not occur in list $\Lambda_l$. One realization of the $L$ rankings
+of the set $X$ is obtained by randomizing the missing items of each
+list. By randomizing a large number of times we can compute
+\eqref{def:sra} for each realization, and then compute the sequential
+rank agreement as the pointwise (for each depth) average of the rank
+agreements. The algorithm is described in detail in
+Algorithm~\ref{sra-algorithm}.
+
+The proposed approach is based on two assumptions: 1) that the most
+interesting items are found in the top of the lists, and 2) that the
+ranks that are missing from the lists provide so little information
+that it is reasonable to assume that they can be represented by a
+random order. The first assumption is justifiable because we have
+already accepted that it is reasonable to rank the items in the first
+place. The second assumption is fair in the light of the first
+assumption provided that we have a ``sufficiently large'' part of the
+top of the lists available.
+
+When the two assumptions are satisfied then it is clear that the
+interesting part of the sequential rank agreement curves is restricted
+to depths where the number of items without ranks available is
+low.
+%Generally, without additional prior knowledge about the
+%underlying distribution of the ranks it seems reasonable to restrict
+%attention of the sequential rank agreement to lists no longer than the
+%longest of the $L$ lists examined.
+Similar to fully observed lists we generally expect the sequential
+rank agreement to start low and then increase unless the lists are
+completely unrelated (in which case the sequential rank agreement will
+be constant at a high level) or if the lists mostly agree on the
+ranking (in which case the sequential rank agreement will also be
+constant but at a low level). For incomplete ranked lists we also
+expect a changepoint around the depth where the lists are become
+incomplete. This is an artefact stemming from the fact that we assume
+that the remainder of the lists can be replaced by a simple
+permutation of the missing items. \added{Note that if an item is
+  ranked highly in a few lists but unranked in the remaining lists
+  then it gets poor rank agreement since we only compare
+  whether the lists agree on their rankings and in this case they
+  clearly do not.}
+
+The right-most plot of Figure~\ref{fig:example1} shows the impact
+of restricting the Golub data such that only top-20 lists are
+available instead of full lists of length 3051 (20 was chosen to
+resemble the list lengths that might be harvested from published manuscripts). The sequential rank agreement increases much quicker
+because the incomplete lists introduce more noise in the estimation of
+the agreement, but it is still possible to see that the top of the
+list has a sequantial rank agreement that is not substantially
+different from the full lists.
+
+% \subsection{Comparison to average overlap}
+% The average overlap is a measure that is widely used in comparisons of
+% two fully observed ranked lists \citep{Fagin2003,Webber2010} and the
+% idea behind the average overlap closely resembles the sequential rank
+% agreement. In this section we present an extension of the average
+% overlap to more that two lists and compare this to sequential rank
+% agreement.
+
+% The overlap among $L$ lists observed until depth $d$ is defined as the
+% proportion of items that are found in all $L$ lists compared to the
+% possible number of items found in all lists observed until depth $d$:
+% \begin{align}
+%   O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}\label{def:overlap}
+% \end{align}
+% Typically the overlap is only considered for pairwise comparisons
+% \citep{Bar-Ilan2006,Boulesteix2009} but formula \ref{def:overlap}
+% is easily extended to the situation where $L>2$ as shown above.
+
+% The average overlap at depth $d$ is defined as the average of the
+% overlaps until depth $d$,
+% \begin{equation}
+% AO(d) = \frac1d\sum_{i=1}^d O(d).
+% \end{equation}
+% This definition directly ensures that further emphasis is put on items
+% in the top of the lists since their item specific overlap contributes
+% to the calculation of average overlap at higher depths.
+
+% There are three main drawbacks of the average overlap method: First it
+% is highly sensitive to single items in the top of one the lists.  The
+% average overlap profile may change dramatically depending on the
+% presence or absence of similar items in the beginning of the
+% lists. For example, if all lists have the same item at the top then
+% the average overlap attains its maximum at 1, but if just one of the
+% lists differ then the overage overlap starts at its minimum of
+% 0. Secondly, when the number of lists, $L$, is high then it might be
+% difficult to obtain a non-zero overlap (and hence a non-zero average
+% overlap) at the beginning of the lists because the overlap will be
+% zero until an item is present in all $L$ lists. Finally, the average
+% overlap has an interpretation in terms of moving averages of
+% percentages (of $L$ sets) which is somewhat less intuitive than our
+% proposed sequential rank distance.
+
+% In the following we compare the average overlap and the sequential
+% rank agreement using the analysis of the Golub data
+% \citep{Golub1999}. We consider a typical application of the average
+% overlap and thus restrict attention to only two gene rankings obtained
+% by the $t$ tests and logistic regression. These two marginal modeling
+% approaches should result in very similar gene rankings. The first two
+% columns of Table~\ref{tab1} show the actual top-10 ranking of the
+% predictors.
+
+% % chunks 8,9
+
+% \begin{figure}[htbp]
+% \begin{center}
+% \includegraphics[width=.49\textwidth]{paper-fig-ao1}
+% \includegraphics[width=.49\textwidth]{paper-fig-ao2}
+% \end{center}
+%  \caption{Average overlap (left figure) and sequential rank agreement
+%  (right figure) for comparing two different analysis methods
+%  (marginal $t$ test and logistic regression)
+%     applied to the Golub data. The black lines are based on the full
+%     set of predictors while the red lines have the two highest
+%     associated predictors removed
+%     from the data before computing the average overlap and sequential
+%     rank agreement.}
+%   \label{fig:case1}
+%  \end{figure}
+
+%  The black lines in Figure~\ref{fig:case1} show the average overlap
+%  and sequential rank agreement for these data. It is clear from both
+%  plots that there is perfect agreement towards the top of the ranked
+%  lists: the average overlap is 1 and the sequential rank agreement is
+%  zero.  However, if we remove those two items from the data (i.e.,
+%  gene/predictor 2124 and 896) and redo the analyses then we get
+%  substantially different curves for the average overlap but not for
+%  sequential rank agreement (red lines in Figure~\ref{fig:case1}). For
+%  the sequential rank agreement we get roughly the same estimate of
+%  agreement from rank 3 as we did from the full dataset. This indicates
+%  that sequential rank is more robust against small perturbations of
+%  the data. For the average overlap the curves completely change when
+%  these two items are removed and which would lead to a different
+%  conclusion about the agreement among the items from rank 3 downwards.
+
+\section{Evaluating sequential rank agreement results}
+To evaluate the sequential rank agreement values we propose two
+different benchmark values corresponding to two different
+hypotheses. We wish to determine if we observe better agreement than
+would be expected if there were no relevant information available in
+the data.
+
+The first reference hypothesis is
+\begin{eqnarray}
+H_0  & : &  \text{The list rankings correspond to completely randomly}\label{eq:permutationHypothesis}\\
+       &  & \text{permuted lists}\nn
+\end{eqnarray}
+which not only assumes that there is no information in the data on
+which the rankings are based but also that the methods used to provide
+the rankings are completely independent.
+
+Alternatively, we can remove the restriction of the independence among
+the methods used to generate the $L$ ranked lists and only require
+that there is no information contained in the rankings but that the
+rankings are all based on applying the method/approaches to the same
+data
+\begin{eqnarray*}
+\widetilde H_0 & :&  \text{The list rankings are based on data that contain}\\
+& &   \text{no association to the outcome.}
+\end{eqnarray*}
+
+This alternative null hypothesis addresses the fact that some ranking
+methods are more likely to provide similar rankings of the same data
+because the ranking methods focus on the same features of the data
+rather than because of any information contained in the data.
+
+\subsection{Permutation-based inference}
+
+$H_0$ is a quite unrealistic null hypothesis but we can easily obtain
+realizations from that null hypothesis simply by permuting the items
+within each list and then computing the sequential rank agreement for
+the permuted lists. In the fully observed case each experiment
+contains $L$ lists of random permutations of the items in $X$. For the
+incomplete case we first permute the items $X_1,\dots,X_P$ and then assign missing ranks
+for list $l$ from  $d_l$ to $P$ (\ie, each list has the same number of
+observed rankings as was observed for list $l$ in the original
+dataset). The sequential rank agreement curve from the original
+lists can then be compared to, say, the pointwise 95\% quantiles of
+the observed rank agreements obtained under $H_0$.
+
+To obtain the distribution under $\widetilde H_0$ the idea is to
+repeat the ranking procedures for unassociated data many times.
+\added{For each resample, we
+  \begin{enumerate}
+  \item first permute the outcome variable in the dataset. This
+    removes any association between the predictor variables and the
+    outcome while keeping the structure in the predictors, and 
+  \item we apply the same methods that was used for the original data to the permuted dataset to generate $L$
+    new rankings and compute the sra for the unassociated data.
+  \end{enumerate}
+} Note that we only permute the outcomes and thus preserve the
+internal structure of the predictors. This randomization approach
+requires that the original data is available and as such it may not be
+possible to evaluate $\widetilde H_0$ in all situations.
+
+If the sequential rank agreement for the original data lies
+substantially below the distribution of the sequential rank agreements
+obtained under either $H_0$ or $\widetilde H_0$ then this suggests that
+the original ranked lists agree \emph{more} than expected
+in data with no information, and therefore that the information in the
+lists is significantly more in agreement than what would be expected.
+
+Figure~\ref{fig:example1} shows the empirical distributions of
+sequential rank agreement under $H_0$ and $\widetilde H_0$ each based
+on $400$ permutations of the Golub data from Section \ref{sec:amfol}.
+\added{ Figure~\ref{fig:example1} indicates that the observed
+  sequential rank agreement for the Golub data is significantly better
+  than what would be expected by chance for data that contain no
+  information since it lies below the reference areas if the lists
+  were just random ($H_0$ corresponding to the blue area). However, if
+  we consider $\widetilde H_0$ then the sequential rank agreement is
+  just inside the red area and we conclude that the agreement seen in
+  the Golub data is \emph{not} significantly better than what we would
+  expect when we remove the association between the predictors and
+  the outcome.}
+
+The incomplete data also suggests that there may be at most 1 or 2
+ranked items towards the top of the lists that yield a result better
+than what would be expected (the bottom-right plot). Not surprisingly,
+the sequential rank agreement under $\widetilde H_0$ is lower than the
+sequential rank agreement under $H_0$ because the four methods used to
+rank the data ($t$ test, logistic regression, elastic net, and MIC)
+generally tend to identify (and rank) similar predictors even if there
+are only spurious associations.
+
+\added{It is important to stress that neither $H_0$ nor
+  $\widetilde H_0$ are related to questions regarding the association
+  between the outcome and the predictors in the dataset. Both
+  hypotheses are purely considering how the rankings agree in a
+  situation where there is no relevant information available in the
+  data used for creating the rankings. It is also worth pointing out,
+  that if the lists are short ($P$ low) and there are few lists ($L$
+  low) then the number of possible different permutations under the
+  null is small and the $p$ value obtained may be fluctuating if the
+  number of permutations is small. We have found that a number of
+  permutations over 500 works well for smaller samples. The number of
+  permutations can be lowered for larger data sizes.}
+
+\subsection{Asymptotic inference of change in agreement}\label{sec:chgpoint}
+In many applications it is of interest to estimate a list depth which
+satisfies a changepoint criterion since that corresponds to a change
+in agreement among the list ranks. \added{In particular, a changepoint
+  will provide a data-driven indicator as to the depth until the lists
+  exhibit a change in rank agreement, and would consequently be an
+  obvious choice for identifying the set of items that the lists agree
+  the upon the most}. In this section we investigate the theoretical
+properties of our proposed method for this specific task. As in
+\citet{hall:schi:2012} we consider an infinite set of lists and study
+the asymptotic behaviour for $L\to\infty$. The list lengths are not
+allowed to change with $L$ since the lengths are fixed in most
+applications.
+
+We start by showing that $\widehat{\textrm{sra}}_L$ is a consistent
+estimator of $\textrm{sra}$ for $L \rightarrow \infty$.
+%The result is stated in Theorem \ref{thm:consistency}.
+
+% below and requires the
+% following assumptions.
+% \begin{assumption}
+% We assume that:
+% \begin{enumerate}[label=\Alph*)]
+% \item{The $L$ lists are independent and identically distributed
+    % samples from a probability distribution $Q$ on the set of lists.}
+% \item{$\E_Q[R(X_p)^2] < \infty$ uniformly in $p$.}
+% \item{$|S(d)| > 0$ uniformly in $d$}.
+% \end{enumerate}
+% \end{assumption}
+% We note that assumption $C$ implies that for every list depth $d$
+% there must be a non-zero probability of ... under the probability
+% measure $Q$.
+
+\begin{theorem}
+  Assume that $\{R_l(X)\}_{l=1}^L$ are independent draws from a
+  probability distribution $Q$ on the set of lists $\Pi$. Then,
+  $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty =
+  o_P(1)$.
+\label{thm:consistency}
+\begin{proof}
+See Appendix A in the supplementary material.%\ref{sec:appA}.
+\end{proof}
+\end{theorem}
+
+We now define the changepoint as the first crossing point of the
+sequential rank agreement and a threshold function
+$q\colon\,\{1,\ldots,P\} \mapsto \mathbb{R}_{\geq 0}$. The values of
+$q$ could be a deterministic constant or, % in the experimental setting,
+for example, the limits-of-agreement obtained in randomly permuted
+lists corresponding to the null-hypothesis in equation
+(\ref{eq:permutationHypothesis}). We define the superlevel set of the
+sequential rank agreement with respect to $q$ as
+\begin{align}
+  \mathcal{L}(q) = \left\{d : \textrm{sra}(d) \geq q(d)\right\}.
+\end{align}
+A changepoint $d^\ast(q)$ in the list agreement is then defined by the position
+\begin{equation}
+  d^\ast(q) =
+  \begin{cases}
+    \inf(\mathcal{L}(q)) & |\mathcal{L}(q)| > 0\\
+  P & |\mathcal{L}(q)| = 0
+  \end{cases}
+\end{equation}
+corresponding to the first list depth where the sequential rank
+agreement exceeds the threshold if such a position exists. Otherwise,
+the full list is in agreement according to $q$ and the changepoint is
+set to the full length of the lists. The empirical superlevel set is
+similarly defined as
+\begin{align}
+ \widehat{\mathcal{L}}_L(\widehat{q}_L) = \left\{d : \widehat{\textrm{sra}}_L(d) \geq \widehat{q}_L(d)\right\}
+\end{align}
+where we allow the threshold function to depend on the sample size as
+well. The estimated changepoint is therefore
+\begin{align}
+  \widehat{d^\ast_L}(\widehat{q}_L) &= 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| > 0)\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) + 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0)P.
+\end{align}
+
+The consistency of the estimated changepoint, $\widehat{d^\ast_L}(\widehat{q}_L)$, follows from Theorem \ref{thm:consistency} by the following corollary. 
+\begin{corollary}\label{thecor}
+Let $\widehat{q}_L$ be a positive threshold function such that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ for some limiting
+function $q$.  Then, $\widehat{d^\ast_L}(\widehat{q}_L) \overset{P}{\longrightarrow} d^\ast(q)$ for $L \rightarrow \infty$.
+\label{cor:consistency}
+\begin{proof}
+See Appendix B in supplementary material.%\ref{sec:appB}.
+\end{proof}
+\end{corollary}
+
+Corollary \ref{thecor} indicates that we can use the threshold
+function $\widehat{q}_L$ estimated under the null hypothesis as
+discussed in the previous section as a limiting threshold function for
+inferring the depth $d$, where the observed sequential rank agreement
+first crosses the threshold of the null threshold, \ie, the depth
+until which the observed ranked lists are in better agreement than
+expected under the null hypothesis. In that sense the threshold
+function serves the same role as the limits of agreement in method
+comparison studies, except that the threshold function is not constant
+but can accommodate the changing nature of the number of items used
+for the computation of the sequential rank agreement for a given depth.
+%
+In practice we can compute an estimate of the threshold function under
+the null using the permutation approach sketched in the previous
+section which makes it relevant even for small sample settings.
+
+% \mcomment{Elaborate more on how to apply this in practice (finite
+  % sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style?  }
+
+\section{Application to ovarian cancer data}
+We now consider an application of the sequential rank agreement to two
+datasets consisting of MALDI-TOF (Matrix-Assisted Laser
+Desorption/Ionization Time Of Flight) mass spectra obtained from blood
+samples from patients with either benign or malignant ovarian
+tumors. The datasets are sub-samples of the Danish MALOVA and DACOVA
+study populations.
+
+The MALOVA study is a multidisciplinary Danish study on ovarian cancer
+\citep{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
+with an ovarian tumor and referred for surgery from the participating
+departments of gynecology were enrolled continuously from December
+1994 to May 1999. For the purpose of illustration we use a random
+sub-sample of $119$ patients with a total of $58$ patients with
+malignant ovarian cancers as cases and $61$ patients with benign
+ovarian tumors as controls. The DACOVA study is another
+multidisciplinary Danish study on ovarian cancer which included about
+$66\%$ of the female population of Denmark
+\citep{bertelsen1991protocol}. The study aimed to continuously enroll
+all patients that were referred to surgery of an ovarian tumor
+clinically suspected to be cancer during the period from 1984 to
+1990. Similarly, we use a random sub-sample from the DACOVA study of
+$113$ patients with a total of $54$ malignant ovarian cancers and $59$
+benign ovarian tumors/gynecologic disorders.
+
+Each spectrum consists of $49642$ samples over a range of mass-to-charge ratios
+between $800$ to $20000$ Dalton which we downsample on an equidistant grid of
+5000 points by linear interpolation. We then preprocess the downsampled
+spectra individually by first removing the slow-varying baseline intensity
+with the SNIP algorithm \citep{ryan1988snip} followed by a
+normalization with respect to the total ion count. Finally, we standardize
+the $5000$ predictors to have column-wise zero mean and unit variance in
+each dataset.
+
+We use the two datasets to illustrate how the sequential rank
+agreement can be applied in two different scenarios. In the first
+scenario we assess the agreement of four different statistical
+classification methods in how they rank the predictors according to
+their importance for distinguishing benign and malignant tumors. In
+the second scenario we assess the agreement among rankings of
+individual predicted risks of having a malignant tumor. The first
+scenario is relevant in the context of biomarker discovery and the
+latter is important e.g., when ranking patients according to immediacy of
+treatment.
+
+Four classification methods are considered: Random Forest
+\citep{breiman2001random} implemented in the R package
+\texttt{randomForest} \citep{liaw2002classification}, logistic Lasso
+\citep{tibshirani1996regression} and Ridge regression
+\citep{segerstedt1992ordinary} both implemented in the R package
+\texttt{glmnet} \citep{friedman2010regularization}, and Partial Least
+Squares Discriminant Analysis (PLS-DA) \citep{boulesteix2004pls}
+implemented in the R package \texttt{caret} \citep{Jed-Wing:2014aa}.
+%
+All four methods depend on a tuning parameter. The tuning parameter for
+Lasso and Ridge regression is the degree of penalization, and for PLS-DA
+it is the number of components (the dimensionality of the subspace).
+We estimate these separately for each sub-sample by a 20 times repeated
+5-fold cross-validation procedure. For the Random Forest we grow a fixed
+number of $5000$ trees and let the tuning parameter be the number of
+predictors randomly sampled at each split. We estimate this by a binary
+search with respect to minimizing the Out-of-Bag classification error estimate.
+
+In both scenarios we use the MALOVA data to train the statistical
+models, and in both situations the agreements are assessed with
+respect to perturbations of the training data in the following
+manner. We repeatedly draw a random sub-sample (without replication)
+consisting of 90\% of the MALOVA observations and train the four
+models on each sub-sample. We use 1000 iterations for the sub-sampling
+procedure.
+
+The implementation of Lasso and Ridge regression in the
+\texttt{glmnet} package offers three different cross-validated
+optimization criteria for the penalty parameter: total deviance,
+classification accuracy and area under ROC. We apply all three
+criteria to our data to investigate their effect on the agreements.
+Note also that the Lasso models produce incomplete lists depending on
+the value of the penalty parameter.
+
+\subsection{Agreement of predictor rankings}
+For each of the four methods, each of the 1000 models trained on
+the 1000 sub-samples of the MALOVA data produces a ranking of the 5000
+predictors according to their importance for discriminating between
+the tumor types. For the Random Forest classifier the predictors are ranked according
+to the Gini index, while for the logistic Lasso and Ridge regression
+models we order by absolute magnitude of the estimated
+regression coefficients. For the PLS-DA model the importance of the
+predictors is based on a weighted sum of the absolute coefficients
+where the weights are proportional to the reduction in the sums of
+squares across the components.
+
+The right panel of Figure \ref{fig:app1} shows the sequential rank
+agreement of the estimated importance of the 5000 predictors.  For
+clarity of presentation we zoom in on the agreement up to list depth
+600. At deeper list depths all agreement curves are approximately constant.
+%
+As expected, most of the sequential rank agreement curves start low,
+indicating good agreement, followed by an increase until they
+approximately become constant. This has the interpretation that the
+agreement across the different sub-samples is higher in the top as
+compared to the tail of the lists for all these classification
+methods. The changepoints where the curves become approximately
+constant are the list depths where the ranks of the remaining items
+become close to uniformly random.
+
+A not expected shape of the agreement curves is seen for the Ridge
+models for all three tuning criteria. They all show higher
+disagreement in the top of the lists followed by a decrease. The
+reason behind this behavior is rather subtle. Looking at the
+distribution of the absolute value of the regression coefficients we
+see that a large proportion of them are numerically very close to zero
+and have almost equal absolute value. This is a general feature of the
+Ridge models in this dataset and seen for all the 1000 trained models.
+This implies that when predictors are ranked according to the
+magnitude of their coefficients, their actual order becomes more
+uncertain and more close to a random permutation. This problem can be
+alleviated by truncating all predictors with absolute coefficient
+values below a given threshold thereby introducing an artificial
+incompletion of the lists.  For the Ridge models tuned with the
+deviance criterion, Figure \ref{fig:app2} (left) shows the the
+sequential rank agreement where for each of the 1000 trained models
+the predictors were artificially censored when their absolute
+coefficient value was lower than the $0.1\%$ quantile of the 5000
+absolute coefficient values. The curve was calculated using Algorithm
+\ref{sra-algorithm} with $B=1000$ and $P=5000$. The corresponding
+curve from Figure \ref{fig:app1} (right panel) is shown for
+comparison.  Even though the number of predictors with missing ranks is very
+small compared to the total number of predictors, the effect on the
+sequential rank agreement is substantial and with the artificial
+censoring the shape of the curves is as expected, starting low and
+then increasing.
+
+Looking at the agreement curves for the Lasso models in Figure
+\ref{fig:app1} (right) we clearly see the effect of the sparsity
+inducing penalization giving rise to incomplete lists. These curves
+were similarly calculated using \ref{sra-algorithm} and $1000$ random
+permutations. Under the deviance optimization criterion the median
+number of non-zero coefficients was 33 (range 16 to 50) and for the
+class accuracy criterion 14 (range 4 to 56). These values correspond
+to the list depths where the agreement curves become constant as a
+result of the subsequent censoring.
+
+\subsection{Agreement of individual risk predictions}
+\label{sec:airp}
+To assess the stability of the individual risk predictions we apply
+the predictors from the DACOVA dataset to each of the models. The
+predicted probabilities are then ranked in decreasing order such that
+the patients with the highest risk of a malignant tumor appears in the
+top of the list. Figure \ref{fig:app1} (left) shows the sra separately
+for each method, based on the 1000 risk predictions obtained from the
+models trained in the same $1000$ random sub-samples of the MALOVA
+data.
+
+Most curves start low and then increase indicating higher agreement
+among high risk patients. This is expected if we rank the individuals
+according to highest risk of disease. However, it is also expected
+that individuals with very low risk also show high agreement. In this
+case we order the patients according to (high) risk prediction but we
+could essentially also have reversed the order to identify the
+patients that have low risk prediction.
+
+An exception is the risk prediction agreement for the Lasso tuned with the AUC criterion which
+shows very low agreement among the high values of the predicted
+risks. The reason is that optimizing the penalty parameter with
+respect to the AUC criterion tends to favor a very high penalty value
+causing only a single predictor to be selected in each of the 1000
+iterations. This results in a lack of generalizability to the DACOVA
+data which gives rise to the higher disagreement in the predicted risks.
+In the extreme case where the penalty becomes so high
+that none of the predictors are selected by the Lasso, the sequential
+rank agreement for the predicted probabilities becomes undefined since
+all the ranks will be ties.
+
+Comparing the left and right panels of Figure \ref{fig:app1} it can
+further be seen that some of the methods show better agreement with
+respect to the predicted probabilities than for ranking the importance
+of the predictors and vice versa. Ridge regression shows higher agreement across training
+sets for the risk predictions than PLS-DA, and PLS-DA shows higher
+agreement for predictor importance than Ridge regression.
+
+Lasso shows similar agreement for ranking the risk predictions as
+PLS-DA (except for the AUC criterion), and poorer agreement for
+ranking predictors. This reason for the latter is the high
+auto-correlation between the intensities in the mass spectra which
+leads to collinearity issues in the regression models. It is
+well-known that variable selection with the Lasso does not
+perform very well when the predictors are highly correlated. The
+collinearity does, however, not affect the agreement of the risk
+predictions (Figure \ref{fig:app1}, left panel), since
+the specific variable selected is not that important for 
+% it is not so important which specific variable that gets selected from
+a group of highly correlated predictors when the purpose is risk
+predictions.
+
+It appears that Ridge regression tuned with the AUC criterion achieves
+the best performance with respect to the stability of ranking the
+individual predicted risk probabilities. It must, however, be stressed
+that the sequential rank agreement in this application is only
+concerned with the agreement of the risk predictions across
+sub-samples and not with the actual accuracy of the risk
+predictions. Thus, we also computed the AUC values for the different models
+based on the DACOVA data. The distributions across the $1000$
+sub-samples for a selection of the models is shown in the right panel
+of Figure \ref{fig:app2}.  Here we see that PLS-DA attains the highest
+AUC values with a median value of $0.70$ while the Ridge model with
+the AUC criterion attains a median AUC of $0.49$. This implies that
+while Ridge regression optimized with respect to the AUC criterion
+achieves the best sequential rank agreement, it performs similar to a
+random coin toss with respect to classifying the DACOVA patients.
+In practice both concerns are of importance.
+
+\section{Simulation study --- comparison of list agreements}
+
+We present results from a simulation study where we investigated the
+small sample properties of the sequential rank agreement and compared
+it to the topK method \citep{hall:schi:2012} with respect to list
+depth agreement.
+
+\added{The purpose of the simulations were twofold: First we want to
+  investigate the rank agreement as it changes with threshold $q$ and
+  number of lists $L$. Secondly, we want to compare the results (and
+  conclusions) from sra with the topK for a realistic situation where
+  the true, underlying agreement is not governed by a simple
+  probability distribution. Thus, we are interested in two features of
+  the methods: the depth until which they agree, and the number of
+  unique predictors found in that set.
+
+  To define the depth of agreement for sra we set a constant threshold
+  function to an integer $q$ and report the first crossing point,
+  i.e., the smallest list depth where sra exceeds $q$ (as implemented
+  in the \texttt{sra} function from package \texttt{SuperRanker} using
+  the median absolute distance argument). For topK we use the function
+  \texttt{j0.multi} which is implemented in the \texttt{R} package
+  \texttt{TopKLists}. Specifically, we set the tuning parameter
+  \texttt{v} of \texttt{j0.multi} to the value 6 and the window
+  parameter \texttt{d} to $q$ and report the output parameter
+  \texttt{maxK} as the depth of agreement. Thus, to make this
+  comparison we assume that the first crossing point of sra and the
+  result of the topK method measure the same underlying feature. }
+
+\added{The simulations should mimic a data analysis situation where we
+  have a single dataset and where important features are identified
+  (and ranked) using marginal $t$ tests.  We want to used agreement to
+  understand the stability of the observed feature selection.  } In
+each simulation run, we first generated an ``original'' dataset with
+1000 predictors and 400 observations. The predictors were drawn
+independently from a standard Gaussian distribution with variance 1
+such that
+$$E(y_i) = \sum_{j=1}^{15}x_{ij},$$
+where $y_i$ is the $i$th response, and $x_{ij}$ is the $j$th predictor
+for the $i$th measurement.
+
+For each ``original dataset'' we obtained $L$ ranked lists of the 1000
+predictors by drawing $L$ bootstrap samples (with replacement) of 400
+observations and then ranking the 1000 predictors according to their
+marginal $t$ test statistics. Thus, we assessed the depth of agreement
+among lists that are ranked with the same statistical method on
+bootstrap versions of the same dataset. We report results from two
+scenarios each based on 1000 simulated datasets:
+\begin{enumerate}
+\item[] Scenario I: Fix the number of lists \(L=8\) and vary the threshold  $q\in\{3,4,5,6,7,8,9,10\}$.
+\item[] Scenario II: Fix the threshold \(q=5\) and vary the lists, $L\in\{3,5,10,50\}$.
+\end{enumerate}
+In both scenarios we summarized the distribution of the estimated
+depth of agreement as well as the average number of unique predictors found
+in the set of predictors which is selected by the estimated depth of
+agreement.
+ % Asymptotic results are available for both methods as the number of
+ % lists increases, so this will investigate how much the results might
+ % change as $L$ increases. For a fixed constant threshold of $q_L=5$ we
+ % increase the number of available lists to investigate the sensitivity
+ % of the results to $L$.
+  % Throughout the simulations we
+  % assume a constant, prespecified threshold function, $q_L$, and try
+  % different values for the constant. The approach by
+  % \citet{hall:schi:2012} uses a window size parameter, $K$, to
+  % represent the maximum difference in rankings before they consider an
+  % item in two lists to be in agreement.  Because the standard
+  % deviation --- which is used in the sequential rank agreement
+  % definition of $\hat{A}_L$--- squares its differences, it tends to
+  % give more weight to larger rank differences and less weight to
+  % smaller differences compared to the mean absolute difference.  To
+  % ease comparison between the sequential rank agreement and the
+  % \citet{hall:schi:2012} approach we would like to use the same
+  % threshold for both methods, and in order to do this we use the
+  % median absolute deviance throughout the simulations in lieu of the
+  % pooled standard deviation.
+ % For all parameter combinations we ran 1000 simulations and report the
+ % distribution of the estimated depth of agreement and the average
+ % number of different predictors found at a given depth.
+The results from Scenario I are shown in the left panel of
+Figure~\ref{fig:example99}. The violin plots (with rectangular kernel)
+show the distributions of the estimated depths of agreement for both
+methods. As expected the depth of agreement increased when the
+threshold for agreement/window increased.
+
+\added{
+We see that sra results in a substantially lower depth of agreement
+than the topK method. Also the average numbers of unique predictors
+(bold numbers inside the plots) \added{which ideally should be 15 to
+  reflect the number of true underlying predictors} are markedly
+smaller --- and close to the true value --- for sra. Even larger
+differences were found when we used the Euclidean distance instead of
+the median absolute distance for the sequential rank agreement
+(results not shown).  The right panel of Figure~\ref{fig:example99}
+shows the results from Scenario II. The number of lists has little
+impact on the results, and again the sra is more conservative than
+topK and as a consequence sra includes fewer predictors in the
+selected set where the lists agree.
+
+Note that the effect sizes of the 15 predictors in the model is the
+same and in practice we observe that the majority of the 15 predictors
+are generally picked in each sample but that their individual rankings
+vary substantially in the top 15 within each bootstrap sample. If the
+number of influential predictors is lessened then the variance in
+depth estimation and number of predictors is reduced.}
+
+\section{Discussion}
+In this article we address the problem of comparing ranked lists of
+the same items. Our proposed method can handle both the situation
+where the underlying data to generate the ranked lists are available
+and the situation where the only available data is the actual ranked
+lists. In addition, incomplete ranked lists where only the ranks of the
+top $k$ ranked items are known can be accommodated as well. The
+proposed agreement measure can be interpreted as the average distance
+between an item's rank and the average rank assigned to that item
+across lists.
+
+The sequential rank agreement can be used to determine the depth at
+which the rank agreement becomes too large to be desirable based on
+prior requirements or acceptable differences, or it can be used to
+visually determine when the change in agreement becomes too large.  In
+that regard the investigator can have prior limits on the level of
+agreement that is acceptable.
+
+We have shown that sra is very versatile: it can be used not only to
+compare ranked lists of items produced from different
+samples/populations but that it also can be used to study the ranks
+obtained from different analysis methods on the same data, and to
+evaluate the stability of the ranks by bootstrapping (or sub-sampling)
+the data repeatedly and comparing the ranks obtained from training the
+models on the bootstrapped data.
+
+While the sequential rank agreement is primarily an exploratory tool
+we have suggested two null hypotheses that can be used to evaluate the
+sequential rank agreement obtained. Note that none of the two null
+hypotheses are concerned with the actual ``true ranking'' but are
+purely concerned with consistency/stability of the rankings among the
+lists, and consequently we cannot determine if the rankings are good but only
+whether they agree. The sequential rank agreement curve can be
+compared visually to the curves obtained under either of the null
+distributions and simple point-wise $p$-values can be obtained for
+each depth by counting the number of sequential rank agreements under
+the null hypothesis that is less than or equal to the observed rank
+agreement.
+% Two simple extensions can be pursued to make more formal
+% uniform tests: one would be to use a Kolmogorov-Smirnov-like test and
+% use the largest difference between the variance-weighted sequential
+% rank agreement curve and the mean null rank agreement curve as a test
+% statistic. Alternatively, a changepoint analysis could be made on the
+% sequential rank agreement in order to determine the depths at which
+% there are ``jumps'' in the rank agreement. These jumps would
+% correspond to depths for which the agreement among the lists was
+% substantially worse and could serve as indicators for when the lists
+% no longer agree sufficiently satisfactory.
+
+Finally, we have --- whenever possible --- used all available ranks
+from the lists.  We could choose to restrict attention to the rank of
+items which show evidence for significance in their models. That would
+ensure that there would be put less emphasis on the agreement of the
+non-significant items and it would be easier to identify a change in
+agreement among the items that were deemed to be relevant. In our
+application section we have successfully introduced such an artificial
+censoring for the predictor rankings obtained with ridge regression.
+
+We note that the sequential rank agreement is still marred by problems
+that generally apply to ranking of items and/or
+individuals. Collinearity in particular can be a huge problem when
+bootstrapping data or when comparing different analysis methods. For
+example, marginal analyses where each item is analyzed separately will
+assign similar ranks to two highly correlated predictors while methods
+that provide a sparse solution such as the Lasso will just rank one of
+the two predictors high while the other might have a very low rank.
+Thus in such a scenario we would expect low agreement of the rankings
+from Lasso and marginal analyses simply because of the way correlated
+predictors are handled. This is not a shortcoming of the sequential
+rank agreement but is a problem general to all ranked lists.
+
+Another caveat with the way the sequential rank agreement is defined
+is the use of the standard deviation to measure agreement. The
+standard deviation is an integral part of the limits-of-agreement as
+discussed by \citet{alt:bland:1983}. However, the standard deviation
+can also be unstable when the number of observations is low and
+alternative measures such as the median absolute deviance may prove more
+stable in some situations. However, the current definition using the
+standard deviation is analogous to the approach used for agreement in
+method comparison studies so we have used that.
+
+In conclusion we have introduced a method for evaluation of ranked
+(partial/censored) lists that can be easily interpreted and that can
+be applied to a large number of situations.  The method presented here
+can be adapted further by using it to compare and classify statistical
+analysis methods that agree on the rankings they provide or by using
+the rank agreement to optimize a hyper-parameter in, say, elastic net
+regularized regression where the rank agreement is used to determine
+the mixing proportion between the $L_1$ and the $L_2$ penalty.
+\added{Finally, the proposed method may be adapted (with some
+  additional assumptions) to the situation where there are put equal
+  emphasis on both ends of the lists and not just on the top of the
+  lists.}
+%Finally, the proposed method may be adapted (with some additional
+%assumptions) to the situation where the lists are subsets of each
+%other.
+We will be investigating these extensions further in the future.
+
+\section{Supplementary Materials}
+\label{sec5}
+
+The reader is referred to the on-line Supplementary Materials for
+technical appendices and proofs.
+
+%  \appendix
+
+% \newpage
+% \section{Proof of Theorem \ref{thm:consistency}}
+% \label{sec:appA}
+
+% We start by defining the following cadlag function
+% \begin{align}
+%   N(p; d) = \sum_{s=1}^p 1\left(Q(R(X_s) \leq d) > 0\right), \quad p = 1,\ldots, P
+% \end{align}
+% which runs through the list elements in an arbitrary order and counts
+% how many list elements that have a strictly positive probability to have
+% rank less than or equal to $d$ under the probability measure $Q$.
+% Note that the cardinality of the set $S(d)$ in equation
+% (\ref{eq:sumset}) is equal to the counting process evaluated at the last element: $|S(d)| = N(P;d)$.
+% The empirical counterpart of the counting process is given by
+% \begin{align}
+%   \widehat{N}_L(p; d) &= \sum_{s=1}^p 1\left( \frac{1}{L} \sum_{l=1}^L 1(R_l(X_s) \leq d) > 0\right)\\
+%         &= \sum_{s=1}^p 1\left(\widehat{Q}_L(R(X_s) \leq d) > 0\right)\nn.
+% \end{align}
+% The joint law of the sequence
+% $\{\widehat{N}_L(p;d)\}_{1\leq p \leq P}$ is completely determined by
+% the finite set of the jump times of $N(p;d)$. Since
+% $\{1(R_l(X_s) \leq d); l=1,\dots, L\}$ consists of independent and
+% identically Bernoulli distributed variables with expectation
+% $Q(R(X_s) \leq d)$ and finite variance it follows from the law of large numbers that
+% $\widehat{Q}_L(R(X_p) \leq d) \overset{P}{\longrightarrow} Q(R(X_p)
+% \leq d)$ for every $p$ and $d$ as $L \rightarrow \infty$. 
+% Therefore, we have
+% \begin{align}
+%   \sup_{d \in 1,\ldots,P}\sup_{p \in 1,\ldots, P} \left|\widehat{N}_L(p; d) - N(p; d)\right| = o_P(1).\label{eq:countingConv}
+% \end{align}
+
+% The sequential rank agreement given in equation
+% (\ref{def:sra}) may therefore be rewritten as the following integral
+% with respect to $N$
+% \begin{align}
+%   \textrm{sra}(d) = \int_{1}^P \frac{A(X_p)}{N(P; d)}\mathrm dN(p; d).
+% \end{align}
+% The empirical sequential rank agreement is similarly given by
+% \begin{align}
+%   \widehat{\textrm{sra}}_L(d) &= \int_1^P \frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)}\mathrm d \widehat{N}_L(p; d) 
+% \end{align}
+% and it follows that
+% \begin{align}
+%   \sup_{d \in 1,\ldots,P}\left|\widehat{\textrm{sra}}_L(d) - \textrm{sra}(d)\right| \leq \sup_{d \in 1,\ldots,P}\left|U_L(d)\right| - \sup_{d \in 1,\ldots,P}\left|V_L(d)\right|
+% \end{align}
+% where
+% \begin{align}
+%   U_L(d) &= \int_1^P\frac{A(X_p)}{N(P;d)} \mathrm d\left(\widehat{N}_L(p; d) - N(p; d)\right)\\
+%   V_L(d) &=  \int_1^P \left(\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right)\mathrm d\widehat{N}_L(p; d)
+% \end{align}
+% The conclusion of the proof therefore follows if each of these two
+% terms are uniformly of order $o_P(1)$ in $d$.  For the first term we
+% have that
+% \begin{align}
+%   \left|U_L(d)\right| \leq \left(\sup_{p \in 1,\ldots,P}\left|\frac{A(X_p)}{N(P;d)}\right|\right)\left(\widehat{N}_L(p; d) - N(p; d)\right)\biggr\rvert_1^P
+% \end{align}
+% where the second factor is $o_P(1)$ uniformly in $d$ by equation
+% (\ref{eq:countingConv}). The first factor is $O(1)$ since $N(P;d)>0$
+% uniformly in $d$ and because $A(X_p) = O(P^2)$ uniformly in $p$ by the Cauchy-Schwarz inequality. It
+% thus follows that $\left|U_L(d)\right| = o_P(1)$ uniformly in $d$.
+
+% Similarly we derive an upper bound for $V_L(d)$ by
+% \begin{align}
+%  V_L(d) &\leq \sup_{p \in 1,\ldots,P}\left|\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right| \int_1^P \left|\mathrm d\widehat{N}_L(p; d)\right|\label{eq:secondTerm}
+% \end{align}
+% and it follows directly from the law of large numbers that
+% $\widehat{A_L}(X_p) = A(X_p) + o_P(1)$ uniformly in $p$ since
+% $R_i(X_p)$ in equation (\ref{eq:empDistance}) for $i=1,\ldots,L$ are
+% independent and $Q$-identically distributed random variables with
+% finite second moment. Again, $\widehat{N}_L(P; d) = N(P;d) + o_P(1)$ uniformly in $d$ by equation
+% (\ref{eq:countingConv}) and $N(P;d)>0$ so by the continuous mapping theorem the first factor
+% in equation (\ref{eq:secondTerm}) is of order $o_P(1)$ uniformly in
+% $p$ and $d$. The result then follows by noting that the second factor
+% is bounded by $P$ uniformly in $d$.
+
+% \section{Proof of Corollary \ref{cor:consistency}}
+% \label{sec:appB}
+% Recall that $\mathcal{L}$ is the superlevel set of list positions
+% where the sequential rank agreement exceeds the threshold function
+% $q$. Let $A \triangle B = (A \setminus B) \cup (B \setminus A)$ be the
+% symmetric difference between sets $A$ and $B$. It is then sufficient
+% to show that $\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q) \overset{P}{\longrightarrow} \emptyset$
+% for $L \rightarrow \infty$ where $\emptyset$ denotes the empty set.
+% We have that
+% \begin{align}
+%   \plim_{L \rightarrow \infty}\left(\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q)\right) &=   \left(\plim_{L \rightarrow \infty}\widehat{\mathcal{L}}_L(\widehat{q}_L)\right) \triangle \mathcal{L}(q)\\
+%   &= \left\{d : \plim_{L\rightarrow \infty} \left(\widehat{\textrm{sra}}_L(d) - \widehat{q}_L(d)\right)\geq 0\right\} \triangle \mathcal{L}(q)\nn\\
+%   &= \left\{d : \textrm{sra}(d) - q(d) \geq 0\right\}\triangle \mathcal{L}(q)\nn\\
+%   &= \mathcal{L}(q)\triangle \mathcal{L}(q)\nn\\
+%   &= \emptyset\nn
+% \end{align}
+% as a consequence of Theorem \ref{thm:consistency}, the assumption that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ and the continuous mapping theorem. This completes the proof.
+
+%\bibliographystyle{rss}
+\bibliographystyle{biorefs}
+\bibliography{paperref}
+
+\begin{figure}[htbp]
+   \begin{center}
+ \includegraphics[width=.49\textwidth]{newfig1a}
+ \includegraphics[width=.49\textwidth]{newfig1b}
+% \includegraphics[width=.49\textwidth]{paper-fig1b}
+% \includegraphics[width=.49\textwidth]{paper-fig2b}
+ \end{center}
+ \caption{Left panel: Sequential rank agreement for 4 different
+   analysis methods applied to the 3051 genes in the Golub data (black
+   line). Right panel: Corresponding sequential rank agreement for the
+   same data but where only the top 20 ranked items are available and
+   the rank of the remaining items are not available. The blue and red
+   areas correspond to the independent and randomized reference
+   hypothesis areas, respectively. Note that both the $x$ and $y$ axes
+   are shown on the log scale to ``zoom in'' on the top of the lists.}
+ \label{fig:example1}
+ \end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=.49\textwidth]{pics/riskAgreementPlot}%
+\includegraphics[width=.49\textwidth]{pics/predictorAgreementPlot}
+\end{center}
+\caption{Left panel: Sequential rank agreement of 1000 rankings of the
+  predicted risks of malignant tumor. For each method the different
+  rankings were obtained by first training models in 1000 random
+  sub-samples of the MALOVA data and then predicting the risk of
+  malignant tumor in the 113 DACOVA patients. Right panel: Sequential
+  rank agreement of 1000 rankings of the 5000 predictors. The rankings
+  were obtained from the same 1000 trained models.}
+ \label{fig:app1}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=.49\textwidth]{pics/ridgeCensoring}%
+\includegraphics[width=.49\textwidth]{pics/aucPlot}
+\end{center}
+\caption{Left panel: Sequential rank agreement for Ridge regression
+  obtained by artificially censoring predictor ranks when their absolute
+  coefficient values are lower than the 0.1\% quantile. Right panel:
+  Box plots of AUC values across the $1000$ sub-samples with respect
+  to the known class labels of the DACOVA data.}
+ \label{fig:app2}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+   \begin{center}
+ \includegraphics[width=.49\textwidth]{Ksim.pdf}
+ \includegraphics[width=.49\textwidth]{Lsim}
+ \end{center}
+ \caption{Left panel: Simulation study showing distribution of
+   estimated rank agreements for sra and topK for varying thresholds
+   and fixed number of lists $L=8$. The bold numbers are the average
+   number of unique predictors included in the set where the lists
+   agree. Right panel: Simulation results for varying number of lists
+   and with fixed threshold of $q=5$.}
+ \label{fig:example99}
+\end{figure}
+
+\newpage
+
 \begin{table}[tb]
@@
     each item in each of the three lists and (c) shows the cumulative
-    set of items up to a given depth in the three lists.}
+    set of items up to a given depth in the three lists when $\varepsilon=0$ (i.e., an item is added to $S(d)$ whenever it appears in at least one list).}
 \begin{center}
@@
 
-The agreement among the lists regarding the rank given to an item $X_p$
-can be measured by the variance across the lists
-\begin{align}
-  A(X_p) &= \E_Q\left[\left(R(X_p)-\E_Q R(X_p)\right)^2\right]\\
-   &= \sum_{r\in\Pi} \left(r(X_p)-\E_Q R(X_p)\right)^2 Q(r),\nn
-\end{align}
-where $\Pi$ is the set of all permutations of $X$, $Q$ is a
-probability mass function on $\Pi$, and
-$\E_Q R(X_p)=\sum_{r\in\Pi} r(X_p)Q(r)$. The empirical counterpart is
-\begin{align}
-\widehat{A}_L(X_p) = \frac{1}{L-1}\sum_{i=1}^L (R_i(X_p) - \overline{R}_L(X_p))^2, \quad \overline{R}_L(X_p) = \frac{1}{L}\sum_{i=1}^L R_i(X_p).\label{eq:empDistance}
-\end{align}
-For each item, the function $\widehat{A}_L$ has an interpretation as the expected
-Euclidean distance of the individual rankings from the expected
-ranking over the $L$ lists, and it corresponds to the same measure
-that is used in method comparison studies to compute the limits of
-agreement \citep{alt:bland:1983}.
-
-For an integer $1\le d\le P$ we define the expected set of unique
-items found by merging the first $d$ elements across the possible
-lists:
-\begin{align}
-S(d) &= \left\{X_p; \left(\sum_{r\in\Pi} 1\left(r(X_p)\le d\right) Q(r)\right) >0\right\}\label{eq:sumset}
-\end{align}
-where $1(\cdot)$ denotes the indicator function.
-The empirical counterpart is the set of unique items ranked less than or
-equal to $d$ in any of the $L$ lists:
-\begin{equation}
-\widehat{S}_{L}(d) = \left\{X_p; \left(\frac{1}{L}\sum_{l=1}^L 1\left(R_l(X_p)\le d\right)\right)>0\right\},
-\end{equation}
-which is exemplified in Panel (c) of Table \ref{tab:example}.
-
-We define the \emph{sequential rank agreement} as the weighted
-expected agreement of the items found in the set $S(d)$:
-\begin{align}
-  \textrm{sra}(d) = \frac{1}{|S(d)|}\sum_{p \in S(d)}A(X_p)\label{def:sra}
-\end{align}
-where $|S(d)|$ is the cardinality of the set $S(d)$.
-The empirical counterpart is equivalently given by
-\begin{equation}
-  \widehat{\textrm{sra}}_L(d) = \frac{\sum_{\{p \in \widehat{S}_{L}(d)\}}(L-1)\widehat{A}_L(X_p)}{(L-1)|\widehat {S}_{L}(d)|}\label{def:sraEst}
-\end{equation}
-
-Values of \textrm{sra} close to zero suggests that the lists agree on
-the rankings while larger values suggest disagreement.  Just like in
-method comparison studies we prefer smaller limits of agreement
-because that means that the differences between the rankings is
-small. If the ranked lists are identical then the value of sequential
-rank agreement will be zero for all position or depths $d$ down the
-lists.  The sequential rank agreement is equivalent to the pooled
-variance of the items found in $S(d)$, and it can be interpreted as
-the average distance of the individual rankings of the lists from the
-average ranking for each of the items we have seen until depth $d$.
-For an easier interpretation we suggest to plot
-$\sqrt{\widehat{\textrm{sra}}_L(d)}$ corresponding to the pooled
-standard deviation.
-
-\subsection{Agreement among fully observed lists}
-\label{sec:amfol}
-
-The simplest case occurs when all $L$ lists are fully observed, \ie, when
-we observe the rank of all $P$ items for all $L$ lists. Fully
-observed ranked lists are common and arise, for example, when
-different statistical analysis methods are applied to a single dataset
-to produce lists of predictors ranked according to their importance,
-or if the same analysis method is applied to data from different
-populations.
-
-With fully observed lists we can plot the sequential rank agreement
-\eqref{def:sra} as a function of depth $d$. An example is seen in the top
-panels of Figure~\ref{fig:example1} where four analysis methods were
-used to rank 3051 gene expression values measured on 38 tumor mRNA
-samples in order to classify between acute lymphoblastic leukemia
-(ALL) and acute myeloid leukemia (AML)
-\citep{Golub1999}. Preprocessing of the gene expression data was done
-as described in \citet{Dudoit2002} and the four different analysis
-approaches were: marginal unequal variances two-sample $t$ tests,
-marginal logistic regression analyses, elastic net logistic regression
-\citep{friedman2010regularization}, and marginal maximum information
-content correlations (MIC) \citep{Reshef2011}. For the first two
-methods, the genes were ranked according to minimum $p$ value, for
-logistic regression the genes were ordered by size of the
-corresponding coefficients (after standardization), and MIC was
-ordered by absolute correlation which resulted in the top rankings
-seen in Table~\ref{tab1}.
-
-The sequential rank agreement curve shown in the two left plots of
-Figure~\ref{fig:example1} roughly indicates the average distance in
-ranks for the genes found among the first $d$ positions in the $4$
-lists.  It can be seen that the sequential rank agreement limits are
-better towards the top of the lists (smaller values on the $y$ axis
-corresponds to \emph{better} agreement) than towards the bottom of the
-lists. It is clear from the curve in Figure~\ref{fig:example1} that
-there is a substantial drop in agreement (higher sra) even after the
-first depth. Thus, if we were to restrict attention to a small set of
-predictors then we would focus on predictors 2124, 829, and 378 as
-seen in Table~\ref{tab1}.
-
-% latex table generated in R 3.2.3 by xtable 1.8-2 package
-% Fri Apr  8 10:22:58 2016
+\newpage
+
 \begin{table}[tb]
 \centering
-\caption{List of ranked results from the Golub data. Numbers indicate the predictor/gene for the given ranking and method. Only the top 10 ranks are shown in the table.} 
+\caption{List of ranked results from the Golub data. Numbers indicate
+  the predictor/gene for the given ranking and method. Only the top 10
+  ranks are shown in the table. The ranked lists appear to agree that
+  genes 2124 and 829 are among the most interesting while the highest
+  ranked gene from MIC, gene 378, is not found in the top 10 for two
+  of the other methods.}
 \label{tab1}
@@
 1 & 2124 & 2124 & 829 & 378 \\ 
-  2 & 896 & 896 & 2124 & 829 \\ 
-  3 & 2600 & 829 & 2198 & 896 \\ 
-  4 & 766 & 394 & 1665 & 1037 \\ 
-  5 & 829 & 766 & 1920 & 2124 \\ 
-  6 & 2851 & 2670 & 1042 & 808 \\ 
-  7 & 703 & 2939 & 808 & 108 \\ 
-  8 & 2386 & 2386 & 849 & 515 \\ 
-  9 & 2645 & 1834 & 937 & 2670 \\ 
-  10 & 2002 & 378 & 1995 & 2600 \\ 
+  2 & 896 & 896 & 2198 & 829 \\ 
+  3 & 2600 & 829 & 2124 & 896 \\ 
+  4 & 766 & 394 & 808 & 1037 \\ 
+  5 & 829 & 766 & 1665 & 2124 \\ 
+  6 & 2851 & 2670 & 1920 & 808 \\ 
+  7 & 703 & 2939 & 1389 & 108 \\ 
+  8 & 2386 & 2386 & 1767 & 515 \\ 
+  9 & 2645 & 1834 & 1042 & 2670 \\ 
+  10 & 2002 & 378 & 2600 & 2600 \\ 
    \hline
@@
 
-% chunks 2,3,4,5,6,7
-
-\begin{figure}[htbp]
-   \begin{center}
- \includegraphics[width=.49\textwidth]{paper-fig1a}
- \includegraphics[width=.49\textwidth]{paper-fig2a}
- \includegraphics[width=.49\textwidth]{paper-fig1b}
- \includegraphics[width=.49\textwidth]{paper-fig2b}
- \end{center}
- \caption{Left panels: Sequential rank agreement for 4 different
-   analysis methods applied to the 3051 genes in the Golub data (black
-   line). Right panels: Corresponding sequential rank agreement for
-   the same data but where only the top 20 ranked items are available
-   and the rank of the remaining items are not available. The blue and
-   red areas correspond to the independent and randomized reference
-   hypothesis areas, respectively. The bottom plots are identical to
-   the corresponding top plots but have been zoomed in on the first
-   part of the $x$ axis.}
- \label{fig:example1}
- \end{figure}
-
-% If the lists agree on all items then the sequential rank agreement is
-% zero for all depths. 
- Generally, if there are changes in the level of
- rank agreement then this suggests that there are sets of items that
- are ranked similarly in all lists and other sets of items that have
- been assigned vastly different ranks in the lists. For example, in gene
- association studies we generally expect the agreement among the lists
- to be better towards the top of the lists and worse towards the end
- of the lists. In this case the sequential rank agreement curve will
- start at a low level and then increase until it levels off as seen in
- Figure~\ref{fig:example1}.
-
-\subsection{Analysis of incomplete lists}
-Incomplete or partial lists are a common occurrence that arise, for
-example, in case of missing data (items), when comparing top $d$ list
-results from publications, or when some methods only rank a subset of
-the items. For example, penalized regression based on the Lasso
-provides a sparse set of predictors that have non-zero
-coefficients. There is no obvious ordering of the set of predictors
-whose coefficient has been shrunk to zero and thus we end up with a
-partial ordering.  Incomplete lists also occur if the analyst
-restricts attention to the ranks of items that have been found to be
-statistically significant.
-
-Sequential rank agreement can be generalized to incomplete lists in
-the following way. Let $\Lambda_l\subset X$ be the subset of $d_l$
-items that have been ranked highest in list $l$.  The case where all
-lists are incomplete at the same depth $d$ corresponds to
-$d_1=\cdots=d_L=d$. For incomplete lists the rank function becomes
-\begin{equation}
-	\widetilde R_l(X_p) = \begin{cases}
-	     \{R_l(X_p)\} & \text{for } X_p\in \Lambda_l\\
-	     \{d_l+1,\dots,P\} & \text{for} X_p \not\in \Lambda_l
-        \end{cases}
-\end{equation}
-where we only know that the rank for the unobserved items in list $l$
-must be larger than the largest rank observed in that list.
-
-The agreement, $A(X_p)$, cannot be computed directly for all
-predictors in the presence of incomplete lists because the exact rank
-for some items will be unknown. Also, recall that the rankings within
-a single list are not independent since each rank must appear exactly
-once in each list. Thus, we cannot simply assign the same number
-(\eg, the mean of the unassigned ranks) to the missing items since
-that would result in \emph{less} variation of the ranks and hence less
-variation of the agreement, and it would artificially introduce a
-(downward) bias of agreement for items that are missing in multiple
-lists.
-
-Instead we randomize the ranks $\{d_{l}+1,\dots,P\}$ to the items that
-do not occur in list $\Lambda_l$. One realization of the $L$ rankings
-of the set $X$ is obtained by randomizing the missing items of each
-list. By randomizing a large number of times we can compute
-\eqref{def:sra} for each realization, and then compute the sequential
-rank agreement as the pointwise (for each depth) average of the rank
-agreements. The algorithm is described in detail in
-Algorithm~\ref{sra-algorithm}.
+\newpage
 
@@
 
-The proposed approach is based on two assumptions: 1) that the most
-interesting items are found in the top of the lists, and 2) that the
-ranks that are missing from the lists provide so little information that it is reasonable
-to assume that they can be represented by a random order. The first
-assumption is justifiable because we have already accepted that it is
-reasonable to rank the items in the first place. The second assumption
-is fair in the light of the first assumption provided that we have a
-``sufficiently large'' part of the top of the lists available.
-
-When the two assumptions are satisfied then it is clear that the
-interesting part of the sequential rank agreement curves is restricted
-to the depths where the number of items without ranks available is
-low. Generally, without additional prior knowledge about the
-underlying distribution of the ranks it seems reasonable to restrict
-attention of the sequential rank agreement to lists no longer than the
-longest of the $L$ lists examined.
-
-Like for fully observed lists we generally expect the sequential rank
-agreement to start low and then increase unless the lists are
-completely unrelated (in which case the sequential rank agreement will
-be constant at a high level) or if the lists mostly agree on the
-ranking (in which case the sequential rank agreement will also be
-constant but at a low level). For incomplete ranked lists we also
-expect a changepoint around the depth where the lists are
-become incomplete. This is an artefact stemming from the fact that we assume
-that the remainder of the lists can be replaced by a simple
-permutation of the missing items.
-
-The two right-most plots of Figure~\ref{fig:example1} show the impact
-of restricting the Golub data such that only top-20 lists are
-available instead of full lists of length 1000 (20 was chosen to
-resemble the list lengths that might be published in a
-manuscript). The sequential rank agreement increases much quicker
-because the incomplete lists introduce more noise in the estimation of
-the agreement, but it is still possible to see that the top of the
-list has a sequantial rank agreement that is not substantially
-different from the full lists.
-
-% \subsection{Comparison to average overlap}
-% The average overlap is a measure that is widely used in comparisons of
-% two fully observed ranked lists \citep{Fagin2003,Webber2010} and the
-% idea behind the average overlap closely resembles the sequential rank
-% agreement. In this section we present an extension of the average
-% overlap to more that two lists and compare this to sequential rank
-% agreement.
-
-% The overlap among $L$ lists observed until depth $d$ is defined as the
-% proportion of items that are found in all $L$ lists compared to the
-% possible number of items found in all lists observed until depth $d$:
-% \begin{align}
-%   O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}\label{def:overlap}
-% \end{align}
-% Typically the overlap is only considered for pairwise comparisons
-% \citep{Bar-Ilan2006,Boulesteix2009} but formula \ref{def:overlap}
-% is easily extended to the situation where $L>2$ as shown above.
-
-% The average overlap at depth $d$ is defined as the average of the
-% overlaps until depth $d$,
-% \begin{equation}
-% AO(d) = \frac1d\sum_{i=1}^d O(d).
-% \end{equation}
-% This definition directly ensures that further emphasis is put on items
-% in the top of the lists since their item specific overlap contributes
-% to the calculation of average overlap at higher depths.
-
-% There are three main drawbacks of the average overlap method: First it
-% is highly sensitive to single items in the top of one the lists.  The
-% average overlap profile may change dramatically depending on the
-% presence or absence of similar items in the beginning of the
-% lists. For example, if all lists have the same item at the top then
-% the average overlap attains its maximum at 1, but if just one of the
-% lists differ then the overage overlap starts at its minimum of
-% 0. Secondly, when the number of lists, $L$, is high then it might be
-% difficult to obtain a non-zero overlap (and hence a non-zero average
-% overlap) at the beginning of the lists because the overlap will be
-% zero until an item is present in all $L$ lists. Finally, the average
-% overlap has an interpretation in terms of moving averages of
-% percentages (of $L$ sets) which is somewhat less intuitive than our
-% proposed sequential rank distance.
-
-% In the following we compare the average overlap and the sequential
-% rank agreement using the analysis of the Golub data
-% \citep{Golub1999}. We consider a typical application of the average
-% overlap and thus restrict attention to only two gene rankings obtained
-% by the $t$ tests and logistic regression. These two marginal modeling
-% approaches should result in very similar gene rankings. The first two
-% columns of Table~\ref{tab1} show the actual top-10 ranking of the
-% predictors.
-
-% % chunks 8,9
-
-% \begin{figure}[htbp]
-% \begin{center}
-% \includegraphics[width=.49\textwidth]{paper-fig-ao1}
-% \includegraphics[width=.49\textwidth]{paper-fig-ao2}
-% \end{center}
-%  \caption{Average overlap (left figure) and sequential rank agreement
-%  (right figure) for comparing two different analysis methods
-%  (marginal $t$ test and logistic regression)
-%     applied to the Golub data. The black lines are based on the full
-%     set of predictors while the red lines have the two highest
-%     associated predictors removed
-%     from the data before computing the average overlap and sequential
-%     rank agreement.}
-%   \label{fig:case1}
-%  \end{figure}
-
-%  The black lines in Figure~\ref{fig:case1} show the average overlap
-%  and sequential rank agreement for these data. It is clear from both
-%  plots that there is perfect agreement towards the top of the ranked
-%  lists: the average overlap is 1 and the sequential rank agreement is
-%  zero.  However, if we remove those two items from the data (i.e.,
-%  gene/predictor 2124 and 896) and redo the analyses then we get
-%  substantially different curves for the average overlap but not for
-%  sequential rank agreement (red lines in Figure~\ref{fig:case1}). For
-%  the sequential rank agreement we get roughly the same estimate of
-%  agreement from rank 3 as we did from the full dataset. This indicates
-%  that sequential rank is more robust against small perturbations of
-%  the data. For the average overlap the curves completely change when
-%  these two items are removed and which would lead to a different
-%  conclusion about the agreement among the items from rank 3 downwards.
-
-\section{Evaluating sequential rank agreement}
-To interpret the sequential rank agreement values we propose two
-different benchmark values corresponding to two different
-hypotheses. We wish to determine if we observe better agreement than
-what would be expected if there were no relevant information available
-in the data.
-
-The first reference hypothesis is
-\begin{eqnarray}
-H_0  & : &  \text{The list rankings correspond to completely randomly}\label{eq:permutationHypothesis}\\
-       &  & \text{permuted lists}\nn
-\end{eqnarray}
-which not only assumes that there is no information in the data on
-which the rankings are based but also that the methods used to provide
-the rankings are completely independent.
-
-Alternatively, we can remove the restriction on the independence among
-the methods used to generate the individuals rankings in the $L$ lists
-by only requiring that there is no information contained in the ranking but that
-the rankings are all based on applying the method/approaches to the
-same data
-\begin{eqnarray*}
-\widetilde H_0 & :&  \text{The list rankings are based on data that contain}\\
-& &   \text{no association to the outcome.}
-\end{eqnarray*}
-
-This alternative null hypothesis addresses the fact that some ranking
-methods are more likely to provide similar rankings of the same data
-because the ranking methods focus on the same features of the data
-rather than because of any information contained in the data.
-
-\subsection{Permutation-based inference}
-
-$H_0$ is a quite unrealistic null hypothesis but we can easily obtain
-realizations from that null hypothesis simply by permuting the items
-within each list and then computing the sequential rank agreement for
-the permuted lists. In the fully observed case each experiment
-contains $L$ lists of random permutations of the items in $X$. For the
-incomplete case we first permute the items $X_1,\dots,X_P$ and then assign missing ranks
-for list $l$ from  $d_l$ to $P$ (\ie, each list so it has the same number of
-observed rankings as was observed for list $l$ in the original
-dataset). The sequential rank agreement curve from the original
-lists can then be compared to, say, the pointwise 95\% quantiles of
-the observed rank agreements obtained under $H_0$.
-
-To obtain the distribution under $\widetilde H_0$ the idea is to
-repeat the ranking procedures for unassociated data many times.  Thus,
-we first permute the outcome variable of the dataset on which the
-rankings are based. This removes any association between the predictor
-variables and the outcome. Then we apply the methods to the permuted
-dataset to generate $L$ new rankings and subsequently compute the
-sequential rank agreement. Note that we only permute the outcomes and
-preserve the structure of the candidate predictors. This randomization
-approach requires that we have the original data available and as such
-it may not be possible to evaluate $\widetilde H_0$ in all situations.
-
-If the sequential rank agreement for the original data lies
-substantially below the distribution of the sequential rank agreements
-obtained under either $H_0$ or $\widetilde H_0$ then this suggests that
-the original ranked lists agree \emph{more} than what we would expect
-in data with no information, and therefore that the information in the
-lists is significantly more in agreement than what would be expected.
-
-Figure~\ref{fig:example1} shows the empirical distributions of
-sequential rank agreement under $H_0$ and $\widetilde H_0$ each based
-on $400$ permutations of the Golub data described in Section
-\ref{sec:amfol}. Not surprisingly, the sequential rank agreement under
-$\widetilde H_0$ is lower than the sequential rank agreement under
-$H_0$ because the four methods used to rank the data ($t$ test,
-logistic regression, elastic net, and MIC) generally tend to identify
-(and rank) similar predictors even if there are only spurious
-associations. The two bottom panels in Figure~\ref{fig:example1} also
-indicate that the observed sequential rank agreement (the black line)
-is better than what would be expected by chance for data that contain
-no information since it lies below the reference areas. The incomplete
-data also suggests that there may be at most 1 or 2 ranked items
-towards the top of the lists that yield a result better than what
-would be expected (the bottom-right plot).
-
-It is important to stress that neither $H_0$ nor $\widetilde H_0$ are
-related to questions regarding the association between the outcome and
-the predictors in the dataset. Both hypotheses are purely considering
-how the rankings agree in a situation where there is no relevant
-information available in the data used for creating the rankings.
-
-\subsection{Asymptotic inference of change in agreement}\label{sec:chgpoint}
-In many applications it is of interest to estimate a list depth which
-satisfies a changepoint criterion since that corresponds to a change
-in agreement among the list ranks. In this section we investigate the
-theoretical properties of our proposed method for this specific
-task. As in \cite{hall:schi:2012} we consider an infinite set
-of lists and study the asymptotic behaviour for $L\to\infty$.
-However, we do not allow the length of the lists to change with $L$
-because in most applications the lengths of the lists are fixed.
-
-We start by showing that $\widehat{\textrm{sra}}_L$ is a consistent
-estimator of $\textrm{sra}$ for $L \rightarrow \infty$. The result is
-stated in Theorem \ref{thm:consistency}.
-
-% below and requires the
-% following assumptions.
-% \begin{assumption}
-% We assume that:
-% \begin{enumerate}[label=\Alph*)]
-% \item{The $L$ lists are independent and identically distributed
-    % samples from a probability distribution $Q$ on the set of lists.}
-% \item{$\E_Q[R(X_p)^2] < \infty$ uniformly in $p$.}
-% \item{$|S(d)| > 0$ uniformly in $d$}.
-% \end{enumerate}
-% \end{assumption}
-% We note that assumption $C$ implies that for every list depth $d$
-% there must be a non-zero probability of ... under the probability
-% measure $Q$.
-
-\begin{theorem}
-  Assume that $\{R_l(X)\}_{l=1}^L$ are independent draws from a
-  probability distribution $Q$ on the set of lists $\Pi$. Then,
-  $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty =
-  o_P(1)$.
-\label{thm:consistency}
-\begin{proof}
-See Appendix \ref{sec:appA}.
-\end{proof}
-\end{theorem}
-
-We now define the changepoint as the first crossing point of the
-sequential rank agreement and a threshold function
-$q\colon\,\{1,\ldots,P\} \mapsto \mathbb{R}_{\geq 0}$. The values of
-$q$ could be a deterministic constant or, % in the experimental setting,
-for example, the limits-of-agreement obtained in randomly permuted
-lists corresponding to the null-hypothesis in equation
-(\ref{eq:permutationHypothesis}). We define the superlevel set of the
-sequential rank agreement with respect to $q$ as
-\begin{align}
-  \mathcal{L}(q) = \left\{d : \textrm{sra}(d) \geq q(d)\right\}.
-\end{align}
-A changepoint $d^\ast(q)$ in the list agreement is then defined by the position
-\begin{equation}
-  d^\ast(q) =
-  \begin{cases}
-    \inf(\mathcal{L}(q)) & |\mathcal{L}(q)| > 0\\
-  P & |\mathcal{L}(q)| = 0
-  \end{cases}
-\end{equation}
-corresponding to the first list depth where the sequential rank
-agreement exceeds the threshold if such a position exists. Otherwise,
-the full list is in agreement according to $q$ and the changepoint is
-set to the full length of the lists. The empirical superlevel set is
-similarly defined as
-\begin{align}
- \widehat{\mathcal{L}}_L(\widehat{q}_L) = \left\{d : \widehat{\textrm{sra}}_L(d) \geq \widehat{q}_L(d)\right\}
-\end{align}
-where we allow the threshold function to depend on the sample size as
-well. The estimated changepoint is therefore
-\begin{align}
-  \widehat{d^\ast_L}(\widehat{q}_L) &= 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| > 0)\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) + 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0)P.
-\end{align}
-
-The consistency of the estimated changepoint, $\widehat{d^\ast_L}(\widehat{q}_L)$, follows from Theorem \ref{thm:consistency} by the following corollary. 
-\begin{corollary}\label{thecor}
-Let $\widehat{q}_L$ be a positive threshold function such that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ for some limiting
-function $q$.  Then, $\widehat{d^\ast_L}(\widehat{q}_L) \overset{P}{\longrightarrow} d^\ast(q)$ for $L \rightarrow \infty$.
-\label{cor:consistency}
-\begin{proof}
-See Appendix \ref{sec:appB}.
-\end{proof}
-\end{corollary}
-
-Corollary \ref{thecor} indicates that we can use the threshold
-function $\widehat{q}_L$ estimated under the null hypothesis as
-discussed in the previous section as a limiting threshold function for
-inferring the depth $d$, where the observed sequential rank agreement
-first crosses the threshold of the null threshold, \ie, the depth
-until which the observed ranked lists are in better agreement than
-expected under the null hypothesis. In that sense the threshold
-function serves the same role as the limits of agreement in method
-comparison studies, except that the threshold function is not constant
-but can accommodate the changing nature of the number of items used
-for the computation of the sequential rank agreement for a given depth.
-
-In practice we can compute an estimate of the threshold function under
-the null using the permutation approach sketched in the previous
-section which makes it relevant even in small sample settings.
-
-% \mcomment{Elaborate more on how to apply this in practice (finite
-  % sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style?  }
-
-\section{Application to ovarian cancer data}
-We now consider an application of the sequential rank agreement to two
-datasets consisting of MALDI-TOF (Matrix-Assisted Laser
-Desorption/Ionization Time Of Flight) mass spectra obtained from blood
-samples from patients with either benign or malignant ovarian
-tumors. The datasets are sub-samples of the Danish MALOVA and DACOVA
-study populations.
-
-The MALOVA study is a multidisciplinary Danish study on ovarian cancer
-\citep{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
-with an ovarian tumor and referred for surgery from the participating
-departments of gynecology were enrolled continuously from December
-1994 to May 1999. For the purpose of illustration we use a random
-sub-sample of $119$ patients with a total of $58$ patients with
-malignant ovarian cancers as cases and $61$ patients with benign
-ovarian tumors as controls. The DACOVA study is another
-multidisciplinary Danish study on ovarian cancer which included about
-$66\%$ of the female population of Denmark
-\citep{bertelsen1991protocol}. The study aimed to continuously enroll
-all patients that were referred to surgery of an ovarian tumor
-clinically suspected to be cancer during the period from 1984 to
-1990. Similarly, we use a random sub-sample from the DACOVA study of
-$113$ patients with a total of $54$ malignant ovarian cancers and $59$
-benign ovarian tumors/gynecologic disorders.
-
-Each spectrum consists of $49642$ samples over a range of mass-to-charge ratios
-between $800$ to $20000$ Dalton which we downsample on an equidistant grid of
-5000 points by linear interpolation. We then preprocess the downsampled
-spectra individually by first removing the slow-varying baseline intensity
-with the SNIP algorithm \citep{ryan1988snip} followed by a
-normalization with respect to the total ion count. Finally, we standardize
-the $5000$ predictors to have column-wise zero mean and unit variance in
-each dataset.
-
-We use the two datasets to illustrate how the sequential rank
-agreement can be applied in two different scenarios. In the first
-scenario we assess the agreement of four different statistical
-classification methods in how they rank the predictors according to
-their importance for distinguishing benign and malignant tumors. In
-the second scenario we assess the agreement among rankings of
-individual predicted risks of having a malignant tumor. The first
-scenario is relevant in the context of biomarker discovery and the
-latter is important e.g., when ranking patients according to immediacy of
-treatment.
-
-Four classification methods are considered: Random Forest
-\citep{breiman2001random} implemented in the R package
-\texttt{randomForest} \citep{liaw2002classification}, logistic Lasso
-\citep{tibshirani1996regression} and Ridge regression
-\citep{segerstedt1992ordinary} both implemented in the R package
-\texttt{glmnet} \citep{friedman2010regularization}, and Partial Least
-Squares Discriminant Analysis (PLS-DA) \citep{boulesteix2004pls}
-implemented in the R package \texttt{caret} \citep{Jed-Wing:2014aa}.
-
-In both scenarios we use the MALOVA data to train the statistical
-models, and in both situations the agreements are assessed with
-respect to perturbations of the training data in the following
-manner. We repeatedly draw a random sub-sample (without replication)
-consisting of 90\% of the MALOVA observations and train the four
-models on each sub-sample. We use 1000 iterations for the sub-sampling
-procedure.
-
-All four methods depend on a tuning parameter. The tuning parameter for
-Lasso and Ridge regression is the degree of penalization, and for PLS-DA
-it is the number of components (the dimensionality of the subspace).
-We estimate these separately for each sub-sample by a 20 times repeated
-5-fold cross-validation procedure. For the Random Forest we grow a fixed
-number of $5000$ trees and let the tuning parameter be the number of
-predictors randomly sampled at each split. We estimate this by a binary
-search with respect to minimizing the Out-of-Bag classification error estimate.
-
-The implementation of Lasso and Ridge regression in the
-\texttt{glmnet} package offers three different cross-validated
-optimization criteria for the penalty parameter: total deviance,
-classification accuracy and area under ROC. We apply all three
-criteria to our data to investigate their effect on the agreements.
-Note also that the Lasso models produce incomplete lists depending on
-the value of the penalty parameter.
-
-\subsection{Agreement of predictor rankings}
-For each of the four methods, each of the 1000 models trained on
-the 1000 sub-samples of the MALOVA data produces a ranking of the 5000
-predictors according to their importance for discriminating between
-the tumor types. For the Random Forest classifier the predictors are ranked according
-to the Gini index, while for the logistic Lasso and Ridge regression
-models we order by absolute magnitude of the estimated
-regression coefficients. For the PLS-DA model the importance of the
-predictors is based on a weighted sum of the absolute coefficients
-where the weights are proportional to the reduction in the sums of
-squares across the components.
-
-The right panel of Figure \ref{fig:app1} shows the sequential rank
-agreement of the estimated importance of the 5000 predictors.  For
-clarity of presentation we zoom in on the agreement up to list depth
-600. At deeper list depths all agreement curves are approximately constant.
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.49\textwidth]{pics/riskAgreementPlot}%
-\includegraphics[width=.49\textwidth]{pics/predictorAgreementPlot}
-\end{center}
-\caption{Left panel: Sequential rank agreement of 1000 rankings of the
-  predicted risks of malignant tumor. For each method the different
-  rankings were obtained by first training models in 1000 random
-  sub-samples of the MALOVA data and then predicting the risk of
-  malignant tumor in the 113 DACOVA patients. Right panel: Sequential
-  rank agreement of 1000 rankings of the 5000 predictors. The rankings
-  were obtained from the same 1000 trained models.}
- \label{fig:app1}
-\end{figure}
-
-For most of the sequential rank agreement curves we see, as expected,
-that they start low, indicating good agreement, followed by an
-increase until they approximately become constant. This has the
-interpretation that the agreement across the different sub-samples is
-higher in the top as compared to the tail of the lists for all these
-classification methods. The changepoints where the curves become
-approximately constant are the list depths where the ranks of the
-remaining items become close to uniformly random.
-
-A not expected shape of the agreement curves is seen for the Ridge
-models for all three tuning criteria. They all show higher
-disagreement in the top of the lists followed by a decrease. The
-reason behind this behavior is rather subtle. Looking at the
-distribution of the absolute value of the regression coefficients we
-see that a large proportion of them are numerically very close to zero
-and have almost equal absolute value. This is a general feature of the
-Ridge models in this dataset and seen for all the 1000 trained models.
-This implies that when predictors are ranked according to the
-magnitude of their coefficients, their actual order becomes more
-uncertain and more close to a random permutation. This problem can be
-alleviated by truncating all predictors with absolute coefficient
-values below a given threshold thereby introducing an artificial
-incompletion of the lists.  For the Ridge models tuned with the
-deviance criterion, Figure \ref{fig:app2} (left) shows the the
-sequential rank agreement where for each of the 1000 trained models
-the predictors were artificially censored when their absolute
-coefficient value was lower than the $0.1\%$ quantile of the 5000
-absolute coefficient values. The curve was calculated using Algorithm
-\ref{sra-algorithm} with $B=1000$ and $P=5000$. The corresponding
-curve from Figure \ref{fig:app1} (right panel) is shown for
-comparison.  Even though the number of predictors with missing ranks is very
-small compared to the total number of predictors, the effect on the
-sequential rank agreement is substantial and with the artificial
-censoring the shape of the curves is as expected, starting low and
-then increasing.
-
-Looking at the agreement curves for the Lasso models in Figure
-\ref{fig:app1} (right) we clearly see the effect of the sparsity
-inducing penalization giving rise to incomplete lists. These curves
-were similarly calculated using \ref{sra-algorithm} and $1000$ random
-permutations. Under the deviance optimization criterion the median
-number of non-zero coefficients was 33 (range 16 to 50) and for the
-class accuracy criterion 14 (range 4 to 56). These values correspond
-to the list depths where the agreement curves become constant as a
-result of the subsequent censoring.
-
-\subsection{Agreement of individual risk predictions}
-\label{sec:airp}
-To assess the stability of the individual risk predictions we apply
-the predictors from the DACOVA dataset to each of the models. The
-predicted probabilities are then ranked in decreasing order such that the
-patients with the highest risk of a malignant tumor appears in the top
-of the list. Figure \ref{fig:app1} (left) shows the sequential rank
-agreement separately for each method, based on the 1000 risk predictions
-obtained from the models trained in the same $1000$ random sub-samples
-of the MALOVA data.
-
-Most curves start low and then increase indicating higher agreement
-among high risk patients. This is expected if we rank the individuals
-according to highest risk of disease. However, it is also expected
-that individuals with very low risk also show high agreement. In this
-case we order the patients according to (high) risk prediction but we
-could essentially also have reversed the order to identify the
-patients that have low risk prediction.
-
-An exception is the risk prediction agreement for the Lasso tuned with the AUC criterion which
-shows very low agreement among the high values of the predicted
-risks. The reason is that optimizing the penalty parameter with
-respect to the AUC criterion tends to favor a very high penalty value
-causing only a single predictor to be selected in each of the 1000
-iterations. This results in a lack of generalizability to the DACOVA
-data which gives rise to the higher disagreement in the predicted risks.
-In the extreme case where the penalty becomes so high
-that none of the predictors are selected by the Lasso, the sequential
-rank agreement for the predicted probabilities becomes undefined since
-all the ranks will be ties.
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.49\textwidth]{pics/ridgeCensoring}%
-\includegraphics[width=.49\textwidth]{pics/aucPlot}
-\end{center}
-\caption{Left panel: Sequential rank agreement for Ridge regression
-  obtained by artificially censoring predictor ranks when their absolute
-  coefficient values are lower than the 0.1\% quantile. Right panel:
-  Box plots of AUC values across the $1000$ sub-samples with respect
-  to the known class labels of the DACOVA data.}
- \label{fig:app2}
-\end{figure}
-
-Comparing the left and right panels of Figure \ref{fig:app1} it can
-further be seen that some of the methods show better agreement with
-respect to the predicted probabilities than for ranking the importance
-of the predictors and vice versa. Ridge regression shows higher agreement across training
-sets for the risk predictions than PLS-DA, and PLS-DA shows higher
-agreement for predictor importance than Ridge regression.
-
-Lasso shows similar agreement for ranking the risk predictions than
-PLS-DA (except for the AUC criterion), and poorer agreement for
-ranking predictors. This reason for the latter is the high
-auto-correlation between the intensities in the mass spectra which
-leads to collinearity issues in the regression models. It is
-well-known that variable selection with the Lasso penalty does not
-perform very well when the predictors are highly correlated. The
-collinearity does, however, not affect the agreement of the risk
-predictions, as seen on the left panel of Figure \ref{fig:app1}, since
-it is not so important which specific variable that gets selected from
-a group of highly correlated predictors when the purpose is risk
-predictions.
-
-It appears that Ridge regression tuned with the AUC criterion achieves
-the best performance with respect to the stability of ranking the
-individual predicted risk probabilities. It must, however, be stressed
-that the sequential rank agreement in this application is only
-concerned with the agreement of the risk predictions across
-sub-samples and not with the actual accuracy of the risk
-predictions. Thus, we also computed the AUC values for the different models
-based on the DACOVA data. The distributions across the $1000$
-sub-samples for a selection of the models is shown in the right panel
-of Figure \ref{fig:app2}.  Here we see that PLS-DA attains the highest
-AUC values with a median value of $0.70$ while the Ridge model with
-the AUC criterion attains a median AUC of $0.49$. This implies that
-while Ridge regression optimized with respect to the AUC criterion
-achieves the best sequential rank agreement, it performs similar to a
-random coin toss with respect to classifying the DACOVA patients.
-In practice both concerns are of importance.
-
-\section{Simulation study --- comparison of list agreements}
-
-We present results from a simulation study where we investigated the
-small sample properties of the sequential rank agreement and compared
-it to the topK method \citep{hall:schi:2012} with respect to list
-depth agreement. For the sequential rank agreement we use the function
-\texttt{sra} of the \texttt{R} package \texttt{SuperRanker} with the
-\texttt{"mad"} option (median absolute distance). To define the depth
-of agreement we set a constant threshold function to an integer $q$
-and report the first crossing point, i.e., the smallest list depth
-where sra exceeds $q$. For topK we use the function
-\texttt{j0.multi} which is implemented in the \texttt{R} package
-\texttt{TopKLists}. Specifically, we set the tuning parameter
-\texttt{v} of \texttt{j0.multi} to the value 6 and the window
-parameter \texttt{d} to $q$ and report the output parameter
-\texttt{maxK} as the depth of agreement. Thus, to make this comparison
-we assume that the first crossing point of sra and the result of the
-topK method measure the same underlying feature.
-
-In each simulation run, we generated a dataset with 1000 predictors
-and 400 observations. The predictors were drawn independently from a
-standard Gaussian distribution. The outcome was drawn from a Gaussian
-distribution with variance 1 and mean that depended additively on the
-first 15 predictors, i.e.,
-
-$$y_i = \sum_{j=1}^{15}x_{ij} + \epsilon_i,$$
-
-where $y_i$ is the $i$th response, and $x_{ij}$ is the $j$th predictor
-for the $i$th measurement, and $\epsilon_i\sim N(0, 1)$. In each
-simulation run we obtained $L$ ranked lists of the 1000 predictors.
-This was done by first drawing $L$ bootstrap samples (with
-replacement) and then in each bootstrap sample ranking the 1000
-predictors according to their marginal $t$ test statistics. Thus, we
-assessed the depth of agreement among lists that are ranked with the
-same statistical method on bootstrap versions of the same dataset. We
-report simulation results from two scenarios where each is based on
-1000 simulated datasets:
-\begin{enumerate}
-\item[] Scenario I: Fix \(L=8\) and vary $q\in\{3,4,5,6,7,8,9,10\}$.
-\item[] Scenario II: Fix \(q=5\) and vary $L\in\{3,5,10,50\}$.
-\end{enumerate}
-In both scenarios we summarized the distribution of the estimated
-depth of agreement and the average number of unique predictors found
-in the set of predictors which is selected by the estimated depth of
-agreement.
- % Asymptotic results are available for both methods as the number of
- % lists increases, so this will investigate how much the results might
- % change as $L$ increases. For a fixed constant threshold of $q_L=5$ we
- % increase the number of available lists to investigate the sensitivity
- % of the results to $L$.
-  % Throughout the simulations we
-  % assume a constant, prespecified threshold function, $q_L$, and try
-  % different values for the constant. The approach by
-  % \citet{hall:schi:2012} uses a window size parameter, $K$, to
-  % represent the maximum difference in rankings before they consider an
-  % item in two lists to be in agreement.  Because the standard
-  % deviation --- which is used in the sequential rank agreement
-  % definition of $\hat{A}_L$--- squares its differences, it tends to
-  % give more weight to larger rank differences and less weight to
-  % smaller differences compared to the mean absolute difference.  To
-  % ease comparison between the sequential rank agreement and the
-  % \citet{hall:schi:2012} approach we would like to use the same
-  % threshold for both methods, and in order to do this we use the
-  % median absolute deviance throughout the simulations in lieu of the
-  % pooled standard deviation.
- % For all parameter combinations we ran 1000 simulations and report the
- % distribution of the estimated depth of agreement and the average
- % number of different predictors found at a given depth.
-The results from Scenario I are shown in the left panel of
-Figure~\ref{fig:example99}. The violin plots (with rectangular kernel)
-show the distributions of the estimated depths of agreement for both
-methods. We see that sra results in a substantially lower depth of
-agreement than the topK method. Also the average numbers of unique
-predictors (bold numbers inside the plots) are markedly smaller for
-sra. Even larger differences were found when we used the Euclidean
-distance instead of the median absolute distance for the sequential
-rank agreement (results not shown). As expected the depth of agreement
-increased when the threshold for agreement/window increased.
-\begin{figure}[htbp]
-   \begin{center}
- \includegraphics[width=.49\textwidth]{Ksim.pdf}
- \includegraphics[width=.49\textwidth]{Lsim}
- \end{center}
- \caption{Left panel: Simulation study showing distribution of
-   estimated rank agreements for sra and topK for varying thresholds
-   and fixed number of lists $L=8$. The bold numbers are the average
-   number of unique predictors included in the set where the lists
-   agree. Right panel: Simulation results for varying number of lists
-   and with fixed threshold of $q=5$.}
- \label{fig:example99}
-\end{figure}
-
-The right panel of Figure~\ref{fig:example99} shows the results from
-Scenario II. The number of lists has little impact on the
-results. Generally, sra is more conservative than topK and as a
-consequence sra includes fewer predictors in the selected set where
-the lists agree.
-
-\section{Discussion}
-In this article we address the problem of comparing ranked lists of
-the same items. Our proposed method can handle both the situation
-where the underlying data to generate the ranked lists are available
-and the situation where the only available data is the actual ranked
-lists. In addition, incomplete ranked lists where only the ranks of the
-top $k$ ranked items are known can be accommodated as well. The
-proposed agreement measure can be interpreted as the average distance
-between an item's rank and the average rank assigned to that item
-across lists. Thus the measure determines how well the lists agree on
-the ranks of a specific item.
-
-The sequential rank agreement is extremely versatile. We have shown
-that it can be used not only to compare ranked lists of items produced
-from different samples/populations but that it also can be used to
-study the ranks obtained from different analysis methods on the same
-data as well as to evaluate the stability of the ranks from a single
-method by bootstrapping (or sub-sampling) the data repeatedly and
-comparing the ranks obtained from training the models in the
-bootstrapped data.
-
-The sequential rank agreement can be used to determine the depth at
-which the rank agreement becomes too large to be desirable based on
-prior requirements or acceptable differences, or it can be used to
-visually determine when the change in agreement becomes too large.  In
-that regard the investigator can have prior limits on the level of
-agreement that is acceptable.
-
-While the sequential rank agreement is primarily an exploratory tool
-we have suggested two null hypotheses that can be used to evaluate the
-sequential rank agreement obtained. Note that none of the two null
-hypotheses are concerned with the actual ``true ranking'' but are
-purely concerned with consistency/stability of the rankings among the
-lists. As such we cannot determine if the rankings are good but only
-whether they agree. The sequential rank agreement curve can be
-compared visually to the curves obtained under either of the null
-distributions and simple point-wise $p$-values can be obtained for
-each depth by counting the number of sequential rank agreements under
-the null hypothesis that is less than or equal to the observed rank
-agreement. Two simple extensions can be pursued to make more formal
-uniform tests: one would be to use a Kolmogorov-Smirnov-like test and
-use the largest difference between the variance-weighted sequential
-rank agreement curve and the mean null rank agreement curve as a test
-statistic. Alternatively, a changepoint analysis could be made on the
-sequential rank agreement in order to determine the depths at which
-there are ``jumps'' in the rank agreement. These jumps would
-correspond to depths for which the agreement among the lists was
-substantially worse and could serve as indicators for when the lists
-no longer agree sufficiently satisfactory.
-
-Finally, we have --- whenever possible --- used all available ranks
-from the lists.  We could choose to restrict attention to the rank of
-items which show evidence for significance in their models. That would
-ensure that there would be put less emphasis on the agreement of the
-non-significant items and it would be easier to identify a change in
-agreement among the items that were deemed to be relevant. In our
-application section we have successfully introduced such an artificial
-censoring for the predictor rankings obtained with ridge regression.
-
-We note that the sequential rank agreement is still marred by problems
-that generally apply to ranking of items and/or
-individuals. Collinearity in particular can be a huge problem when
-bootstrapping data or when comparing different analysis methods. For
-example, marginal analyses where each item is analyzed separately will
-assign similar ranks to two highly correlated predictors while methods
-that provide a sparse solution such as the Lasso will just rank one of
-the two predictors high while the other might have a very low rank.
-Thus in such a scenario we would expect low agreement of the rankings
-from Lasso and marginal analyses simply because of the way correlated
-predictors are handled. This is not a shortcoming of the sequential
-rank agreement but is a problem general to all ranked lists.
-
-Another caveat with the way the sequential rank agreement is defined
-is the use of the standard deviation to measure agreement. The
-standard deviation is an integral part of the limits-of-agreement as
-discussed by \citet{alt:bland:1983}. However, the standard deviation
-can also be unstable when the number of observations is low and
-alternative measures such as the median absolute deviance may prove more
-stable in some situations. However, the current definition using the
-standard deviation is analogous to the approach used for agreement in
-method comparison studies so we have used that.
-
-In conclusion we have introduced a method for evaluation of ranked
-(partial/censored) lists that can be easily interpreted and that can be
-applied to a large number of situations.  The method presented here
-can be adapted further by using it to compare and classify statistical
-analysis methods that agree on the rankings they provide or by using
-the rank agreement to optimize a hyper-parameter in, say, elastic net
-regularized regression where the rank agreement is used to determine
-the mixing proportion between the $L_1$ and the $L_2$ penalty.
-%Finally, the proposed method may be adapted (with some additional
-%assumptions) to the situation where the lists are subsets of each
-%other.
-We will be investigating these extensions further in the future.
-
-%\bibliographystyle{rss}
-\bibliographystyle{rss}
-\bibliography{paperref}
-
-\appendix
-
-\newpage
-\section{Proof of Theorem \ref{thm:consistency}}
-\label{sec:appA}
-
-We start by defining the following cadlag function
-\begin{align}
-  N(p; d) = \sum_{s=1}^p 1\left(Q(R(X_s) \leq d) > 0\right), \quad p = 1,\ldots, P
-\end{align}
-which runs through the list elements in an arbitrary order and counts
-how many list elements that have a strictly positive probability to have
-rank less than or equal to $d$ under the probability measure $Q$.
-Note that the cardinality of the set $S(d)$ in equation
-(\ref{eq:sumset}) is equal to the counting process evaluated at the last element: $|S(d)| = N(P;d)$.
-The empirical counterpart of the counting process is given by
-\begin{align}
-  \widehat{N}_L(p; d) &= \sum_{s=1}^p 1\left( \frac{1}{L} \sum_{l=1}^L 1(R_l(X_s) \leq d) > 0\right)\\
-        &= \sum_{s=1}^p 1\left(\widehat{Q}_L(R(X_s) \leq d) > 0\right)\nn.
-\end{align}
-The joint law of the sequence
-$\{\widehat{N}_L(p;d)\}_{1\leq p \leq P}$ is completely determined by
-the finite set of the jump times of $N(p;d)$. Since
-$\{1(R_l(X_s) \leq d); l=1,\dots, L\}$ consists of independent and
-identically Bernoulli distributed variables with expectation
-$Q(R(X_s) \leq d)$ and finite variance it follows from the law of large numbers that
-$\widehat{Q}_L(R(X_p) \leq d) \overset{P}{\longrightarrow} Q(R(X_p)
-\leq d)$ for every $p$ and $d$ as $L \rightarrow \infty$. 
-Therefore, we have
-\begin{align}
-  \sup_{d \in 1,\ldots,P}\sup_{p \in 1,\ldots, P} \left|\widehat{N}_L(p; d) - N(p; d)\right| = o_P(1).\label{eq:countingConv}
-\end{align}
-
-The sequential rank agreement given in equation
-(\ref{def:sra}) may therefore be rewritten as the following integral
-with respect to $N$
-\begin{align}
-  \textrm{sra}(d) = \int_{1}^P \frac{A(X_p)}{N(P; d)}\mathrm dN(p; d).
-\end{align}
-The empirical sequential rank agreement is similarly given by
-\begin{align}
-  \widehat{\textrm{sra}}_L(d) &= \int_1^P \frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)}\mathrm d \widehat{N}_L(p; d) 
-\end{align}
-and it follows that
-\begin{align}
-  \sup_{d \in 1,\ldots,P}\left|\widehat{\textrm{sra}}_L(d) - \textrm{sra}(d)\right| \leq \sup_{d \in 1,\ldots,P}\left|U_L(d)\right| - \sup_{d \in 1,\ldots,P}\left|V_L(d)\right|
-\end{align}
-where
-\begin{align}
-  U_L(d) &= \int_1^P\frac{A(X_p)}{N(P;d)} \mathrm d\left(\widehat{N}_L(p; d) - N(p; d)\right)\\
-  V_L(d) &=  \int_1^P \left(\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right)\mathrm d\widehat{N}_L(p; d)
-\end{align}
-The conclusion of the proof therefore follows if each of these two
-terms are uniformly of order $o_P(1)$ in $d$.  For the first term we
-have that
-\begin{align}
-  \left|U_L(d)\right| \leq \left(\sup_{p \in 1,\ldots,P}\left|\frac{A(X_p)}{N(P;d)}\right|\right)\left(\widehat{N}_L(p; d) - N(p; d)\right)\biggr\rvert_1^P
-\end{align}
-where the second factor is $o_P(1)$ uniformly in $d$ by equation
-(\ref{eq:countingConv}). The first factor is $O(1)$ since $N(P;d)>0$
-uniformly in $d$ and because $A(X_p) = O(P^2)$ uniformly in $p$ by the Cauchy-Schwarz inequality. It
-thus follows that $\left|U_L(d)\right| = o_P(1)$ uniformly in $d$.
-
-Similarly we derive an upper bound for $V_L(d)$ by
-\begin{align}
- V_L(d) &\leq \sup_{p \in 1,\ldots,P}\left|\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right| \int_1^P \left|\mathrm d\widehat{N}_L(p; d)\right|\label{eq:secondTerm}
-\end{align}
-and it follows directly from the law of large numbers that
-$\widehat{A_L}(X_p) = A(X_p) + o_P(1)$ uniformly in $p$ since
-$R_i(X_p)$ in equation (\ref{eq:empDistance}) for $i=1,\ldots,L$ are
-independent and $Q$-identically distributed random variables with
-finite second moment. Again, $\widehat{N}_L(P; d) = N(P;d) + o_P(1)$ uniformly in $d$ by equation
-(\ref{eq:countingConv}) and $N(P;d)>0$ so by the continuous mapping theorem the first factor
-in equation (\ref{eq:secondTerm}) is of order $o_P(1)$ uniformly in
-$p$ and $d$. The result then follows by noting that the second factor
-is bounded by $P$ uniformly in $d$.
-
-\section{Proof of Corollary \ref{cor:consistency}}
-\label{sec:appB}
-Recall that $\mathcal{L}$ is the superlevel set of list positions
-where the sequential rank agreement exceeds the threshold function
-$q$. Let $A \triangle B = (A \setminus B) \cup (B \setminus A)$ be the
-symmetric difference between sets $A$ and $B$. It is then sufficient
-to show that $\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q) \overset{P}{\longrightarrow} \emptyset$
-for $L \rightarrow \infty$ where $\emptyset$ denotes the empty set.
-We have that
-\begin{align}
-  \plim_{L \rightarrow \infty}\left(\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q)\right) &=   \left(\plim_{L \rightarrow \infty}\widehat{\mathcal{L}}_L(\widehat{q}_L)\right) \triangle \mathcal{L}(q)\\
-  &= \left\{d : \plim_{L\rightarrow \infty} \left(\widehat{\textrm{sra}}_L(d) - \widehat{q}_L(d)\right)\geq 0\right\} \triangle \mathcal{L}(q)\nn\\
-  &= \left\{d : \textrm{sra}(d) - q(d) \geq 0\right\}\triangle \mathcal{L}(q)\nn\\
-  &= \mathcal{L}(q)\triangle \mathcal{L}(q)\nn\\
-  &= \emptyset\nn
-\end{align}
-as a consequence of Theorem \ref{thm:consistency}, the assumption that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ and the continuous mapping theorem. This completes the proof.
-
 \end{document}

### 2018-01-24 R/sra.R

@@ -217,7 +217,7 @@ sra.list <- function(object, B=1, na.strings=NULL, nitems=max(sapply(object, len
 #'     censored observations
 #' @param type The type of measure to use. Either sd (standard
 #'     deviation - the default) or mad (median absolute deviance)
-#' @param epsilon A non-negative numeric vector that contains the minimum limit in proportion of lists that must show the item. Defaults to 0. If a single number is provided then the value will be recycles to the number of items.   
+#' @param epsilon A non-negative numeric vector that contains the minimum limit in proportion of lists that must show the item. Defaults to 0. If a single number is provided then the value will be recycles to the number of items. Should usually be low.  
 #' @return A matrix with n columns and the same number of rows as for the input object. Each column contains one
 #' simulated sequential rank agreement curve from one permutation run.
 #' @author Claus Ekstrøm <ekstrom@@sund.ku.dk>
@@ -251,7 +251,7 @@ random_list_sra <- function(object, B=1, n=1, na.strings=NULL, type=c("sd", "mad
         for (j in 1:ncol(object)) {
             object[,j] <- c(sample(nitems, size=notmiss[j]), rep(NA, nitems-notmiss[j]))
         }
-        sra(object, B=B, type=type)
+        sra(object, B=B, type=type, epsilon=epsilon)
     })
     res
 

### 2018-02-25 latex/biosts3.tex
Updated readme with example code (simpel and more advanced)

@@
 In this article we introduce sequential rank agreement for measuring
-agreement among  ranked lists.  The general idea is to
-define agreement based on the sequence of ranks from a subset of the first $d$
-items in each list. As agreement metric we adapt the limits of
-agreement known from agreement between quantitative variables
+agreement among ranked lists.  The general idea is to define agreement
+based on the sequence of ranks from a subset of the first $d$ items in
+each list. As agreement metric we adapt the limits of agreement known
+from agreement between quantitative variables
 \citep{alt:bland:1983,Carstensen2010} but any measure for agreement
-could essentially be used. Our proposed approach allows us to
-compare multiple lists simultaneously, it provides a dynamic
-measure of agreement as a function of the depth in the lists, it places
-higher weight on items at the top of the list, it accommodates partially
+could essentially be used. Our proposed approach allows us to compare
+multiple lists simultaneously, it provides a dynamic measure of
+agreement as a function of the depth in the lists, it places higher
+weight on items at the top of the list, it accommodates partially
 observed lists of varying lengths, and has a natural interpretation
 that directly relates to the ranks.  Graphical illustration of
-sequential rank agreement allows us to infer any
-changepoints, i.e., a list depth where a substantial change in the agreement of
-the lists occur but we also provide asymptotical and randomization-based graphical
-tools to compare the observed rank agreement to the expected agreement
-found in non-informative data.  In this sense our approach is a
-combination and generalization of some of the ideas of
-\citet{Carterette2009} and \citet{Boulesteix2009}. The former compares
-two rankings based on the distance between them as measured by a
-multivariate Gaussian distribution and the latter presents an overview
-of approaches for aggregation of ranked lists including bootstrap and
-leave-one-out jackknife approaches. We show the asymptotic
+sequential rank agreement allows us to infer any changepoints, i.e., a
+list depth where a substantial change in the agreement of the lists
+occur but we also provide asymptotical and randomization-based
+graphical tools to compare the observed rank agreement to the expected
+agreement found in non-informative data.  In this sense our approach
+is a combination and generalization of some of the ideas of
+\added{\citet{iriz:etal:2005}}, \citet{Carterette2009}, and
+\citet{Boulesteix2009}. \citet{Carterette2009} compares two rankings
+based on the distance between them as measured by a multivariate
+Gaussian distribution and the latter presents an overview of
+approaches for aggregation of ranked lists including bootstrap and
+leave-one-out jackknife approaches. \added{\citet{iriz:etal:2005}
+  proposes a plot which is a special case of our setting where the
+  agreement metric is the overlap proportion}. We show the asymptotic
 distribution of the endpoint of agreement and discuss how to infer the
@@
 
-Other recent approaches consider the intersection of lists as the basis
-for a similarity measure. However, simple intersection also places
-equal weights on all depths of the list and therefore \citet{Fagin2003} and
-\citet{Webber2010} proposed weighted intersections which put more
-emphasis on the top of the lists. 
+Other recent approaches consider the intersection of lists as the
+basis for a similarity measure \citep{iriz:etal:2005}. However, simple
+intersection also places equal weights on all depths of the list and
+therefore \citet{Fagin2003}, and \citet{Webber2010} proposed weighted
+intersections which put more emphasis on the top of the lists.
 % One proposal is denoted the average overlap (AO) and used for

### 2018-02-28 latex/biosts3.tex
Shrunk to meet size requirements

@@
 
-  A typical situation where
-  these questions arise is in high-dimensional genomics studies such
-  as genome-wide association studies where several analysis methods
-  (e.g., regression methods, lasso, random forest) can be used
-  identify and rank millions of gene variants according to their
-  association with the outcome. The importance (i.e., ranking) of each
-  gene variant may vary from method to method and a consensus summary
-  of agreement of the findings is needed to determine which gene
-  variants to investigate more closely in subsequent validation
-  studies. To minimiize expenses it is only of interest to consider
-  gene variants that have high ranking across the different methods.
-  Another situation where multiple ranked lists appear is in machine
+  A typical situation where these questions arise is in
+  high-dimensional genomics studies such as genome-wide association
+  studies where several analysis methods (e.g., regression methods,
+  lasso, random forest) can be used identify and rank millions of gene
+  variants according to their association with the outcome. The
+  ranking of each gene variant may vary from method to method and a
+  consensus summary of agreement of the findings is needed to
+  determine which gene variants to investigate more closely in
+  subsequent validation studies. To minimiize expenses it is only of
+  interest to consider gene variants that have high ranking across the
+  different methods.  Multiple ranked lists also appear in machine
   learning where the stability of the ranks produced by a
-  ``black-box''-technique can be evaluted by bootstrapping the data
-  repeatedly and comparing the ranks obtained from training the models
-  on the bootstrapped data. Assessing which items are stable (i.e.,
-  consistant across bootstrap samples) will help to weed out possible
-  fluke findings.}
+  ``black-box''-technique can be evaluted by bootstrapping the
+  data. Assessing which items are stable (i.e., consistant rankings
+  across bootstrap samples) will help to weed out possible fluke
+  findings.}
 
@@
 from agreement between quantitative variables
-\citep{alt:bland:1983,Carstensen2010} but any measure for agreement
+\citep{alt:bland:1983} but any measure for agreement
 could essentially be used. Our proposed approach allows us to compare
@@
 leave-one-out jackknife approaches. \added{\citet{iriz:etal:2005}
-  proposes a plot which is a special case of our setting where the
-  agreement metric is the overlap proportion}. We show the asymptotic
-distribution of the endpoint of agreement and discuss how to infer the
-distribution in small-sample situations using computational
-methods. This enables us to make inferences about the endpoint of
-agreement even for situations where there is no actual changepoint
-present but just a gradual decline in agreement.
+  proposes a plot based on the intersection of lists which is a
+  special case of our setting where the agreement metric is the
+  overlap proportion}. However, simple intersection also places equal
+weights on all depths of the list and therefore \citet{Fagin2003}, and
+\citet{Webber2010} proposed weighted intersections which put more
+emphasis on the top of the lists.  Specifically, \citet{Webber2010}
+define their rank-biased overlap (RBO) by weighting with a converging
+series to ensure that the top is weighted higher than the potentially
+non-informative bottom of the lists.  It is possible to use the
+existing methods to calculate agreement of lists until a given depth,
+i.e., limited to the $d$ items of each list. However, the
+interpretation may not be straightforward, especially in the case of
+more than two lists, and they may not accommodate partial rankings.
 
@@
 % to calculate the similarity measure.
-
-Other recent approaches consider the intersection of lists as the
-basis for a similarity measure \citep{iriz:etal:2005}. However, simple
-intersection also places equal weights on all depths of the list and
-therefore \citet{Fagin2003}, and \citet{Webber2010} proposed weighted
-intersections which put more emphasis on the top of the lists.
-% One proposal is denoted the average overlap (AO) and used for
-% comparison later on. 
-Specifically, \citet{Webber2010} define their rank-biased overlap
-(RBO) by weighting with a converging series to ensure that the top is
-weighted higher than the potentially non-informative bottom of the
-lists.  It is possible to use the existing methods to calculate
-agreement of lists until a given depth, i.e., limited to the $d$ items
-of each list. However, the interpretation may not be straightforward,
-especially in the case of more than two lists, and they may not
-accommodate partial rankings.
 
@@
 sequential rank agreement for multiple ranked lists and discuss how to
-handle incomplete lists. In section 3 we discuss approaches to
-evaluate the results obtained from sequential rank agreement. Finally
-we apply the sequential rank agreement to two Danish ovarian cancer
-studies and compare our method to the method of \citet{hall:schi:2012}
-in a small sample simulation study before we discuss the findings
-along with possible extensions. The approaches presented in this
-manuscript are available in the R package \texttt{SuperRanker} which
-can be found on CRAN.
+handle incomplete lists. In section 3 we show the asymptotic
+distribution of the endpoint of agreement and discuss approaches to
+evaluate the results obtained from sequential rank agreement.
+%infer the
+%distribution in small-sample situations using computational
+%methods. This enables us to make inferences about the endpoint of
+%agreement even for situations where there is no actual changepoint
+%present but just a gradual decline in agreement.
+Finally we apply the sequential rank agreement to two Danish ovarian
+cancer studies and compare our method to the method of
+\citet{hall:schi:2012} in a small sample simulation study before we
+discuss the findings and possible extensions. The approaches
+presented are implemented in the R package \texttt{SuperRanker}  found on CRAN.
 % chunk 1
@@
 and similarly $R_l(X_{34})=1$.
-
+% 
 In all what follows we consider a fixed set of items and consider the
@@
 ranking over the $L$ lists, and it corresponds to the same measure
-that is used in method comparison studies to compute the limits of
+that is used to compute the limits of
 agreement \citep{alt:bland:1983}.
@@
 \end{equation}
-where $|S(d)|$ is the cardinality of the set $S(d)$. \added{As stated, we are only interested in sra($d$) when $|S(d)|>0$ so the value 0 above is somewhat arbitrary.}
+where $|S(d)|$ is the cardinality of the set $S(d)$. \added{As stated, we are only interested in sra($d$) when $|S(d)|>0$.}
 The empirical counterpart when $|S(d)|>0$ is equivalently given by
@@
   frequent among the observed lists and we can conclude that the lists
-  do not agree above the threshold $\varepsilon$.  } If the ranked
-lists are identical then the value of sequential rank agreement will
-be zero for all values of $d$. %position or depths $d$ down the lists.
+  do not agree above the threshold $\varepsilon$.  } The sequential rank agreement will be zero for all values of $d$
+when the ranked
+lists are identical.
+ %position or depths $d$ down the lists.
 
@@
 
-As an example consider the data by \citet{Golub1999} (also found in
+As an example consider the data by \citet{Golub1999} (found in
 \citet{Dudoit2002}) where 3051 gene expression values measured on 38
 tumor mRNA samples were used to classify between acute lymphoblastic
-leukemia (ALL) and acute myeloid leukemia (AML). Several possible
-analysis methods can be used for this for example marginal unequal
+leukemia (ALL) and acute myeloid leukemia (AML). Several 
+analysis methods are possible for these data for example marginal unequal
 variances two-sample $t$ tests, marginal logistic regression analyses,
@@
 and marginal maximum information content correlations (MIC)
-\citep{Reshef2011}, and we would like to identify a set of genes that
+\citep{Reshef2011}, and we want to identify a set of genes that
 consistently are most likely to be associated to leukemia.
@@
 corresponds to \emph{better} agreement) than towards the bottom of the
-lists. It is clear from the curve in Figure~\ref{fig:example1} that
-there is a substantial deterioration in agreement (higher sra) after
+lists. Figure~\ref{fig:example1} shows a substantial deterioration in agreement (higher sra) after
 depth 5. Thus, if we were to restrict attention to a small set of
@@
 are sets of items that are ranked similarly in all lists while other
-items constitute set(s) that have been assigned vastly different
+items constitute set(s) with vastly different
 ranks. When the lists are likely to agree on a few top ranked items
-then the sequential rank agreement curve will start at a low level and
+then the sequential rank agreement curve will start low and
 then increase until it levels off exactly as is seen in
@@
 agreements. The algorithm is described in detail in
-Algorithm~\ref{sra-algorithm}.
+Algorithm~C in the appendix.
 
@@
 Alternatively, we can remove the restriction of the independence among
-the methods used to generate the $L$ ranked lists and only require
-that there is no information contained in the rankings but that the
-rankings are all based on applying the method/approaches to the same
-data
+the methods used to generate the $L$ ranked lists under the null
+hypothesis that any association to the outcome is removed in the
+data.
+
+%and only require
+%that there is no information contained in the rankings but that the
+%rankings are all based on applying the method/approaches to the same
+%data
 \begin{eqnarray*}
@@
 rank the data ($t$ test, logistic regression, elastic net, and MIC)
-generally tend to identify (and rank) similar predictors even if there
-are only spurious associations.
+generally tend to identify the same predictors.
+%even if there are only spurious associations.
 
@@
   between the outcome and the predictors in the dataset. Both
-  hypotheses are purely considering how the rankings agree in a
-  situation where there is no relevant information available in the
-  data used for creating the rankings. It is also worth pointing out,
-  that if the lists are short ($P$ low) and there are few lists ($L$
-  low) then the number of possible different permutations under the
-  null is small and the $p$ value obtained may be fluctuating if the
-  number of permutations is small. We have found that a number of
-  permutations over 500 works well for smaller samples. The number of
-  permutations can be lowered for larger data sizes.}
+  hypotheses are purely considering how the rankings agree in
+  situations where the data contains no information for creating the
+  rankings. It is also worth pointing out, that if the lists are short
+  ($P$ low) and there are few lists ($L$ low) then the number of
+  possible different permutations under the null is small and the $p$
+  value obtained may be fluctuating if the number of permutations is
+  small. We have found that a number of permutations over 500 works
+  well even for smaller samples.}
 
@@
   Assume that $\{R_l(X)\}_{l=1}^L$ are independent draws from a
-  probability distribution $Q$ on the set of lists $\Pi$. Then,
+  probability distribution $Q$ on the set of lists $\Pi$. Then
   $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty =
@@
 \begin{proof}
-See Appendix A in the supplementary material.%\ref{sec:appA}.
+See App. A in the supplement.%ary% material.%\ref{sec:appA}.
 \end{proof}
@@
 \end{align}
-where we allow the threshold function to depend on the sample size as
-well. The estimated changepoint is therefore
+where the threshold function may depend on the sample size.
+% as well.
+The estimated changepoint is% therefore
 \begin{align}
@@
 Let $\widehat{q}_L$ be a positive threshold function such that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ for some limiting
-function $q$.  Then, $\widehat{d^\ast_L}(\widehat{q}_L) \overset{P}{\longrightarrow} d^\ast(q)$ for $L \rightarrow \infty$.
+function $q$.  Then $\widehat{d^\ast_L}(\widehat{q}_L) \overset{P}{\longrightarrow} d^\ast(q)$ for $L \rightarrow \infty$.
 \label{cor:consistency}
 \begin{proof}
-See Appendix B in supplementary material.%\ref{sec:appB}.
+See App. B in supplement.%ary material.%\ref{sec:appB}.
 \end{proof}
@@
 
-The MALOVA study is a multidisciplinary Danish study on ovarian cancer
+The MALOVA study is a Danish study on ovarian cancer
 \citep{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
@@
 ovarian tumors as controls. The DACOVA study is another
-multidisciplinary Danish study on ovarian cancer which included about
+Danish study on ovarian cancer which included about
 $66\%$ of the female population of Denmark
@@
 clinically suspected to be cancer during the period from 1984 to
-1990. Similarly, we use a random sub-sample from the DACOVA study of
-$113$ patients with a total of $54$ malignant ovarian cancers and $59$
-benign ovarian tumors/gynecologic disorders.
+1990. We use a random sub-sample from the DACOVA study of $54$
+malignant ovarian cancers and $59$ benign ovarian tumors/gynecologic
+disorders.
 
@@
 respect to perturbations of the training data in the following
-manner. We repeatedly draw a random sub-sample (without replication)
+manner. We draw 1000 random sub-samples (without replication)
 consisting of 90\% of the MALOVA observations and train the four
-models on each sub-sample. We use 1000 iterations for the sub-sampling
-procedure.
+models on each sub-sample.
+%We use 1000 iterations for the sub-sampling
+%procedure.
 
@@
 classification accuracy and area under ROC. We apply all three
-criteria to our data to investigate their effect on the agreements.
-Note also that the Lasso models produce incomplete lists depending on
+criteria to investigate their effect on the sra.
+Note that these models produce incomplete lists depending on
 the value of the penalty parameter.
@@
 
-The right panel of Figure \ref{fig:app1} shows the sequential rank
+The top right panel of Figure \ref{fig:app1} shows the sequential rank
 agreement of the estimated importance of the 5000 predictors.  For
@@
 As expected, most of the sequential rank agreement curves start low,
-indicating good agreement, followed by an increase until they
-approximately become constant. This has the interpretation that the
-agreement across the different sub-samples is higher in the top as
+indicating good agreement, followed by an increase until they become
+approximately constant. This has the interpretation that the
+agreement across the sub-samples is higher in the top as
 compared to the tail of the lists for all these classification
@@
 incompletion of the lists.  For the Ridge models tuned with the
-deviance criterion, Figure \ref{fig:app2} (left) shows the the
+deviance criterion, Figure \ref{fig:app1} (bottom left) shows the the
 sequential rank agreement where for each of the 1000 trained models
@@
 absolute coefficient values. The curve was calculated using Algorithm
-\ref{sra-algorithm} with $B=1000$ and $P=5000$. The corresponding
-curve from Figure \ref{fig:app1} (right panel) is shown for
+1 from the appendix with $B=1000$ and $P=5000$. The corresponding
+curve from Figure \ref{fig:app1} (top right) is shown for
 comparison.  Even though the number of predictors with missing ranks is very
@@
 Looking at the agreement curves for the Lasso models in Figure
-\ref{fig:app1} (right) we clearly see the effect of the sparsity
+\ref{fig:app1} (top right) we clearly see the effect of the sparsity
 inducing penalization giving rise to incomplete lists. These curves
-were similarly calculated using \ref{sra-algorithm} and $1000$ random
+were similarly calculated using the algorithm and $1000$ random
 permutations. Under the deviance optimization criterion the median
@@
 the patients with the highest risk of a malignant tumor appears in the
-top of the list. Figure \ref{fig:app1} (left) shows the sra separately
+top of the list. Figure \ref{fig:app1} (top left) shows the sra separately
 for each method, based on the 1000 risk predictions obtained from the
@@
 
-Comparing the left and right panels of Figure \ref{fig:app1} it can
+Comparing the top left and right panels of Figure \ref{fig:app1} it can
 further be seen that some of the methods show better agreement with
@@
 collinearity does, however, not affect the agreement of the risk
-predictions (Figure \ref{fig:app1}, left panel), since
+predictions (Figure \ref{fig:app1}, top left), since
 the specific variable selected is not that important for 
@@
 based on the DACOVA data. The distributions across the $1000$
-sub-samples for a selection of the models is shown in the right panel
-of Figure \ref{fig:app2}.  Here we see that PLS-DA attains the highest
+sub-samples for a selection of the models is shown in the bottom right panel
+of Figure \ref{fig:app1}.  Here we see that PLS-DA attains the highest
 AUC values with a median value of $0.70$ while the Ridge model with
@@
 
-We present results from a simulation study where we investigated the
-small sample properties of the sequential rank agreement and compared
-it to the topK method \citep{hall:schi:2012} with respect to list
-depth agreement.
-
+We present results from a simulation study where we compare the small
+sample properties of the sequential rank agreement to the topK method
+\citep{hall:schi:2012}.
 \added{The purpose of the simulations were twofold: First we want to
   investigate the rank agreement as it changes with threshold $q$ and
-  number of lists $L$. Secondly, we want to compare the results (and
-  conclusions) from sra with the topK for a realistic situation where
+  number of lists $L$. Secondly, we want to compare the results from sra with the topK for a realistic situation where
   the true, underlying agreement is not governed by a simple
@@
   the methods: the depth until which they agree, and the number of
-  unique predictors found in that set.
+  unique predictors found.
 
@@
   i.e., the smallest list depth where sra exceeds $q$ (as implemented
-  in the \texttt{sra} function from package \texttt{SuperRanker} using
+  in the \texttt{sra} function from our package \texttt{SuperRanker} using
   the median absolute distance argument). For topK we use the function
@@
 topK and as a consequence sra includes fewer predictors in the
-selected set where the lists agree.
-
-Note that the effect sizes of the 15 predictors in the model is the
+selected set where the lists agree. The effect sizes of the 15 predictors in the model is the
 same and in practice we observe that the majority of the 15 predictors
@@
 compared visually to the curves obtained under either of the null
-distributions and simple point-wise $p$-values can be obtained for
+distributions and point-wise $p$-values can be obtained for
 each depth by counting the number of sequential rank agreements under
@@
 from the lists.  We could choose to restrict attention to the rank of
-items which show evidence for significance in their models. That would
-ensure that there would be put less emphasis on the agreement of the
+items which show evidence for significance in their models. That
+ensures that less emphasis is put on the agreement of the
 non-significant items and it would be easier to identify a change in
-agreement among the items that were deemed to be relevant. In our
-application section we have successfully introduced such an artificial
-censoring for the predictor rankings obtained with ridge regression.
+agreement among the items that were deemed to be relevant.  Artificial
+censoring was successfully introduced in the ridge regression
+application.
+%In our
+%application section we have successfully introduced such an artificial
+%censoring for the predictor rankings obtained with ridge regression.
 
@@
 assign similar ranks to two highly correlated predictors while methods
-that provide a sparse solution such as the Lasso will just rank one of
-the two predictors high while the other might have a very low rank.
-Thus in such a scenario we would expect low agreement of the rankings
-from Lasso and marginal analyses simply because of the way correlated
-predictors are handled. This is not a shortcoming of the sequential
-rank agreement but is a problem general to all ranked lists.
+that provide a sparse solution such as the Lasso will just rank a
+single of the two predictors high.  Thus in such a scenario we would
+expect low agreement of the rankings from Lasso and marginal analyses
+simply because of the way correlated predictors are handled. This is
+general problem for ranked lists and not a shortcoming of the
+sequential rank agreement. %but is a problem
+%general to all ranked lists.
 
@@
 standard deviation is an integral part of the limits-of-agreement as
-discussed by \citet{alt:bland:1983}. However, the standard deviation
-can also be unstable when the number of observations is low and
-alternative measures such as the median absolute deviance may prove more
-stable in some situations. However, the current definition using the
-standard deviation is analogous to the approach used for agreement in
-method comparison studies so we have used that.
+discussed by \citet{alt:bland:1983} which is why we have followed the
+analogous path. However, the standard deviation is unstable
+when the number of observations is low and alternatives such
+as the median absolute deviance may prove more stable in some
+situations.
 
@@
 %other.
-We will be investigating these extensions further in the future.
+%We will be investigating these extensions further in the future.
 
@@
 \includegraphics[width=.49\textwidth]{pics/predictorAgreementPlot}
+\includegraphics[width=.49\textwidth]{pics/ridgeCensoring}%
+\includegraphics[width=.49\textwidth]{pics/aucPlot}
 \end{center}
-\caption{Left panel: Sequential rank agreement of 1000 rankings of the
+\caption{Top left panel: Sequential rank agreement of 1000 rankings of the
   predicted risks of malignant tumor. For each method the different
@@
   sub-samples of the MALOVA data and then predicting the risk of
-  malignant tumor in the 113 DACOVA patients. Right panel: Sequential
+  malignant tumor in the 113 DACOVA patients. Top right panel: Sequential
   rank agreement of 1000 rankings of the 5000 predictors. The rankings
-  were obtained from the same 1000 trained models.}
+  were obtained from the same 1000 trained models.
+Bottom left panel: Sequential rank agreement for Ridge regression
+  obtained by artificially censoring predictor ranks when their absolute
+  coefficient values are lower than the 0.1\% quantile. Bottom right panel:
+  Box plots of AUC values across the $1000$ sub-samples with respect
+  to the known class labels of the DACOVA data.
+}
  \label{fig:app1}
@@
 
-\newpage
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.49\textwidth]{pics/ridgeCensoring}%
-\includegraphics[width=.49\textwidth]{pics/aucPlot}
-\end{center}
-\caption{Left panel: Sequential rank agreement for Ridge regression
-  obtained by artificially censoring predictor ranks when their absolute
-  coefficient values are lower than the 0.1\% quantile. Right panel:
-  Box plots of AUC values across the $1000$ sub-samples with respect
-  to the known class labels of the DACOVA data.}
- \label{fig:app2}
-\end{figure}
+%\newpage
+%
+%\begin{figure}[htbp]
+%\begin{center}
+%\includegraphics[width=.49\textwidth]{pics/ridgeCensoring}%
+%\includegraphics[width=.49\textwidth]{pics/aucPlot}
+%\end{center}
+%\caption{Left panel: Sequential rank agreement for Ridge regression
+%  obtained by artificially censoring predictor ranks when their absolute
+%  coefficient values are lower than the 0.1\% quantile. Right panel:
+%  Box plots of AUC values across the $1000$ sub-samples with respect
+%  to the known class labels of the DACOVA data.}
+% \label{fig:app2}
+%\end{figure}
 
@@
 
-\newpage
-
-\begin{algorithm}
-\caption{Sequential rank agreement algorithm for incomplete lists}
-\label{sra-algorithm}
-\begin{algorithmic}[1]
-  \Procedure{Incomplete list rank agreement}{} \State Let $B$ be the number
-  of permutations to use \For{each $b \in B$} \For{each list
-    $l \in L$}
-  \State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1.7cm}{Permute
-    the unassigned ranks, $\{d_l+1, \ldots, P\}$, and assign them
-    randomly to the items \emph{not} found in the list, i.e.,
-    $\Lambda^\complement=X\setminus\Lambda_l$. Combine the
-    result with $\Lambda_l$ to fill out the list.}
-\EndFor
-\State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1cm}{Let sra($b$) be the sequential rank agreement computed from the
-filled out lists.}
-\EndFor
-\State Return element-wise averages across all $B$ permutations of sra($b$).
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
 \end{document}

### 2023-08-27 inst/topK.Rmd
Updated documentation

---
title: "sra/TopK - simulation comparisons"
author: "Claus Ekstrøm"
date: "16 Mar 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up packages and global parameters

```{r}
library(SuperRanker)
library(TopKLists)
library(MESS)

#Simulate a list of ranks
# K location of jump
# P = total length of predictors
# p = probability that top is permuted
simRank <- function(K, P, p) {
    topPermuted <- rbinom(1, 1, p)
    top <- (1-topPermuted) * (1:K) + topPermuted * sample(1:K)
    bottom <- sample((K+1):P)
    c(top, bottom)
}

blockSim <- function(blocksizes, maxlength) {
  x <- seq(maxlength)
  slut <- pmin(maxlength, c(cumsum(blocksizes), maxlength))
  start <- c(1, pmin(maxlength, cumsum(blocksizes)+1))
  unlist(sapply(seq(length(start)), function(i) { sample(seq(start[i], slut[i]))}))
}

# Example simulation
myrank <- sapply(1:3, function(i) simRank(K=10, P=100, p=0))
```

```{r}
nsim <- 1000
N <- 400
listlength <- 1000
threshold <- 12     # Our threshold. q*
nlists <- 10       # Number of lists available
Klist <- c(3, 5, 8, 15, 20)  # How far the lists agree. The truth
nu <- 6 # Tuning parameter for topK. Should be somewhere between 2 and 10 as per their paper
```

## Run simulations

```{r echo=FALSE, warning=FALSE}
simple <- function(K, nlists=nlists) {

res <- sapply(1:nsim, function(i) {

  # Simulate lists
  # Fixed parameters for now
  simlists <- bootstrappedlists(n=400, 
                                nlists=8, 
                                preal=15, 
                                listlength, 
                                effect=15)

  ## Return the results from sra and topK
  topk <- j0.multi(simlists, d=K, v=nu)$maxK
  myres <- c(
    min(which(sra(simlists, type="mad")>=K),listlength)-1,
    ifelse(is.na(topk), 1, topk)-1  # Becomes 0 later on
    )
  ## Indsæt predictor values her
  c(myres, length(unique(as.vector(simlists[0:myres[1],]))), length(unique(as.vector(simlists[0:myres[2],]))))
  })
res

# Store the results in a data frame

results <- data.frame(y=c(res[1,], res[2,]),
                       x=rep(c("sra", "topK"), times=rep(nsim, 2)),
                       np=c(res[3,], res[4,]),
                       K=K,
                       nlists=nlists)

results
}

Klist <- c(3, 5, 8, 10)
Knames <- c(`3` = "q=3",
            `5` = "q=5",
            `8` = "q=8", 
            `10` = "q=10"
                    )

www <- do.call(rbind, 
               lapply(Klist, function(K) simple(K, nlists=nlists)))

www <- www %>% group_by(x, K) %>% mutate(grp.mean.values = round(mean(np), 2))

p <- ggplot(www, aes(x=factor(x), y=y, fill=factor(x))) +
   geom_violin( scale="width", adjust=.4, bw=.7, kernel="rectangular")  + 
  facet_grid(. ~ K, labeller = as_labeller(Knames)) + xlab("") + ylab("Depth of agreement") + theme(legend.position="none") + theme_bw() + 
  stat_summary(aes(label=grp.mean.values), fun.y=mean, geom="text", size=6, fontface = "bold") + 
  theme(strip.text.x = element_text(size=16, face="bold"),
        text = element_text(size=14),
        panel.border = element_blank(),
        legend.position="none")  +  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))
p

```

Now change the value of $K$ and run simulations again

```{r echo=FALSE, warning=FALSE}
Klist <- c(3, 5, 10, 50)

## Kør for fast antal prædiktorer men varierer antallet af lister
print(Klist)

www <- do.call(rbind, 
               lapply(Klist, function(K) simple(5, nlists=K)))

www <- www %>% group_by(x, nlists) %>% mutate(grp.mean.values = round(mean(np), 2))

save(www, file="www")
#www
# 
```

```{r echo=FALSE}
Knames <- c(`3` = "L=3",
            `5` = "L=5",
            `10` = "L=10", 
            `50` = "L=50"
                    )

p <- ggplot(www, aes(x=factor(x), y=y, fill=factor(x))) +
   geom_violin( scale="width", adjust=.4, bw=.7, kernel="rectangular")  + 
  facet_grid(. ~ nlists, labeller = as_labeller(Knames)) + xlab("") + ylab("Depth of agreement") + theme(legend.position="none") + stat_summary(aes(label=grp.mean.values), fun.y=mean, geom="text", size=6, fontface = "bold") + theme_bw() + 
  theme(strip.text.x = element_text(size=16, face="bold"),
        panel.border = element_blank(),
        text = element_text(size=14), legend.position="none")  +  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))
p
```

### 2023-08-27 latex/R/censoring.Rmd
Updated documentation

---
title: "SuperRanker issues"
author: "Claus Ekstrøm"
date: "2 Jun 2015"
output: html_document
---

# Betydning af censurering

```{r echo=FALSE}
#note: always pass alpha on the 0-255 scale
makeTransparent<-function(someColor, alpha=100)
{
  newColor<-col2rgb(someColor)
  apply(newColor, 2, function(curcoldata){rgb(red=curcoldata[1], green=curcoldata[2],
    blue=curcoldata[3],alpha=alpha, maxColorValue=255)})
}
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(SuperRanker)
library(glmnet)
library(minerva)
library(changepoint)
library(randomForestSRC)

# Read data
library(multtest)
data(golub)

y <- golub.cl
x <- golub

producelists <- function(x, y) {
    nitems <- nrow(x)
    index <- seq(1, nrow(golub))
                                        # d <- data.frame(y, x)

                                        # Marginale t-tests
    mt.p <- sapply(index, function(i) { t.test(x[i,] ~ y)$p.value } )
    list1 <- order(mt.p)

                                        # Marginale logreg-tests
    mlogreg.p <- sapply(index, function(i) { drop1(glm(y ~ x[i,], family=binomial), test="Chisq")[2,5] } )
    list2 <- order(mlogreg.p)

                                        # Elastic net
    X <- scale(t(x))
    enet <- glmnet(X, y, family="binomial", alpha=.8)
    nyres <- cv.glmnet(X, y, family="binomial", alpha=.8)
    coefficients <- coef(enet, s=nyres$lambda.1se)[-1]
    nonzeros <- sum(coefficients!=0)
    list3 <- order(abs(coefficients), decreasing=TRUE)
    if (nonzeros<nitems) {
        list3[(nonzeros+1):nitems] <- NA
    }
                                        # MIC
    MIC <- sapply(index, function(i) { mine(x[i,], y)$MIC})
    list4 <- order(MIC, decreasing=TRUE)

                                        # Random Forest
###    dd <- data.frame(y=factor(y), t(x))
                                        # dd <- data.frame(y=y, t(x))
###    f1 <- rfsrc(y ~ ., data=dd, ntree=100)
###    variables <- abs(f1$importance[,1])
###    num.undecided <- sum(variables==0)
###    list5 <- order(variables, decreasing=TRUE)
###    list5[(length(variables)-num.undecided):length(variables)] <- 0

    cbind(list1,list2,list3,list4)
}

inputmatrix <- producelists(x, y)
colnames(inputmatrix) <- c("T", "LogReg", "ElasticNet", "MIC", "RF")[1:ncol(inputmatrix)]

res <- sqrt(sra(inputmatrix, B=20))

```

Censurerer 4 metoder ved 20 (dvs. 4 top-20 lister). Lader den funde listlængde være hhv. 50, 100, 200

```{r}
im <- inputmatrix[1:20,1:2]

res50 <- sqrt(sra(im, B=200, nitems=50))
res100 <- sqrt(sra(im, B=200, nitems=100))
res200 <- sqrt(sra(im, B=200, nitems=200))
```

```{r}
plot(res50, ylim=c(0, 80))
lines(res100, col="red", lwd=3)
lines(res200, col="blue", lwd=3)
```

```{r}
plot(res100/res50, type="l", ylim=c(0, 10), xlim=c(0, 50))
lines(res200/res50)
```

---

# Randomsiering vs. mean censored rank

```{r}
m <- matrix(c(1:4,1,NA, NA, NA, 1:2, NA, NA), ncol=3)
m
sra(m, B=1000)
```

Her er der to "problemer":

* Hvordan ved vi, hvilke items der skal tages med? Fx. ved dybde 2 skal vi så se på {1, 2} eller {1, 2, ?}?
* A(2) = (sd(c(2,2,2)) + sd(c(2,2,3)) + sd(c(2,2,4))) / 3

```{r}
m <- matrix(c(1:3, 1, NA, NA), ncol=2)
m
sra(m, B=10000)

```

### 2023-08-27 latex/biosts3-supp.tex
Updated documentation

@@
 \usepackage[utf8]{inputenc}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 %\usepackage[]{graphicx}
@@
 
-This document provides proofs of the theorem and corollary shown in the main manuscript.
+This document provides proofs of the theorem and corollary shown in the main manuscript as well as details about the algorithm for incomplete lists.
 
@@
 
+\section{Algorithm for sequential rank agreement for incomplete lists}
+\label{sec:appA}
+
+\begin{algorithm}
+\caption{Sequential rank agreement algorithm for incomplete lists}
+\label{sra-algorithm}
+\begin{algorithmic}[1]
+  \Procedure{Incomplete list rank agreement}{} \State Let $B$ be the number
+  of permutations to use \For{each $b \in B$} \For{each list
+    $l \in L$}
+  \State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1.7cm}{Permute
+    the unassigned ranks, $\{d_l+1, \ldots, P\}$, and assign them
+    randomly to the items \emph{not} found in the list, i.e.,
+    $\Lambda^\complement=X\setminus\Lambda_l$. Combine the
+    result with $\Lambda_l$ to fill out the list.}
+\EndFor
+\State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1cm}{Let sra($b$) be the sequential rank agreement computed from the
+filled out lists.}
+\EndFor
+\State Return element-wise averages across all $B$ permutations of sra($b$).
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
 \end{document}

### 2023-08-27 latex/superranker-manuscript-chunks.Rnw
Updated documentation

--- 2016-03-31 latex/paper.Rnw
@@
-\documentclass[12pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[]{graphicx}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage[]{color}
-\usepackage{subcaption}
-\usepackage{natbib}
-\usepackage{xspace}
-\usepackage{calc}
-\usepackage[noae]{Sweave}
-\usepackage{authblk}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-%\usepackage{geometry}
-%\geometry{a4paper, portrait, margin=1in}
-%
-% New commands
-%
-\newcommand{\ie}{\emph{i.e.}\@\xspace}
-\newcommand{\eg}{\emph{e.g.}\@\xspace}
-
-\DeclareMathOperator{\E}{\mathbb{E}}
-\DeclareMathOperator{\Var}{Var}
-\DeclareMathOperator{\Cov}{Cov}
-\DeclareMathOperator{\arginf}{arginf}
-
-\usepackage[amsthm,thmmarks]{ntheorem}
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}
-\newtheorem{definition}{Definition}
-\newtheorem{proposition}{Proposition}
-\newtheorem{corollary}{Corollary}
-\newtheorem{lemma}{Lemma}
-\newtheorem{assumption}{Assumption}
-\newtheorem{example}{Example}
-\renewcommand\proofSymbol{\ensuremath{\blacksquare}}
-
-\SweaveOpts{prefix.string=paper, keep.source=TRUE}
-
-%%% RSS specific items
-
-\title{Sequential rank agreement methods for comparison of ranked lists}
-\author{Claus Thorn Ekstrøm}
-\author{Thomas Alexander Gerds}
-\author{Andreas Kryger Jensen}
-\author{Kasper Brink-Jensen}
-\affil{Biostatistics, University of Copenhagen}
-\renewcommand{\topfraction}{.85}
-\renewcommand{\bottomfraction}{.7}
-\renewcommand{\textfraction}{.15}
-\renewcommand{\floatpagefraction}{.66}
-\renewcommand{\dbltopfraction}{.66}
-\renewcommand{\dblfloatpagefraction}{.66}
-\setcounter{topnumber}{9}
-\setcounter{bottomnumber}{9}
-\setcounter{totalnumber}{20}
-\setcounter{dbltopnumber}{9}
-
-\begin{document}
-\SweaveOpts{concordance=TRUE}
-
-% Use square brackets for numbering affiliations
-
-\maketitle
-
-\begin{abstract}
-  The comparison of alternative rankings of a set of items is a
-  general and prominent task in applied statistics. Predictor
-  variables are ranked according to magnitude of association with an
-  outcome, prediction models rank subjects according to the
-  personalized risk of an event, and genetic studies rank genes
-  according to their difference in gene expression levels. This
-  article constructs measures of the agreement of two or more ordered
-  lists. We use the standard deviation of the ranks to define a
-  measure of agreement that both provides an intuitive interpretation
-  and can be applied to any number of lists even if some or all are
-  incomplete or censored. The approach can identify change-points in
-  the agreement of the lists and the sequential changes of agreement
-  as a function of the depth of the lists can be compared graphically
-  to a permutation based reference set.  The usefulness of these tools
-  are illustrated using gene rankings, and using data from two Danish
-  ovarian cancer studies where we assess the within and between
-  agreement of different statistical classification methods.
-\end{abstract}
-
-Key words: sequential rank agreement, partial ranking, order
-statistic, gene rankings, methods comparison, variable selection,
-permutation
-
-\section{Introduction}
-
-Ranking of items or results is common in scientific research and
-ranked lists occur naturally as the result of many statistical
-applications. Regression methods rank predictor variables according to
-magnitude of their association with an outcome, prediction models rank
-subjects according to their risk of an event, and genetic studies rank
-genes according to their difference in gene expression levels across
-samples.
-
-When several rankings of the same items are available, a common
-research question is to what extent they agree. In particular, is it
-possible to identify an optimal rank until which the lists agree on
-the items?  A typical situation arises in high-dimensional genomics
-studies when several analysis methods are applied to rank a list of
-genes according to their association with a phenotype, treatment
-effect or other outcome. A measure of agreement of gene rankings
-obtained by different methods could often help to identify which genes
-are worth to pursue in further experiments.
-
-Several approaches exist to measure the ``distance'' between two
-ranked lists. Of these, Kendall's $\tau$ \citep{kendall1948} and
-Spearman’s footrule $\rho$ \citep{Spearman1910} are among the most
-well-known, but these measures do not distinguish between agreement in
-the top versus towards the bottom of the lists and the provide a
-single measure of the overall distance. \citet{Shieh1998} proposed a
-weighted version of Kendall's $\tau$ where each pair of rankings can
-be assigned different weights, and \citet{Yilmaz2008} proposed the
-$\tau_{ap}$ which places higher emphasis on the top of the lists.
-Spearman’s footrule uses the ranks of the variables for calculation of
-the distance and the use of ranks are also employed in the $M$ measure
-of \citet{Bar-Ilan2006} where the reciprocal rank differences are used
-to calculate the similarity measure.
-
-Other recent approaches consider the intersection of lists as the basis
-for a similarity measure. However, simple intersection also places
-equal weights on all depths of the list and therefore \citet{Fagin2003} and
-\citet{Webber2010} proposed weighted intersections which put more
-emphasis on the top of the lists. One proposal is denoted the average
-overlap (AO) and used for comparison later on. Specifically,
-\citet{Webber2010} define their rank-biased overlap (RBO) by weighting
-with a converging series to ensure that the top is weighted higher
-than the potentially non-informative bottom of the lists.
-It is possible to use the existing methods to calculate agreement of
-lists until a given depth, i.e., limited to the $d$ items of each
-list. However, the interpretation may not be straightforward,
-especially in the case of more than two lists, and they may not
-accommodate partial rankings.
-
-In this article we introduce sequential rank agreement for measuring
-agreement among (partially) ranked lists.  The general idea is to
-define agreement based on the sequence of ranks from the first $d$
-elements in each list. As agreement metric we adapt the limits of
-agreement known from agreement between quantitative variables
-\citep{alt:bland:1983,Carstensen2010}. Our proposed approach allows us
-to compare more than two lists simultaneously, it provides a dynamic
-measure of agreement as a function of the depth in the lists, places
-high weight on the top of the list, accommodates censored/incomplete
-lists of varying lengths, and has a natural interpretation that
-directly relates to the ranks. Graphical illustration of sequential
-rank agreement potentially allows us to infer a change-point, i.e., a
-list depth where a change in the agreement of the lists occur but we
-also provide randomization-based graphical tools to compare the
-observed rank agreement to the expected agreement found in
-non-informative data.  In this sense our approach is a combination and
-generalization of some of the ideas of \citet{Carterette2009} and
-\citet{Boulesteix2009}. The former compares two rankings based on the
-distance between them as measured by a multivariate Gaussian
-distribution and the latter presents an overview of approaches for
-aggregation of ranked lists including bootstrap and leave-one-out
-jackknife approaches.
-
-The manuscript is organized as follows: In the next section we define
-sequential rank agreement for multiple ranked lists and discuss how to
-handle incomplete/censored/partial lists. In section 3 we present and
-discuss approaches to evaluate the results obtained from sequential
-rank agreement. Finally we apply the sequential rank agreement to two
-Danish ovarian cancer studies before we discuss the findings along
-with possible extensions. The approaches presented in this manuscript
-are available in the R package \texttt{SuperRanker}.
-
 <<echo=FALSE>>=
@@
 @ %
-
-\section{Methods}
-
-Consider a set of $P$ different items $X=\{X_1,\dots,X_P\}$ and a
-ranking function $R: \{X_1,\dots,X_P\}\to \{1,\dots,P\}$, such that
-$R(X_p)$ is the rank of item $X_p$. The inverse mapping $R^{-1}$ gives
-the item $R^{-1}(r)$ that was assigned to rank $r\in\{1,\dots,P\}$. An
-ordered list is the realization of a ranking function $R$ applied to a
-data sample $X$ and we let $R_1(X),\dots,R_L(X)$, $L\geq2$, denote
-ordered lists obtained from $L$ different ranking functions applied to
-the same data $X$. We will use the same notation for the equivalent
-scenario where we have ordered lists obtained from applying the same
-ranking function to $L$ samples from the same population, i.e., for
-ranked lists $R(X^1), \ldots, R(X^L)$.  Panels (a) and (b) of
-Table~\ref{tab:example} show a schematic example of these
-mappings. Thus if $R_l^{-1}(1)=X_{34}$ then item $X_{34}$ is ranked
-first in list $l$ and similarly $R_l(X_{34})=1$.
-
-\begin{table}[tb]
-  \caption{Example set of ranked lists. (a) shows the ranked list of
-    items for each of three lists, (b) presents the ranks obtained by
-    each item in each of the three lists and (c) shows the cumulative
-    set of items up to a given depth in the three lists.}
-\begin{center}
-\footnotesize{
-  \begin{subtable}{4.5cm}%
-    \caption{}
-      \begin{tabular}{cccc}
-        \hline\hline % & \multicolumn{3}{c}{List} \\
-        Rank & $R^{-1}_1$ & $R^{-1}_2$ & $R^{-1}_3$ \\ \hline
-        1 & A & A & B \\
-        2 & B & C & A \\
-        3 & C & D & E \\
-        4 & D & B & C \\
-        5 & E & E & D \\ \hline
-    \end{tabular}
-  \end{subtable}
-% \hfill
-\hspace{1em}
-%%%
-  \begin{subtable}{4cm}%
-    \caption{}
-    \begin{tabular}{cccc} \hline\hline
-  % & \multicolumn{3}{c}{List} \\
-    Item & $R_1$ & $R_2$ & $R_3$ \\ \hline
-    A & 1 & 1 & 2 \\
-    B & 2 & 4 & 1 \\
-    C & 3 & 2 & 4 \\
-    D & 4 & 3 & 5 \\
-    E & 5 & 5 & 3 \\ \hline
-  \end{tabular}
-\end{subtable}
-% \hfill
-\hspace{1em}
-%%%
-\begin{subtable}{4cm}%
-  \caption{}
-\begin{tabular}{cc} \hline\hline
-% Depth   &     \\
-Depth  &  $S_d$ \\ \hline
-1 & $\{$A, B$\}$\\
-2 & $\{$A, B, C$\}$ \\
-3 & $\{$A, B, C, D, E$\}$ \\
-4 & $\{$A, B, C, D, E$\}$ \\
-5 & $\{$A, B, C, D, E$\}$ \\ \hline
-\end{tabular}
-\end{subtable}
-}
-\end{center}
-\label{tab:example}
-\end{table}
-
-The agreement of the lists regarding the rank given to an item $X_p$ can be
-measured by
-\begin{equation}
-  A(X_p) = f(R_1(X_p), \ldots, R_L(X_p))
-\end{equation}
-for a measurable non-negative real-valued distance function $f$.
-For an integer $1\le d\le P$ we define the unique set of items found
-by merging the first $d$ elements of each of the $L$ lists, i.e., the
-set of items ranked less than or equal to $d$ in any of the lists:
-\begin{equation}
-S_d = \{R_l^{-1}(r) ; r \leq d, l = 1, \ldots, L \},
-\end{equation}
-which is exemplified in Panel (c) of Table \ref{tab:example}.
-We then define the \emph{sequential rank agreement} as the weighted expected agreement of the items
-found in the set $S_d$: [FIX: we have talked about this before, but does that redundant (L-1) factor help wrt interpretation?]
-\begin{align}
-\textrm{sra}(d)&= \frac{\sum_{\{p \in S_d\}}(L-1)\E[A(X_p)]}{(L-1)|S_d|}\label{def:sra}\\
-   &= \sum_{\{p \in S_d\}}|S_d|^{-1} \E[A(X_p)]
-\end{align}
-where $|S_d|$ refers to the cardinality of the set $S_d$, and the expectation
-is taken with respect to the distribution of the rankings of $X_p$ accross the $L$ lists.
-Values of \textrm{sra} close to zero suggests that the lists agree while larger
-values suggest disagreement. Just like in method comparison studies we
-prefer smaller limits of agreement because that means that the
-differences between the methods is small. If the ranked lists are
-identical then the value of sequential rank agreement will be zero for
-all depths $d$.
-
-%Note that the terms that are part of \eqref{def:sra} are not
-%independent since every list contains the ranks from 1 to $P$ exactly
-%once. As a consequence we simply use the pooled standard deviation as
-%a measure of the average distance between rankings and we refrain from
-%using traditional distributional results for this measure.
-
-Throughout the rest of this paper we will use the variance to quantify the
-agreement of the rankings and therefore define
-\begin{align}
-  A(X_p) = (R_i(X_p) -  \E[R_i(X_p)])^2
-\end{align}
-For each item, the function $A$ has an interpretation as the expected
-euclidean distance of the individual rankings from the expected ranking over all the lists, and
-it corresponds to the same measure that is used in method comparison studies to compute the
-limits of agreement \citep{alt:bland:1983}. The sequential rank agreement then becomes equivalent
-to the pooled variance of the items found in $S_d$, and it can be interpreted as
-the average distance of the individual rankings of the lists from the
-average ranking for each of the items we have seen until depth $d$.
-
-Given a sample of $1,\ldots,L$ lists and estimator of estimated sequatial rank agreement is given by
-\begin{equation}
-\widehat{\textrm{sra}}_L(d)= \frac{\sum_{\{p \in S_d\}}(L-1)\widehat{A_L}(X_p)}{(L-1)|S_d|}\label{def:sra}
-\end{equation}
-where
-\begin{align}
-\widehat{A}_L(X_p) = \frac{1}{L-1}\sum_{i=1}^L (R_i(X_p) - \overline{R}(X_p))^2, \quad \overline{R}_L(X_p) = \frac{1}{L}\sum_{i=1}^L R_i(X_p)
-\end{align}
-is the sample variance of the rank of item $X_p$. For an easier interpretation we suggest to plot $\widehat{\textrm{sra}}_L(d)^{1/2}$ corresponding to the pooled standard deviation.
-
-\subsection{Agreement among fully observed lists}
-\label{sec:amfol}
-
-The simplest case occurs when all $L$ lists are fully observed, \ie,
-we have observed the rank of all $P$ items for all $L$ lists. Fully
-observed ranked lists are common and arise, for example, when
-different statistical analysis methods are applied to a single dataset
-to produce lists of predictors ranked according to their importance,
-or if the same analysis method is applied to data from different
-populations.
-
-With fully observed lists we can plot the sequential rank agreement
-\eqref{def:sra} as a function of depth $d$. An example is seen in top
-panels of Figure~\ref{fig:example1} where four analysis methods were
-used to rank 3051 gene expression values measured on 38 tumor mRNA
-samples in order to classify between acute lymphoblastic leukemia
-(ALL) and acute myeloid leukemia (AML)
-\citep{Golub1999}. Preprocessing of the gene expression data was done
-as described in \citet{Dudoit2002} and the four different analysis
-approaches were: marginal two-sample $t$ tests, marginal logistic
-regression analyses, elastic net logistic regression
-\citep{friedman2010regularization}, and marginal maximum information
-content correlations (MIC) \citep{Reshef2011}. For the first two
-methods, the genes were ranked according to minimum $p$ value, for
-logistic regression the genes were ordered by size of the
-corresponding coefficients (after standardization), and MIC was
-ordered by absolute correlation which resulted in the top rankings
-seen in Table~\ref{tab1}.
-
-The sequential rank agreement plotted in Figure~\ref{fig:example1}
-shows how many ranks apart we on average expect the genes found in the
-top $d$ part of the lists to be for each depth $d$.  It can be seen
-that the sequential rank agreement limits are better towards the top
-of the lists (smaller values on the $y$ axis) corresponding to
-\emph{better} agreement than in the tails of the lists. It is clear
-from the curve in Figure~\ref{fig:example1} that there is a
-substantial drop in agreement even after the first depth. Thus, if we
-were to restrict attention to a small set of predictors then we would
-focus on predictors 2124, 829, and 378 as seen in Table~\ref{tab1}.
 
@@
 
-\begin{figure}[htbp]
-   \begin{center}
- \includegraphics[width=.49\textwidth]{paper-fig1a}
- \includegraphics[width=.49\textwidth]{paper-fig2a}
- \includegraphics[width=.49\textwidth]{paper-fig1b}
- \includegraphics[width=.49\textwidth]{paper-fig2b}
- \end{center}
- \caption{Left panels: Sequential rank agreement for 4 different
-   analysis methods applied to the 3051 genes in the Golub data /black
-   line). Right
-   panels: Corresponding sequential rank agreement for the same data
-   but where only the top 20 ranked items are available. The blue and
-   red areas correspond to the independent and randomized reference
-   hypothesis areas, respectively. The bottom plots are identical to
-   the corresponding top plots but have been zoomed in on the first
-   part of the $x$ axis.}
- \label{fig:example1}
- \end{figure}
-
- If the lists agree on all items then the sequential rank agreement is
- zero for all depths. Generally, if there are changes in the level of
- rank agreement then this suggests that there are sets of items that
- are ranked similarly in all lists and other sets of items that have
- been assigned vastly different ranks in the lists. In gene
- association studies we generally expect the agreement among the lists
- to be better towards the top of the lists and worse towards the end
- of the lists. In this case the sequential rank agreement curve will
- start at a low level and then increase until it levels off as in
- Figure~\ref{fig:example1}.
-
-\subsection{Analysis of partial/censored lists}
-
-Incomplete or partial lists are also a common occurrence. They arise
-for example in case of missing data (items), when comparing top $d$
-list results from publications, or when some methods only rank a
-subset of the items. For example, penalized regression based on the
-Lasso provides a sparse set of predictors that have non-zero
-coefficients. There is no obvious ordering of the set of predictors
-whose coefficient has been shrunk to zero and thus we end up with a
-partial ordering. Incomplete lists also occur when the analyst decides
-to censor all lists at depth $d$ such that the measurement of
-agreement is restricted to the first $d$ elements of each list, or
-when the analyst only wants to consider the ranking of items that have
-been found to be statistically significant.
-
-Sequential rank agreement can be generalized to partial lists in the
-following way. Let $\Lambda_l\subset X$ be the subset of $d_l$ items
-that have been ranked highest in list $l$.  The case where all lists are
-censored at the same depth $d$ corresponds to $d_1=\cdots=d_L=d$. For
-incomplete/censored lists the rank function becomes
-\begin{equation}
-  \widetilde R_l(X_p) = \left\{\begin{array}{cl} \{R_l(X_p)\} & \text{ for } X_p\in \Lambda_l \\
-      \{d_l+1,\dots,P\} & \text{ for } X_p \not\in \Lambda_l\end{array}\right.
-\end{equation}
-where we only know that the rank for the unobserved items in list $l$
-must be larger than the largest rank observed in that list.
-
-The agreement, $A(X_p)$, cannot be computed directly for all
-predictors in the presence of censored lists because the exact rank
-for some items will be unknown. Also, recall that the rankings within
-a single list are not independent since each rank must appear exactly
-once in each list. Thus, we cannot simply assign the same number
-(e.g., the mean of the unassigned ranks) to the censored items since
-that would result in less variation of the ranks and hence less
-variation of the agreement, and it would artificially introduce a
-(downward) bias of agreement for items that are censored in multiple
-lists.
-
-Instead we randomize the ranks $\{d_{l}+1,\dots,P\}$ to the items that
-do not occur in list $\Lambda_l$. One realization of the $L$ rankings
-of the set $X$ is obtained by randomizing the missing items of each
-list. By randomizing a large number of times we can compute
-\eqref{def:sra} for each realization, and then compute the sequential
-rank agreement as the pointwise (for each depth) average of the rank
-agreements. The algorithm is described in detail in
-Algorithm~\ref{sra-algorithm}.
-
-\begin{algorithm}
-\caption{Sequential rank agreement algorithm for censored lists}
-\label{sra-algorithm}
-\begin{algorithmic}[1]
-  \Procedure{Censored rank agreement}{} \State Let $B$ be the number
-  of permutations to use \For{each $b \in B$} \For{each censored list
-    $l \in L$}
-  \State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1.7cm}{Permute
-    the unassigned ranks, $\{d_l+1, \ldots, P\}$, and assign them
-    randomly to the items \emph{not} found in the list, i.e.,
-    $\Lambda^\complement=X\setminus\Lambda_l$. Combine the
-    result with $\Lambda_l$ to fill out the list.}
-\EndFor
-\State \parbox[t]{\dimexpr\linewidth-\algorithmicindent-1cm}{Let sra($b$) be the sequential rank agreement computed from the
-filled out lists.}
-\EndFor
-\State Return element-wise averages across all $B$ permutations of sra($b$).
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-The proposed approach is based on two assumptions: 1) that the most
-interesting items are found in the top of the lists, and 2) that the
-censored rankings provide so little information that it is reasonable
-to assume that they can be represented by a random order. The first
-assumption is justifiable because we have already accepted that it is
-reasonable to rank the items in the first place. The second assumption
-is fair in the light of the first assumption provided that we have a
-``sufficiently large'' part of the top of the lists available.
-
-When the two assumptions are satisfied then it is clear that the
-interesting part of the sequential rank agreement curves is restricted
-to the depths where the number of censored items is
-low. Generally, without additional prior knowledge about the
-distribution of the censored items it seems reasonable to restrict the
-attention of the sequential rank agreement to depths smaller than
-$\max(d_1, \ldots, d_L)$ where at least one list provides an
-uncensored item.
-
-Like for fully observed lists we generally expect the sequential rank
-agreement to start low and then increase unless the lists are
-completely unrelated (in which case the sequential rank agreement will
-be constant at a high level) or if the lists mostly agree on the
-ranking (in which case the sequential rank agreement will also be
-constant but at a low level). For partially ranked lists we also
-expect a change-point around the depth where the lists are
-censored. This is an artefact stemming from the fact that we assume
-that the remainder of the lists can be replaced by a simple
-permutation of the missing items.
-
-\section{Evaluating sequential rank agreement}
-
-To interpret the sequential rank agreement values we propose two
-different benchmark values corresponding to two different
-hypotheses. We wish to determine if we observe better agreement than
-what would be expected if there were no relevant information available
-in the data.
-
-The first reference hypothesis is
-\begin{eqnarray*}
-H_0  & : &  \text{The list rankings correspond to complete randomly}\\
-       &  & \text{permuted lists},
-\end{eqnarray*}
-which not only assumes that there is no information in the data on
-which the rankings are based but also that the methods are completely
-independent.
-
-Alternatively, we can remove the restriction on the independence among
-the methods by only requiring that there is no information in the
-ranking but that the rankings are all based on applying the
-method/approaches to the same data
-\begin{eqnarray*}
-\widetilde H_0 & :&  \text{The list rankings are based on data containing}\\
-& &   \text{no association to the outcome.}
-\end{eqnarray*}
-$H_0$ is a quite unrealistic null hypothesis but we can easily obtain
-realizations from that null hypothesis simply by permuting the items
-within each list and then computing the sequential rank agreement for
-the permuted lists. In the fully observed case each experiment
-contains $L$ lists of random permutations of the items in $X$. For the
-censored/partially observed case we first permute the items
-$X_1,\dots,X_P$ and then censor list $l$ at $d_l$. The sequential rank
-agreement curve from the the original lists can then be compared to,
-say, the pointwise 95\% quantiles of the observed rank agreements
-obtained under $H_0$.
-
-To obtain the distribution under $\widetilde H_0$ the idea is to
-repeat the ranking procedures for unassociated data many times.  Thus,
-we first permute the outcome variable of the data set on which the
-rankings are based. This removes any association between the predictor
-variables and the outcome. Then we apply the methods to the permuted
-data set to generate $L$ new rankings and subsequently compute the
-sequential rank agreement. Note that we only permute the outcomes and
-preserve the structure of the candidate predictors. The randomization
-approach requires that we have the original data available and as such
-it may not be possible to evaluate $\widetilde H_0$ in all situations.
-
-If the sequential rank agreement for the original data lies
-substantially below the distribution of the sequential rank agreements
-obtained under either $H_0$ or $\widetilde H_0$ then this suggests that
-the original ranked lists agree \emph{more} than what we would expect
-in data with no information, and therefore that the information in the
-lists is significantly more in agreement than what would be expected.
-
-Figure~\ref{fig:example1} shows the empirical distributions of
-sequential rank agreement under $H_0$ and $\widetilde H_0$ each based
-on $400$ permutations of the Golub data described in Section
-\ref{sec:amfol}. Not surprisingly, the sequential rank agreement under
-$\widetilde H_0$ is lower than the sequential rank agreement under
-$H_0$ because the four methods used to rank the data ($t$ test,
-logistic regression, elastic net, and MIC) generally tend to identify
-(and rank) similar predictors even if there are only spurious
-associations. The two bottom panels in Figure~\ref{fig:example1} also
-indicate that the observed sequential rank agreement (the black line)
-is better than what would be expected by chance for data that contain
-no information since it lies below the reference areas. However, the
-censored data also suggests that there may be at most 1 or 2 ranked
-items towards the top of the lists that yield a result better than
-what would be expected (the bottom-right plot).
-
-It is important to stress that neither $H_0$ nor $\widetilde H_0$ are
-related to questions regarding the association between the outcome and
-the predictors in the data set. Both hypotheses are purely considering
-how the rankings agree in a situation where there is no relevant
-information available in the data used for creating the rankings.
-
-\section{Change-point analysis}
-In many applications it is of interest to estimate a list depths which
-satisfies a change point criterion. In this section we investigate the
-theoretical properties of our proposed method for this specific
-task. As Hall and Schimek, we work with an infinite set of lists
-$\mathcal{L}$ and study the asymptotic behaviour for
-$L\to\infty$. However, we do not allow the length of the lists to
-change with $L$ as in most applications the length of the lists is
-fixed [FIXME: for finite $L$ are there only finitely many forskelige
-lists (aber dass stoert keinen grossen geist, siger karlsson)]. First,
-we define the parameter of interest. For any given threshold value $q$
-a change-point $d^\ast(q)$ in the list agreement is defined by
-\begin{equation}
-  d^\ast(q) = \arginf_{d\in\{1,\dots,P\}}\textrm{sra}(d)>q.
-\end{equation}
-The value of $q$ could for example be the agreement of randomly
-permuted lists [FIXME: there is only one SRA value for infinitely many
-random permutations, no skygge!].
-The corresponding estimated change-point is similarly defined by
-\begin{equation}
-\widehat{d^\ast_L}(q) = \arginf_{d\in\{1,\dots,P\}}\widehat{\textrm{sra}}_L(d)>q
-\end{equation}
-
-We start by showing that $\widehat{\textrm{sra}}_L(d)$ is consistent for $\textrm{sra}(d)$ at each
-$d \in \left\{1,\ldots,P\right\}$ for $L \rightarrow \infty$. The consistency of $\widehat{d^\ast_L}(q)$
-then follows by a simple corollary.
-
-\begin{theorem}
-hej
-\begin{proof}
-We start by noting that we can rewrite the sequential rank agreement in equation \ref{def:sra} as
-\begin{equation}
-\textrm{sra}(d)= \frac 1{\sum_{p=1}^P \sum_{l=1}^L 1\{R_l(X_p)\le d\}}\sum_{p=1}^P(1-\prod_{l=1}^L 1\{R_l(X_p)>d\}) \E[A(X_p)]
-\end{equation}
-
-[FIXME: Andreas details the proof with Cauchy-Schwarz.]
-[FIXME: Det er lidt kedelig for saadan en fin en som Andreas: Skal theoremet og bevis ikke vaere for en general agreement metric, ikke nodvendigvis den i equation (2), og kan vi charakterisere hvad den skal opfylde?]
-\end{proof}
-\label{thm:consistency}
-\end{theorem}
-
-\begin{corollary}
-For fixed $q$ it follows from theorem \ref{thm:consistency} that
-$\widehat{d^\ast_L}(q) \overset{P}{\longrightarrow} d^\ast(q)$
-for $L \rightarrow \infty$.
-\begin{proof}
- hej
-\end{proof}
-\end{corollary}
-
-\section{Comparison to average overlap}
-
-The average overlap is a measure that is widely used in comparisons of
-two fully observed ranked lists \citep{Fagin2003,Webber2010} and the
-idea behind the average overlap closely resembles the sequential rank
-agreement. In this section we present an extension of the average
-overlap to more that two lists and compare this to sequential rank
-agreement.
-
-The overlap among $L$ lists observed until depth $d$ is defined as the
-proportion of items that are found in all $L$ lists compared to the
-possible number of items found in all lists observed until depth $d$:
-\begin{equation}
-  O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}. \label{def:overlap}
-\end{equation}
-Typically the overlap is only considered for pairwise comparisons
-\citep{Bar-Ilan2006,Boulesteix2009} but formula \ref{def:overlap}
-is easily extended to the situation where $L>2$ as shown above.
-
-The average overlap at depth $d$ is defined as the average of the
-overlaps until depth $d$,
-\begin{equation}
-AO(d) = \frac1d\sum_{i=1}^d O(d).
-\end{equation}
-This definition directly ensures that further emphasis is put on items
-in the top of the lists since their item specific overlap contributes
-to the calculation of average overlap at higher depths.
-
-There are three main drawbacks of the average overlap method: First it
-is highly sensitive to single items in the top of one the lists.  The
-average overlap profile may change dramatically depending on the
-presence or absence of similar items in the beginning of the
-lists. For example, if all lists have the same item at the top then
-the average overlap attains its maximum at 1, but if just one of the
-lists differ then the overage overlap starts at its minimum of
-0. Secondly, when the number of lists, $L$, is high then it might be
-difficult to obtain a non-zero overlap (and hence a non-zero average
-overlap) at the beginning of the lists because the overlap will be
-zero until an item is present in all $L$ lists. Finally, the average
-overlap has an interpretation in terms of moving averages of
-percentages (of $L$ sets) which is somewhat less intuitive than our
-proposed sequential rank distance.
-
-In the following we compare the average overlap and the sequential
-rank agreement using the analysis of the Golub data
-\citep{Golub1999}. We consider a typical application of the average
-overlap and thus restrict attention to only two gene rankings obtained
-by the $t$ tests and logistic regression. These two marginal modeling
-approaches should result in very similar gene rankings. The first two
-columns of Table~\ref{tab1} show the actual top-10 ranking of the
-predictors.
-
 <<fig-ao1, echo=FALSE, fig=TRUE,include=FALSE, eval=TRUE>>=
@@
 @
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.49\textwidth]{paper-fig-ao1}
-\includegraphics[width=.49\textwidth]{paper-fig-ao2}
-\end{center}
- \caption{Average overlap (left figure) and sequential rank agreement
- (right figure) for comparing two different analysis methods
- (marginal $t$ test and logistic regression)
-    applied to the Golub data. The black lines are based on the full
-    set of predictors while the red lines have the two highest
-    associated predictors removed
-    from the data before computing the average overlap and sequential
-    rank agreement.}
-  \label{fig:case1}
- \end{figure}
-
- The black lines in Figure~\ref{fig:case1} show the average overlap
- and sequential rank agreement for these data. It is clear from both
- plots that there is perfect agreement towards the top of the ranked
- lists: the average overlap is 1 and the sequential rank agreement is
- zero.  However, if we remove those two items from the data (i.e.,
- gene/predictor 2124 and 896) and redo the analyses then we get
- substantially different curves for the average overlap but not for
- sequential rank agreement (red lines in Figure~\ref{fig:case1}). For
- the sequential rank agreement we get roughly the same estimate of
- agreement from rank 3 as we did from the full dataset. This indicates
- that sequential rank is more robust against small perturbations of
- the data. For the average overlap the curves completely change when
- these two items are removed and which would lead to a different
- conclusion about the agreement among the items from rank 3 downwards.
-
-\section{Application to ovarian cancer data}
-We now consider an application of the sequential rank agreement to two
-data sets consisting of MALDI-TOF (Matrix-Assisted Laser
-Desorption/Ionization Time Of Flight) mass spectra obtained from blood
-samples from patients with either benign or malignant ovarian
-tumors. The data sets are sub-samples of the Danish MALOVA and DACOVA
-study populations.
-
-The MALOVA study is a multidisciplinary Danish study on ovarian cancer
-\citep{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
-with an ovarian tumor and referred for surgery from the participating
-departments of gynecology were enrolled continuously from December
-1994 to May 1999. For the purpose of illustration we use a random
-sub-sample of $119$ patients with a total of $58$ patients with
-malignant ovarian cancers as cases and $61$ patients with benign
-ovarian tumors as controls. The DACOVA study is another
-multidisciplinary Danish study on ovarian cancer which included about
-$66\%$ of the female population of Denmark
-\citep{bertelsen1991protocol}. The study aimed to continuously enroll
-all patients that were referred to surgery of an ovarian tumor
-clinically suspected to be cancer during the period from 1984 to
-1990. Similarly, we use a random sub-sample from the DACOVA study of
-$113$ patients with a total of $54$ malignant ovarian cancers and $59$
-benign ovarian tumors/gynecologic disorders.
-
-Each spectrum consists of $49642$ samples over a range of mass-to-charge ratios
-between $800$ to $20000$ Dalton which we downsample on an equidistant grid of
-5000 points by linear interpolation. We then preprocess the downsampled
-spectra individually by first removing the slow-varying baseline intensity
-with the SNIP algorithm \citep{ryan1988snip} followed by a
-normalization with respect to the total ion count. Finally, we standardize
-the $5000$ predictors to have column-wise zero mean and unit variance in
-each data set.
-
-We use the two data sets to illustrate how the sequential rank
-agreement can be applied in two different scenarios. In the first
-scenario we assess the agreement of four different statistical
-classification methods in how they rank the predictors according to
-their importance for distinguishing benign and malignant tumors. In
-the second scenario we assess the agreement among rankings of
-individual predicted risks of having a malignant tumor. The first
-scenario is relevant in the context of biomarker discovery and the
-latter is important e.g., when ranking patients according to immediacy of
-treatment.
-
-Four classification methods are considered: Random Forest
-\citep{breiman2001random} implemented in the R package
-\texttt{randomForest} \citep{liaw2002classification}, logistic Lasso
-\citep{tibshirani1996regression} and Ridge regression
-\citep{segerstedt1992ordinary} both implemented in the R package
-\texttt{glmnet} \citep{friedman2010regularization}, and Partial Least
-Squares Discriminant Analysis (PLS-DA) \citep{boulesteix2004pls}
-implemented in the R package \texttt{caret} \citep{Jed-Wing:2014aa}.
-
-In both scenarios we use the MALOVA data to train the statistical
-models, and in both situations the agreements are assessed with
-respect to perturbations of the training data in the following
-manner. We repeatedly draw a random sub-sample (without replication)
-consisting of 90\% of the MALOVA observations and train the four
-models on each sub-sample. We use 1000 iterations for the sub-sampling
-procedure.
-
-All four methods depend on a tuning parameter. The tuning parameter for
-Lasso and Ridge regression is the degree of penalization, and for PLS-DA
-it is the number of components (the dimensionality of the subspace).
-We estimate these separately for each sub-sample by a 20 times repeated
-5-fold cross-validation procedure. For the Random Forest we grow a fixed
-number of $5000$ trees and let the tuning parameter be the number of
-predictors randomly sampled at each split. We estimate this by a binary
-search with respect to minimizing the Out-of-Bag classification error estimate.
-
-The implementation of Lasso and Ridge regression in the
-\texttt{glmnet} package offers three different cross-validated
-optimization criteria for the penalty parameter: total deviance,
-classification accuracy and area under ROC. We apply all three
-criteria to our data to investigate their effect on the agreements.
-Note also that the Lasso models produce censored lists depending on
-the value of the penalty parameter.
-
-\subsection{Agreement of predictor rankings}
-For each of the four methods, each of the 1000 models trained on
-the 1000 sub-samples of the MALOVA data produces a ranking of the 5000
-predictors according to their importance for discriminating between
-the tumor types. For the Random Forest classifier the predictors are ranked according
-to the Gini index, while for the logistic Lasso and Ridge regression
-models we order by absolute magnitude of the estimated
-regression coefficients. For the PLS-DA model the importance of the
-predictors is based on a weighted sum of the absolute coefficients
-where the weights are proportional to the reduction in the sums of
-squares across the components.
-
-The right panel of Figure \ref{fig:app1} shows the sequential rank
-agreement of the estimated importance of the 5000 predictors.  For
-clarity of presentation we zoom in on the agreement up to list depth
-600. At deeper list depths all agreement curves are approximately constant.
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.49\textwidth]{pics/riskAgreementPlot}%
-\includegraphics[width=.49\textwidth]{pics/predictorAgreementPlot}
-\end{center}
-\caption{Left panel: Sequential rank agreement of 1000 rankings of the
-  predicted risks of malignant tumor. For each method the different
-  rankings were obtained by first training models in 1000 random
-  sub-samples of the MALOVA data and then predicting the risk of
-  malignant tumor in the 113 DACOVA patients. Right panel: Sequential
-  rank agreement of 1000 rankings of the 5000 predictors. The rankings
-  were obtained from the same 1000 trained models.}
- \label{fig:app1}
-\end{figure}
-
-For most of the sequential rank agreement curves we see, as expected,
-that they start low, indicating good agreement, followed by an
-increase until they approximately become constant. This has the
-interpretation that the agreement across the different sub-samples is
-higher in the top as compared to the tail of the lists for all these
-classification methods. The change-points where the curves become
-approximately constant are the list depths where the ranks of the
-remaining items become close to uniformly random.
-
-A not expected shape of the agreement curves is seen for the Ridge
-models for all three tuning criteria. They all show higher
-disagreement in the top of the lists followed by a decrease. The
-reason behind this behavior is rather subtle. Looking at the
-distribution of the absolute value of the regression coefficients we see that a large
-proportion of them are numerically very close to zero and have almost
-equal absolute value. This is a general feature of the Ridge models in
-this data set and seen for all the 1000 trained models.
-This implies that when predictors are ranked according to the
-magnitude of their coefficients, their actual order becomes more
-uncertain and more close to a random permutation. This problem can be
-alleviated by truncating all predictors with absolute coefficient values
-below a given threshold thereby introducing an artificial censoring of the lists.
-For the Ridge models tuned with the deviance criterion, Figure \ref{fig:app2} (left) shows the
-the sequential rank agreement where for each of the 1000 trained
-models the predictors were artificially censored when their absolute
-coefficient value was lower than the $0.1\%$ quantile of the 5000
-absolute coefficient values. The curve was calculated
-using Algorithm \ref{sra-algorithm} with $B=1000$ and $P=5000$. The corresponding
-curve from Figure \ref{fig:app1} (right panel) is shown for comparison.
-Even though the number of censored predictors is very small compared to the total
-number of predictors, the effect on the sequential rank agreement is
-substantial and with the artificial censoring the shape of the curves
-is as expected, starting low and then increasing.
-
-Looking at the agreement curves for the Lasso models in Figure \ref{fig:app1}
-(right) we clearly see the effect of the sparsity inducing penalization giving
-rise to censored lists. These curves were similarly calculated using \ref{sra-algorithm}
-and $1000$ random permutations. Under the deviance optimization criterion the
-median number of non-zero coefficients was 33 (range 16 to 50) and for the
-class accuracy criterion 14 (range 4 to 56). These values correspond to
-the list depths where the agreement curves become constant as a result of
-the subsequent censoring.
-
-\subsection{Agreement of individual risk predictions}
-\label{sec:airp}
-To assess the stability of the individual risk predictions we apply
-the predictors from the DACOVA data set to each of the models. The
-predicted probabilities are then ranked in decreasing order such that the
-patients with the highest risk of a malignant tumor appears in the top
-of the list. Figure \ref{fig:app1} (left) shows the sequential rank
-agreement separately for each method, based on the 1000 risk predictions
-obtained from the models trained in the same $1000$ random sub-samples
-of the MALOVA data.
-
-Most curves start low and then increase indicating higher agreement
-among high risk patients. This is expected if we rank the individuals
-according to highest risk of disease. However, it is also expected
-that individuals with very low risk also show high agreement. In this
-case we order the patients according to (high) risk prediction but we
-could essentially also have reversed the order to identify the
-patients that have low risk prediction.
-
-An exception is the risk prediction agreement for the Lasso tuned with the AUC criterion which
-shows very low agreement among the high values of the predicted
-risks. The reason is that optimizing the penalty parameter with
-respect to the AUC criterion tends to favor a very high penalty value
-causing only a single predictor to be selected in each of the 1000
-iterations. This results in a lack of generalizability to the DACOVA
-data which gives rise to the higher disagreement in the predicted risks.
-In the extreme case where the penalty becomes so high
-that none of the predictors are selected by the Lasso, the sequential
-rank agreement for the predicted probabilities becomes undefined since
-all the ranks will be ties.
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.49\textwidth]{pics/ridgeCensoring}%
-\includegraphics[width=.49\textwidth]{pics/aucPlot}
-\end{center}
-\caption{Left panel: Sequential rank agreement for Ridge regression
-  obtained by artificially censoring predictors when their absolute
-  coefficient values are lower than the 0.1\% quantile. Right panel:
-  Box plots of AUC values across the $1000$ sub-samples with respect
-  to the known class labels of the DACOVA data.}
- \label{fig:app2}
-\end{figure}
-
-Comparing the left and right panels of Figure \ref{fig:app1} it can
-further be seen that some of the methods show better agreement with
-respect to the predicted probabilities than for ranking the importance
-of the predictors and vice versa. Ridge regression shows higher agreement across training
-sets for the risk predictions than PLS-DA, and PLS-DA shows higher
-agreement for predictor importance than Ridge regression.
-
-Lasso shows similar agreement for ranking the risk predictions than
-PLS-DA (unless with the AUC criterion), and poorer agreement for ranking
-predictors. This reason for the latter is the high auto-correlation
-between the intensities in the mass spectra which leads to
-collinearity issues in the regression models. It is well-known that
-variable selection with the Lasso penalty does not preform very well
-when the predictors are highly correlated. The collinearity does,
-however, not affect the agreement of the risk predictions, as seen on
-the left panel of Figure \ref{fig:app1}, since it is not so important
-which specific variable that gets selected from a group of highly correlated
-predictors when the purpose is risk predictions.
-
-It appears that Ridge regression tuned with the AUC criterion achieves
-the best performance with respect to the stability of ranking the
-individual predicted risk probabilities. It must, however, be stressed
-that the sequential rank agreement in this application is only
-concerned with the agreement of the risk predictions across
-sub-samples and not with the actual accuracy of the risk
-predictions. Thus, we also computed the AUC values for the different models
-based on the DACOVA data. The distributions across the $1000$
-sub-samples for a selection of the models is shown in the right panel
-of Figure \ref{fig:app2}.  Here we see that PLS-DA attains the highest
-AUC values with a median value of $0.70$ while the Ridge model with
-the AUC criterion attains a median AUC of $0.49$. This implies that
-while Ridge regression optimized with respect to the AUC criterion
-achieves the best sequential rank agreement, it performs similar to a
-random coin toss with respect to classifying the DACOVA patients.
-In practice both concerns are of importance.
-
-\section{Discussion}
-In this article we address the problem of comparing ranked lists of
-the same items. Our proposed method can handle both the situation
-where the underlying data to generate the ranked lists are available
-and the situation where the only available data is the actual ranked
-lists. In addition, censored ranked lists where only the ranks of the
-top $k$ ranked items are known can be accommodated as well. The
-proposed agreement measure can be interpreted as the average distance
-between an item's rank and the average rank assigned to that item
-across lists. Thus the measure determines how well the lists agree on
-the ranks of a specific item.
-
-The sequential rank agreement is extremely versatile. We have shown
-that it can be used not only to compare ranked lists of items produced
-from different samples/populations but that it also can be used to
-study the ranks obtained from different analysis methods on the same
-data as well as to evaluate the stability of the ranks from a single
-method by bootstrapping (or sub-sampling) the data repeatedly and
-comparing the ranks obtained from training the models in the
-bootstrapped data.
-
-The sequential rank agreement can be used to determine the depth at
-which the rank agreement becomes too large to be desirable based on
-prior requirements or acceptable differences, or it can be used to
-visually determine when the change in agreement becomes too large.  In
-that regard the investigator can have prior limits on the level of
-agreement that is acceptable.
-
-While the sequential rank agreement is primarily an exploratory tool
-we have suggested two null hypotheses that can be used to evaluate the
-sequential rank agreement obtained. Note that none of the two null
-hypotheses are concerned with the actual ``true ranking'' but are
-purely concerned with consistency/stability of the rankings among the
-lists. As such we cannot determine if the rankings are good but only
-whether they agree. The sequential rank agreement curve can be
-compared visually to the curves obtained under either of the null
-distributions and simple point-wise $p$-values can be obtained for
-each depth by counting the number of sequential rank agreements under
-the null hypothesis that is less than or equal to the observed rank
-agreement. Two simple extensions can be pursued to make more formal
-uniform tests: one would be to use a Kolmogorov-Smirnov-like test and
-use the largest difference between the variance-weighted sequential
-rank agreement curve and the mean null rank agreement curve as a test
-statistic. Alternatively, a change-point analysis could be made on the
-sequential rank agreement in order to determine the depths at which
-there are ``jumps'' in the rank agreement. These jumps would
-correspond to depths for which the agreement among the lists was
-substantially worse and could serve as indicators for when the lists
-no longer agree sufficiently satisfactory.
-
-Finally, we have --- whenever possible --- used all available ranks
-from the lists.  We could choose to censor the rankings of all the
-items that were deemed not to be significant if information on both
-rankings and the corresponding evidence for significance was
-available. That would ensure that there would be put less emphasis on
-the agreement of the non-significant items and it would be easier to
-identify a change in agreement among the items that were deemed to
-be relevant. In our application section we have successfully introduced
-such an artificial censoring for the predictor rankings obtained with
-ridge regression.
-
-We note that the sequential rank agreement is still marred by problems
-that generally apply to ranking of items and/or
-individuals. Collinearity in particular can be a huge problem when
-bootstrapping data or when comparing different analysis methods. For
-example, marginal analyses where each item is analyzed separately will
-assign similar ranks to two highly correlated predictors while methods
-that provide a sparse solution such as the Lasso will just rank one of
-the two predictors high while the other might have a very low rank.
-Thus in such a scenario we would expect low agreement of the rankings
-from Lasso and marginal analyses simply because of the way correlated
-predictors are handled. This is not a shortcoming of the sequential
-rank agreement but is a problem general to all ranked lists.
-
-Another caveat with the way the sequential rank agreement is defined
-is the use of the standard deviation to measure agreement. The
-standard deviation is an integral part of the limits-of-agreement as
-discussed by \citet{alt:bland:1983}. However, the standard deviation
-can also be unstable when the number of observations is low and
-alternative measures such as the median absolute deviance may prove more
-stable in some situations. However, the current definition using the
-standard deviation is analogous to the approach used for agreement in
-method comparison studies so we have used that.
-
-In conclusion we have introduced a method for evaluation of ranked
-(censored) lists that can be easily interpreted and that can be
-applied to a large number of situations.  The method presented here
-can be adapted further by using it to compare and classify statistical
-analysis methods that agree on the rankings they provide or by using
-the rank agreement to optimize a hyper-parameter in, say, elastic net
-regularized regression where the rank agreement is used to determine
-the mixing proportion between the $L_1$ and the $L_2$ penalty.
-%Finally, the proposed method may be adapted (with some additional
-%assumptions) to the situation where the lists are subsets of each
-%other.
-We will be investigating these extensions further in the future.
-
-%\bibliographystyle{rss}
-\bibliographystyle{rss}
-\bibliography{paperref}
-
-\end{document}

### 2023-08-27 R/SuperRanker-package.R

@@ -1,3 +1,7 @@
+#' @keywords internal 
+"_PACKAGE"
+
+
 #' Functions related to comparison of ranked lists
 #'
 #' SuperRanker allows you to estimate the agreement between two or more rankings of the same items.

### 2023-08-27 src/RcppExports.cpp

@@ -5,6 +5,11 @@
 
 using namespace Rcpp;
 
+#ifdef RCPP_USE_GLOBAL_ROSTREAM
+Rcpp::Rostream<true>&  Rcpp::Rcout = Rcpp::Rcpp_cout_get();
+Rcpp::Rostream<false>& Rcpp::Rcerr = Rcpp::Rcpp_cerr_get();
+#endif
+
 // overlap
 NumericVector overlap(IntegerMatrix rankMat);
 RcppExport SEXP _SuperRanker_overlap(SEXP rankMatSEXP) {

### 2023-08-27 R/SuperRanker-package.R
Documentation update

@@ -1,13 +1,10 @@
-#' @keywords internal 
-"_PACKAGE"
-
-
 #' Functions related to comparison of ranked lists
 #'
 #' SuperRanker allows you to estimate the agreement between two or more rankings of the same items.
 #'
 #' @docType package
 #' @name SuperRanker
+#' @aliases SuperRanker-package
 #' @useDynLib SuperRanker
 #' @importFrom Rcpp sourceCpp
 #' @importFrom stats quantile
